### 2015-07-31 hz_math_notes.tex
initial import

\documentclass[12pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
\usepackage{hyperref}
\usepackage{amsmath,amssymb}
% See the ``Article customise'' template for come common customisations
\usepackage{color}
\newcommand{\gc}[1]{{\em \color{blue} {Graham: #1} }}
\newcommand{\as}[1]{{\em \color{red} {Alisa: #1}}}
\newcommand{\yb}[1]{{\em \color{green} {Yaniv #1}}}

\def\Var{\mathop{\rm Var}}
\def\Cov{\mathop{\rm Cov}}

\title{}
\author{}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

This is the equation describing the distribution of allele frequency across geographic space (Slatkin 73, eq. 11):

\begin{equation}
A(\xi) = -2 + 3\tanh^2(\xi/2 + \tanh^{-1}(\sqrt{2/3}))
\end{equation}
where
\begin{equation}
\xi = x/l_c
\end{equation}
where 
\begin{equation}
l_c = \sigma/\sqrt{s}
\end{equation}
the characteristic length.

$$A = p-q = p - 1+p = 2p-1$$

Which means:

\begin{equation}
p(\xi) = \frac{A(\xi)+1}{2}
\end{equation}

If $x$ is negative:

\begin{equation}
p = \frac{3}{2}(1-\tanh^2(-\xi/2 + \tanh^{-1}(\sqrt{2/3})))
\end{equation}

\subsubsection{Approximations of movement and recom. probs in tails}
Furthermore, if $x$ (or $\xi$) is really negative:

\begin{equation}
p = \frac{3}{2}(1-\tanh^2( -\xi/2))
\end{equation}

Let $-\xi/2 = z$:

\begin{equation}
\tanh(z) = \frac{\exp(z) - \exp(-z)}{\exp(-z) + \exp(z)}
\end{equation}

Divide through by $\exp(z)$:
\begin{eqnarray}
= &\frac{1 - \exp(-2z)}{\exp(-2z) + 1}\\
 = &\frac{1}{\exp(-2z) +1} - \frac{\exp(-2z)}{\exp(-2z) +1}
\end{eqnarray}

Since $-z$ is actually positive, the denominator is approximately $\exp(-2z)$. So:
\begin{eqnarray}
& = \frac{1}{\exp(-2z) } - \frac{\exp(-2z)}{\exp(-2z) }\\
 &= \frac{1}{\exp(-2z) } -1 \\
& = {\exp(2z) } -1 \\
\tanh(z) & \approx \exp(2z) - 1
\end{eqnarray}

Which means that:
\begin{equation}
\tanh^2(z) \approx 1 - 2\exp(2z) + \exp(4z) \approx 1 - 2\exp(2z)
\end{equation}
since $\exp(4z)$ goes to $0$ faster than  $\exp(2z)$\\

Sub into equation for $p$:
\begin{eqnarray}
 p &\approx 3/2(1-  1 + 2\exp(2z)) = 3\exp(2z)
 p &\approx  6\exp(2z)/2 = 3\exp(2z)\\
\end{eqnarray}
Which makes the derivative:
\begin{equation}
\frac{dp}{dz} = 6\exp(2z)
\end{equation}
Making
\begin{equation}
\frac{dp}{dz}/p = 2 \label{scaled_speed}
\end{equation}
A constant drift to the right in the tail.

So that in unscaled units the `drift' towards the right-side of the
hybrid zone in the tails  is
\begin{equation}
c= 2/l_c   \label{unscaled_speed}
\end{equation}
%\gc{Do I have the $l_c$ right way up?}\as{I think you do}
That means that in the tails the change in position per generation is
distributed as:
\begin{equation}
N(c,\sigma)
\end{equation}
%\gc{need to check we aren't missing some factor here, e.g. check there is no $\delta t$ needed}
\paragraph{An approximate probability of the probability we've not  recombined off}
Assuming that we can ignore the variance $c \gg \sigma$ (?) then if we are on the wrong background we move right at constant rare give\eqref{unscaled_speed} (at least in the tail). Assuming that if we recombine before we make it to thecenter of the hybrid zone, that we automatically recombine onto the good background. We sample a lineage a geographic distance $L$ away form the hybrid zone, at a locus a genetic distance $r$ from selected locus. Then if $\tau$ is the time that the zone establishes, the probability that our lineage stays on the wrong background is
\begin{eqnarray}
\approx \exp( - L r / c ), ~\textrm{for~}  L/c <\tau\\
0,~ \textrm{otherwise}\\
\end{eqnarray}
%\gc{need to think about the meaning of the too far from the hybrid  zone cutoff here, do we need $\exp(-r\tau)$ there instead. This  could reflect the fact that our lineage might not make it to the  zone on it's own but will be sucked back up if it survives without recombining back to time $\tau$} \paragraph{Perhaps a route to a better approximation?}

Taken from Barton and Etheridge spatial hitchhiking paper page 25-26:

\emph{The probability that the time for a Brownian motion with drift −c hits
zero for the first time at time $t$ starting from $x$ is}
\begin{equation}
\phi(t,x) = \frac{x}{\sigma\sqrt{(2\pi t^3)}} \exp
\left(-\frac{(x-ct)^2}{2\sigma^2 t} \right)
\end{equation}
\emph{see e.g. Cox and Miller (1977) p.221, Equation 74}
\href{http://books.google.com/books?hl=en&lr=&id=76QOAAAAQAAJ&oi=fnd&pg=PR9&dq=The+Theory+of+Stochastic+Processes+cox&ots=cBTJktzRam&sig=kc_3OsLxXMyVRXx9lIUJCCAEys4#v=onepage&q=The%20Theory%20of%20Stochastic%20Processes%20cox&f=false}{link}
.\emph{ Notice that, at
least if $x \gg \sigma$, this is essentially a Gaussian density, centred on $t =
x/c$, the time since the wave was at the current location of the
lineage, with standard deviation $\sigma/c$.}

In combination with the approx. linear drift \ref{unscaled_speed} we
could use this to obtain an approximate distribution on the time it
takes to reach the cline center. 
We could try to approximate prob of not recombining off bad background 
before we reach hybrid zone center. Assuming that if we recombine,
that we'll fall off the bad background.
 The probability that we are
still on the bad background by the time we reach hybrid zone center,
over a time $\tau$, is
\begin{equation}
\approx \int_0^\tau \phi(x,t) e^{-rt} dt
\end{equation}

\subsection{Covariance between ancestry in neutral non-equib. zone}

The probability of recombining at time $t$ is:
$$re^{-rt}$$
After which point the two lineages are independent with variance (and covariance) of positions after time $\tau-t$ is:
$$ (\tau-t)\sigma^2$$
We can take the integral over $t$ to get:

$$ \Var(X) = \int (\tau-t)re^{-rt} \,dt $$
Where X is the position of a lineage at time $\tau$.

Now suppose that a lineage sampled at time $0$ recombines at time $t$ when the lineage is at position $Z$. At time $\tau$ the two lineages after recombination are at positions $x + Z$ and $y + Z$, where $x$ and $y$ are the displacement during time $(\tau-t)$.
$$\Cov(x+Z,y+Z)=\Cov(x,Z)+\Cov(y,Z)+\Cov(x,y)+\Cov(Z,Z)$$
which comes down to:
	$$\Cov(Z,Z) = \Var(Z,Z) = \sigma^2 t$$
We then integrate across $t$ to get:
	$$\int \sigma^2tre^{-rt} \,dt$$
Which gives us: 
$$\sigma^2(\tau - \frac{1-e^{-r\tau}}{r})$$

We now consider the case in which recombination does not occur in time $\tau$. This happens with probability $e^{-r\tau}$. Since under this scenario, $X=Y$, $\Cov(X,Y) = \Var(X) = \sigma\tau$. This gives us a covariance of:
$$\sigma\tau e^{-r\tau} + \sigma^2(\tau - \frac{1-e^{-r\tau}}{r})$$

Now, let us take a look at the covariance between $X$ and $Y$, where $H_X$ and $H_Y$ are indicator variables such that  $H_X = 1$ if $X$ is on the RHS of the HZ, and $H_X = 0$ if $X$ is on the LHS of the HZ. Suppose that we start with a lineage sampled on the LHS of the HZ, a distance $L$ to the left of position 0 ($-L$. Under this scenario:

$$\Cov(H_X,H_Y) = 	\mathbb{E}[H_XH_Y] - \mathbb{E}[H_X][H_Y]$$
$$= \Pr(H_X = 1, H_Y=1) - \mathbb{E}[H_X]^2$$
Since $\mathbb{E}[H_X] = \mathbb{E}[H_Y]$.
$$\mathbb{E}[H_X] = \Pr(H_X = 1) = \int_L^\infty \! x\sim\mathcal{N}(0,\sigma\sqrt{t}) \,dx$$
where $x$ is the distance moved from starting point $L$. If recombination happens at time $t$ at position $-L+z$ (i.e. lineage has moved distance $z$ when recombination occurs)
$$\Pr(H_X = 1, H_Y=1) $$
$$= \int_0^\tau  re^{-rt}\int_{-\infty}^\infty z\sim\mathcal{N}(0,\sigma\sqrt{t})  \iint \limits_{L-z}^\infty   (x\sim\mathcal{N}(0,\sigma\sqrt{\tau - t}))  (y\sim\mathcal{N}(0,\sigma\sqrt{\tau - t}))                             \,dxdy  \, dz \, dt$$

%\gc{So I'll lay out here how  started to think about this, and then  let you work it through the remains of the way. I think this helps  with both cases. We imagine we start out at location zero, and that the  geographic location of our lineages at our two genomic locations at time  $\tau$ are X and Y. With prob. $e^{-r\tau}$ $X=Y$. Then let's consider case where we recombine at time $t$ (which occurs  with prob. $re^{rt}$), at a location Z (which occurs with prob. $Z  \sim N(0,\sigma^2 t)$). What is $E(XY | t,Z)$?  Then we can integrate out over t and Z }

%\gc{I see where you are going with this but I think there is a slip  over where Z is in the above example}

This integral doesn't converge numerically in R or mathematica. We can simplify the expression to a single integral by expressing the normal distributions as a bivariate normal with vac-cov matrix:

\[\left (\begin{array}{cc}
\tau\sigma^2 & t\sigma^2\\
t\sigma^2 & \tau\sigma^2\end{array}\right)\]

This reduces the quadruple integral to:
\begin{equation}
\int_0^\tau r\tau e^{-r\tau \frac{t}{\tau}}\int_{L}^\infty \int_L^\infty f_t(x,y) \,dxdy \,dt
\end{equation}

Where 
$$
f(x,y)= \frac{1}{2\pi\tau\sigma^2\sqrt{1-(\frac{t}{\tau})^2}}\exp\left(-\frac{x^2-2\frac{t}{\tau}xy-y^2}{2\tau\sigma^2(1-(\frac{t}{\tau})^2)}\right)$$

Equation 26 is then
%\[\begin{array}{rcl}
\begin{equation}
 e^{-r t}\int_L^\infty\int_L^\infty f_t(x,y)\,dxdy|_{t=0}^{t=\tau} + \int_0^\tau e^{-r\tau t}\int_L^\infty\int_L^\infty\frac{\partial f_t}{\partial t} \,dydx \,dt
 \end{equation}
%\end{array}\]

We can calculated the double integral in the first term of equation 27 by interpreting that $t=0$ means that recombination has happened at time $0$ and the probabilities of  both lineages moving $L$ units are independent. Similarly, when $t=1$, recombination does not happen until time $\tau$ and so we only need the probability that one lineage has moved $L$ units. This gives us.
\begin{equation}
\left(1-\Phi\left(\frac{L}{\sqrt{\tau}\sigma}\right)\right)^2 - e^{-r\tau}\left(1-\Phi\left( \frac{L}{\sqrt{\tau}\sigma}\right)\right)
\end{equation}

To solve the second term of equation 27, we use the fact that 
$$\frac{\partial f_t}{\partial t} = \frac{\partial^2f_t}{\partial x \partial y}$$

Which gives us 
\[\begin{array}{rcl}
&&\int_0^\tau e^{-rt}f_t(L,L) \,dt\\
&&\\
&=& \int_0^\tau e^{-rt}\frac{1}{2\pi\tau\sigma^2\sqrt{1-(\frac{t}{\tau})^2}}\exp\left(-\frac{L^2(\frac{t}{\tau}-1)}{\tau\sigma^2(1-(\frac{t}{\tau})^2)}\right) \,dt
\end{array}\]
This integral converges with no trouble in R. Finally, since:

$$\Cov(X,Y) = (1-e^{-r\tau})\mathbb{E}[XY|\textrm{no rec.}] - (1-e^{-r\tau})\mathbb{E}[X| \textrm{no rec.}]^2 + e^{-r\tau}\Cov(X,X|\textrm{rec.})$$

We assemble all components of this to get:
\[\begin{array}{rl}
\Cov(X,Y)  = &  \int_0^\tau e^{-rt}\frac{1}{2\pi\tau\sigma^2\sqrt{1-(\frac{t}{\tau})^2}}\exp\left(-\frac{L^2(\frac{t}{\tau}-1)}{\tau\sigma^2(1-(\frac{t}{\tau})^2)}\right) \,dt \\
& + \left(1-\Phi\left(\frac{L}{\sqrt{\tau}\sigma}\right)\right)^2 - e^{-r\tau}\left(1-\Phi\left( \frac{L}{\sqrt{\tau}\sigma}\right)\right)\\
& - (1-e^{-r\tau})\left( 1 - \Phi(\frac{L}{\sigma\sqrt{\tau}})\right)\\
& + e^{-r\tau}\left( 1 - \Phi\left(\frac{L}{\sigma\sqrt{\tau}}\right)\right)\Phi\left(\frac{L}{\sigma\sqrt{\tau}}\right)
\end{array}\]

\subsection{With selection:}

Now consider two linked loci on a 'bad' background. If recomb. occurs on the LHS, one of the loci will end up on the 'correct' background, and the other will stay on the 'bad' one (remember factor of 2!). If, on the other hand, recomb happens on the RHS, both loci will be on what is now the 'correct' background.

On a stable cline, the distance traveled in time $t$ becomes

$$L_t \sim \mathcal{N}(tc,t\sigma^2)$$

If recombination occurs on the LHS and locus 1($y_1$) recombines onto the 'good' bg, then the distribution of locus $y_1$ is $\mathcal{N}(-tc,t\sigma^2)$ and the distribution of $y_2$ is $\mathcal{N}(tc,t\sigma^2)$

Not quite sure about how to get to a covariance with conditioning here....

Similar logic can be applied to the binary scenario. Is this on the right track??

\end{document}

### 2015-09-09 notes/brownian-local-times.Rmd
.

---
title: "Brownian local times"
date: "`r date()`"
---

$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
$

It states in [Takacs](https://projecteuclid.org/euclid.aoap/1177004703) that 
if $\tau(x)$ is the local time at $x$ for Brownian motion run for time 1 and begun at 0,
i.e.,
$
\begin{align*}
    \tau(x) = \lim_{\epsilon \to 0} \frac{1}{\epsilon}
        \text{measure}\{t \,:\, x \le B_t \le x+\epsilon, \, 0 \le t \le 1\}
\end{align*}
$
then
$
\begin{align*}
    \P\{ \tau(x) \le u \} = 2 \Phi(|x|+u)-1,
\end{align*}
$
where $\Phi(x)=\P\{B_1 \le x\}$ is the Gaussian CDF.

Let's check this:
```{r sim_rw}
nwalks <- 1e4
ntimes <- 1e4
# want total time to be 1, so:
eps <- 1/sqrt(ntimes)
rws <- matrix( rnorm(nwalks*ntimes)*eps, nrow=ntimes, ncol=nwalks )
rws <- rbind(0,apply(rws,2,cumsum))
matplot(rws[,1:20],type='l', ylab='position')
```
Here we've simulated `r nwalks` random walks for a total of `r ntimes` steps,
and scaled steps to have this be a total Brownian time of 1.
Let's check:
```{r final_dist}
xh <- hist(rws[ntimes,],breaks=50)
lines( xh$mids, sum(xh$counts)*diff(pnorm(xh$breaks)) )
```

Now, let's check on the local time formula.
The number of steps the random walk spends at position $k$ on a grid with spacing $\epsilon$
and a time step of \gamma
should be $\epsilon \times \tau(k\epsilon) / \gamma$.
First, the local time at 0:
```{r local_times_0}
xh <- hist(colSums(abs(rws)<=eps)/ntimes/(2*eps),main="local time at 0", breaks=30)
lines( xh$mids, sum(xh$counts)*diff( 2*pnorm(xh$breaks)-1 ), col='red' )
```
What about the local time at 1/2?
Note there is an atom at 0, now:
```{r local_times}
layout(matrix(1:4,nrow=2))
for (loc in c(0.1, 0.2, 0.4, 0.6)) {
    lts <- colSums(abs(rws-loc)<=eps)/ntimes/(2*eps)
    zeros <- (lts==0)
    pred.zero <- 2*pnorm(abs(loc))-1
    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=c(0,pred.zero*nwalks))
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc)+xh$breaks)-1 ), col='red' )
    points(c(0,0), c(sum(zeros),pred.zero*nwalks), col=1:2, pch=20, cex=2)
}
```

Its Laplace transform
=====================

Now suppose that something happens at rate $r$ relative to the local time at $x$
(e.g., recombination to the other background).
Then the chance that this has not happened by time 1 is
$\begin{align*}
\E\left[ e^{-r\tau(x)} \right]
&=
\int_0^\infty e^{-rt} 2 \phi(x+t) dt \\
&=
e^{rx} \int_x^\infty e^{-rt} 2 \phi(t) dt \\
&=
2 e^{rx} \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - rt - t^2/2 \right) dt \\
&=
2 e^{rx+r^2/2} \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - (t+r)^2/2 \right) dt \\
&=
2 e^{rx+r^2/2} 2 \left( 1-\Phi(x+r) \right) .
\end{align*}$

### 2015-09-10 notes/brownian-local-times.Rmd
.

@@
 
-$
+$$
 \newcommand{\E}{\mathbb{E}}
 \renewcommand{\P}{\mathbb{P}}
-$
+\DeclareMathOperator{\sgn}{sgn}
+$$
+
+```{r doc_setup, include=FALSE}
+fig.dim <- 5
+knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
+```
+
+Killing at Brownian local times
+===============================
 
@@
 i.e.,
-$
+$$
 \begin{align*}
@@
 \end{align*}
-$
+$$
 then
-$
+$$
 \begin{align*}
@@
 \end{align*}
-$
+$$
 where $\Phi(x)=\P\{B_1 \le x\}$ is the Gaussian CDF.
@@
 Let's check this:
-```{r sim_rw}
+```{r sim_rw, cache=TRUE}
 nwalks <- 1e4
-ntimes <- 1e4
+ntimes <- 1e3
 # want total time to be 1, so:
@@
 First, the local time at 0:
-```{r local_times_0}
-xh <- hist(colSums(abs(rws)<=eps)/ntimes/(2*eps),main="local time at 0", breaks=30)
+```{r local_times_0, cache=TRUE, depends="sim_rw"}
+xh <- hist(colSums(abs(rws)<=1.5*eps)/ntimes/(3*eps),main="local time at 0", breaks=30)
 lines( xh$mids, sum(xh$counts)*diff( 2*pnorm(xh$breaks)-1 ), col='red' )
 ```
-What about the local time at 1/2?
-Note there is an atom at 0, now:
-```{r local_times}
+What about the local time at other locations?
+Note there is an atom at 0, now,
+and that the discretization could trip us up;
+the colors (red/green/blue) use the (mid/upper/lower) points of the interval that we count in.
+```{r local_times, cache=TRUE, depends="sim_rw"}
 layout(matrix(1:4,nrow=2))
 for (loc in c(0.1, 0.2, 0.4, 0.6)) {
-    lts <- colSums(abs(rws-loc)<=eps)/ntimes/(2*eps)
+    lts <- colSums(abs(rws-loc)<=1.5*eps)/ntimes/(3*eps)
     zeros <- (lts==0)
-    pred.zero <- 2*pnorm(abs(loc))-1
-    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=c(0,pred.zero*nwalks))
+    pred.zero <- 2*pnorm(abs(loc)+c(0,1.5*eps,-1.5*eps))-1
+    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=range(c(0,pred.zero*nwalks)))
     lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc)+xh$breaks)-1 ), col='red' )
-    points(c(0,0), c(sum(zeros),pred.zero*nwalks), col=1:2, pch=20, cex=2)
+    points(rep(0,4), c(sum(zeros),pred.zero*nwalks), col=1:4, pch=20, cex=2)
+    # lower bound
+    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc+1.5*eps)+xh$breaks)-1 ), col=3 )
+    # upper bound
+    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc-1.5*eps)+xh$breaks)-1 ), col=4 )
 }
 ```
+Clearly, the discretization is still tripping us up a bit,
+but it's pretty good.
 
@@
 Then the chance that this has not happened by time 1 is
-$\begin{align*}
+$$
+\begin{align*}
 \E\left[ e^{-r\tau(x)} \right]
 &=
-\int_0^\infty e^{-rt} 2 \phi(x+t) dt \\
+\int_0^\infty e^{-rt} 2 \phi(|x|+t) dt \\
 &=
-e^{rx} \int_x^\infty e^{-rt} 2 \phi(t) dt \\
+e^{r|x|} \int_{|x|}^\infty e^{-rt} 2 \phi(t) dt \\
 &=
-2 e^{rx} \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - rt - t^2/2 \right) dt \\
+2 e^{r|x|} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - rt - t^2/2 \right) dt \\
 &=
-2 e^{rx+r^2/2} \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - (t+r)^2/2 \right) dt \\
+2 e^{r|x|+r^2/2} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - (t+r)^2/2 \right) dt \\
 &=
-2 e^{rx+r^2/2} 2 \left( 1-\Phi(x+r) \right) .
-\end{align*}$
+2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
+\end{align*}
+$$
+At $r=0$ this should equal 1, since it's the chance that nothing happens.
+Same for $|x| \to \infty$.
+Note that at $x=r=0$ this is equal to 1, as it should be.
+Also, for large $x$, since $1-\Phi(x) \sim e^{-x^2/2}/\sqrt{2\pi}x$,
+this is approximately 
+$2 \exp( (r+1)|x|+r^2/2 )/\sqrt{2\pi}(|x|+r}$,
+and, of course, goes to zero.
+
+Let's check that, too: why not?
+```{r killing, cache=TRUE, depends="sim_rw"}
+loc <- 0.1
+compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
+        visits <- colSums(abs(rws-loc)<=1.5*eps)
+        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) > 0 )
+        c( 
+          rate=rate,
+          simulation=mean(killed), 
+          theory=1-2 * exp(rate*loc+rate^2/2) * pnorm(loc+rate,lower.tail=FALSE),
+          SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
+          )
+    } ) ) )
+with(compare.rates, {
+         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=range(c(simulation,theory)) ) 
+         points( rate, simulation, pch=20 )
+         arrows( x0=rate, y0=simulation-2*SE, y1=simulation+2*SE, angle=90, code=3 )
+    } )
+```
+Hm, again, something probably in the discretization is throwing us off a bit.
+
+What this says about clines
+===========================
+
+The above restricted to standard Brownian motion on one unit of time.
+Really, we will want to use this across $\tau$ generations,
+modeling a random walk with variance $\sigma^2$ per generation.
+
+The key quantity we want is the probability that a Brownian motion with variance $\sigma^2$ killed at rate $r$
+with respect to its local time at $x$ was still alive at time $\tau$.
+Denote this by
+$$
+\begin{align*}
+    \zeta(x,r,\tau,\sigma).
+\end{align*}
+$$
+By Brownian scaling,
+since $B_{ct}$ has the same distribution as $\sqrt{c}B_t$,
+$$
+\begin{align*}
+    \zeta(x,r,\tau,\sigma) 
+    &=
+    \zeta(x,r\tau,1,\sqrt{\tau}\sigma) \\
+    &=
+    \zeta(x/\sqrt{\tau\sigma^2},r\tau,1,1) \\
+    &=
+    2 \exp\left(r\tau\left(\frac{|x|}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{|x|}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) .
+\end{align*}
+$$
+
+Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
+at a location at recombination distance $r$ from the selected site
+inherits from the left-hand side of the barrier
+(i.e., an ancestor on $(-\infty,0)$)
+$\tau$ generations ago
+is the probability no recombination at the barrier has occurred
+plus one-half the probability recombination has occurred.
+This is
+$$
+\begin{align*}
+    p(x,r,\tau,A) = % \zeta 1_{x\le0} + (1-\zeta)/2 = (1-(1-2*1_{x\le0})\zeta)/2
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
+    % \frac{1}{2}\left\{ 1 - \sgn(x) 2 \exp\left(r\tau\left(\frac{x}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{x}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) \right\}
+\end{align*}
+$$
+
+Does this actually make a cline?
+```{r show_cline}
+zeta <- function (x,r,tau=1,sigma=1) {
+    sign(x)*2 * exp(r*tau*(abs(x)/sqrt(tau*sigma)+r*tau/2) + pnorm(abs(x)/sqrt(tau*sigma^2)+r*tau,lower.tail=FALSE,log.p=TRUE))
+}
+pcline <- function (...) {
+    (1/2)*( 1 - zeta(...) )
+}
+xx <- seq(-5,5,length.out=500)
+plot(xx, pcline(xx,r=1), type='l' )
+```

### 2015-09-10 notes/brownian-local-times.Rmd
.

@@
 where $\Phi(x)=\P\{B_1 \le x\}$ is the Gaussian CDF.
+Differentiating this, $\P\{\tau \in du\} = \partial_u \P\{\tau \le u\}$, so
+$$
+\begin{align*}
+    \P\{ \tau(x) = 0 \} &= 2 \Phi(|x|)-1 \\
+    \P\{ \tau(x) \in du \} &= 2 \phi(|x|+u) ,
+\end{align*}
+$$
+where $\phi$ is the Gaussian density.
 
@@
 &=
+2\Phi(|x|)-1 + 
 \int_0^\infty e^{-rt} 2 \phi(|x|+t) dt \\
 &=
+2\Phi(|x|)-1 + 
 e^{r|x|} \int_{|x|}^\infty e^{-rt} 2 \phi(t) dt \\
 &=
+2\Phi(|x|)-1 + 
 2 e^{r|x|} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - rt - t^2/2 \right) dt \\
 &=
+2\Phi(|x|)-1 + 
 2 e^{r|x|+r^2/2} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - (t+r)^2/2 \right) dt \\
 &=
+2\Phi(|x|)-1 + 
 2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
@@
 $$
-At $r=0$ this should equal 1, since it's the chance that nothing happens.
-Same for $|x| \to \infty$.
-Note that at $x=r=0$ this is equal to 1, as it should be.
-Also, for large $x$, since $1-\Phi(x) \sim e^{-x^2/2}/\sqrt{2\pi}x$,
+At $r=0$ this equals 1, as it should, since it's the chance that nothing happens.
+For large $x$, since $1-\Phi(x) \sim e^{-x^2/2}/\sqrt{2\pi}x$,
 this is approximately 
-$2 \exp( (r+1)|x|+r^2/2 )/\sqrt{2\pi}(|x|+r}$,
-and, of course, goes to zero.
+$1 + 2 \exp( r|x| + r^2/2 - (|x|+r)^2/2 )$,
+which goes to 1, as it should.
 
@@
 loc <- 0.1
+zeta <- function (x,r) {
+    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
+}
 compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
         visits <- colSums(abs(rws-loc)<=1.5*eps)
-        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) > 0 )
+        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) == 0 )
         c( 
@@
           simulation=mean(killed), 
-          theory=1-2 * exp(rate*loc+rate^2/2) * pnorm(loc+rate,lower.tail=FALSE),
+          theory=zeta(loc,rate),
           SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
@@
 with(compare.rates, {
-         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=range(c(simulation,theory)) ) 
+         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=range(c(simulation,theory)), ylim=c(0,1) ) 
          points( rate, simulation, pch=20 )
@@
 ```
-Hm, again, something probably in the discretization is throwing us off a bit.
 
@@
     &=
-    \zeta(x/\sqrt{\tau\sigma^2},r\tau,1,1) \\
+    \zeta(x/\sqrt{\tau\sigma^2},r\tau,1,1) ,
+    % &=
+    % 2 \exp\left(r\tau\left(\frac{|x|}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{|x|}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) .
+\end{align*}
+$$
+and
+$$
+\begin{align*}
+    \zeta(x,r,1,1)
     &=
-    2 \exp\left(r\tau\left(\frac{|x|}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{|x|}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) .
+    2\Phi(|x|)-1 + 
+    2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
 \end{align*}
@@
 ```{r show_cline}
-zeta <- function (x,r,tau=1,sigma=1) {
-    sign(x)*2 * exp(r*tau*(abs(x)/sqrt(tau*sigma)+r*tau/2) + pnorm(abs(x)/sqrt(tau*sigma^2)+r*tau,lower.tail=FALSE,log.p=TRUE))
-}
 pcline <- function (...) {

### 2015-09-10 notes/brownian-local-times.Rmd
.

@@
 with(compare.rates, {
-         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=range(c(simulation,theory)), ylim=c(0,1) ) 
+         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=c(0,1) ) 
          points( rate, simulation, pch=20 )
@@
 ```
+
+This is very steep at $x=0$.
+Notice that as $x$ goes to 0,
+$$
+\begin{align*}
+    \lim_{x \to 0} \zeta(x,r) 
+    &= 2 e^{r^2/2} (1-\Phi(r)) \\
+    &= \frac{2}{r\sqrt{2\pi}} + O(r^{-3}) .
+\end{align*}
+$$
+so that the difference in the left and right limits is
+$$
+\begin{align*}
+    \lim_{x \to 0-} p(x,r) 
+    -\lim_{x \to 0+} p(x,r) 
+    &= 
+    \frac{1}{2}+\frac{2}{r\sqrt{2\pi}} 
+    - \frac{1}{2}-\frac{2}{r\sqrt{2\pi}} 
+      +O(r^{-3}) \\
+    &=
+    \frac{4}{r\sqrt{2\pi}}
+      +O(r^{-3}) 
+\end{align*}
+$$
+
+Let's also look at the slope of $p$ at $x=0$.
+Since $\partial_x |x| = \sgn(x)$,
+$$
+\begin{align*}
+    \partial_x \zeta(x,r)
+    &=
+    2 \sgn(x) \left(
+    \phi(|x|)
+    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
+    - e^{r|x|+r^2/2} \phi(|x|+r)
+      \right)
+\end{align*}
+$$
+and so
+$$
+\begin{align*}
+    -\partial_x p(x)
+    &=
+    \phi(|x|)
+    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
+    - e^{r|x|+r^2/2} \phi(|x|+r) \\
+    &\to
+    \frac{1}{\sqrt{2\pi}}(1-e^{r^2/2}e^{-r^2/2})
+    + r e^{r^2/2} (1-\Phi(r))  \qquad \text{as } x \to 0 \\
+    &=
+    r e^{r^2/2} (1-\Phi(r)) \\
+    &=
+    \frac{1}{\sqrt{2\pi}} + O(\frac{1}{r}) .
+\end{align*}
+$$
 
@@
     &=
-    2\Phi(|x|)-1 + 
+    1 - 2(1-\Phi(|x|)) + 
     2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
@@
 (i.e., an ancestor on $(-\infty,0)$)
-$\tau$ generations ago
-is the probability no recombination at the barrier has occurred
-plus one-half the probability recombination has occurred.
-This is
+$\tau$ generations ago.
+For an individual to the left of the barrier ($x<0$),
+this is the probability no recombination at the barrier has occurred
+plus one-half the probability recombination has occurred,
+i.e., $\zeta+(1-\zeta)/2=(1+\zeta)/2$.
+If $x>0$ this is the one-half times probability a recombination has occurred,
+i.e., $(1-\zeta)/2$.
+Combining these,
+the probability of ending up on the left-hand side is
 $$
 \begin{align*}
-    p(x,r,\tau,A) = % \zeta 1_{x\le0} + (1-\zeta)/2 = (1-(1-2*1_{x\le0})\zeta)/2
+    p(x,r,\tau,A) = 
     \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
-    % \frac{1}{2}\left\{ 1 - \sgn(x) 2 \exp\left(r\tau\left(\frac{x}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{x}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) \right\}
 \end{align*}
@@
 ```{r show_cline}
-pcline <- function (...) {
-    (1/2)*( 1 - zeta(...) )
+pcline <- function (x,r,tau=1,sigma=1) {
+    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
 }
 xx <- seq(-5,5,length.out=500)
-plot(xx, pcline(xx,r=1), type='l' )
+plot(xx, pcline(xx,r=1), type='l', xlab="distance from cline center", ylab="prob of inherited from the left" )
+rr <- seq(0,2,length.out=11)
+pp <- sapply( rr, function (r) { pcline(xx,r=r) } )
+matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
+legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
 ```

### 2015-09-11 notes/brownian-local-times.Rmd
added density of hets

@@
 It states in [Takacs](https://projecteuclid.org/euclid.aoap/1177004703) that 
-if $\tau(x)$ is the local time at $x$ for Brownian motion run for time 1 and begun at 0,
+if $\ell(x)$ is the local time at $x$ for Brownian motion run for time 1 and begun at 0,
 i.e.,
@@
 \begin{align*}
-    \tau(x) = \lim_{\epsilon \to 0} \frac{1}{\epsilon}
+    \ell(x) = \lim_{\epsilon \to 0} \frac{1}{\epsilon}
         \text{measure}\{t \,:\, x \le B_t \le x+\epsilon, \, 0 \le t \le 1\}
@@
 \begin{align*}
-    \P\{ \tau(x) \le u \} = 2 \Phi(|x|+u)-1,
+    \P\{ \ell(x) \le u \} = 2 \Phi(|x|+u)-1,
 \end{align*}
@@
 where $\Phi(x)=\P\{B_1 \le x\}$ is the Gaussian CDF.
-Differentiating this, $\P\{\tau \in du\} = \partial_u \P\{\tau \le u\}$, so
+Differentiating this, $\P\{\ell \in du\} = \partial_u \P\{\ell \le u\}$, so
 $$
 \begin{align*}
-    \P\{ \tau(x) = 0 \} &= 2 \Phi(|x|)-1 \\
-    \P\{ \tau(x) \in du \} &= 2 \phi(|x|+u) ,
+    \P\{ \ell(x) = 0 \} &= 2 \Phi(|x|)-1 \\
+    \P\{ \ell(x) \in du \} &= 2 \phi(|x|+u) ,
 \end{align*}
@@
 and a time step of \gamma
-should be $\epsilon \times \tau(k\epsilon) / \gamma$.
+should be $\epsilon \times \ell(k\epsilon) / \gamma$.
 First, the local time at 0:
@@
 \begin{align*}
-\E\left[ e^{-r\tau(x)} \right]
+\E\left[ e^{-r\ell(x)} \right]
 &=
@@
 
+Also note that for large $|x|$, by the asymptotics for $1-\Phi(x)$,
+$$
+\begin{align*}
+    \zeta(x,r) = 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) .
+\end{align*}
+$$
+
 What this says about clines
@@
     &=
-    \zeta(x/\sqrt{\tau\sigma^2},r\tau,1,1) ,
-    % &=
-    % 2 \exp\left(r\tau\left(\frac{|x|}{\sqrt{\tau}\sigma}+\frac{r\tau}{2}\right)\right) \left( 1-\Phi\left(\frac{|x|}{\sqrt{\tau\sigma^2}}+r\tau \right) \right) .
+    \zeta(x/\sqrt{\tau\sigma^2},r\tau) ,
 \end{align*}
 $$
-and
+where
 $$
 \begin{align*}
-    \zeta(x,r,1,1)
+    \zeta(x,r)
     &=
@@
 
+**Cline flattening:**
+immediately, this tells us how a cline gets wider as time progresses:
+for a fixed $r$, the width grows as $\sqrt{\tau}$.
+It also tells us that clines with the same $r\tau$ have the same shape,
+after rescaling the width by $\sqrt{\tau}$.
+Also, this tells us we only have a cline for $r$ of order at least $1/\tau$:
+for smaller $r$, the locus is still completely linked to the selected locus,
+and hence a step function.
+
+More generally, here's what $\zeta$ looks like:
+```{r plot_zeta, fig.height=2*fig.dim}
+layout(1:2)
+xvals <- seq(0,5,length.out=150)
+rvals <- seq(0,10,length.out=100)
+zetavals <- outer( xvals, rvals, zeta )
+matplot(xvals,zetavals,type='l',xlab="x",ylab=expression(zeta),col=rainbow(length(rvals)))
+lrvals <- seq(1,length(rvals),length.out=6)
+legend("bottomright",lty=1,col=rainbow(length(rvals))[lrvals],legend=paste("r=",round(rvals[lrvals],digits=2)))
+matplot(rvals,t(zetavals),type='l',xlab="r",ylab=expression(zeta),col=rainbow(length(xvals)))
+lxvals <- seq(1,length(xvals),length.out=6)
+legend("bottomright",lty=1,col=rainbow(length(xvals))[lxvals],legend=paste("x=",round(xvals[lxvals],digits=2)),bg='white')
+```
+
 Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
@@
 $\tau$ generations ago.
+Let $p_H$ denote the density of heterozygotes in the cline
+-- this should be $p_H = (1-s)/2$).
 For an individual to the left of the barrier ($x<0$),
-this is the probability no recombination at the barrier has occurred
-plus one-half the probability recombination has occurred,
-i.e., $\zeta+(1-\zeta)/2=(1+\zeta)/2$.
-If $x>0$ this is the one-half times probability a recombination has occurred,
-i.e., $(1-\zeta)/2$.
+the probability of inheriting from the left of the barrier at $\tau$
+is the probabiltiy that no recombination at the barrier has occurred,
+plus one-half the probability that recombination in a heterozygote has occurred,
+i.e., $\zeta p_H +(1-\zeta p_H)/2=(1+\zeta p_H)/2$.
+If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
+i.e., $(1-\zeta p_H)/2$.
 Combining these,
@@
     p(x,r,\tau,A) = 
-    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) p_H \right\} .
 \end{align*}
@@
 ```{r show_cline}
-pcline <- function (x,r,tau=1,sigma=1) {
-    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
+pcline <- function (x,r,tau=1,sigma=1,pH=1/2) {
+    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau)*pH )
 }
-xx <- seq(-5,5,length.out=500)
-plot(xx, pcline(xx,r=1), type='l', xlab="distance from cline center", ylab="prob of inherited from the left" )
-rr <- seq(0,2,length.out=11)
-pp <- sapply( rr, function (r) { pcline(xx,r=r) } )
+xx <- seq(-15,15,length.out=500)
+for (tau in c(5,50)) {
+    plot(xx, pcline(xx,r=1), type='n', 
+            xlab="distance from cline center", ylab="prob of inherited from the left", 
+            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
+    rr <- seq(0,0.2,length.out=11)
+    pp <- sapply( rr, function (r) { pcline(xx,r=r,tau=tau) } )
     matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
     legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
+}
 ```

### 2015-09-11 notes/lineage-movement.tex
notes on movement in a skinny zone

\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\deriv}[1]{\frac{d}{d#1}}
\newcommand{\dderiv}[1]{\frac{d^2}{d^2#1}}
\newcommand{\given}{\;\vert\;}

\begin{document}

\section{Discrete model}

Consider a discrete model in which the total number of individuals at location $x$ is $N(x)$,
each individual is of type either B or b at a selected locus,
and the number of individuals of type B and location $x$ and time $t$ is $n(x,t)$
(but, we often neglect the $t$).
Suppose that type B individuals at location $x$ reproduce at rate $s_B(x)$, 
and likewise type b at rate $s_b(x)$.
At reproduction, individuals recombine with others in the same location,
and the offspring choose a new location according to the kernel $m(x,y)$.
The population dynamics are random, but suppose that $N(x)$ is sufficiently large that these do not vary with time;
instead we want to follow a lineage backwards in time,
at a locus separated from the selected locus by recombination fraction $r$.

Suppose this is a Moran model.
There are four things that can happen:
\begin{enumerate}
    \item[$x\xrightarrow{BB}y$] One type B individual at location $x$ reproduces, 
        either does not recombine or recombines with another type B,
        and sends the offspring to $y$.
    \item[$x\xrightarrow{Bb}y$] An individual at location $x$ reproduces, 
        recombines with the other type,
        and sends to $y$ an offspring
        who inherits at the selected locus from the type b parent 
        and the at the neutral locus from the type B parent.
    \item[$x\xrightarrow{bB}y$] An individual at location $x$ reproduces, 
        recombines with the other type,
        and sends to $y$ an offspring
        who inherits at the selected locus from the type B parent 
        and the at the neutral locus from the type b parent.
    \item[$x\xrightarrow{bb}y$] One type b individual at location $x$ reproduces, 
        either does not recombine or recombines with another type b,
        and sends the offspring to $y$.
\end{enumerate}
Let $p(x) = n(x)/N(x)$.
These four things happen at rates:
\begin{align}
    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= p(x) s_B(x) \left(1 - r (1-p(x)) \right)  N(x) m(x,y) \\
    & x\xrightarrow{Bb}y & \qquad w_{Bb}(x,y) &= r p(x) (1-p(x)) \frac{s_B(x)+s_b(x)}{2} N(x) m(x,y) \\
    & x\xrightarrow{bB}y & \qquad w_{bB}(x,y) &= r p(x) (1-p(x)) \frac{s_B(x)+s_b(x)}{2} N(x) m(x,y) \\
    & x\xrightarrow{bb}y & \qquad w_{bb}(x,y) &= (1-p(x)) s_b(x) \left(1 - r p(x) \right) N(x) m(x,y) 
\end{align}

Note that at equilibrium, we require
\begin{align}
  0 = \sum_y N(y) m(y,x) \left\{ p(y)s_B(y)(1-p(x)) - (1-p(y))s_b(y)p(x) \right\} .
\end{align}

\subsection{Lineage movement}

These rates tell us the rates at which a lineage will move, backwards in time.
For instance, the rate at which a lineage at the selected locus
currently in a type B individual at location $x$
jumps to another type B individual at location $y$ is equal to the rate of influx of migrants from $y$
divided by the number of B alleles at $x$,
or
\begin{align}
  r_A(x,y) &= \frac{w_{BB}(y,x)}{p(x)N(x)} \\
  &= N(y) p(y)  s_B(y) m(y,x) \frac{ 1 }{ N(x) p(x) }
\end{align}

If we follow a linked locus until the first time it is found in a type b individual,
it moves according to these jump rates, and at location $x$ recombines at rate
\begin{align}
  \gamma_B(x) &= \frac{1}{N(x)p(x)}\sum_y w_{bB}(y,x)  \\
  &= r p(y) (1-p(y)) \frac{s_B(y)+s_b(y)}{2} N(y) m(y,x) \frac{1}{N(x)p(x)}.
\end{align}
Suppose that we kill the lineage when it recombines onto type b.
Let $X_t$ denote the position of the lineage at time $t$ in the past,
with $X_t = \rho$ if it has recombined onto b,
and let $f$ be test function with $f(\rho)=0$.
If we let $M_r(y) = s_B(y) \left(1 - r (1-p(y)) \right)$, then
\begin{align}
    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y r_{B}(y,x) ( f(y)-f(x) ) - \gamma_B(x) f(x) \\
    &= \frac{1}{N(x)p(x)} \sum_y  N(y) p(y) M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_B(x) f(x) . \label{eqn:discrete_generator}
\end{align}

\section{Diffusion limit}

Now suppose that $m(x,y)$ is symmetric, and depends on a parameter $\sigma$ so that as $\sigma \to 0$,
the associate random walk converges to Brownian motion, so that
\begin{align}
    \lim_{\sigma \to 0} \sum_y \frac{ m(x,y) ( f(y) - f(x) ) }{\sigma^2} = \frac{1}{2} \dderiv{x} f(x) .
\end{align}
Write $f'(x) = \deriv{x}f(x)$, and note that
\begin{align}
    \frac{1}{\sigma^2} \sum_y g(y) m(y,x) (f(y)-f(x)) 
    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) + (g(x)-g(y)) f(x) \right) \\
    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) \right) \\
    & \qquad - f(x) \frac{1}{\sigma^2} \sum_y m(y,x) (g(y)-g(x)) \\
    &\xrightarrow{\sigma \to 0} \frac{1}{2} \dderiv{x}\left( g(x)f(x) \right) - \frac{1}{2} f(x) \dderiv{x} g(x) \\
    &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) \\
    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) . \label{eqn:deriv_limit}
\end{align}

If we also rescale recombination, so that $\rho = r/\sigma^2$,
then 
\begin{align}
    \frac{1}{\sigma^2} \gamma_B(x) \xrightarrow{\sigma \to 0} \rho (1-p(x)) \frac{s_B(x)+s_b(x)}{2} ,
\end{align}
which we can define to be equal to $\rho k(x)$.
In this case, note that since $r = \rho \sigma^2$,
\begin{align}
    M_r(x) \xrightarrow{\sigma \to 0} s_B(x) .
\end{align}
Under these assumptions, combining \eqref{eqn:discrete_generator} and \eqref{eqn:deriv_limit},
\begin{align}
    \deriv{t} \E[f(X_{t/\sigma^2}) \given X_0=x ] &\xrightarrow{\sigma \to 0} 
    \frac{1}{2} s_B(x) \dderiv{x} f(x) + \frac{1}{N(x)p(x)} \deriv{x} \left\{ N(x)p(x) s_B(x) \right\}  \deriv{x} f(x) - \rho k(x) f(x) \\
    &= \frac{1}{2} s_B(x) f''(x) + \left( s_B'(x) + \log(N(x)p(x))' s_B(x) \right) f'(x)  - \rho k(x) f(x) ,
\end{align}
i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with drift $\frac{1}{ N(x)p(x)} \deriv{x} ( N(x) p(x) s_B(x) )$ and killed at rate $\rho k(x)$.

\paragraph{Stationary distribution}
If $X$ has a stationary distribution then its density $\pi(x)$ solves the forward Kolmogorov equation,
which is here (letting $n(x) = N(x) p(x)$ for convenience)
\begin{align}
  0 &= \frac{1}{2} \dderiv{x} \left( s_B(x) \pi(x) \right) - \deriv{x} \left( \frac{\pi(x)}{n(x)} \deriv{x}(n(x) s_B(x)) \right) .
\end{align}
It is easy to check that $\pi(x) \propto n(x)^2 s(x) $ is a solution,
and since it is nonnegative, the stationary distribution,
i.e.
\begin{align}
  \pi(x) = \frac{ \left( N(x) p(x) \right)^2 s_B(x) }{ \int_y \left( N(y) p(y) \right)^2 s_B(y) dy } ,
\end{align}
if the denominator (below denoted by $Z$) is finite.

In other words, the stationary distribution is proportional to the local rate of production of $B$ alleles, \emph{squared}.

\paragraph{Speed and scale}
It will be useful to think about the Dubins-Schwartz construction of this diffusion.
Note that if we define
\begin{align}
  z(x) = \int_{0}^x \frac{ dy }{ n(y)^2 s_B(y)^2 },
\end{align}
where $n(x) = N(x) p(x)$ as before,
then
\begin{align}
  \deriv{t} \E^x[z(X_t)] &= \frac{1}{2} s_B(x) \deriv{x} \frac{1}{ n(x)^2 s_B(x)^2 } + \frac{1}{n(x)} \frac{ \deriv{x}(n(x)s_B(x)) }{  n(x)^2 s_B(x)^2 } \\
  &= - s_B(x) \frac{ \deriv{x} (n(x)s_B(x)) }{ n(x)^3 s_B(x)^3 } + \frac{1}{n(x)} \frac{ \deriv{x}(n(x)s_B(x)) }{  n(x)^2 s_B(x)^2 } \\
  &= 0 ,
\end{align}
that $Z_t = z(X_t)$ is a martingale.

Since
\begin{align}
  dX_t = \frac{\deriv{x}(n(X_t)s_B(X_t))}{n(X_t)} dt + \sqrt{s_B(X_t)} dB_t,
\end{align}
then if $s_B(x)$ is bounded below, the lineage won't get stuck,
and will always come back to the patch.
Since $z'(x) = 1/(n(x)s_B(x))^2$,
applying It\^o's formula,
\begin{align}
  dZ_t &=  z'(X_t) dX_t + \frac{1}{2} z''(X_t) s_B(X_t) dt \\
  &= \frac{ dB_t }{ n^2(X_t) s_B^{3/2}(X_t) } ,
\end{align}
i.e.\ $Z_t$ ``looks like Brownian motion moving at speed $1/(n^4(X_t) s_B^3(X_t))$''.
This happens because the transformation $z(x)$ squooshes up locations far away from the patch,
where the lineage moves at a more-or-less constant speed;
so that $Z_t$ has to move faster in those squooshed-up regions.

\paragraph{Rescaling selection out}
Suppose for the moment that $s_b(x)=1$. 
It hasn't entered here much anyhow.
First note that if we change spatial variables to
\begin{align}
  s(x) = \int_0^x \sqrt{s_B(y)} dy
\end{align}
then $ds = \sqrt{s(x)} dx$, and the generator for the process becomes
\begin{align}
  f \mapsto 
    \frac{1}{2} f''(s) + \log\left( N(s) p(s) s_B(s) \right)' f'(s) - \rho k(s) f(s) .
\end{align}
At first sight this looks like the usual scaling of $\sigma/\sqrt{s}$,
but it isn't, because in the notation I'm using there, $s_B$ is around 1,
i.e.\ $s_B = 1+s$ if $s$ is the usual selection coefficient.

\paragraph{Recombination time}
Let $\tau_\rho$ be the first time  the lineage recombines onto type b.
If $\rho$ is much smaller than the mixing time of a type B lineage,
then the time to recombination is roughly Exponential,
with mean
\begin{align}
    \E[\tau_\rho] \approx \frac{1}{ \rho \int \pi(x) k(x) dx } .
\end{align}
The product $\pi(x) k(x)$ is:
\begin{align}
    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 s_B(x) (1-p(x)) (s_B(x)+s_b(x))/2 .
\end{align}

More generally, if we define
\begin{align}
    h(x) = \E[\tau_\rho \given X_0 = x]
\end{align}
then $h$ solves the equation
\begin{align}
  -1 = \frac{1}{n(x)}(n(x)s_B(x))' h'(x) + \frac{1}{2} s_B(x) h''(x) - \rho k(x) h(x) .
\end{align}

The probability that a segment of length $r = \rho \sigma^2$ to one side of the selected locus
does not recombine across time $t$ is $\P\{\tau_\rho > t\}$.
Therefore, if $L_t$ is the length of the nonrecombined segment to one side of the selected locus
after time $t$, then
\begin{align}
    \E[ L_t \given X_0=x ] &= \sigma^2 \int_0^\infty \P\{ \tau_\rho > t \} d\rho \\
\end{align}

The appropriate thing to do is to look at $\E[L_T]$, where $T$ is the coalescence time.
Generally, then
\begin{align}
  \P\{ L_t > \rho \} = \P\{ \tau_\rho > T \} .
\end{align}
Do do this properly, we need to run two diffusions.
However, if we let $T$ instead be independent Exponential($\lambda$), then 
\begin{align}
  \ell(x) = \P\{ \tau_\rho > T \given X_0 = x\} 
\end{align}
solves the equation
\begin{align}
  \frac{1}{n(x)}(n(x)s_B(x))' h'(x) + \frac{1}{2} s_B(x) h''(x) - (\lambda + \rho k(x)) h(x) = \lambda .
\end{align}

%%%%%%%%%
\section{Integrals against $p$}

Recall that if $N(x)$ is constant, the equilibrium frequency $p(x)$ solves
\[ \frac{1}{2} p''(x) + s(x) p(x)(1-p(x)) = 0 \quad ,\]
where $s(x) = s_B(x)-s_b(x)$,
and that if $s(x) = s$ is constant on a region, the \emph{inverse} solves
\begin{align}
  \frac{dx}{dp} = \frac{1}{\sqrt{sp^2(2p/3-1) + K}},
\end{align}
where $K$ is a constant to make the boundary conditions come out right.

If this is the case, then we can change variables:
\begin{align}
  \int_a^b f(p(x)) dx = \int_{p(a)}^{p(b)} \frac{f(p) dp }{\sqrt{sp^2(2p/3-1) + K}} .
\end{align}
so for instance, 
the denominator for the stationary distribution (dropping the $N$'s)
\begin{align}
  Z &= \int p(y)^2 s_B(y) dy \\
  &= \int \frac{ p (1+s) dp }{ \sqrt{ s (2p/3-1) + K } } .
\end{align}
Note, however, that the first expression makes it clear that if we make the patch bigger, that $Z$ will increase roughly proportionally
-- since this increases the area over which $p(y)\approx 1$.
To make this occur in the integral over $p$ on the second line, 
it must be the case that $K$ is approximately $s/3$ plus a term inversely proportional to the area of the patch.

The integral of $\pi(x) k(x)$ is
\begin{align}
  \frac{1}{Z} \int_a^b p(x)^2(1-p(x))s_B(x)(s_b(x)+s_B(x)) dx = \frac{1}{Z} \int_{p(a)}^{p(b)} \frac{p (1-p) (1+s)(1+s/2) dp }{\sqrt{s(2p/3-1) + K}} .
\end{align}
I have the feeling Barton does similar things in one of his papers,
but haven't been able to track it down.

%%%%%%
\section{Haplotype lengths}

Some statements above about ``haplotype length'' aren't quite right.
Consider the haplotype shared about two B alleles;
by the time these coalesce, their ``shared haplotype'' in the above sense may be broken up into a number of blocks --
all coalescing linked to other B alleles.
More properly, the length of the contiguous haplotype
is determined by branching diffusions.
The single-lineage version is thus:
begin a diffusion with mass equal to genetic length (of something, rescaled);
then, any remaining particles split in two at rate proportional to their length,
always labeling the particle corresponding to the selected allele;
when this happens any particles other than the labeled particle may die
or may continue as an independent diffusion.

\section{Questions:} 
\begin{enumerate}

    \item Do we get the same system using, say, a Wright--Fisher model?  (I assume so?)

    \item Can we prove convergence of the finite model to this one, under suitable assumptions?  A main difficulty here would be showing that the lineage spends all its time in the part of the range where $p(x)$ does not fluctuate significantly, i.e.\ is not too small\ldots but if this is not true, it would be even more interesting.

\end{enumerate}

\end{document}

### 2015-09-11 notes/skinny-zone.Rmd
notes on movement in a skinny zone

---
title: "Lineages around a narrow hybrid zone"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator{\sgn}{sgn}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
```

Suppose that heterozygotes have fitness $(1-s)$;
we know that allele frequencies around a hybrid zone behave as
$$
\begin{align} \label{eqn:pfreq}
    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) (1-p(t,x)) (1-2p(t,x)) ,
\end{align}
$$
and from `lineage-movement.tex`,
if $A$ is the allele whose frequency is $p$,
an $A$-lineage follows a diffusion with generator 
$$
\begin{align*}
    G_A f(x)
    &= 
    \frac{\sigma^2}{2} (1-s(1-p(x))) \partial_x^2 f(x)
    + (2sp(x) + (1-s)) ( \partial_x \log p(x) ) \partial_x f(x) ,
\end{align*}
$$
i.e., with diffusion speed $\sigma^2 (1-s(1-p))$
and drift $(2sp(x) + (1-s)) ( \partial_x \log p(x) )$.

If we move a large distance in the direction where the $A$ allele is disadvantageous,
the $p \to 0$,
and the equation \eqref{eqn:pfreq}
becomes 
$$
\begin{align*}
    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) .
\end{align*}
$$
In other words, lineages are moving by heat flow with drift.

### 2015-09-11 notes/lineage-movement.tex
.

@@
 \newcommand{\deriv}[1]{\frac{d}{d#1}}
-\newcommand{\dderiv}[1]{\frac{d^2}{d^2#1}}
+\newcommand{\dderiv}[1]{\frac{d^2}{d#1^2}}
 \newcommand{\given}{\;\vert\;}
@@
 Now suppose that $m(x,y)$ is symmetric, and depends on a parameter $\sigma$ so that as $\sigma \to 0$,
-the associate random walk converges to Brownian motion, so that
+the associated random walk converges to Brownian motion, so that
 \begin{align}

### 2015-09-11 notes/skinny-zone.Rmd
.

@@
 \renewcommand{\P}{\mathbb{P}}
+\newcommand{\bone}{\mathbf{1}}
+\newcommand{\given}{\mid}
 \DeclareMathOperator{\sgn}{sgn}
+\DeclareMathOperator{\cov}{cov}
 $$
@@
 
+**In this document take $\sigma=1$.**
+
+A narrow hybrid zone
+--------------------
+
 Suppose that heterozygotes have fitness $(1-s)$;
@@
 $$
-\begin{align} \label{eqn:pfreq}
-    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) (1-p(t,x)) (1-2p(t,x)) ,
+\begin{align}
+    \partial_t p(t,x) &= \frac{1}{2} \partial_x^2 p(t,x) - s p(t,x) (1-p(t,x)) (1-2p(t,x)) ,
 \end{align}
@@
     &= 
-    \frac{\sigma^2}{2} (1-s(1-p(x))) \partial_x^2 f(x)
+    \frac{1}{2} (1-s(1-p(x))) \partial_x^2 f(x)
     + (2sp(x) + (1-s)) ( \partial_x \log p(x) ) \partial_x f(x) ,
@@
 $$
-i.e., with diffusion speed $\sigma^2 (1-s(1-p))$
+i.e., with diffusion speed $(1-s(1-p))$
 and drift $(2sp(x) + (1-s)) ( \partial_x \log p(x) )$.
@@
 If we move a large distance in the direction where the $A$ allele is disadvantageous,
-the $p \to 0$,
-and the equation \eqref{eqn:pfreq}
-becomes 
+then $p \to 0$,
+and the equation for $p$ becomes 
 $$
 \begin{align*}
-    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) .
+    \partial_t p(t,x) &= \frac{1}{2} \partial_x^2 p(t,x) - s p(t,x) .
 \end{align*}
 $$
-In other words, lineages are moving by heat flow with drift.
+At stationarity, so the left side is zero,
+along with the boundary condition $\lim_{x \to \infty} p(x) = 0$,
+this is solved by
+$$
+\begin{align*}
+    p(x) \simeq C \exp\left( - x \sqrt{2s} \right) ,
+\end{align*}
+$$
+and so for $x$ where $p$ is small,
+$$
+\begin{align*}
+    \partial_x \log p(x) = - \sqrt{2s}
+\end{align*}
+$$
+and so the diffusion has generator
+$$
+\begin{align*}
+    \tilde G_A f(x)
+    &= 
+    \frac{1}{2} (1-s) \partial_x^2 f(x)
+    - (1-s) \sqrt{2s} \partial_x f(x) .
+\end{align*}
+$$
+
+Let $B_t$ be a standard Brownian motion
+and let
+$$
+\begin{align*}
+X_t = x + B_{(1-s)t} - (1-s) \sqrt{2s} t
+\end{align*}
+$$
+be the diffusion whose generator is $\tilde G_A$.
+We'd like to know the length of the haplotype carried with the lineage,
+i.e., if recombination occurs at rate $r$,
+whether $X$ hits 0 before an independent Exponential($r$) time.
+Applying Brownian scaling, 
+rescaling time by $(1-s)$ and space by $\sqrt{1-s}$,
+this is the same thing as asking
+whether a Brownian motion begun at 0 with drift $c$
+and killed at rate $\lambda$
+ever hits $x/(\sqrt{1-s})$,
+where
+$$
+\begin{align*}
+    c 
+    &=
+    (1-s) \sqrt{2s} \times \frac{1}{ \sqrt{(1-s)} } \\
+    &=
+    \sqrt{2s(1-s)} .
+\end{align*}
+$$
+and
+$$
+\begin{align*}
+    \lambda
+    &=
+    \frac{ r }{ (1-s) } .
+\end{align*}
+$$
+By Borodin \& Salminen 2.1.1.2,
+for a Brownian motion $W^{(c)}$ with drift $c$ and independent Exponential($\lambda$) time $\tau$,
+$$
+\begin{align*}
+    \P\{ \sup_{0\le t\le \tau} W^{(c)}_t \ge x \}
+    &=
+    \exp\left( - x \left( \sqrt{2\lambda+c^2} - c \right) \right) .
+\end{align*}
+$$
+
+Haplotype length around a rare migrant selected allele
+------------------------------------------------------
+
+Let $L$ be the **length of the haplotype** carried by the migrant lineage.
+The above tells us that
+$$
+\begin{align*}
+    \P^x\{ L > r \}
+    &= 
+    \exp\left( - x \left( \sqrt{2r+2s(1-s)} - \sqrt{2s(1-s)} \right) \right) .
+\end{align*}
+$$
+Rearranging,
+$$
+\begin{align*}
+    \P^x\{ L > (u/x\sqrt{2}+\sqrt{s(1-s)})^2-s(1-s) \}
+    &= 
+    \P^x\{ x\sqrt{2}\left(\sqrt{L+s(1-s)}-\sqrt{s(1-s)}\right) > u \} \\
+    &= 
+    \exp\left( - u \right) .
+\end{align*}
+$$
+In other words,
+if $Y$ is an Exponential with rate $x \sqrt{2}$,
+then $L$ and $(Y+\sqrt{s(1-s)})^2-s(1-s)$ have the same distribution.
+(This is a shifted Weibull.)
+Therefore,
+$$
+\begin{align*}
+    \E[L] 
+    &= 
+    \frac{1}{2x} + \frac{\sqrt{2s(1-s)}}{x} 
+\end{align*}
+$$
+and we can work out the variance also.
+
+Also note that (for drift $c$ to the right) that the furthest to the left
+that $W^{(c)}$ *ever* goes is Exponential($c$).
+In our case, that means that rare migrants are usually descended from some ancestor
+that's an additional $\sqrt{s(1-s)}$ *further* away from where they came from.
+
+Covariance in ancestry
+----------------------
+
+Now suppose that we have two loci, $g_1$ and $g_2$, linked to the selected site,
+at genetic distances $r$ and $r+\ell$ respectively,
+as in *(A)* of the following figure:
+
+![Configuration of linked loci.](linked-sites.png)
+
+Let $I_1$ be the ancestry of a sampled individual at locus $g_1$,
+and likewise $I_2$,
+i.e., $I_1=B$ if the sampled individual inherited at $g_1$
+from an ancestor living to the right of the barrier at time $\tau$ in the past,
+and $I_1=A$ otherwise.
+Note that
+$$
+\cov[\bone_B(I_1),\bone_B(I_2)]
+=
+\P\{ I_1 = B \} \left( \P\{ I_2 = B \given I_1 = B \} - \P\{ I_2 = B \} \right) .
+$$
+We may want to work with the covariance,
+or maybe the difference of the conditional probability itself.
+
+If there has been no recombination between $g_1$ and $g_2$,
+then $I_1=I_2$.
+If there has been, then we need to condition on where and when it occurred.
+Taking from [this document](brownian-local-times.html),
+define ($h=1-\zeta$)
+$$
+\begin{align*}
+    h(x,r,\tau) 
+    &:= \P^x\{ I_1 = B \} \\
+    &= 2(1-\Phi(|x|/\sqrt{\tau})) - 2 e^{r\tau|x|+r^2\tau^2/2} (1-\Phi(|x|/\sqrt{\tau}+r\tau)) .
+\end{align*}
+$$
+Then
+$$
+\begin{align*}
+    \P\{ I_2 = B \given I_1 = B \}
+    &=
+    e^{-r\tau}
+    +
+    \E\left[
+    \int_0^\tau \ell e^{-\ell t} h(X_t,r+\ell,\tau-t) dt
+    \given
+    I_1=B
+    \right] ,
+\end{align*}
+$$
+where $X_t$ is the location of the $g_1$-lineage (then conditioned on $I_1=B$).
+If we let $q_t(x,y)$ be the probability density of $X$ conditioned on $I_1=B$,
+then
+$$
+\begin{align*}
+    \P\{ I_2 = B \given I_1 = B \}
+    &=
+    e^{-r\tau}
+    +
+    \int_0^\tau 
+    \int_{-\infty}^\infty
+    q_t(x,y)
+    \ell e^{-\ell t} h(y,r+\ell,\tau-t) 
+    dy
+    dt .
+\end{align*}
+$$
+What is the motion of $X$ given $I_1=B$?
+$X$ moves as a Brownian motion that is sometimes reflected at zero.

### 2015-09-12 yaniv_sims.R
Create yaniv_sims.R
R script for fwd time simulations.

@@ -1,0 +1,211 @@
+#setwd("~/Desktop/swarmz")
+#library(multicore)
+# MAKE STARTING SAMPLES
+
+
+	getRuns = function(index){
+	    #This returns an array in which rows are the the beggingns and end of a TRUE/FALSE vector
+	    starts  = which(c(0,index)[-1]==1 & c(0,index)[-length(index)]!=1)
+	    stops   = which(c(index,0)[-1]!=1 & c(index,0)[-(1+length(index))]==1)
+	    if(index[length(index)] == 1 & index[(length(index)-1)] !=1){starts = c(starts,length(index))}
+	    return (cbind(starts,stops))
+	}
+
+
+	namesList = function(these.names){
+		# this is a way to get a list where each elelment has a name equalto its value, so that when I lapply, I get back a [correctly] named list
+		all.things = as.list(these.names)
+		names(all.things)=these.names
+		return(all.things)
+	}
+	
+	#get breaks between species' ancestry
+	spBreaks = function(sim){
+		ind.ancest = lapply( sim$inds,function(IND){ lapply(IND,function(CHR){lapply(CHR,function(PAR){
+				if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
+				sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
+				sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
+				tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
+				tmp = tmp[order(tmp$starts),]
+				tmp$starts = PAR$bp[tmp$starts]
+				tmp$stops = c(PAR$bp[-1],1)[tmp$stops]
+				return(tmp)
+		})})})
+		ancest.prop = t(sapply(ind.ancest,function(IND){
+			tmp = t(sapply(IND,function(CHR){ c(X1=with(CHR$X1,sum(sp2*(stops-starts))), X2 = with(CHR$X2,sum(sp2*(stops-starts))))  }))
+			return(colSums(tmp)/nrow(tmp))
+		}))		
[…]

### 2015-09-12 yaniv_sims.R
Update yaniv_sims.R

@@ -198,14 +198,46 @@ sims.sums = lapply(sims.s0.1,spBreaks)
 
 
 
-#####
-#set total range width to -7SIGMA,7SIGMA
-#NEW MIGRATION KERNEL: Sample from gaussian, rnorm(ninds,0.5,2.5)
+
+loci = seq(0.5,1,0.1)
+
+#MAKE file.snp
+    info = data.frame(cbind(1,round(loci*1000000),0,loci))
+    colnames(info) = NULL
+    #write.csv(info,row.names=F,file="file.snp")
 
 
-#sim1.5 = forqs(n=100, sp.ids = rep(c("A","B"),each=50),deme = rep(1:10,each=10),n.chr=1,n.gen=5, QTL = qtl ,mig = MIG)
-#sim1.sum.100 = spBreaks(sim1.100)
-#save(sim1, file = "5gen.Robj")
+#GENOTYPE  [ASSUMES A ONE CHR GENOME WHERE WE HAVE PHASED DIPLOID DATA]
+    genoInd = function(IND,loci){sapply(IND[[1]],function(CHR){as.numeric(CHR$sp2[as.numeric(cut(loci,c(CHR$starts,1)))])})}
 
-mig1=pnorm(-c(0.5,1,1.5,2))
-mig2 = pnorm(-c(1,1.5,2,Inf))
+#GENOTYPE ALL MY INDS
+my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, genoInd, loci)))})
+   #   colnames(my.genos) = NULL
+
+#MAKE OUR FILE [NOTE THIS TAKES A LONG TIME AND IT MIGHT BE SMARTER TO WRITE THE THING ABOVE TO FILE ONE LINE AT A TIME]
+   # write.table(my.genos, file = "file.geno",row.names = FALSE,sep="")
+
+
+genotypes = lapply(my.genos,function(Z){(Z[,seq(1,ncol(Z),2)]+ Z[,seq(2,ncol(Z),2)])/2})
+
+inds.per.deme = 10
+freqs = lapply(genotypes,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),inds.per.deme)),mean)})})
+
+######COMPARE TO THEORY:
+
+xx <- seq(-5,5,length.out=500)
[…]

### 2015-09-12 sims/run-sim.Rmd
began tidying

---
title: "Running the simulation"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
```

Setting up the simulation
-------------------------

First we need to decide where the loci under selection are.
Here `qtl` is a named list, one element per chromosome,
and each element of the list is a data frame
whose columns give the name of the locus,
the selection coefficient,
and the position in Morgans.

```{r setup_qtl}
qtl=list(
         chr1=data.frame(
                    traits = c("underdominant"),
                    s = c(0.1),
                    pos = c(0.5)
                )
    )
```

Now, let's set up the population,
which means specifying the number of individuals,
the number of demes,
the width of the zone in real units,
and the dispersal SD ($\sigma$).
Our locus has a selection coefficient of `r qtl[[1]]$s`,
so the cline width will be about `r round(1/sqrt(qtl[[1]]$s),digits=1)`$\sigma$ wide;
so we'll take our zone to be four times that,
and measure distance in units of $\sigma$:
```{r setup_pop}
zone.width <- ceiling(4/sqrt(qtl[[1]]$s))  # in real units
nind <- 1000
ndeme <- 50
sigma <- 1 # in real units
deme.spacing <- zone.width/ndeme
sigma.demes <- sigma/deme.spacing # in units of demes
```
The discretized dispersal kernel moves as $\lfloor Z+1/2 \rfloor$,
where $Z \sim N(0,\sigma_\text{demes}^2)$;
let's check this has SD close to $\sigma$:
```{r check_sigma}
zz <- floor( rnorm( 1e6, mean=1/2, sd=sigma.demes ) )*deme.spacing
sd(zz)
```
That's pretty good.

Running a simulation
--------------------

To run the simulation,

### 2015-09-12 sims/sim-fns.R

@@ -1,0 +1,179 @@
+
+getRuns = function(index){
+    #This returns an array in which rows are the the beggingns and end of a TRUE/FALSE vector
+    starts  = which(c(0,index)[-1]==1 & c(0,index)[-length(index)]!=1)
+    stops   = which(c(index,0)[-1]!=1 & c(index,0)[-(1+length(index))]==1)
+    if(index[length(index)] == 1 & index[(length(index)-1)] !=1){starts = c(starts,length(index))}
+    return (cbind(starts,stops))
+}
+
+
+namesList = function(these.names){
+    # this is a way to get a list where each elelment has a name equalto its value, so that when I lapply, I get back a [correctly] named list
+    all.things = as.list(these.names)
+    names(all.things)=these.names
+    return(all.things)
+}
+
+#get breaks between species' ancestry
+spBreaks = function(sim){
+    ind.ancest = lapply( sim$inds,function(IND){ lapply(IND,function(CHR){lapply(CHR,function(PAR){
+            if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
+            sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
+            sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
+            tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
+            tmp = tmp[order(tmp$starts),]
+            tmp$starts = PAR$pos[tmp$starts]
+            tmp$stops = c(PAR$pos[-1],1)[tmp$stops]
+            return(tmp)
+    })})})
+    ancest.prop = t(sapply(ind.ancest,function(IND){
+        tmp = t(sapply(IND,function(CHR){ c(X1=with(CHR$X1,sum(sp2*(stops-starts))), X2 = with(CHR$X2,sum(sp2*(stops-starts))))  }))
+        return(colSums(tmp)/nrow(tmp))
+    }))		
+    return(list(ind.ancest = ind.ancest, ancest.prop = ancest.prop))
+}
+
+
+initializeInds = function(arr,n.chr=n.chr){	
+    return(lapply(seq_along(arr[,1]),function(IND){
[…]

### 2015-09-12 sims/yaniv_sims.R

@@ -1,0 +1,63 @@
+#setwd("~/Desktop/swarmz")
+#library(multicore)
+# MAKE STARTING SAMPLES
+
+source("sim-fns.R")
+
+
+
+qtl=list(chr1=data.frame(traits=c("underdominant"), s = c(0.1),pos=c(0.5)))
+
+
+#sigma = n*sqrt(s)/2/sqrt(tau). Define the width of the range by the number of demes. sigma/sqrt(s) is the width of the cline. sqrt(tau) is rate at which cline width is increasing (should also depend on rqtl=list(data.frame(traits=c("underdominant"), s = c(0.1),pos=c(0.5)))
+
+sims.s0.1 = lapply(c(50),sim.zone,nind=2000,ndeme=200,sigma=2.5)
+
+sims.sums = lapply(sims.s0.1,spBreaks)
+#save(sims.sums,file="~/Projects/HybridZones/ClineProjects/sims_sums_s0.1.Robj")
+
+
+
+
+loci = seq(0.5,1,0.1)
+
+#MAKE file.snp
+    info = data.frame(cbind(1,round(loci*1000000),0,loci))
+    colnames(info) = NULL
+    #write.csv(info,row.names=F,file="file.snp")
+
+
+#GENOTYPE  [ASSUMES A ONE CHR GENOME WHERE WE HAVE PHASED DIPLOID DATA]
+    genoInd = function(IND,loci){sapply(IND[[1]],function(CHR){as.numeric(CHR$sp2[as.numeric(cut(loci,c(CHR$starts,1)))])})}
+
+#GENOTYPE ALL MY INDS
+my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, genoInd, loci)))})
+   #   colnames(my.genos) = NULL
+
+#MAKE OUR FILE [NOTE THIS TAKES A LONG TIME AND IT MIGHT BE SMARTER TO WRITE THE THING ABOVE TO FILE ONE LINE AT A TIME]
+   # write.table(my.genos, file = "file.geno",row.names = FALSE,sep="")
+
+
+genotypes = lapply(my.genos,function(Z){(Z[,seq(1,ncol(Z),2)]+ Z[,seq(2,ncol(Z),2)])/2})
+
+inds.per.deme = 10
[…]

### 2015-09-12 yaniv_sims.R

@@ -1,243 +1,0 @@
-#setwd("~/Desktop/swarmz")
-#library(multicore)
-# MAKE STARTING SAMPLES
-
-
-	getRuns = function(index){
-	    #This returns an array in which rows are the the beggingns and end of a TRUE/FALSE vector
-	    starts  = which(c(0,index)[-1]==1 & c(0,index)[-length(index)]!=1)
-	    stops   = which(c(index,0)[-1]!=1 & c(index,0)[-(1+length(index))]==1)
-	    if(index[length(index)] == 1 & index[(length(index)-1)] !=1){starts = c(starts,length(index))}
-	    return (cbind(starts,stops))
-	}
-
-
-	namesList = function(these.names){
-		# this is a way to get a list where each elelment has a name equalto its value, so that when I lapply, I get back a [correctly] named list
-		all.things = as.list(these.names)
-		names(all.things)=these.names
-		return(all.things)
-	}
-	
-	#get breaks between species' ancestry
-	spBreaks = function(sim){
-		ind.ancest = lapply( sim$inds,function(IND){ lapply(IND,function(CHR){lapply(CHR,function(PAR){
-				if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
-				sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
-				sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
-				tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
-				tmp = tmp[order(tmp$starts),]
-				tmp$starts = PAR$bp[tmp$starts]
-				tmp$stops = c(PAR$bp[-1],1)[tmp$stops]
-				return(tmp)
-		})})})
-		ancest.prop = t(sapply(ind.ancest,function(IND){
-			tmp = t(sapply(IND,function(CHR){ c(X1=with(CHR$X1,sum(sp2*(stops-starts))), X2 = with(CHR$X2,sum(sp2*(stops-starts))))  }))
-			return(colSums(tmp)/nrow(tmp))
-		}))		
[…]

### 2015-09-13 sims/run-sim.Rmd
added rmarkdown on using sims

@@
 knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
+source("sim-fns.R")
 ```
@@
 zone.width <- ceiling(4/sqrt(qtl[[1]]$s))  # in real units
-nind <- 1000
-ndeme <- 50
+n.ind <- 1000
+n.deme <- 50
 sigma <- 1 # in real units
-deme.spacing <- zone.width/ndeme
+deme.spacing <- zone.width/n.deme
+deme.locs <- (seq_len(n.deme)-0.5-(n.deme/2))*deme.spacing
 sigma.demes <- sigma/deme.spacing # in units of demes
@@
 ```
-That's pretty good.
+That's pretty close.
 
@@
 
-To run the simulation,
+The simulation doesn't know how we're measuring things in real units,
+so the sigma we pass needs to be in units of demes.
+It returns all the information about the individuals,
+but then we use `spBreaks()` to extract which individual is which ancestry.
+```{r run_quick_sim}
+sim <- sim.zone( n.ind=n.ind, n.deme=n.deme, n.gen=10, sigma=sigma.demes, quiet=TRUE )
+sim.sums <- spBreaks(sim)
+```
+
+OK, now let's genotype these individuals at some loci:
+```{r genotype_sim, fig.width=3*fig.dim, fig.height=1.5*fig.dim}
+loci <- seq(0,1,0.05)
+loci.cols <- rainbow(12)[cut(abs(loci-0.5),breaks=min((length(loci)-1)/2,8))]
+info <- data.frame(
+               chr=1,
+               bp=floor(loci*1e6),
+               pos=loci
+           )
+genotypes <- geno.pop( sim.sums$ind.ancest, list(chr1=loci) )
+```
+
+Now, let's
+compute ancestry frequencies,
+and plot the resulting clines:
+```{r genotype_freqs, fig.width=3*fig.dim, fig.height=1.5*fig.dim}
+freqs <- do.call( cbind, tapply( 1:ncol(genotypes), sim$deme, function (kk) {
+                    rowMeans(genotypes[,kk])
+                } ) )
+matplot(deme.locs, t(freqs), type='l', col=loci.cols, xlab="spatial position", ylab="ancestry frequency")
+legend("topleft",lty=1,legend=loci,col=loci.cols,title="locus position")
+```
+
+We can also look at the local frequencies of the diploid genotypes at each locus:
+```{r diploid_freqs}
+diploid.freqs <- lapply( seq_along(loci), function (locus) {
+                do.call( rbind, tapply( 1:ncol(genotypes), sim$deme, function (kk) {
+                    as.vector( tabulate( 1+genotypes[locus,kk], nbins=3 ) )
+                } ) )
+           } )
+# one end
+matplot(diploid.freqs[[1]],type='l',main=paste("locus at",loci[1]),lty=1,col=c('red','black','green'))
+legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
+# the selected locus
+sel.locus <- match( qtl$chr1$pos, loci )
+matplot(diploid.freqs[[sel.locus]],type='l',main=paste("selected locus at",loci[sel.locus]),lty=1,col=c('red','black','green'))
+legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
+# the other end
+matplot(diploid.freqs[[length(loci)]],type='l',main=paste("locus at",loci[length(loci)]),lty=1,col=c('red','black','green'))
+legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
+```

### 2015-09-13 sims/sim-fns.R

@@ -17,16 +17,20 @@ namesList = function(these.names){
 
 #get breaks between species' ancestry
 spBreaks = function(sim){
-    ind.ancest = lapply( sim$inds,function(IND){ lapply(IND,function(CHR){lapply(CHR,function(PAR){
-            if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
-            sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
-            sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
-            tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
-            tmp = tmp[order(tmp$starts),]
-            tmp$starts = PAR$pos[tmp$starts]
-            tmp$stops = c(PAR$pos[-1],1)[tmp$stops]
-            return(tmp)
-    })})})
+    ind.ancest = lapply( sim$inds,function(IND){ 
+            lapply(IND,function(CHR){
+               lapply(CHR,function(PAR){
+                    if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
+                    sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
+                    sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
+                    tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
+                    tmp = tmp[order(tmp$starts),]
+                    tmp$starts = PAR$pos[tmp$starts]
+                    tmp$stops = c(PAR$pos[-1],1)[tmp$stops]
+                    return(tmp)
+                })
+            })
+        })
     ancest.prop = t(sapply(ind.ancest,function(IND){
         tmp = t(sapply(IND,function(CHR){ c(X1=with(CHR$X1,sum(sp2*(stops-starts))), X2 = with(CHR$X2,sum(sp2*(stops-starts))))  }))
         return(colSums(tmp)/nrow(tmp))
@@ -107,20 +111,10 @@ makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),m
     if(is.null(QTL)){ w_abs = rep(1,length(inds)) }
[…]

### 2015-09-13 sims/yaniv_sims.R

@@ -11,7 +11,7 @@ qtl=list(chr1=data.frame(traits=c("underdominant"), s = c(0.1),pos=c(0.5)))
 
 #sigma = n*sqrt(s)/2/sqrt(tau). Define the width of the range by the number of demes. sigma/sqrt(s) is the width of the cline. sqrt(tau) is rate at which cline width is increasing (should also depend on rqtl=list(data.frame(traits=c("underdominant"), s = c(0.1),pos=c(0.5)))
 
-sims.s0.1 = lapply(c(50),sim.zone,nind=2000,ndeme=200,sigma=2.5)
+sims.s0.1 = lapply(c(50),sim.zone,n.ind=2000,n.deme=200,sigma=2.5)
 
 sims.sums = lapply(sims.s0.1,spBreaks)
 #save(sims.sums,file="~/Projects/HybridZones/ClineProjects/sims_sums_s0.1.Robj")
@@ -27,11 +27,8 @@ loci = seq(0.5,1,0.1)
     #write.csv(info,row.names=F,file="file.snp")
 
 
-#GENOTYPE  [ASSUMES A ONE CHR GENOME WHERE WE HAVE PHASED DIPLOID DATA]
-    genoInd = function(IND,loci){sapply(IND[[1]],function(CHR){as.numeric(CHR$sp2[as.numeric(cut(loci,c(CHR$starts,1)))])})}
-
 #GENOTYPE ALL MY INDS
-my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, genoInd, loci)))})
+my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
    #   colnames(my.genos) = NULL
 
 #MAKE OUR FILE [NOTE THIS TAKES A LONG TIME AND IT MIGHT BE SMARTER TO WRITE THE THING ABOVE TO FILE ONE LINE AT A TIME]

### 2015-09-14 notes/brownian-local-times.Rmd
l

@@
 $\tau$ generations ago.
-Let $p_H$ denote the density of heterozygotes in the cline
--- this should be $p_H = (1-s)/2$).
+(Note that $r$ should be scaled by the density and fecundity of heterozygotes in the cline --
+see below.)
 For an individual to the left of the barrier ($x<0$),
 the probability of inheriting from the left of the barrier at $\tau$
-is the probabiltiy that no recombination at the barrier has occurred,
+is the probability that no recombination at the barrier has occurred,
 plus one-half the probability that recombination in a heterozygote has occurred,
-i.e., $\zeta p_H +(1-\zeta p_H)/2=(1+\zeta p_H)/2$.
+i.e., $\zeta +(1-\zeta)/2=(1+\zeta)/2$.
 If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
-i.e., $(1-\zeta p_H)/2$.
+i.e., $(1-\zeta)/2$.
 Combining these,
@@
     p(x,r,\tau,A) = 
-    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) p_H \right\} .
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
 \end{align*}
@@
 
-Does this actually make a cline?
+Here's the cline shape:
 ```{r show_cline}
-pcline <- function (x,r,tau=1,sigma=1,pH=1/2) {
-    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau)*pH )
+pcline <- function (x,r,tau=1,sigma=1) {
+    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
 }
@@
     plot(xx, pcline(xx,r=1), type='n', 
-            xlab="distance from cline center", ylab="prob of inherited from the left", 
+            xlab="distance from cline center", ylab="prob inherited from the left", 
             main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
@@
 ```
+
+Adjusting for cline width
+=========================
+
+Above, we were approximating the time that a reflecting lineage
+spends in a small interval near zero
+by the local time at zero multiplied by the width of the interval.
+In fact, when the lineage is at $x$, 
+the lineage recombines to the other background
+in a given generation
+with probability equal to the recombination rate 
+multiplied by the probability of being found in a heterozygote.
+If $p_{AB}(x)$ is the local fraction of heterozygotes,
+and heterozygotes have relative fecundity $1-s$,
+then the probability of being in a heterozygote is
+$(1-s)p_{AB}(x)/((1-s)p_{AB}(x)+2p_{AA}(x))$;
+if diploid genotypes are assorted randomly,
+then $p_{AB} = 2p(1-p)$ and $p_AA=p^2$, so the probability that an $A$ lineage
+passes through a heterozygote in a given generation is
+$$
+\begin{align*}
+f_H(x) 
+&= 
+\frac{ (1-s)p(x)(1-p(x)) }{ p(x)^2 + (1-s) p(x) (1-p(x)) }
+&= 
+\frac{ (1-s)(1-p(x)) }{ p(x) + (1-s) (1-p(x)) } \\
+&= 
+\frac{ (1-s)(1-p(x)) }{ 1 - s (1-p(x)) } .
+\end{align*}
+$$
+
+Bazykin's approximation to a cline is
+$$
+p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
+$$
+Here's what that looks like:
+```{r tanh_cline}
+tanh_cline <- function (x,s,sigma=1) {
+    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+}
+plot(xx, tanh_cline(xx,s=0.04) )
+```
+
+A lineage does something a bit more complicated,
+but let's just run a random walk with drift equal to the logarithmic derivative of $p(x)$,
+i.e.,
+$$
+\begin{align*}
+    \partial_x \log(p(x)) 
+    &= -\frac{2\sqrt{s}}{\sigma} 
+        \frac{1-\tanh^2(-2x\sqrt{s}/σ)}{1+\tanh(-2x\sqrt{s}/σ)} .
+\end{align*}
+$$
+
+```{r sim_lineage}
+drift <- function (x,s,sigma=1) {
+    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
+}
+nwalks <- 1e4
+ntimes <- 1e3
+s <- 0.1
+# want total time to be 1, so:
+eps <- 1/sqrt(ntimes)
+rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=(ntimes-1), ncol=nwalks ))
+for (k in 2:ntimes) {
+    rws[k,] <- rws[k-1,]+rws[k,]+drift(rws[k-1,],s)
+}
+layout( t(1:2), widths=c(3,1) )
+matplot(rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
+plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
+lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
+```

### 2015-10-02 drafts/hybrid_zone.tex
Rename hybrd_zone.tex to hybrid_zone.tex

\documentclass[12pt]{amsart}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry

% See the ``Article customise'' template for come common customisations

\title{}
\author{}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

\section{Introduction}
-Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation. Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone). \\
-Gene flow across a hybrid zone may be impeded by selection against migrants. In older contact zones, loci that are associated with reduced fitness in heterozygotes will have steeper clines than the rest of the genome. There is probably a lot to learn from geographic patterns at loci neighboring selected loci (e.g. small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)\\ 
-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
-We consider the non-equilibrium state of hybrid zones, and patterns of linkage around selected loci. This will become increasingly relevant as studiers of HZs get genomic data and can reliably estimate block lengths. \\

\section{Methods}
\subsection{Model}
We consider a model in which two isolated subspecies, species $A$ and species $B$ come into contact at time $\tau$ generations in the past. Individuals move forward in time by sampling a displacement from a Gaussian distribution. In the simplest case of under-dominance, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.\\

We consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past. The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that the two species are fixed for alternate alleles at this locus). For every other locus, we keep track of $a$, the identity at $L$, across time and space. If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise. Recombination is modeled as a poisson process with rate $r$. The identity $a$ at locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$. 
\\

As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone. This means that recombination events that occur away from the center of the zone do not result in a change in $a$.\\

\subsection{}

\subsection{Simulations}
Forward in time simulations were used to asses theoretical predictions. We simulate discrete demes with nearest-neighbor migration and non-overlapping generations. Each generation, an individual moves to a neighboring deme with probability $m$. The direction in which it moves is determined by a coin-flip, except at edge demes, where an individual moves in one direction with probability $m/2$. Following migration, the next generation is produced by replacing each individual in a deme by randomly choosing two parents with probability weighted by the fitness of their genotypes. The genotype of the new individual is generated by simulating a recombined chromosome from each parent. This simulation keeps track of the ancestry at each location in the chromosome for all individuals. Each individual has 1 pair of chromosomes of length 1 Morgan.

\end{document}

### 2015-10-05 notes/brownian-local-times.Rmd
added html

--- 2015-09-11 notes/brownian-local-times.Rmd
@@
 $$
-\begin{align*}
+\begin{aligned}
     \ell(x) = \lim_{\epsilon \to 0} \frac{1}{\epsilon}
         \text{measure}\{t \,:\, x \le B_t \le x+\epsilon, \, 0 \le t \le 1\}
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P\{ \ell(x) \le u \} = 2 \Phi(|x|+u)-1,
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P\{ \ell(x) = 0 \} &= 2 \Phi(|x|)-1 \\
     \P\{ \ell(x) \in du \} &= 2 \phi(|x|+u) ,
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
 \E\left[ e^{-r\ell(x)} \right]
@@
 2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \lim_{x \to 0} \zeta(x,r) 
@@
     &= \frac{2}{r\sqrt{2\pi}} + O(r^{-3}) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \lim_{x \to 0-} p(x,r) 
@@
       +O(r^{-3}) 
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \partial_x \zeta(x,r)
@@
       \right)
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     -\partial_x p(x)
@@
     \frac{1}{\sqrt{2\pi}} + O(\frac{1}{r}) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \zeta(x,r) = 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \zeta(x,r,\tau,\sigma).
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \zeta(x,r,\tau,\sigma) 
@@
     \zeta(x/\sqrt{\tau\sigma^2},r\tau) ,
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \zeta(x,r)
@@
     2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     p(x,r,\tau,A) = 
     \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) p_H \right\} .
-\end{align*}
+\end{aligned}
 $$

### 2015-10-05 notes/skinny-zone.Rmd
more html

@@
 $$
-\begin{align}
+\begin{aligned}
     \partial_t p(t,x) &= \frac{1}{2} \partial_x^2 p(t,x) - s p(t,x) (1-p(t,x)) (1-2p(t,x)) ,
-\end{align}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     G_A f(x)
@@
     + (2sp(x) + (1-s)) ( \partial_x \log p(x) ) \partial_x f(x) ,
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \partial_t p(t,x) &= \frac{1}{2} \partial_x^2 p(t,x) - s p(t,x) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     p(x) \simeq C \exp\left( - x \sqrt{2s} \right) ,
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \partial_x \log p(x) = - \sqrt{2s}
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \tilde G_A f(x)
@@
     - (1-s) \sqrt{2s} \partial_x f(x) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
 X_t = x + B_{(1-s)t} - (1-s) \sqrt{2s} t
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     c 
@@
     \sqrt{2s(1-s)} .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \lambda
@@
     \frac{ r }{ (1-s) } .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P\{ \sup_{0\le t\le \tau} W^{(c)}_t \ge x \}
@@
     \exp\left( - x \left( \sqrt{2\lambda+c^2} - c \right) \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P^x\{ L > r \}
@@
     \exp\left( - x \left( \sqrt{2r+2s(1-s)} - \sqrt{2s(1-s)} \right) \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P^x\{ L > (u/x\sqrt{2}+\sqrt{s(1-s)})^2-s(1-s) \}
@@
     \exp\left( - u \right) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \E[L] 
@@
     \frac{1}{2x} + \frac{\sqrt{2s(1-s)}}{x} 
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     h(x,r,\tau) 
@@
     &= 2(1-\Phi(|x|/\sqrt{\tau})) - 2 e^{r\tau|x|+r^2\tau^2/2} (1-\Phi(|x|/\sqrt{\tau}+r\tau)) .
-\end{align*}
+\end{aligned}
 $$
@@
 $$
-\begin{align*}
+\begin{aligned}
     \P\{ I_2 = B \given I_1 = B \}
@@
     \right] ,
-\end{align*}
+\end{aligned}
 $$
@@
     dt .
-\end{align*}
+\end{aligned}
 $$

### 2015-10-06 notes/brownian-local-times.Rmd
more html

--- 2015-09-14 notes/brownian-local-times.Rmd
@@
 $$
-\begin{align*}
-\E\left[ e^{-r\ell(x)} \right]
+\begin{aligned}
+\zeta(x,r) &:=
+\E\left[ e^{-r\ell(x)} \right] \\
 &=
@@
 $$
-\begin{align*}
+\begin{aligned}
     p(x,r,\tau,A) = 
     \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
-\end{align*}
+\end{aligned}
 $$

### 2015-10-14 notes/brownian-local-times.Rmd
another approx
.

@@
 knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
+```
+
+*Useful functions used below:*
+```{r setup_fns}
+tanh_cline <- function (x,s,sigma=1) {
+    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+}
+pcline <- function (x,r,tau=1,sigma=1) {
+    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
+}
 ```
@@
 
-Also note that for large $|x|$, by the asymptotics for $1-\Phi(x)$,
+Also note that by the asymptotics for $1-\Phi(x)$,
 $$
 \begin{aligned}
-    \zeta(x,r) = 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) .
+    \zeta(x,r) \approx 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) \qquad \text{for large }|x| .
 \end{aligned}
@@
 ```{r show_cline}
-pcline <- function (x,r,tau=1,sigma=1) {
-    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
-}
+pcline
 xx <- seq(-15,15,length.out=500)
@@
 ```{r tanh_cline}
-tanh_cline <- function (x,s,sigma=1) {
-    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
-}
+tanh_cline
 plot(xx, tanh_cline(xx,s=0.04) )
@@
 s <- 0.1
+xx <- seq(-15,15,length.out=500)
 # want total time to be 1, so:

### 2015-10-15 notes/brownian-local-times.Rmd
.

--- 2015-10-06 notes/brownian-local-times.Rmd
@@
 rws <- rbind(0,apply(rws,2,cumsum))
-matplot(rws[,1:20],type='l', ylab='position')
+matplot(seq(0,1,length.out=ntimes),rws[,1:20],type='l', ylab='position')
 ```
@@
 and scaled steps to have this be a total Brownian time of 1.
-Let's check:
+Let's check;
+this should just be a standard Gaussian:
 ```{r final_dist}
-xh <- hist(rws[ntimes,],breaks=50)
+xh <- hist(rws[ntimes,],breaks=50,main="final position")
 lines( xh$mids, sum(xh$counts)*diff(pnorm(xh$breaks)) )
@@
 What about the local time at other locations?
-Note there is an atom at 0, now,
+Note there is an atom at 0 now
+(that is depicted by the dots in the figures below),
 and that the discretization could trip us up;
@@
 Here's the cline shape:
-```{r show_cline}
+```{r show_cline, cache=TRUE}
 pcline <- function (x,r,tau=1,sigma=1) {
@@
 if diploid genotypes are assorted randomly,
-then $p_{AB} = 2p(1-p)$ and $p_AA=p^2$, so the probability that an $A$ lineage
+then $p_{AB} = 2p(1-p)$ and $p_{AA}=p^2$, so the probability that an $A$ lineage
 passes through a heterozygote in a given generation is
@@
 Here's what that looks like:
-```{r tanh_cline}
+```{r tanh_cline, cache=TRUE}
 tanh_cline <- function (x,s,sigma=1) {
@@
 }
-plot(xx, tanh_cline(xx,s=0.04) )
+fH <- function (x,s,pfn=tanh_cline,...) {
+    px <- pfn(x,s=s,...)
+    ( (1-s)*(1-px) )/( 1 - s*(1-px) )
+}
+plot(xx, tanh_cline(xx,s=0.04), type='l' )
 ```
@@
 
-```{r sim_lineage}
+```{r sim_lineage, cache=TRUE, depends=c("tanh_cline","show_cline","killing")}
 drift <- function (x,s,sigma=1) {
@@
 nwalks <- 1e4
-ntimes <- 1e3
+ntimes <- 3e3
+total.time <- 10
 s <- 0.1
-# want total time to be 1, so:
-eps <- 1/sqrt(ntimes)
-rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=(ntimes-1), ncol=nwalks ))
+dt <- total.time/ntimes
+# put the noise in rws
+rws <- rbind( rnorm(nwalks), # initial conditions
+             matrix( rnorm(nwalks*(ntimes-1)), nrow=(ntimes-1), ncol=nwalks ) )
+# and then integrate over it
 for (k in 2:ntimes) {
-    rws[k,] <- rws[k-1,]+rws[k,]+drift(rws[k-1,],s)
+    rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],s)
 }
 layout( t(1:2), widths=c(3,1) )
-matplot(rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
-plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
+matplot(seq(0,total.time,length.out=ntimes),rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
+legend("topright",lwd=2,legend=c("prob of het","cline","drift"),col=c("black","green","red"))
+plot( (-1)*drift(xx,s=s), xx, col='red', lwd=2, type='l', ylim=range(rws), ylab='', xlab='drift or prob' )
+lines( tanh_cline(xx,s=s), xx, col='green', lwd=2 )
 lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
 ```
+The cline is in green and the local probability of being in a heterozygote is in black;
+the drift (multiplied by $-1$) is in red.
+
+Let's just check that's doing the right thing:
+```{r sim_lineage_check, cache=TRUE, depends="sim_lineage"}
+xbreaks <- seq(-3,3,length.out=100)
+xmids <- xbreaks[-1]-diff(xbreaks)/2
+dX <- apply(rws,2,diff)
+Xmeans <- tapply( dX, cut(rws[-nrow(rws),],breaks=xbreaks), mean )
+Xvar <- tapply( dX, cut(rws[-nrow(rws),],breaks=xbreaks), var )
+layout(1:2)
+plot(xmids,Xmeans/dt,ylab="mean dX")
+lines(xmids, drift(xmids,s), col='red', lwd=2)
+plot(xmids,Xvar/dt,ylab="variance dX")
+abline(h=1,col='red',lwd=2)
+```
+
+**Comparison to zero-width cline:**
+Let's start a bunch of random walks sorta far from the cline center,
+and look at the distribution of the time spend in heterozygotes.
+```{r sim_het_time, cache=TRUE, depends="sim_lineage"}
+nwalks <- 1e4
+ntimes <- 3e3
+total.time <- 100
+s <- 0.1
+dt <- total.time/ntimes
+# put the noise in rws
+rws <- rbind( -10, # initial conditions
+             matrix( rnorm(nwalks*(ntimes-1)), nrow=(ntimes-1), ncol=nwalks ) )
+# and then integrate over it
+for (k in 2:ntimes) {
+    rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],s)
+}
+layout( t(1:2), widths=c(3,1) )
+matplot(seq(0,total.time,length.out=ntimes),rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
+legend("topright",lwd=2,legend=c("prob of het","cline","drift"),col=c("black","green","red"))
+plot( (-1)*drift(xx,s=s), xx, col='red', lwd=2, type='l', ylim=range(rws), ylab='', xlab='drift or prob' )
+lines( tanh_cline(xx,s=s), xx, col='green', lwd=2 )
+lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
+```
+Here is the distribution of $T=\int_0^{`r total.time`} f_H(X_t) dt$:
+```{r dist_het_time, cache=TRUE, depends="sim_het_time"}
+het.time <- dt*colSums( fH(rws,s=s) )
+layout(1:2)
+hist( het.time, breaks=50, xlab="time in hets", freq=FALSE, main="histogram of T" ) 
+hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
+```
+
+Brownian motion with drift above zero
+=====================================
+
+A better model for the process might be Brownian motion
+that, when above zero, has negative drift,
+and is killed at some rate.
+In other words, 
+if $dX_t = -\theta(X_t) dt + dB_t$,
+where $\theta(x)=\theta$ for $x \ge 0$, then
+we want to know
+$$\begin{aligned}
+    u(x;\lambda,\theta):=\E^x \left[ \exp\left( - \lambda \int_0^\infty X_t dt \right) \right] .
+\end{aligned}$$
+(*Small problem:*  this isn't guarenteed to be on the correct size of zero when we want it to.)
+Well, by Girsanov's theorem, we know that
+$\E[f(X_t)] = \E[f(B_t)\zeta(t)]$
+where
+$$\begin{aligned}
+\zeta(t) = \exp\left[ \int_0^t \theta(B_u) dB_t - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+\end{aligned}$$
+and since by Tanaka's formula,
+$$\begin{aligned}
+\theta \max(0,B_t) - \max(0,B_0) = \int_0^t \theta(B_u) du + \theta \ell_t(0) ,
+\end{aligned}$$
+thus
+$$\begin{aligned}
+\zeta(t) = \exp\left[ \theta \max(0,B_t) - \theta \max(0,B_0) - \theta \ell_t(0) - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+\end{aligned}$$
+
+More than we need to know
+-------------------------
+
+According to Karatzas & Shreve (1984)
+(with their $\delta=\theta_1=0$),
+if $dX_t = -\theta(X_t) dt + dB_t$,
+where $\theta(x)=\theta$ for $x \ge 0$, then
+if we define
+$$
+\begin{aligned}
+h(s;x,\mu) = \frac{|x|}{\sqrt{2\pi s^3}} \exp\left[-\frac{(x+\mu s)^2}{2s}\right] \qquad x>0 ,
+\end{aligned}
+$$
+(this is the density of the first time a BM with drift $\mu$ started at $x$ hits 0),
+then
+$$
+\begin{aligned}
+p_t(x,z) 
+&= \P\{ X_t = z \mid X_0 = x \}\\
+&= \begin{cases}
+    2 \int_0^\infty \int_0^t h(t-\tau;b-z,0) h(\tau;x+b,\theta) d\tau db
+        & \qquad x \ge 0, \quad z\le 0 \\
+    2 \int_0^\infty \int_0^t e^{2\theta z} h(t-\tau;b,0) h(\tau;x+b,\theta) d\tau db & \\
+    \qquad {} + \frac{1}{\sqrt{2\pi t}} 
+        \left\{ 
+            \exp\left[ -\frac{ (x-z+\theta t)^2 }{2t} \right]
+            +\exp\left[ -\frac{ (x+z-\theta t)^2 }{2t} - 2 \theta x \right]
+        \right\}
+        & \qquad x \ge 0, \quad z > 0 .
+\end{cases}
+\end{aligned}
+$$
+Also, $p_t(x,z;\theta)=p_t(-x,-z;-\theta)$ gets the same expressions for $x<0$.

### 2015-10-15 notes/diffusion-local-times.Rmd
;.

--- 2015-10-14 notes/brownian-local-times.Rmd
@@
 ---
-title: "Brownian local times"
+title: "Diffusion local times and the time until recombination"
 date: "`r date()`"
@@
 ```{r setup_fns}
+pcline <- function (x,r,tau=1,sigma=1) {
+    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
+}
 tanh_cline <- function (x,s,sigma=1) {
@@
 }
-pcline <- function (x,r,tau=1,sigma=1) {
-    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
+tanh_drift <- function (x,s,sigma=1) {
+    # log-derivative of tanh_cline
+    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
 }
+tanh_het <- function (x,s,sigma=1) {
+    # prob of being in a het = (1-s)(1-p)/(1-s(1-p))
+    pp <- tanh_cline(x,s,sigma=sigma)
+    (1-s)*(1-pp)/(1-s*(1-pp))
+}
+```
+Approximations, generally
+=========================
+
+We have a function $p(x)$ (the local proportion of $A$ alleles),
+and another function $f_H(x)$ (the local probability an $A$ is in a heterozygote),
+which happens to equal $(1-s)(1-p(x))/(1-s(1-p(x)))$ but we might not want to use that fact;
+then $p$ defines a diffusion $X$ by $dX_t = \log(p(x))' dt + dB_t$
+and then we define $H_t = \int_0^t f_H(X_s) ds$.
+We want to know $\E^x[\exp(-r H_t)]$ as a function of $x$, $r$, and $t$.
+
+More concretely,
+we want the distribution of the time to first recombination in a heterozygote;
+denote this by $T$, so that
+$$
+\begin{aligned}
+\P\{ T > t \} 
+&= \exp\left\{ - \int_0^t r f_H(X_s) ds \right\} \\
+&= \exp\left\{ - \int_{-\infty}^{\infty} r f_H(x) \ell_t(x) dx \right\}  ,
+\end{aligned}
+$$
+where $\ell$ is the local time density for $X$.
+This is the Laplace transform of 
+the heterozygote-weighted local time, $H_t = \int f_H(x) \ell_t(x) dx$,
+so it means that we want to come up with a reasonable approximation to this quantity.
+
+This suggests two possible approximations:
+
+1. A lineage doesn't go past the hybrid zone,
+   and only recombines when it's in the hybrid zone,
+   which has width something,
+   and where the probability of being in a heterozygote is 1/2.
+
+2. A lineage spends little time in the hybrid zone,
+   and recombines when it's past the hybrid zone,
+   where it's type is rare,
+   thus replacing both the drift $\partial_x \log(p(x))$
+   and the probability of being in a heterozygote $f_H(x)$ with step functions.
+
+The first one seems more natural,
+but it requires two parameters: 
+what the width should be and where the barrier should be.
+
+The right model
+===============
+
+When the lineage is at $x$, 
+the lineage recombines to the other background
+in a given generation
+with probability equal to the recombination rate 
+multiplied by the probability of being found in a heterozygote.
+If $p_{AB}(x)$ is the local fraction of heterozygotes,
+and heterozygotes have relative fecundity $1-s$,
+then the probability of being in a heterozygote is
+$(1-s)p_{AB}(x)/((1-s)p_{AB}(x)+2p_{AA}(x))$;
+if diploid genotypes are assorted randomly,
+then $p_{AB} = 2p(1-p)$ and $p_{AA}=p^2$, so the probability that an $A$ lineage
+passes through a heterozygote in a given generation is
+$$
+\begin{align*}
+f_H(x) 
+&= 
+\frac{ (1-s)p(x)(1-p(x)) }{ p(x)^2 + (1-s) p(x) (1-p(x)) }
+&= 
+\frac{ (1-s)(1-p(x)) }{ p(x) + (1-s) (1-p(x)) } \\
+&= 
+\frac{ (1-s)(1-p(x)) }{ 1 - s (1-p(x)) } .
+\end{align*}
+$$
+
+Bazykin's approximation to a cline is
+$$
+p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
+$$
+Here's what that looks like:
+```{r tanh_cline}
+s <- 0.1
+tanh_cline
+xx <- seq(-15,15,length.out=500)
+plot(xx, (-1)*tanh_drift(xx,s=s), type='l', col='red', xlab="space", ylab='proportion or drift')
+lines(xx, tanh_cline(xx,s=s), col='black')
+lines(xx, 2*tanh_cline(xx,s=s)*(1-tanh_cline(xx,s=s)), col='blue')
+lines(xx, tanh_het(xx,s=s), col='green')
+legend("topleft",lty=1,col=c("black","red","blue","green"),
+        legend=c("p(x)","-log(p(x))'","2p(1-p)","prob of het"))
+```
+
+A lineage does something a bit more complicated,
+but let's just run a random walk with drift equal to the logarithmic derivative of $p(x)$,
+i.e.,
+$$
+\begin{align*}
+    \partial_x \log(p(x)) 
+    &= -\frac{2\sqrt{s}}{\sigma} 
+        \frac{1-\tanh^2(-2x\sqrt{s}/σ)}{1+\tanh(-2x\sqrt{s}/σ)} .
+\end{align*}
+$$
+(Really, the drift should be the logarithmic derivative of $p(x)(1-sf_H(x))$.)
+
+```{r sim_lineage,cache=TRUE}
+tanh_drift
+nwalks <- 1e4
+ntimes <- 1e3
+s <- 0.1
+tau <- 10
+xx <- seq(-15,15,length.out=500)
+# want total time to be 1, so:
+dt <- tau/ntimes
+tt <- seq(0,tau,length.out=ntimes)
+rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*sqrt(dt), nrow=(ntimes-1), ncol=nwalks ))
+for (k in 2:ntimes) {
+    rws[k,] <- rws[k-1,]+rws[k,]+dt*tanh_drift(rws[k-1,],s)
+}
+layout( t(1:2), widths=c(3,1) )
+matplot(tt, rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
+plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
+lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
+```
+
+Now let's get local times for those walks.
+```{r lineage_local_times,cache=TRUE,depends="sim_lineage"}
+xx <- seq( min(rws), max(rws), length.out=200 )
+xmids <- xx[-1]-diff(xx)/2
+lts <- dt*apply( matrix(findInterval(rws,xx),nrow=nrow(rws)), 2, tabulate, nbins=length(xx)-1 )
+layout( t(1:2), widths=c(3,1) )
+matplot(lts[,1:30],xmids,type='l',xlab="local time", ylab="space")
+plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
+lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
 ```
@@
 ===============================
+
+Now, let's look at the analytics for the first approximation.
 
@@
 ```
-
-Adjusting for cline width
-=========================
-
-Above, we were approximating the time that a reflecting lineage
-spends in a small interval near zero
-by the local time at zero multiplied by the width of the interval.
-In fact, when the lineage is at $x$, 
-the lineage recombines to the other background
-in a given generation
-with probability equal to the recombination rate 
-multiplied by the probability of being found in a heterozygote.
-If $p_{AB}(x)$ is the local fraction of heterozygotes,
-and heterozygotes have relative fecundity $1-s$,
-then the probability of being in a heterozygote is
-$(1-s)p_{AB}(x)/((1-s)p_{AB}(x)+2p_{AA}(x))$;
-if diploid genotypes are assorted randomly,
-then $p_{AB} = 2p(1-p)$ and $p_AA=p^2$, so the probability that an $A$ lineage
-passes through a heterozygote in a given generation is
-$$
-\begin{align*}
-f_H(x) 
-&= 
-\frac{ (1-s)p(x)(1-p(x)) }{ p(x)^2 + (1-s) p(x) (1-p(x)) }
-&= 
-\frac{ (1-s)(1-p(x)) }{ p(x) + (1-s) (1-p(x)) } \\
-&= 
-\frac{ (1-s)(1-p(x)) }{ 1 - s (1-p(x)) } .
-\end{align*}
-$$
-
-Bazykin's approximation to a cline is
-$$
-p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
-$$
-Here's what that looks like:
-```{r tanh_cline}
-tanh_cline
-plot(xx, tanh_cline(xx,s=0.04) )
-```
-
-A lineage does something a bit more complicated,
-but let's just run a random walk with drift equal to the logarithmic derivative of $p(x)$,
-i.e.,
-$$
-\begin{align*}
-    \partial_x \log(p(x)) 
-    &= -\frac{2\sqrt{s}}{\sigma} 
-        \frac{1-\tanh^2(-2x\sqrt{s}/σ)}{1+\tanh(-2x\sqrt{s}/σ)} .
-\end{align*}
-$$
-
-```{r sim_lineage}
-drift <- function (x,s,sigma=1) {
-    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
-}
-nwalks <- 1e4
-ntimes <- 1e3
-s <- 0.1
-xx <- seq(-15,15,length.out=500)
-# want total time to be 1, so:
-eps <- 1/sqrt(ntimes)
-rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=(ntimes-1), ncol=nwalks ))
-for (k in 2:ntimes) {
-    rws[k,] <- rws[k-1,]+rws[k,]+drift(rws[k-1,],s)
-}
-layout( t(1:2), widths=c(3,1) )
-matplot(rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
-plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
-lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
-```

### 2015-10-15 notes/diffusion-local-times.Rmd
.

@@
 }
+sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
+    # want total time to be tau, so:
+    dt <- tau/ntimes
+    # put the noise in rws
+    rws <- matrix( c( rep_len(xinit,nwalks), # initial conditions
+                 rnorm(nwalks*(ntimes-1)) ), nrow=ntimes, ncol=nwalks, byrow=TRUE )
+    # and then integrate over it
+    for (k in 2:ntimes) {
+        rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],...)
+    }
+    attr(rws,"dt") <- dt
+    return(rws)
+}
 ```
+
 Approximations, generally
@@
 xx <- seq(-15,15,length.out=500)
-# want total time to be 1, so:
-dt <- tau/ntimes
 tt <- seq(0,tau,length.out=ntimes)
-rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*sqrt(dt), nrow=(ntimes-1), ncol=nwalks ))
-for (k in 2:ntimes) {
-    rws[k,] <- rws[k-1,]+rws[k,]+dt*tanh_drift(rws[k-1,],s)
-}
+rws <- sim_diffusion( tanh_drift, tau, nwalks, ntimes, s=s, xinit=rnorm(nwalks) )
 layout( t(1:2), widths=c(3,1) )
 matplot(tt, rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
-plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
+legend("topright",lwd=2,legend=c("prob of het","cline","drift"),col=c("black","green","red"))
+plot( (-1)*tanh_drift(xx,s=s), xx, col='red', lwd=2, type='l', ylim=range(rws), ylab='', xlab='drift or prob' )
+lines( tanh_cline(xx,s=s), xx, col='green', lwd=2 )
 lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
 ```
+The cline is in green and the local probability of being in a heterozygote is in black;
+the drift (multiplied by $-1$) is in red.
 
@@
 xmids <- xx[-1]-diff(xx)/2
-lts <- dt*apply( matrix(findInterval(rws,xx),nrow=nrow(rws)), 2, tabulate, nbins=length(xx)-1 )
+lts <- attr(rws,"dt")*apply( matrix(findInterval(rws,xx),nrow=nrow(rws)), 2, tabulate, nbins=length(xx)-1 )
 layout( t(1:2), widths=c(3,1) )
@@
 lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
+```
+
+Here is the distribution of $T=\int_0^{`r tau`} f_H(X_t) dt$:
+```{r dist_het_time, cache=TRUE, depends="sim_het_time"}
+het.time <- weighted_time( rws, tanh_het, s=s )
+layout(1:2)
+hist( het.time, breaks=50, xlab="time in hets", freq=FALSE, main="histogram of T" ) 
+hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
+```
+
+Comparison, by simulation
+=========================
+
+```{r sim_general, fig.height=3*fig.dim}
+weighted_time <- function (rws, weight.fn, ...) {
+    colSums( weight.fn(rws,...) ) * attr(rws,"dt")
+}
+nwalks <- 1e4
+ntimes <- 1e3
+s <- 0.1
+tau <- 10
+lt.comparison <- list(
+        tanh=weighted_time( 
+               sim_diffusion( tanh_drift,tau,nwalks,ntimes,s=s ),
+               tanh_het, s=s
+           ),
+        reflected=weighted_time(
+                sim_diffusion( function(...){0}, tau, nwalks, ntimes ),
+                function (x,s) { ifelse(abs(x)<1/(2*sqrt(s)), 1/2, 0) }, s=s
+            ),
+        reflected_tanh=weighted_time(
+                sim_diffusion( function(...){0}, tau, nwalks, ntimes ),
+                tanh_het, s=s
+            ),
+        pushed=weighted_time(
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s ),
+                function (x) { x>0 }
+            ),
+        pushed_tanh=weighted_time(
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s ),
+                tanh_het, s=s
+            )
+        )
+layout(seq_along(lt.comparison))
+for (k in seq_along(lt.comparison)) {
+    hist(lt.comparison[[k]],xlim=range(lt.comparison),main=names(lt.comparison)[k],breaks=50,xlab="time in hets")
+}
 ```
@@
 ntimes <- 1e3
-# want total time to be 1, so:
+# want total time to be tau, so:
 eps <- 1/sqrt(ntimes)
@@
 Here's the cline shape:
-```{r show_cline}
+```{r show_cline, cache=TRUE}
 pcline
@@
 ```
+
+Brownian motion with drift above zero
+=====================================
+
+A better model for the process might be Brownian motion
+that, when above zero, has negative drift,
+and is killed at some rate.
+In other words, 
+if $dX_t = -\theta(X_t) dt + dB_t$,
+where $\theta(x)=\theta$ for $x \ge 0$, then
+we want to know
+$$\begin{aligned}
+    u(x;\lambda,\theta):=\E^x \left[ \exp\left( - \lambda \int_0^\infty X_t dt \right) \right] .
+\end{aligned}$$
+(*Small problem:*  this isn't guarenteed to be on the correct size of zero when we want it to.)
+Well, by Girsanov's theorem, we know that
+$\E[f(X_t)] = \E[f(B_t)\zeta(t)]$
+where
+$$\begin{aligned}
+\zeta(t) = \exp\left[ \int_0^t \theta(B_u) dB_t - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+\end{aligned}$$
+and since by Tanaka's formula,
+$$\begin{aligned}
+\theta \max(0,B_t) - \max(0,B_0) = \int_0^t \theta(B_u) du + \theta \ell_t(0) ,
+\end{aligned}$$
+thus
+$$\begin{aligned}
+\zeta(t) = \exp\left[ \theta \max(0,B_t) - \theta \max(0,B_0) - \theta \ell_t(0) - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+\end{aligned}$$
+
+More than we need to know
+-------------------------
+
+According to Karatzas & Shreve (1984)
+(with their $\delta=\theta_1=0$),
+if $dX_t = -\theta(X_t) dt + dB_t$,
+where $\theta(x)=\theta$ for $x \ge 0$, then
+if we define
+$$
+\begin{aligned}
+h(s;x,\mu) = \frac{|x|}{\sqrt{2\pi s^3}} \exp\left[-\frac{(x+\mu s)^2}{2s}\right] \qquad x>0 ,
+\end{aligned}
+$$
+(this is the density of the first time a BM with drift $\mu$ started at $x$ hits 0),
+then
+$$
+\begin{aligned}
+p_t(x,z) 
+&= \P\{ X_t = z \mid X_0 = x \}\\
+&= \begin{cases}
+    2 \int_0^\infty \int_0^t h(t-\tau;b-z,0) h(\tau;x+b,\theta) d\tau db
+        & \qquad x \ge 0, \quad z\le 0 \\
+    2 \int_0^\infty \int_0^t e^{2\theta z} h(t-\tau;b,0) h(\tau;x+b,\theta) d\tau db & \\
+    \qquad {} + \frac{1}{\sqrt{2\pi t}} 
+        \left\{ 
+            \exp\left[ -\frac{ (x-z+\theta t)^2 }{2t} \right]
+            +\exp\left[ -\frac{ (x+z-\theta t)^2 }{2t} - 2 \theta x \right]
+        \right\}
+        & \qquad x \ge 0, \quad z > 0 .
+\end{cases}
+\end{aligned}
+$$
+Also, $p_t(x,z;\theta)=p_t(-x,-z;-\theta)$ gets the same expressions for $x<0$.

### 2015-10-15 notes/diffusion-local-times.Rmd
more comparisons

@@
 fig.dim <- 5
-knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
+knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
 ```
@@
     return(rws)
+}
+weighted_time <- function (rws, weight.fn, ...) {
+    colSums( weight.fn(rws,...) ) * attr(rws,"dt")
 }
@@
 ```{r tanh_cline}
+tanh_cline
 s <- 0.1
-tanh_cline
 xx <- seq(-15,15,length.out=500)
@@
 
-```{r sim_general, fig.height=3*fig.dim}
-weighted_time <- function (rws, weight.fn, ...) {
-    colSums( weight.fn(rws,...) ) * attr(rws,"dt")
-}
+Here are distributions of $T=\int_0^{`r tau`} f_H(X_t) dt$
+for two different starting points: $x=0$ and $x=-3$,
+and five different models:
+
+- **tanh:** the ``correct'' model;
+- **reflected:** a Brownian motion reflected at zero and counting one-half time in a neighborhood of size $\sigma/\sqrt{s}$;
+- **reflected tanh:** a Brownian motion reflected at zero and counting time weighted by the $\tanh$ cline;
+- **pushed:** a Brownian motion pushed back at rate $4\sigma\sqrt{s}$ above zero and counting time above zero;
+- **pushed tanh:** a Brownian motion pushed back at rate $4\sigma\sqrt{s}$ above zero and counting time weighted by the $\tanh$ cline.
+
+```{r sim_general, fig.height=3*fig.dim, cache=TRUE}
 nwalks <- 1e4
@@
 tau <- 10
-lt.comparison <- list(
+initvals <- c(0,-3)
+lt.comparison <- lapply( initvals, function (xinit) { list(
         tanh=weighted_time( 
-               sim_diffusion( tanh_drift,tau,nwalks,ntimes,s=s ),
+               sim_diffusion( tanh_drift,tau,nwalks,ntimes,s=s, xinit=xinit ),
                tanh_het, s=s
@@
         reflected=weighted_time(
-                sim_diffusion( function(...){0}, tau, nwalks, ntimes ),
+                -abs(sim_diffusion( function(...){0}, tau, nwalks, ntimes, xinit=xinit )),
                 function (x,s) { ifelse(abs(x)<1/(2*sqrt(s)), 1/2, 0) }, s=s
@@
         reflected_tanh=weighted_time(
-                sim_diffusion( function(...){0}, tau, nwalks, ntimes ),
+                -abs(sim_diffusion( function(...){0}, tau, nwalks, ntimes, xinit=xinit )),
                 tanh_het, s=s
@@
         pushed=weighted_time(
-                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s ),
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s, xinit=xinit ),
                 function (x) { x>0 }
@@
         pushed_tanh=weighted_time(
-                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s ),
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s, xinit=xinit ),
                 tanh_het, s=s
             )
+        ) } )
+layout(matrix(1:(length(lt.comparison)*length(lt.comparison[[1]])),ncol=2))
+for (k in seq_along(lt.comparison)) {
+    for (j in seq_along(lt.comparison[[k]])) {
+        hist(lt.comparison[[k]][[j]],xlim=range(lt.comparison),main=names(lt.comparison[[k]])[j],breaks=seq(0,tau,length.out=50),xlab="time in hets")
+    }
+}
+```
+
+Of course, we aren't actually trying to estimate the distribution of $T$ itself,
+but rather the probabilities of recombination (the Laplace transform).
+Here are these probabilities, as a function of $r$
+(the second plots are on a log-log scale):
+```{r recomb_prob_comparisons, cache=TRUE, depends="sim_general"}
+rvals <- exp(seq(log(.001),log(0.5),length.out=400))
+precombs <- lapply( lt.comparison, function (lts) {
+        sapply( lts, function (lt) sapply( rvals, function (r) { mean((-1)*expm1(-r*lt)) } ) )
+    } )
+layout(t(1:2))
+for (k in seq_along(precombs)) {
+    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
+    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+}
+for (k in seq_along(precombs)) {
+    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), log='xy', main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1  )
+    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+}
+```
+
+Here's the same thing for a much longer time:
+```{r sim_general_longer, fig.height=3*fig.dim, cache=TRUE}
+nwalks <- 1e3
+ntimes <- 1e4
+s <- 0.1
+tau <- 100
+initvals <- c(0,-3)
+lt.comparison <- lapply( initvals, function (xinit) { list(
+        tanh=weighted_time( 
+               sim_diffusion( tanh_drift,tau,nwalks,ntimes,s=s, xinit=xinit ),
+               tanh_het, s=s
+           ),
+        reflected=weighted_time(
+                -abs(sim_diffusion( function(...){0}, tau, nwalks, ntimes, xinit=xinit )),
+                function (x,s) { ifelse(abs(x)<1/(2*sqrt(s)), 1/2, 0) }, s=s
+            ),
+        reflected_tanh=weighted_time(
+                -abs(sim_diffusion( function(...){0}, tau, nwalks, ntimes, xinit=xinit )),
+                tanh_het, s=s
+            ),
+        pushed=weighted_time(
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s, xinit=xinit ),
+                function (x) { x>0 }
+            ),
+        pushed_tanh=weighted_time(
+                sim_diffusion( function(x,s){ifelse(x>0,-4*sqrt(s),0)}, tau, nwalks, ntimes, s=s, xinit=xinit ),
+                tanh_het, s=s
             )
-layout(seq_along(lt.comparison))
-for (k in seq_along(lt.comparison)) {
-    hist(lt.comparison[[k]],xlim=range(lt.comparison),main=names(lt.comparison)[k],breaks=50,xlab="time in hets")
+        ) } )
+tbreaks <- seq(0,tau,length.out=50)
+Tdists <- lapply( lt.comparison, function (lts) {
+        sapply( lts, function (lt) tabulate(findInterval(lt,tbreaks),nbins=length(tbreaks)-1) )
+    } )
+rvals <- exp(seq(log(.001),log(0.5),length.out=400))
+precombs <- lapply( lt.comparison, function (lts) {
+        sapply( lts, function (lt) sapply( rvals, function (r) { mean((-1)*expm1(-r*lt)) } ) )
+    } )
+layout(t(1:2))
+for (k in seq_along(precombs)) {
+    matplot( tbreaks[-1]-diff(tbreaks)/2, Tdists[[k]], type='l', col=1:ncol(Tdists[[k]]), main=paste("x =",initvals[k]), ylim=range(Tdists), xlab="time in hets", lty=1 )
+    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+}
+for (k in seq_along(precombs)) {
+    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
+    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
 }

### 2015-10-16 notes/diffusion-local-times.Rmd
.

@@
 knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
+library(ReacTran)
+library(deSolve)
 ```
@@
 hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
+```
+
+The exact solution
+------------------
+
+Let
+$$
+\zeta(x,t,r) = \P\{ T_r > t \} .
+$$
+Then the Kolmogorov forward equation tells us that $\zeta$ solves the following differential equation:
+$$\begin{aligned}
+    \frac{d}{dt} \zeta(x,t,r) 
+    &= 
+    \frac{\sigma^2}{2} \frac{d^2}{dx^2} \zeta(x,t,r) + \frac{d}{dx} \log(p(x)) \frac{d}{dx} \zeta(x,t,r)  - r f_H(x) \zeta(x,t,r)  \\
+    \zeta(x,0,r) = 0 .
+\end{aligned}$$
+
+So, let's solve that, numerically.
+We'll use the `ReacTran` library:
+```{r solve_pde}
+thegrid <- setup.grid.1D(x.up=-10, x.down=10, N = 1000)
+driftfn <- setup.prop.1D(grid=thegrid, func=tanh_drift, s=s)
+hetfn <- setup.prop.1D(grid=thegrid, func=tanh_het, s=s)
+# ode.1D( )
 ```

### 2015-10-21 sims/yaniv_sims.R
added sim figures

@@ -55,6 +55,10 @@ my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.an
 	#my.genos is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.
 freqs = lapply(my.genos,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
 
+matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
+matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+
 ###########################
 ###########################
 #To get distribution of chunks of ancestry B to the (wlog) right of the selected locus:
@@ -77,6 +81,25 @@ get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,P
 
 #output is a nind*2 matrix, where each column is a chromosome. 
 #EXAMPLE: testB = lapply(1:50,function(X){get.deme.chunks(DEME=X,ancA=F)})
+#D = 23; hist(testB[[D]][which(testB[[D]]<1 & testB[[D]]>0)], col="black",breaks=seq(0,1,0.05))
+
+add.alpha <- function(col, alpha=1){
+if(missing(col))
+stop("Please provide a vector of colours.")
+apply(sapply(col, col2rgb)/255, 2, 
+function(x) 
+rgb(x[1], x[2], x[3], alpha=alpha)) 
+
+}
+
+transparent_rainbow = add.alpha(rainbow(ndemes),0.75)
+
+D = 1; hist(testB[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")
+
+for(D in seq(3,ndemes,5)){
+	hist(testB[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
+}
+
 ######################
 ######################
 #Get LD between two specified loci:
[…]

### 2015-10-21 notes/diffusion-local-times.Rmd

@@
 \DeclareMathOperator{\sgn}{sgn}
+\newcommand{\grad}{\nabla}
 $$
@@
 }
-weighted_time <- function (rws, weight.fn, ...) {
-    colSums( weight.fn(rws,...) ) * attr(rws,"dt")
+weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
+    colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
 }
@@
 $$\begin{aligned}
-    \frac{d}{dt} \zeta(x,t,r) 
+    \frac{\partial}{\partial t} \zeta(x,t,r) 
     &= 
-    \frac{\sigma^2}{2} \frac{d^2}{dx^2} \zeta(x,t,r) + \frac{d}{dx} \log(p(x)) \frac{d}{dx} \zeta(x,t,r)  - r f_H(x) \zeta(x,t,r)  \\
-    \zeta(x,0,r) = 0 .
+    \frac{\sigma^2}{2} \frac{\partial^2}{\partial x^2} \zeta(x,t,r) + \frac{\partial }{\partial x} \log(p(x)) \frac{\partial }{\partial x} \zeta(x,t,r)  - r f_H(x) \zeta(x,t,r)  \\
+    \zeta(x,0,r) = 1 .
 \end{aligned}$$
@@
 So, let's solve that, numerically.
-We'll use the `ReacTran` library:
-```{r solve_pde}
-thegrid <- setup.grid.1D(x.up=-10, x.down=10, N = 1000)
-driftfn <- setup.prop.1D(grid=thegrid, func=tanh_drift, s=s)
+We'll use the `ReacTran` library,
+which in 1D solves equations of the form
+$$\begin{aligned}
+   \frac{\partial C}{\partial t} 
+   &= 
+   \frac{1}{A} \grad \cdot A \left( D \grad C - v C \right) \\
+   &= 
+   \left( \frac{1}{A} \grad AD - v \right) \cdot \grad C 
+   - \frac{1}{A} \grad A \cdot v C 
+   + D \Delta C \\
+\end{aligned}$$
+which is the same as what we have if we let $D=\sigma^2/2$ and $A=p$ and $v=0$.
+```{r solve_pde, cache=TRUE}
+thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 1000)
+clinefn <- setup.prop.1D(grid=thegrid, func=tanh_cline, s=s)
 hetfn <- setup.prop.1D(grid=thegrid, func=tanh_het, s=s)
-# ode.1D( )
+xx <- seq(-15,15,length.out=500)
+pde.fn <- function (t,y,parms,...) {
+    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+    list( tran - r * hetfn$mid * y )
+}
+yinit <- rep(1.0,thegrid$N)
+tt <- seq(0,100,length.out=500)
+r <- 1
+pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
+image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
 ```
+
+It's reassuring the solution is between 0 and 1.
+Let's spot check that with the simulation.
+Here it is at $t=`r tau`$:
+```{r check_pde_soln, cache=TRUE}
+check.t <- function (alpha) {
+    t.tau <- which.min(abs(tt-alpha*tau))
+    probs <- exp( - weighted_time( rws, tanh_het, s=s, use.t=1:floor(nrow(rws)*alpha) ) )
+    plot(rws[1,],probs,xlab='initial location', ylab=paste('probability of no recomb by t=',alpha*tau),ylim=c(0,1))
+    lines(lowess(rws[1,],probs,f=0.05),col='red',lwd=2)
+    lines(thegrid$x.int,pde.soln[t.tau,],col='green',lwd=2)
+    legend("topright",lwd=2,col=c("red","green"),legend=c("mean","numerical solution"))
+}
+check.t(1)
+```
+and, at other values of $t$:
+```{r check_pde_soln_2, cache=TRUE, depends="check_pde_soln"}
+layout(t(1:2))
+check.t(0.05)
+check.t(0.1)
+check.t(0.25)
+check.t(0.5)
+```
+**Looks pretty good.**
 
@@
 
-- **tanh:** the ``correct'' model;
+- **tanh:** the "correct" model;
 - **reflected:** a Brownian motion reflected at zero and counting one-half time in a neighborhood of size $\sigma/\sqrt{s}$;

### 2015-10-21 sims/yaniv_sims.R
marking down simulations

@@ -55,9 +55,12 @@ my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.an
 	#my.genos is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.
 freqs = lapply(my.genos,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
 
+
+#pdf(file="freqplot.pdf")
 matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
 legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
 matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+#dev.off()
 
 ###########################
 ###########################

### 2015-10-21 sims/run-sim-Alisa.Rmd
marking down simulations

--- 2015-09-13 sims/run-sim.Rmd
@@
 ---
-title: "Running the simulation"
+title: "Running simulations by Alisa"
+author: Alisa
 date: "`r date()`"
+output: html_document
 ---
@@
 -------------------------
+To simulate, we read in files from sim-fns.R
+```{r load_fns}
+	source("sim-fns.R")
+```
+
+Start off with simulation of 50 generation zone with 50 demes of 500 individuals each and sigma of 1:
+```{r set_deme_parameters}
+	zone_age = 50
+	ndemes = 50
+	deme_size = 500
+	ninds = ndemes*deme_size
+	SIGMA = 1
+```
+note that zone_age above can be a vector of times.
+
+Now to define selection. Here we selection against heterozygotes to be strong (0.5).
+```{r set_selection}
+	zone_age = c(50)
+	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
+```
+`qtl` is a named list, one element per chromosome,
+and each element of the list is a data frame
+whose columns give the name of the locus,
+the selection coefficient,
+and the position in Morgans.
+
+Running the simulation
+-------------------------
+
+This is a two-step process. First we simulate genotypes:
+```{r run_sim_step1}
+	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
+```
+The output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
+	where:
+		"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+		"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
+		"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
+		"pars" is an integer that gives the number of generatsion of the simulation
+		"QTL" returns the QTL vector used in the simulation
+
+The second step is to parse the simulated genotypes:
+```{r run_sim_step2}
+	sims.sums = lapply(sims.s0.1,spBreaks)
+```
+The output here is a list: [[zone_age]][[c("ind.ancestry","ancest.prop")]]
+		"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
+		"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
+
+This resulting files can get quite large, but we've saved them here
+```{r save_file}
+	save(sims.sums,file="~/Projects/HybridZones/ClineProjects/sims_sums_s0.5_sigma1_50*500_tau50.Robj")
+```
+
+Getting relevant statistics
+-------------------------
+Let's start by defining a vector of loci at which we will measure things:
+
+```{r focal_loci}
+	loci = list(seq(0.5,1,0.05))
+```
+
+Now, we get frequencies at each of the sites in the above vector, 
+first by getting genotypes at all loci, then using those to calculate an allele frequency for each deme:
+```{r get_frequencies}
+	genotypes = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
+	freqs = lapply(my.genos,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
+
+		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
+		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+
+```
+(`genotypes` is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
+Each line in the plot represents a locus a given distance from the selected locus.
+
+We can compare these frequencies to predictions made by theory :
+
+```{r peters_predictions}
+	zeta <- function(x,r){
+	2*pnorm(abs(x)) - 1 + 2 * exp(r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE))	
+	}
+
+	pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
+	(1/2)*(1-sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau))	
+	}
+
+	rr=(loci[[1]]-0.5)*SIGMA/sqrt(S) # loci at which frequency is measured
+	xx <- (1:ndemes)-0.5-ndemes/2	 # deme positions
+	pp <- do.call(cbind,lapply(rr,function(R){pcline(xx,r=R,tau=zone_age,sigma=SIGMA)})) #allele frequency
+
+		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
+		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+
+```
+
+Next, we think about tract lengths in these populations
 

### 2015-10-21 notes/diffusion-local-times.Rmd

@@
 ```
+
+The goal
+========
+
+The key quantity we compute with diffusion theory 
+is the probability that a lineage, 
+begun at location $x$ and killed at rate $r$ when in heterozygotes,
+is still alive at time $\tau$.
+Denote this by
+$$
+\begin{aligned}
+    \zeta(x,r,\tau,\sigma).
+\end{aligned}
+$$
+
+Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
+at a location at recombination distance $r$ from the selected site
+inherits from the left-hand side of the barrier
+(i.e., an ancestor on $(-\infty,0)$)
+$\tau$ generations ago.
+(Note that $r$ should be scaled by the density and fecundity of heterozygotes in the cline --
+see below.)
+For an individual to the left of the barrier ($x<0$),
+the probability of inheriting from the left of the barrier at $\tau$
+is the probability that no recombination at the barrier has occurred,
+plus one-half the probability that recombination in a heterozygote has occurred,
+i.e., $\zeta +(1-\zeta)/2=(1+\zeta)/2$.
+If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
+i.e., $(1-\zeta)/2$.
+Combining these,
+the probability of ending up on the left-hand side is
+$$
+\begin{aligned}
+    p(x,r,\tau,A) = 
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
+\end{aligned}
+$$
 
@@
 nwalks <- 1e4
-ntimes <- 1e3
+ntimes <- 4e3
 s <- 0.1
-tau <- 10
+tau <- 5
 xx <- seq(-15,15,length.out=500)
@@
 We'll use the `ReacTran` library,
-which in 1D solves equations of the form
+which in 1D solves equations with advection-diffusion parts of the form:
 $$\begin{aligned}
@@
 which is the same as what we have if we let $D=\sigma^2/2$ and $A=p$ and $v=0$.
-```{r solve_pde, cache=TRUE}
-thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 1000)
+Here is the setup:
+```{r solve_pde_setup, cache=TRUE}
+thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
 clinefn <- setup.prop.1D(grid=thegrid, func=tanh_cline, s=s)
 hetfn <- setup.prop.1D(grid=thegrid, func=tanh_het, s=s)
-xx <- seq(-15,15,length.out=500)
+yinit <- rep(1.0,thegrid$N)
+tt <- seq(0,tau,length.out=100)
+r <- 1
+```
+
+First, two quick checks: First, by setting $f_H=1$,
+we should have the solution $u(x,t,r) = \exp(-tr)$.
+Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
+```{r solve_pde_check, cache=TRUE, depends="solve_pde_setup"}
+layout(t(1:2))
+test.hetfn <- setup.prop.1D(grid=thegrid, func=function(x){rep(1,length(x))})
+pde.fn <- function (t,y,parms,...) {
+    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+    list( tran - r * test.hetfn$mid * y )
+}
+pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
+matplot(pde.soln,type='l',ylab="probability of no recomb", col='black',log='y')
+lines(tt,exp(-r*tt),lwd=2,lty=2,col='red')
+## no diffusion
+pde.fn <- function (t,y,parms,...) {
+    tran <- tran.1D(C=y, A=clinefn, D=1e-6, dx=thegrid)$dC
+    list( tran - r * hetfn$mid * y )
+}
+pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
+plot.x <- sort(match(sample(thegrid$x.mid[thegrid$x.mid>-5 & thegrid$x.mid<5],12),thegrid$x.mid))
+matplot(pde.soln[,plot.x],type='l',ylab="probability of no recomb", col=rainbow(length(plot.x)), lty=1 )
+matpoints( exp(outer( tt, -r*hetfn$mid[plot.x], "*" )), lty=4, col=rainbow(length(plot.x)), pch=20 )
+```
+**Looks good.**
+
+Now, to solve the real problem:
+```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
 pde.fn <- function (t,y,parms,...) {
@@
 }
-yinit <- rep(1.0,thegrid$N)
-tt <- seq(0,100,length.out=500)
-r <- 1
 pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
@@
 layout(t(1:2))
-check.t(0.05)
 check.t(0.1)
-check.t(0.25)
-check.t(0.5)
+check.t(0.2)
+check.t(0.4)
+check.t(0.8)
 ```
@@
     matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
-    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+    legend("topright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
 }
@@
     matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), log='xy', main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1  )
-    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+    legend("bottomright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
 }
@@
 Here's the same thing for a much longer time:
-```{r sim_general_longer, fig.height=3*fig.dim, cache=TRUE}
+```{r sim_general_longer, fig.height=fig.dim, cache=TRUE}
 nwalks <- 1e3

### 2015-10-22 notes/diffusion-local-times.Rmd

@@
 ```{r setup_fns}
-pcline <- function (x,r,tau=1,sigma=1) {
-    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
+pcline <- function (x,r,recombfn,...) {
+    (1/2)*( 1 - sign(x)*recombfn(x,2*r,...) )
 }
@@
     (1-s)*(1-pp)/(1-s*(1-pp))
+}
+zeta.reflected <- function (x,r) {
+    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
 }
@@
 }
+clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),...) {
+    # plot output of deSolve as clines in a nice way
+    graphics::matplot( x, t(soln[times,-1]), type='l', col=rainbow(length(times)), lty=1, 
+            xlab="space", ylab=ylab, ylim=ylim, ...)
+    legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
+    return( invisible( t(soln[times,-1]) ) )
+}
 ```
@@
 is the probability that a lineage, 
-begun at location $x$ and killed at rate $r$ when in heterozygotes,
+begun at location $x$ and killed at rate $k$ when in heterozygotes,
 is still alive at time $\tau$.
@@
 \begin{aligned}
-    \zeta(x,r,\tau,\sigma).
+    \zeta(x,k,\tau).
 \end{aligned}
 $$
-
-Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
+Now, let $q(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
 at a location at recombination distance $r$ from the selected site
@@
 $\tau$ generations ago.
-(Note that $r$ should be scaled by the density and fecundity of heterozygotes in the cline --
-see below.)
-For an individual to the left of the barrier ($x<0$),
+The probability that a recombination occurs 
+between the lineage and the selected locus on either pair of chromatids
+is $2r$,
+and if this happens, the lineage ends up on either background with probability $1/2$.
+Therefore,
+for an individual to the left of the barrier ($x<0$),
 the probability of inheriting from the left of the barrier at $\tau$
@@
 plus one-half the probability that recombination in a heterozygote has occurred,
-i.e., $\zeta +(1-\zeta)/2=(1+\zeta)/2$.
+so if $\zeta = \zeta(x,2r,\tau)$, then
+$q=\zeta +(1-\zeta)/2=(1+\zeta)/2$.
 If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
-i.e., $(1-\zeta)/2$.
+i.e., $q=(1-\zeta)/2$.
 Combining these,
@@
 \begin{aligned}
-    p(x,r,\tau,A) = 
-    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
+    q(x,r,\tau,A) = 
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,2r,\tau) \right\} .
 \end{aligned}
@@
 }
-pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
+pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
 image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
@@
 **Looks pretty good.**
+
+The even more right solution
+============================
+
+Actually, the cline gets set up
+in a time-inhomogeneous way.
+Let's do that, by numerical solution,
+then run the lineage back through it.
+The forwards-time evolution of the cline,
+with isotropic density and migration,
+solves the PDE (Barton 1979):
+$ \begin{align*}
+  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 + sp(x,t)(1-p(x,t))(2p(x,t)-1) .
+\end{align*} $
+```{r fwds_pde, cache=TRUE}
+thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
+yinit <- ifelse( thegrid$x.mid>0, 0, 1 )
+tt <- seq(0,tau,length.out=100)
+fwds.pde <- function (t,y,parms,...) {
+    tran <- tran.1D(C=y, D=1/2, dx=thegrid)$dC
+    list( tran + s * y * (1-y) * (2*y-1) )
+}
+fwds.soln <- ode.1D( y=yinit, times=tt, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
+image(fwds.soln,legend=TRUE,main="mean cline development", yaxt='n', ylab='space')
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+```
+Now, let's run the killed lineage backwards in it:
+```{r fwds_back_pde, cache=TRUE, depends="fwds_pde"}
+yinit <- rep(1.0,thegrid$N)
+cline.interp <- function (T) {
+    # return the interpolated cline for a particular time
+    which.t <- findInterval(T,fwds.soln[,1],rightmost.closed=TRUE)
+    lu <- fwds.soln[c(which.t,which.t+1),1]
+    alpha <- (T-lu[1])/diff(lu)
+    return( (1-alpha)*fwds.soln[which.t,-1] + alpha*fwds.soln[which.t+1,-1]  )
+}
+rev.pde <- function (t,y,parms,...) {
+    p <- cline.interp(tau-t)
+    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline.interp(tau-t)), interp="linear")
+    tran <- tran.1D(C=y, A=p, D=1/2, dx=thegrid)$dC
+    list( tran - r * (1-p) * y )
+}
+rev.soln <- ode.1D( y=yinit, times=tt, func=rev.pde, nspec=1, tcrit=tau ) # note FIRST COLUMN IS TIME
+image(rev.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+```
+
+Let's compare the case where the cline is set up instantly or not:
+```{r compare_rightness, cache=TRUE,depends=c("fwds_back_pde","solve_pde_setup","setup_fns")}
+layout(t(1:2))
+fb.p <- clineplot(rev.soln,main="forwards-backwards")
+st.p <- clineplot(pde.soln,main="stationary cline")
+layout(1)
+plot(fb.p,st.p,asp=1,pch=20,cex=0.5,xlab="forwards-backwards",ylab="stationary",col=rainbow(ncol(fb.p))[col(fb.p)])
+abline(0,1,lwd=3)
+```
 
@@
 ```{r killing, cache=TRUE, depends="sim_rw"}
+zeta.reflected
 loc <- 0.1
-zeta <- function (x,r) {
-    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
-}
 compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
@@
           simulation=mean(killed), 
-          theory=zeta(loc,rate),
+          theory=zeta.reflected(loc,rate),
           SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
@@
 rvals <- seq(0,10,length.out=100)
-zetavals <- outer( xvals, rvals, zeta )
+zetavals <- outer( xvals, rvals, zeta.reflected )
 matplot(xvals,zetavals,type='l',xlab="x",ylab=expression(zeta),col=rainbow(length(rvals)))
@@
 
-Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
-at a location at recombination distance $r$ from the selected site
-inherits from the left-hand side of the barrier
-(i.e., an ancestor on $(-\infty,0)$)
-$\tau$ generations ago.
-(Note that $r$ should be scaled by the density and fecundity of heterozygotes in the cline --
-see below.)
-For an individual to the left of the barrier ($x<0$),
-the probability of inheriting from the left of the barrier at $\tau$
-is the probability that no recombination at the barrier has occurred,
-plus one-half the probability that recombination in a heterozygote has occurred,
-i.e., $\zeta +(1-\zeta)/2=(1+\zeta)/2$.
-If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
-i.e., $(1-\zeta)/2$.
-Combining these,
-the probability of ending up on the left-hand side is
+Recall from above that the predicted cline is
 $$
 \begin{aligned}
-    p(x,r,\tau,A) = 
-    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
+    q(x,r,\tau,A) = 
+    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,2r,\tau) \right\} .
 \end{aligned}
@@
 ```{r show_cline, cache=TRUE}
-pcline
+zeta.pcline <- function (x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) {
+    (1/2)*( 1 - sign(x)*zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau) )
+}
 xx <- seq(-15,15,length.out=500)
 for (tau in c(5,50)) {
-    plot(xx, pcline(xx,r=1), type='n', 
+    plot(xx, zeta.pcline(xx,r=1), type='n', 
             xlab="distance from cline center", ylab="prob inherited from the left", 
@@
     rr <- seq(0,0.2,length.out=11)
-    pp <- sapply( rr, function (r) { pcline(xx,r=r,tau=tau) } )
+    pp <- sapply( rr, function (r) { zeta.pcline(xx,r=r,tau=tau) } )
     matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)

### 2015-10-22 notes/diffusion-local-times.Rmd
.

@@
 solves the PDE (Barton 1979):
-$ \begin{align*}
-  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 + sp(x,t)(1-p(x,t))(2p(x,t)-1) .
-\end{align*} $
+$$\begin{aligned*}
+  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
+\end{aligned*}$$
 ```{r fwds_pde, cache=TRUE}
@@
 ```
+
+Just forwards
+-------------
+
+An alternate route to this would have been to
+solve the coupled PDE for the frequencies of the selected locus 
+and the linked locus.
 
@@
     matplot( tbreaks[-1]-diff(tbreaks)/2, Tdists[[k]], type='l', col=1:ncol(Tdists[[k]]), main=paste("x =",initvals[k]), ylim=range(Tdists), xlab="time in hets", lty=1 )
-    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+    legend("bottomright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
 }
@@
     matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
-    legend("topleft",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
+    legend("bottomright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
 }
@@
 ```{r show_cline, cache=TRUE}
-zeta.pcline <- function (x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) {
-    (1/2)*( 1 - sign(x)*zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau) )
-}
+zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1)
 xx <- seq(-15,15,length.out=500)

### 2015-10-22 sims/run-sim-Alisa.Rmd
got .rmd to work-ish!

--- 2015-09-12 sims/run-sim.Rmd
@@
 -------------------------
+To simulate, we read in files from sim-fns.R
+```{r load_fns}
+	source("sim-fns.R")
+```
 
-First we need to decide where the loci under selection are.
-Here `qtl` is a named list, one element per chromosome,
+Start off with simulation of 50 generation zone with 50 demes of 500 individuals each and sigma of 1:
+```{r set_deme_parameters}
+	zone_age	= 50
+	ndemes 		= 50
+	deme_size 	= 500
+	ninds 		= ndemes*deme_size
+	SIGMA 		= 1
+```
+note that zone_age above can be a vector of times.
+
+Now to define selection. Here we selection against heterozygotes to be strong (0.5).
+```{r set_selection}
+	S  = 0.5
+	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
+```
+`qtl` is a named list, one element per chromosome,
 and each element of the list is a data frame
@@
 
-```{r setup_qtl}
-qtl=list(
-         chr1=data.frame(
-                    traits = c("underdominant"),
-                    s = c(0.1),
-                    pos = c(0.5)
-                )
-    )
+Running the simulation
+-------------------------
+
+This is a two-step process. First we simulate genotypes:
+```{r run_sim_step1}
+	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
+	#This can take a while, 
+```
+The output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
+	where:
+		"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+		"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
+		"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
+		"pars" is an integer that gives the number of generatsion of the simulation
+		"QTL" returns the QTL vector used in the simulation
+
+The second step is to parse the simulated genotypes:
+```{r run_sim_step2}
+	sims.sums = lapply(sims.s0.1,spBreaks)
+```
+The output here is a list: [[zone_age]][[c("ind.ancestry","ancest.prop")]]
+		"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
+		"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
+
+This resulting files can get quite large, but we've saved them here
+```{r save_file,cache=TRUE}
+	save(sims.s0.1,file="sims_sums_s0.5_sigma1_50by500_tau50.Robj")
 ```
 
-Now, let's set up the population,
-which means specifying the number of individuals,
-the number of demes,
-the width of the zone in real units,
-and the dispersal SD ($\sigma$).
-Our locus has a selection coefficient of `r qtl[[1]]$s`,
-so the cline width will be about `r round(1/sqrt(qtl[[1]]$s),digits=1)`$\sigma$ wide;
-so we'll take our zone to be four times that,
-and measure distance in units of $\sigma$:
-```{r setup_pop}
-zone.width <- ceiling(4/sqrt(qtl[[1]]$s))  # in real units
-nind <- 1000
-ndeme <- 50
-sigma <- 1 # in real units
-deme.spacing <- zone.width/ndeme
-sigma.demes <- sigma/deme.spacing # in units of demes
+Getting relevant statistics
+-------------------------
+Let's start by defining a vector of loci at which we will measure things:
+
+```{r focal_loci}
+	loci = list(seq(0.5,1,0.05))
 ```
-The discretized dispersal kernel moves as $\lfloor Z+1/2 \rfloor$,
-where $Z \sim N(0,\sigma_\text{demes}^2)$;
-let's check this has SD close to $\sigma$:
-```{r check_sigma}
-zz <- floor( rnorm( 1e6, mean=1/2, sd=sigma.demes ) )*deme.spacing
-sd(zz)
+
+Now, we get frequencies at each of the sites in the above vector, 
+first by getting genotypes at all loci, then using those to calculate an allele frequency for each deme:
+```{r get_frequencies}
+	genotypes = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
+	freqs = lapply(genotypes,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
+	xx <- (1:ndemes)-0.5-ndemes/2
+		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
+		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+
 ```
-That's pretty good.
+(`genotypes` is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
+Each line in the plot represents a locus a given distance from the selected locus.
 
-Running a simulation
---------------------
+We can compare these frequencies to predictions made by theory :
 
-To run the simulation,
+```{r peters_predictions}
+	zeta <- function(x,r){
+	2*pnorm(abs(x)) - 1 + 2 * exp(r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE))	
+	}
+
+	pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
+	(1/2)*(1-sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau))	
+	}
+
+	rr=(loci[[1]]-0.5)*SIGMA/sqrt(S) # loci at which frequency is measured
+	xx <- (1:ndemes)-0.5-ndemes/2	 # deme positions
+	pp <- do.call(cbind,lapply(rr,function(R){pcline(xx,r=R,tau=zone_age,sigma=SIGMA)})) #allele frequency
+
+		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
+		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
+
+```
+
+Next, we think about tract lengths in these populations. The following generates the distribution of lengths 
+of continuous ancestry B to the right of the specified locus (`focal_sites`) (in this case the selected locus). 
+Note that if the specified locus is of ancestry A, then the length of ancestry B is 0. The first function finds
+and measures the chunk in an individual that contains the focal site. The second function generates a 
+distribution for each deme.
+
+```{r tract_length_functions}
+	focal_sites = lapply(qtl,function(Z){Z$pos})
+
+	get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
+		#return the interval containing focal site for a individual	
+		chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+		identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+		replace(chunk,which(identity==ancA),0)
+	}
+
+	get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
+		#get distribution of chunks within a deme
+		INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
+		intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
+		return(intervals)
+	}
+
+```
+the output of get.deme.chunks is a nind*2 matrix, where each column is a chromosome. To run:
+
+```{r get_tract_lengths}
+	tracts = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=F)})
+
+		#function to increase transparency of colors
+		add.alpha <- function(col, alpha=1){
+		if(missing(col))
+		stop("Please provide a vector of colours.")
+		apply(sapply(col, col2rgb)/255, 2, 
+		function(x) 
+		rgb(x[1], x[2], x[3], alpha=alpha)) 
+		}
+
+	transparent_rainbow = add.alpha(rainbow(ndemes),0.75)
+
+		D = 1; hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")
+
+		for(D in seq(1,ndemes,5)){
+		hist(tracts[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
+		}
+```
+To get LD in these populations, we define `get.genotype` to get the genotypes at two loci in individuals, and `get.LD` to calculate LD in a deme:
+```{r LD_functions}
+ 	get.genotype = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS1=0.5,POS2=0.5){
+	site1 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS1 & X$stops>POS1),3]})	
+	site2 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS2 & X$stops>POS2),3]})
+	return(rbind(site1,site2))
+	}
+
+	get.LD = function(IND_DATA= sims.sums[[1]]$ind.ancest,DEME=1,CHR=1,POS1 = 0.5, POS2=0.1){
+	INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
+	genotypes = do.call(cbind,lapply(INDS,get.genotype,CHR=CHR,POS1=POS1,POS2=POS2))
+	freqB1 = length(which(genotypes[1,]))/length(genotypes[1,])
+	freqB2 = length(which(genotypes[2,]))/length(genotypes[2,])
+	freqB12 = length(which(genotypes[1,] & genotypes[2,]))/length(genotypes[1,])
+	LD = freqB12 - freqB2*freqB1
+	return(LD)
+}
+
+```
+We can just calculate LD over demes and loci pairs. Here's an example of LD between pairs of loci where one member of the
+pair is selected on. This gives a plot of decay in population LD as we get further away from the center of the zone.
+
+```
+LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:ndemes,get.LD,POS1=0.5,POS2=Z,CHR=1,IND_DATA=sims.sums[[1]]$ind.ancest)}))
+	matplot(xx,LD_matrix,type="l",col=rainbow(20),xlim=c(-10,10),lty=1)
+
+```
+
+Simulate a realistic population (with Populus as a model)
+-------------------------
+We propose that the density of Populus along a one-dimensional transect is one every 10 meters. Additionally, based on a
+literature search that suggests most pollen only makes it to the nearest neighbor and seeds do not disperse much, with 
+potentially a long tail, we propose a sigma of 100m. Furthermore, based on HZ width of 20km for most of the geome, we 
+propose that width at selected loci may be 1km, giving a selection coefficient of 0.01 (based on width = 2*sigma/sqrt(S)). 
+We space demes ever 1km for 100km, with 100 individuals per deme (given above density of 1/10 inds/meter). In units of demes, sigma is 10.
+
+```{r set_populus_parameters,cache=TRUE}
+	zone_age	= 10
+	ndemes 		= 100
+	deme_size 	= 100
+	ninds 		= ndemes*deme_size
+	SIGMA 		= 0.1
+	S 			= 0.01
+	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
+
+	sims.populus_like.10gens = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
+	sims.sums.populus_like.10gens = lapply(sims.populus_like.10gens,spBreaks)
+	save(sims.populus_like.10gens,file="simspopulus_like.10gens.Robj")
+```

### 2015-10-22 notes/diffusion-local-times.Rmd
boundary conditions

@@
 }
-clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),...) {
+clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),legend=TRUE,...) {
     # plot output of deSolve as clines in a nice way
@@
             xlab="space", ylab=ylab, ylim=ylim, ...)
-    legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
+    if (legend) legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
     return( invisible( t(soln[times,-1]) ) )
@@
 eps <- 1/sqrt(ntimes)
-rws <- matrix( rnorm(nwalks*ntimes)*eps, nrow=ntimes, ncol=nwalks )
-rws <- rbind(0,apply(rws,2,cumsum))
+rws <- rbind(rep(0,nwalks), matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=ntimes-1, ncol=nwalks ) )
+rws <- apply(rws,2,cumsum)
 matplot(seq(0,1,length.out=ntimes),rws[,1:20],type='l', ylab='position')
@@
 
-This is very steep at $x=0$.
-Notice that as $x$ goes to 0,
-$$
-\begin{aligned}
-    \lim_{x \to 0} \zeta(x,r) 
-    &= 2 e^{r^2/2} (1-\Phi(r)) \\
-    &= \frac{2}{r\sqrt{2\pi}} + O(r^{-3}) .
-\end{aligned}
-$$
-so that the difference in the left and right limits is
-$$
-\begin{aligned}
-    \lim_{x \to 0-} p(x,r) 
-    -\lim_{x \to 0+} p(x,r) 
-    &= 
-    \frac{1}{2}+\frac{2}{r\sqrt{2\pi}} 
-    - \frac{1}{2}-\frac{2}{r\sqrt{2\pi}} 
-      +O(r^{-3}) \\
-    &=
-    \frac{4}{r\sqrt{2\pi}}
-      +O(r^{-3}) 
-\end{aligned}
-$$
-
-Let's also look at the slope of $p$ at $x=0$.
-Since $\partial_x |x| = \sgn(x)$,
-$$
-\begin{aligned}
-    \partial_x \zeta(x,r)
-    &=
-    2 \sgn(x) \left(
-    \phi(|x|)
-    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
-    - e^{r|x|+r^2/2} \phi(|x|+r)
-      \right)
-\end{aligned}
-$$
-and so
-$$
-\begin{aligned}
-    -\partial_x p(x)
-    &=
-    \phi(|x|)
-    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
-    - e^{r|x|+r^2/2} \phi(|x|+r) \\
-    &\to
-    \frac{1}{\sqrt{2\pi}}(1-e^{r^2/2}e^{-r^2/2})
-    + r e^{r^2/2} (1-\Phi(r))  \qquad \text{as } x \to 0 \\
-    &=
-    r e^{r^2/2} (1-\Phi(r)) \\
-    &=
-    \frac{1}{\sqrt{2\pi}} + O(\frac{1}{r}) .
-\end{aligned}
-$$
-
-Also note that by the asymptotics for $1-\Phi(x)$,
-$$
-\begin{aligned}
-    \zeta(x,r) \approx 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) \qquad \text{for large }|x| .
-\end{aligned}
-$$
-
-What this says about clines
-===========================
+Other values of time
+--------------------
 
@@
 
-**Cline flattening:**
-immediately, this tells us how a cline gets wider as time progresses:
+Solving the PDE
+---------------
+
+Let's check this agrees with the solution for the PDE.
+According to Knight 1981 (Thm 7.4.3), $\zeta'(0) = - 2 r \zeta(0)$;
+ReacTran lets you specify the 'flux' at the boundary is proportional to $\zeta(0)$,
+where the flux is in our case $\sigma^2/2 \zeta'$.
+First note that the solution depends on where we put the boundaries:
+```{r reflected_pde_boundaries}
+solve.reflected <- function (x.up=-20) {
+    x.eps <- 5*eps
+    thegrid <- setup.grid.1D(x.up=x.up, x.down=0, N = 5000)
+    hetfn <- setup.prop.1D(grid=thegrid, func=function(x){abs(x)<x.eps})
+    yinit <- rep(1.0,thegrid$N)
+    tau <- 40
+    tt <- seq(0,tau,length.out=20)
+    r <- 1
+    pde.fn <- function (t,y,parms,...) {
+        # flux = zeta'(0)/2 = a.bl.down*( zeta(0) - C.down )
+        tran <- tran.1D(C=y, A=1, D=1/2, dx=thegrid, C.down=0, a.bl.down=4*r)$dC
+        list( tran - r * hetfn$mid * y )
+    }
+    soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
+    attr(soln,"x") <- thegrid$x.mid
+    return(soln)
+}
+pde.soln.0 <- solve.reflected(x.up=-5)
+pde.soln.1 <- solve.reflected(x.up=-8)
+pde.soln.2 <- solve.reflected(x.up=-20)
+clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),times=seq_along(tt),xlim=c(-5,0))
+clineplot(pde.soln.1,x=attr(pde.soln.1,"x"),add=TRUE)
+clineplot(pde.soln.2,x=attr(pde.soln.2,"x"),add=TRUE)
+```
+
+And, the PDE solution doesn't agree with the simulations very well: 
+clearly, we need a different boundary condition at zero.
+This is broken in ReacTran; I've modified the function to fix this.
+```{r reflected_pde,cache=TRUE}
+source("tran.1D.R")
+x.eps <- 5*eps
+thegrid <- setup.grid.1D(x.up=-20, x.down=0, N = 5000)
+hetfn <- setup.prop.1D(grid=thegrid, func=function(x){abs(x)<x.eps})
+yinit <- rep(1.0,thegrid$N)
+tau <- 40
+tt <- seq(0,tau,length.out=20)
+r <- 1
+pde.fn <- function (t,y,parms,...) {
+    # flux = zeta'(0)/2 = a.bl.down*( zeta(0) - C.down )
+    tran <- tran.1D(C=y, A=1, D=1/2, dx=thegrid, C.down=0, a.bl.down=2*r )$dC
+    list( tran - r * hetfn$mid * y )
+}
+pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
+ps <- clineplot(pde.soln,times=seq_along(tt),xlim=c(-8,0),legend=FALSE)
+matlines( thegrid$x.mid, outer( thegrid$x.mid, tt, function(x,t) { zeta.reflected(x/sqrt(t),r*t*x.eps) }), lty=3, col=rainbow(length(tt)) )
+legend("bottomleft",lty=1:2,legend=c("PDE","theory"))
+```
+Still not right...
+
+The associated cline shape
+--------------------------
+
+The Brownian rescaling above tells us how a cline gets wider as time progresses:
 for a fixed $r$, the width grows as $\sqrt{\tau}$.
@@
 
+Asymptotics
+-----------
+
+This is very steep at $x=0$.
+Notice that as $x$ goes to 0,
+$$
+\begin{aligned}
+    \lim_{x \to 0} \zeta(x,r) 
+    &= 2 e^{r^2/2} (1-\Phi(r)) \\
+    &= \frac{2}{r\sqrt{2\pi}} + O(r^{-3}) .
+\end{aligned}
+$$
+so that the difference in the left and right limits is
+$$
+\begin{aligned}
+    \lim_{x \to 0-} q(x,r) 
+    -\lim_{x \to 0+} q(x,r) 
+    &= 
+    \frac{1}{2}+\frac{2}{r\sqrt{2\pi}} 
+    - \frac{1}{2}-\frac{2}{r\sqrt{2\pi}} 
+      +O(r^{-3}) \\
+    &=
+    \frac{4}{r\sqrt{2\pi}}
+      +O(r^{-3}) 
+\end{aligned}
+$$
+
+Let's also look at the slope of $q$ at $x=0$.
+Since $\partial_x |x| = \sgn(x)$,
+$$
+\begin{aligned}
+    \partial_x \zeta(x,r)
+    &=
+    2 \sgn(x) \left(
+    \phi(|x|)
+    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
+    - e^{r|x|+r^2/2} \phi(|x|+r)
+      \right)
+\end{aligned}
+$$
+and so
+$$
+\begin{aligned}
+    -\partial_x q(x)
+    &=
+    \phi(|x|)
+    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
+    - e^{r|x|+r^2/2} \phi(|x|+r) \\
+    &\to
+    \frac{1}{\sqrt{2\pi}}(1-e^{r^2/2}e^{-r^2/2})
+    + r e^{r^2/2} (1-\Phi(r))  \qquad \text{as } x \to 0 \\
+    &=
+    r e^{r^2/2} (1-\Phi(r)) \\
+    &=
+    \frac{1}{\sqrt{2\pi}} + O(\frac{1}{r}) .
+\end{aligned}
+$$
+
+Also note that by the asymptotics for $1-\Phi(x)$,
+$$
+\begin{aligned}
+    \zeta(x,r) \approx 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) \qquad \text{for large }|x| .
+\end{aligned}
+$$
+
 Brownian motion with drift above zero
@@
 if $dX_t = -\theta(X_t) dt + dB_t$,
-where $\theta(x)=\theta$ for $x \ge 0$, then
-we want to know
+where $\theta(x)=\theta$ for $x \ge 0$, 
+and we define
 $$\begin{aligned}
-    u(x;\lambda,\theta):=\E^x \left[ \exp\left( - \lambda \int_0^\infty X_t dt \right) \right] .
+\Gamma_t = \text{meas}\{ t : X_t > 0 \} ,
 \end{aligned}$$
-(*Small problem:*  this isn't guarenteed to be on the correct size of zero when we want it to.)
+then we want to know
+$$\begin{aligned}
+    u(x;\lambda,\theta):=\E^x \left[ \exp\left( - \lambda \Gamma_t \right) \right] .
+\end{aligned}$$
+(*Small problem:*  $X$ isn't guarenteed to be on the correct side of zero when we want it to be.)
 Well, by Girsanov's theorem, we know that
-$\E[f(X_t)] = \E[f(B_t)\zeta(t)]$
+$\E[f(X_t)] = \E[f(B_t)Z(t)]$
 where
 $$\begin{aligned}
-\zeta(t) = \exp\left[ \int_0^t \theta(B_u) dB_t - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+Z(t) = \exp\left[ \int_0^t \theta(B_u) dB_t - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
 \end{aligned}$$
@@
 $$\begin{aligned}
-\theta \max(0,B_t) - \max(0,B_0) = \int_0^t \theta(B_u) du + \theta \ell_t(0) ,
+\theta \left(\max(0,B_t) - \max(0,B_0)\right) 
+&= 
+\int_0^t \theta(B_u) du + \theta \ell_t(0) \\
+&= 
+\theta \Gamma_t + \theta \ell_t(0) ,
 \end{aligned}$$
@@
 $$\begin{aligned}
-\zeta(t) = \exp\left[ \theta \max(0,B_t) - \theta \max(0,B_0) - \theta \ell_t(0) - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+Z(t) = \exp\left[ \theta \max(0,B_t) - \theta \max(0,B_0) - \theta \ell_t(0) - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
+\end{aligned}$$
+and we want to know
+$$\begin{aligned}
+u(x;\lambda,\theta)
+&=
+\E^x \left[
+\exp\left( 
+- \lambda \Gamma_t
+\right)
+\right]
 \end{aligned}$$

### 2015-10-22 notes/tran.1D.R

@@ -1,0 +1,285 @@
+tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL, 
+    flux.down = NULL, a.bl.up = NULL, a.bl.down = NULL, D = 0, 
+    v = 0, AFDW = 1, VF = 1, A = 1, dx, full.check = FALSE, full.output = FALSE) 
+{
+    if (is.null(dx)) 
+        stop("error: dx should be inputted ")
+    N <- length(C)
+    if (N == 0) 
+        stop("C should be a vector with numeric values")
+    if (is.null(C.up)) 
+        C.up <- C[1]
+    if (is.null(C.down)) 
+        C.down <- C[N]
+    if (!is.list(AFDW)) 
+        AFDW <- list(int = rep_len(AFDW,N+1))
+    if (!is.list(D)) 
+        D <- list(int = rep_len(D,N+1))
+    if (!is.list(v)) 
+        v <- list(int = rep_len(v,N+1))
+    if (!is.list(VF)) 
+        VF <- list(int = rep(VF, length.out = (N + 1)), mid = 0.5 * 
+            (rep(VF, length.out = (N + 1))[1:N] + rep(VF, length.out = (N + 
+                1))[2:(N + 1)]))
+    if (!is.list(A)) 
+        A <- list(int = rep(A, length.out = (N + 1)), mid = 0.5 * 
+            (rep(A, length.out = (length(C) + 1))[1:N] + rep(A, 
+                length.out = (N + 1))[2:(N + 1)]))
+    if (is.list(dx)) 
+        grid <- dx
+    if (!is.list(dx)) 
+        grid <- list(dx = rep(dx, length.out = N), dx.aux = 0.5 * 
+            (c(0, rep(dx, length.out = N)) + c(rep(dx, length.out = N), 
+                0)))
+    if (full.check) {
+        gn <- names(AFDW)
+        if (!"int" %in% gn) 
+            stop("error: AFDW should be a list that contains 'int', the AFDW values at the interface of the grid cell ")
+        if (is.null(AFDW$int)) 
+            stop("error: AFDW is NULL, should contain (numeric) values")
+        if (!is.null(AFDW$int)) {
+            if (!((length(AFDW$int) == 1) || (length(AFDW$int) == 
+                (N + 1)))) 
+                stop("error: AFDW should be a vector of length 1 or N+1")
+        }
+        if (any(AFDW$int < 0) || any(AFDW$int > 1)) 
[…]

### 2015-10-23 sims/run-sim-Alisa.Rmd
peter fiddling

@@
 knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
-source("sim-fns.R")
 ```
@@
 
-Now to define selection. Here we selection against heterozygotes to be strong (0.5).
+Now to define selection. 
 ```{r set_selection}
@@
 ```
+Here we selection against heterozygotes to be strong (`r S`).
 `qtl` is a named list, one element per chromosome,
@@
 This is a two-step process. First we simulate genotypes:
-```{r run_sim_step1}
+```{r run_sim_step1,cache=TRUE,depends="set_selection"}
 	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
@@
 ```
-The output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
-	where:
-		"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
-		"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
-		"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
-		"pars" is an integer that gives the number of generatsion of the simulation
-		"QTL" returns the QTL vector used in the simulation
+The output is: a list with hierarchy: `[[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]` where:
+
+	- `inds` is a list of length `ninds`, such that each `[[ind]][[chr]][[X1/X2]]` is a data.frame with columns
+        * `pos` : coordinates of rec.,  (*what does this mean?*)
+        * `ancest` : ancestor ID
+	- `sp.ids` is a character vector of length `ninds`, giving the ancestry given an ancestor ID
+	- `deme` is an integer vector of length `ninds`, giving the deme number of an ancestor, given their ancestor ID
+	- `pars` is an integer that gives the number of generations of the simulation
+	- `QTL` is the QTL vector used in the simulation
 
 The second step is to parse the simulated genotypes:
-```{r run_sim_step2}
+```{r run_sim_step2,cache=TRUE,depends="run_sim_step1"}
 	sims.sums = lapply(sims.s0.1,spBreaks)
 ```
-The output here is a list: [[zone_age]][[c("ind.ancestry","ancest.prop")]]
-		"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
-		"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
+The output here is a list: `[[zone_age]][[c("ind.ancestry","ancest.prop")]]`
+
+    * `ind.ancestry` is a list of length ninds: each `[[ind][[chr]]][[X1/X2]]` is a data.frame with columns
+        * `starts` = (chunk) starts, 
+        * `stops` = (chunk) stops, 
+        * `sp2` = sp2(T/F) (*what is this?*)
+    * `ancest.prop` is a $2 \times$`ninds` matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
 
 This resulting files can get quite large, but we've saved them here
-```{r save_file,cache=TRUE}
+```{r save_file,cache=TRUE,depends="run_sim_step1"}
 	save(sims.s0.1,file="sims_sums_s0.5_sigma1_50by500_tau50.Robj")
@@
 		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
-
 ```
-(`genotypes` is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
+(`genotypes` is a list of `nloci`$\times$`ninds` dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
 Each line in the plot represents a locus a given distance from the selected locus.
@@
     }
-
     pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
@@
 
-		#function to increase transparency of colors
-		add.alpha <- function(col, alpha=1){
-		if(missing(col))
-		stop("Please provide a vector of colours.")
-		apply(sapply(col, col2rgb)/255, 2, 
-		function(x) 
-		rgb(x[1], x[2], x[3], alpha=alpha)) 
-		}
+	transparent_rainbow = adjustcolor(rainbow(ndemes),0.75)
 
-	transparent_rainbow = add.alpha(rainbow(ndemes),0.75)
-
-		D = 1; hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")
-
-		for(D in seq(1,ndemes,5)){
+    D = 1 
+    hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")
+    for (D in seq(1,ndemes,5)) {
         hist(tracts[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
@@
 
-```
+```{r ld_matrix}
 LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:ndemes,get.LD,POS1=0.5,POS2=Z,CHR=1,IND_DATA=sims.sums[[1]]$ind.ancest)}))

### 2015-10-23 notes/diffusion-local-times.Rmd
pde clines

@@
 ```{r setup_fns}
-pcline <- function (x,r,recombfn,...) {
-    (1/2)*( 1 - sign(x)*recombfn(x,2*r,...) )
-}
-tanh_cline <- function (x,s,sigma=1) {
-    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
-}
-tanh_drift <- function (x,s,sigma=1) {
-    # log-derivative of tanh_cline
-    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
-}
-tanh_het <- function (x,s,sigma=1) {
-    # prob of being in a het = (1-s)(1-p)/(1-s(1-p))
-    pp <- tanh_cline(x,s,sigma=sigma)
-    (1-s)*(1-pp)/(1-s*(1-pp))
-}
-zeta.reflected <- function (x,r) {
-    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
-}
-sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
-    # want total time to be tau, so:
-    dt <- tau/ntimes
-    # put the noise in rws
-    rws <- matrix( c( rep_len(xinit,nwalks), # initial conditions
-                 rnorm(nwalks*(ntimes-1)) ), nrow=ntimes, ncol=nwalks, byrow=TRUE )
-    # and then integrate over it
-    for (k in 2:ntimes) {
-        rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],...)
-    }
-    attr(rws,"dt") <- dt
-    return(rws)
-}
-weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
-    colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
-}
-clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),legend=TRUE,...) {
-    # plot output of deSolve as clines in a nice way
-    graphics::matplot( x, t(soln[times,-1]), type='l', col=rainbow(length(times)), lty=1, 
-            xlab="space", ylab=ylab, ylim=ylim, ...)
-    if (legend) legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
-    return( invisible( t(soln[times,-1]) ) )
-}
+source("cline-fns.R")
 ```
@@
 and if this happens, the lineage ends up on either background with probability $1/2$.
-Therefore,
+At this point, its probability of originally coming from one or the other background
+still depends on its location,
+but if we **assume this is 1/2**,
 for an individual to the left of the barrier ($x<0$),
@@
 Combining these,
-the probability of ending up on the left-hand side is
+the probability of ending up on the left-hand side is *approximated* as
 $$
@@
 yinit <- rep(1.0,thegrid$N)
-cline.interp <- function (T) {
-    # return the interpolated cline for a particular time
-    which.t <- findInterval(T,fwds.soln[,1],rightmost.closed=TRUE)
-    lu <- fwds.soln[c(which.t,which.t+1),1]
-    alpha <- (T-lu[1])/diff(lu)
-    return( (1-alpha)*fwds.soln[which.t,-1] + alpha*fwds.soln[which.t+1,-1]  )
-}
 rev.pde <- function (t,y,parms,...) {
-    p <- cline.interp(tau-t)
+    p <- cline.interp(tau-t,soln=fwds.soln)
     # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline.interp(tau-t)), interp="linear")
@@
 
-Killing at Brownian local times
-===============================
+Reflected Brownian motion, killed according to local time at zero
+=================================================================
 
@@
 Its Laplace transform
-=====================
+---------------------
 
@@
 ```{r reflected_pde_boundaries}
+source("tran.1D.R")
 solve.reflected <- function (x.up=-20) {
@@
 ```
-Still not right...
+Still not right... hm.
 
@@
 ```{r show_cline, cache=TRUE}
-zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1)
+zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) }
 xx <- seq(-15,15,length.out=500)
@@
 In other words, 
-if $dX_t = -\theta(X_t) dt + dB_t$,
-where $\theta(x)=\theta$ for $x \ge 0$, 
-and we define
+$$\begin{aligned}
+dX_t = -\theta(X_t) dt + dB_t,
+\end{aligned}$$
+where $\theta(x)=\theta$ for $x \ge 0$, and $\theta(x)=0$ for $x<0$.
+
+Solving the PDE
+---------------
+
+Above we simulated the clines deriving from this process.
+```{r sim_pushed, cache=TRUE}
+rvals <- exp(seq(log(.001),log(0.5),length.out=10))
+pushfn <- function(x,s){ifelse(x>0,-4*sqrt(s),0)}
+nwalks <- 1e4
+ntimes <- 1e3
+s <- 0.1
+tau <- 10
+xinit <- seq(-8,8,length.out=nwalks)
+sim.rw <- sim_diffusion( pushfn, tau, nwalks, ntimes, s=s, xinit=xinit )
+sim.wt <- weighted_time( sim.rw, function (x) { x>0 } )
+# distribution of time above zero depending on starting point
+plot(xinit,sim.wt,xlab="initial position", ylab="time above zero")
+```
+We can also get the predicted clines by solving the PDE.
+Note that here we are approximating $p(x) = \min(1,\exp(-4x\sqrt{s}))$.
+```{r pushed_pde,cached=TRUE}
+thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
+clinefn <- setup.prop.1D(grid=thegrid, func=function(x,s){ifelse(x>0,exp(-4*x*sqrt(s)),1.0)}, s=s)
+hetfn <- setup.prop.1D(grid=thegrid, func=function(x){ifelse(x>0,1,0)})
+yinit <- rep(1.0,thegrid$N)
+tt <- seq(0,tau,length.out=100)
+pde.solns <- lapply( rvals, function (r) {
+        pde.fn <- function (t,y,parms,...) {
+            tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+            if (any(is.na(tran))) { browser() }
+            list( tran - r * hetfn$mid * y )
+        }
+        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
+      } )
+layout(t(1:2))
+for (k in c(6,10)) {
+    clineplot(pde.solns[[k]],main=sprintf("r=%0.2f",tau,rvals[k]))
+}
+```
+
+Here are the two compared,
+by smoothing the probabilities for the simulation:
+```{r sim_clines, cache=TRUE, depends=c("sim_pushed","pushed_pde")}
+prob.no.recomb <- sapply( rvals, function (r) { fitted(loess( exp(-r*sim.wt) ~ xinit, span=0.1)) } )
+matplot(xinit,prob.no.recomb,type='l',ylab="prob of no recomb",
+        xlab="space",col=rainbow(length(rvals)),lty=1,main=sprintf("t=%d",tau))
+legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals,digits=2)))
+for (k in seq_along(rvals)) {
+    lines( thegrid$x.mid, pde.solns[[k]][findInterval(tau,pde.solns[[k]][,1]),-1], 
+          lty=3, lwd=3, col=rainbow(length(rvals))[k] )
+}
+```
+Well, that looks reasonable, but hard to tell if the discrepancy is due to the loess or not.
+
+And, finally, here's some actual clines predicted by this model:
+```{r pushed_cline,cache=TRUE,depends="pushed_pde"}
+tau <- 100
+tt <- seq(0,tau,length.out=100)
+pde.solns <- lapply( rvals, function (r) {
+        pde.fn <- function (t,y,parms,...) {
+            tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+            if (any(is.na(tran))) { browser() }
+            list( tran - r * hetfn$mid * y )
+        }
+        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
+      } )
+plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
+layout(t(1:2))
+for (tval in plot.tvals) {
+    matplot( thegrid$x.mid, 
+            pcline(thegrid$x.mid, p=sapply( pde.solns, cline.interp, T=tval ) ), 
+                   type='l', main=sprintf("t=%d",tval) 
+        )
+    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
+}
+```
+
+Analytical form
+---------------
+
+If we define
 $$\begin{aligned}

### 2015-10-23 notes/pde-clines.Rmd
pde clines

---
title: "Clines from PDE"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\grad}{\nabla}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
```

*Useful functions used below:*
```{r setup_fns}
source("cline-fns.R")
```

The goal
========

Suppose that a lineage moves in a frequency profile $p(x)$ of allele $A$,
switching between backgrounds at rate $r p(x)$ or $r (1-p(x))$.
Let $(X_t,Z_t)$ denote this process,
so that
$$\begin{aligned}
    dX_t &= \grad \log(p(X_t)) dt + dB_t  \qquad \text{if } Z_t = A \\
         &= \grad \log(1-p(X_t)) dt + dB_t  \qquad \text{if } Z_t = B \\
    \P\{ Z_{t+\epsilon} = B | Z_t = A \} &= \epsilon r (1-p(X_t)) + O(\epsilon^2) \\
    \P\{ Z_{t+\epsilon} = A | Z_t = B \} &= \epsilon r p(X_t) + O(\epsilon^2) .
\end{aligned}$$
(Note that $p$ may also depend on $t$.)
We would like to find the probability that 
a lineage which we sample at location $x$ on background $z$,
derived from background $A$ $\tau$ units of time ago,
which we denote by
$$
q(x,z,\tau,r) = \P\{ Z_\tau = A \} .
$$
(Note that this gives us linkage to the selected site,
and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau) q(x,B,\tau,r)$,
but it is easier to compute this.)

Let $f$ be a function from $\R \times \{A,B\} \to \R$.
The process above has generator
$$\begin{aligned}
    Gf(x,z) 
    &:= \partial_t \E^{(x,z)}\left[ f(X_t,Z_t) \right] \vert_{t=0} \\
    &= \begin{cases}
    \grad \log(p(x)) \cdot \grad f(x,A) + \frac{\sigma^2}{2} \Delta f(x,A) + r (1-p(x)) (f(x,B)-f(x,A)) \qquad &\text{if } z=A \\
    \grad \log(1-p(x)) \cdot \grad f(x,B) + \frac{\sigma^2}{2} \Delta f(x,B) + r p(x) (f(x,A)-f(x,B)) \qquad &\text{if } z=B .
    \end{cases}
\end{aligned}$$

We know that 
$$\begin{aligned}
q(x,A,0,r) = 1  \\
q(x,B,0,r) = 0  \\
\end{aligned}$$
and that $\partial_t q = G q$, i.e.,
(writing $q(x,A,t,r) = q_A(x,t)$ for the moment):
$$\begin{aligned}
    \partial_t q_A(x,t)
        &= 
        \grad \log(p(x)) \cdot \grad q_A(x,t) + \frac{\sigma^2}{2} \Delta q_A(x,t) + r (1-p(x)) (q_B(x,t)-q_A(x,t)) \\
    \partial_t q_B(x,t)
        &= 
        \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) \\
\end{aligned}$$

Let's solve this numerically.
```{r solve_pde_setup, cache=TRUE}
s <- 0.1
tau <- 100
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
cline.A <- setup.prop.1D(grid=thegrid, func=tanh_cline, s=s)
cline.B <- setup.prop.1D(grid=thegrid, func=function(x,s)1-tanh_cline(x,s), s=s)
yinit <- c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) )
tt <- seq(0,tau,length.out=100)
r <- 1
pde.fn <- function (t,y,parms,...) {
    yA <- y[1:thegrid$N]
    yB <- y[thegrid$N+(1:thegrid$N)]
    tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=thegrid)$dC
    tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=thegrid)$dC
    list( c( 
            tran.A + r * cline.B$mid * (yB-yA),
            tran.B + r * cline.A$mid * (yA-yB)
            ) )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
A.inds <- 1+(1:thegrid$N)
B.inds <- 1+thegrid$N+(1:thegrid$N)
pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*cline.A$mid[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*cline.B$mid[col(pde.soln[,B.inds])] )
attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
attr(pde.cline,"nspec")<-1
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```

### 2015-10-23 notes/cline-fns.R

@@ -1,0 +1,50 @@
+pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
+    (1/2)*( 1 - sign(x)*p )
+}
+tanh_cline <- function (x,s,sigma=1) {
+    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+}
+tanh_drift <- function (x,s,sigma=1) {
+    # log-derivative of tanh_cline
+    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
+}
+tanh_het <- function (x,s,sigma=1) {
+    # prob of being in a het = (1-s)(1-p)/(1-s(1-p))
+    pp <- tanh_cline(x,s,sigma=sigma)
+    (1-s)*(1-pp)/(1-s*(1-pp))
+}
+zeta.reflected <- function (x,r) {
+    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
+}
+sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
+    # want total time to be tau, so:
+    dt <- tau/ntimes
+    # put the noise in rws
+    rws <- matrix( c( rep_len(xinit,nwalks), # initial conditions
+                 rnorm(nwalks*(ntimes-1)) ), nrow=ntimes, ncol=nwalks, byrow=TRUE )
+    # and then integrate over it
+    for (k in 2:ntimes) {
+        rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],...)
+    }
+    attr(rws,"dt") <- dt
+    return(rws)
+}
+weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
+    colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
+}
+clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),legend=TRUE,...) {
+    # plot output of deSolve as clines in a nice way
+    graphics::matplot( x, t(soln[times,-1]), type='l', col=rainbow(length(times)), lty=1, 
+            xlab="space", ylab=ylab, ylim=ylim, ...)
[…]

### 2015-10-23 notes/pde-clines.Rmd
changed r

@@
 tt <- seq(0,tau,length.out=100)
-r <- 1
+r <- 0.1
 pde.fn <- function (t,y,parms,...) {

### 2015-10-23 notes/pde-clines.Rmd
.

--- 2015-10-23 notes/pde-clines.Rmd
@@
         &= 
-        \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) \\
+        \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) .
 \end{aligned}$$
 
-Let's solve this numerically.
+Paving the way for analytic approximations
+------------------------------------------
+
+Note that $p$ appears twice here:
+first pushing lineages, and second causing recombination.
+We may want to decouple these for the purposes of approximation:
+we'll consider the following system of PDE:
+$$\begin{aligned}
+    \partial_t q_A(x,t)
+        &= 
+        \grad \log(u(x)) \cdot \grad q_A(x,t) + \frac{\sigma^2}{2} \Delta q_A(x,t) + r (1-v(x)) (q_B(x,t)-q_A(x,t)) \\
+    \partial_t q_B(x,t)
+        &= 
+        \grad \log(1-u(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r v(x) (q_A(x,t)-q_B(x,t)) .
+\end{aligned}$$
+
+Here's a function to solve this numerically:
 ```{r solve_pde_setup, cache=TRUE}
-s <- 0.1
-tau <- 100
 thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
-cline.A <- setup.prop.1D(grid=thegrid, func=tanh_cline, s=s)
-cline.B <- setup.prop.1D(grid=thegrid, func=function(x,s)1-tanh_cline(x,s), s=s)
+solve.pde <- function ( u, v=u, r, times, ... ) {
+    vfn <- function(x,t){v(x=x,t=t,...)}
+    cline.A <- setup.prop.1D(grid=thegrid, func=u, ...)
+    cline.B <- setup.prop.1D(grid=thegrid, func=function(x,...)1-u(x,...), ...)
     yinit <- c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) )
-tt <- seq(0,tau,length.out=100)
-r <- 1
+    r <- 0.1
     pde.fn <- function (t,y,parms,...) {
@@
         list( c( 
-            tran.A + r * cline.B$mid * (yB-yA),
-            tran.B + r * cline.A$mid * (yA-yB)
+                tran.A + r * (1-vfn(x=thegrid$x.mid,t=t)) * (yB-yA),
+                tran.B + r * vfn(x=thegrid$x.mid,t=t) * (yA-yB)
                 ) )
     }
-pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
+    ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
+}
+```
+
+Numerical solutions
+-------------------
+
+Let's solve this numerically.
+```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
+s <- 0.1
+tau <- 100
+tt <- seq(0,tau,length.out=100)
+r <- 0.1
+pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s )
 image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))

### 2015-10-23 notes/cline-fns.R

@@ -1,7 +1,7 @@
 pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
     (1/2)*( 1 - sign(x)*p )
 }
-tanh_cline <- function (x,s,sigma=1) {
+tanh_cline <- function (x,s,sigma=1,...) {
     (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
 }
 tanh_drift <- function (x,s,sigma=1) {

### 2015-10-23 notes/diffusion-local-times.Rmd
minor fixes

@@
 library(ReacTran)
-library(deSolve)
 ```
@@
 solves the PDE (Barton 1979):
-$$\begin{aligned*}
+$$\begin{aligned}
   \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
-\end{aligned*}$$
+\end{aligned}$$
 ```{r fwds_pde, cache=TRUE}
@@
 First note that the solution depends on where we put the boundaries:
-```{r reflected_pde_boundaries}
-source("tran.1D.R")
+```{r reflected_pde_boundaries,cache=TRUE}
+# NOTE: must install ReacTran from R-forge, e.g. 
+#   install.packages("ReacTran",repos="http://r-forge.r-project.org")
+# or else
+#   source("tran.1D.R")
+# to fix bug in using convective boundary conditions.
+x.eps <- 5*mean(thegrid$dx)
 solve.reflected <- function (x.up=-20) {
-    x.eps <- 5*eps
     thegrid <- setup.grid.1D(x.up=x.up, x.down=0, N = 5000)
@@
 pde.soln.2 <- solve.reflected(x.up=-20)
-clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),times=seq_along(tt),xlim=c(-5,0))
+clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),xlim=c(-5,0))
 clineplot(pde.soln.1,x=attr(pde.soln.1,"x"),add=TRUE)
@@
 ```{r reflected_pde,cache=TRUE}
-source("tran.1D.R")
-x.eps <- 5*eps
 thegrid <- setup.grid.1D(x.up=-20, x.down=0, N = 5000)

### 2015-10-23 notes/pde-clines.Rmd
minor fixes

@@
         &= 
-        \grad \log(1-u(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r v(x) (q_A(x,t)-q_B(x,t)) .
+        \grad \log(1-u(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r v(x) (q_A(x,t)-q_B(x,t)) \\
+    q_A(x,0) &= 1 \\
+    q_A(x,t) &= 1-q_B(-x,t) .
 \end{aligned}$$
@@
 B.inds <- 1+thegrid$N+(1:thegrid$N)
-pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*cline.A$mid[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*cline.B$mid[col(pde.soln[,B.inds])] )
+sel.cline <- tanh_cline(thegrid$x.mid,s=s)
+pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
 attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )

### 2015-10-23 notes/pde-clines.Rmd
began comparing to sims

@@
 s <- 0.1
-tau <- 100
-tt <- seq(0,tau,length.out=100)
+zone.age <- 100  # this is tau
+tt <- seq(0,zone.age,length.out=100)
 r <- 0.1
@@
 ```
+
+Comparison to simulation
+------------------------
+
+We'll setup, run, and summarize simulations
+as described in [this file](../sims/run-sims.html).
+```{r setup_sims, cache=TRUE}
+source("../sims/sim-fns.R")
+zone.age	= 50
+ndemes 		= 500
+deme.size 	= 10
+ninds 		= ndemes*deme.size
+SIGMA 		= 1
+s           = 0.1
+qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
+# spatial locations
+xx <- (1:ndemes)-0.5-ndemes/2
+# run the simulation
+sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
+# post-process
+sim.sums <- spBreaks(sim)
+```
+and, postprocess:
+```{r genotype_sims, cache=TRUE, depends="setup_sims"}
+rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
+loci <- list(0.5+rvals)
+genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
+freqs <- apply(genotypes,1,function(Z){
+            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
+        })
+```
+
+OK, now let's match the PDE solution to the simulation.
+First, does the cline at the selected locus match the the tanh cline?
+```{r match_cline}
+plot(xx,freqs[,1],type='l',xlab="spatial location",ylab="proportion")
+lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red')
+```
+Looks pretty reasonable.
+
+```{r match_pde}
+tt <- seq(0,zone.age,length.out=100)
+pde.solns <- lapply(rvals, function (r) { solve.pde( u=tanh_cline, r=r, times=tt, s=s ) })
+```

### 2015-10-23 sims/run-sim-Alisa.Rmd
began comparing to sims

@@
 ```
-note that zone_age above can be a vector of times.
+note that `zone_age` above can be a vector of times.
 

### 2015-10-24 notes/pde-clines.Rmd
modified simulation function to work with small deme sizes

--- 2015-10-23 notes/pde-clines.Rmd
@@
 thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
-solve.pde <- function ( u, v=u, r, times, ... ) {
+solve.pde <- function ( u, v=u, r, times, grid, ... ) {
     vfn <- function(x,t){v(x=x,t=t,...)}
-    cline.A <- setup.prop.1D(grid=thegrid, func=u, ...)
-    cline.B <- setup.prop.1D(grid=thegrid, func=function(x,...)1-u(x,...), ...)
-    yinit <- c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) )
+    cline.A <- setup.prop.1D(grid=grid, func=u, ...)
+    cline.B <- setup.prop.1D(grid=grid, func=function(x,...)1-u(x,...), ...)
+    yinit <- c( rep(1.0,grid$N), rep(0.0,grid$N) )
     r <- 0.1
     pde.fn <- function (t,y,parms,...) {
-        yA <- y[1:thegrid$N]
-        yB <- y[thegrid$N+(1:thegrid$N)]
-        tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=thegrid)$dC
-        tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=thegrid)$dC
+        yA <- y[1:grid$N]
+        yB <- y[grid$N+(1:grid$N)]
+        tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=grid)$dC
+        tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=grid)$dC
         list( c( 
-                tran.A + r * (1-vfn(x=thegrid$x.mid,t=t)) * (yB-yA),
-                tran.B + r * vfn(x=thegrid$x.mid,t=t) * (yA-yB)
+                tran.A + r * (1-vfn(x=grid$x.mid,t=t)) * (yB-yA),
+                tran.B + r * vfn(x=grid$x.mid,t=t) * (yA-yB)
                 ) )
     }
-    ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
+    soln <- ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
+    attr(soln,"u") <- u
+    attr(soln,"v") <- v
+    return(soln)
+}
+cline.from.soln <- function (soln, grid, clinefn=attr(soln,"u"), ... ) {
+    A.inds <- 1+(1:grid$N)
+    B.inds <- 1+grid$N+(1:grid$N)
+    sel.cline <- clinefn(grid$x.mid,s=s)
+    pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
+    attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
+    attr(pde.cline,"nspec")<-1
+    return(pde.cline)
 }
@@
 s <- 0.1
-tau <- 100
-tt <- seq(0,tau,length.out=100)
+zone.age <- 100  # this is tau
+tt <- seq(0,zone.age,length.out=100)
 r <- 0.1
-pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s )
+pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s, grid=thegrid )
 image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
@@
 ```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
-A.inds <- 1+(1:thegrid$N)
-B.inds <- 1+thegrid$N+(1:thegrid$N)
-sel.cline <- tanh_cline(thegrid$x.mid,s=s)
-pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
-attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
-attr(pde.cline,"nspec")<-1
+pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
 image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
@@
 ```
+
+Comparison to simulation
+------------------------
+
+We'll setup, run, and summarize simulations
+as described in [this file](../sims/run-sims.html).
+```{r setup_sims, cache=TRUE}
+source("../sims/sim-fns.R")
+zone.age	= 50
+ndemes 		= 500
+deme.size 	= 10
+ninds 		= ndemes*deme.size
+SIGMA 		= 10
+s           = 0.1
+qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
+# spatial locations
+xx <- (1:ndemes)-0.5-ndemes/2
+# run the simulation
+sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
+# post-process
+sim.sums <- spBreaks(sim)
+```
+and, postprocess:
+```{r genotype_sims, cache=TRUE, depends="setup_sims"}
+rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
+loci <- list(0.5+rvals)
+genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
+freqs <- apply(genotypes,1,function(Z){
+            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
+        })
+```
+
+OK, now let's match the PDE solution to the simulation.
+First, does the cline at the selected locus match the the tanh cline?
+```{r match_cline}
+plot(xx,freqs[,1],type='l',xlab="spatial location",ylab="proportion")
+lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red')
+```
+Looks pretty reasonable.
+
+```{r match_pde}
+tt <- seq(0,zone.age,length.out=100)
+pde.solns <- lapply(rvals, function (r) { solve.pde( u=tanh_cline, r=r, times=tt, s=s ) })
+```

### 2015-10-24 sims/sim-fns.R

@@ -23,7 +23,7 @@ spBreaks = function(sim){
                     if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
                     sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
                     sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
-                    tmp = rbind(data.frame(sp1,sp2=rep(F,nrow(sp1))),data.frame(sp2,sp2=rep(T,nrow(sp2))))
+                    tmp = rbind(data.frame(sp1,sp2=rep(FALSE,nrow(sp1))),data.frame(sp2,sp2=rep(TRUE,nrow(sp2))))
                     tmp = tmp[order(tmp$starts),]
                     tmp$starts = PAR$pos[tmp$starts]
                     tmp$stops = c(PAR$pos[-1],1)[tmp$stops]
@@ -90,6 +90,8 @@ genoAll = function(inds,QTL,sp.ids){
 #######    MAKING A NEW GEN     #######
 
 makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),min.deme=min(demes)){
+    pop.sizes <- table(demes) # number of individuals in each deme
+    chrnames <- namesList( names(inds[[1]]) )
     #genotyping
     if(!is.null(QTL)){ 
         genos = lapply(genoAll(inds=inds,QTL=qtl,sp.ids=sp.ids) ,function(IND){do.call(rbind,IND)})  
@@ -116,16 +118,33 @@ makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),m
     # new locations of individuals
     inds.df$new.deme = pmin(max.deme, pmax(min.deme, inds.df$deme + sample_move ) )
     inds.df$id = seq_along(inds.df$new.deme)
-    new.inds.df = do.call(rbind,lapply(  unique(inds.df$deme)  , function(d){			
-        #find parents
-        a = with(inds.df[inds.df$new.deme==d,], sample(id,size=length(which(demes==d)), prob=w/sum(w) , replace=T))
-        # MODIFY B|A TO ALLOW FOR ASSORTATIVE MATING & eg no selfing
-        b = with(inds.df[inds.df$new.deme==d,], sample(id, size=length(which(demes==d)), prob=w/sum(w) , replace=T))
[…]

### 2015-10-25 notes/pde-clines.Rmd
allowed logA in tran.1D

@@
 ```{r solve_pde_setup, cache=TRUE}
+source("cline-fns.R")
 thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
-solve.pde <- function ( u, v=u, r, times, grid, ... ) {
-    vfn <- function(x,t){v(x=x,t=t,...)}
-    cline.A <- setup.prop.1D(grid=grid, func=u, ...)
-    cline.B <- setup.prop.1D(grid=grid, func=function(x,...)1-u(x,...), ...)
-    yinit <- c( rep(1.0,grid$N), rep(0.0,grid$N) )
-    r <- 0.1
-    pde.fn <- function (t,y,parms,...) {
-        yA <- y[1:grid$N]
-        yB <- y[grid$N+(1:grid$N)]
-        tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=grid)$dC
-        tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=grid)$dC
-        list( c( 
-                tran.A + r * (1-vfn(x=grid$x.mid,t=t)) * (yB-yA),
-                tran.B + r * vfn(x=grid$x.mid,t=t) * (yA-yB)
-                ) )
-    }
-    soln <- ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
-    attr(soln,"u") <- u
-    attr(soln,"v") <- v
-    return(soln)
-}
-cline.from.soln <- function (soln, grid, clinefn=attr(soln,"u"), ... ) {
-    A.inds <- 1+(1:grid$N)
-    B.inds <- 1+grid$N+(1:grid$N)
-    sel.cline <- clinefn(grid$x.mid,s=s)
-    pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
-    attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
-    attr(pde.cline,"nspec")<-1
-    return(pde.cline)
-}
+solve.pde
+cline.from.soln
 ```
@@
 source("../sims/sim-fns.R")
-zone.age	= 50
+zone.age	= 20
 ndemes 		= 500
-deme.size 	= 10
+deme.size 	= 5
 ninds 		= ndemes*deme.size
@@
 
+Here's the PDE solution:
 ```{r match_pde}
 tt <- seq(0,zone.age,length.out=100)
-pde.solns <- lapply(rvals, function (r) { solve.pde( u=tanh_cline, r=r, times=tt, s=s ) })
+pde.solns <- lapply(rvals, function (r) { solve.pde( u=function(...)1-tanh_cline(...), r=r, times=tt, s=s, grid=thegrid ) })
+pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
+match.clines <- sapply(pde.clines,function(x)x[length(tt),])
 ```
+Now, comparing the PDE solution to the others:
+```{r match_cline_pde}
+for (k in seq_along(rvals)) {
+    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
+    lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red',lwd=3)
+    lines(1-thegrid$x.mid*SIGMA,match.clines[,k],col='green',lwd=3)
+}
+```

### 2015-10-25 notes/reactran-notes.Rmd
allowed logA in tran.1D

---
title: "Using ReacTran"
date: "`r date()`"
---

`ReacTran::tran.1D` returns the reaction-diffusion part of an equation that has a term like
$$\begin{aligned}
    \frac{1}{A} \frac{d}{dx}  A f(x)  ,
\end{aligned}$$
where $f$ is the flux.
This can run into numerical difficulties if $A$ is very small (numerically zero).
In this case, it makes sense to solve, equivalently,
$$\begin{aligned}
    \frac{d}{dx} f(x) + f(x) \frac{d}{dx} \log A 
\end{aligned}$$
I have modified `tran.1D` to do this (in [tran.1D.R](tran.1D.R)),
allowing us to pass in $\log A$ instead of $A$.

Now we need a good way to compute the $\log$ of the cline function 
$$\begin{aligned}
    c(x) = \frac{1}{2}(1+\tanh(-2 x \sqrt(s)/\sigma)) .
\end{aligned}$$
Suppose that $x$ is  large and positive (where the problem is).
We can rewrite
$$\begin{aligned}
1+\tanh(-x) 
&= 1+\frac{e^{-x}-e^{x}}{e^{-x}+e^{x}} \\
&= \frac{2e^{-x}}{e^x+e^{-x}} \\
&= \frac{2e^{-x}}{e^x(1+e^{-2x})} 
&= \frac{2e^{-2x}}{1+e^{-2x}} 
\end{aligned}$$
and so
$$\begin{aligned}
\log(1+\tanh(x) )
&= \log(2) - 2x - \log(1+e^{-2x}) .
\end{aligned}$$

Here's a quick check:
```{r log_tanh}
tanh_cline <- function (x,s,sigma=1,log=FALSE) {
    if (log) {
        - (4*x*sqrt(s)/sigma) - log1p(exp(-4*x*sqrt(s)/sigma))
    } else {
        (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
    }
}
xx <- seq(-30,50,length.out=200)
# not log-scale
plot(xx,tanh_cline(xx,s=0.1),type='l')
lines(xx,exp(tanh_cline(xx,s=0.1,log=TRUE)),col='red',lty=3,lwd=2)
# log-scale
plot(xx,log(tanh_cline(xx,s=0.1)),type='l',ylim=range(tanh_cline(xx,s=0.1,log=TRUE)))
lines(xx,(tanh_cline(xx,s=0.1,log=TRUE)),col='red',lty=3,lwd=2)
```

### 2015-10-25 notes/cline-fns.R

@@ -1,21 +1,26 @@
 pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
     (1/2)*( 1 - sign(x)*p )
 }
+
 tanh_cline <- function (x,s,sigma=1,...) {
     (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
 }
+
 tanh_drift <- function (x,s,sigma=1) {
     # log-derivative of tanh_cline
     -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
 }
+
 tanh_het <- function (x,s,sigma=1) {
     # prob of being in a het = (1-s)(1-p)/(1-s(1-p))
     pp <- tanh_cline(x,s,sigma=sigma)
     (1-s)*(1-pp)/(1-s*(1-pp))
 }
+
 zeta.reflected <- function (x,r) {
     2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
 }
+
 sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
     # want total time to be tau, so:
     dt <- tau/ntimes
@@ -29,9 +34,11 @@ sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
     attr(rws,"dt") <- dt
     return(rws)
 }
+
 weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
     colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
 }
+
 clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),legend=TRUE,...) {
     # plot output of deSolve as clines in a nice way
     graphics::matplot( x, t(soln[times,-1]), type='l', col=rainbow(length(times)), lty=1, 
@@ -39,12 +46,49 @@ clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=nt
     if (legend) legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
     return( invisible( t(soln[times,-1]) ) )
 }
-cline.interp <- function (T,soln) {
+
[…]

### 2015-10-25 notes/tran.1D.R

@@ -1,6 +1,8 @@
 tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL, 
     flux.down = NULL, a.bl.up = NULL, a.bl.down = NULL, D = 0, 
-    v = 0, AFDW = 1, VF = 1, A = 1, dx, full.check = FALSE, full.output = FALSE) 
+    v = 0, AFDW = 1, VF = 1, A = 1, dx, 
+    log.A = FALSE,   # TRUE if A is given in log units
+    full.check = FALSE, full.output = FALSE) 
 {
     if (is.null(dx)) 
         stop("error: dx should be inputted ")
@@ -94,15 +96,14 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
         if (is.null(A$int) || is.null(A$mid)) 
             stop("error: the surface area A is NULL, should contain (numeric) values")
         if (!is.null(A$int)) {
-            if (!((length(A$int) == 1) || (length(A$int) == (N + 
-                1)))) 
+            if (!((length(A$int) == 1) || (length(A$int) == (N + 1)))) 
                 stop("error: A$int should be a vector of length 1 or N+1")
         }
         if (!is.null(A$mid)) {
             if (!((length(A$mid) == 1) || (length(A$mid) == (N)))) 
                 stop("error: A$mid should be a vector of length 1 or N")
         }
-        if (any(A$int < 0) || any(A$mid < 0)) 
+        if ( (any(A$int <= 0) || any(A$mid <= 0)) || (!log.A && ( any(A$int==0) || any(A$mid==0) )) )
             stop("error: the area A should always be positive")
         gn <- names(grid)
         if (!"dx" %in% gn) 
@@ -273,7 +274,13 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
         flux[1] <- flux.up
     if (!is.null(flux.down)) 
         flux[N + 1] <- flux.down
-    dC <- -diff(A$int * flux)/A$mid/VF$mid/grid$dx
+    
+    if (!log.A) {
+        dC <- -diff(A$int * flux)/A$mid/VF$mid/grid$dx
+    } else {
+        dC <- -( diff(A$int)*flux + diff(flux) )/VF$mid/grid$dx
+    }
+    if (any(is.na(dC))) { browser() }
     if (!full.output) {
         return(list(dC = dC, flux.up = flux[1], flux.down = flux[length(flux)]))
     }

### 2015-10-25 notes/diffusion-local-times.Rmd
got log-scale tran.1D working

@@
 rev.pde <- function (t,y,parms,...) {
-    p <- cline.interp(tau-t,soln=fwds.soln)
-    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline.interp(tau-t)), interp="linear")
+    p <- cline_interp(tau-t,soln=fwds.soln)
+    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
     tran <- tran.1D(C=y, A=p, D=1/2, dx=thegrid)$dC
@@
     matplot( thegrid$x.mid, 
-            pcline(thegrid$x.mid, p=sapply( pde.solns, cline.interp, T=tval ) ), 
+            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
                    type='l', main=sprintf("t=%d",tval) 

### 2015-10-25 notes/pde-clines.Rmd
got log-scale tran.1D working

@@
 ```{r solve_pde_setup, cache=TRUE}
-source("cline-fns.R")
-thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
+thegrid <- setup.grid.1D(x.up=-50, x.down=50, N = 5000)
 solve.pde
@@
 r <- 0.1
-pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s, grid=thegrid )
+log_u <- function(x,...){ tanh_cline(x=-x,...,log=TRUE) }
+log_1mu <- function(x,...){ tanh_cline(x=x,...,log=TRUE) }
+pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, s=s, grid=thegrid, log.u=TRUE )
 image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
@@
 ```
+**Note:** the range should be wide enough these are *going to zero*
+at the ends, or else we have to worry about the boundary conditions.
 

### 2015-10-25 notes/reactran-notes.Rmd
got log-scale tran.1D working

@@
 ```
+
+Proof of concept
+----------------
+
+```{r check_reactran}
+library(ReacTran)
+orig.tran.1D <- tran.1D
+# modified version with log.A option
+source("tran.1D.R")
+
+thegrid <- setup.grid.1D(x.up=-30, x.down=30, N = 5000)
+
+Avec <- tanh_cline(thegrid$x.mid,s=1)
+
+# original form
+pde.fn <- function (t,y,parms,...) {
+    tran <- orig.tran.1D(C=y, A=Avec, D=1/2, dx=thegrid)$dC
+    list( tran + (1-Avec) * y )
+}
+
+yinit <- rep(1.0,thegrid$N)
+times <- seq(0,1,length.out=100)
+soln <- ode.1D( y=yinit, times=times, func=pde.fn, nspec=1 )
+image(soln,legend=TRUE)
+
+logAvec <- tanh_cline(thegrid$x.mid,log=TRUE,s=1)
+
+pde.fn.log <- function (t,y,parms,...) {
+    tran <- tran.1D(C=y, A=logAvec, D=1/2, dx=thegrid, log.A=TRUE)$dC
+    list( tran + (1-Avec) * y )
+}
+
+soln <- ode.1D( y=yinit, times=times, func=pde.fn.log, nspec=1 )
+image(soln,legend=TRUE)
+```

### 2015-10-25 notes/cline-fns.R

@@ -2,8 +2,12 @@ pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
     (1/2)*( 1 - sign(x)*p )
 }
 
-tanh_cline <- function (x,s,sigma=1,...) {
-    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+tanh_cline <- function (x,s,sigma=1,log=FALSE,...) {
+    if (log) {
+        - (4*x*sqrt(s)/sigma) - log1p(exp(-4*x*sqrt(s)/sigma))
+    } else {
+        (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+    }
 }
 
 tanh_drift <- function (x,s,sigma=1) {
@@ -56,19 +60,23 @@ cline_interp <- function (T,soln) {
     return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
 }
 
-solve.pde <- function ( u, v=u, r, times, grid, ... ) {
+solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE, 
+                       um1=function(...){1-u(...)}, ... ) {
     # Solve the pde for a lineage 
     #  where 'u' gives the drift and 'v' the rate of recombination.
+    #  and if log.u is TRUE then u is actually log(proportion)
     vfn <- function(x,t){v(x=x,t=t,...)}
-    u.vec <- u(x=grid$x.mid,t=t,...)
+    u.vec <- u(x=grid$x.int,t=t,...)
+    um1.vec <- um1(x=grid$x.int,t=t,...) # 1-u.vec, but maybe log(1-u)
     v.vec <- v(x=grid$x.mid,t=t,...)
-    yinit <- c( rep(1.0,grid$N), rep(0.0,grid$N) )
+    if (log.u) { v.vec <- exp(v.vec) }
+    yinit <- c( rep(0.0,grid$N), rep(1.0,grid$N) )
     r <- 0.1
     pde.fn <- function (t,y,parms,...) {
         yA <- y[1:grid$N]
         yB <- y[grid$N+(1:grid$N)]
-        tran.A <- tran.1D(C=yA, A=u.vec, D=1/2, dx=grid)$dC
-        tran.B <- tran.1D(C=yB, A=1-u.vec, D=1/2, dx=grid)$dC
+        tran.A <- tran.1D(C=yA, A=u.vec, D=1/2, dx=grid, log.A=log.u)$dC
[…]

### 2015-10-25 notes/tran.1D.R

@@ -1,3 +1,4 @@
+
 tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL, 
     flux.down = NULL, a.bl.up = NULL, a.bl.down = NULL, D = 0, 
     v = 0, AFDW = 1, VF = 1, A = 1, dx, 
@@ -24,9 +25,8 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
             (rep(VF, length.out = (N + 1))[1:N] + rep(VF, length.out = (N + 
                 1))[2:(N + 1)]))
     if (!is.list(A)) 
-        A <- list(int = rep(A, length.out = (N + 1)), mid = 0.5 * 
-            (rep(A, length.out = (length(C) + 1))[1:N] + rep(A, 
-                length.out = (N + 1))[2:(N + 1)]))
+        A <- list(int = rep(A, length.out = (N + 1)), 
+                  mid = 0.5 * (rep(A, length.out = (length(C) + 1))[1:N] + rep(A, length.out = (N + 1))[2:(N + 1)]))
     if (is.list(dx)) 
         grid <- dx
     if (!is.list(dx)) 
@@ -278,9 +278,10 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
     if (!log.A) {
         dC <- -diff(A$int * flux)/A$mid/VF$mid/grid$dx
     } else {
-        dC <- -( diff(A$int)*flux + diff(flux) )/VF$mid/grid$dx
+        # (e^x f(x) - e^y f(y))/e^z = e^(x-z) f(x) - e^(y-z) f(y)
+        dC <- -( exp(A$int[-1]-A$mid)*flux[-1] - exp(A$int[-grid$N]-A$mid)*flux[-grid$N] )/VF$mid/grid$dx
     }
-    if (any(is.na(dC))) { browser() }
+    # if (any(is.na(dC))) { browser() }
     if (!full.output) {
         return(list(dC = dC, flux.up = flux[1], flux.down = flux[length(flux)]))
     }

### 2015-10-25 notes/pde-clines.Rmd
matching sims sorta?

--- 2015-10-25 notes/pde-clines.Rmd
@@
 library(ReacTran)
+source("tran.1D.R") # fix for log.A
 ```
@@
 ```{r solve_pde_setup, cache=TRUE}
-source("cline-fns.R")
-thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
+thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
+log_u <- function(x,...){ tanh_cline(x=-x,...,log=TRUE) }
+log_1mu <- function(x,...){ tanh_cline(x=x,...,log=TRUE) }
+cline.from.soln
 solve.pde
-cline.from.soln
 ```
@@
 ```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
+r <- 0.1
 s <- 0.1
+sigma <- 10
 zone.age <- 100  # this is tau
 tt <- seq(0,zone.age,length.out=100)
-r <- 0.1
-pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s, grid=thegrid )
+pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
 image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
@@
 These combine to give the predicted cline:
-```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
+```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
 pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
@@
 source("../sims/sim-fns.R")
-zone.age	= 20
 ndemes 		= 500
@@
 ninds 		= ndemes*deme.size
-SIGMA 		= 10
-s           = 0.1
-qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
+qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s, pos=c(0.5)))
 # spatial locations
@@
 # run the simulation
-sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
+sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=sigma,quiet=TRUE)
 # post-process
@@
 plot(xx,freqs[,1],type='l',xlab="spatial location",ylab="proportion")
-lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red')
+lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red')
 ```
@@
 Here's the PDE solution:
-```{r match_pde}
+```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
 tt <- seq(0,zone.age,length.out=100)
-pde.solns <- lapply(rvals, function (r) { solve.pde( u=function(...)1-tanh_cline(...), r=r, times=tt, s=s, grid=thegrid ) })
+pde.solns <- lapply(rvals, function (r) { solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
 pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
-match.clines <- sapply(pde.clines,function(x)x[length(tt),])
+match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
 ```
@@
     plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
-    lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red',lwd=3)
-    lines(1-thegrid$x.mid*SIGMA,match.clines[,k],col='green',lwd=3)
+    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
+    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
 }

### 2015-10-25 notes/cline-fns.R

@@ -71,7 +71,6 @@ solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE,
     v.vec <- v(x=grid$x.mid,t=t,...)
     if (log.u) { v.vec <- exp(v.vec) }
     yinit <- c( rep(0.0,grid$N), rep(1.0,grid$N) )
-    r <- 0.1
     pde.fn <- function (t,y,parms,...) {
         yA <- y[1:grid$N]
         yB <- y[grid$N+(1:grid$N)]
@@ -90,6 +89,7 @@ solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE,
             function(x,t) {u(x=x,...)} 
         }
     attr(soln,"v") <- vfn
+    attr(soln,"r") <- r
     return(soln)
 }
 cline.from.soln <- function (soln, grid, clinefn=attr(soln,"u") ) {

### 2015-10-25 notes/diffusion-local-times.Rmd
fixed bug in tran.1D fix

@@
 pde.fn <- function (t,y,parms,...) {
-    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
     list( tran - r * test.hetfn$mid * y )
@@
 pde.fn <- function (t,y,parms,...) {
-    tran <- tran.1D(C=y, A=clinefn, D=1e-6, dx=thegrid)$dC
+    tran <- tran.1D(C=y, A=clinefn, D=1e-6, flux.up=0, flux.down=0, dx=thegrid)$dC
     list( tran - r * hetfn$mid * y )
@@
 pde.fn <- function (t,y,parms,...) {
-    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
     list( tran - r * hetfn$mid * y )
@@
     # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
-    tran <- tran.1D(C=y, A=p, D=1/2, dx=thegrid)$dC
+    tran <- tran.1D(C=y, A=p, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
     list( tran - r * (1-p) * y )
@@
 # or else
-#   source("tran.1D.R")
+source("tran.1D.R")
 # to fix bug in using convective boundary conditions.
@@
 ```{r show_cline, cache=TRUE}
-zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) }
+zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r,...){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) }
 xx <- seq(-15,15,length.out=500)
@@
         pde.fn <- function (t,y,parms,...) {
-            tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
+            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
             if (any(is.na(tran))) { browser() }

### 2015-10-25 notes/pde-clines.Rmd
fixed bug in tran.1D fix

@@
 $$
-q(x,z,\tau,r) = \P\{ Z_\tau = A \} .
+q(x,z,\tau,r) = \P^x\{ Z_\tau = A \} .
 $$
 (Note that this gives us linkage to the selected site,
-and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau) q(x,B,\tau,r)$,
+and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau))q(x,B,\tau,r)$,
 but it is easier to compute this.)
+Also note that by symmetry, we should have that
+if $r>0$, then
+$$\begin{aligned}
+    q(x,z,\tau,r) \to \frac{1}{2} \qquad \text{as } \tau \to \infty .
+\end{aligned}$$
 
@@
 sigma <- 10
-zone.age <- 100  # this is tau
+zone.age <- 200  # this is tau
 tt <- seq(0,zone.age,length.out=100)
@@
 # same thing, plotted with lines
-clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
+clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
 ```
@@
 
+The smaller $r$ is, the slower the per-background probabilities smooth out:
+```{r small_r_sim}
+r <- 0.01
+pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
+image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
+clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
+```
+
+And, the cline flattens out *very* slowly:
+```{r longer_t_sim}
+r <- 0.1
+zone.age <- 2000  # this is tau
+tt <- seq(0,zone.age,length.out=100)
+pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
+image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
+clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
+```
+This should be symmetric.  Let's check:
+```{r check_symmetry}
+pde.soln <- solve.pde( um1=log_u, u=log_1mu, r=r, times=tt, 
+                yinit=c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) ),
+                s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
+image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
+clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
+```
+
 Comparison to simulation
-------------------------
+========================
 
@@
 match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
+for (k in floor(seq(1,length(rvals),length.out=3))) {
+    image(pde.solns[[k]],legend=TRUE,main=sprintf("r=%f",rvals[k]))
+}
 ```

### 2015-10-25 notes/reactran-notes.Rmd
fixed bug in tran.1D fix

@@
 ---
+
+The interface area
+------------------
 
@@
 ```
+
+<!--
+Boundary conditions
+-------------------
+
+We want the function `tran.1D` to return the generator of a Markov process
+applied to a function
+-- not a sub-Markov process, with killing,
+as that appears in the reaction term.
+What should the boundary conditions be?
+Let $G$ be the generator,
+which we write in the form
+$$\begin{aligned}
+Gf = \frac{1}{A} \grad \cdot A D \grad f,
+\end{aligned}$$
+where $D \grad f$ is the *flux*.
+We know that $G\textbf{1}=0$,
+where $\textbf{1}$ is the constant function.
+Well,
+$$\begin{aligned}
+    G\textbf{1} 
+    &=
+\end{aligned}$$
+
+-->

### 2015-10-25 notes/cline-fns.R

@@ -60,22 +60,43 @@ cline_interp <- function (T,soln) {
     return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
 }
 
-solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE, 
-                       um1=function(...){1-u(...)}, ... ) {
+forwards.pde <- function (s,times,grid) {
+    # solve the pde for frequencies, forwards in time
+    yinit <- ifelse( grid$x.mid>0, 0, 1 )
+    fwds.pde <- function (t,y,parms,...) {
+        tran <- tran.1D(C=y, D=1/2, dx=grid)$dC
+        list( tran + s * y * (1-y) * (2*y-1) )
+    }
+    soln <- ode.1D( y=yinit, times=times, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
+    attr(soln,"s") <- s
+    return(soln)
+}
+
+solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE, um1, 
+                yinit=c( rep(0.0,grid$N), rep(1.0,grid$N) ), ... ) {
     # Solve the pde for a lineage 
+    #  moving between backgrounds
     #  where 'u' gives the drift and 'v' the rate of recombination.
     #  and if log.u is TRUE then u is actually log(proportion)
+    # Note that um1 should be provided if u is in log scale.
     vfn <- function(x,t){v(x=x,t=t,...)}
     u.vec <- u(x=grid$x.int,t=t,...)
+    if (missing(um1)) {
+        if (log.u) {
+            warning("Computing 1-u from u even though u is in log scale.")
+            um1 <- function(...){-expm1(u(...))}
+        } else {
+            um1 <- function(...) u(...)
+        }
+    }
     um1.vec <- um1(x=grid$x.int,t=t,...) # 1-u.vec, but maybe log(1-u)
     v.vec <- v(x=grid$x.mid,t=t,...)
     if (log.u) { v.vec <- exp(v.vec) }
-    yinit <- c( rep(0.0,grid$N), rep(1.0,grid$N) )
     pde.fn <- function (t,y,parms,...) {
         yA <- y[1:grid$N]
         yB <- y[grid$N+(1:grid$N)]
-        tran.A <- tran.1D(C=yA, A=u.vec, D=1/2, dx=grid, log.A=log.u)$dC
[…]

### 2015-10-25 notes/tran.1D.R

@@ -104,6 +104,7 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
                 stop("error: A$mid should be a vector of length 1 or N")
         }
         if ( (any(A$int <= 0) || any(A$mid <= 0)) || (!log.A && ( any(A$int==0) || any(A$mid==0) )) )
+        if ( !log.A && (any(A$int < 0) || any(A$mid < 0)) )
             stop("error: the area A should always be positive")
         gn <- names(grid)
         if (!"dx" %in% gn) 
@@ -274,13 +275,14 @@ tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL,
         flux[1] <- flux.up
     if (!is.null(flux.down)) 
         flux[N + 1] <- flux.down
-    
+
     if (!log.A) {
         dC <- -diff(A$int * flux)/A$mid/VF$mid/grid$dx
     } else {
         # (e^x f(x) - e^y f(y))/e^z = e^(x-z) f(x) - e^(y-z) f(y)
-        dC <- -( exp(A$int[-1]-A$mid)*flux[-1] - exp(A$int[-grid$N]-A$mid)*flux[-grid$N] )/VF$mid/grid$dx
+        dC <- -( exp(A$int[-1]-A$mid)*flux[-1] - exp(A$int[-(grid$N+1)]-A$mid)*flux[-(grid$N+1)] )/VF$mid/grid$dx
     }
+
     # if (any(is.na(dC))) { browser() }
     if (!full.output) {
         return(list(dC = dC, flux.up = flux[1], flux.down = flux[length(flux)]))

### 2015-10-25 notes/pde-clines.Rmd
another bug

@@
 $$
-(Note that this gives us linkage to the selected site,
-and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau))q(x,B,\tau,r)$,
-but it is easier to compute this.)
 Also note that by symmetry, we should have that
@@
         \grad \log(1-p(x)) \cdot \grad q_B(x,t) + \frac{\sigma^2}{2} \Delta q_B(x,t) + r p(x) (q_A(x,t)-q_B(x,t)) .
+\end{aligned}$$
+
+The cline shape
+---------------
+
+The PDE above gives probabilities given the initial background,
+so therefore *linkage* to the selected site.
+For the cline shape we need to average over choice of background,
+which we denote
+$$\begin{aligned}
+p(x,\tau,r) =  p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau))q(x,B,\tau,r) .
+\end{aligned}$$
+Note that
+$$\begin{aligned}
+q(x,A,\tau,r) &= 1-q(-x,B,\tau,r) \\
+p(x,\tau) &= 1-p(-x,\tau) 
+\end{aligned}$$
+and so
+$$\begin{aligned}
+p(x,\tau,r) =  p(x,\tau) q(x,A,\tau,r) + p(-x,\tau) (1-q(-x,A,\tau,r))A .
 \end{aligned}$$
@@
 ```{r longer_t_sim}
-r <- 0.1
+r <- 1.0
 zone.age <- 2000  # this is tau
@@
 pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
-                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
-image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
-axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
-pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
-clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
-```
-This should be symmetric.  Let's check:
-```{r check_symmetry}
-pde.soln <- solve.pde( um1=log_u, u=log_1mu, r=r, times=tt, 
-                yinit=c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) ),
                   s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
@@
 rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
-loci <- list(0.5+rvals)
+loci <- list(0.5+c(rev(rvals[-1]),rvals))
 genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
@@
 ```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
+zone.age <- sim$pars["n.gen"]
 tt <- seq(0,zone.age,length.out=100)

### 2015-10-25 notes/cline-fns.R

@@ -120,8 +120,8 @@ cline.from.soln <- function (soln, grid, clinefn=attr(soln,"u") ) {
     A.inds <- 1+(1:grid$N)
     B.inds <- 1+grid$N+(1:grid$N)
     sel.cline <- outer( grid$x.mid, soln[nrow(soln),1], clinefn )
-    pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
-    attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
+    pde.cline <- cbind( soln[,1], soln[,A.inds]*sel.cline[col(soln[,A.inds])] + soln[,B.inds]*(1-sel.cline)[col(soln[,B.inds])] )
+    attributes(pde.cline) <- c( attributes(pde.cline), attributes(soln)[setdiff(names(attributes(soln)),c("dim","dimnames"))] )
     attr(pde.cline,"nspec")<-1
     return(pde.cline)
 }

### 2015-10-26 notes/pde-clines.Rmd
made different Rmd files not interfere with each other

@@
 log_1mu <- function(x,...){ tanh_cline(x=x,...,log=TRUE) }
-cline.from.soln
-solve.pde
+cline_from_soln
+solve_pde
 ```
@@
 s <- 0.1
-sigma <- 10
+sigma <- 5
 zone.age <- 200  # this is tau
 tt <- seq(0,zone.age,length.out=100)
-pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+pde.soln <- solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                   s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
@@
 ```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
-pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
-image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
+pde.cline <- cline_from_soln(pde.soln,grid=thegrid)
+image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r),yaxt='n')
+axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
 # same thing, plotted with lines
@@
 r <- 0.01
-pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+pde.soln <- solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                   s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
@@
 axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
-pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
+pde.cline <- cline_from_soln(pde.soln,grid=thegrid)
+layout(1)
 clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
@@
 tt <- seq(0,zone.age,length.out=100)
-pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+pde.soln <- solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                   s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
@@
 source("../sims/sim-fns.R")
+zone.age    = 50
 ndemes 		= 500
@@
 tt <- seq(0,zone.age,length.out=100)
-pde.solns <- lapply(rvals, function (r) { solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
+pde.solns <- lapply(rvals, function (r) { solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                 s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
-pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
+pde.clines <- lapply( pde.solns, cline_from_soln, grid=thegrid )
 match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
-for (k in floor(seq(1,length(rvals),length.out=3))) {
+for (k in floor(seq(2,length(rvals),length.out=3))) {
     image(pde.solns[[k]],legend=TRUE,main=sprintf("r=%f",rvals[k]))

### 2015-10-26 notes/cline-fns.R

@@ -60,11 +60,11 @@ cline_interp <- function (T,soln) {
     return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
 }
 
-forwards.pde <- function (s,times,grid) {
+forwards_pde <- function (s,times,grid,sigma=1) {
     # solve the pde for frequencies, forwards in time
     yinit <- ifelse( grid$x.mid>0, 0, 1 )
     fwds.pde <- function (t,y,parms,...) {
-        tran <- tran.1D(C=y, D=1/2, dx=grid)$dC
+        tran <- tran.1D(C=y, D=sigma^2/2, dx=grid)$dC
         list( tran + s * y * (1-y) * (2*y-1) )
     }
     soln <- ode.1D( y=yinit, times=times, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
@@ -72,14 +72,38 @@ forwards.pde <- function (s,times,grid) {
     return(soln)
 }
 
-solve.pde <- function ( u, v=u, r, times, grid, log.u=FALSE, um1, 
+forwards_backwards_pde <- function (s, times, grid, r, sigma=1,
+        fwds.soln=forwards_pde(s=s,times=times,grid=grid),
+        yinit=c( rep(0.0,grid$N), rep(1.0,grid$N) ), ... ) {
+    # solve the pde for a lineage 
+    # run backwards in the forwards-time profile
+    rev.pde.fn <- function (t,y,parms,...) {
+        yA <- y[1:grid$N]
+        yB <- y[grid$N+(1:grid$N)]
+        p <- cline_interp(max(times)-t,soln=fwds.soln)
+        tran.A <- tran.1D(C=yA, A=p, D=sigma^2/2, dx=grid, flux.up=0, flux.down=0)$dC
+        tran.B <- tran.1D(C=yB, A=1-p, D=sigma^2/2, dx=grid, flux.up=0, flux.down=0)$dC
+        # if (any(is.na(tran.A)|is.na(tran.B))) { browser() }
+        list( c( 
+                tran.A + r * (1-p) * (yB-yA),
+                tran.B + r * p * (yA-yB)
+                ) )
+    }
+    rev.soln <- ode.1D( y=yinit, times=times, func=rev.pde.fn, nspec=2, tcrit=max(times) ) # note FIRST COLUMN IS TIME
+    attr(rev.soln,"r") <- r
+    attr(rev.soln,"s") <- s
+    attr(rev.soln,"s") <- s
+}
+
[…]

### 2015-10-26 notes/diffusion-local-times.Rmd
.

@@
 Now let's get local times for those walks.
-```{r lineage_local_times,cache=TRUE,depends="sim_lineage"}
+```{r lineage_local_times,cache=TRUE,dependson="sim_lineage"}
 xx <- seq( min(rws), max(rws), length.out=200 )
@@
 Here is the distribution of $T=\int_0^{`r tau`} f_H(X_t) dt$:
-```{r dist_het_time, cache=TRUE, depends="sim_het_time"}
+```{r dist_het_time, cache=TRUE, dependson="sim_het_time"}
 het.time <- weighted_time( rws, tanh_het, s=s )
@@
 Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
-```{r solve_pde_check, cache=TRUE, depends="solve_pde_setup"}
+```{r solve_pde_check, cache=TRUE, dependson="solve_pde_setup"}
 layout(t(1:2))
@@
 Now, to solve the real problem:
-```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
+```{r solve_pde, cache=TRUE, dependson="solve_pde_setup"}
 pde.fn <- function (t,y,parms,...) {
@@
 and, at other values of $t$:
-```{r check_pde_soln_2, cache=TRUE, depends="check_pde_soln"}
+```{r check_pde_soln_2, cache=TRUE, dependson="check_pde_soln"}
 layout(t(1:2))
@@
 Now, let's run the killed lineage backwards in it:
-```{r fwds_back_pde, cache=TRUE, depends="fwds_pde"}
+```{r fwds_back_pde, cache=TRUE, dependson="fwds_pde"}
 yinit <- rep(1.0,thegrid$N)
@@
 Let's compare the case where the cline is set up instantly or not:
-```{r compare_rightness, cache=TRUE,depends=c("fwds_back_pde","solve_pde_setup","setup_fns")}
+```{r compare_rightness, cache=TRUE,dependson=c("fwds_back_pde","solve_pde_setup","setup_fns")}
 layout(t(1:2))
@@
 (the second plots are on a log-log scale):
-```{r recomb_prob_comparisons, cache=TRUE, depends="sim_general"}
+```{r recomb_prob_comparisons, cache=TRUE, dependson="sim_general"}
 rvals <- exp(seq(log(.001),log(0.5),length.out=400))
@@
 First, the local time at 0:
-```{r local_times_0, cache=TRUE, depends="sim_rw"}
+```{r local_times_0, cache=TRUE, dependson="sim_rw"}
 xh <- hist(colSums(abs(rws)<=1.5*eps)/ntimes/(3*eps),main="local time at 0", breaks=30)
@@
 the colors (red/green/blue) use the (mid/upper/lower) points of the interval that we count in.
-```{r local_times, cache=TRUE, depends="sim_rw"}
+```{r local_times, cache=TRUE, dependson="sim_rw"}
 layout(matrix(1:4,nrow=2))
@@
 Let's check that, too: why not?
-```{r killing, cache=TRUE, depends="sim_rw"}
+```{r killing, cache=TRUE, dependson="sim_rw"}
 zeta.reflected
@@
 by smoothing the probabilities for the simulation:
-```{r sim_clines, cache=TRUE, depends=c("sim_pushed","pushed_pde")}
+```{r sim_clines, cache=TRUE, dependson=c("sim_pushed","pushed_pde")}
 prob.no.recomb <- sapply( rvals, function (r) { fitted(loess( exp(-r*sim.wt) ~ xinit, span=0.1)) } )
@@
 And, finally, here's some actual clines predicted by this model:
-```{r pushed_cline,cache=TRUE,depends="pushed_pde"}
+```{r pushed_cline,cache=TRUE,dependson="pushed_pde"}
 tau <- 100

### 2015-10-26 notes/pde-clines.Rmd
.

@@
 Let's solve this numerically.
-```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
+```{r solve_pde, cache=TRUE, dependson="solve_pde_setup"}
 r <- 0.1
@@
 These combine to give the predicted cline:
-```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
+```{r show_pde_cline,cache=TRUE,dependson=c("solve_pde_setup","solve_pde")}
 pde.cline <- cline_from_soln(pde.soln,grid=thegrid)
@@
 and, postprocess:
-```{r genotype_sims, cache=TRUE, depends="setup_sims"}
+```{r genotype_sims, cache=TRUE, dependson="setup_sims"}
 rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
-loci <- list(0.5+c(rev(rvals[-1]),rvals))
+loci <- list(0.5+c(rev(rvals[-1]),-rvals))
 genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
@@
 Here's the PDE solution:
-```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
+```{r match_pde,cache=TRUE,dependson="solve_pde_setup"}
 zone.age <- sim$pars["n.gen"]
@@
 for (k in seq_along(rvals)) {
-    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
+    matplot(xx,freqs[,which(abs(rvals[k]-loci[[1]]+0.5)<1e-9)],type='l',
+            xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
     lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)

### 2015-10-30 notes/diffusion-local-times.Rmd
.

@@
 
-And, finally, here's some actual clines predicted by this model:
-```{r pushed_cline,cache=TRUE,dependson="pushed_pde"}
-tau <- 100
-tt <- seq(0,tau,length.out=100)
-pde.solns <- lapply( rvals, function (r) {
-        pde.fn <- function (t,y,parms,...) {
-            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
-            if (any(is.na(tran))) { browser() }
-            list( tran - r * hetfn$mid * y )
-        }
-        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
-      } )
-plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
-layout(t(1:2))
-for (tval in plot.tvals) {
-    matplot( thegrid$x.mid, 
-            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
-                   type='l', main=sprintf("t=%d",tval) 
-        )
-    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
-}
-```
-
 Analytical form

### 2015-10-30 notes/rough-calculations.Rmd
.

---
title: "Clines from PDE"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\grad}{\nabla}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
source("tran.1D.R") # fix for log.A
```

*Useful functions used below:*
```{r setup_fns}
source("cline-fns.R")
```

In the following, $x$ is the distance from the cline center,
$r$ is the recombination rate to the selected locus,
$\tau$ is the time since secondary contact.
The two alleles are $A$ and $B$.
We would like to know:

1.  What is the width of the cline at a linked, neutral locus?
    How long does it take to flatten out?
    How far from the cline do you have to go to find mostly one type?

2.  How long are the haplotypes around the rare $B$ alleles on the $A$ side?
    This should be constant in time.

3.  How long, and how common, are chunks of $B$ haplotype *not* linked to $B$ alleles on the $A$ side?
    These will accumulate as time progresses.

4.  How does neighborhood size affect the results?
    Does the mean cline, averaged across low-population-density simulations,
    match the cline obtained by a high population density, i.e., the PDE?

5.  What about the variance??

### 2015-10-30 notes/rough-calculations.Rmd
notes

@@
 ---
-title: "Clines from PDE"
+title: "Rough calculations"
 date: "`r date()`"
@@
 5.  What about the variance??
+
+6.  For a pair of sites, what is the correlation in ancestral state at those loci,
+    given the state at the selected locus
+    (and, also, averaged over choice of selected locus).
+
+    a.  Allow for ascertainment of SNPs in extremal individuals
+        by estimating probability of coalescence in ancestral populations
+        from the extremal individuals.
+        For each pair of loci, correlation of these two loci with extremal individuals.
+
+How a neutral cline flattens
+============================
+
+1.  *What is the width of the cline at a linked, neutral locus?*
+    *How long does it take to flatten out?*
+    *How far from the cline do you have to go to find mostly one type?*

### 2015-10-31 notes/diffusion-local-times.Rmd
rough calcs

@@
 \renewcommand{\P}{\mathbb{P}}
+\newcommand{\calE}{\mathcal{E}}
 \DeclareMathOperator{\sgn}{sgn}
@@
 Also, $p_t(x,z;\theta)=p_t(-x,-z;-\theta)$ gets the same expressions for $x<0$.
+
+Notes on excursion theory
+=========================
+
+Let $n(e)$ denote the excursion measure of Brownian motion,
+on $\calE = \{ f : [0,\infty) \to [0,\infty) \, \text{s.t.} f(s)=0, t>s \Rightarrow f(t)=0 \}$.
+For $f \in \calE$ denote $\zeta(f) = \sup\{ t : f(t)>0\}$.
+Then [Rogers 1989](http://www.bibsonomy.org/bibtex/201686d46785a10af443e4d48a5196088/peter.ralph)
+$$\begin{aligned}
+n(f : \sup_t f(t) > a ) = \frac{1}{2a} .
+\end{aligned}$$
+and
+$$\begin{aligned}
+n(f : \zeta f > t ) = \sqrt{\frac{2}{\pi t}} .
+\end{aligned}$$

### 2015-10-31 notes/rough-calculations.Rmd
rough calcs

@@
 1.  *What is the width of the cline at a linked, neutral locus?*
-    *How long does it take to flatten out?*
-    *How far from the cline do you have to go to find mostly one type?*
+    *How long does it take to flatten out?* *How far from the cline do you have to go to find mostly one type?*
+
+**Answer:**
+
+a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
+b.  The selected cline takes time $s$ to spread out to its full width.
+c.  Cline behaves like the selected cline for $r < 1/\tau$ (more tentatively, for $r < \sqrt{\frac{2}{s\tau}}$).
+d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.
+
+*Rationale:*
+
+a.  The maximum of Brownian motion until time $t$ has mean $\sqrt{2t/\pi}$,
+    and has the same marginal distribution as the absolute value of Brownian motion.
+    Therefore, the probability a lineage from $x$ hasn't got to the center of the cline *at all* by time $\tau$ 
+    is about $2\Phi(x/\sqrt{\sigma^2 \tau})-1$;
+    and therefore, if $x/\sigma > 2\sqrt{\tau}$ then the cline cannot have started to form, regardless of $r$.
+
+b.  The selected cline spreads out (in forwards time!) with diffusion,
+    and so will reach maximum width $\sigma/\sqrt{s}$ in time $s$.
+    Let $q(t,x)$ be the solution to
+    $$\begin{aligned}
+    \dot q = \frac{1}{2} q'' + q (1-q) (2q-1) .
+    \end{aligned}$$
+    If $u=at$ and $y=bx$ and let $p(t,x) = q(at,bx)$ then
+    $$\begin{aligned}
+    \dot p(t,x)
+        &= \partial_t q(at,bx) \\
+        &= a \dot q(at,bx) \\
+        &= a \left\{ \frac{1}{2} q''(at,bx) + q (1-q) (2q-1) \right\} \\
+        &= a \left\{ \frac{1}{2b^2} \partial^2_x q(at,bx) + q (1-q) (2q-1) \right\} \\
+        &= \frac{a}{2b^2} p''(t,x) + a p (1-p) (2p-1)  .
+    \end{aligned}$$
+    If we choose $a=s$ and $b=\sigma/\sqrt{s}$ then $a/b^2=\sigma^2$, so
+    $$\begin{aligned}
+    \dot p = \frac{\sigma^2}{2} p'' + s p (1-p) (2p-1) .
+    \end{aligned}$$
+    We know that the unitless cline $q$ will spread out in time 1;
+    so our cline will spread out in time $s$.
+
+c.  The cline will be narrower if $r$ is small.
+    Consider a lineage begun near the center of the cline.
+    The most time it can spend in heterozygotes is $\tau$,
+    so if $r \ll 1/\tau$, the chance of recombining is small.
+    So, the cline will look like the cline at the selected locus if $r < 1/\tau$.
+
+d.  By L\'evy's identity, the local time of a Brownian motion at zero
+    has the same distribution (pathwise!) as the running maximum of Brownian motion.
+    Therefore, the local time at zero up until $\tau$ grows like $\sqrt{\tau}$.
+    The same is true for Brownian motion reflected at zero.
+    The time a lineage spends in heterozygotes should be roughly proportional to the local time
+    (each time it visits zero from the left, it spends some time in hets before leaving again; 
+    the constant of proportionality is the mean time spent per visit).
+    Therefore, there will be appreciable chance of recombination if $r C \sqrt{\tau}>1$,
+    where $C$ is the constant;
+    this says that if clines down to $r$ are flat at $\tau$, then at time $2\tau$ they'll be flat down to $r/4$.
+
+    But, what is the constant $C$?
+    Consider the approximation of a lineage that acts like Brownian motion with a constant drift $s$ above zero.
+    As noted in [this file](diffusion-local-times.html),
+    the excursion measure of Brownian motion assigns mass $\sqrt{\frac{1}{2 \pi t^3}}$ to excursions of length $t$;
+    we must tilt this by $e^{-ts}$ for the selection above zero;
+    so that the mean rate at which this semi-pushed motion accumulates time above zero,
+    with respect to local time at zero,
+    is  (letting $x=\sqrt{2st}$)
+    $$\begin{aligned}
+    \int_0^\infty t \frac{e^{-st}}{ \sqrt{2 \pi t^3 } } dt 
+    &=
+    \int_0^\infty \frac{e^{-st}}{ \sqrt{2 \pi t } } dt \\
+    &=
+    \sqrt{2s} \int_0^\infty \frac{e^{-x^2/2}}{ \sqrt{2 \pi} } dx \\
+    &=
+    \sqrt{\frac{s}{2}} .
+    \end{aligned}$$
+    This suggests taking $C=\sqrt{s/2}$.

### 2015-10-31 notes/rough-calculations.Rmd
testing

@@
 source("tran.1D.R") # fix for log.A
-```
-
-*Useful functions used below:*
-```{r setup_fns}
-source("cline-fns.R")
 ```
@@
 b.  The selected cline takes time $s$ to spread out to its full width.
-c.  Cline behaves like the selected cline for $r < 1/\tau$ (more tentatively, for $r < \sqrt{\frac{2}{s\tau}}$).
+c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2}{s\tau}}$ if $\tau>1/s$).
 d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.
@@
 
-d.  By L\'evy's identity, the local time of a Brownian motion at zero
+d.  By Lévy's identity, the local time of a Brownian motion at zero
     has the same distribution (pathwise!) as the running maximum of Brownian motion.
@@
     This suggests taking $C=\sqrt{s/2}$.
+
+Simulation check
+----------------
+
+We'll check these with the PDE solutions
+(and, then with simulations):
+
+```{r setup_fns}
+source("cline-fns.R")
+```
+
+a.  Look at space-time frequency images at $r=1$ and check they spread out as $\sqrt{t}$.
+
+    ```{r unlinked_clines, cache=TRUE}
+    thegrid <- setup.grid.1D(x.up=-50, x.down=50, N = 5000)
+    s <- 0.01
+    r <- 1.0
+    tt <- seq(0,1000,length.out=250)
+    unlinked.solns <- forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE)
+    # image(unlinked.solns,legend=TRUE,zlim=c(0,1),yaxt='n')
+    # axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+    unlinked.cline  <- cline_from_soln( unlinked.solns, grid=thegrid )
+    image(unlinked.cline,legend=TRUE,zlim=c(0,1),yaxt='n',main=sprintf("r=%f, s=%f",r,s))
+    axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+    matlines(tt,0.5+outer(sqrt(tt)/diff(range(thegrid$x.mid)),(1:4)/2,"*"),lwd=2,col='green')
+    matlines(tt,0.5-outer(sqrt(tt)/diff(range(thegrid$x.mid)),(1:4)/2,"*"),lwd=2,col='green')
+    ```
+
+b.  Look at space-time frequency images at $r=0$ and several values of $s$
+    to check they become stationary after about $1/s$.
+    The clines are colored red for times less than $1/s$, and black afterwards;
+    the green dotted lines are at $\pm 1/\sqrt{s}$.
+    Indeed, cline flattening has slowed considerably by $1/s$, 
+    and the limiting cline reaches $(1/4,3/4)$ at $\pm 1/\sqrt{s}$.
+
+    ```{r forwards_selected_cline,cache=TRUE}
+    thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
+    svals <- exp(seq(log(1e-4),log(0.5),length.out=6))
+    fwds.solns <- lapply(svals, function(s) {
+            tt <- seq(0,3/s,length.out=40)
+            forwards_pde(s=s,times=tt,grid=thegrid) 
+        } )
+    for (k in seq_along(svals)) {
+        clineplot( fwds.solns[[k]], main=sprintf("s=%f",svals[k]), col=ifelse( fwds.solns[[k]][,1]>1/svals[k], adjustcolor("black",0.5), adjustcolor("red",0.5) ) )
+        abline(v=c(-1,1)/sqrt(svals[k]),lty=3,col='green',lwd=2)
+    }
+    ```
+
+c.  Look at space-recombination frequency images at a bunch of times
+    to check that flattening occurs around $r=\sqrt{2/s\tau}$.

### 2015-10-31 notes/cline-fns.R

@@ -43,11 +43,14 @@ weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
     colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
 }
 
-clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,ylab="probability",ylim=c(0,1),legend=TRUE,...) {
+clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,
+                       ylab="probability", ylim=c(0,1), legend=TRUE, lty=1,
+                       col=rainbow(length(times)),
+                       ...) {
     # plot output of deSolve as clines in a nice way
-    graphics::matplot( x, t(soln[times,-1]), type='l', col=rainbow(length(times)), lty=1, 
+    graphics::matplot( x, t(soln[times,-1]), type='l', col=col, lty=lty, 
             xlab="space", ylab=ylab, ylim=ylim, ...)
-    if (legend) legend("bottomleft", lty=1, col=rainbow(length(times)), legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
+    if (legend) legend("bottomleft", lty=lty, col=col, legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
     return( invisible( t(soln[times,-1]) ) )
 }
 
@@ -74,15 +77,26 @@ forwards_pde <- function (s,times,grid,sigma=1) {
 
 forwards_backwards_pde <- function (s, times, grid, r, sigma=1,
         fwds.soln=forwards_pde(s=s,times=times,grid=grid),
-        yinit=c( rep(0.0,grid$N), rep(1.0,grid$N) ), ... ) {
+        yinit=c( rep(0.0,grid$N), rep(1.0,grid$N) ), 
+        log.p=TRUE, eps=1e-16, ... ) {
     # solve the pde for a lineage 
     # run backwards in the forwards-time profile
+    #   ... note that we solve the PDE in the forwards-time direction in *both* cases.
+    # HOWEVER: note that this becomes singular near t=0,
+    #   to deal with this we introduce 'eps'
     rev.pde.fn <- function (t,y,parms,...) {
[…]

### 2015-11-01 notes/rough-calculations.Rmd
fixed bug in cline_from_soln

@@
 b.  The selected cline takes time $s$ to spread out to its full width.
-c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2}{s\tau}}$ if $\tau>1/s$).
-d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.
+c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2s}{\tau}}$ if $\tau>1/s$).
+d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2s}{\tau}}$.
 
@@
     &=
-    \sqrt{\frac{s}{2}} .
+    \frac{1}{2\sqrt{2s}} .
     \end{aligned}$$
@@
 a.  Look at space-time frequency images at $r=1$ and check they spread out as $\sqrt{t}$.
+    The green lines are at multiples of $\pm \sqrt{t}$.
 
@@
 c.  Look at space-recombination frequency images at a bunch of times
-    to check that flattening occurs around $r=\sqrt{2/s\tau}$.
+    to check that flattening occurs around $r=\sqrt{2s/\tau}$.
+    Note that the system is [stiff](https://en.wikipedia.org/wiki/Stiff_equation), 
+    as diffusive systems [usually are](http://scicomp.stackexchange.com/questions/891/the-definition-of-stiff-ode-system),
+    particularly for small $r$ at the times when each set of probabilities becomes steep;
+    setting $s=0.1$ this is a problem, but not with smaller (less steep) $s$.
+
+    ```{r space_recomb, cache=TRUE}
+    s <- 0.01
+    rvals <- exp(seq(log(1e-3),log(0.5),length.out=20))
+    tt <- seq(0,200,length.out=10)
+    thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
+    fwds.soln <- forwards_pde(s=s,times=seq(0,1.2*max(tt),length.out=1000),grid=thegrid)
+    r.solns <- lapply(rvals, function(r) {
+            forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+        } )
+    unlinked.clines  <- lapply(r.solns, cline_from_soln, grid=thegrid )
+    xr <- lapply( seq_along(tt), function (k) { do.call( cbind, lapply( unlinked.clines, function (x) x[k,-1] ) ) } )
+    ```
+    ```{r plot_sr}
+    for (k in 2:length(xr)) {
+        matplot(thegrid$x.mid, xr[[k]], type='l', col=rainbow(2*length(rvals)), lty=1, main=sprintf("t=%f",tt[k]) )
+        legend("bottomright",lty=1, col=rainbow(2*length(rvals)), legend=sprintf("r=%f : %f",rvals,rvals*sqrt(tt[k]/2/s)))
+    }
+    ```

### 2015-11-01 notes/cline-fns.R

@@ -63,9 +63,11 @@ cline_interp <- function (T,soln) {
     return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
 }
 
-forwards_pde <- function (s,times,grid,sigma=1) {
-    # solve the pde for frequencies, forwards in time
-    yinit <- ifelse( grid$x.mid>0, 0, 1 )
+forwards_pde <- function (s,times,grid,sigma=1,
+                      yinit=ifelse( grid$x.mid>0, 0, 1 )) {
+    # Solve the pde for frequencies, forwards in time:
+    #   by default, yinit=1 to the *left* of zero;
+    #   so finds the probability the selected site initiates on the left
     fwds.pde <- function (t,y,parms,...) {
         tran <- tran.1D(C=y, D=sigma^2/2, dx=grid)$dC
         list( tran + s * y * (1-y) * (2*y-1) )
@@ -76,26 +78,35 @@ forwards_pde <- function (s,times,grid,sigma=1) {
 }
 
 forwards_backwards_pde <- function (s, times, grid, r, sigma=1,
-        fwds.soln=forwards_pde(s=s,times=times,grid=grid),
-        yinit=c( rep(0.0,grid$N), rep(1.0,grid$N) ), 
+        fwds.soln=forwards_pde(s=s,times=times,grid=grid,yinit=ifelse(grid$x.mid<0,yinit[1:grid$N],yinit[grid$N+(1:grid$N)])),
+        yinit=c( rep(1.0,grid$N), rep(0.0,grid$N) ), 
         log.p=TRUE, eps=1e-16, ... ) {
-    # solve the pde for a lineage 
-    # run backwards in the forwards-time profile
+    # solve the pde for a lineage run backwards in the forwards-time profile
     #   ... note that we solve the PDE in the forwards-time direction in *both* cases.
     # HOWEVER: note that this becomes singular near t=0,
-    #   to deal with this we introduce 'eps'
+    #   so we push p away from the boundary by 'eps'.
+    # By default
+    #  - when the cline forms, A is on the left and B is on the right
+    #  - fwds.soln is the frequency of the type A selected locus,
+    #          i.e., the prob the selected lineage begins on the *left*
+    #  - yA, the first N numbers, is the probability that a locus linked to a type A selected locus
[…]

### 2015-11-01 notes/rough-calculations.Rmd
.

@@
     for (k in 2:length(xr)) {
-        matplot(thegrid$x.mid, xr[[k]], type='l', col=rainbow(2*length(rvals)), lty=1, main=sprintf("t=%f",tt[k]) )
-        legend("bottomright",lty=1, col=rainbow(2*length(rvals)), legend=sprintf("r=%f : %f",rvals,rvals*sqrt(tt[k]/2/s)))
+        matplot(thegrid$x.mid, xr[[k]], type='l', col=rainbow(2*length(rvals)), lty=1, main=sprintf("t=%f",tt[k]), ylab='probability' )
+        legend("bottomright",lty=1, col=rainbow(2*length(rvals)), legend=sprintf("r=%f : %f",rvals,rvals*sqrt(tt[k]/2/s)),cex=0.5)
     }

### 2015-11-02 notes/diffusion-local-times.Rmd
.

@@
 Let's compare the case where the cline is set up instantly or not:
-```{r compare_rightness, cache=TRUE,dependson=c("fwds_back_pde","solve_pde_setup","setup_fns")}
+```{r compare_rightness, cache=TRUE,dependson=c("fwds_back_pde","solve_pde_setup")}
 layout(t(1:2))

### 2015-11-02 notes/pde-clines.Rmd
.

@@
 $$\begin{aligned}
-q(x,A,0,r) = 1  \\
-q(x,B,0,r) = 0  \\
+q(x,A,0,r) &= 1  \\
+q(x,B,0,r) &= 0  \\
 \end{aligned}$$
@@
 
+Haplotype lengths
+=================
+
+There are various haplotype-related statistics we could compute:
+
+1.  Distance to closest recombination in a heterozygote
+
+Distance to closest recombination in a heterozygote
+---------------------------------------------------
+
+The first haplotype-related statistic
+is the distance along the chromosome from an $A$ allele until the closest recombination
+that occurred in an $AB$ heterozygote.
+Let $H$ be this distance;
+then, 
+$$\begin{aligned}
+    u(x,\tau,r)=\P^(x,\tau)\{H>r\}
+\end{aligned}$$
+solves the following PDE:
+$$\begin{aligned}
+    u(x,0,r) &= 1  \\
+    \partial_t u(x,t,r)
+        &= 
+        \grad \log(p(x)) \cdot \grad u(x,t,r) + \frac{\sigma^2}{2} \Delta u(x,t,r) - r (1-p(x)) u(x,t,r) .
+\end{aligned}$$
+
 Comparison to simulation

### 2015-11-02 notes/rough-calculations.Rmd
.

--- 2015-10-31 notes/rough-calculations.Rmd
@@
 
-1.  What is the width of the cline at a linked, neutral locus?
+1.  **How a neutral cline flattens:**  
+    What is the width of the cline at a linked, neutral locus?
     How long does it take to flatten out?
@@
 
-2.  How long are the haplotypes around the rare $B$ alleles on the $A$ side?
+2.  **How do haplotypes look around the selected locus in the cline:**
+    How long are the haplotypes around the rare $B$ alleles on the $A$ side?
     This should be constant in time.
 
-3.  How long, and how common, are chunks of $B$ haplotype *not* linked to $B$ alleles on the $A$ side?
+3.  **How does introgression happen:**
+    How long, and how common, are chunks of $B$ haplotype *not* linked to $B$ alleles on the $A$ side?
     These will accumulate as time progresses.
 
-4.  How does neighborhood size affect the results?
-    Does the mean cline, averaged across low-population-density simulations,
-    match the cline obtained by a high population density, i.e., the PDE?
-
-5.  What about the variance??
-
-6.  For a pair of sites, what is the correlation in ancestral state at those loci,
+4.  **What can we learn from linkage?**
+    For a pair of sites, what is the correlation in ancestral state at those loci,
     given the state at the selected locus
@@
         For each pair of loci, correlation of these two loci with extremal individuals.
+
+5.  How does neighborhood size affect the results?
+    Does the mean cline, averaged across low-population-density simulations,
+    match the cline obtained by a high population density, i.e., the PDE?
+
+6.  What about the variance??
 
@@
 a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
-b.  The selected cline takes time $s$ to spread out to its full width.
-c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2}{s\tau}}$ if $\tau>1/s$).
-d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.
+b.  The selected cline takes time $1/s$ to spread out to its full width.
+c.  Cline behaves like the selected cline for $r < 1/\tau$, and begin to flatten for $r>1/\tau$.
+d.  For large distances and times (but with $x<\sigma\sqrt{\tau}$),
+    the cline may stay steep out to
+    $r = \sqrt{\frac{2s}{\tau}}$ rather than $r=1/\tau$,
+    but this seems hard to actually observe.
 
-*Rationale:*
+**Rationale:**
 
@@
 b.  The selected cline spreads out (in forwards time!) with diffusion,
-    and so will reach maximum width $\sigma/\sqrt{s}$ in time $s$.
+    and so will reach maximum width $\sigma/\sqrt{s}$ in time $1/s$.
     Let $q(t,x)$ be the solution to
@@
     We know that the unitless cline $q$ will spread out in time 1;
-    so our cline will spread out in time $s$.
+    so our cline will spread out in time $1/s$.
 
@@
 
-c.  Look at space-recombination frequency images at a bunch of times
-    to check that flattening occurs around $r=\sqrt{2/s\tau}$.
+c.  To check that flattening occurs around $r=1/\tau$ (i.e., at $\tau=1/r$)
+    we'll look at when a neutral allele linked to the $A$ allele
+    near the center starts to get close to 1/2.
+
+    (*Technical note:* the system of ODE is [stiff](https://en.wikipedia.org/wiki/Stiff_equation), 
+    as diffusive systems [usually are](http://scicomp.stackexchange.com/questions/891/the-definition-of-stiff-ode-system),
+    particularly for small $r$ at the times when each set of probabilities becomes steep;
+    setting $s=0.1$ this is a problem, but not with smaller (less steep) $s$.)
+
+    ```{r space_recomb, cache=TRUE}
+    s <- 0.01
+    rvals <- c(0,exp(seq(log(.0025),log(0.05),length.out=8)))
+    tt <- seq(0,1000,length.out=100)
+    thegrid <- setup.grid.1D(x.up=-50, x.down=50, N = 2000)
+    fwds.soln <- forwards_pde(s=s,times=seq(0,1.2*max(tt),length.out=1000),grid=thegrid)
+    r.solns <- lapply(rvals, function(r) {
+            forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+        } )
+    linked.clines  <- lapply(r.solns, cline_from_soln, grid=thegrid )
+    xr <- lapply( seq_along(tt), function (k) { do.call( cbind, lapply( linked.clines, function (x) x[k,-1] ) ) } )
+    ```
+    The vertical lines are at $1/r$ (only on the righthand plots, due to the limitations of `filled.contour`):
+    ```{r plot_sr}
+    for (k in seq_along(rvals)) {
+        image( r.solns[[k]], legend=TRUE, zlim=c(0,1), yaxt='n', ylab='position', main=sprintf("r=%f, s=%f",rvals[k],s=s) )
+        axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
+        abline(v=1/rvals[k],lwd=2,lty=3)
+    }
+    ```
+
+d.  To look for the $r=1/\sqrt{\tau}$ behaviour, we could look at positions outside the selected cline 
+    ($x$ larger than $C/\sqrt{2s}$, for C=4, say) 
+    and times large enough that unlinked clines have started to flatten out there 
+    and that the selected cline has got to full width
+    ($\tau$ larger than $C^2/s$).
+    But: if $\tau=C^2/s$ then the critical value is $r=(s/C)^2$,
+    which at $s=0.01$ is $r=6.25\times10^{-6}$
+    **Moral:** the narrower cline due to linkage may be only visible only very nearby to the selected site.
+    *To-do:* figure out parameter values where this behaviour becomes visible: need to be much further away?
+
+<!--
+    ```{r long_time,cache=TRUE}
+    s <- 0.01
+    r <- 0.0001
+    tt <- seq(0,30000,length.out=500)
+    thegrid <- setup.grid.1D(x.up=-200, x.down=200, N = 2000)
+    fwds.soln <- forwards_pde(s=s,times=seq(0,1.2*max(tt),length.out=1000),grid=thegrid)
+    rev.soln <- forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+    linked.cline  <- cline_from_soln( rev.soln, grid=thegrid )
+    ```
+-->
+
+How do haplotypes look around the selected locus in the cline
+=============================================================
+
+2.  How long are the haplotypes around the rare $B$ alleles on the $A$ side?
+
+**Answer:**
+
+a.  The haplotype of a $B$ allele at distance $x$ into the $A$ side should be of length about $\sigma/(sx)$.
+
+**Rationale:**
+
+a.  This is (probably) the Laplace transform of the time since the $B$ allele
+    last had an ancestor on the $B$ side,
+    i.e., for a $B$ lineage to make it back to the $B$ side.
+    Since lineages on the wrong side move at speed $\sigma/s$,
+    if the allele is at position $x$, 
+    this will take about $sx/\sigma$;
+    and so the haplotype should be about $\sigma/(sx)$.

### 2015-11-03 notes/pde-clines.Rmd
c code?

@@
 library(ReacTran)
+library(Matrix)
 source("tran.1D.R") # fix for log.A
@@
 $$\begin{aligned}
-p(x,\tau,r) =  p(x,\tau) q(x,A,\tau,r) + p(-x,\tau) (1-q(-x,A,\tau,r))A .
+p(x,\tau,r) =  p(x,\tau) q(x,A,\tau,r) + p(-x,\tau) (1-q(-x,A,\tau,r)) .
 \end{aligned}$$
@@
 The smaller $r$ is, the slower the per-background probabilities smooth out:
-```{r small_r_sim}
+```{r small_r_sim,cache=TRUE}
 r <- 0.01
@@
 And, the cline flattens out *very* slowly:
-```{r longer_t_sim}
+```{r longer_t_sim,cache=TRUE}
 r <- 1.0
@@
 
-There are various haplotype-related statistics we could compute:
+Consider a genomic region $[a,b]$, with $a\le 0\le b$,
+with the selected site at 0.
+Recombination without coalescence makes the history of this region
+a branching process with conserved branching rates.
+Each branch is labeled by a region of the genome,
+and moves independently according to generator $G_r$, 
+where $r$ is the distance from the region to the selected site,
+and branches at rate $w$, where $w$ is its width.
+We would like to know the probability that 
+*all* branches end up on the same side of the barrier at time $\tau$ in the past.
 
-1.  Distance to closest recombination in a heterozygote
+Suppose that the region includes a type $A$ allele,
+and we want the probability that it inherits entirely from the $A$ side.
+Let
+$$\begin{aligned}
+    g(x,A,a,b,\tau) = \P^{x,A}\{ [a,b] \text{ inherits from $A$, $\tau$ generations ago }\} .
+\end{aligned}$$
+If $W(t)=\{W_1(t),W_2(t),\ldots,W_{n(t)}(t)\}$ is the state of the process $t$ generations ago,
+where $W_k=(x_k,z_k,a_k,b_k)$,
+with $x_k$ the position and $W_k\in\{A,B\}$ the allele,
+then by independence of branches,
+$$\begin{aligned}
+    g(x,A,a,b,0) &= 1 \\
+    g(x,B,a,b,0) &= 0 \\
+    g(x,B,a,b,\tau) &= 0 \qquad \text{if } 0 \in [a,b] \\
+    g(x,A,a,b,\tau)
+    &=
+    \E\left[
+        \prod_{i=1}^{n(t)} g(x_k,z_k,a_k,b_k,\tau-t)
+    \right] \qquad \text{for each } t \le \tau .
+\end{aligned}$$
+Taking $t = \epsilon$ in the last equation,
+so that the probability of branching is $1-e^{-(b-a)\epsilon}=\epsilon(b-a)+O(\epsilon^2)$,
+if $0\le a < b$ this is
+$$\begin{aligned}
+g(x,A,a,b,\tau)
+    &=
+    \epsilon 
+    \left\{
+    G_{r(a,b)} g(x,A,a,b,\tau)
+    +
+     \int_a^b g(x,A,a,\theta,\tau-\epsilon)
+     \left(
+     p(x,\tau-\epsilon)g(x,A,\theta,b,\tau-\epsilon) 
+     + (1-p(x,\tau-\epsilon))g(x,B,\theta,b,\tau-\epsilon) 
+     \right)d\theta
+    \right\}
+    + O(\epsilon^2),
+\end{aligned}$$
+where 
+$$\begin{aligned}
+r(a,b)=\begin{cases}
+    \min(|a|,|b|) \qquad &\text{if } ab \ge 0 \\
+    0 \qquad &\text{if } 0 \in [a,b] .
+\end{cases}
+\end{aligned}$$
+and $G_r f(x,z)$ is defined above.
+On the other hand, if $a<0<b$, then
+$$\begin{aligned}
+g(x,A,a,b,\tau)
+    &=
+    \epsilon 
+    \left\{
+    G_{r(a,b)} g(x,A,a,b,\tau)
+    +
+     \int_0^b g(x,A,a,\theta,\tau-\epsilon)
+     \left(
+     p(x,\tau-\epsilon)g(x,A,\theta,b,\tau-\epsilon)
+     + (1-p(x,\tau-\epsilon))g(x,B,\theta,b,\tau-\epsilon) 
+     \right)d\theta
+    +
+     \int_a^0 \left(
+     p(x,\tau-\epsilon)g(x,A,a,\theta,\tau-\epsilon)
+     + (1-p(x,\tau-\epsilon))g(x,B,a,\theta,\tau-\epsilon) 
+     \right) g(x,A,\theta,b,\tau-\epsilon) d\theta
+    \right\}
+    + O(\epsilon^2),
+\end{aligned}$$
+
+Therefore, assuming $G$ is sufficiently nice,
+$$\begin{aligned}
+\frac{d}{d\tau} g(x,z,a,b,\tau)
+&=
+G_{r(a,b)} g(x,z,a,b,\tau)
++
+\int_a^b
+\left(
+    g(x,z,a,\theta,\tau)g(x,z,\theta,b,\tau) 
+    - g(x,z,a,b,\tau)
+\right)
+d\theta
+\end{aligned}$$
+Since $z \in \{A,B\}$, this is a system of two integro-differential equations 
+in three variables plus time ($x$, $a$, $b$, and $\tau$; four if $x\in\R^2$).
+
+Numerical solutions will discretize; what discretization of the genome might be good?
+If we want to solve out to time $\tau$,
+then on a chromosome of length 1 there will be about $\tau$ breakpoints,
+each of which approximately Exponential($\tau$),
+and so the smallest distance will have mean $1/\tau^2$.
+But, suppose we've solved it for time $\tau$, and want to solve it to time $2\tau$.
+We can use the same discretization, with initial conditions from the first round.
+<!--
+Quick check:
+```{r smallest_check}
+f <- function (t) { min(diff(c(0,sort(runif(rpois(1,lambda=t))),1))) }
+fk <- sapply( 2:200, function (k) { replicate(n=100,f(k)) } )
+plot(1+col(fk),fk,col=rainbow(300)[col(fk)],log='y',ylab='smallest interval')
+points(2:200,colMeans(fk),pch=20,cex=2)
+lines(2:200,1/(2:200)^2,col='purple',lwd=2)
+```
+-->
+
+Note that if we think of $u(a,b)$ as a kernel, 
+then the map $u(a,b) \mapsto \int_a^b u(a,\theta) u(\theta,b) d\theta$
+is just the product of the kernel to itself;
+on discretization this becomes a matrix power,
+so the operator $u(a,b) \mapsto \int_a^b u(a,\theta)u(\theta,b) - u(a,b) d\theta$
+```{r solve_branching}
+xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=100)
+agrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=30)
+# determines the order the solution is in:
+xab <- subset( expand.grid( x=xgrid$x.mid, a=agrid$x.mid, b=agrid$x.mid ), ( b>0 ) & (abs(a)<=abs(b)) )
+find_ab <- function (ab) {
+    # map into the subset of (a,b) contained in xab
+    flip <- ab[,1]>ab[,2] 
+    ab[flip,] <- ab[flip,] %*% cbind( c(0,1), c(1,0) )
+    flop <- ( ab[,1]<0 & abs(ab[,1])>=abs(ab[,2]) )
+    ab[flop,] <- ab[flop,] %*% cbind( c(0,-1), c(-1,0) )
+    return(ab)
+}
+integral_map <- function (y,ab) {
+    # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
+    # The key step is to do something like:
+    #     1 | x y .
+    #   a 2 | . . u
+    #     3 | . . v
+    #       +------
+    #     b   1 2 3
+    # --> should  return x*u + y*v,
+    # the inner product of a partial row with a partial column
+
+}
+```
 
@@
 
-The first haplotype-related statistic
+A simpler statistic, that might approximate the correct quantity well,
 is the distance along the chromosome from an $A$ allele until the closest recombination
@@
 $$\begin{aligned}
-    u(x,\tau,r)=\P^(x,\tau)\{H>r\}
+    u(x,\tau,r)=\P^{(x,\tau)}\{H>r\}
 \end{aligned}$$

### 2015-11-03 notes/test-unifprod.R

@@ -1,0 +1,14 @@
+dyn.load("unifprod.so")
+unifprod <- function (x,y) { 
+    .C("unifprod",
+            x=as.numeric(x),
+            nrx=NROW(x),
+            ncx=NCOL(x),
+            y=as.numeric(y),
+            nry=NROW(y),
+            ncy=NCOL(y),
+            z=matrix(0.0,nrow=NROW(x),ncol=NCOL(y))
+        )[[7]] 
+}
+unifprod(diag(3)*1.0,diag(3)*1.0)
+

### 2015-11-03 notes/unifprod.c

@@ -1,0 +1,89 @@
+#include <R.h>
+#include <Rinternals.h>
+
+/* modified from ./src/main/array.c in R-3.2.2 source */
+
+/*
+ dyn.load("unifprod.so")
+ unifprod <- function (x,y) { .C("unifprod",x=as.numeric(x),nrx=NROW(x),ncx=NCOL(x),y=as.numeric(y),nry=NROW(y),ncy=NCOL(y),z=matrix(0.0,nrow=NROW(x),ncol=NCOL(y)))[[7]] }
+ unifprod(diag(3)*1.0,diag(3)*1.0)
+ */
+
+void unifprod(double *x, int *nrx, int *ncx,
+                    double *y, int *nry, int *ncy, double *z)
+{
+    /* compute the matrix product (A,B) -> C where
+     *   C[i,j] = A[i]*B[i+1] + A[i+1]*B[i+2] + ... + A[j-1]*B[j]
+     * and so if A is (m x n) and B is (n x o) then C is (m x o) and
+     *   C[i,i] = 0
+     */
+    double sum;
+    R_xlen_t NRX = *nrx, NRY = *nry;
+    R_xlen_t NCX = *ncx, NCY = *ncy;
+
+    if (NRX > 0 && NCX > 0 && NRY > 0 && NCY > 0) {
+        for (int i = 0; i < NRX; i++)
+        for (int k = 0; k < NCY; k++) {
+            sum = 0.0;
+            for (int j = i; j < NCX-1; j++) {
+                sum += x[i + j * NRX] * y[j + 1 + k * NRY];
+            }
+            z[i + k * NRX] = (double) sum;
+        }
+    } else { /* zero-extent operations should return zeroes */
+        for(R_xlen_t i = 0; i < NRX*NCY; i++) z[i] = 0;
+    }
+}
+
+
+// // "ORIGINAL"
+// static void unifprod(double *x, int nrx, int ncx,
+//                     double *y, int nry, int ncy, double *z)
+// {
+//     /* compute the matrix product (A,B) -> C where
+//      *   C[i,j] = A[i]*B[i+1] + A[i+1]*B[i+2] + ... + A[j-1]*B[j]
+//      * and so if A is (m x n) and B is (n x o) then C is (m x o) and
+//      *   C[i,i] = 0
+//      */
+//     double sum;
+//     R_xlen_t NRX = nrx, NRY = nry;
+// 
+//     if (nrx > 0 && ncx > 0 && nry > 0 && ncy > 0) {
[…]

### 2015-11-04 notes/pde-clines.Rmd

@@
 then the map $u(a,b) \mapsto \int_a^b u(a,\theta) u(\theta,b) d\theta$
-is just the product of the kernel to itself;
-on discretization this becomes a matrix power,
-so the operator $u(a,b) \mapsto \int_a^b u(a,\theta)u(\theta,b) - u(a,b) d\theta$
+is similar to the product of the kernel to itself;
+on discretization this becomes like a matrix power.
+This is implemented in some C code.
 ```{r solve_branching}
+source("unifprod.R")  # for the C part
+integral_map <- function (x,eps) {
+    # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
+    # x is a square matrix,
+    #   eps is the grid spacing between locations
+    x2 <- eps*unifprod(x)
+}
 xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=100)
 agrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=30)
+# we'll save the solution as an array,
+#  with dimensions (x,a,b)
+# Note we only use the upper triangle a<=b, but oh well.
+soln.array <- array(0.0,dim=c(xgrid$N,agrid$N,agrid$N))
+
 # determines the order the solution is in:
@@
     return(ab)
-}
-integral_map <- function (y,ab) {
-    # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
-    # The key step is to do something like:
-    #     1 | x y .
-    #   a 2 | . . u
-    #     3 | . . v
-    #       +------
-    #     b   1 2 3
-    # --> should  return x*u + y*v,
-    # the inner product of a partial row with a partial column
-
 }

### 2015-11-06 notes/unifprod.R
upper tri unifprod

@@ -35,36 +35,76 @@ unifprod.cpp <- 'NumericMatrix unifprod(NumericMatrix x) {
     return z;
 }'
 
+unifprod.ut.cpp <- 'NumericVector unifprod_ut(NumericVector y, double ncx) {
+    NumericVector z(y.size());
+    int index = 0;
+    int pindex, outindex;
+    for (int i=0; i<y.size(); i++) {
+        z[i] = 0.0;
+    }
+    for (int j=1; j<=ncx; j++) {
+        for (int i=1; i<=j; i++) {
+            pindex = (j+1)*(j+2)/2 - 1;
+            outindex = index + j;
+            z[outindex-j] -= (j-i)*y[index];
+            // printf("i=%i, j=%i, index=%i, yi=%f \\n",i,j,index,y[index]);
+            if ( (i<ncx) && (j<ncx) ) {
+                for (int k=j+1; k<=ncx; k++) {
+                    // printf("i=%i, j=%i, k=%i, index=%i, pindex=%i, outindex=%i yy=%f \\n",i,j,k,index+1,pindex+1,outindex+1,y[index]*y[pindex]);
+                    z[outindex] += y[index]*y[pindex];
+                    pindex += k;
+                    outindex += k;
+                }
+            }
+            index++;
+        }
+    }
+    return z;
+}'
+
+
+
 cppFunction(unifprod.cpp)
+cppFunction(unifprod.ut.cpp)
 
 
 if (FALSE) { ## testing
-    unifprod_R <- function (x) {
+
+    unifprod_R <- function (x,subtract=TRUE) {
         z <- matrix(0,nrow=NROW(x),ncol=NCOL(x))
         stopifnot( NCOL(x) == NROW(x) )
         for (i in seq(1,length.out=nrow(z)-1)) {
             for (k in seq(i+1,ncol(z))) {
-                z[i,k] <- sum(x[i,seq(i,k-1)]*x[seq(i+1,k),k]-x[i,k])
+                if (subtract) {
+                    z[i,k] <- sum(x[i,seq(i,k-1)]*x[seq(i+1,k),k]-x[i,k])
+                } else {
+                    z[i,k] <- sum(x[i,seq(i,k-1)]*x[seq(i+1,k),k])
+                }
             }
         }
         return(z)
     }
 
-    unifprod_ut_R <- function (x,ncx) {
-        # same, but for upper triangular ordered x
-        z <- matrix(0,nrow=ncx,ncol=ncx)
-        index <- 0  # where are we in the vector
[…]

### 2015-11-06 notes/unifprod.c

@@ -25,10 +25,48 @@ SEXP unifprod2(SEXP x, SEXP Rnrx, SEXP Rncx)
                 sum += X[i + j * nrx] * X[j + 1 + k * nrx] - xik;
             }
             REAL(z)[i + k * nrx] = (double) sum;
-            REAL(z)[i + k * nrx] = (double) sum;
         }
     } else { /* zero-extent operations should return zeroes */
         for(int i = 0; i < nrx*ncx; i++) REAL(z)[i] = 0.0;
+    }
+    UNPROTECT(1);
+    return z;
+}
+
+
+SEXP unifprod_ut2(SEXP y, SEXP Rncx)
+{
+    /* compute the matrix transformation A -> C where
+     *   C[i,j] = A[i,i]*A[i+1,j] + A[i,i+1]*A[i+2,j] + ... + A[i,j-1]*A[j,j] - (j-i) A[i,j]
+     *   C[i,i] = 0
+     */
+    int ncx = INTEGER(Rncx)[0];
+    double *Y = REAL(y);
+    int ylen = ncx*(ncx+1)/2;
+    SEXP z;
+    PROTECT(z = allocVector(REALSXP, ylen));
+
+    int index = 0;
+    int pindex, outindex;
+    for (int i=0; i<ylen; i++) {
+        REAL(z)[i] = 0.0;
+    }
+    for (int j=1; j<=ncx; j++) {
+        for (int i=1; i<=j; i++) {
+            pindex = (j+1)*(j+2)/2 - 1;
+            outindex = index + j;
+            REAL(z)[outindex-j] -= (j-i)*Y[index];
+            // printf("i=%i, j=%i, index=%i, yi=%f \\n",i,j,index,Y[index]);
+            if ( (i<ncx) && (j<ncx) ) {
+                for (int k=j+1; k<=ncx; k++) {
+                    // printf("i=%i, j=%i, k=%i, index=%i, pindex=%i, outindex=%i yy=%f \\n",i,j,k,index+1,pindex+1,outindex+1,Y[index]*Y[pindex]);
+                    REAL(z)[outindex] += Y[index]*Y[pindex];
+                    pindex += k;
+                    outindex += k;
+                }
+            }
+            index++;
+        }
     }
     UNPROTECT(1);
     return z;

### 2015-11-06 notes/Schematic_figure.Rmd

---
title: "Thinking about ancestry in real data"
date: "`r date()`"
---

The Model
========
This is to make a figure of the ancestry of a haplotype. Recombinations within h haplotype fragment is depicted as a branching event, producing two smaller haplotype fragments. Recombinations between the fragment and selected locus can occur along the branch, and can result in a change in which selected locus the fragment is linked to. All recombinations are depicted as red dots.

If a fragment is colored blue, it is linked to allele B (ancestrally RHS), and orange if linked to A. If a fragment is linked to ancestry B, it moves with drift D if it is on the RHS, and with no drift otherwise. The reverse is true for a fragment linked to ancestry A. A recombination event will put a fragment on ancestry A if it happens on the LHS, and on B if on the RHS of the zone center.

```{r plot}
split_times = c(15,50,70)
recombinations = c(30,60,90,70,100)

make.path = function(LENGTH,START,BACKGROUND,D=0.5){ #BACKGROUND = 1 for LHS ancestry
	path = vector(length=LENGTH)
	path[1] = START

	for(i in 2:LENGTH){
		if ((path[i-1]<0 & BACKGROUND>0) | (path[i-1]>0 & BACKGROUND<0)){path[i]=rnorm(1,BACKGROUND*D,0.5)+path[i-1] } else{path[i] = rnorm(1,0,0.5)+path[i-1]}
	}

	return(path)
}

plot.path = function(START_TIME, STOP_TIME,START_POS){
	path = make.path(STOP_TIME-START_TIME+1,START_POS,sign(START_POS))
	points(path,START_TIME:STOP_TIME,type="l",col=ifelse(START_POS<0,"orange","blue"),lwd=1.2)
	points(START_POS,START_TIME,pch=19,col="red")
	return(path)
}

backbone = make.path(100,-5,1)
plot(backbone,1:100,type="l",xlim=c(-10,10))

branch1 = plot.path(split_times[1],recombinations[1],backbone[split_times[1]])
rec1.1 = plot.path(recombinations[1],recombinations[2],branch1[length(branch1)])
rec1.2 = plot.path(recombinations[2],recombinations[3],rec1.1[length(rec1.1)])
rec1.3 = plot.path(recombinations[3],recombinations[5],rec1.2[length(rec1.2)])

branch2 = plot.path(split_times[2],recombinations[4],backbone[split_times[2]])
rec2.1 = plot.path(recombinations[4],recombinations[5],branch2[length(branch2)])

abline(v=0,lty=2,col="grey",lwd=0.8)
legend("bottomright",legend = c("selected locus","linked to A", "linked to B","rec. event"),col=c("black","orange","blue","red"),lty=c(rep(1,3),NA),pch=c(rep(NA,3),19))

```
In practice, the drift parameter, and probability of recombining on a particular ancestry is a function in space of the selected allele frequency.

### 2015-11-06 notes/pde-clines.Rmd
minor changes

@@
 
-Consider a genomic region $[a,b]$, with $a\le 0\le b$,
+Consider a genomic region $[a,b]$,
 with the selected site at 0.
@@
 
-Therefore, assuming $G$ is sufficiently nice,
+Therefore, assuming $g$ is sufficiently nice,
 $$\begin{aligned}
@@
 +
-\int_a^b
-\left(
-    g(x,z,a,\theta,\tau)g(x,z,\theta,b,\tau) 
-    - g(x,z,a,b,\tau)
-\right)
-d\theta
+I g(x,z,a,b,\tau),
 \end{aligned}$$
+where $I$ is the integral operator defined above.
 Since $z \in \{A,B\}$, this is a system of two integro-differential equations 

### 2015-11-11 notes/pde-clines.Rmd
fix to pde

--- 2015-11-04 notes/pde-clines.Rmd
@@
 so that the probability of branching is $1-e^{-(b-a)\epsilon}=\epsilon(b-a)+O(\epsilon^2)$,
-if $0\le a < b$ this is
+if $0\le a < b$ this implies that
 $$\begin{aligned}
+\frac{d}{d\tau} 
 g(x,A,a,b,\tau)
     &=
-    \epsilon 
-    \left\{
     G_{r(a,b)} g(x,A,a,b,\tau)
     +
-     \int_a^b g(x,A,a,\theta,\tau-\epsilon)
+     \int_a^b g(x,A,a,\theta,\tau)
      \left(
-     p(x,\tau-\epsilon)g(x,A,\theta,b,\tau-\epsilon) 
-     + (1-p(x,\tau-\epsilon))g(x,B,\theta,b,\tau-\epsilon) 
+     p(x,\tau)g(x,A,\theta,b,\tau) 
+     + (1-p(x,\tau))g(x,B,\theta,b,\tau) 
      \right)d\theta
-    \right\}
-    + O(\epsilon^2),
 \end{aligned}$$
@@
 $$\begin{aligned}
+\frac{d}{d\tau} 
 g(x,A,a,b,\tau)
     &=
-    \epsilon 
-    \left\{
     G_{r(a,b)} g(x,A,a,b,\tau)
     +
-     \int_0^b g(x,A,a,\theta,\tau-\epsilon)
+     \int_0^b g(x,A,a,\theta,\tau)
      \left(
-     p(x,\tau-\epsilon)g(x,A,\theta,b,\tau-\epsilon)
-     + (1-p(x,\tau-\epsilon))g(x,B,\theta,b,\tau-\epsilon) 
+     p(x,\tau)g(x,A,\theta,b,\tau)
+     + (1-p(x,\tau))g(x,B,\theta,b,\tau) 
      \right)d\theta
@@
      \int_a^0 \left(
-     p(x,\tau-\epsilon)g(x,A,a,\theta,\tau-\epsilon)
-     + (1-p(x,\tau-\epsilon))g(x,B,a,\theta,\tau-\epsilon) 
-     \right) g(x,A,\theta,b,\tau-\epsilon) d\theta
-    \right\}
-    + O(\epsilon^2),
+     p(x,\tau)g(x,A,a,\theta,\tau)
+     + (1-p(x,\tau))g(x,B,a,\theta,\tau) 
+     \right) g(x,A,\theta,b,\tau) d\theta
 \end{aligned}$$
 
+<!--
 Therefore, assuming $G$ is sufficiently nice,
@@
 \end{aligned}$$
+-->
 Since $z \in \{A,B\}$, this is a system of two integro-differential equations 

### 2015-11-11 notes/pde-clines.Rmd

@@
 
-<!--
-Therefore, assuming $G$ is sufficiently nice,
+Therefore, assuming $g$ is sufficiently nice,
 $$\begin{aligned}
@@
 +
-\int_a^b
-\left(
-    g(x,z,a,\theta,\tau)g(x,z,\theta,b,\tau) 
-    - g(x,z,a,b,\tau)
-\right)
-d\theta
+I g(x,z,a,b,\tau),
 \end{aligned}$$
--->
+where $I$ is the integral operator defined above.
 Since $z \in \{A,B\}$, this is a system of two integro-differential equations 

### 2015-11-11 notes/rough-calculations.Rmd

@@
 
-a.  The haplotype of a $B$ allele at distance $x$ into the $A$ side should be of length about $\sigma/(sx)$.
+a.  The haplotype of a $B$ allele at distance $x$ into the $A$ side should be of length about $\sigma/(sx)$, as long as this is less than $1/\tau$. (this is fairly restrictive)
 
@@
     and so the haplotype should be about $\sigma/(sx)$.
+    This gives the size of the chunk carried with it on this excursion;
+    this will be of the same type with high probability at least if the length is less than $1/\tau$,
+    as above.

### 2015-11-17 drafts/hybrid_zone.tex
took a stab at writing some intro

@@
 \section{Introduction}
--Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation.
+The process of speciation often involves allopatric divergence in relative isolation.
+When formerly isolated populations come back into contact, a hybrid zone can form, and may be maintained stably by selection.
+In such instances, selection on some subset of the genome prevents mixing of genotypes between species.
+We refer to such hybrid zones as tension zones (as opposed to contact zones which reflect a non-equilibrium gradient in ancestry that will eventually be eroded by migration throughout the whole genome).
+Because the population genetic signatures around loci experiencing selection is going to differ from other regions of the genome, natural hybrid zones present exciting opportunities to understand species maintenance and the process of speciation.\\
+
+Gene-flow across a hybrid zone is impeded by selection against migrant, or hybrid, genotypes, and as a result, produces a cline in ancestry that is expected to be steepest at and nearby the selected locus.
+Elsewhere in the genome, clines in ancestry are wider, and are more heavily influenced by hybrid zone age and rates of migration.
+Historically, this has been the most relied-upon population genetic signature, and it is only recently that abundant genomic data has opened the possibility of studying patterns of linkage within hybrid zones.
+For example, patterns of linkage disequilibrium (LD) or tract length distribution across geographic space can provide information about zone age \cite{Sedghifar2015}.
+In tensions zones geographic patterns of linkage at loci neighboring targets of selection can potentially be a powerful source of information.\\
+
+Many of the traditional models of genetic variation in hybrid zones to date have focused on the equilibrium state of hybrid zones, at both selected and linked loci (lots of Barton).
+
+%-Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation.
 Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone).
 \\
--Gene flow across a hybrid zone may be impeded by selection against migrants.
+%-Gene flow across a hybrid zone may be impeded by selection against migrants.
 In older contact zones, loci that are associated with reduced fitness in heterozygotes will have steeper clines than the rest of the genome.
@@
 small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)\\
--There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
+%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
 -We consider the non-equilibrium state of hybrid zones, and patterns of linkage around selected loci.

### 2015-11-21 notes/pde-clines.Rmd
bivariate integral transform

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\one}{\mathbf{1}}
 \DeclareMathOperator{\sgn}{sgn}
@@
     g(x,B,a,b,0) &= 0 \\
-    g(x,B,a,b,\tau) &= 0 \qquad \text{if } 0 \in [a,b] \\
     g(x,A,a,b,\tau)
@@
      \right)d\theta
+     -(b-a)g(x,A,a,b,\tau)
 \end{aligned}$$
@@
      + (1-p(x,\tau))g(x,B,\theta,b,\tau) 
-     \right)d\theta
+     \right)d\theta \\
+   &\qquad {}
     +
@@
      + (1-p(x,\tau))g(x,B,a,\theta,\tau) 
-     \right) g(x,A,\theta,b,\tau) d\theta
+     \right) g(x,A,\theta,b,\tau) d\theta  \\
+   &\qquad {}
+     -(b-a)g(x,A,a,b,\tau)
 \end{aligned}$$
+
+Writing $g_A$ and $g_B$ for the moment, and dropping $x$ and $\tau$,
+the generator of the recombination part is, 
+where $x\wedge y=\max(x,y)$ and $x \vee y = \min(x,y)$,
+$$\begin{aligned}
+Ig_A(a,b)
+    &= 
+        p \int_{a\wedge 0}^b g_A(a,\theta) g_A(\theta,b) d\theta
+        + (1-p) \int_{a\wedge 0}^b g_A(a,\theta) g_B(\theta,b) d\theta
+        + p \int_{a}^{b \vee 0} g_A(a,\theta) g_A(\theta,b) d\theta
+        + (1-p) \int_{a\wedge 0}^b g_B(a,\theta) g_A(\theta,b) d\theta 
+       \\ &\qquad {}
+         -(b-a)g_A(a,b) \\
+    &= 
+    \int_a^b
+        p \one_{\theta>0} g_A(a,\theta) g_A(\theta,b) 
+        + (1-p) \one_{\theta>0} g_A(a,\theta) g_B(\theta,b) 
+        + p \one_{\theta\le0} g_A(a,\theta) g_A(\theta,b) 
+        + (1-p) \one_{\theta\le0} g_B(a,\theta) g_A(\theta,b) d\theta
+       \\ &\qquad {}
+         -(b-a)g_A(a,b) \\
+    &= 
+    \int_a^b
+    \left\{
+        p g_A(a,\theta) g_A(\theta,b) 
+        + (1-p) \left( 
+            \one_{\theta>0} g_A(a,\theta) g_B(\theta,b) 
+            + \one_{\theta\le0} g_B(a,\theta) g_A(\theta,b) 
+        \right) 
+        -g_A(a,b) \right\}
+        d\theta   \\
+    &= 
+    p \int_a^b
+        \left( g_A(a,\theta) g_A(\theta,b)  - g_A(a,b) \right) d\theta 
+    \\ &\qquad {}
+    + (1-p) \int_a^b \one_{\theta>0} \left( g_A(a,\theta) g_B(\theta,b) - g_A(a,b) \right) d\theta
+    \\ &\qquad {}
+    + (1-p) \int_a^b \one_{\theta\le0} \left( g_B(a,\theta) g_A(\theta,b) - g_A(a,b) \right) d\theta  .
+\end{aligned}$$
+$Ig_B$ is the same after exchanging $A$ with $B$, and $p$ with $1-p$.
 
@@
 in three variables plus time ($x$, $a$, $b$, and $\tau$; four if $x\in\R^2$).
+
+*Note:* We also know that
+$$\begin{aligned}
+    g(x,B,a,b,\tau) &= 0 \qquad \text{if } 0 \in [a,b] \\
+\end{aligned}$$
+but this will have to be arranged by the cline function.
 
@@
 This is implemented in some C code.
+<!--
 ```{r solve_branching}
 source("unifprod.R")  # for the C part
-integral_map <- function (x,eps) {
+recomb_generator <- function (x,eps) {
     # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
-    # x is a square matrix,
+    # x is the upper triangular part of a square matrix
     #   eps is the grid spacing between locations
-    x2 <- eps*unifprod(x)
+    x2 <- eps*unifprod_ut(x)
 }
-xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=100)
-agrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=30)
+xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=10)
+agrid <- setup.grid.1D(x.up=0.0, x.down=0.5, N=5)
+# function to find a,b coordinates from the upper triangular coordinates
+ab.ord <- matrix( 0, nrow=agrid$N, ncol=agrid$N )
+avals <- agrid$x.mid[row(ab.ord)[upper.tri(ab.ord,diag=TRUE)]]
+bvals <- agrid$x.mid[col(ab.ord)[upper.tri(ab.ord,diag=TRUE)]]
+rvals <- ifelse( avals*bvals<0, 0, pmin(avals,bvals) )
+ab.coord <- function (ab) { cbind( a=avals[ab], b=bvals[ab] ) }
 # we'll save the solution as an array,
-#  with dimensions (x,a,b)
-# Note we only use the upper triangle a<=b, but oh well.
-soln.array <- array(0.0,dim=c(xgrid$N,agrid$N,agrid$N))
+#  with dimensions (x,ab)
+# Here ab is upper-triangular ordered
+
+s <- 0.1
+tt <- seq(0,3,length.out=5)
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+
+soln.array <- matrix(0.0,nrow=2*xgrid$N,ncol=agrid$N*(agrid$N+1)/2)
+
+    sigma=1
+    yinit=rep( c( rep(1.0,xgrid$N), rep(0.0,xgrid$N) ), ncol(soln.array) )
+
+    eps <- 1e-16
+    rev.pde.fn <- function (t,y,parms,...) {
+        # need to apply tran.1D to columns
+        # and recomb_generator to rows
+        dim(y) <- dim(soln.array)
+        p <- cline_interp(t,soln=fwds.soln) # freq of B
+        log.p <- log(pmax(p,min(eps,min(p[p>0]))))
+        log.1mp <- log(pmax(1-p,min(eps,min((1-p)[p<1]))))
+        diffusion <- sapply( 1:ncol(y), function (kcol) {
+                r <- rvals[kcol]
+                yA <- y[1:xgrid$N,kcol]
+                yB <- y[xgrid$N+(1:xgrid$N),kcol]
+                tran.A <- tran.1D(C=yA, A=log.p, D=sigma^2/2, dx=xgrid, flux.up=0, flux.down=0, log.A=TRUE)$dC
+                tran.B <- tran.1D(C=yB, A=log.1mp, D=sigma^2/2, dx=xgrid, flux.up=0, flux.down=0, log.A=TRUE)$dC
+                c( 
+                        tran.A + r * (1-p) * (yB-yA),
+                        tran.B + r * p * (yA-yB)
+                        )
+             } )
+        recombination <- apply( y, 1, recomb_generator, p=p )
+        return( list( diffusion + recombination ) )
+    }
+    rev.soln <- ode.1D( y=yinit, times=times, func=rev.pde.fn, nspec=2*ncol(soln.array), tcrit=max(fwds.soln[,1]) )
 
@@
 ```
+
+-->
 

### 2015-11-21 notes/biprod_ut.cpp

@@ -1,0 +1,36 @@
+NumericVector biprod_ut(
+        NumericVector yA, 
+        NumericVector yB,
+        int zeroind,
+        double ncx) {
+    NumericVector z(yA.size());
+    int index = 0;
+    int pindex, outindex;
+    for (int i=0; i<yA.size(); i++) {
+        z[i] = 0.0;
+    }
+    for (int j=1; j<=ncx; j++) {
+        for (int i=1; i<=j; i++) {
+            pindex = (j+1)*(j+2)/2 - 1;
+            outindex = index + j;
+            z[outindex-j] -= (j-i)*yA[index];
+            // printf("i=%i, j=%i, index=%i, yi=%f \\n",i,j,index,yA[index]);
+            if ( (i<ncx) && (j<ncx) ) {
+                for (int k=j+1; k<=ncx; k++) {
+                    // printf("i=%i, j=%i, k=%i, index=%i, pindex=%i, outindex=%i yy=%f \\n",i,j,k,index+1,pindex+1,outindex+1,yA[index]*yA[pindex]);
+                    if ( j < zeroind ) {
+                        z[outindex] += yB[index]*yA[pindex];
+                    } else {
+                        z[outindex] += yA[index]*yB[pindex];
+                    }
+                    pindex += k;
+                    outindex += k;
+                }
+            }
+            index++;
+        }
+    }
+    return z;
+}
+
+

### 2015-11-21 notes/tests/test-unifprod.R

@@ -48,6 +48,70 @@ source("../unifprod.R",chdir=TRUE)
         return(z)
     }
 
+    biprod_R <- function (xA,xB,zeroind,subtract=TRUE) {
+        z <- matrix(0,nrow=NROW(x),ncol=NCOL(x))
+        stopifnot( NCOL(x) == NROW(x) )
+        for (i in seq(1,length.out=nrow(z)-1)) {
+            for (k in seq(i+1,ncol(z))) {
+                j1 <- seq(i,k-1)
+                j2 <- j1+1
+                if (i>=zeroind) {  # both to the right of zero
+                    xy <- xA[i,j1]*xB[j2,k]
+                } else if (k<=zeroind) {  # both to the left of zero
+                    xy <- xB[i,j1]*xA[j2,k]
+                } else {  # i <= 0 <= k
+                    xy <- ifelse(j1>=zeroind, xA[i,j1]*xB[j2,k], xB[i,j1]*xA[j2,k] )
+                }
+                if (subtract) {
+                    z[i,k] <- sum(xy-xA[i,k])
+                } else {
+                    z[i,k] <- sum(xy)
+                }
+            }
+        }
+        return(z)
+    }
+
+
+    biprod_ut_R <- function (yA,yB,zeroind,ncx,subtract=TRUE) {
+        # discretization of  y -> z where
+        #   z[a,b] = \int_a^0 ( yA(a,x) yB(x,b) - yA(a,b) dx + \int_0^b ( yB(a,x) yA(x,b) - yA(a,b) dx 
+        # to discretize this we sum over recombinations occuring between marker (k-1) and marker k
+        #   and zeroind is the index of the marker at zero
+        #   so whichever interval has zeroind in it should be yA
+        z <- numeric(length(yA))
+        index <- 0    # where are we in the input vector
+        for (j in seq(1,ncx)) {    # column
+            for (i in seq(1,j)) {  # row
+                index <- index + 1
+                pindex <- (j+1)*(j+2)/2    # index of first partner at [j+1,j+1] in product
+                outindex <- index + j     # where we are in the output (also equal to j*(j+1)/2 + i)
+                if (subtract) { 
+                    # cat("right: i=", i, ", j=", j, ", index=", index, "subtr=", (j-i)*yA[index], "\n")
[…]

### 2015-11-21 notes/unifprod.R

@@ -1,7 +1,9 @@
 library(Rcpp)
 
+# see test-unifprod.R for more details.
+
 # Discretization of u(a,b) -> \int_a^b (u(a,t) u(t,b) - u(a,b)) dt
-#  so output is y[i,k] = sum( x[i,i:(k-1)]*x[(i+1):k,k] )
+#  so output is y[i,k] = sum( x[i,i:(k-1)]*x[(i+1):k,k] ) - (k-1)*x[i,k]
 #
 # The key step is to do something like:
 #     1 | x y z
@@ -9,8 +11,8 @@ library(Rcpp)
 #     3 | . . v
 #       +------
 #     b   1 2 3
-# --> should  return x*u + y*v
-# the inner product of a partial row with a partial column
+# --> [1,3]th entry of output is x*u + y*v - 2*z
+# the first part is the inner product of a partial row with a partial column
 
 # makes unifprod(), which works on matrices
 unifprod.cpp <- paste(readLines("unifprod.cpp"),collapse="\n")
@@ -19,9 +21,12 @@ unifprod.cpp <- paste(readLines("unifprod.cpp"),collapse="\n")
 # upper-triangular portions of matrices, with diagonals
 unifprod.ut.cpp <- paste(readLines("unifprod_ut.cpp"),collapse="\n")
 
+# and does the two-argument version; see tests/test-unifprod.R.
+biprod.ut.cpp <- paste(readLines("biprod_ut.cpp"),collapse="\n")
 
 
 cppFunction(unifprod.cpp)
 cppFunction(unifprod.ut.cpp)
+cppFunction(biprod.ut.cpp)
 
 

### 2015-11-22 notes/pde-clines.Rmd
solving haplotypes

@@
 library(Matrix)
+library(parallel)
 source("tran.1D.R") # fix for log.A
@@
         -g_A(a,b) \right\}
-        d\theta   \\
-    &= 
-    p \int_a^b
-        \left( g_A(a,\theta) g_A(\theta,b)  - g_A(a,b) \right) d\theta 
-    \\ &\qquad {}
-    + (1-p) \int_a^b \one_{\theta>0} \left( g_A(a,\theta) g_B(\theta,b) - g_A(a,b) \right) d\theta
-    \\ &\qquad {}
-    + (1-p) \int_a^b \one_{\theta\le0} \left( g_B(a,\theta) g_A(\theta,b) - g_A(a,b) \right) d\theta  .
+        d\theta   
 \end{aligned}$$
@@
 This is implemented in some C code.
-<!--
-```{r solve_branching}
-source("unifprod.R")  # for the C part
-recomb_generator <- function (x,eps) {
-    # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
-    # x is the upper triangular part of a square matrix
-    #   eps is the grid spacing between locations
-    x2 <- eps*unifprod_ut(x)
-}
-xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=10)
-agrid <- setup.grid.1D(x.up=0.0, x.down=0.5, N=5)
-# function to find a,b coordinates from the upper triangular coordinates
-ab.ord <- matrix( 0, nrow=agrid$N, ncol=agrid$N )
-avals <- agrid$x.mid[row(ab.ord)[upper.tri(ab.ord,diag=TRUE)]]
-bvals <- agrid$x.mid[col(ab.ord)[upper.tri(ab.ord,diag=TRUE)]]
-rvals <- ifelse( avals*bvals<0, 0, pmin(avals,bvals) )
-ab.coord <- function (ab) { cbind( a=avals[ab], b=bvals[ab] ) }
-# we'll save the solution as an array,
-#  with dimensions (x,ab)
-# Here ab is upper-triangular ordered
+```{r solve_branching, cache=TRUE}
+recomb_generator
 
 s <- 0.1
-tt <- seq(0,3,length.out=5)
-fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+tt <- seq(0,10,length.out=20)
 
-soln.array <- matrix(0.0,nrow=2*xgrid$N,ncol=agrid$N*(agrid$N+1)/2)
+xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=10)
+rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=50)
 
-    sigma=1
-    yinit=rep( c( rep(1.0,xgrid$N), rep(0.0,xgrid$N) ), ncol(soln.array) )
-
-    eps <- 1e-16
-    rev.pde.fn <- function (t,y,parms,...) {
-        # need to apply tran.1D to columns
-        # and recomb_generator to rows
-        dim(y) <- dim(soln.array)
-        p <- cline_interp(t,soln=fwds.soln) # freq of B
-        log.p <- log(pmax(p,min(eps,min(p[p>0]))))
-        log.1mp <- log(pmax(1-p,min(eps,min((1-p)[p<1]))))
-        diffusion <- sapply( 1:ncol(y), function (kcol) {
-                r <- rvals[kcol]
-                yA <- y[1:xgrid$N,kcol]
-                yB <- y[xgrid$N+(1:xgrid$N),kcol]
-                tran.A <- tran.1D(C=yA, A=log.p, D=sigma^2/2, dx=xgrid, flux.up=0, flux.down=0, log.A=TRUE)$dC
-                tran.B <- tran.1D(C=yB, A=log.1mp, D=sigma^2/2, dx=xgrid, flux.up=0, flux.down=0, log.A=TRUE)$dC
-                c( 
-                        tran.A + r * (1-p) * (yB-yA),
-                        tran.B + r * p * (yA-yB)
-                        )
-             } )
-        recombination <- apply( y, 1, recomb_generator, p=p )
-        return( list( diffusion + recombination ) )
-    }
-    rev.soln <- ode.1D( y=yinit, times=times, func=rev.pde.fn, nspec=2*ncol(soln.array), tcrit=max(fwds.soln[,1]) )
-
-# determines the order the solution is in:
-xab <- subset( expand.grid( x=xgrid$x.mid, a=agrid$x.mid, b=agrid$x.mid ), ( b>0 ) & (abs(a)<=abs(b)) )
-find_ab <- function (ab) {
-    # map into the subset of (a,b) contained in xab
-    flip <- ab[,1]>ab[,2] 
-    ab[flip,] <- ab[flip,] %*% cbind( c(0,1), c(1,0) )
-    flop <- ( ab[,1]<0 & abs(ab[,1])>=abs(ab[,2]) )
-    ab[flop,] <- ab[flop,] %*% cbind( c(0,-1), c(-1,0) )
-    return(ab)
-}
+# took 24min with xgrid$N=10, rgrid$N=20 and 50 time steps up to t=3, :(
+hap.soln <- forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=rgrid)
 ```
 
--->
+```{r plot_hap_probs, cache=TRUE, depends="solve_branching"}
+hap.probs <- get_hap_probs( hap.soln, left=-.025, right=.025 )
+image( hap.probs, zlim=c(0,1), legend=TRUE, yaxt='n', 
+      main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+```
 

### 2015-11-22 notes/biprod_ut.cpp

@@ -2,7 +2,7 @@ NumericVector biprod_ut(
         NumericVector yA, 
         NumericVector yB,
         int zeroind,
-        double ncx) {
+        int ncx) {
     NumericVector z(yA.size());
     int index = 0;
     int pindex, outindex;

### 2015-11-22 notes/cline-fns.R

@@ -1,3 +1,5 @@
+source("unifprod.R")  # for the C part
+
 pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
     (1/2)*( 1 - sign(x)*p )
 }
@@ -193,3 +195,100 @@ cline_from_soln <- function (soln, grid, clinefn=attr(soln,"u") ) {
     return(pde.cline)
 }
 
+
+####
+## Linkage-related things
+####
+
+recomb_generator <- function (yA,yB,p,zeroind,nabvals,eps) {
+    # the discretization of the map y(a,b) -> \int_a^b (h(a,u)*h(u,b)-h(a,b)) du
+    # yA and yA is the upper triangular part of square matrices
+    #   of dimensions nabvals x nabvals
+    #   eps is the grid spacing between locations
+    #   and zeroind is the index corresponding to the marker at zero
+    eps * ( 
+      p * unifprod_ut(y=yA,ncx=nabvals)
+      + (1-p) * biprod_ut(yA=yA,yB=yB,zeroind=zeroind,ncx=nabvals)
+    )
+}
+
+forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
+        fwds.soln=forwards_pde(s=s,times=times,grid=xgrid,yinit=ifelse(xgrid$x.mid<0,yinit[1:xgrid$N],yinit[xgrid$N+(1:xgrid$N)])),
+        yinit=rep( c( rep(1.0,xgrid$N), rep(0.0,xgrid$N) ), rgrid$N*(rgrid$N+1)/2 ),
+        eps=1e-16, ... ) {
+    ###
+    # solves the system of PDE for the probability that a haplotype across genomic region [a,b]
+    # sampled at time t and location x and linked to selected allele z
+    # derives entirely from side A at time 0.
+    ###
+    # s = selection coefficient
+    # times = times to solve at
+    # xgrid = grid of spatial locations
+    # rgrid = grid of recombination values
+    ###
+    this.lapply <- if ("parallel" %in% .packages() && parallel::detectCores()>1) { function (...) mclapply(...,mc.cores=detectCores()) } else { lapply }
+
+    # precompute some things
+    rgrid$eps <- unique(rgrid$dx)
+    rgrid$zeroind <- which(rgrid$x.mid>0)[1]
+
[…]

### 2015-11-22 notes/unifprod_ut.cpp

@@ -1,4 +1,4 @@
-NumericVector unifprod_ut(NumericVector y, double ncx) {
+NumericVector unifprod_ut(NumericVector y, int ncx) {
     NumericVector z(y.size());
     int index = 0;
     int pindex, outindex;

### 2015-11-22 notes/pde-clines.Rmd
LD decay

@@
 library(Matrix)
-library(parallel)
 source("tran.1D.R") # fix for log.A
@@
 
+Here are haplotype probabilities centered on the selected site:
 ```{r plot_hap_probs, cache=TRUE, depends="solve_branching"}
-hap.probs <- get_hap_probs( hap.soln, left=-.025, right=.025 )
+for (haplen in c(0.02, 0.26, 0.46)) {
+    hap.probs <- get_hap_probs( hap.soln, left=-haplen/2, right=haplen/2 )
     image( hap.probs, zlim=c(0,1), legend=TRUE, yaxt='n', 
@@
     axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
 ```
 
+And, here are haplotype probabilities for a fairly short haplotype moving away from the selected site:
+```{r plot_hap_probs_locs, cache=TRUE, depends="solve_branching"}
+haplen <- 0.04
+for (haploc in c(0.01, 0.09, 0.21, 0.31)) {
+    hap.probs <- get_hap_probs( hap.soln, left=haploc, right=haploc+haplen )
+    image( hap.probs, zlim=c(0,1), legend=TRUE, yaxt='n', 
+          main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+    locator(1)
+}
+```
+
+Here's all the solution traces, i.e., how probabilities of inheriting an $A$ haplotype
+of all lengths, positions, and spatial locations changes with time
+```{r plot_hap_everything, cache=TRUE, depends="solve_branching"}
+matplot(hap.soln[,1],hap.soln[,-1],type='l',xlab='time',ylab='probability')
+```
+
+Here's how these look across space, at a fixed time,
+with line type showing haplotype length
+and color showing distance to the selected site.
+```{r plot_hap_clines, cache=TRUE, depends="solve_branching"}
+plot.t <- 3
+hap.clines <- hap.soln[which(hap.soln[,1]>=plot.t)[1],-1]
+dim(hap.clines) <- attr(hap.soln,"soln.dims")
+r.df <- data.frame(attr(hap.soln,"r"))
+r.df$length <- r.df$right-r.df$left
+r.df$dist <- ifelse(r.df$right*r.df$left>0,pmin(abs(r.df$right),abs(r.df$left)),0)
+r.df$col <- as.numeric(cut(r.df$dist,6))
+r.df$lty <- as.numeric(cut(r.df$length,5))
+matplot(xgrid$x.mid,hap.clines[1:xgrid$N,],type='l',main=paste("A: t=",plot.t),ylim=c(0,1),xlab="space",col=r.df$col,lty=r.df$lty,ylab="probability")
+legend("bottomleft",lty=unique(r.df$lty),legend=paste("length=",round(tapply(r.df$length,r.df$lty,mean),2)))
+matplot(xgrid$x.mid,hap.clines[xgrid$N+1:xgrid$N,],type='l',main=paste("B: t=",plot.t),ylim=c(0,1),xlab="space",col=r.df$col,lty=r.df$lty,ylab="probability")
+legend("topright",lty=1,col=unique(r.df$col),legend=paste("dist=",round(tapply(r.df$dist,r.df$col,mean),2)))
+```
+
+And, here's a heatmap of haplotype length across space and time:
+shown is the longest haplotype length that has probability of at least 50% of coming entirely from $A$:
+```{r plot_hap_lens, cache=TRUE, depends=c("solve_branching","plot_hap_clines")}
+xfac <- rep(1:xgrid$N,nrow(r.df))
+ABfac <- rep( rep(c("A","B"),each=xgrid$N), nrow(r.df) )
+hap.lens <- lapply( c(A="A",B="B"), function (allele) {
+    do.call( rbind, lapply( 1:nrow(hap.soln), function (k) {
+            tapply( hap.soln[k,1+which(ABfac==allele)], xfac, function (xx) {
+                   if(any(xx>0.5)) { max(r.df$length[xx>0.5]) } else { 0.0 }
+                } )
+        } ) )
+    } )
+filled.contour(hap.soln[,1],xgrid$x.mid,hap.lens$A,xlab="time",ylab="space",zlim=c(0,max(hap.lens$B)),main="A-linked A haplotype length")
+filled.contour(hap.soln[,1],xgrid$x.mid,hap.lens$B,xlab="time",ylab="space",zlim=c(0,max(hap.lens$B)),main="B-linked A haplotype length")
+```
+
+Linkage
+=======
+
+The way the discretization for haplotypes is implemented
+effectively only keeps track of a finite number of markers.
+To do two-site ancestral LD, then,
+we just need to use the above code at only two sites on the genome.
+
+First, we compute per-allele solutions:
+```{r solve_LD, cache=TRUE}
+s <- 0.1
+tt <- seq(0,100,length.out=21)
+xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
+rr <- seq(0,0.5,length.out=20)[-1]
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid,yinit=ifelse(xgrid$x.mid<0,1,0))
+LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )
+```
+... but then we must extract linkage from these.
+Below we report for recombination distance $r$,
+the probability that two markers at distance $r$ 
+sampled in one individual at that location and time
+(with one end fixed to the selected site)
+are both of the same ancestral type.
+If the local proportion of type $A$ is $p$ and the probability when linked to $A$ is $u_A$,
+we want $p u_A(r_0,r_0+r) + (1-p) u_B(r_0,r_0+r) - (p u_A(r_0) + (1-p) u_A(r_0))(p u_A(r_0+r) + (1-p) u_A(r_0+r))$:
+(**is that right?**)
+```{r compute_LD, cache=TRUE, depends="solve_LD"}
+pfun <- function (t,x) { 
+        fp <- function (x,t) { approx(xgrid$x.mid,cline_interp(t,soln=fwds.soln),xout=x)$y }
+        ans <- numeric(length(x))
+        for (tval in unique(t)) {
+            dothese <- (rep_len(t,length(x))==tval)
+            ans[dothese] <- fp(x=x[dothese],t=tval)
+        }
+        return(ans)
+}
+ABfac <- rep( rep(c("A","B"),each=xgrid$N), 3 )  # three combinations of two recomb points
+rfac <- rep( c("11","12","22"), each=xgrid$N*2 )
+LDs <- lapply( LD.solns, function (lds) {
+            # returns a time x space matrix
+            pp <- outer( lds[,1], xgrid$x.mid, pfun )
+            return( 
+                   pp * lds[,1+which(ABfac=="A"&rfac=="12")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="12")] 
+                   - ( pp * lds[,1+which(ABfac=="A"&rfac=="11")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="11")] )
+                       * ( pp * lds[,1+which(ABfac=="A"&rfac=="22")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="22")] )
+               )
+    } )
+plot.x <- xgrid$x.mid[which.min(abs(xgrid$x.mid))]
+matplot( tt, sapply( LDs, function (x) { x[,match(plot.x,xgrid$x.mid)] } ), type='l', ylab="ancestral LD", xlab="time" )
+plot.t <- tt[which.min(abs(tt-20))]
+matplot( xgrid$x.mid, sapply( LDs, function (x) { x[match(plot.t,tt),] } ), type='l', ylab="ancestral LD", xlab="space" )
+matplot( rr, t(sapply( LDs, function (x) { x[match(plot.t,tt),] } )), type='l', ylab="ancestral LD", xlab="recomb dist", main=paste("t=",round(plot.t,2)), col=rainbow(xgrid$N), lty=1 )
+legend("topright",lty=1,col=rainbow(xgrid$N)[floor(seq(1,xgrid$N,length.out=6))],legend=paste("x=",round(xgrid$x.mid[floor(seq(1,xgrid$N,length.out=6))],2)))
+```
+Hm, the right/left spatial positions should look the same, right?  Is something off?
+
 Distance to closest recombination in a heterozygote
----------------------------------------------------
+===================================================
 
-A simpler statistic, that might approximate the correct quantity well,
+A simpler statistic, that might approximate haplotype statistics well,
 is the distance along the chromosome from an $A$ allele until the closest recombination

### 2015-11-22 notes/cline-fns.R

@@ -226,10 +226,10 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
     # xgrid = grid of spatial locations
     # rgrid = grid of recombination values
     ###
-    this.lapply <- if ("parallel" %in% .packages() && parallel::detectCores()>1) { function (...) mclapply(...,mc.cores=detectCores()) } else { lapply }
+    # note: doing parallel lapply below is actually a good bit slower.
 
     # precompute some things
-    rgrid$eps <- unique(rgrid$dx)
+    rgrid$eps <- unique(diff(rgrid$x.mid))
     rgrid$zeroind <- which(rgrid$x.mid>0)[1]
 
     # function to find a,b coordinates from the upper triangular coordinates
@@ -253,7 +253,7 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
         p <- cline_interp(t,soln=fwds.soln) # freq of B
         log.p <- log(pmax(p,min(eps,min(p[p>0]))))
         log.1mp <- log(pmax(1-p,min(eps,min((1-p)[p<1]))))
-        diffusion <- do.call( cbind, this.lapply( 1:ncol(y), function (kcol) {
+        diffusion <- do.call( cbind, lapply( 1:ncol(y), function (kcol) {
                 r <- rvals[kcol]
                 yA <- y[1:xgrid$N,kcol]
                 yB <- y[xgrid$N+(1:xgrid$N),kcol]
@@ -264,7 +264,7 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
                         tran.B + r * p * (yA-yB)
                         )
              } ) )
-        recombination <- do.call( rbind, this.lapply( 1:nrow(y), function (krow) {
+        recombination <- do.call( rbind, lapply( 1:nrow(y), function (krow) {
                    if (krow<=xgrid$N) {
                        recomb_generator( yA=y[krow,], yB=y[krow+xgrid$N,], p=p[krow], zeroind=rgrid$zeroind, nabvals=rgrid$N, eps=rgrid$eps )
                    } else {
@@ -278,6 +278,16 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
     attr(hap.soln,"s") <- s
     attr(hap.soln,"sigma") <- sigma
     attr(hap.soln,"soln.dims") <- soln.dims
[…]

### 2015-11-24 notes/pde-clines.Rmd
pde clines
.

@@
 xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=10)
-rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=50)
+rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=40)
 
@@
 
+**Note:** do the same with fixed distance between the two and changing distance to selected locus.
+
 Distance to closest recombination in a heterozygote

### 2015-11-30 notes/pde-clines.Rmd
.

--- 2015-11-22 notes/pde-clines.Rmd
@@
 plot.x <- xgrid$x.mid[which.min(abs(xgrid$x.mid))]
-matplot( tt, sapply( LDs, function (x) { x[,match(plot.x,xgrid$x.mid)] } ), type='l', ylab="ancestral LD", xlab="time" )
+matplot( tt, sapply( LDs, function (x) { x[,match(plot.x,xgrid$x.mid)] } ), type='l', ylab="ancestral LD", xlab="time", main=paste("location=",round(plot.x,2)) )
 plot.t <- tt[which.min(abs(tt-20))]

### 2015-11-30 drafts/hybrid_zone.tex

@@
 % \geometry{landscape} % rotated page geometry
+\usepackage{natbib}
+\usepackage{color}
+% See the ``Article customise'' template for come common customisations
+\newcommand{\alisa}[1]{{\em \color{red} #1}}
 
-% See the ``Article customise'' template for come common customisations
-
-\title{}
+\title{Ancestry linkage in hybrid zones}
 \author{}
@@
 \section{Introduction}
-The process of speciation often involves allopatric divergence in relative isolation.
-When formerly isolated populations come back into contact, a hybrid zone can form, and may be maintained stably by selection.
+The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}.
+When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \cite{Barton1985}.
+Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing differential selection \alisa{better way to say this?} may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \cite{Barton1986}.\\
+
+Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
+These models predict cline width to be narrowest at and nearby selected loci, with cline width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
+Population genetic patterns at these loci are important because, in practice, determining whether a hybrid zone has reached equilibrium can be challenging.
+In particular, even if clines at selected loci are stable, the rest of the genome may still contain additional information.
+
+\alisa{Are we mainly just arguing that we get more robust information from looking at linkage around selected loci? Is there anything else?}
+
+In addition to patterns of allele frequency, the extent of linkage in hybrid zones can also provide information about important parameters, such as age and rates of gene flow, in hybrid zones.
+As denser genotype data becomes more readily available, it is now possible to utilize patterns of linkage in the analysis of hybrid zones.
+Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore can provide information about the age of the hybrid zone, and the strength of selection at nearby loci.\\
+
+Using an equilibrium model. We often don't know if hybrid zones are at equilibrium.
+
+%In the absence of selection, genome-wide covariance in ancestry can be used to estimate timing and gene-flow in hybrid zones \cite{Sedghifar2015}.
+
+%When formerly isolated populations come back into contact, a hybrid zone can form, and may be maintained stably by selection.
 In such instances, selection on some subset of the genome prevents mixing of genotypes between species.
@@
 
-Gene-flow across a hybrid zone is impeded by selection against migrant, or hybrid, genotypes, and as a result, produces a cline in ancestry that is expected to be steepest at and nearby the selected locus.
+%Gene-flow across a hybrid zone is impeded by selection against migrant, or hybrid, genotypes, and as a result, produces a cline in ancestry that is expected to be steepest at and nearby the selected locus.
 Elsewhere in the genome, clines in ancestry are wider, and are more heavily influenced by hybrid zone age and rates of migration.
@@
 
-Many of the traditional models of genetic variation in hybrid zones to date have focused on the equilibrium state of hybrid zones, at both selected and linked loci (lots of Barton).
+%Many of the traditional models of genetic variation in hybrid zones to date have focused on the equilibrium state of hybrid zones, at both selected and linked loci (lots of Barton).
 
@@
 
-As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.
+%As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.
 This means that recombination events that occur away from the center of the zone do not result in a change in $a$.\\
@@
 \subsection{Simulations}
-Forward in time simulations were used to asses theoretical predictions.
+Forward in time simulations were used to assess theoretical predictions.
 We simulate discrete demes with nearest-neighbor migration and non-overlapping generations.
@@
 
+\subsection{Data}
+We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m.
+musculus/M.
+m.
+domesticus} \cite{Turner2011,Turner2014}.
+The genomic data generated by \citep{Turner2014} was for a mapping population generated by mating wild-caught individuals.
+Because some individuals were mated multiple times, the mapping population comprises many closely related individuals.
+We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals).\\
+
+Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}.
+
 \end{document}

### 2015-12-03 notes/diffusion-local-times.Rmd
time scales

@@
 layout(t(1:2))
-fb.p <- clineplot(rev.soln,main="forwards-backwards")
-st.p <- clineplot(pde.soln,main="stationary cline")
+fb.p <- clineplot(rev.soln,thegrid,main="forwards-backwards")
+st.p <- clineplot(pde.soln,thegrid,main="stationary cline")
 layout(1)
@@
 pde.soln.2 <- solve.reflected(x.up=-20)
-clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),xlim=c(-5,0))
-clineplot(pde.soln.1,x=attr(pde.soln.1,"x"),add=TRUE)
-clineplot(pde.soln.2,x=attr(pde.soln.2,"x"),add=TRUE)
+clineplot(pde.soln.0,thegrid,x=attr(pde.soln.0,"x"),xlim=c(-5,0))
+clineplot(pde.soln.1,thegrid,x=attr(pde.soln.1,"x"),add=TRUE)
+clineplot(pde.soln.2,thegrid,x=attr(pde.soln.2,"x"),add=TRUE)
 ```
@@
 pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
-ps <- clineplot(pde.soln,times=seq_along(tt),xlim=c(-8,0),legend=FALSE)
+ps <- clineplot(pde.soln,thegrid,times=seq_along(tt),xlim=c(-8,0),legend=FALSE)
 matlines( thegrid$x.mid, outer( thegrid$x.mid, tt, function(x,t) { zeta.reflected(x/sqrt(t),r*t*x.eps) }), lty=3, col=rainbow(length(tt)) )
@@
 for (k in c(6,10)) {
-    clineplot(pde.solns[[k]],main=sprintf("r=%0.2f",tau,rvals[k]))
+    clineplot(pde.solns[[k]],thegrid,main=sprintf("r=%0.2f",tau,rvals[k]))
 }

### 2015-12-03 notes/rough-calculations.Rmd
time scales

@@
     for (k in seq_along(svals)) {
-        clineplot( fwds.solns[[k]], main=sprintf("s=%f",svals[k]), col=ifelse( fwds.solns[[k]][,1]>1/svals[k], adjustcolor("black",0.5), adjustcolor("red",0.5) ) )
+        clineplot( fwds.solns[[k]], thegrid, main=sprintf("s=%f",svals[k]), col=ifelse( fwds.solns[[k]][,1]>1/svals[k], adjustcolor("black",0.5), adjustcolor("red",0.5) ) )
         abline(v=c(-1,1)/sqrt(svals[k]),lty=3,col='green',lwd=2)
@@
 
-a.  The haplotype of a $B$ allele at distance $x$ into the $A$ side should be of length about $\sigma/(sx)$, as long as this is less than $1/\tau$. (this is fairly restrictive)
+a.  The haplotype of a $B$ allele at distance $x$ into the $A$ side should be of length about $s \sigma/x$, as long as this is less than $1/\tau$. (this is fairly restrictive)
 
@@
     i.e., for a $B$ lineage to make it back to the $B$ side.
-    Since lineages on the wrong side move at speed $\sigma/s$,
+    Since lineages on the wrong side move at speed $\sigma s$,
     if the allele is at position $x$, 
-    this will take about $sx/\sigma$;
-    and so the haplotype should be about $\sigma/(sx)$.
+    this will take about $x/(s\sigma)$;
+    and so the haplotype should be about $s\sigma/x$.
     This gives the size of the chunk carried with it on this excursion;

### 2015-12-03 notes/time-scales.Rmd
time scales

---
title: "Time scales for cline flattening"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
library(Matrix)
source("tran.1D.R") # fix for log.A
source("cline-fns.R")
```

Here we would like to explore what happens over various time scales.
We'll do:

-  tens of generations
-  hundreds of generations
-  thousands of generations

At each time scale, we'll look at:

-  linked clines
-  ancestry disequilibrium
-  haplotype lengths

Tens of generations
===================

First, the setup:
```{r setup_tens, cache=TRUE}
s <- 0.1
tt <- seq(0,40,length.out=21)
xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
rr <- seq(0,0.2,length.out=21)
rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid,yinit=ifelse(xgrid$x.mid<0,1,0))
```
We'll solve it up until time `r max(tt)` and on a grid of `r xgrid$N` sites from `r min(xgrid$x.mid)` to `r max(xgrid$x.mid)`.

Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
```{r fwds_soln_plot, cache=TRUE, depends="setup_tens"}
image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
add_grid_axis(xgrid)
```

Linked clines
-------------

```{r tens_linked_clines, cache=TRUE, depends="setup_tens"}
linked.solns <- lapply( rr, function (r) {
            forwards_backwards_pde(s=s,times=tt,grid=xgrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
        } )
linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
```

Here are linked clines for a few time steps:
```{r plot_tens_linked_clines, cache=TRUE, depends="tens_linked_clines"}
for (plot.t in c(4,10,20,40)) {
    k <- which.min(abs(tt-plot.t))
    lclines <- sapply( linked.clines, cline_interp, T=tt[k] )
    matplot( xgrid$x.mid, lclines, type='l', ylab="proportion", xlab='space', 
            main=sprintf("t=%0.0f",tt[k]),
           lty=1, col=rcols )
    legend("topright", lty=1, col=rcols, legend=sprintf("r=%0.2f",rr) )
}
```

Ancestry disequilibrium
-----------------------

Now we'll look at ancestry disequilibrium across the region.

```{r tens_ld_solutions, cache=TRUE}
rr <- seq(-0.2,0.2,length.out=21)
rcols <- rainbow(length(unique(abs(rr)))*1.2)[match(abs(rr),unique(abs(rr)))]
rpairs <- as.matrix(expand.grid(r1=rr,r2=rr))
LD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), mc.cores=parallel::detectCores() )
```

Here are heatmaps of ancestry LD at a few different times.
The selected site is in the center.
```{r tens_ld_heatmaps, cache=TRUE, depends="tens_ld_solutions"}
```

### 2015-12-03 notes/cline-fns.R

@@ -45,15 +45,21 @@ weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
     colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
 }
 
-clineplot <- function (soln,x=thegrid$x.mid,times=seq(1,nrow(soln),length.out=ntimes),ntimes=20,
-                       ylab="probability", ylim=c(0,1), legend=TRUE, lty=1,
-                       col=rainbow(length(times)),
-                       ...) {
+clineplot <- function (soln, thegrid, x=thegrid$x.mid,
+               times=seq(1,nrow(soln),length.out=ntimes), ntimes=20,
+               ylab="probability", ylim=c(0,1), legend=TRUE, lty=1,
+               col=rainbow(length(times)),
+           ...) {
     # plot output of deSolve as clines in a nice way
     graphics::matplot( x, t(soln[times,-1]), type='l', col=col, lty=lty, 
             xlab="space", ylab=ylab, ylim=ylim, ...)
     if (legend) legend("bottomleft", lty=lty, col=col, legend=paste("t=",round(soln[times,1],digits=2)), cex=0.5)
     return( invisible( t(soln[times,-1]) ) )
+}
+
+add_grid_axis <- function (thegrid,side=2,...) {
+    axis(side=side,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),
+            labels=pretty(thegrid$x.mid))
 }
 
 cline_interp <- function (T,soln) {

### 2015-12-03 notes/pde-clines.Rmd
time scales
fixed yinit at zero

--- 2015-11-24 notes/pde-clines.Rmd
@@
 ---
-
-$$
-\newcommand{\E}{\mathbb{E}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\one}{\mathbf{1}}
-\DeclareMathOperator{\sgn}{sgn}
-\newcommand{\grad}{\nabla}
-$$
 
@@
 rr <- seq(0,0.5,length.out=20)[-1]
-fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid,yinit=ifelse(xgrid$x.mid<0,1,0))
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
 LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )
@@
 we want $p u_A(r_0,r_0+r) + (1-p) u_B(r_0,r_0+r) - (p u_A(r_0) + (1-p) u_A(r_0))(p u_A(r_0+r) + (1-p) u_A(r_0+r))$:
-(**is that right?**)
 ```{r compute_LD, cache=TRUE, depends="solve_LD"}

### 2015-12-03 notes/time-scales.Rmd
fixed yinit at zero

@@
 rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
-fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid,yinit=ifelse(xgrid$x.mid<0,1,0))
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+pfun <- cline_fun(fwds.soln)
 ```
@@
     k <- which.min(abs(tt-plot.t))
-    lclines <- sapply( linked.clines, cline_interp, T=tt[k] )
+    lclines <- sapply( linked.clines, cline_interp, t=tt[k] )
     matplot( xgrid$x.mid, lclines, type='l', ylab="proportion", xlab='space', 
@@
 rcols <- rainbow(length(unique(abs(rr)))*1.2)[match(abs(rr),unique(abs(rr)))]
-rpairs <- as.matrix(expand.grid(r1=rr,r2=rr))
-LD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), mc.cores=parallel::detectCores() )
+rpairs <- as.matrix(subset(expand.grid(r1=rr,r2=rr),r1<r2))
+LD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
+               forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), 
+           mc.cores=parallel::detectCores() )
+LD.values <- lapply( LD.solns, AD_from_soln, xgrid=xgrid )
 ```
 
-Here are heatmaps of ancestry LD at a few different times.
+Here are heatmaps of ancestry LD at a few different times and locations
 The selected site is in the center.
-```{r tens_ld_heatmaps, cache=TRUE, depends="tens_ld_solutions"}
+```{r tens_ld_heatmaps, cache=TRUE, depends="tens_ld_solutions", fig.width=3*fig.dim}
+plot.x <- c(0,5,10)
+these.x <- 
+layout(t(1:3))
+for (plot.t in c(4,10,20,40)) {
+    image( 
+}
 ```

### 2015-12-03 notes/cline-fns.R

@@ -62,17 +62,33 @@ add_grid_axis <- function (thegrid,side=2,...) {
             labels=pretty(thegrid$x.mid))
 }
 
-cline_interp <- function (T,soln) {
+cline_interp <- function (t,soln) {
     # return the interpolated cline for a particular time
     # from the output of deSolve
-    which.t <- findInterval(T,soln[,1],rightmost.closed=TRUE)
-    lu <- soln[c(which.t,which.t+1),1]
-    alpha <- (T-lu[1])/diff(lu)
-    return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
+    sapply(t, function (TT) {
+            which.t <- findInterval(TT,soln[,1],rightmost.closed=TRUE)
+            lu <- soln[c(which.t,which.t+1),1]
+            alpha <- (TT-lu[1])/diff(lu)
+            return( (1-alpha)*soln[which.t,-1] + alpha*soln[which.t+1,-1]  )
+        } )
+}
+
+cline_fun <- function (soln) {
+    # return a function that interpolates the solution to (t,x)
+    function (t,x) { 
+            fp <- function (x,t) { approx(xgrid$x.mid,cline_interp(t,soln=soln),xout=x)$y }
+            ans <- numeric(length(x))
+            for (tval in unique(t)) {
+                dothese <- (rep_len(t,length(x))==tval)
+                ans[dothese] <- fp(x=x[dothese],t=tval)
+            }
+            return(ans)
+    }
 }
 
 forwards_pde <- function (s,times,grid,sigma=1,
-                      yinit=ifelse( grid$x.mid>0, 0, 1 )) {
+                      yinit=ifelse( grid$x.mid==0, 0.5, ifelse( grid$x.mid>0, 0, 1 )) 
+              ) {
     # Solve the pde for frequencies, forwards in time:
     #   by default, yinit=1 to the *left* of zero;
     #   so finds the probability the selected site initiates on the left
@@ -86,7 +102,8 @@ forwards_pde <- function (s,times,grid,sigma=1,
 }
 
 forwards_backwards_pde <- function (s, times, grid, r, sigma=1,
[…]

### 2015-12-03 notes/time-scales.Rmd
.

@@
 rpairs <- as.matrix(subset(expand.grid(r1=rr,r2=rr),r1<r2))
-LD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
+AD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
                forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), 
            mc.cores=parallel::detectCores() )
-LD.values <- lapply( LD.solns, AD_from_soln, xgrid=xgrid )
+AD.values <- lapply( AD.solns, AD_from_soln, xgrid=xgrid )
 ```
@@
 plot.x <- c(0,5,10)
-these.x <- 
 layout(t(1:3))
 for (plot.t in c(4,10,20,40)) {
-    image( 
+    for (k in seq_along(plot.x)) {
+        this.t <- match(plot.t,tt)
+        this.x <- match(plot.x[k],xgrid$x.mid)
+        ADmat <- matrix(NA,nrow=length(rr),ncol=length(rr))
+        ADmat[cbind(match(rpairs[,1],rr),match(rpairs[,2],rr))] <- sapply( AD.values, "[", cbind(this.t,this.x) )
+        image(ADmat,zlim=c(0,max(unlist(AD.values))))
+    }
 }

### 2015-12-03 notes/time-scales.Rmd
more times

@@
 ```
+
+Hundreds of generations
+===================
+
+First, the setup:
+```{r setup_hunds, cache=TRUE}
+s <- 0.1
+tt <- seq(0,400,length.out=21)
+xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
+rr <- seq(0,0.2,length.out=21)
+rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+pfun <- cline_fun(fwds.soln)
+```
+We'll solve it up until time `r max(tt)` and on a grid of `r xgrid$N` sites from `r min(xgrid$x.mid)` to `r max(xgrid$x.mid)`.
+
+Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
+```{r fwds_soln_plot, cache=TRUE, depends="setup_hunds"}
+image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
+add_grid_axis(xgrid)
+```
+
+Linked clines
+-------------
+
+```{r hunds_linked_clines, cache=TRUE, depends="setup_hunds"}
+linked.solns <- lapply( rr, function (r) {
+            forwards_backwards_pde(s=s,times=tt,grid=xgrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+        } )
+linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
+```
+
+Here are linked clines for a few time steps:
+```{r plot_hunds_linked_clines, cache=TRUE, depends="hunds_linked_clines"}
+for (plot.t in 10*c(4,10,20,40)) {
+    k <- which.min(abs(tt-plot.t))
+    lclines <- sapply( linked.clines, cline_interp, t=tt[k] )
+    matplot( xgrid$x.mid, lclines, type='l', ylab="proportion", xlab='space', 
+            main=sprintf("t=%0.0f",tt[k]),
+           lty=1, col=rcols )
+    legend("topright", lty=1, col=rcols, legend=sprintf("r=%0.2f",rr) )
+}
+```
+
+Ancestry disequilibrium
+-----------------------
+
+Now we'll look at ancestry disequilibrium across the region.
+
+```{r hunds_ld_solutions, cache=TRUE}
+rr <- seq(-0.2,0.2,length.out=21)
+rcols <- rainbow(length(unique(abs(rr)))*1.2)[match(abs(rr),unique(abs(rr)))]
+rpairs <- as.matrix(subset(expand.grid(r1=rr,r2=rr),r1<r2))
+AD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
+               forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), 
+           mc.cores=parallel::detectCores() )
+AD.values <- lapply( AD.solns, AD_from_soln, xgrid=xgrid )
+```
+
+Here are heatmaps of ancestry LD at a few different times and locations
+The selected site is in the center.
+```{r hunds_ld_heatmaps, cache=TRUE, depends="hunds_ld_solutions", fig.width=3*fig.dim}
+plot.x <- c(0,5,10)
+layout(t(1:3))
+for (plot.t in 10*c(4,10,20,40)) {
+    for (k in seq_along(plot.x)) {
+        this.t <- match(plot.t,tt)
+        this.x <- match(plot.x[k],xgrid$x.mid)
+        ADmat <- matrix(NA,nrow=length(rr),ncol=length(rr))
+        ADmat[cbind(match(rpairs[,1],rr),match(rpairs[,2],rr))] <- sapply( AD.values, "[", cbind(this.t,this.x) )
+        image(ADmat,zlim=c(0,max(unlist(AD.values))))
+    }
+}
+```
+
+Thousands of generations
+===================
+
+First, the setup:
+```{r setup_thous, cache=TRUE}
+s <- 0.1
+tt <- seq(0,4000,length.out=21)
+xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
+rr <- seq(0,0.2,length.out=21)
+rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
+fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+pfun <- cline_fun(fwds.soln)
+```
+We'll solve it up until time `r max(tt)` and on a grid of `r xgrid$N` sites from `r min(xgrid$x.mid)` to `r max(xgrid$x.mid)`.
+
+Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
+```{r fwds_soln_plot, cache=TRUE, depends="setup_thous"}
+image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
+add_grid_axis(xgrid)
+```
+
+Linked clines
+-------------
+
+```{r thous_linked_clines, cache=TRUE, depends="setup_thous"}
+linked.solns <- lapply( rr, function (r) {
+            forwards_backwards_pde(s=s,times=tt,grid=xgrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+        } )
+linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
+```
+
+Here are linked clines for a few time steps:
+```{r plot_thous_linked_clines, cache=TRUE, depends="thous_linked_clines"}
+for (plot.t in 100*c(4,10,20,40)) {
+    k <- which.min(abs(tt-plot.t))
+    lclines <- sapply( linked.clines, cline_interp, t=tt[k] )
+    matplot( xgrid$x.mid, lclines, type='l', ylab="proportion", xlab='space', 
+            main=sprintf("t=%0.0f",tt[k]),
+           lty=1, col=rcols )
+    legend("topright", lty=1, col=rcols, legend=sprintf("r=%0.2f",rr) )
+}
+```
+
+Ancestry disequilibrium
+-----------------------
+
+Now we'll look at ancestry disequilibrium across the region.
+
+```{r thous_ld_solutions, cache=TRUE}
+rr <- seq(-0.2,0.2,length.out=21)
+rcols <- rainbow(length(unique(abs(rr)))*1.2)[match(abs(rr),unique(abs(rr)))]
+rpairs <- as.matrix(subset(expand.grid(r1=rr,r2=rr),r1<r2))
+AD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
+               forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), 
+           mc.cores=parallel::detectCores() )
+AD.values <- lapply( AD.solns, AD_from_soln, xgrid=xgrid )
+```
+
+Here are heatmaps of ancestry LD at a few different times and locations
+The selected site is in the center.
+```{r thous_ld_heatmaps, cache=TRUE, depends="thous_ld_solutions", fig.width=3*fig.dim}
+plot.x <- c(0,5,10)
+layout(t(1:3))
+for (plot.t in 100*c(4,10,20,40)) {
+    for (k in seq_along(plot.x)) {
+        this.t <- match(plot.t,tt)
+        this.x <- match(plot.x[k],xgrid$x.mid)
+        ADmat <- matrix(NA,nrow=length(rr),ncol=length(rr))
+        ADmat[cbind(match(rpairs[,1],rr),match(rpairs[,2],rr))] <- sapply( AD.values, "[", cbind(this.t,this.x) )
+        image(ADmat,zlim=c(0,max(unlist(AD.values))))
+    }
+}
+```

### 2015-12-03 notes/time-scales.Rmd
.

@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r fwds_soln_plot, cache=TRUE, depends="setup_tens"}
+```{r tens_fwds_soln_plot, cache=TRUE, depends="setup_tens"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r fwds_soln_plot, cache=TRUE, depends="setup_hunds"}
+```{r hunds_fwds_soln_plot, cache=TRUE, depends="setup_hunds"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r fwds_soln_plot, cache=TRUE, depends="setup_thous"}
+```{r thous_fwds_soln_plot, cache=TRUE, depends="setup_thous"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")

### 2015-12-05 notes/computing-AD.Rmd
begun computing AD

---
title: "Measures of correlation in ancestry"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

Goal
=====

Compute something like correlation in ancestry between nearby markers,
in different locations,
as a measure of typical haplotype length in the (spatial and/or genomic) region.

**Nomenclature:** Yaniv followed John in calling it "ancestry disequilibrium".
This admirably follows population genetics tradition
in choosing a name that is in comprehensible to outsider.
I propose calling it something that uses "correlation" instead of "disequilibrium"?
(*Todo:* figure out what the measures below are correlations of, and say that.)

Suggestion
==========

Fix spatial locations $x$, $y$, and $z$.
and loci $i$ and $j$.
Suppose we sample individuals at $x$, $y$, and $z$,
and then in each of these samples, at each locus, pick a random allele;
let $(X_i,X_j)$ be the alleles sampled thus at $x$, and likewise for $Y$ and $Z$.
(If we have phased data, instead pick a random haplotype.)
Then, define
\begin{align} \label{eqn:Dx_defn}
    D_x(i,j) = \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} - \P\{ X_i=Y_i \neq Z_i \} \P\{ X_j=Y_j \neq Z_j \} .
\end{align}

Now, how to estimate this?

Suppose $I$ has alleles $a$ and $A$, and $J$ has alleles $b$ and $B$;
and suppose that at location $x$ we have sampled $n_x$ individuals,
of which $n_x(aa,bb)$ have diploid genotype $a/a$, $b/b$, etcetera.
Define $p_x = n_x/n$, and so forth.
The first term is
\begin{align} \label{eqn:Dx_defn}
    \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} 
\end{align}

### 2015-12-09 drafts/hybrid_zone.tex
added to draft

@@
 When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \cite{Barton1985}.
-Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing differential selection \alisa{better way to say this?} may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}.
+Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing selection may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \cite{Barton1986}.\\
@@
 Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
-These models predict cline width to be narrowest at and nearby selected loci, with cline width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
-Population genetic patterns at these loci are important because, in practice, determining whether a hybrid zone has reached equilibrium can be challenging.
-In particular, even if clines at selected loci are stable, the rest of the genome may still contain additional information.
-
-\alisa{Are we mainly just arguing that we get more robust information from looking at linkage around selected loci? Is there anything else?}
-
-In addition to patterns of allele frequency, the extent of linkage in hybrid zones can also provide information about important parameters, such as age and rates of gene flow, in hybrid zones.
-As denser genotype data becomes more readily available, it is now possible to utilize patterns of linkage in the analysis of hybrid zones.
-Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore can provide information about the age of the hybrid zone, and the strength of selection at nearby loci.\\
-
-Using an equilibrium model. We often don't know if hybrid zones are at equilibrium.
+These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
+As denser genotype data becomes more readily available, we can begin to utilize patterns of linkage in the analysis of hybrid zones.
+Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
+It is our goal here to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
+This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.
+In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one.
+If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact.
+In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
+\\
 
@@
 %-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
--We consider the non-equilibrium state of hybrid zones, and patterns of linkage around selected loci.
-This will become increasingly relevant as studiers of HZs get genomic data and can reliably estimate block lengths.
-\\
 
@@
 Because some individuals were mated multiple times, the mapping population comprises many closely related individuals.
-We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals).\\
+We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals).
+Approximately 280,000 SNPs were represented in the final dataset.
 
-Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}.
+Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}. \\
 
+Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization.
+Specifically, the high LD between quite distantly linked SNPs in the FS population suggests that a relatively high proportion of early-generation hybrids, whereas the rapid decay of LD in the the TU population reflects a much older hybridization event.
+Importance of geographic distribution of popgen patterns.
+
+\begin{table}
+\begin{tabular}{l|c|c}
+Individual ID \citep{Turner2014} & Population \citep{Turner2011} & Longitude(E)\\
+\hline
+FP112 & FS&11.665\\
+FP67 & FS &11.665\\
+FP114 & FS &11.665\\
+FP79 & FS &11.665\\
+FP152 &  FS &11.665\\
+FP11 & GL &11.965\\
+FP111&GL &11.965\\
+FP59 & HA  &11.721\\
+FP92 & HA &11.721\\
+FP7 &  HA  & 11.721\\
+FP45 & HA&11.721\\
+FP244& HO &11.693\\
+FP9&   RE&11.994\\
+FP5 &  RF&11.767\\
+FP78 &  SO &11.539\\
+FP52 &  SO &11.539\\
+FP180& SO &11.539\\
+FP14 & ST&11.539\\
+FP40 & ST&11.539\\
+FP145 & TU&11.748\\
+FP29&  TU&11.748\\
+FP135& TU&11.748\\
+
+\end{tabular}
+\end{table}
 \end{document}

### 2015-12-11 notes/pde-clines.Rmd
fixed symmetry bug?

@@
 rr <- seq(0,0.5,length.out=20)[-1]
-fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+fwds.soln <- forwards_pde(s=s,times=tt,grid=extend_grid(xgrid))
 LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )

### 2015-12-11 notes/tests/check-symmetry.R

@@ -1,0 +1,27 @@
+library(ReacTran)
+library(Matrix)
+source("../tran.1D.R",chdir=TRUE) # fix for log.A
+source("../cline-fns.R",chdir=TRUE)
+s <- 0.1
+tt <- seq(0,400,length.out=21)
+xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
+rr <- c(0.05,0.2)
+rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
+fwds.soln <- forwards_pde(s=s,times=tt,grid=extend_grid(xgrid))
+pfun <- cline_fun(fwds.soln)
+linked.solns <- lapply( rr, function (r) {
+            forwards_backwards_pde(s=s,times=tt,grid=xgrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
+        } )
+linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
+
+# check forwards solution is symmetric
+xx <- fwds.soln[,-1]
+yy <- 1-xx[,ncol(xx):1]
+stopifnot( all( abs(as.vector(xx)-as.vector(yy)) < 1e-12 ) )
+
+# and reverse?
+lclines <- sapply( linked.clines, cline_interp, t=200 )
+stopifnot( all( abs(lclines-(1-lclines[nrow(lclines):1,])) < 1e-12 ) )
+
+
+cat("Everything checks out!\n")

### 2015-12-11 data_analysis/plot_chromopainter_copyprobs.R
added copyprobs.R

@@ -1,0 +1,24 @@
+#plot chromopainter copyprob plot.
+
+setwd("/hdd/HZ_selection/mouse_data_HZ/Unpruned_mouse_data/unpruned_mouse_by_chrom/")
+files =list.files()
+probfiles = files[which(grepl("unpruned",files)==F)]
+
+files_inds = sapply(probfiles,function(X){paste("X_",strsplit(strsplit(X,"_")[[1]][2],"ind")[[1]][1],sep="")})
+
+
+indtable = read.table("/hdd/HZ_selection/mouse_data_HZ/mouse_by_chrom/mousedata_indfile",as.is=T)
+inds = indtable$V1
+
+pdf(file="/hdd/HZ_selection/mouse_data_HZ/Unpruned_mouse_data/chromopaintings.pdf",width=13,height=3)
+
+for(i in 1:length(probfiles)){
+
+	data=read.table(probfiles[i],as.is=T)
+	plot(data$V1,data$V2,type="l",main=paste("chr",probfiles[i],"_pop",indtable$V2[which(inds==files_inds[i])],sep=""),ylab="Pr(ancA)",xlab="position (bp)")
+
+	with(data,polygon(c(V1,V1[length(V1)],V1[1]), c(V2,0,0),col="red" ))
+}
+
+dev.off()
+

### 2015-12-11 notes/time-scales.Rmd

@@
 rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
-fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
+fwds.soln <- forwards_pde(s=s,times=tt,grid=extend_grid(xgrid))
 pfun <- cline_fun(fwds.soln)
@@
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
-add_grid_axis(xgrid)
+add_grid_axis(attr(fwds.soln,"grid"))
 ```
@@
             main=sprintf("t=%0.0f",tt[k]),
-           lty=1, col=rcols )
+           lty=1, col=rcols, ylim=c(0,1) )
     legend("topright", lty=1, col=rcols, legend=sprintf("r=%0.2f",rr) )

### 2015-12-11 notes/computing-AD.Rmd
estimating AD

@@
 This admirably follows population genetics tradition
-in choosing a name that is in comprehensible to outsider.
+in choosing a name that is incomprehensible to outsider.
 I propose calling it something that uses "correlation" instead of "disequilibrium"?
@@
 of which $n_x(aa,bb)$ have diploid genotype $a/a$, $b/b$, etcetera.
-Define $p_x = n_x/n$, and so forth.
-The first term is
-\begin{align} \label{eqn:Dx_defn}
+First, compute the two-locus haplotype probabilities:
+\begin{align}
+    p_x(a,b)
+    &=
+    \frac{1}{n_x} \left( 
+        n_x(aa,bb) 
+        + \frac{1}{2} n_x(aa,bB)
+        + \frac{1}{2} n_x(aA,bb)
+        + \frac{1}{4} n_x(aA,bB)
+     \right)
+\end{align}
+and so forth.
+We can get these by defining $p_x(a,b) = n_x(a,b)/n_x$, where expected haplotype counts are defined by:
+\begin{align}
+    \begin{bmatrix} n(a,b) \\ n(A,b) \\ n(a,B) \\ n(A,B) \end{bmatrix}
+    &=
+    \begin{bmatrix}     1    & \frac{1}{2} &    0     & \frac{1}{2} & \frac{1}{4} &    0        &    0      &    0        &     0   \\ % ab
+                        0    & \frac{1}{2} &    1     &    0        & \frac{1}{4} & \frac{1}{2} &    0      &    0        &     0   \\ % Ab
+                        0    &     0       &    0     & \frac{1}{2} & \frac{1}{4} &    0        &    1      & \frac{1}{2} &     0   \\ % aB
+                        0    &     0       &    0     &    0        & \frac{1}{4} & \frac{1}{2} &    0      & \frac{1}{2} &     1      % AB
+    \end{bmatrix}
+    \begin{bmatrix} n(aa,bb) \\ n(aA,bb)  \\ n(AA,bb) \\ n(aa,bB)  \\ n(aA,bB)   \\   n(AA,bB) \\ n(aa,BB) \\ n(aA,BB)   \\ n(AA,BB) \end{bmatrix}
+\end{align}
+Denote by $Q$ the matrix above, and note that columns of $Q$ sum to 1.
+
+The first term is estimated by
+\begin{align} 
     \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} 
+    &=
+    p_x(a,b) p_y(a,b) p_z(A,B)
+    + p_x(A,b) p_y(A,b) p_z(a,B)
+      \\ & \qquad {}
+    + p_x(a,B) p_y(a,B) p_z(A,b)
+    + p_x(A,B) p_y(A,B) p_z(a,b)
 \end{align}
+(In this ordering, reversing the vector complements it, so in R this would be `p[,x]*p[,y]*rev(p[,z])`.)
+
+We would like to estimate the probability in the second term
+by independent, i.e., *distinct* samples,
+so it is a bit more involved.
+First, suppose that we draw two individuals from our sample from $x$
+*without replacement*, and pick a haplotype in each.
+If the first genotype is $(aa,bB)$, for instance, then the probability of obtaining a $b$ on the second draw is
+$$ \frac{n_x(aa,bb) + n_x(aA,bb) + \frac{1}{2} (n_x(aa,bB)-1) + \frac{1}{2} n_x(aA,bB) }{ n_x } = \frac{ n_x(\cdot,b) - 1/2 }{ n_x-1 } . $$
+Similar calculations show that 
+the probability that we get an $a$ at the first locus in the first draw and a $b$ at the second locus in the second draw is
+$$
+    \tilde p_x(a,b) = \frac{n_x}{n_x-1} p_x(a,\cdot) p_x(\cdot,b) - \frac{1}{n_x-1} \sum_g p_x(g) H_g(b)^2,
+$$
+where the sum is over diploid genotypes $g$, and
+$H_g(b)$ is the first row of $H_g$, obtained by multiplying $Q$ on the left by a $2 \times 4$ matrix:
+\begin{align}
+H_g = \begin{bmatrix}
+    1 & 1 & 1 & 1/2 & 1/2 & 1/2 & 0 & 0 & 0 \\
+    0 & 0 & 0 & 1/2 & 1/2 & 1/2 & 1 & 1 & 1 
+\end{bmatrix} .
+\end{align}
+Using this definition, the second term is then estimated as
+\begin{align} 
+    \P\{ X_i=Y_i \neq Z_i \} \P\{ X_j=Y_j \neq Z_j \}
+    &=
+    \tilde p_x(a,b) \tilde p_y(a,b) \tilde p_z(A,B)
+    + \tilde p_x(A,b) \tilde p_y(A,b) \tilde p_z(a,B)
+      \\ & \qquad {}
+    + \tilde p_x(a,B) \tilde p_y(a,B) \tilde p_z(A,b)
+    + \tilde p_x(A,B) \tilde p_y(A,B) \tilde p_z(a,b)
+\end{align}

### 2015-12-16 drafts/hybrid_zone.tex
added some simulation results

@@
-\documentclass[12pt]{amsart}
+\documentclass[12pt]{article}
 \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}.
-When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \cite{Barton1985}.
-Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing selection may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \cite{Barton1986}.\\
+When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \citep{Barton1985}.
+Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing selection may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \citep{Barton1986}.\\
+
+The mixing of ancestries is brought about by the the process of recombination that takes place every generation.
+The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.
+Here we investigate expected genome-wide patterns of co-ancestry in the presence of selection.
+Selection against migrant ancestries is expected to reduce admixture proportions at and surrounding a target of selection.
+The effect on patterns of ancestral block lengths is, however, less well understood.
+Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.\\
 
@@
 We simulate discrete demes with nearest-neighbor migration and non-overlapping generations.
-Each generation, an individual moves to a neighboring deme with probability $m$.
-The direction in which it moves is determined by a coin-flip, except at edge demes, where an individual moves in one direction with probability $m/2$.
+Each generation, the dispersal distance of an individual is sampled from a normal distribution, and then discretized to correspond to a deme that is a corresponding distance away.
 Following migration, the next generation is produced by replacing each individual in a deme by randomly choosing two parents with probability weighted by the fitness of their genotypes.
@@
 Each individual has 1 pair of chromosomes of length 1 Morgan.
+We simulate a single selected locus at position 1M along the chromosome.
+
+Hybrid zones of ages 10, 100 and 1000 generations were simulated under conditions of $\sigma=1$ and $s=0.1$.
+From this we measure unbroken tract lengths to the right of the selected locus.
+These were compared to the output of neutral simulations with $s=0$
 
@@
 
-Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}. \\
+Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\
 
@@
 \begin{tabular}{l|c|c}
-Individual ID \citep{Turner2014} & Population \citep{Turner2011} & Longitude(E)\\
+Individual ID  & Population \citep{Turner2011} & Longitude(E)\\
+\citep{Turner2014}& \citep{Turner2011}&\\
 \hline
@@
 \end{table}
+
+\section{Discussion}
+Topics include\\
+- ignoring coalescent\\
+- single locus model, selection is STRONG
+
+\bibliographystyle{dinat}
+\bibliography{library}
+
 \end{document}

### 2015-12-16 sims/yaniv_sims.R

@@ -10,10 +10,12 @@ SIGMA = 1
 ndemes = 50
 deme_size = 500
 ninds = ndemes*deme_size
-S = 0.5
+S = 0.1
 
 
-zone_age = c(50)
+zone_age = c(100)
+
+outfile = sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_tau%s",SIGMA,ninds,ndemes,S,zone_age)
 
 #PARAMS for parsing:
 loci = list(seq(0.5,1,0.05))
@@ -33,6 +35,7 @@ sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
 			#"pars" is an integer that gives the number of generatsion of the simulation
 			#"QTL" returns the QTL vector used in the simulation
 
+save(sims.s0.1,file=paste(outfile,"_rawoutput.Robj",sep=""))
 			
 #PARSE the simulated chromosomes
 sims.sums = lapply(sims.s0.1,spBreaks)
@@ -41,7 +44,7 @@ sims.sums = lapply(sims.s0.1,spBreaks)
 		#"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
 		
 
-#save(sims.sums,file="~/Projects/HybridZones/ClineProjects/sims_sums_s0.1.Robj")
+save(sims.sums,file=paste(outfile,"_simsums.Robj",sep=""))
 
 
 ######################
@@ -57,7 +60,7 @@ freqs = lapply(my.genos,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X)
 
 
 #pdf(file="freqplot.pdf")
-matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
+matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main=paste(zone_age,"generations"),ylab="freq",xlab="deme")
 legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
 matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
 #dev.off()
@@ -65,6 +68,8 @@ matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
 ###########################
 ###########################
[…]

### 2016-01-01 notes/time-scales.Rmd
dependson not depends, grrr

@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r tens_fwds_soln_plot, cache=TRUE, depends="setup_tens"}
+```{r tens_fwds_soln_plot, cache=TRUE, dependson="setup_tens"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
@@
 
-```{r tens_linked_clines, cache=TRUE, depends="setup_tens"}
+```{r tens_linked_clines, cache=TRUE, dependson="setup_tens"}
 linked.solns <- lapply( rr, function (r) {
@@
 Here are linked clines for a few time steps:
-```{r plot_tens_linked_clines, cache=TRUE, depends="tens_linked_clines"}
+```{r plot_tens_linked_clines, cache=TRUE, dependson="tens_linked_clines"}
 for (plot.t in c(4,10,20,40)) {
@@
 The selected site is in the center.
-```{r tens_ld_heatmaps, cache=TRUE, depends="tens_ld_solutions", fig.width=3*fig.dim}
+```{r tens_ld_heatmaps, cache=TRUE, dependson="tens_ld_solutions", fig.width=3*fig.dim}
 plot.x <- c(0,5,10)
@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r hunds_fwds_soln_plot, cache=TRUE, depends="setup_hunds"}
+```{r hunds_fwds_soln_plot, cache=TRUE, dependson="setup_hunds"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
@@
 
-```{r hunds_linked_clines, cache=TRUE, depends="setup_hunds"}
+```{r hunds_linked_clines, cache=TRUE, dependson="setup_hunds"}
 linked.solns <- lapply( rr, function (r) {
@@
 Here are linked clines for a few time steps:
-```{r plot_hunds_linked_clines, cache=TRUE, depends="hunds_linked_clines"}
+```{r plot_hunds_linked_clines, cache=TRUE, dependson="hunds_linked_clines"}
 for (plot.t in 10*c(4,10,20,40)) {
@@
 The selected site is in the center.
-```{r hunds_ld_heatmaps, cache=TRUE, depends="hunds_ld_solutions", fig.width=3*fig.dim}
+```{r hunds_ld_heatmaps, cache=TRUE, dependson="hunds_ld_solutions", fig.width=3*fig.dim}
 plot.x <- c(0,5,10)
@@
 Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
-```{r thous_fwds_soln_plot, cache=TRUE, depends="setup_thous"}
+```{r thous_fwds_soln_plot, cache=TRUE, dependson="setup_thous"}
 image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
@@
 
-```{r thous_linked_clines, cache=TRUE, depends="setup_thous"}
+```{r thous_linked_clines, cache=TRUE, dependson="setup_thous"}
 linked.solns <- lapply( rr, function (r) {
@@
 Here are linked clines for a few time steps:
-```{r plot_thous_linked_clines, cache=TRUE, depends="thous_linked_clines"}
+```{r plot_thous_linked_clines, cache=TRUE, dependson="thous_linked_clines"}
 for (plot.t in 100*c(4,10,20,40)) {
@@
 The selected site is in the center.
-```{r thous_ld_heatmaps, cache=TRUE, depends="thous_ld_solutions", fig.width=3*fig.dim}
+```{r thous_ld_heatmaps, cache=TRUE, dependson="thous_ld_solutions", fig.width=3*fig.dim}
 plot.x <- c(0,5,10)

### 2016-01-20 drafts/hybrid_zone.tex
some figures and code

@@
 When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \citep{Barton1985}.
-Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing selection may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+If genetic material can be exchanged via hybridization, clines in ancestry throughout the genome may diminish over time; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \citep{Barton1986}.\\
@@
 
+Some words about how new genomic data will allow us to focus on tracts and not just individual loci.
+
 Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
 These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
-As denser genotype data becomes more readily available, we can begin to utilize patterns of linkage in the analysis of hybrid zones.
+As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
+
  Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
@@
 \\
+
+Expectations:
+Conditional  on having identity B on the wrong side, far away from the center of the zone, the length of unbroken ancestry surrounding the selected locus should be relatively long.
+This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
 
@@
 This simulation keeps track of the ancestry at each location in the chromosome for all individuals.
-Each individual has 1 pair of chromosomes of length 1 Morgan.
-We simulate a single selected locus at position 1M along the chromosome.
+Each individual has 2 pairs of chromosomes each of length 1 Morgan such that chromosome 1 contains a selected locus at position 0.5M, and chromosome 2 contains no selected loci.
 
@@
 From this we measure unbroken tract lengths to the right of the selected locus.
-These were compared to the output of neutral simulations with $s=0$
+These were compared to the output of neutral simulations with $s=0$\\
 
@@
 
+\section{Results}
+\subsection{The distribution of ancestry tract lengths}
+We are interested specifically in the distribution of the length of continuous tracts of ancestry surrounding the selected locus.
+Far away from the zone center variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, however, when present an unfit migrant haplotype is expected to be longer than a neutral migrant haplotype.
+Figure X shows the estimated probability density (pdf), of unbroken tract of ancestry B surrounding the selected locus, conditioning on a haplotype being of ancestry B at the selected locus.
+Compared to the expected density of tract lengths surrounding neutral loci, tracts of ancestry B around the selected locus are longer.\\
+
+\subsection{Number of ancestors}
+Not sure what the result is here yet -- unexpected
+
 \section{Discussion}
@@
 - ignoring coalescent\\
-- single locus model, selection is STRONG
+- single locus model, selection is STRONG\\
+
+Some words about how confidently calling tracts is the way of the future.\\
+How can this be applied to detecting selection?
 

### 2016-01-20 sims/plot_chromosome_schematic.R

@@ -1,0 +1,114 @@
+#Want to produce a plot of the haplotypes
+setwd("~/Documents/Hybrid_Zones/clinal-lineages/")
+
+plot.chroms = function(DATA=sims.sums,INDS=random_inds,CHR="chr1"){
+	
+		plot(0:1,0:1,col="white",bty="n",xaxt="n",yaxt="n",xlab="",ylab="")
+
+	for(i in 1:length(random_inds)){
+		data=DATA[[1]][["ind.ancest"]][[INDS[i]]][[CHR]]$X1
+		rect(data$starts,(i-1)/50+0.001,data$stops,i/50-0.001,col=ifelse(data$sp2,"orange","red"),border=FALSE)
+		#segments(0,i/50,1,i/50,col="orange",lwd=5,lend="butt")
+		#if(length(which(data$sp2))>0){
+		#segments(data$starts[data$sp2],i/50,data$stops[data$sp2],i/50,col="red",lwd=5,lend="butt")	}
+	} 	
+}
+
+ndemes=50
+random_inds = sapply(1:ndemes,function(X){ind_indices=1:500+(X-1)*500;return(sample(ind_indices,size=1))})
+
+
+#TAU=100
+
+chroms_sel= get(load("sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"))
+chroms_neut = get(load("sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"))
+	#output is:  a list: [[zone_age]][[c("ind.ancest","ancest.prop")]]
+		#"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
+		#"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
+
+pdf(file=c("sims/plot_chromosomes_tau100.pdf"))
+par(mfrow=c(1,2),mar=c(0,0,0,0),oma=c(4,4,2,1))
+
[…]

### 2016-01-20 sims/plot_compare_chunk_ecdf.R

@@ -1,0 +1,60 @@
+load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums_chunks.Robj")
+#this is a list of [selected,near,unlinked chunk distribution].The object name is "chunks"
+
+load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau1000_chunks_by_deme.Robj")
+#this is only around the selected locus. The object name is "chunks_by_deme"
+
+
+add.alpha <- function(col, alpha=1){
+if(missing(col))
+stop("Please provide a vector of colours.")
+apply(sapply(col, col2rgb)/255, 2, 
+function(x) 
+rgb(x[1], x[2], x[3], alpha=alpha)) 
+
+}
+transparent_rainbow = add.alpha(rainbow(ndemes),0.7)
+
+
+empty_deme = rep(0,2*deme_size)
+
+chunks_ecdf = lapply(chunks,function(X){
+	lapply(X,function(D){
+		segregating =which(D<1 & D>0)
+		if(length(segregating)>0){
+		return(ecdf(D[segregating]))}else return(ecdf(D))	
+	})})
+
+
+noSelection_ecdf = lapply(chunks_by_deme,function(D){segregating =which(D<1 & D>0)
+		if(length(segregating)>0){
+		return(ecdf(D[segregating]))}else return(ecdf(D))})
+
+#vanilla ecdfs	
+pdf(file="simulation_SIGMA1_Ninds25000_ndemes50_s0_tau1000_ecdf_comparisons.pdf")	
+par(mfrow=c(2,2),mar=c(2,2,2,2))
+for(type in names(chunks)){
+	plot(1, type="n",ylim=c(0,1),xlim=c(0,0.2),main=type,ylab="1-ecdf(x)",xlab="x=chunk length")	
+	#legend("topright",legend=seq(-24.5,24.5,5),col=transparent_rainbow[seq(1,50,5)],lty=1)
+	for(i in rev(1:ndemes)){points(1-chunks_ecdf[[type]][[i]](seq(0,1,0.005))~seq(0,1,0.005),ylim=c(0,1),type="l",col=transparent_rainbow[i])}
+}
+
[…]

### 2016-01-20 sims/sim-fns.R

@@ -201,7 +201,7 @@ sim.zone <- function (n.ind, n.deme, ...) {
 	sim.gens( n=n.ind, 
              sp.ids = rep(c("A","B"),each=n.ind/2), 
              deme=rep(1:n.deme,each=floor(n.ind/n.deme)),
-             n.chr=1,
+             n.chr=length(qtl),
              QTL = qtl,...)
 }
 
@@ -215,10 +215,10 @@ geno.ind <- function(IND,loci) {
     # Each individual is a list of chromosomes; each chromosome is a two-column data frame of starts, stops, and ancestral identities
     # Returns diploid genotypes;
     #  as a vector, with all chromosomes squooshed together
-    if (length(IND)!=length(loci)) { stop("Number of chromsomes not matching.") }
-    unlist( lapply( seq_along(IND), function (CHR) {
+    if (length(IND)!=length(loci)) { stop("Number of chromosomes not matching.") }
+     lapply( seq_along(IND), function (CHR) {
             rowSums( sapply(IND[[CHR]],function(HAP){
                    HAP$sp2[ findInterval( loci[[CHR]], c(HAP$starts,1), rightmost.closed=TRUE ) ]
               }) )
-        }) )
+        }) 
 }

### 2016-01-20 sims/yaniv_sims.R

@@ -4,6 +4,16 @@
 
 source("sim-fns.R")
 
+add.alpha <- function(col, alpha=1){
+if(missing(col))
+stop("Please provide a vector of colours.")
+apply(sapply(col, col2rgb)/255, 2, 
+function(x) 
+rgb(x[1], x[2], x[3], alpha=alpha)) 
+
+}
+
+
 #PARAMS for simulation:
 
 SIGMA = 1
@@ -12,18 +22,20 @@ deme_size = 500
 ninds = ndemes*deme_size
 S = 0.1
 
+transparent_rainbow = add.alpha(rainbow(ndemes),0.7)
 
 zone_age = c(100)
 
 outfile = sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_tau%s",SIGMA,ninds,ndemes,S,zone_age)
 
 #PARAMS for parsing:
-loci = list(seq(0.5,1,0.05))
+loci = list(seq(0.5,1,0.05),seq(0.5,1,0.05))
+xx <- (1:ndemes)-0.5-ndemes/2
 
 
 
 #DEFINE selection against genotypes:
-qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
+qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)),chr2=data.frame(traits=c("underdominant"), s = 0,pos=c(0.5)))
 
 #SIMULATE the populations (NB: Can't make per-deme population size too small, otherwise deme may end up empty and throws up an error.)
 sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
@@ -54,15 +66,18 @@ save(sims.sums,file=paste(outfile,"_simsums.Robj",sep=""))
 ###NOW get frequencies at each site. 
 
 #GENOTYPE ALL MY INDS
-my.genos = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
-	#my.genos is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.
[…]

### 2016-01-20 notes/pde-clines.Rmd

@@
 Here are haplotype probabilities centered on the selected site:
-```{r plot_hap_probs, cache=TRUE, depends="solve_branching"}
+```{r plot_hap_probs, cache=TRUE, dependson="solve_branching"}
 for (haplen in c(0.02, 0.26, 0.46)) {
@@
 And, here are haplotype probabilities for a fairly short haplotype moving away from the selected site:
-```{r plot_hap_probs_locs, cache=TRUE, depends="solve_branching"}
+```{r plot_hap_probs_locs, cache=TRUE, dependson="solve_branching"}
 haplen <- 0.04
@@
 of all lengths, positions, and spatial locations changes with time
-```{r plot_hap_everything, cache=TRUE, depends="solve_branching"}
+```{r plot_hap_everything, cache=TRUE, dependson="solve_branching"}
 matplot(hap.soln[,1],hap.soln[,-1],type='l',xlab='time',ylab='probability')
@@
 and color showing distance to the selected site.
-```{r plot_hap_clines, cache=TRUE, depends="solve_branching"}
+```{r plot_hap_clines, cache=TRUE, dependson="solve_branching"}
 plot.t <- 3
@@
 shown is the longest haplotype length that has probability of at least 50% of coming entirely from $A$:
-```{r plot_hap_lens, cache=TRUE, depends=c("solve_branching","plot_hap_clines")}
+```{r plot_hap_lens, cache=TRUE, dependson=c("solve_branching","plot_hap_clines")}
 xfac <- rep(1:xgrid$N,nrow(r.df))
@@
 we want $p u_A(r_0,r_0+r) + (1-p) u_B(r_0,r_0+r) - (p u_A(r_0) + (1-p) u_A(r_0))(p u_A(r_0+r) + (1-p) u_A(r_0+r))$:
-```{r compute_LD, cache=TRUE, depends="solve_LD"}
+```{r compute_LD, cache=TRUE, dependson="solve_LD"}
 pfun <- function (t,x) { 

### 2016-01-25 drafts/hybrid_zone.tex
made code and plot for statistics

@@
 
-\subsection{}
-
 \subsection{Simulations}
@@
 These were compared to the output of neutral simulations with $s=0$\\
+
+\subsection{Statistics}
+With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci.
+In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection.
+The most straightforward measure is $l_I(b)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site.
+
+Another statistic is the cumulative probability $C_I(b)$ of a sampled chunk encompassing position $b$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
+
+Additionally, we look at the mean value of these statistics in the two chunks, $b-$ and $b+$, that flank the one containing $b$.
 

### 2016-01-25 sims/potential_statistics.R

@@ -1,0 +1,151 @@
+load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums.Robj")
+
+deme_ID = do.call(c,lapply(1:50,function(X){rep(X,500)}))
+
+#1. For each site, add up scores of all overlapping haplotypes
+#2. Ways to score:
+#	1. length
+#	2. score = $\log$ of ECDF of length (i.e. log of prob that haplotype is longer than this one)
+
+#lengths:
+#lengths given a particular ancestry
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR="chr1",POS=0.5,ancB=TRUE){
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancA),0)
+	
+}
+
+intervalSizes = lapply(seq(0,1,0.05),function(POS){
+				do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
+				get.interval.size(IND_DATA=IND,CHR="chr1",POS=POS,ancB=TRUE)	
+				}))
+				})
+				
+#lengths of any ancestry:
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR="chr1",POS=0.5,ancB=TRUE){
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+}
+
+intervalSizes_allAncs = lapply(seq(0,1,0.05),function(POS){
+				do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
+				get.interval.size(IND_DATA=IND,CHR="chr1",POS=POS,ancB=TRUE)	
+				}))
+				})
+
+#[[POS]][ind,chr]
+
+#COMPARE:
+pdf(file="statistic_length.pdf",height = 3,width=5)
+#mean length across ALL indivduals in the zone (including non-recombinant chroms) vs chromosome position
+
[…]

### 2016-01-28 drafts/hybrid_zone.tex
a bunch of new figures

@@
 The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}.
-When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \citep{Barton1985}.
-If genetic material can be exchanged via hybridization, clines in ancestry throughout the genome may diminish over time; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \citep{Barton1986}.\\
+When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting a hybrid zone \citep{Barton1985}.
+In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD).
+If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.\\
 
-The mixing of ancestries is brought about by the the process of recombination that takes place every generation.
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation.
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.
-Here we investigate expected genome-wide patterns of co-ancestry in the presence of selection.
-Selection against migrant ancestries is expected to reduce admixture proportions at and surrounding a target of selection.
+Here we investigate expected genome-wide patterns of co-ancestry across a hybrid zone in the presence of selection.
+Selection against migrant ancestries in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection.
 The effect on patterns of ancestral block lengths is, however, less well understood.
@@
 
-Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
+While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
 These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
@@
  Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
-It is our goal here to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
+Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
 This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.
@@
 
-\subsection{Data}
-We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m.
+%\subsection{Data}
+%We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m.
 musculus/M.
@@
 
-Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\
+%Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\
 
-Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization.
+%Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization.
 Specifically, the high LD between quite distantly linked SNPs in the FS population suggests that a relatively high proportion of early-generation hybrids, whereas the rapid decay of LD in the the TU population reflects a much older hybridization event.
@@
 
-\begin{table}
-\begin{tabular}{l|c|c}
-Individual ID  & Population \citep{Turner2011} & Longitude(E)\\
-\citep{Turner2014}& \citep{Turner2011}&\\
-\hline
-FP112 & FS&11.665\\
-FP67 & FS &11.665\\
-FP114 & FS &11.665\\
-FP79 & FS &11.665\\
-FP152 &  FS &11.665\\
-FP11 & GL &11.965\\
-FP111&GL &11.965\\
-FP59 & HA  &11.721\\
-FP92 & HA &11.721\\
-FP7 &  HA  & 11.721\\
-FP45 & HA&11.721\\
-FP244& HO &11.693\\
-FP9&   RE&11.994\\
-FP5 &  RF&11.767\\
-FP78 &  SO &11.539\\
-FP52 &  SO &11.539\\
-FP180& SO &11.539\\
-FP14 & ST&11.539\\
-FP40 & ST&11.539\\
-FP145 & TU&11.748\\
-FP29&  TU&11.748\\
-FP135& TU&11.748\\
+%\begin{table}
+%\begin{tabular}{l|c|c}
+%Individual ID  & Population \citep{Turner2011} & Longitude(E)\\
+%\citep{Turner2014}& \citep{Turner2011}&\\
+%\hline
+%FP112 & FS&11.665\\
+%FP67 & FS &11.665\\
+%FP114 & FS &11.665\\
+%FP79 & FS &11.665\\
+%FP152 &  FS &11.665\\
+%FP11 & GL &11.965\\
+%FP111&GL &11.965\\
+%FP59 & HA  &11.721\\
+%FP92 & HA &11.721\\
+%FP7 &  HA  & 11.721\\
+%FP45 & HA&11.721\\
+%FP244& HO &11.693\\
+%FP9&   RE&11.994\\
+%FP5 &  RF&11.767\\
+%FP78 &  SO &11.539\\
+%FP52 &  SO &11.539\\
+%FP180& SO &11.539\\
+%FP14 & ST&11.539\\
+%FP40 & ST&11.539\\
+%FP145 & TU&11.748\\
+%FP29&  TU&11.748\\
+%FP135& TU&11.748\\
 
-\end{tabular}
-\end{table}
+%\end{tabular}
+%\end{table}
 

### 2016-01-28 sims/potential_statistics.R

@@ -1,6 +1,20 @@
 load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums.Robj")
 
-deme_ID = do.call(c,lapply(1:50,function(X){rep(X,500)}))
+chromosome = "chr2"
+plot_start = 0.45
+plot_stop=0.55
+increments = 0.005
+Ngen = 1000
+sel=0.1
+SIGMA=1
+Ninds=25000
+Ndemes = 50
+
+positions = seq(plot_start,plot_stop,increments)
+
+outstring = sprintf("simulation_SIGMA%d_Ninds%d_ndemes50_s%g_tau%d_%s_start%g_stop%g_by%g",SIGMA,Ninds,sel,Ngen,chromosome,plot_start,plot_stop,increments)
+
+deme_ID = do.call(c,lapply(1:Ndemes,function(X){rep(X,Ninds/Ndemes)}))
 
 #1. For each site, add up scores of all overlapping haplotypes
 #2. Ways to score:
@@ -9,40 +23,40 @@ deme_ID = do.call(c,lapply(1:50,function(X){rep(X,500)}))
 
 #lengths:
 #lengths given a particular ancestry
-get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR="chr1",POS=0.5,ancB=TRUE){
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=chromosome,POS=0.5,ancB=TRUE){
 	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})
 	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
 	replace(chunk,which(identity==ancA),0)
 	
 }
 
-intervalSizes = lapply(seq(0,1,0.05),function(POS){
+intervalSizes = lapply(positions,function(POS){
 				do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
-				get.interval.size(IND_DATA=IND,CHR="chr1",POS=POS,ancB=TRUE)	
+				get.interval.size(IND_DATA=IND,CHR=chromosome,POS=POS,ancB=TRUE)	
 				}))
 				})
 				
 #lengths of any ancestry:
[…]

### 2016-01-29 drafts/hybrid_zone.tex
added files so latex compiles

@@
 If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.\\
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.
 
@@
 The effect on patterns of ancestral block lengths is, however, less well understood.
-Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.\\
+Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
 
@@
 In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
-\\
 
@@
 We refer to such hybrid zones as tension zones (as opposed to contact zones which reflect a non-equilibrium gradient in ancestry that will eventually be eroded by migration throughout the whole genome).
-Because the population genetic signatures around loci experiencing selection is going to differ from other regions of the genome, natural hybrid zones present exciting opportunities to understand species maintenance and the process of speciation.\\
+Because the population genetic signatures around loci experiencing selection is going to differ from other regions of the genome, natural hybrid zones present exciting opportunities to understand species maintenance and the process of speciation.
 
@@
 For example, patterns of linkage disequilibrium (LD) or tract length distribution across geographic space can provide information about zone age \cite{Sedghifar2015}.
-In tensions zones geographic patterns of linkage at loci neighboring targets of selection can potentially be a powerful source of information.\\
+In tensions zones geographic patterns of linkage at loci neighboring targets of selection can potentially be a powerful source of information.
 
@@
 Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone).
-\\
 %-Gene flow across a hybrid zone may be impeded by selection against migrants.
@@
 There is probably a lot to learn from geographic patterns at loci neighboring selected loci (e.g.
-small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)\\
-%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
+small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)
+%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci.
 
@@
 Individuals move forward in time by sampling a displacement from a Gaussian distribution.
-In the simplest case of under-dominance, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.\\
+In the simplest case of under-dominance, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
 
@@
 The identity $a$ at locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
-\\
 
 %As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.
-This means that recombination events that occur away from the center of the zone do not result in a change in $a$.\\
+This means that recombination events that occur away from the center of the zone do not result in a change in $a$.
 
@@
 From this we measure unbroken tract lengths to the right of the selected locus.
-These were compared to the output of neutral simulations with $s=0$\\
+These were compared to the output of neutral simulations with $s=0$
 
@@
 
-%Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\
+%Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.
 
@@
 %FP135& TU&11.748\\
-
 %\end{tabular}
@@
 Figure X shows the estimated probability density (pdf), of unbroken tract of ancestry B surrounding the selected locus, conditioning on a haplotype being of ancestry B at the selected locus.
-Compared to the expected density of tract lengths surrounding neutral loci, tracts of ancestry B around the selected locus are longer.\\
+Compared to the expected density of tract lengths surrounding neutral loci, tracts of ancestry B around the selected locus are longer.
 
@@
 
-\bibliographystyle{dinat}
+\bibliographystyle{molecularEcology}
 \bibliography{library}

### 2016-01-29 drafts/hybrid_zone.tex
minor changes

@@
 \section{Introduction}
-The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}.
-When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting a hybrid zone \citep{Barton1985}.
+The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.
+When formerly isolated populations come back into contact, a gradient of ancestry can form across geography, resulting a hybrid zone \citep{Barton1985}.
 In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD).
@@
 While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.
-These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
+These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
 As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
@@
 In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
+
+Something about how we just study the selection against hets model,
+but we expect general conclusions to apply to other situations.
 

### 2016-01-30 drafts/hybrid_zone.tex
started with theory

@@
 
+To fix notation, we will say that species $A$ was on the ``left'' of the zone of contact,
+which corresponds to spatial positions $x<0$.
+
 We consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past.
@@
 
+%%%%%%%% %%%%%%%%%%
+\subsection{Analysis}
+
+Consider a single locus, with alternate alleles fixed in the two parental populations,
+for which heterozygotes have fitness $1-s$, and homozygotes are of equal fitness.
+The deterministic theory predicts that after secondary contact,
+these alleles will form a stable cline.
+To study gene flow past this cline we first need to understand how the cline itself is formed.
+Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
+Suppose that secondary contact occurred at time $t=0$, 
+and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
+(In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
+Let $\sigma^2$ be the mean squared distance between parent and offspring.
+Assuming local random assortment of alleles into diploids,
+that the habitat and dispersal are homogeneous,
+and that $s$ and $\sigma$ are small, 
+the commonly-used equation that $p$ solves is
+\begin{align} \label{eqn:cline_pde}
+    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
+\end{align}
+where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
+as derived in \citep{bazykin}; for discussion of diploidy see \citep{diploidcline}.
+There is not an exact analytic solution to this equation, 
+but a good approximation to the equilibrium solution is that
+$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{bazykin}.
+Considering the numerous approximations going into this,
+the important attributes are that 
+the stable cline has width of order $\sigma/\sqrt{s}$,
+and decays exponentially.
+Furthermore, the cline is established through diffusion.
+This is slowed down somewhat by selection against heterozygotes,
+but is still diffusive, and so it will take time of order $1/s$
+to reach its stable width.
+(To confirm this, 
+rescale space by $\sigma/s$ and time by $1/s$ in equation \eqref{eqn:cline_pde},
+and see that the result doesn't depend on $\sigma$ or on $s$.)
+
+Now suppose that the frequency profile of the selected allele $p(x,t)$ is given.
+We will understand how lineages of sampled individuals behave,
+both at and away from the selected site,
+by following lineages as they move between backgrounds,
+as done for instance in \citet{selectioncoal,ralph2015patchy}.
+Consider the collection of $A$ alleles found at location $x$.
+The expected number of such alleles produced by location $y$ is proportional to 
+the number of $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
+looking in the other direction, this is also the probability that an $A$ allele at $x$ had a parent at $y$.
+In other words, lineages move as the random walk determined by dispersal,
+but biased towards areas that produce more of the selected allele that they carry.
+Making the same assumptions underlying equation \eqref{eqn:cline_pde},
+we show in Appendix XXX
+that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
+with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$).
+Since $p=1$ far from the zone to the left, 
+and is proportional to $\exp(-x\sqrt{s}/\sigma)$ to the right,
+then roughly, lineages on ``their own'' side wander randomly,
+while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
+back towards the side where they are more common.
+Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
+at the time of secondary contact,
+this push must get more intense the closer it is to the time of secondary contact.
+Lineages of $B$ alleles do the same thing, of course, but in the opposite direction.
+
+The behavior of a lineage at a linked locus is similar,
+except that when in heterozygotes the lineage may move between backgrounds:
+if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
+it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
+but if it spends time in the clinal region where recombination may happen,
+it may be found to have been inherited from a $B$-carrying individual,
+whose ancestors will tend to be more from the right.
+Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
+If $X_t$ is the location of its ancestor $t$ generations ago,
+and $Z_t$ is the identity of the selected allele that ancestor carried,
+then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$,
+and that $Z$ jumps between $A$ and $B$ at rate either $r p$ or $r(1-p)$,
+depending on the identity of $Z$.
+In symbols,
+\begin{align}
+    \begin{aligned}
+        dX_t &= \begin{cases}
+            & s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
+            & s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
+        \end{cases} \\
+        \P\{ Z_{t+\epsilon} &= B \given Z_t = A \} = \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
+        \P\{ Z_{t+\epsilon} &= A \given Z_t = B \} = \epsilon r p(X_t,t) + O(\epsilon^2)  .
+    \end{aligned}
+\end{align}
+
+\textbf{illustrative figure of a lineage moving back and forth?}
+
+Clines at linked sites are given by
+the probability that 
+an allele sampled $t$ generations after secondary contact at location $x$,
+at recombination distance $r$ to a selected allele of type $z$
+is inherited from an individual of ancestry $A$,
+where $z$ can be either $A$ or $B$.
+Denote this probability $q_r(x,t,z)$.
+In other words,
+\begin{align}
+    q_r(x,t,z) = \P^x \{Z_t = A\} .
+\end{align}
+The description above implies that $q_r$ solves the equation
+\begin{align}
+    \dot q_r(x,t,A) &= \grad \log(p(x,t)) \cdot \grad q_r(x,t,A) 
+            + \frac{\sigma^2}{2} \Delta q_r(x,t,A)
+            + r (1-p(x,t))(q_r(x,t,B)-q_r(x,t,A)) \\
+    \dot q_r(x,t,B) &= \grad \log(1-p(x,t)) \cdot \grad q_r(x,t,B) 
+            + \frac{\sigma^2}{2} \Delta q_r(x,t,B)
+            + r (1-p(x,t))(q_r(x,t,A)-q_r(x,t,B))  ,
+\end{align}
+with boundary conditions $q_r(x,0,A)=1$ and $q_r(x,0,B)=0$.
+
+%%%%%%%% %%%%%%%%%%
 \subsection{Simulations}

### 2016-01-30 notes/cline-fns.R

@@ -272,7 +272,7 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
         fwds.grid=extend_grid(xgrid),  
         fwds.soln=forwards_pde(s=s,times=times,grid=fwds.grid,
             yinit=ifelse(fwds.grid$x.mid==0,(yinit[1:fwds.grid$N]+yinit[fwds.grid$N+(1:fwds.grid$N)])/2,
-                         ifelse(fwds.grid$x.mid<0,yinit[1:fwds.grid$N],yinit[fwds.grid$N+(1:fwds.fwds.grid$N)])) 
+                         ifelse(fwds.grid$x.mid<0,yinit[1:fwds.grid$N],yinit[fwds.grid$N+(1:fwds.grid$N)])) 
         ),
         yinit=rep( c( rep(1.0,xgrid$N), rep(0.0,xgrid$N) ), rgrid$N*(rgrid$N+1)/2 ),
         eps=1e-16, ... ) {

### 2016-01-30 drafts/hybrid_zone.tex
more analysis

@@
 \geometry{a4paper} % or letter or a5paper or ... etc
-% \geometry{landscape} % rotated page geometry
+\usepackage{graphicx}
+\usepackage{amsmath,amssymb}
 \usepackage{natbib}
 \usepackage{color}
-% See the ``Article customise'' template for come common customisations
+
 \newcommand{\alisa}[1]{{\em \color{red} #1}}
+\newcommand{\plr}[1]{{\em \color{blue} #1}}
+
+\newcommand{\E}{\mathbb{E}}
+\renewcommand{\P}{\mathbb{P}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\one}{\mathbf{1}}
+\DeclareMathOperator{\sgn}{sgn}
+\newcommand{\grad}{\nabla}
+
+\newcommand{\given}{\,\vert\,}
+\newcommand{\st}{\,\colon\,}
+\renewcommand{\and}{\,\&\,}
 
@@
 but we expect general conclusions to apply to other situations.
+
+The tracking lineages given selection trick is an example of the more general idea of duality for particle systems.
+It also assumes a large pop density limit that excludes recombination and makes lineages Markov;
+if things are very patchy this might not be a good one.
+Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
 
@@
 
+\plr{The aim here is to give both the heuristics, for how a lineage moves around
+and what that tells us about the scale on which clines happen;
+as well as writing down the PDE.  These should probably be better separated?}
+
+Perhaps start out with the main points:
+the cline takes time $s$ to spread out,
+and after time $\tau$ will have clines out to genetic distance of order $1/\tau$,
+maybe more.
+
+\paragraph{Establishment of the selected cline}
 Consider a single locus, with alternate alleles fixed in the two parental populations,
@@
 
+\paragraph{Lineages at the selected locus}
 Now suppose that the frequency profile of the selected allele $p(x,t)$ is given.
@@
 Making the same assumptions underlying equation \eqref{eqn:cline_pde},
-we show in Appendix XXX
+we show in Appendix \ref{apx:lineage_derivation}
 that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
@@
 
+\paragraph{Lineages at linked loci}
 The behavior of a lineage at a linked locus is similar,
@@
 \begin{align}
-    \begin{aligned}
+    \begin{aligned} \label{eqn:lineage_motion}
         dX_t &= \begin{cases}
-            & s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
-            & s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
+             s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
+             s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
         \end{cases} \\
-        \P\{ Z_{t+\epsilon} &= B \given Z_t = A \} = \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
-        \P\{ Z_{t+\epsilon} &= A \given Z_t = B \} = \epsilon r p(X_t,t) + O(\epsilon^2)  .
+        \P\{ Z_{t+\epsilon} = B \given Z_t = A \} &= \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
+        \P\{ Z_{t+\epsilon} = A \given Z_t = B \} &= \epsilon r p(X_t,t) + O(\epsilon^2)  .
     \end{aligned}
@@
 
-\textbf{illustrative figure of a lineage moving back and forth?}
+\textbf{figure of a lineage moving back and forth?}
 
+\paragraph{How fast do linked clines relax?}
+Although loci not under selection can in principle spread across the cline unimpeded,
+in practice it can take quite some time for even unlinked neutral loci to homogenize,
+due to the decreased fitness of heterozygotes \citep{bartonbarrier}
+and the relative slowness of diffusive movement \citep{sedghifar}.
+Suppose we sample across at cline $\tau$ units of time after secondary contact.
+How far away from the selected locus do we expect to see clines that have begun to flatten out more
+than the selected cline?
+A linked cline will have begun to flatten out if there is an appreciable chance that
+a lineage at the linked locus initially linked to an $A$ allele 
+at some point has recombined from an individual carrying the $B$ allele.
+This implies that loci closer than $1/\tau$ to the selected site
+certainly won't have a substantially wider cline than the selected cline.
+\plr{Figure showing this?}
+Of course, this calculation assumes that every generation provides an opportunity for recombination 
+onto the other background,
+but only recombinations in heterozygotes will move the linked lineage between selected backgrounds.
+Since heterozygotes for the selected allele are only found at high frequency in the cline,
+and lineages are generally pushed away from the cline but have no bias far away,
+the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
+and so the width of the genomic region showing clines about the selected locus could be 
+substantially larger than $1/\tau$.
+However, this distinction appears hard to observe for realistic parameter values.
+
+\paragraph{PDE for linked clines}
 Clines at linked sites are given by
@@
 where $z$ can be either $A$ or $B$.
-Denote this probability $q_r(x,t,z)$.
+Denote this probability $q_z(x,t,r)$.
 In other words,
 \begin{align}
-    q_r(x,t,z) = \P^x \{Z_t = A\} .
+    q_z(x,t,r) = \P^x \{Z_t = A\} .
 \end{align}
@@
 \begin{align}
-    \dot q_r(x,t,A) &= \grad \log(p(x,t)) \cdot \grad q_r(x,t,A) 
-            + \frac{\sigma^2}{2} \Delta q_r(x,t,A)
-            + r (1-p(x,t))(q_r(x,t,B)-q_r(x,t,A)) \\
-    \dot q_r(x,t,B) &= \grad \log(1-p(x,t)) \cdot \grad q_r(x,t,B) 
-            + \frac{\sigma^2}{2} \Delta q_r(x,t,B)
-            + r (1-p(x,t))(q_r(x,t,A)-q_r(x,t,B))  ,
+    \begin{aligned}  \label{eqn:q_pde}
+    \dot q_A(x,t,r) 
+            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
+                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
+            \\ &\qquad {} + 
+                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
+    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
+            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
+            \\ &\qquad {} + 
+            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
+    \end{aligned} \\
 \end{align}
-with boundary conditions $q_r(x,0,A)=1$ and $q_r(x,0,B)=0$.
+with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$.
+We will have more use for the differential operators on the right-hand sides of these equations,
+so write these also as 
+$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$
+and 
+$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$.
+
+\paragraph{Numerical computation}
+To determine how the cline at a linked locus is expected to relax,
+we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically,
+producing, for instance, a heatmap of expected local frequencies of $A$ ancestry
+across space and time, as in Figure \ref{fig:pde_clines}.
+Shown is the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
+which we denote $p(x,t,r)$ and can compute as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$.
+
+%%%%%%
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.7\textwidth]{figs/example_cline}
+    \end{center}
+    \caption{
+        \textbf{Numerically computed clines,} at
+        \textbf{(A)} the selected site, which forms over time $1/s$ and then remains stable, and
+        \textbf{(B)} a linked site at distance $r=0.1$, which continues to smooth out as $\sqrt{t}$.
+        Other parameters.
+        \plr{Do we want this figure? This one is just a placeholder; I'd put in two, that showed the formation and spread better.
+        Maybe putting in simulation results would be better, with comparison to numerical solutions? But, this is easier to look at?}
+        \label{fig:pde_clines}
+    }
+\end{figure}
+
+\paragraph{Haplotype lengths}
+We can also find the probability of seeing an entire block of genome (haplotype)
+of a single ancestry at a given location and time.
+As in \citet{sedghifar},
+to do this we need only keep track of both lineages when a recombination in the middle of a block occurs,
+so that lineages will act like a particular class of labeled, branching diffusions,
+where the total branching rate is conserved.
+We will not consider subsequent coalescence.
+The general description of the process is as follows.
+The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
+with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
+(If the selected site lies inside the segment, $r=0$.)
+Additionally, at rate equal to the genetic length of the segment,
+a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
+each of which procedes as before, independently.
+
+\paragraph{Spatial haplotype patterns}
+Furthermore, we expect that stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
+because they have had more opportunities to recombine with $B$ haplotypes.
+However, stretches of $A$ ancestry that contain the selected site are expected to be longer
+than those that do not at the same spatial location,
+because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
+concretely, they move at speed roughly $\sigma/\sqrt{s}$,
+so we expect (selected) $A$ alleles at distance $x$ from the cline center
+to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago,
+and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
+\plr{(check predicted decay of mean length $1/x$ in simulations)}
+For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
+
+\paragraph{General haplotype probabilities}
+To quantify haplotype abundances,
+we'd like to know the probability that the entire segment of genome between positions $a$ and $b$ relative to the selected site
+in an individual sampled at spatial location $x$ and time $t$ after the cline has formed
+is of ancestry $A$.
+Writing this probability $g_A(x,t;a,b)$,
+branching adds a term to equation \eqref{eqn:q_pde}, so that 
+\plr{write this equation more compactly?}
+\begin{align}
+    \begin{aligned} \label{eqn:g_pde}
+        \dot g_A(x,t;a,b) 
+            &= G_{r(a,b)}^A g_A(x,t;a,b) 
+            - (b-a) g_A(x,t;a,b) 
+            \\ {} & \qquad 
+            + p(x,t) \int_a^b {
+                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
+                - g_A(x,t;a,b) .
+            } d\theta
+            \\ {} & \qquad 
+            + (1-p(x,t)) \int_a^0 {
+                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
+            } d\theta
+            \\ {} & \qquad 
+            + (1-p(x,t)) \int_0^b {
+                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
+            } d\theta,
+    \end{aligned}
+\end{align}
+where the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
+and $r(a,b)$ is the distance from the segment to the selected site:
+$r(a,b)=0$ if $ab<0$ and $r(a,b)=\min(|a|,|b|)$ otherwise.
+The equation for $g_B(x,t;a,b)$ is identical after exchanging $A \leftrightarrow B$ and
+$p \leftrightarrow (1-p)$.
+The boundary conditions are $g_A(x,0;a,b)=1$ and $g_B(x,0;a,b)=0$.
+
+\paragraph{Numerical solutions}
+Notice that equations \eqref{eqn:g_pde}
+are hierarchical in $[a,b]$:
+the equation for haplotype identity probabilities on a segment $[a,b]$ depends only on those probabilities for smaller segments.
+This allows for efficient numerical solutions,
+described in more detail in Appendix \ref{apx:haplotype_calcs}.
+
+%%%%%%
+\begin{figure}
+    \begin{center}
+    \end{center}
+    \caption{
+        \textbf{Haplotype length distributions},
+        comparing numerical solutions to simulations.
+        \plr{maybe?}
+        \label{fig:haplotype_lengths}
+    }
+\end{figure}
+
+\paragraph{Correlations in ancestry}
+To compute correlations in local ancestry
+(i.e., ``ancestry disequilibrium'', as in \citet{pool,brandvainthisvolume}),
+we need only follow lineages at two sites, instead of an entire region,
+the model described above requires only minor modifications,
+and our numerical code can be used directly.
 
@@
 
+Numerical solution for the PDE can be used on real, non-flat landscapes that change with time;
+we have code and notes on solving them on github.
+
 \bibliographystyle{molecularEcology}
@@
 
+\appendix
+
+\section{Rescaling a dicrete model to obtain the lineage motion}
+\label{apx:lineage_derivation}
+
+Here put derivation of the PDE for $p(x,t)$ and also the SDE of a lineage motion.
+
+\section{Nuemrical solutions for haplotype probabilities}
+\label{apx:haplotype_calcs}.
+
+Here put some details about the numerical solutions of haplotype probabilities.
+
 \end{document}

### 2016-01-30 notes/rough-calculations.Rmd
more analysis

@@
     where $C$ is the constant;
-    this says that if clines down to $r$ are flat at $\tau$, then at time $2\tau$ they'll be flat down to $r/4$.
+    this says that if clines down to $r$ are flat at $\tau$, then at time $4\tau$ they'll be flat down to $r/2$.
 

### 2016-02-01 sims/potential_statistics.R
smaller selection figures

@@ -50,7 +50,7 @@ intervalSizes_allAncs = lapply(positions,function(POS){
 #[[POS]][ind,chr]
 
 #COMPARE:
-pdf(file=paste(outstring,"statistic_length.pdf",sep=""),height = 3,width=5)
+pdf(file=paste(outstring,"statistic_length.pdf",sep="_"),height = 3,width=5)
 #mean length across ALL indivduals in the zone (including non-recombinant chroms) vs chromosome position
 
 plot(positions,sapply(intervalSizes_allAncs,mean),type="l",xlab="physical position",ylab="mean length",main="mean length, whole zone")
@@ -79,7 +79,7 @@ for(i in 1:50){points(mean_per_deme_AncB[,i]/mean(mean_per_deme_AncB[,i],na.rm=T
 
 dev.off()
 
-pdf(file=paste(outstring,"statistic_ecdf.pdf",sep=""),height = 3,width=5)
+pdf(file=paste(outstring,"statistic_ecdf.pdf",sep="_"),height = 3,width=5)
 
 #	2. score = $\log$ of ECDF of length (i.e. log of prob that haplotype is longer than this one)
 #relative to ALL individuals in ALL populations:
@@ -126,7 +126,7 @@ dev.off()
 
 #For ALL inds:
 
-pdf(file=paste(outstring,"statistic_length_of_adjacent_blocks.pdf",sep=""),height = 3,width=5)
+pdf(file=paste(outstring,"statistic_length_of_adjacent_blocks.pdf",sep="_"),height = 3,width=5)
 get.flanking.blocks.all = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=chromosome,POS=0.5,ancB=TRUE){
 	focal_chunks = do.call(rbind,lapply(IND_DATA[[CHR]],function(X){
 		FOCUS = which(X$starts<POS & X$stops>POS);

### 2016-02-03 drafts/hybrid_zone.tex
smaller selection figures
notes from talking to alisa

@@
 certainly won't have a substantially wider cline than the selected cline.
-\plr{Figure showing this?}
+\plr{Figure showing this? 
+    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
+    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
+    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
+}
 Of course, this calculation assumes that every generation provides an opportunity for recombination 
@@
         \textbf{(A)} the selected site, which forms over time $1/s$ and then remains stable, and
-        \textbf{(B)} a linked site at distance $r=0.1$, which continues to smooth out as $\sqrt{t}$.
+        \textbf{(B)} a linked site at distance $r=0.01$, which continues to smooth out as $\sqrt{t}$.
+        \textbf{(C)} a linked site at $r=0.1$
+        \textbf{(D)} an unlinked site.
         Other parameters.
        \plr{Do we want this figure? This one is just a placeholder; I'd put in two, that showed the formation and spread better.
-        Maybe putting in simulation results would be better, with comparison to numerical solutions? But, this is easier to look at?}
+       This is from the PDE (we don't have time courses from sims).}
         \label{fig:pde_clines}
@@
         comparing numerical solutions to simulations.
+        At one time, plot 
+        probability of seeing a haplotype of length at least $r$ covering that position on the genome,
+        against (a) position on the genome, at a few spatial locations;
+        or (b) spatial location, at a few positions on the genome.
         \plr{maybe?}
@@
 
+\subsection{Identifying sites under selection}
+
+Plots with x-axis the genome and having peaks at the selected site.
+
+Question: how much resolution do we have?
+
 \section{Discussion}

### 2016-02-04 drafts/hybrid_zone.tex

@@
 \section{Methods}
-\subsection{Model}
+\subsection*{Model}
 We consider a model in which two isolated subspecies, species $A$ and species $B$ come into contact at time $\tau$ generations in the past.
@@
 %%%%%%%% %%%%%%%%%%
-\subsection{Analysis}
+\subsection*{Analysis}
 
@@
 %%%%%%%% %%%%%%%%%%
-\subsection{Simulations}
+\subsection*{Simulations}
 Forward in time simulations were used to assess theoretical predictions.
@@
 
-\subsection{Statistics}
+\subsection*{Statistics}
 With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci.
@@
 
-%\subsection{Data}
+%\subsection*{Data}
 %We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m.
@@
 \section{Results}
-\subsection{The distribution of ancestry tract lengths}
+\subsection*{The distribution of ancestry tract lengths}
 We are interested specifically in the distribution of the length of continuous tracts of ancestry surrounding the selected locus.
@@
 
-\subsection{Number of ancestors}
+\subsection*{Number of ancestors}
 Not sure what the result is here yet -- unexpected
 
-\subsection{Identifying sites under selection}
+\subsection*{Identifying sites under selection}
 

### 2016-02-04 sims/compare-to-theory.Rmd

---
title: "Comparison of simulation to theory"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
source("sim-fns.R")
```

Here's the things that we'd like to have plots of simulation compared to theory for,
and might put in the paper:

1.  **Establishment of the selected cline:** the frequency of the selected allele, $p(x,t)$,
    across space, at a few times.

2.  **Conditional ancestry probabilities:** the conditional frequency of ancestry $A$
    given being linked to a selected allele of ancestry $z$, denoted $q_z(x,t,r)$,
    across space, at a few times or distances and for each $z$.

3.  **Relaxation of linked clines:** the frequency of ancestry $A$, denoted $p(x,t,r)$, 
    across space, at a few times and one or two recombination distances. 
    (This comes after $q$ because it is computed using $q$.)

4.  **Haplotype probabilities:** the probability of seeing a haplotype of length at least $\ell$
    covering position $r$ on the genome at location $x$ in space: for fixed $\ell$,
    against $r$ and/or $x$.

### 2016-02-04 sims/run-sim.Rmd

@@
 
+Here `sim` is a list with five named components:
+
+- **inds** is a list of length `ninds`, 
+    - each element of which is a list, one element per chromosome,
+        - each element of which is a list with names "X1" and "X2", corresponding to the two copies of the chromosome,
+            - each element of which is a data frame, with columns:
+                - **pos**: coordinates of recombination
+                - **ancest**: ancestor ID
+- **sp.inds** is a character vector of length `ninds`, giving the ancestral type of each ancestor (i.e., ancestries corresponding to the the ancestor IDs)
+- **deme** is an integer vector of length `ninds`, giving the deme number of each individual (including the ancestors)
+- **pars** is an integer that gives the number of generations of the simulation
+- **QTL** returns the QTL vector used in the simulation
+
+and `sim.sums` is a named list with components:
+
+- **ind.ancestry**: a list of length `ninds`
+    - each element of which is a list, one element per chromosome,
+        - each element of which is a list with names "X1" and "X2", corresponding to the two copies of the chromosome,
+            - each element of which is a data frame, with variables
+                - **starts**: start position of a chunk of ancestry
+                - **stops**: end position of the chunk of ancestry
+                - **sp2**: logical variable saying whether the chunk is of ancestry $A$ or not
+- **ancest.prop**: is a `ninds*2` matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
+
 OK, now let's genotype these individuals at some loci:

### 2016-02-04 drafts/hybrid_zone.tex
simulation methods rewritten

@@
 \subsection*{Simulations}
-Forward in time simulations were used to assess theoretical predictions.
-We simulate discrete demes with nearest-neighbor migration and non-overlapping generations.
-Each generation, the dispersal distance of an individual is sampled from a normal distribution, and then discretized to correspond to a deme that is a corresponding distance away.
-Following migration, the next generation is produced by replacing each individual in a deme by randomly choosing two parents with probability weighted by the fitness of their genotypes.
-The genotype of the new individual is generated by simulating a recombined chromosome from each parent.
-This simulation keeps track of the ancestry at each location in the chromosome for all individuals.
-Each individual has 2 pairs of chromosomes each of length 1 Morgan such that chromosome 1 contains a selected locus at position 0.5M, and chromosome 2 contains no selected loci.
+
+We implemented forwards-time simulations of diploid individuals on a one-dimensional row of demes
+with nonoverlapping generations and fixed population sizes (a Wright-Fisher model),
+and a single locus with selection against heterozygotes in the center of one of two chromosomes, each of length 1 Morgan.
+Each deme has exactly $n_d$ diploid individuals at the start of each generation.
+Each generation,
+first, each individual migrated to a new deme offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
+where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
+(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
+Migrants past either end of the range remained at the ends.
+Then, 
+fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
+and in each deme $n_d$ pairs of parents are chosen, with replacement,
+with probability proportional to their fitness.
+Selfing is allowed.
+Since migration is not conservative, demes may have no available parents;
+in this case, parents are chosen from other demes with probability proportional to fitness
+multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
+The next generation is formed by carrying out meiosis in each parent
+and combining the resulting gametes.
+Meiosis results in alternating blocks of the gamete's chromsome
+being inherited from the two parental chromosomes,
+with the blocks separated by a Poisson(1) number of uniformly chosen recombination points along the chromosome,
+and the order of the parental chromosomes chosen randomly.
+The simulation software works by recording for each chromosome
+a list of ancestry breakpoints and the index of the ancestor in the first generation
+that the chromosome inherited that region of genome from;
+genotyping at individual loci was performed by looking up the ancestor's genotype.
 

### 2016-02-04 sims/compare-to-theory.Rmd
simulation methods rewritten

@@
     against $r$ and/or $x$.
+
+# Parameters from a simulation
+
+We'll use some simulation results produced by [yaniv_sims.R](yaniv_sims.R),
+and need to convert those parameters to those from theory.
+Dispersal in the simulation moves as $\lfloor Z + 1/2 \rfloor$,
+where $Z \sim N(0,\sigma^2_\text{demes})$;
+we'll use this fact to compute the actual $\sigma$.
+(Yes, that's an unbiased walk, thanks to the floor.)
+Migrants past the endpoints stay there (they don't reflect).
+
+```{r sim_results}
+simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums_chunks.Robj"
+sim.params <- list( sigma=1, ninds=25000, ndemes=50, s=0.1, tau=1000 )
+
+theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau )
+```

### 2016-02-04 sims/plot_compare_chunk_ecdf.R

@@ -5,15 +5,7 @@ load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000
 #this is only around the selected locus. The object name is "chunks_by_deme"
 
 
-add.alpha <- function(col, alpha=1){
-if(missing(col))
-stop("Please provide a vector of colours.")
-apply(sapply(col, col2rgb)/255, 2, 
-function(x) 
-rgb(x[1], x[2], x[3], alpha=alpha)) 
-
-}
-transparent_rainbow = add.alpha(rainbow(ndemes),0.7)
+transparent_rainbow = adjustcolor(rainbow(ndemes),0.7)
 
 
 empty_deme = rep(0,2*deme_size)

### 2016-02-04 sims/sim-fns.R

@@ -89,12 +89,12 @@ genoAll = function(inds,QTL,sp.ids){
 
 #######    MAKING A NEW GEN     #######
 
-makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),min.deme=min(demes)){
+makeNewGen=function(inds,QTL,sp.ids,demes,sigma,max.deme=max(demes),min.deme=min(demes)){
     pop.sizes <- table(demes) # number of individuals in each deme
     chrnames <- namesList( names(inds[[1]]) )
     #genotyping
     if(!is.null(QTL)){ 
-        genos = lapply(genoAll(inds=inds,QTL=qtl,sp.ids=sp.ids) ,function(IND){do.call(rbind,IND)})  
+        genos = lapply(genoAll(inds=inds,QTL=qtl,sp.ids=sp.ids), function(IND){do.call(rbind,IND)})  
         QTL=do.call(rbind,QTL)
         #make fitness from genotypes
         w_abs = sapply(genos,function(IND){
@@ -112,7 +112,7 @@ makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),m
     }
     if(is.null(QTL)){ w_abs = rep(1,length(inds)) }
     rel.w = w_abs/mean(w_abs)
-    inds.df = data.frame(w=rel.w,deme=demes,edge=ifelse(demes%in%range(demes),2,1))
+    inds.df = data.frame(w=rel.w,deme=demes)
     # sample the displacements of individuals
     sample_move = floor(rnorm(NROW(inds.df),0.5,sigma))
     # new locations of individuals
@@ -120,11 +120,12 @@ makeNewGen=function(inds,QTL,sp.ids,demes,edge="M/2",sigma,max.deme=max(demes),m
     inds.df$id = seq_along(inds.df$new.deme)
     new.inds.df = do.call(rbind,lapply(  unique(inds.df$deme), function(d){	# iterate over demes
                 these.parents = which(inds.df$new.deme==d)
-                # HACK: if there's no available parents, pick ome
+                # if there's no available parents, pick one from a Gaussian displacement
                 while (length(these.parents)==0) {
                     parent.probs = inds.df$w*exp(-(inds.df$deme-d)^2/sigma)
                     these.parents = sample.int(nrow(inds.df),1,prob=parent.probs)
                 }
+                # otherwise, pick parents from the same deme
[…]

### 2016-02-04 sims/yaniv_sims.R

@@ -4,16 +4,6 @@
 
 source("sim-fns.R")
 
-add.alpha <- function(col, alpha=1){
-if(missing(col))
-stop("Please provide a vector of colours.")
-apply(sapply(col, col2rgb)/255, 2, 
-function(x) 
-rgb(x[1], x[2], x[3], alpha=alpha)) 
-
-}
-
-
 #PARAMS for simulation:
 
 SIGMA = 1
@@ -22,7 +12,7 @@ deme_size = 500
 ninds = ndemes*deme_size
 S = 0.1
 
-transparent_rainbow = add.alpha(rainbow(ndemes),0.7)
+transparent_rainbow = adjustcolor(rainbow(ndemes),0.7)
 
 zone_age = c(100)
 
@@ -35,10 +25,18 @@ xx <- (1:ndemes)-0.5-ndemes/2
 
 
 #DEFINE selection against genotypes:
-qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)),chr2=data.frame(traits=c("underdominant"), s = 0,pos=c(0.5)))
+qtl=list(
+        chr1=data.frame(
+                traits=c("underdominant"), 
+                s=S,
+                pos=c(0.5)),
+        chr2=data.frame(
+                traits=c("underdominant"), 
+                s = 0,pos=c(0.5))
+        )
 
-#SIMULATE the populations (NB: Can't make per-deme population size too small, otherwise deme may end up empty and throws up an error.)
-sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
+#SIMULATE the populations
+sims.s0.1 = lapply(zone_age, sim.zone, n.ind=ninds, n.deme=ndemes, sigma=SIGMA)
 	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
 		#where:
 			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]

### 2016-02-04 sims/compare-to-theory.Rmd
.

@@
 
-# Parameters from a simulation
+## Parameters from a simulation
 
@@
 Migrants past the endpoints stay there (they don't reflect).
+We'll work in spatial units equal to one deme spacing;
+since generations are nonoverlapping, one generation is indeed one time step in the simulation.
 
@@
 simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums_chunks.Robj"
-sim.params <- list( sigma=1, ninds=25000, ndemes=50, s=0.1, tau=1000 )
-
-theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau )
+load(simchunks.file)
+get_simparams <- function (fname) {
+    strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
+    as.numeric( sapply( strings, function (x) { gsub(sprintf(".*_%s([0-9.]+)_.*",x),"\\1",simchunks.file) } ) )
+}
+sim.params <- get_simparams( simchunks.file )
+theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau, 
+       density=sim.params$ninds/sim.params$ndemes,
+       width=sim.params$ndemes )
+theory.params
 ```

### 2016-02-04 sims/sim-fns.R

@@ -57,9 +57,10 @@ meiosis = function(CHR,r=1){
     B= breaks[seq_along(breaks)%%2!=start]
     X = rbind(
         do.call(rbind,lapply(A,function(a){
-            with(CHR[[1]],c(pos=a,ancest=ancest[max(which(pos<a))]))})),
+                    with(CHR[[1]],c(pos=a,ancest=ancest[max(which(pos<a))]))})),
             CHR[[1]][(findInterval(CHR[[1]][,"pos"],c(-1,breaks,1.2))%%2)!=start,],
-            do.call(rbind,lapply(B,function(b){with(CHR[[2]],c(pos=b,ancest=ancest[max(which(pos<b)) ]))})),
+            do.call( rbind, lapply(B,function(b){
+                    with(CHR[[2]],c(pos=b,ancest=ancest[max(which(pos<b)) ]))})),
             CHR[[2]][(findInterval(CHR[[2]][,"pos"],c(-1,breaks,1.2))%%2)==start,]
         )
     return(		X[order(X$pos),]		)

### 2016-02-04 sims/compare-to-theory.Rmd
fix in string-to-params fn

--- 2016-02-04 sims/compare-to-theory.Rmd
@@
 simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums_chunks.Robj"
-sim.params <- list( sigma=1, ninds=25000, ndemes=50, s=0.1, tau=1000 )
+load(simchunks.file)  # provides 'chunks'
+get_simparams <- function (fname) {
+    strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
+    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*_%s([0-9.]+)_.*",x),"\\1",simchunks.file)) } )
+}
+sim.params <- get_simparams( simchunks.file )
+theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau, 
+       density=sim.params$ninds/sim.params$ndemes,
+       width=sim.params$ndemes )
+theory.params
+```
 
-theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau )
+# Selected cline
+
+```{r selected_cline_sim}
+
 ```

### 2016-02-05 sims/compare-to-theory.Rmd
begun comparison to theory

--- 2016-02-04 sims/compare-to-theory.Rmd
@@
 source("sim-fns.R")
+library(ReacTran)
+library(Matrix)
+source("../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../notes/cline-fns.R",chdir=TRUE)
 ```
@@
 Migrants past the endpoints stay there (they don't reflect).
+We'll work in spatial units equal to one deme spacing;
+since generations are nonoverlapping, one generation is indeed one time step in the simulation.
 
 ```{r sim_results}
-simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums_chunks.Robj"
-sim.params <- list( sigma=1, ninds=25000, ndemes=50, s=0.1, tau=1000 )
+simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums_chunks.Robj"
+load(simchunks.file)  # provides 'chunks'
+get_simparams <- function (fname) {
+    strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
+    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*_%s([0-9.]+)_.*",x),"\\1",simchunks.file)) } )
+}
+sim.params <- get_simparams( simchunks.file )
+# recombination distances, hard-coded
+sim.rvals <- c(0,0.01,0.5)
+# spatial locations of the demes
+xx <- seq_len(sim.params$ndemes)-sim.params$ndemes/2
+theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau, 
+       density=sim.params$ninds/sim.params$ndemes,
+       width=sim.params$ndemes )
+theory.params
+```
 
-theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau )
+*Note:* We should also work out what the actual value of **s** is.
+
+Here are some functions we'll use to extract what we want
+from data in the `chunks` format:
+```{r helper_fns}
+genotype.counts <- function (ch) { 
+    gcounts <- t(sapply( ch, function (x) { tabulate( 1+rowSums( x>0 ), nbins=3 ) } )) 
+    colnames(gcounts) <- c("AA","AB","BB")
+    return(gcounts)
+}
+dip.to.hap <- rbind( A=c(AA=2,AB=1,BB=0), B=c(AA=0,AB=1,BB=2) )
 ```
+
+# Selected cline
+
+To get the value predicted from theory, we'll solve the corresponding PDE:
+```{r selected_cline_theory}
+tt <- seq(0,theory.params$tau,length.out=21)
+xgrid <- setup.grid.1D(x.up=-theory.params$width/2, x.down=theory.params$width/2, N=2*theory.params$width)
+fgrid <- extend_grid(xgrid)
+fwds.soln <- forwards_pde(s=theory.params$s, times=tt, grid=fgrid, sigma=theory.params$sigma )
+```
+... and, compare it to the observed values:
+```{r selected_cline_sim}
+gcounts <- genotype.counts(chunks[[1]])
+# sim frequencies
+sfreqs <- sweep( tcrossprod(gcounts,dip.to.hap), 1, 2*rowSums(gcounts), "/" )
+# theory frequencies
+tfreqs <- cbind( A=fwds.soln[nrow(fwds.soln),-1], B=1-fwds.soln[nrow(fwds.soln),-1] )
+matplot(xx, sfreqs, type='l', lty=1, main="Selected allele frequency", ylab="frequency", xlab='geographic position')
+matlines(fgrid$x.mid, tfreqs, lty=4, lwd=2)
+legend("topright", legend=outer(colnames(sfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
+```
+
+For fun, let's also look at genotype frequencies (is there HW?):
+```{r selected_geno_freqs}
+matplot(xx, sweep(gcounts,1,rowSums(gcounts),"/"), lty=1, type='l', main="Selected genotype frequencies", ylab="frequency", xlab='geographic position')
+matlines(fgrid$x.mid, cbind(AA=tfreqs[,"A"]^2,AB=2*tfreqs[,"A"]*tfreqs[,"B"],BB=tfreqs[,"B"]^2), lty=4, lwd=2)
+```
+Looks pretty similar, just shifted.

### 2016-02-05 notes/pde-clines.Rmd
bug

@@
 rr <- seq(0,0.5,length.out=20)[-1]
-fwds.soln <- forwards_pde(s=s,times=tt,grid=extend_grid(xgrid))
+fgrid <- extend_grid(xgrid)
+fwds.soln <- forwards_pde(s=s,times=tt,grid=fgrid)
 LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )
@@
 pfun <- function (t,x) { 
-        fp <- function (x,t) { approx(xgrid$x.mid,cline_interp(t,soln=fwds.soln),xout=x)$y }
+        fp <- function (x,t) { approx(fgrid$x.mid,cline_interp(t,soln=fwds.soln),xout=x)$y }
         ans <- numeric(length(x))

### 2016-02-05 sims/compare-to-theory.Rmd
more on comparison to theory
comparison to theory; changed "F" to "FALSE"

@@
 ```{r helper_fns}
-genotype.counts <- function (ch) { 
+genotype_counts <- function (ch) { 
     gcounts <- t(sapply( ch, function (x) { tabulate( 1+rowSums( x>0 ), nbins=3 ) } )) 
@@
 }
+conditioned_genotype_counts <- function (x,y) { 
+    # returns joint genotype counts across space from two chunks objects
+    #  a (ndemes) x 4 x 4 array (4 phased genotypes)
+    deme <- rep(seq_along(x),sapply(x,NROW))
+    x.alleles <- ifelse(do.call(rbind,x)==0,"A","B")
+    y.alleles <- ifelse(do.call(rbind,y)==0,"A","B")
+    x.genotypes <- factor( paste0(x.alleles[,1],x.alleles[,2]), levels=c("AA","AB","BA","BB") )
+    y.genotypes <- factor( paste0(y.alleles[,1],y.alleles[,2]), levels=c("AA","AB","BA","BB") )
+    # x.genotypes <- factor( c("AA","AB","BB")[1+rowSums( do.call(rbind, x)>0 )], levels=c("AA","AB","BB") )
+    # y.genotypes <- factor( c("AA","AB","BB")[1+rowSums( do.call(rbind, y)>0 )], levels=c("AA","AB","BB") )
+    return( table( deme, x=x.genotypes, y=y.genotypes ) )
+}
 dip.to.hap <- rbind( A=c(AA=2,AB=1,BB=0), B=c(AA=0,AB=1,BB=2) )
+bidip.to.bihap <- rbind(
+            "A-A" = c( 2, 1, 1, 0, 1, 1, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0 ),
+            "A-B" = c( 0, 0, 0, 0, 1, 0, 1, 0, 1, 1, 0, 0, 2, 1, 1, 0 ),
+            "B-A" = c( 0, 1, 1, 2, 0, 0, 1, 1, 0, 1, 0, 1, 0, 0, 0, 0 ),
+            "B-B" = c( 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 0, 1, 1, 2 ) )
+colnames( bidip.to.bihap ) <- as.vector( outer( c("AA","AB","BA","BB"), c("AA","AB","BA","BB"), paste, sep="-" ) )
+conditional_haplotype_counts <- function (gcounts) {
+    # uses output of conditioned_genotype_counts to get conditional haplotype counts
+    # a (ndemes) x 2 x 2 array
+    dim(gcounts) <- c(dim(gcounts)[1],prod(dim(gcounts)[2:3]))
+    out <- tcrossprod( gcounts, bidip.to.bihap )
+    dim(out) <- c(dim(gcounts)[1],2,2)
+    dimnames(out) <- list( dimnames(gcounts)[[1]], c("A","B"), c("A","B") )
+    return(out)
+}
+conditional_freqs <- function (gcounts) {
+    # uses output of conditioned_genotype_counts to get conditional allele frequencies
+    # a (ndemes) x 2 matrix
+    hcounts <- conditional_haplotype_counts(gcounts)
+    cbind( A = hcounts[,"A","A"]/(hcounts[,"A","A"]+hcounts[,"A","B"]),
+           B = hcounts[,"A","B"]/(hcounts[,"A","B"]+hcounts[,"B","B"]) )
+}
 ```
@@
 ```{r selected_cline_sim}
-gcounts <- genotype.counts(chunks[[1]])
+gcounts <- genotype_counts(chunks[[1]])
 # sim frequencies
@@
 matplot(xx, sfreqs, type='l', lty=1, main="Selected allele frequency", ylab="frequency", xlab='geographic position')
-matlines(fgrid$x.mid, tfreqs, lty=4, lwd=2)
+matlines(fgrid$x.mid, tfreqs, lty=2, lwd=2)
 legend("topright", legend=outer(colnames(sfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
@@
 Looks pretty similar, just shifted.
+
+**Question:** Is that shift due to stochasticity, or have I got the center in the wrong place on the sims?
+
+# Ancestry probabilities
+
+First, compute theoretical predictions at the values of $r$ in the simulation:
+```{r linked_soln}
+linked.solns <- lapply( sim.rvals, function (r) {
+        forwards_backwards_pde(s=theory.params$s, times=tt, grid=xgrid, r=r, 
+                               sigma=theory.params$sigma, fwds.soln=fwds.soln, log.p=TRUE)
+    } )
+```
+... and compare to the observed values:
+```{r compare_linked}
+for (kr in seq_along(sim.rvals)[-1]) {
+    gcounts <- conditioned_genotype_counts(chunks[[kr]], chunks[[1]])
+    # frequencies from simulation
+    cfreqs <- conditional_freqs(gcounts)
+    # frequencies from theory
+    tfreqs <- linked.solns[[kr]][nrow(linked.solns[[kr]]),-1]
+    dim(tfreqs) <- c( length(xgrid$x.mid), 2 )
+    matplot(xx, cfreqs, type='l', lty=1, main=sprintf("Frequency of A ancestry by selected background, r=%f",sim.rvals[rk]), ylab="frequency", xlab='geographic position')
+    matlines(xgrid$x.mid, tfreqs, lty=2, lwd=2)
+    legend("topright", legend=outer(colnames(cfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
+}
+```

### 2016-02-05 sims/yaniv_sims.R

@@ -101,12 +101,12 @@ get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,P
 }
 
 #output is a nind*2 matrix, where each column is a chromosome. 
-#EXAMPLE: testB = lapply(1:50,function(X){get.deme.chunks(DEME=X,ancA=F)})
+#EXAMPLE: testB = lapply(1:50,function(X){get.deme.chunks(DEME=X,ancA=FALSE)})
 #D = 23; hist(testB[[D]][which(testB[[D]]<1 & testB[[D]]>0)], col="black",breaks=seq(0,1,0.05))
 
-testB = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=F)})
-testB_far = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=F,POS=0.01)})
-testB_unlinked = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=F,POS=0.5,CHR=2)})
+testB = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=FALSE)})
+testB_far = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=FALSE,POS=0.01)})
+testB_unlinked = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=FALSE,POS=0.5,CHR=2)})
 
 chunks = list(selected=testB,distant=testB_far,unlinked=testB_unlinked)
 save(chunks,file = paste(outfile,"simsums_chunks.Robj",sep=""))

### 2016-02-05 sims/compare-to-theory.Rmd
tested chunks fns

--- 2016-02-05 sims/compare-to-theory.Rmd
@@
 ```{r helper_fns}
-genotype.counts <- function (ch) { 
-    gcounts <- t(sapply( ch, function (x) { tabulate( 1+rowSums( x>0 ), nbins=3 ) } )) 
-    colnames(gcounts) <- c("AA","AB","BB")
-    return(gcounts)
-}
-dip.to.hap <- rbind( A=c(AA=2,AB=1,BB=0), B=c(AA=0,AB=1,BB=2) )
+source("chunks_fns.R")
 ```
@@
 tfreqs <- cbind( A=fwds.soln[nrow(fwds.soln),-1], B=1-fwds.soln[nrow(fwds.soln),-1] )
-matplot(xx, sfreqs, type='l', lty=1, main="Selected allele frequency", ylab="frequency", xlab='geographic position')
-matlines(fgrid$x.mid, tfreqs, lty=4, lwd=2)
+matplot(xx, sfreqs, type='l', lty=1, main=sprintf("Selected allele frequency, t=%d",sim.params$tau), ylab="frequency", xlab='geographic position')
+matlines(fgrid$x.mid, tfreqs, lty=2, lwd=2)
 legend("topright", legend=outer(colnames(sfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
@@
 ```{r selected_geno_freqs}
-matplot(xx, sweep(gcounts,1,rowSums(gcounts),"/"), lty=1, type='l', main="Selected genotype frequencies", ylab="frequency", xlab='geographic position')
+matplot(xx, sweep(gcounts,1,rowSums(gcounts),"/"), lty=1, type='l', main=sprintf("Selected genotype frequencies, t=%d",sim.params$tau), ylab="frequency", xlab='geographic position')
 matlines(fgrid$x.mid, cbind(AA=tfreqs[,"A"]^2,AB=2*tfreqs[,"A"]*tfreqs[,"B"],BB=tfreqs[,"B"]^2), lty=4, lwd=2)
@@
 Looks pretty similar, just shifted.
+
+**Question:** Is that shift due to stochasticity, or have I got the center in the wrong place on the sims?
+
+# Ancestry probabilities
+
+First, compute theoretical predictions at the values of $r$ in the simulation:
+```{r linked_soln}
+linked.solns <- lapply( sim.rvals, function (r) {
+        forwards_backwards_pde(s=theory.params$s, times=tt, grid=xgrid, r=r, 
+                           sigma=theory.params$sigma, fwds.soln=fwds.soln, log.p=TRUE)
+    } )
+```
+... and compare to the observed values:
+```{r compare_linked}
+for (kr in seq_along(sim.rvals)[-1]) {
+    gcounts <- conditional_genotype_counts(chunks[[kr]], chunks[[1]])
+    # frequencies from simulation
+    cfreqs <- conditional_freqs(gcounts)
+    # frequencies from theory
+    tfreqs <- linked.solns[[kr]][nrow(linked.solns[[kr]]),-1]
+    dim(tfreqs) <- c( length(xgrid$x.mid), 2 )
+    matplot(xx, cfreqs, type='l', lty=1, 
+            main=sprintf("Frequency of A ancestry by selected background, t=%d, r=%0.3f",sim.params$tau, sim.rvals[kr]), 
+            ylab="frequency", xlab='geographic position')
+    matlines(xgrid$x.mid, tfreqs, lty=2, lwd=2)
+    legend("topright", legend=outer(colnames(cfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
+}
+```
+
+To help interpret this, here's the raw counts of the four phased haplotypes across the range:
+(labels are "linked-selected", for instance, "B-A" has B ancestry linked to the selected A allele)
+```{r hap_counts}
+for (kr in seq_along(sim.rvals)[-1]) {
+    hcounts <- conditional_haplotype_counts( conditional_genotype_counts(chunks[[kr]], chunks[[1]]) )
+    hc.names <- outer( dimnames(hcounts)[[2]], dimnames(hcounts)[[3]], paste, sep="-")
+    dim(hcounts) <- c( dim(hcounts)[1], prod(dim(hcounts)[2:3]) )
+    colnames(hcounts) <- hc.names
+    matplot(xx, hcounts, type='l', 
+            main=sprintf("Haplotype counts (linked-selected), t=%d, r=%0.3f", sim.params$tau, sim.rvals[kr]), 
+            ylab="count", xlab="geographic position")
+    legend('topleft', lty=1:5, col=1:6, legend=hc.names )
+}
+```

### 2016-02-05 sims/chunks_fns.R

@@ -1,0 +1,113 @@
+genotype_counts <- function (ch) { 
+    gcounts <- t(sapply( ch, function (x) { tabulate( 1+rowSums( x>0 ), nbins=3 ) } )) 
+    colnames(gcounts) <- c("AA","AB","BB")
+    return(gcounts)
+}
+
+conditional_genotype_counts <- function (x,y) { 
+    # returns joint genotype counts across space from two chunks objects
+    #  a (ndemes) x 4 x 4 array (4 phased genotypes)
+    deme <- rep(seq_along(x),sapply(x,NROW))
+    x.alleles <- ifelse(do.call(rbind,x)==0,"A","B")
+    y.alleles <- ifelse(do.call(rbind,y)==0,"A","B")
+    x.genotypes <- factor( paste(x.alleles[,1],x.alleles[,2],sep=""), levels=c("AA","AB","BA","BB") )
+    y.genotypes <- factor( paste(y.alleles[,1],y.alleles[,2],sep=""), levels=c("AA","AB","BA","BB") )
+    return( as.array( table( deme, x=x.genotypes, y=y.genotypes ) ) )
+}
+
+dip.to.hap <- rbind( A=c(AA=2,AB=1,BB=0), B=c(AA=0,AB=1,BB=2) )
+
+bidip.to.bihap <- rbind(
+            "A-A" = c( 2, 1, 1, 0, 1, 1, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0 ),
+            "B-A" = c( 0, 1, 1, 2, 0, 0, 1, 1, 0, 1, 0, 1, 0, 0, 0, 0 ),
+            "A-B" = c( 0, 0, 0, 0, 1, 0, 1, 0, 1, 1, 0, 0, 2, 1, 1, 0 ),
+            "B-B" = c( 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 0, 1, 1, 2 ) )
+colnames( bidip.to.bihap ) <- as.vector( outer( c("AA","AB","BA","BB"), c("AA","AB","BA","BB"), paste, sep="-" ) )
+
+conditional_haplotype_counts <- function (gcounts) {
+    # uses output of conditional_genotype_counts to get conditional haplotype counts
+    # a (ndemes) x 2 x 2 array
+    dim(gcounts) <- c(dim(gcounts)[1],prod(dim(gcounts)[2:3]))
+    out <- tcrossprod( gcounts, bidip.to.bihap )
[…]

### 2016-02-05 notes/pde-clines.Rmd
comparison to theory; changed "F" to "FALSE"
comparison to theory progressing

@@
 
-xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=10)
+xgrid <- setup.grid.1D(x.up=-20, x.down=20, N=21)
 rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=40)
 
-# took 24min with xgrid$N=10, rgrid$N=20 and 50 time steps up to t=3, :(
 hap.soln <- forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=rgrid)

### 2016-02-05 sims/chunks_fns.R

@@ -37,7 +37,7 @@ conditional_haplotype_counts <- function (gcounts) {
 conditional_freqs <- function (gcounts) {
     # uses output of conditional_genotype_counts to get conditional allele frequencies
     # of A alleles at the FIRST locus, given the identity of the allele at the SECOND locus
-    # (so, selected allele should be second)
+    # (so, selected allele should be SECOND argument to conditional_genotype_counts)
     # a (ndemes) x 2 matrix
     hcounts <- conditional_haplotype_counts(gcounts)
     cbind( A = hcounts[,"A","A"]/(hcounts[,"A","A"]+hcounts[,"B","A"]),
@@ -94,16 +94,16 @@ true.genotypes <- list( x1=c( "BB-AB", "AA-BB", "AA-AA" ),
                          x2=c( "BA-BB", "BB-BB", "AA-AB" ) )
 true.genocounts <- do.call( table, c( list( rep(c("x1","x2"),each=3), 
                             factor(sapply(strsplit(unlist(true.genotypes),"-"),"[",1), levels=c("AA","AB","BA","BB")),
-                            factor(sapply(strsplit(unlist(true.genotypes),"-"),"[",2),levels=c("AA","AB","BA","BB")) ) ) )
+                            factor(sapply(strsplit(unlist(true.genotypes),"-"),"[",2), levels=c("AA","AB","BA","BB")) ) ) )
 stopifnot( all.equal( test.cgcounts, true.genocounts, check.attributes=FALSE ) )
 
 # test conditional_haplotype_counts
 test.chcounts <- conditional_haplotype_counts( test.cgcounts )
-true.haplotypes <- list( x1=c("BA","BB","AB","AB","AA","AA"),
-                         x2=c("BB","AB","BB","BB","AA","AB") )
+true.haplotypes <- list( x1=c("B-A","B-B","A-B","A-B","A-A","A-A"),
+                         x2=c("B-B","A-B","B-B","B-B","A-A","A-B") )
 true.haplocounts <- do.call( table, c( list( rep(c("x1","x2"),each=6), 
-                            factor(sapply(strsplit(unlist(true.haplotypes),""),"[",1), levels=c("A","B")),
-                            factor(sapply(strsplit(unlist(true.haplotypes),""),"[",2),levels=c("A","B")) ) ) )
+                            factor(sapply(strsplit(unlist(true.haplotypes),"-"),"[",1), levels=c("A","B")),
[…]

### 2016-02-05 sims/compare-to-theory.Rmd
added linked clines

@@
        width=sim.params$ndemes )
-theory.params
+unlist(theory.params)
 ```
@@
 ```
+
+# Linked clines
+
+This will just be a combination of the previous results,
+since it is local proportions of each type without conditioning on the allele at the selected locus.
+```{r linked_clines}
+for (kr in seq_along(sim.rvals)[-1]) {
+    gcounts <- genotype_counts(chunks[[kr]])
+    # sim frequencies
+    sfreqs <- sweep( tcrossprod(gcounts,dip.to.hap), 1, 2*rowSums(gcounts), "/" )
+    # theory frequencies
+    cline.soln <- cline_from_soln(linked.solns[[kr]],grid=xgrid)
+    tfreqs <- cbind( A=cline.soln[nrow(cline.soln),-1], B=1-cline.soln[nrow(cline.soln),-1] )
+    matplot(xx, sfreqs, type='l', lty=1, main=sprintf("Linked allele frequency, t=%d, r=%0.3f",sim.params$tau, sim.rvals[kr]), 
+            ylab="frequency", xlab='geographic position')
+    matlines(xgrid$x.mid, tfreqs, lty=2, lwd=2)
+    legend("topright", legend=outer(colnames(sfreqs),c("sim","theory"),paste,sep=" : "), lty=rep(1:2,each=2), col=rep(1:ncol(sfreqs),2))
+}
+```
+
+# Haplotype probabilities

### 2016-02-06 drafts/hybrid_zone.tex
added yanivs words

@@
 
+%% yaniv says:
+\paragraph{Signals of block length}
+\plr{or, perhaps for introduction?}
+In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived.
+Put differently, blocks of 'wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
+
+A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
+
+\paragraph{Solutions to PDE}
 Numerical solution for the PDE can be used on real, non-flat landscapes that change with time;

### 2016-02-06 drafts/hybrid_zone.tex
added Figure1

@@
 
+\begin{figure}
+\includegraphics{figs/BM_schematic}
+\caption{A cartoon representation of the model.
+The black line represents the path, backwards in time, of a selected site that is sampled in the present day ($\tau=0$).
+The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A is less fit at $x>0$ and ancestry B is less fit when $x<0$.
+In this figure, the sampled allele under selection is of ancestry B, and as a result migrates rapidly back to $x>0$ from a position $x<0$.
+Red dots represent recombination events along the entire haplotype, and here lineages that recombine off the same initial haplotype as the selected allele are represented by colored lines.
+Here, the initial haplotype is only broken up into 3 chunks over 100 generations.
+Recombination events can cause these chunks to change linkage to the allele at the selected locus (e.g.
+a chunk that was initially linked to a selected allele of ancestry A can become linked to a selected allele of ancestry B).
+Here, yellow lineages are linked to an allele of ancestry A at the selected locus, and blue lineages are linked to an allele of ancestry B.
+The ancestry of the blue and yellow lineages are determined by their position at $\tau=100$.
+Here the left-most yellow lineage is of ancestry A and the blue lineage is of ancestry B.
+}\label{Fig:schematic}
+\end{figure}
 %As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.

### 2016-02-07 sims/compare-to-theory.Rmd
Removed old chunk file as @asedghifar said; added frequency sanity check.

@@
 # spatial locations of the demes
-xx <- seq_len(sim.params$ndemes)-sim.params$ndemes/2
+xx <- seq_len(sim.params$ndemes)-(sim.params$ndemes+1)/2
 theory.params <- list( sigma=trueSigma(sim.params$sigma), s=sim.params$s, tau=sim.params$tau, 

### 2016-02-07 sims/frequency_sanity_check.R

@@ -1,0 +1,15 @@
+pdf('cline_centers.pdf', width=6,height=4, pointsize=10)
+
+for (x in list.files(".","simul.*_chunks.Robj")) {
+
+    load(x)
+
+    B_freq = do.call(cbind,lapply(chunks,function(C){sapply(C,function(X){length(which(X>0))})}))
+
+
+    matplot(B_freq/1000,xlab="geo.pos",ylab="p_B", type='l', lty=1, main=x, cex.main=0.5)
+    matlines(1-B_freq/1000, lty=2)
+    abline(v=25.5)
+}
+
+dev.off()

### 2016-02-07 sims/run_sims.R
minor modification of yaniv_sims.R to run_sims.R to output to a directory and save parameters

@@ -1,0 +1,141 @@
+#!/usr/bin/Rscript
+
+source("sim-fns.R")
+
+#PARAMS for simulation:
+
+params = list(
+    SIGMA = 1,
+    ndemes = 100,
+    deme_size = 10,
+    S = 0.1,
+    zone_age = 10
+)
+params$ninds = params$ndemes * params$deme_size
+
+outfile = with( list2env(params), sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_tau%s_dir/results",SIGMA,ninds,ndemes,S,zone_age) )
+dir.create(dirname(outfile))
+
+#PARAMS for parsing:
+loci = list(seq(0.5,1,0.05),seq(0.5,1,0.05))
+xx = (1:params$ndemes)-0.5-params$ndemes/2
+
+
+#DEFINE selection against genotypes:
+qtl=list(
+        chr1=data.frame(
+                traits=c("underdominant"), 
+                s=params$S,
+                pos=c(0.5)),
+        chr2=data.frame(
+                traits=c("underdominant"), 
+                s = 0,pos=c(0.5))
+        )
+
+# save out parameters
+save( params, outfile, loci, xx, qtl, file=paste0(outfile,"_params.Robj") )
+
+#SIMULATE the populations
+sims.full = lapply(params$zone_age, sim.zone, n.ind=params$ninds, n.deme=params$ndemes, sigma=params$SIGMA)
+	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
+		#where:
+			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+			#"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
+			#"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
+			#"pars" is an integer that gives the number of generatsion of the simulation
+			#"QTL" returns the QTL vector used in the simulation
+
+save(sims.full,file=paste(outfile,"_rawoutput.Robj",sep=""))
+			
+#PARSE the simulated chromosomes
+sims.sums = lapply(sims.full,spBreaks)
[…]

### 2016-02-08 drafts/hybrid_zone.tex
added ancestry frequency figure

@@
 
+\begin{figure}
+\includegraphics{figs/alleleFrequencies_sim}
+\caption{Frequency of ancestry B at a given genomic and physical position at $\tau=100$.
+Every line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
+Grey lines represent the same positions from a similar simulation with no selection.}\label{alleleFreq_tau100}
+\end{figure}
+
 \subsection*{Statistics}

### 2016-02-08 sims/compare-to-theory.Rmd
minor modification of yaniv_sims.R to run_sims.R to output to a directory and save parameters
templated comparison to theory

--- 2016-02-05 sims/compare-to-theory.Rmd
@@
 
+*Note:* This can be run as a script, by specifying `simchunks.file` first,
+and then rendering this document, e.g.:
+```
+Rscript -e "library(templater); simchunks.file='simulation_SIGMA1_Ninds5000_ndemes100_s0.1_dir/tau20_dir/results_simsums_chunks.Robj'; render_template('compare-to-theory.Rmd',output='simulation_SIGMA1_Ninds5000_ndemes100_s0.1_dir/tau20_dir/results_comparison-to-theory.html')"
+```
+
 ## Parameters from a simulation
@@
 ```{r sim_results}
+if (!exists("simchunks.file")) { 
     simchunks.file <- "simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums_chunks.Robj" 
+}
 load(simchunks.file)  # provides 'chunks'
@@
     strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
-    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*_%s([0-9.]+)_.*",x),"\\1",simchunks.file)) } )
+    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",simchunks.file)) } )
 }
@@
        width=sim.params$ndemes )
+simchunks.file
 unlist(theory.params)
@@
 
-**Question:** Is that shift due to stochasticity, or have I got the center in the wrong place on the sims?
+**Note:** Verified by looking at simulations with small $\tau$,
+the shift in cline center is due to the cline drifting.
 

### 2016-02-08 sims/compare-to-theory.Rmd
oh, wrong recomb distance

@@
 # recombination distances, hard-coded
-sim.rvals <- c(0,0.01,0.5)
+sim.rvals <- c(0,0.5-0.01,0.5)
 # spatial locations of the demes

### 2016-02-08 drafts/hybrid_zone.tex
yb began editing intro

@@
 \newcommand{\plr}[1]{{\em \color{blue} #1}}
+\newcommand{\yb}[1]{{\em \color{magenta} #1}}
 
@@
 The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.
-When formerly isolated populations come back into contact, a gradient of ancestry can form across geography, resulting a hybrid zone \citep{Barton1985}.
+When formerly isolated populations come back into contact \yb{the exchange of genes across a geographic hybrid zone can establish a gradient of ancestry across geography, \citep{Barton1985}}.
 In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD).
@@
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.
+\yb{It is clear that selection against admixed ancestry at one locus will influence patterns of ancestry at physically linked loci, however the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
 
@@
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.
-Here we investigate expected genome-wide patterns of co-ancestry across a hybrid zone in the presence of selection.
-Selection against migrant ancestries in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection.
+Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
+Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
 The effect on patterns of ancestral block lengths is, however, less well understood. 
-Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
+Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
 
@@
 These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
+\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}
 As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
 
-Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
-Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
+% Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
+%YB Genome-wide sounds wrong to me.
+is this ok?
+Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
+%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
+% YB I don't think we are actually asking about density of selected sites, are we?
+Our goal here is  to understand how strength of \yb{selection, recombination} rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
 This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.

### 2016-02-09 drafts/hybrid_zone.tex
added some text

--- 2016-02-08 drafts/hybrid_zone.tex
@@
 
+We were interested in determining whether targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
+To demonstrate this, we simulated a chromosome of length 1M with a selected site at position 0.5M.
+For a set of positions along the chromosome, we computed the length of the surrounding unbroken block of ancestry.
+There is a detectable signal of increased block length surrounding selected loci, especially when conditioning on the less-fit ancestry (FigureX) at geographic sites far from the center of the zone.
+
 \subsection*{Number of ancestors}

### 2016-02-09 sims/sim-fns.R

@@ -252,4 +252,28 @@ getChunk = function( IND_DATA, CHR, POS ) {
 }
 
 
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
+	#return the interval containing focal site for a individual	
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancA),0)
+}
+
+
+get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
+	#get distribution of chunks within a deme
+	INDS = IND_DATA[which(demeID==DEME)]
+	intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
+	return(intervals)
+}
+
+
+get.chunks.at.positions = function(VECTOR_OF_POS,CHR){
+	lapply(VECTOR_OF_POS,function(POS){lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=FALSE,POS=POS,CHR=CHR)})})
+}
+
+get.ancestry.freqs = function(LIST_OF_CHUNKS){#outputs a ndeme x length(pos) matrix
+	do.call(cbind,lapply(LIST_OF_CHUNKS,function(C){sapply(C,function(X){length(which(X>0))})}))
+}
+
 

### 2016-02-09 sims/compare-to-theory.Rmd
adjusted r vals (and saved them out) in run_sims

--- 2016-02-08 sims/compare-to-theory.Rmd
@@
 }
-load(simchunks.file)  # provides 'chunks'
+load(simchunks.file)  # provides 'chunks' and 'chunks.rvals'
 get_simparams <- function (fname) {
@@
 sim.params <- get_simparams( simchunks.file )
-# recombination distances, hard-coded
-sim.rvals <- c(0,0.01,0.5)
+# recombination distances
+# sim.rvals <- c(0,0.5-0.01,0.5)
+sim.rvals <- sapply( chunks.rvals, "[", 2 )
 # spatial locations of the demes
@@
 # Haplotype probabilities
+
+The probability of seeing a haplotype of length at least $\ell$
+covering position $r$ on the genome at location $x$ in space: for fixed $\ell$,
+against $r$ and/or $x$.
+
+We'll do this against $x$ at a few values of $r$ because that's what we have.
+```{r theory_haps, cache=TRUE}
+rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=41)
+hap.soln <- forwards_backwards_haplotypes(s=theory.params$s, times=tt, xgrid=xgrid, rgrid=rgrid,
+                           sigma=theory.params$sigma, 
+                           fwds.grid=fgrid, fwds.soln=fwds.soln )
+```

### 2016-02-10 drafts/hybrid_zone.tex
added some block length figures

@@
 
-We were interested in determining whether targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
+Of particular interest is whether or targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
 To demonstrate this, we simulated a chromosome of length 1M with a selected site at position 0.5M.

### 2016-02-10 drafts/figs/make_length_along_chr.R
added some block length figures

@@ -1,0 +1,36 @@
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_chr1_start0.48_stop0.52_by0.001_intervalSizes_allAncs.Robj")
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_chr1_start0.48_stop0.52_by0.001_intervalSizes.Robj")
+
+Ndemes=50
+deme_ID = do.call(c,lapply(1:Ndemes,function(X){rep(X,Ninds/Ndemes)}))
+
+positions=seq(0.48,0.52,0.001)
+
+pdf(height=4, width=6.25,file="blocksAlongChromNoConditioning.pdf")
+par(mar=c(3.5,3.5,0.5,0.5))
+mean_per_deme = do.call(rbind, lapply(intervalSizes_allAncs,function(P){tapply(1:25000,deme_ID,function(Z){mean(P[Z,])})}))
+plot(mean_per_deme[,1]/mean(mean_per_deme[,1],na.rm=T),x=positions,ylim=c(0.5,3),type="l",lty=1,col=c(rainbow(25),rev(rainbow(25))),main="",cex.main=0.8,xlab="",ylab="",yaxt="n",xaxt="n")
+for(i in 1:50){points(mean_per_deme[,i]/mean(mean_per_deme[,i],na.rm=T),x=positions,type="l",col=rainbow(50)[i])}
+axis(1,at=seq(0.48,0.52,0.01),labels=seq(-0.02,0.02,0.01),cex.axis=0.9)
+axis(2,at=seq(0.5,3,0.5),cex.axis=0.9,las=2)
+mtext("Distance from selected locus (M)",side=1,line=2.5)
+mtext("Normalized mean block length",side=2,line=2.5)
+legend("topright",title="distance from center:",legend = rev(c(seq(-25,-5,5),seq(5,25,5))),cex=0.7,col=rainbow(50)[seq(1,50,5)],lty=1,xjust=1)
+dev.off()
+
+pdf(height=4, width=6.25,file="blocksAlongChromAncBConditioning.pdf")
+par(mar=c(3.5,3.5,0.5,0.5))
+
+mean_per_deme_AncB = do.call(rbind, lapply(intervalSizes,function(P){tapply(1:25000,deme_ID,function(Z){all_sites = P[Z,]; mean(all_sites[which(all_sites>0)])})}))
[…]

### 2016-02-10 drafts/figs/make_length_along_chr.R
fixed block length figures

@@ -1,3 +1,41 @@
+image.scale <- function(z, zlim, col = heat.colors(12),
+breaks, horiz=TRUE, ylim=NULL, xlim=NULL, ...){
+ if(!missing(breaks)){
+  if(length(breaks) != (length(col)+1)){stop("must have one more break than colour")}
+ }
+ if(missing(breaks) & !missing(zlim)){
+  breaks <- seq(zlim[1], zlim[2], length.out=(length(col)+1)) 
+ }
+ if(missing(breaks) & missing(zlim)){
+  zlim <- range(z, na.rm=TRUE)
+  zlim[2] <- zlim[2]+c(zlim[2]-zlim[1])*(1E-3)#adds a bit to the range in both directions
+  zlim[1] <- zlim[1]-c(zlim[2]-zlim[1])*(1E-3)
+  breaks <- seq(zlim[1], zlim[2], length.out=(length(col)+1))
+ }
+ poly <- vector(mode="list", length(col))
+ for(i in seq(poly)){
+  poly[[i]] <- c(breaks[i], breaks[i+1], breaks[i+1], breaks[i])
+ }
+ xaxt <- ifelse(horiz, "s", "n")
+ yaxt <- ifelse(horiz, "n", "s")
+ if(horiz){YLIM<-c(0,1); XLIM<-range(breaks)}
+ if(!horiz){YLIM<-range(breaks); XLIM<-c(0,1)}
+ if(missing(xlim)) xlim=XLIM
+ if(missing(ylim)) ylim=YLIM
+ plot(1,1,t="n",ylim=ylim, xlim=xlim, xaxt=xaxt, yaxt=yaxt, xaxs="i", yaxs="i", ...)  
+ for(i in seq(poly)){
+  if(horiz){
+   polygon(poly[[i]], c(0,0,1,1), col=col[i], border=NA)
+  }
+  if(!horiz){
+   polygon(c(0,0,1,1), poly[[i]], col=col[i], border=NA)
+  }
+ }
+}
+
+
+
+
 load("simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_chr1_start0.48_stop0.52_by0.001_intervalSizes_allAncs.Robj")
[…]

### 2016-02-10 drafts/hybrid_zone.tex
more block length figures!

@@
 \usepackage{color}
+\usepackage{amssymb}
 
@@
 \date{} % delete this line to display the current date
+
+\usepackage{xr}
+\externaldocument{supplement}
 
@@
 In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection.
-The most straightforward measure is $l_I(b)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site.
+The most straightforward measure is $l_I(m)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site.
+Based on our expectation of the distribution of block lengths, we compute $l_B(m)$ for $m$ across a chromosome.
 
-Another statistic is the cumulative probability $C_I(b)$ of a sampled chunk encompassing position $b$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
+Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
 
-Additionally, we look at the mean value of these statistics in the two chunks, $b-$ and $b+$, that flank the one containing $b$.
+Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$, that flank the block of unbroken ancestry containing $m$.
 
@@
 To demonstrate this, we simulated a chromosome of length 1M with a selected site at position 0.5M.
-For a set of positions along the chromosome, we computed the length of the surrounding unbroken block of ancestry.
-There is a detectable signal of increased block length surrounding selected loci, especially when conditioning on the less-fit ancestry (FigureX) at geographic sites far from the center of the zone.
+For a set of positions $\mathbf{m}$ along the chromosome, we computed the length $l_B(m_i)$
+of the surrounding unbroken block of ancestry.
+There is a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
+A signal is also present, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
+
+\plr{or, perhaps for introduction?}
+In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived.
+Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
+A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
+This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
+
+\begin{figure}
+\includegraphics{figs/blocksAlongChromAncBConditioning}
+\caption{Mean block length surrounding positions along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
+Here each line represents mean block lengths in a given deme, and is normalized by mean block length along the chromosome in the deme.}\label{Fig:blockLengths}
+\end{figure}
 
@@
 \paragraph{Signals of block length}
-\plr{or, perhaps for introduction?}
-In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
-We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived.
-Put differently, blocks of 'wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
-
-A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
-Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
 

### 2016-02-10 drafts/figs/make_length_along_chr.R

@@ -1,41 +1,3 @@
-image.scale <- function(z, zlim, col = heat.colors(12),
-breaks, horiz=TRUE, ylim=NULL, xlim=NULL, ...){
- if(!missing(breaks)){
-  if(length(breaks) != (length(col)+1)){stop("must have one more break than colour")}
- }
- if(missing(breaks) & !missing(zlim)){
-  breaks <- seq(zlim[1], zlim[2], length.out=(length(col)+1)) 
- }
- if(missing(breaks) & missing(zlim)){
-  zlim <- range(z, na.rm=TRUE)
-  zlim[2] <- zlim[2]+c(zlim[2]-zlim[1])*(1E-3)#adds a bit to the range in both directions
-  zlim[1] <- zlim[1]-c(zlim[2]-zlim[1])*(1E-3)
-  breaks <- seq(zlim[1], zlim[2], length.out=(length(col)+1))
- }
- poly <- vector(mode="list", length(col))
- for(i in seq(poly)){
-  poly[[i]] <- c(breaks[i], breaks[i+1], breaks[i+1], breaks[i])
- }
- xaxt <- ifelse(horiz, "s", "n")
- yaxt <- ifelse(horiz, "n", "s")
- if(horiz){YLIM<-c(0,1); XLIM<-range(breaks)}
- if(!horiz){YLIM<-range(breaks); XLIM<-c(0,1)}
- if(missing(xlim)) xlim=XLIM
- if(missing(ylim)) ylim=YLIM
- plot(1,1,t="n",ylim=ylim, xlim=xlim, xaxt=xaxt, yaxt=yaxt, xaxs="i", yaxs="i", ...)  
- for(i in seq(poly)){
-  if(horiz){
-   polygon(poly[[i]], c(0,0,1,1), col=col[i], border=NA)
-  }
-  if(!horiz){
-   polygon(c(0,0,1,1), poly[[i]], col=col[i], border=NA)
-  }
- }
-}
-
-
-
-
 load("simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_chr1_start0.48_stop0.52_by0.001_intervalSizes_allAncs.Robj")
 load("simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_chr1_start0.48_stop0.52_by0.001_intervalSizes.Robj")
 
[…]

### 2016-02-10 drafts/hybrid_zone.tex
more text to draft

@@
 There is a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
-A signal is also present, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
+A signal is also present for, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
 
@@
 Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
-This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
+This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
 
@@
 \includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{Mean block length surrounding positions along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
-Here each line represents mean block lengths in a given deme, and is normalized by mean block length along the chromosome in the deme.}\label{Fig:blockLengths}
+\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
+Here each line represents  $l_B(\mathbf{m})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
 \end{figure}

### 2016-02-10 drafts/supplement.tex
added a supplement

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
%\usepackage{figcaps}
\usepackage[nomarkers,notablist,nofiglist,tablesfirst]{endfloat}
\usepackage{amsfonts}
\usepackage{sectsty}
%\usepackage{subfigure}
%\usepackage{subcaption}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%\allsectionsfont{\sffamily}
%% THE MANUSCRIPT TITLE

\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}

\usepackage{bibentry}

\usepackage{xr}
\externaldocument{hybrid_zone}

\newcommand{\alisa}[1]{{\em \color{red} #1}}
\newcommand{\plr}[1]{{\em \color{blue} #1}}
\newcommand{\yb}[1]{{\em \color{magenta} #1}}

\newcommand{\given}{\,\vert\,}
\newcommand{\st}{\,\colon\,}
\renewcommand{\and}{\,\&\,}

\begin{document}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}
\includegraphics{figs/blocksAlongChromHeatmapAncBConditioning}
\caption{Heatmap of mean block length $l_B(m_i)$ along a simulated chromosome. }\label{Supp:blockLengthHeatmap}
\end{figure}

\begin{figure}
\includegraphics{figs/blocksAlongChromNoConditioning}
\caption{Heatmap of mean block length $l_\bullet(m_i)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, \tau=1000$). }\label{Supp:blockLengthNoAnc}
\end{figure}

\begin{figure}
\includegraphics{figs/blocksAlongChromHeatmap}
\caption{Heatmap of mean block length $l_\bullet(m_i)$ along a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
\end{figure}

\begin{figure}
\includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning}
\caption{Ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and adjacent block lengths across a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$).  Each line represents a deme and is normalized by mean block length across the chromosome in the deme.}\label{Supp:ratioBlockAdjacent}
\end{figure}

\begin{figure}
\includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning}
\caption{Heatmap of $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$). }\label{Supp:ratioBlockAdjacentHeatmap}
\end{figure}

\end{document}

### 2016-02-10 sims/compare-to-theory.Rmd
updated htmls

--- 2016-02-08 sims/compare-to-theory.Rmd
@@
 sim.params <- get_simparams( simchunks.file )
-# recombination distances, hard-coded
-sim.rvals <- c(0,0.01,0.5)
+# recombination distances
+# sim.rvals <- c(0,0.5-0.01,0.5)
+sim.rvals <- sapply( chunks.rvals, "[", 2 ) - 0.5
 # spatial locations of the demes
@@
 # Haplotype probabilities
+
+The probability of seeing a haplotype of length at least $\ell$
+covering position $r$ on the genome at location $x$ in space: for fixed $\ell$,
+against $r$ and/or $x$.
+
+We'll do this against $x$ at a few values of $r$ because that's what we have.
+```{r theory_haps, cache=TRUE}
+rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=41)
+# hap.soln <- forwards_backwards_haplotypes(s=theory.params$s, times=tt, xgrid=xgrid, rgrid=rgrid,
+#                            sigma=theory.params$sigma, 
+#                            fwds.grid=fgrid, fwds.soln=fwds.soln )
+```

### 2016-02-10 sims/compare-to-theory.Rmd
fixed htmls again

@@
 # sim.rvals <- c(0,0.5-0.01,0.5)
-sim.rvals <- sapply( chunks.rvals, "[", 2 ) - 0.5
+sim.chroms <- sapply( chunks.rvals, "[", 1 )
+sim.rvals <- ifelse( sim.chroms==1, sapply( chunks.rvals, "[", 2 ) - 0.5, 0.5 )
 # spatial locations of the demes

### 2016-02-12 drafts/hybrid_zone.tex
made bib nicer (missing a bunch of citations)
added figures for family size

@@
 \usepackage{amsmath,amssymb}
-\usepackage{natbib}
+\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 \usepackage{color}
@@
 due to the decreased fitness of heterozygotes \citep{bartonbarrier}
-and the relative slowness of diffusive movement \citep{sedghifar}.
+and the relative slowness of diffusive movement \citep{Sedghifar2015}.
 Suppose we sample across at cline $\tau$ units of time after secondary contact.
@@
 of a single ancestry at a given location and time.
-As in \citet{sedghifar},
+As in \citet{Sedghifar2015},
 to do this we need only keep track of both lineages when a recombination in the middle of a block occurs,
@@
 \subsection*{Number of ancestors}
-Not sure what the result is here yet -- unexpected
+Although the theoretical model presented here does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
+In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
+This is a reflection of the average size of a haplotype family.
+We find that, as distance from the zone center increases, the average family size of an ancestor from time $\tau$ decreases (Figure~\ref{Fig:family_size}).
+One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.
+
+\begin{figure}
+\includegraphics{figs/number_of_ancestors_tau1000}
+\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space. Each line }\label{Fig:family_size}
+\end{figure}
 
@@
 \section{Discussion}
+Here we present an expectation of ancestry block lengths in hybrid zones.
+We find that block lengths surrounding selected loci are, on average, longer than those surrounding neutral loci .
+This simply reflects the intuition that regions surrounding selected loci are going to have a harder time introgressing, and that further away from the center of the hybrid zone a haplotype containing a disadvantageous allele is likely to have been a relatively recent migrant and therefore will have experienced fewer recombination events.
+
 Topics include\\

### 2016-02-12 drafts/figs/make_length_along_chr.R

@@ -192,3 +192,10 @@ par(mar=c(0.5,3.5,2,0.5))
 	mtext("Distance from HZ center",side=2,line=2.5)
 dev.off()
 
+
+
+#########################
+##########NOW DO NEUTRAL:
+#########################
+
+

### 2016-02-12 drafts/figs/number_of_inds.R

@@ -1,0 +1,76 @@
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_run2016-01-27_rawoutput.Robj")	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.ids","deme","pars","QTL")]]
+		#where:
+			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+			#"sp.ids" is a character vector of length ninds, giving the ancestry given an ancestor ID
+			#"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
+			#"pars" is an integer that gives the number of generations of the simulation
+			#"QTL" returns the QTL vector used in the simulation
+
+
+#Let's get the number of ancestors represented in chunks in each deme:
+
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
+	#return the interval containing focal site for a individual	
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancA),0)
+}
+
+
+get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
+	#get distribution of chunks within a deme
+	INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
+	intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
+	return(intervals)
+}
+
+
+
+
+#returns table (ancestor at locus, and whether it is AncB (T/F))
+ancestry.at.position = function(IND=1,DATA=sims.s0.1,POS = 0.5,CHROM="chr1",SP="B"){
+	distance_from_focal_pos = lapply(DATA[[1]][["inds"]][[IND]][[CHROM]],function(X){d = POS-X$pos;return(X[max(which(d>0)),])})
[…]

### 2016-02-12 sims/potential_statistics.R

@@ -1,11 +1,10 @@
-load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums.Robj")
-
-chromosome = "chr2"
-plot_start = 0.45
-plot_stop=0.55
-increments = 0.005
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_run2016-01-27_simsums.Robj")
+chromosome = "chr1"
+plot_start = 0.48
+plot_stop=0.52
+increments = 0.001
 Ngen = 1000
-sel=0.1
+sel=0
 SIGMA=1
 Ninds=25000
 Ndemes = 50
@@ -26,7 +25,7 @@ deme_ID = do.call(c,lapply(1:Ndemes,function(X){rep(X,Ninds/Ndemes)}))
 get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=chromosome,POS=0.5,ancB=TRUE){
 	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})
 	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
-	replace(chunk,which(identity==ancA),0)
+	replace(chunk,which(identity==ancB),0)
 	
 }
 
@@ -177,14 +176,14 @@ for(i in 1:50){points(flanking.blocks.deme.matrix.AncB[i,]/mean(flanking.blocks.
 
 matplot((mean_per_deme)/t(flanking.blocks.deme.matrix),type="l",col=rainbow(50),lty=1,main="mean_focal/mean_flanking length ALL" ,cex.main=0.7)
 
-matplot((mean_per_deme_AncB)/t(flanking.blocks.deme.matrix.AncB),type="l",col=rainbow(50),lty=1,main="mean_focal/mean_flanking length AncB" ,cex.main=0.7)
+matplot((mean_per_deme_AncB)/t(flanking.blocks.deme.matrix.AncB[50:1,]),type="l",col=rainbow(50),lty=1,main="mean_focal/mean_flanking length AncB" ,cex.main=0.7)
 
 image(t(flanking.blocks.deme.matrix),x=positions,y=-25:25,ylab="deme",xlab="phys.pos",main="heatmap ALL (log)")
[…]

### 2016-02-12 drafts/hybrid_zone.tex
made some edits to the intro

@@
 Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-
-Some words about how new genomic data will allow us to focus on tracts and not just individual loci.
+\alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
+}
 
@@
 \yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}
-As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
+%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
 
@@
 is this ok?
-Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
+Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci.
 %Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
 % YB I don't think we are actually asking about density of selected sites, are we?
-Our goal here is  to understand how strength of \yb{selection, recombination} rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
-This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.
+Our goal here is  to understand how strength of \yb{selection, recombination} rate, and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
+This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.
 In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one.
 If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact.
-In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
+In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation.
+%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
 
-Something about how we just study the selection against hets model,
-but we expect general conclusions to apply to other situations.
+Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
+This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
+Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
 
-The tracking lineages given selection trick is an example of the more general idea of duality for particle systems.
-It also assumes a large pop density limit that excludes recombination and makes lineages Markov;
-if things are very patchy this might not be a good one.
-Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
-
-Expectations:
-Conditional  on having identity B on the wrong side, far away from the center of the zone, the length of unbroken ancestry surrounding the selected locus should be relatively long.
-This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
+\alisa{Moved some points here to the discussion.}
 
@@
 In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
-We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
 Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
@@
 \includegraphics{figs/number_of_ancestors_tau1000}
-\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space. Each line }\label{Fig:family_size}
+\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space.
+Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$).
+Yellow lines are corresponding positions on an unlinked chromosome with no selected loci, and grey lines are corresponding positions in a simulation with no selection.
+Bold lines depict the target of selection when present, and corresponding position on chromosomes not harboring any selected sites.}\label{Fig:family_size}
 \end{figure}
@@
 This simply reflects the intuition that regions surrounding selected loci are going to have a harder time introgressing, and that further away from the center of the hybrid zone a haplotype containing a disadvantageous allele is likely to have been a relatively recent migrant and therefore will have experienced fewer recombination events.
+
+Something about how we just study the selection against hets model,
+but we expect general conclusions to apply to other situations.
+
+The tracking lineages given selection trick is an example of the more general idea of duality for particle systems.
+It also assumes a large pop density limit that excludes recombination and makes lineages Markov;
+if things are very patchy this might not be a good one.
+Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
 

### 2016-02-13 drafts/hybrid_zone.tex
got to the end of 'Establishment of selected cline'

@@
 \usepackage{amssymb}
+
+\usepackage[normalem]{ulem}
 
@@
 \section{Methods}
-\subsection*{Model}
+\subsection*{The Model}
 We consider a model in which two isolated subspecies, species $A$ and species $B$ come into contact at time $\tau$ generations in the past.
 Individuals move forward in time by sampling a displacement from a Gaussian distribution.
-In the simplest case of under-dominance, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
+As in \cite{Barton??}, we consider a simple case of under-dominance.
+Specifically, here the fitness of an individual is  determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
 
-To fix notation, we will say that species $A$ was on the ``left'' of the zone of contact,
+To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact,
 which corresponds to spatial positions $x<0$.
-
-We consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past.
-The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that the two species are fixed for alternate alleles at this locus).
+Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
+The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
 For every other locus, we keep track of $a$, the identity at $L$, across time and space.
 If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise.
-Recombination is modeled as a poisson process with rate $r$.
+Recombination between two loci is modeled as a poisson process with rate $r$.
 The identity $a$ at locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
@@
 \includegraphics{figs/BM_schematic}
-\caption{A cartoon representation of the model.
-The black line represents the path, backwards in time, of a selected site that is sampled in the present day ($\tau=0$).
-The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A is less fit at $x>0$ and ancestry B is less fit when $x<0$.
-In this figure, the sampled allele under selection is of ancestry B, and as a result migrates rapidly back to $x>0$ from a position $x<0$.
-Red dots represent recombination events along the entire haplotype, and here lineages that recombine off the same initial haplotype as the selected allele are represented by colored lines.
-Here, the initial haplotype is only broken up into 3 chunks over 100 generations.
-Recombination events can cause these chunks to change linkage to the allele at the selected locus (e.g.
-a chunk that was initially linked to a selected allele of ancestry A can become linked to a selected allele of ancestry B).
-Here, yellow lineages are linked to an allele of ancestry A at the selected locus, and blue lineages are linked to an allele of ancestry B.
-The ancestry of the blue and yellow lineages are determined by their position at $\tau=100$.
-Here the left-most yellow lineage is of ancestry A and the blue lineage is of ancestry B.
+\caption{A cartoon representation of a haplotype under the model.
+The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day ($\tau=0$).
+The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$.
+Here, because of negative selection, $L_B$ migrates rapidly back to $x>0$ from a position $x<0$.
+Red dots represent recombination events along the entire haplotype, and lineages that recombine off the same haplotype as $L_B$ are represented by colored lines.
+(here, the initial haplotype is broken up into 3 chunks over 100 generations).
+Subsequent recombination events can cause these chunks to change which allele they are linked to at the selected site, and here yellow lineages are linked to an allele of ancestry $A$ at the selected locus, and blue lineages are linked to a selected allele of ancestry $B$.
+From right to left, the chunks at $\tau=100$ are of ancestry $(A, B, B)$.
 }\label{Fig:schematic}
@@
 and after time $\tau$ will have clines out to genetic distance of order $1/\tau$,
-maybe more.
+maybe more. \alisa{has this not already been stated by Nick somewhere?}
 
-\paragraph{Establishment of the selected cline}
+\paragraph{Establishment of  \sout{the selected cline} \alisa{a cline at the selected locus}}
 Consider a single locus, with alternate alleles fixed in the two parental populations,
@@
 Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
-Suppose that secondary contact occurred at time $t=0$, 
+Suppose that secondary contact occurred at time $t=0$, \alisa{(Will switching between backwards and forwards in time not get confusing?)}
 and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
@@
 and decays exponentially.
-Furthermore, the cline is established through diffusion.
-This is slowed down somewhat by selection against heterozygotes,
+Furthermore, the cline is established through diffusion \alisa{of individuals?}, which is slowed down somewhat by selection against heterozygotes,
 but is still diffusive, and so it will take time of order $1/s$
-to reach its stable width.
-(To confirm this, 
-rescale space by $\sigma/s$ and time by $1/s$ in equation \eqref{eqn:cline_pde},
-and see that the result doesn't depend on $\sigma$ or on $s$.)
+to reach its stable width \alisa{(is this something that has been shown before?)} This can be shown by rescaling space by $\sigma/s$ and time by $1/s$ in equation \eqref{eqn:cline_pde}, and seeing that the result \alisa{(what result?)} doesn't depend on $\sigma$ or on $s$.
 
@@
 
+Rapid drift (as in BM-drift) back to the `correct' side, conditional on being the wrong ancestry.
+
 Something about how we just study the selection against hets model,

### 2016-02-13 drafts/hybrid_zone.tex
yb working on page 1

@@
 The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
-When formerly isolated populations come back into contact \yb{the exchange of genes across a geographic hybrid zone can establish a gradient of ancestry across geography, \citep{Barton1985}}.
-In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD).
-If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
+\yb{Thus, ancestry originating from the two diverged populations spreads from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
+%In addition to steep clines in ancestry,
+\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}
+\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both of the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.}
+However, subsets of the genome that are under selection can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
-\yb{It is clear that selection against admixed ancestry at one locus will influence patterns of ancestry at physically linked loci, however the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
+\yb{However the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation.
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher; Chapman \& Thomson; Barid et al)}.   
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
-Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
+Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
 The effect on patterns of ancestral block lengths is, however, less well understood. 
@@
 \alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
-}
+} %GREAT SENTENCE!!
 
-While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
-These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
+The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
+These models predict cline width to be narrowest \yb{(i.e.
+the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider.
+%and more heavily influenced by demography.} %hybrid zone age and rates of gene flow.
 \yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}

### 2016-02-13 drafts/hybrid_zone.tex
yb working on page 2

@@
 
-% Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.
-%YB Genome-wide sounds wrong to me.
-is this ok?
 Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci.
@@
 % YB I don't think we are actually asking about density of selected sites, are we?
-Our goal here is  to understand how strength of \yb{selection, recombination} rate, and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
-This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions.
+Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}.
+%in present day populations.
+This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
 In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact.
-In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation.
+If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \yb{(CITE: hapmix, globetrotter)}.
+\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
+\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci (CITE: clinefit, gbc), but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
+%In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
 %In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
@@
 Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
+\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
 

### 2016-02-13 drafts/hybrid_zone.tex
yb working on model

@@
 \subsection*{The Model}
-We consider a model in which two isolated subspecies, species $A$ and species $B$ come into contact at time $\tau$ generations in the past.
+We consider two isolated populations --- labeled \emph{Species $A$} and \emph{Species $B$} ---  come into contact $\tau$ generations in the past. 
 Individuals move forward in time by sampling a displacement from a Gaussian distribution. 
-As in \cite{Barton??}, we consider a simple case of under-dominance.
+We consider a simple case of under-dominance. 
+\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection (CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993) which also generate clines like those degenerated under complex epistatic systems}.
 Specifically, here the fitness of an individual is  determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
+To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.
 
-To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact,
-which corresponds to spatial positions $x<0$.
 Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}).

### 2016-02-14 drafts/hybrid_zone.tex
got to the end of methods

--- 2016-02-13 drafts/hybrid_zone.tex
@@
 The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.
-When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
-\yb{Thus, ancestry originating from the two diverged populations spreads from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
-%In addition to steep clines in ancestry,
-\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}
-\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both of the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.}
-However, subsets of the genome that are under selection can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+When formerly isolated populations come back into contact \yb{the exchange of genes across a geographic hybrid zone can establish a gradient of ancestry across geography, \citep{Barton1985}}.
+In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD).
+If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.
-\yb{However the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
+\yb{It is clear that selection against admixed ancestry at one locus will influence patterns of ancestry at physically linked loci, however the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher; Chapman \& Thomson; Barid et al)}.   
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation.
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
-Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
+Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
 The effect on patterns of ancestral block lengths is, however, less well understood. 
@@
 \alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
-} %GREAT SENTENCE!!
+}
 
-The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
-These models predict cline width to be narrowest \yb{(i.e.
-the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider.
-%and more heavily influenced by demography.} %hybrid zone age and rates of gene flow.
+While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
+These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow.
 \yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}
@@
 that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$).
+with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
 Since $p=1$ far from the zone to the left, 
@@
 and that $Z$ jumps between $A$ and $B$ at rate either $r p$ or $r(1-p)$,
-depending on the identity of $Z$.
+depending on the identity of $Z$ \alisa{is $t$ the time of recombination?}.
 In symbols,
@@
 
-\textbf{figure of a lineage moving back and forth?}
+\textbf{figure of a lineage moving back and forth?} \alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}
 
@@
 substantially larger than $1/\tau$.
-However, this distinction appears hard to observe for realistic parameter values.
+However, this distinction appears hard to observe for realistic parameter values. \alisa{perhaps move some of this to discussion?}
 
@@
 \end{align}
-The description above implies that $q_r$ solves the equation
+The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation
 \begin{align}
@@
 $\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$.
+\alisa{A few words about what $G$ is?}
 
@@
 producing, for instance, a heatmap of expected local frequencies of $A$ ancestry
-across space and time, as in Figure \ref{fig:pde_clines}.
-Shown is the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
-which we denote $p(x,t,r)$ and can compute as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$.
+across space and time, as in Figure \ref{fig:pde_clines} \alisa{which shows}
+ the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
+\alisa{denoted} $p(x,t,r)$, and can be computed as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. \alisa{Some details on numerical solution (e.g. R package?)}
 
@@
 As in \citet{Sedghifar2015},
-to do this we need only keep track of both lineages when a recombination in the middle of a block occurs,
+to do this we need only keep track of both lineages when a recombination in the middle of a block occurs (Figure~\ref{Fig:schematic}),
 so that lineages will act like a particular class of labeled, branching diffusions,
@@
 a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
-each of which procedes as before, independently.
+each of which proceeds as before, independently.
 
 \paragraph{Spatial haplotype patterns}
-Furthermore, we expect that stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
+We expect that\alisa{, on average,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
 because they have had more opportunities to recombine with $B$ haplotypes.
@@
 The boundary conditions are $g_A(x,0;a,b)=1$ and $g_B(x,0;a,b)=0$.
+\alisa{Is the selected site at $\theta=0$?}
 
@@
 We implemented forwards-time simulations of diploid individuals on a one-dimensional row of demes
-with nonoverlapping generations and fixed population sizes (a Wright-Fisher model),
-and a single locus with selection against heterozygotes in the center of one of two chromosomes, each of length 1 Morgan.
-Each deme has exactly $n_d$ diploid individuals at the start of each generation.
-Each generation,
-first, each individual migrated to a new deme offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
+with non-overlapping generations and fixed population sizes (a Wright-Fisher model).
+Individuals are diploid, with haploid number $n=2$ chromosomes, each of length 1 Morgan.
+One chromosome pair harbors in its center (at position 0.5M) a single locus that reduces fitness by $s$ in heterozygotes, while the other contains no sites under direct selection.
+
+Each deme has exactly $N_d$ diploid individuals at the start of each generation.
+At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
 where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
 (The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
-Migrants past either end of the range remained at the ends.
+Migrants past either end of the range remained at the terminal demes.
 Then, 
 fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
-and in each deme $n_d$ pairs of parents are chosen, with replacement,
-with probability proportional to their fitness.
-Selfing is allowed.
+and in each deme $N_d$ pairs of parents are chosen, with replacement,
+with probability proportional to their fitness and selfing allowed.
 Since migration is not conservative, demes may have no available parents;
@@
 The next generation is formed by carrying out meiosis in each parent
-and combining the resulting gametes.
+and combining the resulting gametes such that each pair of parents leaves one descendant.
 Meiosis results in alternating blocks of the gamete's chromsome
@@
 
-Hybrid zones of ages 10, 100 and 1000 generations were simulated under conditions of $\sigma=1$ and $s=0.1$.
-From this we measure unbroken tract lengths to the right of the selected locus.
+Hybrid zones of ages 10, 100 and 1000 generations \alisa{these numbers may need to be revised?} were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$.
+From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$).
 These were compared to the output of neutral simulations with $s=0$

### 2016-02-14 drafts/hybrid_zone.tex
edited the results

@@
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher; Chapman \& Thomson; Barid et al)}.   
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher1954, Chapman \& Thomson; Barid et al)}.   
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
@@
 In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \yb{(CITE: hapmix, globetrotter)}.
+If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \yb{(CITE: hapmix, Hellenthal2014)}.
 \yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
@@
 \section{Results}
-\subsection*{The distribution of ancestry tract lengths}
-We are interested specifically in the distribution of the length of continuous tracts of ancestry surrounding the selected locus.
+\subsection*{The distribution of ancestry tract lengths across space}
+We are interested specifically in the distribution of continuous tracts of ancestry surrounding the selected locus.
+As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).
+From previous results we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
+
+When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
+This pattern provides a potential avenue for identifying targets of selection by looking for a signal of longer-than-average tracts surrounding a position, compared to a genome-wide background.
+(Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
+A signal is also present for, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
+
+\subsection*{Lengths of adjacent blocks}
+We find that the length of blocks adjacent to blocks of `wrong' ancestry surrounding a selected site are slightly shorter than average (Figure~\ref{Supp:adjacentBlocks}).
+This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
+
+\alisa{Moved Yaniv's explanation to the discussion.}
+%This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
+
+%We are interested specifically in the distribution of the length of continuous tracts of ancestry surrounding the selected locus.
 Far away from the zone center variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, however, when present an unfit migrant haplotype is expected to be longer than a neutral migrant haplotype.
@@
 
-Of particular interest is whether or targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
+%Of particular interest is whether or targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
 To demonstrate this, we simulated a chromosome of length 1M with a selected site at position 0.5M.
-For a set of positions $\mathbf{m}$ along the chromosome, we computed the length $l_B(m_i)$
-of the surrounding unbroken block of ancestry.
+For a set of positions $\mathbf{m}$ along the chromosome, we computed the length $l_B(m_i)$ of the surrounding unbroken block of ancestry.
 There is a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
@@
 
-\plr{or, perhaps for introduction?}
-In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
-We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
-Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
-A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
-Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
-This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
+\begin{figure}
+\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000}
+\caption{Randomly sampled chromosomes  across a hybrid zone of age $\tau=1000$.
+Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
+Red blocks along the chromosome denote ancestry $B$, and orange blocks are ancestry $A$.
+Regions surrounding the selected site are failing to introgress.
+ }\label{Fig:resistanceToIntrogression1000g}
+\end{figure}
 
@@
 
-\subsection*{Number of ancestors}
+\subsection*{The number of `unfit' ancestors}
 Although the theoretical model presented here does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
@@
 \end{figure}
-
-\subsection*{Identifying sites under selection}
-
-Plots with x-axis the genome and having peaks at the selected site.
-
-Question: how much resolution do we have?
 
@@
 
+\plr{or, perhaps for introduction?}
+In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
+Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
+A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
+
 \paragraph{Solutions to PDE}

### 2016-02-14 drafts/supplement.tex
made some edits to the intro
edited the results

@@
 \begin{figure}
+\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau100}
+\caption{Randomly sampled chromosomes  across a hybrid zone of age $\tau=100$.
+Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
+Red blocks along the chromosome denote ancestry $B$, and orange blocks are ancestry $A$.
+Regions surrounding the selected site are failing to introgress.
+ }\label{Fig:resistanceToIntrogression100g}
+\end{figure}
+
+\begin{figure}
 \includegraphics{figs/blocksAlongChromHeatmapAncBConditioning}
@@
 \begin{figure}
+\includegraphics{figs/adjacentBlocksAlongChromAncBConditioning}
+\caption{Mean block length of $l_A(m_i\cdot)$ across chromosome with single under dominant site ($s=0.01, \tau=1000$)}\label{Supp:adjacentBlocks}
+\end{figure}
+
+\begin{figure}
 \includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning}

### 2016-02-14 sims/plot_chromosome_schematic.R

@@ -1,114 +1,0 @@
-#Want to produce a plot of the haplotypes
-setwd("~/Documents/Hybrid_Zones/clinal-lineages/")
-
-plot.chroms = function(DATA=sims.sums,INDS=random_inds,CHR="chr1"){
-	
-		plot(0:1,0:1,col="white",bty="n",xaxt="n",yaxt="n",xlab="",ylab="")
-
-	for(i in 1:length(random_inds)){
-		data=DATA[[1]][["ind.ancest"]][[INDS[i]]][[CHR]]$X1
-		rect(data$starts,(i-1)/50+0.001,data$stops,i/50-0.001,col=ifelse(data$sp2,"orange","red"),border=FALSE)
-		#segments(0,i/50,1,i/50,col="orange",lwd=5,lend="butt")
-		#if(length(which(data$sp2))>0){
-		#segments(data$starts[data$sp2],i/50,data$stops[data$sp2],i/50,col="red",lwd=5,lend="butt")	}
-	} 	
-}
-
-ndemes=50
-random_inds = sapply(1:ndemes,function(X){ind_indices=1:500+(X-1)*500;return(sample(ind_indices,size=1))})
-
-
-#TAU=100
-
-chroms_sel= get(load("sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"))
-chroms_neut = get(load("sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"))
-	#output is:  a list: [[zone_age]][[c("ind.ancest","ancest.prop")]]
-		#"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
-		#"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.
-
-pdf(file=c("sims/plot_chromosomes_tau100.pdf"))
-par(mfrow=c(1,2),mar=c(0,0,0,0),oma=c(4,4,2,1))
-
[…]

### 2016-02-15 drafts/hybrid_zone.tex
yb goes at discusion

@@
 \section{Discussion}
-Here we present an expectation of ancestry block lengths in hybrid zones.
-We find that block lengths surrounding selected loci are, on average, longer than those surrounding neutral loci .
-This simply reflects the intuition that regions surrounding selected loci are going to have a harder time introgressing, and that further away from the center of the hybrid zone a haplotype containing a disadvantageous allele is likely to have been a relatively recent migrant and therefore will have experienced fewer recombination events.
 
-Rapid drift (as in BM-drift) back to the `correct' side, conditional on being the wrong ancestry.
+Most attempts to identify loci under selection in hybrid zones do so by identifying exception allele frequencies as a function of geographic position or genome-wide admixture proportion \yb{(CITE THINGS)}.
+Current genomic approaches to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX, LAMP and fineSTRUCTURE \yb{(CITE)} provides researchers with a measure of the length of ancestry tracts across the genome.
+The information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture.   
+However, how selection influences the length of tracts of ancestry around selected sites in hybrid zones has received much less attention. 
+We we present an expectation of ancestry block lengths in hybrid zone. 
+We find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
 
-Something about how we just study the selection against hets model,
-but we expect general conclusions to apply to other situations.
+Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to have a harder time introgressing. 
+Thus we can interpret this pattern as Brownian motion with drift in which ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the "correct" side of the hybrid zone.
+Selected alleles that have, by chance, encroached deep into the other side of the hybrid have likely to have done so quite recently, as they cannot survive long in such hostile territory.   
+The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
+\yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
+This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.}
+
+\subsection{Additional signatures around selected loci}
+
+We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a large number of independent ancestors   (Figure~\ref{Fig:family_size}).
+% is this right
+This result also reflects the recency of ancestry on the "wrong" side of the hybrid zone --- any one lineage will rapidly retreat back in time to the correct side of the zone.   
+Thus we predict
+
+In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
+Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
+A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
+
+These two novel finding may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown. 
+However, we as sequencing prices continue to drop and long phased reads become more common \yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
+ressarchers will soon have access to information pertaining to both of these findings and thus these results  could further aid in identifying selected sites in hybrid zones.
+
+\subsection{Inference}
+Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
+Specifically, a novel statistic, $l_I(m)$ -- the genetic length of ancestry surrounding a site of a given lineage -- could prove a simple and informative summary. 
+This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to stand genomes from hybrid zones to identify putatively selected loci.
+
+\subsection{Assumptions}
+We made numerous simplifying assumptions to derive our analytical and simulation results. 
+Thus before our results can be profitably used for inference, it is worthwhile to  reflect on our assumptions and their consequences.
+
+One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
+Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
+We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
+An additional assumption of our analytical model is that we ignore coalescence.  
+The broad agreement between our analytical and theoretical results suggests that this assumption will also have negligible effect on our results.
+
+However, other assumptions of our model may be of greater consequence. 
+Importantly, our work considered only a single selected site.   
+While a small number of selected sites will likely have little influence on our results, a higher density of selected sites could. 
+Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \yb{(CITE: (BARTON 1983; Kruuk 1989)}.
+Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting this great $s$ and the map distance between these linked incompatibilities.
+
+Another important assumption is the smooth and one dimensional spatial structure of the geographic landscape.  
+Patchily connected populations, or other esoteric forms of population structure may predominate on the edges of species ranges' where hybrid zones may occur.  
+While, it is still likely that the genomic signature of longer chunks of ancestry around selected sites will hold in these situations, the strength of this expectation is unclear.
 
@@
 
+\subsection{Islands of divergence}  
+The past years have witnessed evocative metaphors f seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
+The metaphors of "islands", "continents" and "archipelagoes" of divergence have inspired much research (and controversy) speciation genomics, but have been largely descriptive.  
+We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
+% in terms of the processes of migration, and selection influencing patterns of ancestry across the genome of hybridizing populations.
+Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.   
+Overall focusing on lengths of ancestry blocks across the genome brings focus onto the processes of migration and selection rather than high-level summaries far severely abstracted from the evolutionary process.
+
+\subsection{Future directions}
+
 Some words about how confidently calling tracts is the way of the future.\\
@@
 \plr{or, perhaps for introduction?}
-In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
-We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
-Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
-A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
-Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
 

### 2016-02-15 drafts/hybrid_zone.tex
some edits to discussion

@@
 When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
-\yb{Thus, ancestry originating from the two diverged populations spreads from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
+\yb{Thus, \alisa{ancestries} originating from the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
 %In addition to steep clines in ancestry,
 \yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}
-\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both of the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.}
-However, subsets of the genome that are under selection can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.}
+Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
-\yb{However the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}
+\yb{The expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
 
@@
 %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow.
-\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}
+\yb{The focus on cline width provides an expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim is to extend this to obtain an expectation for how selection shapes covariance in ancestry over \emph{genomic space}}
+
+%provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
 %As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
@@
 In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \yb{(CITE: hapmix, Hellenthal2014)}.
+If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
 \yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
-\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci (CITE: clinefit, gbc), but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
+\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
 %In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
@@
 \subsection*{The Model}
-We consider two isolated populations --- labeled \emph{Species $A$} and \emph{Species $B$} ---  come into contact $\tau$ generations in the past. 
+We consider two isolated populations --- labeled \emph{Species $A$} and \emph{Species $B$} ---  that come into contact $\tau$ generations in the past. 
 Individuals move forward in time by sampling a displacement from a Gaussian distribution. 
 We consider a simple case of under-dominance. 
-\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection (CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993) which also generate clines like those degenerated under complex epistatic systems}.
+\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, BARTON and GALE 1993} which also generate clines like those degenerated under complex epistatic systems}.
 Specifically, here the fitness of an individual is  determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
@@
 
-Most attempts to identify loci under selection in hybrid zones do so by identifying exception allele frequencies as a function of geographic position or genome-wide admixture proportion \yb{(CITE THINGS)}.
-Current genomic approaches to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX, LAMP and fineSTRUCTURE \yb{(CITE)} provides researchers with a measure of the length of ancestry tracts across the genome.
-The information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture.   
-However, how selection influences the length of tracts of ancestry around selected sites in hybrid zones has received much less attention. 
+Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Gompert2012,OTHERS}.
+Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX, LAMP and fineSTRUCTURE \yb{(CITE)} provides researchers with a measure of the length of ancestry tracts across the genome.
+\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture.
+\alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
+Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}
+
+%However, how selection influences the length of tracts of ancestry around selected sites in hybrid zones has received much less attention.
 We we present an expectation of ancestry block lengths in hybrid zone.
@@
 
-Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to have a harder time introgressing. 
-Thus we can interpret this pattern as Brownian motion with drift in which ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the "correct" side of the hybrid zone.
-Selected alleles that have, by chance, encroached deep into the other side of the hybrid have likely to have done so quite recently, as they cannot survive long in such hostile territory.   
-The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
+Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing.
+Thus we can interpret this pattern as Brownian motion with drift in which an ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the "correct" side of the hybrid zone.
+In other words, selected alleles that have, by chance, encroached deep into the other side of the hybrid zone have likely done so quite recently, as they cannot survive long in such hostile territory.
+\alisa{As a result, the hapolytype that the selected locus resides on has few opportunities for recombination on the `wrong' side, which is reflected in long blocks of contiguous ancestry.}
+%The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
 \yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
@@
 
-We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a large number of independent ancestors   (Figure~\ref{Fig:family_size}).
+We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a relatively large number of independent ancestors per individual of that ancestry.
+(Figure~\ref{Fig:family_size}).
 % is this right
-This result also reflects the recency of ancestry on the "wrong" side of the hybrid zone --- any one lineage will rapidly retreat back in time to the correct side of the zone.   
-Thus we predict
+This result also reflects the fact that, on average, haplotypes on the ``wrong'' side of the hybrid zone tend to be recent migrants and have therefore had few descendants.
 
-In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding.
+In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
 We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
@@
 A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
-Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly.
 
-These two novel finding may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown. 
-However, we as sequencing prices continue to drop and long phased reads become more common \yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
-ressarchers will soon have access to information pertaining to both of these findings and thus these results  could further aid in identifying selected sites in hybrid zones.
+These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown.
+However, as sequencing prices continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.
+\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
+\alisa{In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available data, surrounding selected loci could offer an approximation to these findings.}
 
-\subsection{Inference}
+\subsection{Paths to inference}
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to stand genomes from hybrid zones to identify putatively selected loci.
+\alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, }
 

### 2016-02-15 drafts/hybrid_zone.tex
added missing figures

@@
 This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to stand genomes from hybrid zones to identify putatively selected loci.
-\alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, }
+\alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, with some simplifying approximations, it is possible that a future framework could be constructed for inference of migration rate, hybrid zone age and selection.??}
 
 \subsection{Assumptions}
-We made numerous simplifying assumptions to derive our analytical and simulation results. 
-Thus before our results can be profitably used for inference, it is worthwhile to  reflect on our assumptions and their consequences.
+We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations.
+%Thus before our results can be profitably used for inference, it is worthwhile to  reflect on our assumptions and their consequences.
+Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
+This situation is rare in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}
 
-One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
-Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
-We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
-An additional assumption of our analytical model is that we ignore coalescence.  
-The broad agreement between our analytical and theoretical results suggests that this assumption will also have negligible effect on our results.
+The assumption about the density of selected loci, however, may have greater consequences.
+Our model, which focuses on a single selected site is likely to be still relevant to cases in which there are only a small number of selected loci scattered throughout the genome.
+However, our predictions may differ significantly from scenarios in which the density of selected sites is high.
+Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \yb{(CITE: (BARTON 1983; Kruuk 1989)}.
+Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting  greater $s$ and relatively short map distance between linked incompatibilities -- \alisa{and will impact expectations of genome-wide patterns of clines and block lengths.
+In terms of detecting targets of selection, such sites will be a common limitation among many approaches to identifying selection.}
 
-However, other assumptions of our model may be of greater consequence. 
-Importantly, our work considered only a single selected site.   
-While a small number of selected sites will likely have little influence on our results, a higher density of selected sites could. 
-Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \yb{(CITE: (BARTON 1983; Kruuk 1989)}.
-Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting this great $s$ and the map distance between these linked incompatibilities.
+For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape.
+
+Lastly, the analytic model presented here lacks coalescence.
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy.} The broad agreement between our analytic predictions and simulations suggests that this assumption does not impact our general results.
+
+%One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
+%Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
+%We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
+%An additional assumption of our analytical model is that we ignore coalescence.  
+%The broad agreement between our analytical and theoretical results suggests that this assumption will also have negligible effect on our results.
 
@@
 Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
+
+\alisa{epistasis}
 

### 2016-02-16 drafts/hybrid_zone.tex
fixed some references

@@
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher1954, Chapman \& Thomson; Barid et al)}.   
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
 The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
@@
 We consider a simple case of under-dominance. 
-\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, BARTON and GALE 1993} which also generate clines like those degenerated under complex epistatic systems}.
+\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those degenerated under complex epistatic systems}.
 Specifically, here the fitness of an individual is  determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
@@
 where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as derived in \citep{bazykin}; for discussion of diploidy see \citep{diploidcline}.
+as derived in \citep{Bazykin1969}; for discussion of diploidy see \citep{diploidcline} \alisa{not sure which reference this is}.
 There is not an exact analytic solution to this equation, 
 but a good approximation to the equilibrium solution is that
-$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{bazykin}.
+$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}.
 Considering the numerous approximations going into this,
@@
 by following lineages as they move between backgrounds,
-as done for instance in \citet{selectioncoal,ralph2015patchy}.
+as done for instance in \citet{selectioncoal,Ralph2015}.
 Consider the collection of $A$ alleles found at location $x$.
@@
 in practice it can take quite some time for even unlinked neutral loci to homogenize,
-due to the decreased fitness of heterozygotes \citep{bartonbarrier}
+due to the decreased fitness of heterozygotes \citep{Barton1986}
 and the relative slowness of diffusive movement \citep{Sedghifar2015}.
@@
 Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Gompert2012,OTHERS}.
-Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX, LAMP and fineSTRUCTURE \yb{(CITE)} provides researchers with a measure of the length of ancestry tracts across the genome.
+Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
 \alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture.
@@
 However, as sequencing prices continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.
-\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
+%\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
 \alisa{In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available data, surrounding selected loci could offer an approximation to these findings.}
@@
 This situation is rare in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
-\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989; Barton1993}
 
@@
 However, our predictions may differ significantly from scenarios in which the density of selected sites is high.
-Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \yb{(CITE: (BARTON 1983; Kruuk 1989)}.
+Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \citep{Barton1983; Kruuk1999)}.
 Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting  greater $s$ and relatively short map distance between linked incompatibilities -- \alisa{and will impact expectations of genome-wide patterns of clines and block lengths.
@@
 For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape.
+This allows us to derive analytic expressions that track lineages, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}.
+However, this assumes a large population density limit that excludes recombination and makes linages Markov \alisa{not sure about what point this is making?
+Is there a better way to state this for our intended audience?}.
+On the other hand, our simulations model regularly spaced demes.
+In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.\\
 
 Lastly, the analytic model presented here lacks coalescence.
-\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy.} The broad agreement between our analytic predictions and simulations suggests that this assumption does not impact our general results.
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.} The broad agreement between our analytic predictions and simulations suggests that this assumption does not impact our general results.
 
@@
 
-Another important assumption is the smooth and one dimensional spatial structure of the geographic landscape.  
-Patchily connected populations, or other esoteric forms of population structure may predominate on the edges of species ranges' where hybrid zones may occur.  
-While, it is still likely that the genomic signature of longer chunks of ancestry around selected sites will hold in these situations, the strength of this expectation is unclear.
-
-The tracking lineages given selection trick is an example of the more general idea of duality for particle systems.
-It also assumes a large pop density limit that excludes recombination and makes lineages Markov;
-if things are very patchy this might not be a good one.
 Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
@@
 \alisa{epistasis}
-
-Topics include\\
-- ignoring coalescent\\
-- single locus model, selection is STRONG\\
+\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
 
 \subsection{Islands of divergence}  
-The past years have witnessed evocative metaphors f seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
+The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
 The metaphors of "islands", "continents" and "archipelagoes" of divergence have inspired much research (and controversy) speciation genomics, but have been largely descriptive.  

### 2016-02-16 drafts/hybrid_zone.tex
fixed some references

@@
 To compute correlations in local ancestry
-(i.e., ``ancestry disequilibrium'', as in \citet{pool,brandvainthisvolume}),
+(i.e., ``ancestry disequilibrium'', as in \citet{Pool2015,brandvainthisvolume}),
 we need only follow lineages at two sites, instead of an entire region,

### 2016-02-16 drafts/hybrid_zone.tex
yb wrote more discussion

@@
 
-\subsection{Additional signatures around selected loci}
+\subsection*{Additional signatures around selected loci}
 
@@
 
-\subsection{Paths to inference}
+\subsection*{Paths to inference}
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 
-\subsection{Assumptions}
+\subsection*{Assumptions}
 We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations.
@@
 
-Lastly, the analytic model presented here lacks coalescence.
-\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.} The broad agreement between our analytic predictions and simulations suggests that this assumption does not impact our general results.
+Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
+The broad agreement between our analytic predictions and simulations that includee an explicit pefdigree structure suggests that this assumption does not have a sever qualitative impact on our major results.
 
@@
 
-\subsection{Future directions}
+\subsection*{Future directions}
+Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
+We suggest that such signatures could be combined in an inference framework that aims to identify incompatibilities from genome scans of hybrid zones.  
+However, doing so requires careful evaluation of potential signals of an incompatibility. 
+For example, should we consider jointly lengths of ancestry and allele frequencies to identify incompatibilities?
+Will dividing haplotype length by the length of adjacent haplotypes add power to tests because of the extra information contained in this measure, or add noise because of chance variation in these statistics.
+Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
 
-Some words about how confidently calling tracts is the way of the future.\\
-How can this be applied to detecting selection?
+Throughout, we focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \yb{(CITE: M.L.
+Arnold Transfer and origin of adaptations through natural hybridization: Were Anderson and Stebbins right?
+Plant Cell, 16 (2004), pp.
+562?570)}.
+The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin .\yb{CITE http://www.sciencedirect.com/science/article/pii/S0960982211007160}.
+Future work could dress expected genomic signatures of adaptive introgression in hybridizing populations using the number and length of haplotypes of non-native ancestry. 
+Doing so requires an understanding of how positive selection shapes ancestry block lengths in these cases.
+
+\subsection*{Conclusion}
+SOMETHING AWESOME
 

### 2016-02-16 drafts/hybrid_zone.tex
minor changes

--- 2016-02-16 drafts/hybrid_zone.tex
@@
 To compute correlations in local ancestry
-(i.e., ``ancestry disequilibrium'', as in \citet{pool,brandvainthisvolume}),
+(i.e., ``ancestry disequilibrium'', as in \citet{Pool2015,Schumer2016}),
 we need only follow lineages at two sites, instead of an entire region,

### 2016-02-16 drafts/hybrid_zone.tex

--- 2016-02-16 drafts/hybrid_zone.tex
@@
 To compute correlations in local ancestry
-(i.e., ``ancestry disequilibrium'', as in \citet{Pool2015,brandvainthisvolume}),
+(i.e., ``ancestry disequilibrium'', as in \citet{Pool2015,Schumer2016}),
 we need only follow lineages at two sites, instead of an entire region,

### 2016-02-16 drafts/hybrid_zone.tex
edits

@@
 \usepackage{amssymb}
+\usepackage{setspace}
 
@@
 
-\title{Ancestry linkage in hybrid zones}
+\title{Blocks of extended ancestry around selected loci in hybrid zones}
 \author{}
@@
 
+\usepackage{lineno}
+
+\usepackage{authblk}
+
+\begin{document}
+
+\thispagestyle{empty}
+\author[*]{Alisa Sedghifar}
+\author[$\S$]{Yaniv Brandvain}
+\author[$\dag$]{Peter Ralph}
+
+\affil[*]{\small{Lewis-Sigler Institute for Integrative Genomics and Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, 08544}}
+%\affil[2]{Department of Evolution and Ecology, University of California, Davis, Davis, California, United Sates of America}
+\affil[$\S$]{\small{Department of Plant Biology, University of Minnesota, St.~Paul, Minnesota, 55108}}
+\affil[$\dag$]{\small{Department of Molecular and Computational Biology, University of Southern California, Los Angeles, California, 90089}}
+
 %%% BEGIN DOCUMENT
-\begin{document}
 
@@
 
-\section{Introduction}
+\abstract{Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation.
+Many hybrid zones are characterized by some form of selection against hybrid genotypes, and here we use a combination of theory and simulations a to gain a better understanding of how this impacts patterns of introgression within the genome, and throughout space.
+Based on a simple model of single-locus under dominance, we find that ancestral blocks surrounding loci experiencing selection are longer than average, especially when removed from the center of the hybrid zone.
+The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.}
+
+\linenumbers
+\doublespacing
+
+\section*{Introduction}
 The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
-When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
-\yb{Thus, \alisa{ancestries} originating from the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
+When formerly isolated populations come back into contact \yb{before reproductive isolation is complete, gene flow is still possible.
+Thus, \alisa{ancestries} originating from each of the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
 %In addition to steep clines in ancestry,
-\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}
-\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.}
-Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
-\yb{The expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
+\yb{The recent exchange of genetic material between diverged populations leaves a distinct pattern of variation across the genome, in which long tracts of divergent haplotypes from each ancestral population  are broken up by historical recombination events.}
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
-The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
+\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
+While these patterns of co-ancestry have been described in multi-locus systems, the  \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
+
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
+The expected distribution of lengths of unbroken ancestry in admixed populations have been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
-Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
-The effect on patterns of ancestral block lengths is, however, less well understood. 
+Variation in selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton 1979}.
+The effect on patterns of ancestral block lengths is, however, less well explored theoretically and, \alisa{while empirical patterns of block length in hybrid zones have also not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
+} %GREAT SENTENCE!!
+
 Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-\alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
-} %GREAT SENTENCE!!
 

### 2016-02-17 drafts/hybrid_zone.tex
edits to intro

@@
 
-Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-
 The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
@@
 %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow.
-\yb{The focus on cline width provides an expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim is to extend this to obtain an expectation for how selection shapes covariance in ancestry over \emph{genomic space}}
+\yb{This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim here  is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}}.
+
+Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
+Our goal is to understand how selection against hybrids interacts with dispersal and recombination interact to produce observed patterns ancestry around selected sites in recently established hybrid zones.
+In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
+In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one.
+If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and  patterns of introgression can provide information about selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
+\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
+
+Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
+We therefore expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
+As a corollary to this , conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
+This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
+
+Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
+\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
 
@@
 
-Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci.
+%Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci.
 %Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
 % YB I don't think we are actually asking about density of selected sites, are we?
-Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}.
+%Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}.
 %in present day populations.
-This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
-In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
-\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
-\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
+%This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
+%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
+%If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+
 %In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
@@
 
-Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
-This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events. 
-Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
-\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
-
-\alisa{Moved some points here to the discussion.}
+%Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
+% is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
 

### 2016-02-17 drafts/hybrid_zone.tex
modified intro

@@
 \abstract{Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation.
-Many hybrid zones are characterized by some form of selection against hybrid genotypes, and here we use a combination of theory and simulations a to gain a better understanding of how this impacts patterns of introgression within the genome, and throughout space.
-Based on a simple model of single-locus under dominance, we find that ancestral blocks surrounding loci experiencing selection are longer than average, especially when removed from the center of the hybrid zone.
+\yb{Some form of selection against hybrid genotypes is likely common in many hybrid zones.
+Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space.
+\yb{We find that ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks}, especially when removed from the center of the hybrid zone.
 The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.}
@@
 Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
-The expected distribution of lengths of unbroken ancestry in admixed populations have been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.
+The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
-Variation in selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton 1979}.
-The effect on patterns of ancestral block lengths is, however, less well explored theoretically and, \alisa{while empirical patterns of block length in hybrid zones have also not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
+Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton 1979}.
+The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
 } %GREAT SENTENCE!!
@@
 We therefore expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-As a corollary to this , conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
+As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
 This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.

### 2016-02-17 drafts/hybrid_zone.tex
started on the methods

--- 2016-02-17 drafts/hybrid_zone.tex
@@
 \documentclass[12pt]{article}
-\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
+\usepackage[margin=0.8in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
 \geometry{a4paper} % or letter or a5paper or ... etc
@@
 We consider two isolated populations --- labeled \emph{Species $A$} and \emph{Species $B$} ---  that come into contact $\tau$ generations in the past.
-Individuals move forward in time by sampling a displacement from a Gaussian distribution. 
-We consider a simple case of under-dominance. 
+To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.
+Dispersal is  such that individuals sample their displacement from a Gaussian distribution.
+This has the desirable property that we can consider movement of lineages both backwards and forwards in time.
+In terms of selection, we consider a simple case of under dominance at a single locus.
+Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
 \yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those degenerated under complex epistatic systems}.
-Specifically, here the fitness of an individual is  determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
-To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.
 
@@
 \paragraph{Establishment of  \sout{the selected cline} \alisa{a cline at the selected locus}}
-Consider a single locus, with alternate alleles fixed in the two parental populations,
+Our simple model of underdominance considers a single locus, with alternate alleles fixed in the two parental populations,
 for which heterozygotes have fitness $1-s$, and homozygotes are of equal fitness.
 The deterministic theory predicts that after secondary contact,
-these alleles will form a stable cline.
-To study gene flow past this cline we first need to understand how the cline itself is formed.
+these alleles will form a stable cline \citep{Barton1979}.
+To study gene flow past this cline we \alisa{turn our attention to} how the cline itself is formed.
 Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
-Let $\sigma^2$ be the mean squared distance between parent and offspring.
-Assuming local random assortment of alleles into diploids,
-that the habitat and dispersal are homogeneous,
-and that $s$ and $\sigma$ are small, 
-the commonly-used equation that $p$ solves is
+Let $\sigma^2$ be the mean squared distance between parent and offspring \alisa{(i.e. the variance of the distribution of dispersal distances)}.
+Assuming that: (i) alleles locally assort into diploids randomly, (ii) habitat and dispersal are homogeneous, and (iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
+
 \begin{align} \label{eqn:cline_pde}
@@
 as derived in \citep{Bazykin1969}; for discussion of diploidy see \citep{diploidcline} \alisa{not sure which reference this is}.
-There is not an exact analytic solution to this equation, 
-but a good approximation to the equilibrium solution is that
+Although there is not an exact analytic solution to this equation, a good approximation to the equilibrium solution is that
 $\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}.
-Considering the numerous approximations going into this,
-the important attributes are that 
-the stable cline has width of order $\sigma/\sqrt{s}$,
+This relies on numerous approximations, but importantly, the stable cline has width of order $\sigma/\sqrt{s}$,
 and decays exponentially.
-Furthermore, the cline is established through diffusion \alisa{of individuals?}, which is slowed down somewhat by selection against heterozygotes,
-but is still diffusive, and so it will take time of order $1/s$
-to reach its stable width \alisa{(is this something that has been shown before?)} This can be shown by rescaling space by $\sigma/s$ and time by $1/s$ in equation \eqref{eqn:cline_pde}, and seeing that the result \alisa{(what result?)} doesn't depend on $\sigma$ or on $s$.
+Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} doesn't affect this solution, and so the cline establishes over a timescale of the order $1/s$.
+In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes.
+\alisa{I don't think I understand the independence from $\sigma$?}
 

### 2016-02-17 drafts/hybrid_zone.tex
some methods

@@
 
-\paragraph{Establishment of  \sout{the selected cline} \alisa{a cline at the selected locus}}
+\paragraph{Establishment of  \alisa{a cline at the selected locus}}
 Our simple model of underdominance considers a single locus, with alternate alleles fixed in the two parental populations,
@@
 \paragraph{Lineages at the selected locus}
-Now suppose that the frequency profile of the selected allele $p(x,t)$ is given.
+Now suppose that the frequency profile of the selected allele $p(x,t)$ is known.
 We will understand how lineages of sampled individuals behave,
@@
 as done for instance in \citet{selectioncoal,Ralph2015}.
+
 Consider the collection of $A$ alleles found at location $x$.
-The expected number of such alleles produced by location $y$ is proportional to 
-the number of $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
-looking in the other direction, this is also the probability that an $A$ allele at $x$ had a parent at $y$.
-In other words, lineages move as the random walk determined by dispersal,
-but biased towards areas that produce more of the selected allele that they carry.
+The expected number of such alleles that had a parent at location $y$ is proportional to the number of  $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
+In other words, lineages move as a random walk determined by dispersal,
+but biased towards locations where the selected allele they carry is at higher frequency.
+
 Making the same assumptions underlying equation \eqref{eqn:cline_pde},

### 2016-02-17 drafts/hybrid_zone.tex
moar edits

@@
 \documentclass[12pt]{article}
-\usepackage[margin=0.8in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
+\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
 \geometry{a4paper} % or letter or a5paper or ... etc
@@
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
-Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton 1979}.
+Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}.
 The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
@@
 
-\section{Methods}
+\section*{Methods}
 \subsection*{The Model}
@@
 both at and away from the selected site,
-by following lineages as they move between backgrounds,
+by following lineages as they move between selected backgrounds,
 as done for instance in \citet{selectioncoal,Ralph2015}.
@@
 Consider the collection of $A$ alleles found at location $x$.
-The expected number of such alleles that had a parent at location $y$ is proportional to the number of  $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
-In other words, lineages move as a random walk determined by dispersal,
+The expected number of such alleles that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$.
+In other words, lineages move as a random walk determined by the dispersal kernel,
 but biased towards locations where the selected allele they carry is at higher frequency.
 
-Making the same assumptions underlying equation \eqref{eqn:cline_pde},
-we show in Appendix \ref{apx:lineage_derivation}
-that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
-Since $p=1$ far from the zone to the left, 
-and is proportional to $\exp(-x\sqrt{s}/\sigma)$ to the right,
+\alisa{Maybe describe, in a few words, how you're constructing the model that is in the Appendix.}
+Making the same assumptions underlying equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
+how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
+with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill \alisa{better word?} on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
+At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
+Therefore, roughly, lineages on 
 then roughly, lineages on ``their own'' side wander randomly,
 while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
-back towards the side where they are more common.
+back towards the side where they are more common. \alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
 Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
-at the time of secondary contact,
-this push must get more intense the closer it is to the time of secondary contact.
+at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
 Lineages of $B$ alleles do the same thing, of course, but in the opposite direction.
+\alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}
 
+\alisa{Where did the line numbers go??}
 \paragraph{Lineages at linked loci}
-The behavior of a lineage at a linked locus is similar,
+The behavior of a lineage at a linked locus is similar to a selected locus,
 except that when in heterozygotes the lineage may move between backgrounds:
@@
 whose ancestors will tend to be more from the right.
-Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
-If $X_t$ is the location of its ancestor $t$ generations ago,
+
+Suppose we sample an allele today ($t=0$) at a locus at recombination distance $r$ to the selected site.
+If $X_t$ is the geographic location of its ancestor $t$ generations ago,
 and $Z_t$ is the identity of the selected allele that ancestor carried,
-then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$,
-and that $Z$ jumps between $A$ and $B$ at rate either $r p$ or $r(1-p)$,
-depending on the identity of $Z$ \alisa{is $t$ the time of recombination?}.
+then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ (same as previously described movement of a selected allele),
+and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$)or $r p$ (for $B$ to $A$).
+
 In symbols,
@@
 
-\textbf{figure of a lineage moving back and forth?} \alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}
-
-\paragraph{How fast do linked clines relax?}
-Although loci not under selection can in principle spread across the cline unimpeded,
-in practice it can take quite some time for even unlinked neutral loci to homogenize,
-due to the decreased fitness of heterozygotes \citep{Barton1986}
-and the relative slowness of diffusive movement \citep{Sedghifar2015}.
-Suppose we sample across at cline $\tau$ units of time after secondary contact.
-How far away from the selected locus do we expect to see clines that have begun to flatten out more
-than the selected cline?
-A linked cline will have begun to flatten out if there is an appreciable chance that
-a lineage at the linked locus initially linked to an $A$ allele 
-at some point has recombined from an individual carrying the $B$ allele.
-This implies that loci closer than $1/\tau$ to the selected site
-certainly won't have a substantially wider cline than the selected cline.
-\plr{Figure showing this? 
-    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
-    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
-    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
-}
-Of course, this calculation assumes that every generation provides an opportunity for recombination 
-onto the other background,
-but only recombinations in heterozygotes will move the linked lineage between selected backgrounds.
-Since heterozygotes for the selected allele are only found at high frequency in the cline,
-and lineages are generally pushed away from the cline but have no bias far away,
-the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
-and so the width of the genomic region showing clines about the selected locus could be 
-substantially larger than $1/\tau$.
-However, this distinction appears hard to observe for realistic parameter values. \alisa{perhaps move some of this to discussion?}
+\alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}
 
@@
 where $z$ can be either $A$ or $B$.
-Denote this probability $q_z(x,t,r)$.
+We denote this probability $q_z(x,t,r)$.
 In other words,
@@
 \end{align}
-The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation
+The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation. \alisa{Is the above equation true? Shouldn't it depend on $r$?}
 \begin{align}
@@
 \end{align}
-with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$.
+with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$. \alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
 We will have more use for the differential operators on the right-hand sides of these equations,
 so write these also as 
-$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$
+$$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$$
 and 
-$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$.
+$$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$$.
 \alisa{A few words about what $G$ is?}
@@
 %%%%%%
-\begin{figure}
-    \begin{center}
-       \includegraphics[width=0.7\textwidth]{figs/example_cline}
-    \end{center}
-    \caption{
-        \textbf{Numerically computed clines,} at
-        \textbf{(A)} the selected site, which forms over time $1/s$ and then remains stable, and
-        \textbf{(B)} a linked site at distance $r=0.01$, which continues to smooth out as $\sqrt{t}$.
-        \textbf{(C)} a linked site at $r=0.1$
-        \textbf{(D)} an unlinked site.
-        Other parameters.
-       \plr{Do we want this figure? This one is just a placeholder; I'd put in two, that showed the formation and spread better.
-       This is from the PDE (we don't have time courses from sims).}
-        \label{fig:pde_clines}
-    }
-\end{figure}
+\alisa{MOVED figure here to supplement}
 
@@
 As in \citet{Sedghifar2015},
-to do this we need only keep track of both lineages when a recombination in the middle of a block occurs (Figure~\ref{Fig:schematic}),
+to do this we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
 so that lineages will act like a particular class of labeled, branching diffusions,
-where the total branching rate is conserved.
-We will not consider subsequent coalescence.
-The general description of the process is as follows.
+where the total branching rate is conserved. We will not consider subsequent coalescence.
+The general description of the process is as follows:
 The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
@@
 
-\paragraph{Spatial haplotype patterns}
-We expect that\alisa{, on average,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
-because they have had more opportunities to recombine with $B$ haplotypes.
-However, stretches of $A$ ancestry that contain the selected site are expected to be longer
-than those that do not at the same spatial location,
-because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
-concretely, they move at speed roughly $\sigma/\sqrt{s}$,
-so we expect (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago,
-and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
-\plr{(check predicted decay of mean length $1/x$ in simulations)}
-For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
+\alisa{indeed, regions surrounding the selected locus have a hard time introgressing (Figure~\ref{Fig:resistanceToIntrogression100g})}
 
@@
 To quantify haplotype abundances,
-we'd like to know the probability that the entire segment of genome between positions $a$ and $b$ relative to the selected site
-in an individual sampled at spatial location $x$ and time $t$ after the cline has formed
-is of ancestry $A$.
-Writing this probability $g_A(x,t;a,b)$,
-branching adds a term to equation \eqref{eqn:q_pde}, so that 
+we'd like to know the probability $g_A(x,t;a,b)$ that the entire segment of genome between positions $a$ and $b$, relative to the selected site,
+in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.
+This is similar to equation \eqref{eqn:q_pde} with a term added for branching, so that 
 \plr{write this equation more compactly?}
@@
 \end{align}
-where the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
+
+\alisa{where $\theta$ is the genomic position of a recombination event?}, and the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
 and $r(a,b)$ is the distance from the segment to the selected site:
@@
 The boundary conditions are $g_A(x,0;a,b)=1$ and $g_B(x,0;a,b)=0$.
-\alisa{Is the selected site at $\theta=0$?}
 
@@
 described in more detail in Appendix \ref{apx:haplotype_calcs}.
-
-%%%%%%
-\begin{figure}
-    \begin{center}
-    \end{center}
-    \caption{
-        \textbf{Haplotype length distributions},
-        comparing numerical solutions to simulations.
-        At one time, plot 
-        probability of seeing a haplotype of length at least $r$ covering that position on the genome,
-        against (a) position on the genome, at a few spatial locations;
-        or (b) spatial location, at a few positions on the genome.
-        \plr{maybe?}
-        \label{fig:haplotype_lengths}
-    }
-\end{figure}
 
@@
 These were compared to the output of neutral simulations with $s=0$
+
+The simulations are executed in R, with scripts available on
+
+ https://github.com/asedghifar/hybridzones.
 
@@
 
-\section{Results}
-\subsection*{The distribution of ancestry tract lengths across space}
-We are interested specifically in the distribution of continuous tracts of ancestry surrounding the selected locus.
+\section*{Results}
+
+\section*{Single-locus behavior}
+
+From previous results we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
+
+\paragraph{The timescale over which linked clines relax:}
+Although loci not under selection can in principle spread across the cline unimpeded,
+in practice it can take quite some time for even unlinked neutral loci to homogenize,
+due to the decreased fitness of heterozygotes \citep{Barton1986}
+and the relative slowness of diffusive movement \citep{Sedghifar2015}.
+We are interested, for samples across a cline $\tau$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
+
+A linked cline will begin to flatten out if there is an appreciable chance that a lineage at the linked locus initially linked to an $A$ allele 
+at some point recombined from an individual carrying the $B$ allele.
+This implies that loci closer than $1/\tau$ to the selected site won't have a substantially wider cline than the selected cline.
+Of course, this calculation assumes that every generation provides an opportunity for recombination on the other background, but the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote.
+
+Since heterozygotes for the selected allele are only found at high frequency in the cline,
+and lineages are generally pushed away from the cline but have no bias far away,
+the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
+and so the width of the genomic region showing clines about the selected locus could be 
+substantially larger than $1/\tau$. However, this distinction appears hard to observe for realistic parameter values.
+
+\plr{Figure showing this? 
+    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
+    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
+    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
+}
+
+\subsection*{The distribution of ancestry tract lengths}
+The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone.
+We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome.
 As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).
-From previous results we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
 
@@
 
+In the context of our model, we expect that\alisa{, in the absence of selection,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
+because they have had more opportunities to recombine with $B$ haplotypes. 
+However, stretches of $A$ ancestry that contain a selected site are expected to be longer
+than those that do not at the same spatial location,
+because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
+concretely, they move at speed roughly $\sigma/\sqrt{s}$,
+so we expect (selected) $A$ alleles at distance $x$ from the cline center
+to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago \alisa{(compared to $\sigma$?? for a neutral allele)},
+and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
+\plr{(check predicted decay of mean length $1/x$ in simulations)}
+For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
+
+%%%%%%
+\begin{figure}
+    \begin{center}
+    \end{center}
+    \caption{
+        \textbf{Haplotype length distributions},
+        comparing numerical solutions to simulations.
+        At one time, plot 
+        probability of seeing a haplotype of length at least $r$ covering that position on the genome,
+        against (a) position on the genome, at a few spatial locations;
+        or (b) spatial location, at a few positions on the genome.
+        \plr{maybe?}
+        \label{fig:haplotype_lengths}
+    }
+\end{figure}
+
 \subsection*{Lengths of adjacent blocks}
@@
 
-\alisa{Moved Yaniv's explanation to the discussion.}
 %This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
@@
 
-\section{Discussion}
+\section*{Discussion}
 
@@
 This situation is rare in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
-\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989; Barton1993}
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
 
@@
 However, our predictions may differ significantly from scenarios in which the density of selected sites is high.
-Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \citep{Barton1983; Kruuk1999)}.
+Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \citep{Barton1983,Kruuk1999}.
 Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting  greater $s$ and relatively short map distance between linked incompatibilities -- \alisa{and will impact expectations of genome-wide patterns of clines and block lengths.
@@
 
-Throughout, we focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \yb{(CITE: M.L.
-Arnold Transfer and origin of adaptations through natural hybridization: Were Anderson and Stebbins right?
-Plant Cell, 16 (2004), pp.
-562?570)}.
-The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin .\yb{CITE http://www.sciencedirect.com/science/article/pii/S0960982211007160}.
+Throughout, we focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
+The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin.
+\citep{Song2011}.
 Future work could dress expected genomic signatures of adaptive introgression in hybridizing populations using the number and length of haplotypes of non-native ancestry. 

### 2016-02-17 drafts/supplement.tex
moar edits

@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+
+\begin{figure}
+    \begin{center}
+       \includegraphics[width=0.7\textwidth]{figs/example_cline}
+    \end{center}
+    \caption{
+        \textbf{Numerically computed clines,} at
+        \textbf{(A)} the selected site, which forms over time $1/s$ and then remains stable, and
+        \textbf{(B)} a linked site at distance $r=0.01$, which continues to smooth out as $\sqrt{t}$.
+        \textbf{(C)} a linked site at $r=0.1$
+        \textbf{(D)} an unlinked site.
+        Other parameters.
+       \plr{Do we want this figure? This one is just a placeholder; I'd put in two, that showed the formation and spread better.
+       This is from the PDE (we don't have time courses from sims).}
+        \label{fig:pde_clines}
+    }
+\end{figure}
 

### 2016-02-17 drafts/hybrid_zone.tex
minor edits

@@
 Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}.
-The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
+%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
+\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
 } %GREAT SENTENCE!!
@@
 This has the desirable property that we can consider movement of lineages both backwards and forwards in time.
-In terms of selection, we consider a simple case of under dominance at a single locus.
+
+We model selection as a simple case of under dominance at a single locus.
 Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
@@
 
-Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
+Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
 The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
@@
 and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
-\plr{(check predicted decay of mean length $1/x$ in simulations)}
+\plr{(check predicted decay of mean length $1/x$ in simulations).}
 For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

### 2016-02-20 drafts/hybrid_zone.tex
edits throughout

@@
-\documentclass[12pt]{article}
+\documentclass[11pt,letterpaper]{article}
 \usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
@@
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
+
 Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}.
@@
 Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
-Our goal is to understand how selection against hybrids interacts with dispersal and recombination interact to produce observed patterns ancestry around selected sites in recently established hybrid zones.
+Our goal is to understand how selection against hybrids interacts with dispersal and recombination to produce observed patterns ancestry around selected sites in recently established hybrid zones.
 In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
@@
 Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
-\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those degenerated under complex epistatic systems}.
+\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems}.
 
-Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at time $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
+Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
 The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
-For every other locus, we keep track of $a$, the identity at $L$, across time and space.
+For every other locus, we keep track of $z$, the identity at $L$, across time and space.
 If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise.
 Recombination between two loci is modeled as a poisson process with rate $r$.
-The identity $a$ at locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
+The identity $z$ for locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
 
@@
 Here, because of negative selection, $L_B$ migrates rapidly back to $x>0$ from a position $x<0$.
-Red dots represent recombination events along the entire haplotype, and lineages that recombine off the same haplotype as $L_B$ are represented by colored lines.
-(here, the initial haplotype is broken up into 3 chunks over 100 generations).
-Subsequent recombination events can cause these chunks to change which allele they are linked to at the selected site, and here yellow lineages are linked to an allele of ancestry $A$ at the selected locus, and blue lineages are linked to a selected allele of ancestry $B$.
-From right to left, the chunks at $\tau=100$ are of ancestry $(A, B, B)$.
-}\label{Fig:schematic}
+Red dots represent recombination events along the entire haplotype.
+When a recombination separates two chunks of the sampled haplotype takes place, a branching event occurs and these splitting lineages are represented by colored lines (here the initial haplotype is broken up into 3 chunks over 100 generations).
+Subsequent recombination events can switch the selected background to which a chunk is linked.
+Here, yellow lineages are linked to $L_A$ , and blue lineages are linked to $L_B$.
+Based on position at time $\tau=100$, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
 \end{figure}
@@
 
-Consider the collection of $A$ alleles found at location $x$.
-The expected number of such alleles that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$.
+Consider the collection of $A$ alleles found at geographic location $x$.
+The expected number, among these alleles, that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$.
 In other words, lineages move as a random walk determined by the dispersal kernel,
@@
 \alisa{Maybe describe, in a few words, how you're constructing the model that is in the Appendix.}
-Making the same assumptions underlying equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
+Making the same assumptions as for  equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
 how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
@@
 
-\alisa{Where did the line numbers go??}
 \paragraph{Lineages at linked loci}
@@
 with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$. \alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
+This can be expressed as differential operators, and written as:
+
 We will have more use for the differential operators on the right-hand sides of these equations,
@@
 of a single ancestry at a given location and time.
-As in \citet{Sedghifar2015},
-to do this we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
+To quantify haplotype abundances, we start by considering a segment of the genome between positions $a$ and $b$, relative to the selected site.
+We are interested in the probability  $g_A(x,t;a,b)$ that the entire segment $(a,b)$ in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.
+
+\paragraph{General haplotype probabilities}
+As in \citet{Sedghifar2015}, we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
 so that lineages will act like a particular class of labeled, branching diffusions,
@@
 a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
-each of which proceeds as before, independently.
+each of which proceeds as before, independently. $g_A(x,t;a,b)$ is then the  probability that all branches are found on the same side when the cline formed.
 
-\alisa{indeed, regions surrounding the selected locus have a hard time introgressing (Figure~\ref{Fig:resistanceToIntrogression100g})}
-
-\paragraph{General haplotype probabilities}
-To quantify haplotype abundances,
-we'd like to know the probability $g_A(x,t;a,b)$ that the entire segment of genome between positions $a$ and $b$, relative to the selected site,
-in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.
 This is similar to equation \eqref{eqn:q_pde} with a term added for branching, so that 
@@
 Then, 
-fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
+fitnesses were computed (heterozygotes at the selected locus have fitness $1-s$; homozygotes have fitness 1),
 and in each deme $N_d$ pairs of parents are chosen, with replacement,
@@
 
-Hybrid zones of ages 10, 100 and 1000 generations \alisa{these numbers may need to be revised?} were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$.
+Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$.
 From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$).
@@
 
-\subsection*{Statistics}
+\subsection*{Measures of introgression}
 With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci.
@@
 Since heterozygotes for the selected allele are only found at high frequency in the cline,
-and lineages are generally pushed away from the cline but have no bias far away,
+and lineages \alisa{on the ``correct" side?} are generally pushed away from the cline but have no bias far away,
 the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
@@
 We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome.
+
 As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).
+When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
+This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
+A signal is also present for, but less well resolved when considering $l_\cdot$ (i.e., when ancestry is not taken into account, Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
 
-When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
-This pattern provides a potential avenue for identifying targets of selection by looking for a signal of longer-than-average tracts surrounding a position, compared to a genome-wide background.
-(Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
-A signal is also present for, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
-
-In the context of our model, we expect that\alisa{, in the absence of selection,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
+In the context of our model, we expect that\alisa{, in the absence of selection,} $l_A$ across the genome will tend to be shorter the further one into the $B$ side of the cline,
 because they have had more opportunities to recombine with $B$ haplotypes. 
@@
 \subsection*{Lengths of adjacent blocks}
-We find that the length of blocks adjacent to blocks of `wrong' ancestry surrounding a selected site are slightly shorter than average (Figure~\ref{Supp:adjacentBlocks}).
-This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
+We find that $l_A(k\pm)$, the length of blocks adjacent to blocks containing $k$, are shorter than average when $k$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side" of the side of the zone.
+(Figure~\ref{Supp:adjacentBlocks}).
+This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
 
@@
 \includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
-Here each line represents  $l_B(\mathbf{m})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
+\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
+Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
 \end{figure}
@@
 One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.
+This can potentially provide additional signal for characterizing lineages strongly influenced by selection.
 
@@
 These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown.
-However, as sequencing prices continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.
+However, as the const of sequencing continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.
 %\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
@@
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
-Specifically, a novel statistic, $l_I(m)$ -- the genetic length of ancestry surrounding a site of a given lineage -- could prove a simple and informative summary. 
-This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to stand genomes from hybrid zones to identify putatively selected loci.
+Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary.
+This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
 \alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, with some simplifying approximations, it is possible that a future framework could be constructed for inference of migration rate, hybrid zone age and selection.??}
@@
 Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
-This situation is rare in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
-\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci shares similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
+This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
 
@@
 Is there a better way to state this for our intended audience?}.
-On the other hand, our simulations model regularly spaced demes.
+In contrast, our simulations model regularly spaced demes.
 In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
@@
 \alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
-The broad agreement between our analytic predictions and simulations that includee an explicit pefdigree structure suggests that this assumption does not have a sever qualitative impact on our major results.
+The broad agreement between our analytic predictions and simulations that includee an explicit pefdigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
 
@@
 
-Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
+%Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
 
-\alisa{epistasis}
-\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
+%\alisa{epistasis}
+%\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
 
@@
 The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-The metaphors of "islands", "continents" and "archipelagoes" of divergence have inspired much research (and controversy) speciation genomics, but have been largely descriptive.  
+The metaphors of ``islands", ``continents" and ``archipelagoes" of divergence have inspired much research (and controversy) speciation genomics, but have been largely descriptive.  
 We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
@@
 Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.   
-Overall focusing on lengths of ancestry blocks across the genome brings focus onto the processes of migration and selection rather than high-level summaries far severely abstracted from the evolutionary process.
+Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection rather than high-level summaries that are somewhat abstracted from the evolutionary process.
 
@@
 Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
+We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $\tau$
 
-Throughout, we focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
+Throughout, we have focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
 The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin.
@@
 SOMETHING AWESOME
+
+Despite being a fundamental to evolution, the process of speciation remain poorly understood.
+There is great excitement that population genomic analysis of hybrid populations will provide a means by which to learn more about the biology of speciation.
+In particular, the work we present can be applied towards learning more about gene flow between species and how this is impacted by selection.
 

### 2016-02-22 drafts/hybrid_zone.tex
minor fixes to intro

@@
 
-\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+\yb{To the extent that  genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
 The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
@@
 
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
-The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.
+Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.
+The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
 
-Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}.
+Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}.
 %The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
@@
 
-The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
+The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \citep{Barton1979a,Barton1986}.  
 These models predict cline width to be narrowest \yb{(i.e.
@@
 Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
-We therefore expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
+We therefore expect \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
 As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
@@
 Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
-\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
+\yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
 

### 2016-02-22 drafts/hybrid_zone.tex
minor fixes to methods/results

@@
 
-Similar to the model presented by \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
+Similar to our recent model of a neutral hybrid zone \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
 The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
@@
 these alleles will form a stable cline \citep{Barton1979}.
-To study gene flow past this cline we \alisa{turn our attention to} how the cline itself is formed.
+To study gene flow past this cline, we \alisa{turn our attention to} how the cline itself is formed.
 Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 and decays exponentially.
-Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} doesn't affect this solution, and so the cline establishes over a timescale of the order $1/s$.
+Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} does not affect this solution, and so the cline establishes over a timescale of the order $1/s$.
 In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes.
@@
 
-From previous results we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
+From previous results, we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
 
@@
 
-A linked cline will begin to flatten out if there is an appreciable chance that a lineage at the linked locus initially linked to an $A$ allele 
+A linked cline will begin to flatten out if there is an appreciable chance that a lineage at the linked locus initially coupled with an $A$ allele 
 at some point recombined from an individual carrying the $B$ allele.
-This implies that loci closer than $1/\tau$ to the selected site won't have a substantially wider cline than the selected cline.
+This implies that loci closer than $1/\tau$ to the selected site will not have a substantially wider cline than the selected cline.
 Of course, this calculation assumes that every generation provides an opportunity for recombination on the other background, but the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote.
@@
 because they have had more opportunities to recombine with $B$ haplotypes. 
-However, stretches of $A$ ancestry that contain a selected site are expected to be longer
+However, we expect that stretches of $A$ ancestry containing a selected site will be longer
 than those that do not at the same spatial location,
-because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
-concretely, they move at speed roughly $\sigma/\sqrt{s}$,
+because lineages containing the selected site must have been inherited from the $A$ side of the cline recently. 
+Concretely, these lineages move at speed roughly $\sigma/\sqrt{s}$,
 so we expect (selected) $A$ alleles at distance $x$ from the cline center
@@
 \subsection*{The number of `unfit' ancestors}
-Although the theoretical model presented here does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
+Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
 In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.

### 2016-02-22 drafts/hybrid_zone.tex
minor fixes to discussion

@@
 
-Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Gompert2012,OTHERS}.
+Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
 Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
-\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture.
+\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g.
+][]{Hellenthal2014}.
 \alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
@@
 In other words, selected alleles that have, by chance, encroached deep into the other side of the hybrid zone have likely done so quite recently, as they cannot survive long in such hostile territory.
-\alisa{As a result, the hapolytype that the selected locus resides on has few opportunities for recombination on the `wrong' side, which is reflected in long blocks of contiguous ancestry.}
+\alisa{As a result, the haplotype that the selected locus resides on has few opportunities for recombination on the `wrong' side, which is reflected in long blocks of contiguous ancestry.}
 %The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
@@
 Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly.
+\yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}
 
@@
 \alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
-The broad agreement between our analytic predictions and simulations that includee an explicit pefdigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
+The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
 
@@
 
-\subsection{Islands of divergence}  
+\subsection{*Islands of divergence}  
 The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-The metaphors of ``islands", ``continents" and ``archipelagoes" of divergence have inspired much research (and controversy) speciation genomics, but have been largely descriptive.  
+The metaphors of ``islands", ``continents" and ``archipelagoes" of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
 We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.

### 2016-02-22 drafts/hybrid_zone.tex
.

@@
 \usepackage{setspace}
+\usepackage[hidelinks]{hyperref}
 
@@
 
-\abstract{Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation.
+\abstract{
+Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation.
 \yb{Some form of selection against hybrid genotypes is likely common in many hybrid zones.
@@
 \yb{We find that ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks}, especially when removed from the center of the hybrid zone.
-The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.}
+The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
+}
 
@@
 
+Main points:
+\begin{itemize}
+    \item 
+\end{itemize}
+
+%%%%% %%%%%% %%%%%%%
 \section*{Introduction}
+
+% summary of intro:
+%  clines happen
+%  but then decay unless selection
+%  which affects haplotypes and LD
+%  most work so far just looks at clines
+%  but haplotype blocks should be helpful
+%  here we look at that stuff
+
 The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
-When formerly isolated populations come back into contact \yb{before reproductive isolation is complete, gene flow is still possible.
-Thus, \alisa{ancestries} originating from each of the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}.
+When formerly isolated populations come back into contact 
+before reproductive isolation is complete, gene flow is still possible. 
+As two such populations interbreed,
+a gradient
+
+Thus, \alisa{ancestries} originating from each of the two diverged populations spread from their point of contact across a geographic hybrid zone 
+--- establishing a gradient of ancestry across geography, \citep{Barton1985}. 
 %In addition to steep clines in ancestry,
@@
 Assuming that: (i) alleles locally assort into diploids randomly, (ii) habitat and dispersal are homogeneous, and (iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
-
 \begin{align} \label{eqn:cline_pde}
@@
 
+For completeness, here we provide one possible derivation of equation \eqref{eqn:cline_pde},
+which gives the expected time evolution of $p(x,t)$
+under a particular concrete model.
+Other derivations are given in \citet{thesepapers};
+see especially \citet{thatone}, which does not assume locally random mating.
+A derivation in the same style is found in \citet{ralph2015patchy}.
+
+For concreteness, we will discuss a population of outcrossing plants.
+Suppose that the population is composed of a finite number of discrete demes,
+each having $N$ diploid individuals.
+Suppose that each individual at $x$ chooses to reproduce at random, 
+exponentially distributed intervals,
+producing a random number $X$ of seeds
+using pollen coming from a random location $x+R_p$.
+Each seed then disperses to a random location $x+R_s$,
+where it replaces a randomly chosen extant individual.
+Suppose that
+
+Let $\Phi_{AA}(x,t)$ be the proportion of individuals at location $x$ and time $t$
+with genotype $AA$, and likewise for $\Phi_{AB}$ and $\Phi_{BB}$.
+\begin{align*}
+    \frac{\partial}{\partial t}
+    \E[ f(\Phi_t^N) ]
+    &=
+    N \sum_x
+    \left\{
+        \E\left[ 
+            \Phi_t(x-R_s) \Phi_t(x-R_s-R_p)  # AA
+
+            + \left( \Phi_t(x-R_s) (1-\Phi_t(x-R_s-R_p)) + (1-\Phi_t(x-R_s)) \Phi_t(x-R_s-R_p) \right)
+            + (1-\Phi_t(x-R_s)) (1-\Phi_t(x-R_s-R_p))
+        \right]
+    \right\}
+\end{align*}
+
 Here put derivation of the PDE for $p(x,t)$ and also the SDE of a lineage motion.
 
-\section{Nuemrical solutions for haplotype probabilities}
+\section{Numerical solutions for haplotype probabilities}
 \label{apx:haplotype_calcs}.

### 2016-02-22 drafts/hybrid_zone.tex
.

@@
 As two such populations interbreed,
-a gradient
+individuals' genomes can be labeled according to which ancestral population they are inherited from,
+and local interbreeding and migration of such populations
+creates a gradient of ancestry assignments across geography,
+centered on their point of contact \citep{Barton1985}. 
+\plr{(check that neutral clines discussed in barton '85)}
+If the populations are sufficiently diverged,
+this process leaves a distinct pattern of variation across the genome, 
+in which long tracts of divergent haplotypes from each ancestral population  
+are broken up by historical recombination events,
+leading to substantial correlations in ancestry between distant markers
+(termed ``ancestry LD'').
+\plr{(but, ``ancestry linkage disequilibrium'' doesn't work;
+how about just ``correlations in ancestry''?}
 
-Thus, \alisa{ancestries} originating from each of the two diverged populations spread from their point of contact across a geographic hybrid zone 
---- establishing a gradient of ancestry across geography, \citep{Barton1985}. 
-%In addition to steep clines in ancestry,
-\yb{The recent exchange of genetic material between diverged populations leaves a distinct pattern of variation across the genome, in which long tracts of divergent haplotypes from each ancestral population  are broken up by historical recombination events.}
+Gene flow will eventually erase 
+both geographic clines in ancestry, and strong correlations in ancestry.
+Subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
+The mixing of ancestries of a given genome is mediated by 
+the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
+The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.
+\plr{(check Gravel; isn't that approximate?)}
 
-\yb{To the extent that  genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
-While these patterns of co-ancestry have been described in multi-locus systems, the  \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
-
-Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.
-The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.
+% Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}.
+% While these patterns of co-ancestry have been described in multi-locus systems, the \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
+\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
+} %GREAT SENTENCE!!
 Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
 
-Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}.
-%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
-\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
-} %GREAT SENTENCE!!
-
 The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \citep{Barton1979a,Barton1986}.  
-These models predict cline width to be narrowest \yb{(i.e.
-the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider.
-%and more heavily influenced by demography.} %hybrid zone age and rates of gene flow.
-\yb{This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim here  is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}}.
+Since stronger selection reduces the amount of genetic mixing,
+these models predict the narrowest clines 
+(i.e., the most rapid changes in allele frequencies)
+at loci close to selected loci, and wider clines at less tightly linked loci.
+This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}.
+Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}.
 
@@
 
-%provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
-%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.
+\plr{explain relationship to \citet{Sedghifar2015}}
 
-%Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci.
-%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.
-% YB I don't think we are actually asking about density of selected sites, are we?
-%Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}.
-%in present day populations.
-%This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
-%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-%If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+%%%%%%% %%%%%%%%% %%%%%%%%
+\section*{Methods}
 
-%In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
-%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.
+%%%%%%% %%%%%%%%% 
+\subsection*{The Model}
 
-%Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
-% is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
+We consider two isolated populations
+--- labeled \emph{Species $A$} and \emph{Species $B$} ---  
+that come into contact $\tau$ generations in the past.  
+To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.  
+After contact, the populations live across continuous geographical space,
+with random, local dispersal that we take to be Gaussian
+(although this should not affect the conclusions if the true dispersal distribution is not too fat-tailed).
 
-%In the absence of selection, genome-wide covariance in ancestry can be used to estimate timing and gene-flow in hybrid zones \cite{Sedghifar2015}.
+We model selection through only a single, underdominant locus:
+at this locus, there are two alleles, each fixed in one of the ancestral populations,
+and heterozygous individuals produce on average $1-s$ times as many offspring as either homozygote.
+Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems.
 
-%When formerly isolated populations come back into contact, a hybrid zone can form, and may be maintained stably by selection.
-In such instances, selection on some subset of the genome prevents mixing of genotypes between species.
-We refer to such hybrid zones as tension zones (as opposed to contact zones which reflect a non-equilibrium gradient in ancestry that will eventually be eroded by migration throughout the whole genome).
-Because the population genetic signatures around loci experiencing selection is going to differ from other regions of the genome, natural hybrid zones present exciting opportunities to understand species maintenance and the process of speciation.
-
-%Gene-flow across a hybrid zone is impeded by selection against migrant, or hybrid, genotypes, and as a result, produces a cline in ancestry that is expected to be steepest at and nearby the selected locus.
-Elsewhere in the genome, clines in ancestry are wider, and are more heavily influenced by hybrid zone age and rates of migration.
-Historically, this has been the most relied-upon population genetic signature, and it is only recently that abundant genomic data has opened the possibility of studying patterns of linkage within hybrid zones.
-For example, patterns of linkage disequilibrium (LD) or tract length distribution across geographic space can provide information about zone age \cite{Sedghifar2015}.
-In tensions zones geographic patterns of linkage at loci neighboring targets of selection can potentially be a powerful source of information.
-
-%Many of the traditional models of genetic variation in hybrid zones to date have focused on the equilibrium state of hybrid zones, at both selected and linked loci (lots of Barton).
-
-%-Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation.
-Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone).
-%-Gene flow across a hybrid zone may be impeded by selection against migrants.
-In older contact zones, loci that are associated with reduced fitness in heterozygotes will have steeper clines than the rest of the genome.
-There is probably a lot to learn from geographic patterns at loci neighboring selected loci (e.g.
-small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)
-%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci.
-
-\section*{Methods}
-\subsection*{The Model}
-We consider two isolated populations --- labeled \emph{Species $A$} and \emph{Species $B$} ---  that come into contact $\tau$ generations in the past.
-To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.
-Dispersal is  such that individuals sample their displacement from a Gaussian distribution.
-This has the desirable property that we can consider movement of lineages both backwards and forwards in time.
-
-We model selection as a simple case of under dominance at a single locus.
-Specifically, the fitness of an individual is determined by its genotype at a particular locus $L$, such that individuals that are heterozygous at locus $L$ have fitness reduced by $s>0$ compared to homozygotes.
-\yb{Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems}.
-
-Similar to our recent model of a neutral hybrid zone \cite{Sedghifar2015}, we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}).
+As in \citet{Sedghifar2015}, 
+we gain substantial insight by considering where the ancestors of modern-day individuals lived,
+i.e., where their linages fall across geography as we look further back towards the time of first contact.
+we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}). 
 The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
@@
 \end{figure}
+
 %As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.
@@
         \E\left[ 
-            \Phi_t(x-R_s) \Phi_t(x-R_s-R_p)  # AA
-
+            \Phi_t(x-R_s) \Phi_t(x-R_s-R_p)  % AA
             + \left( \Phi_t(x-R_s) (1-\Phi_t(x-R_s-R_p)) + (1-\Phi_t(x-R_s)) \Phi_t(x-R_s-R_p) \right)

### 2016-02-23 drafts/hybrid_zone.tex
peter comments and section in discussion

@@
 
-Main points:
-\begin{itemize}
-    \item 
-\end{itemize}
+\plr{
+Main points, make sure they appear in the abstract:
+clines establish over time $1/s$, extend along genome to at least $1/T$;
+locally disadvantageous alleles exist as recent, small families (moving at speed $\sigma/\sqrt{s}$);
+haplotype chunks longest around introgressing selected bits, flanked by shorter chunks.
+}
 
@@
 --- labeled \emph{Species $A$} and \emph{Species $B$} ---  
-that come into contact $\tau$ generations in the past.  
-To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.  
+that came into contact $T$ generations in the past.  
+We will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.  
 After contact, the populations live across continuous geographical space,
@@
 We model selection through only a single, underdominant locus:
-at this locus, there are two alleles, each fixed in one of the ancestral populations,
-and heterozygous individuals produce on average $1-s$ times as many offspring as either homozygote.
-Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems.
+at this locus, there are two alleles, one corresponding to each of the ancestral populations,
+and individuals who are heterozygous at this locus produce on average $1-s$ times as many offspring than either homozygote.
+Although most selection in hybrid zones is likely more complex than this simple situation, 
+previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993},
+or complex epistatic systems \plr{CITE}.
 
@@
 we gain substantial insight by considering where the ancestors of modern-day individuals lived,
-i.e., where their linages fall across geography as we look further back towards the time of first contact.
-we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}). 
-The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
-For every other locus, we keep track of $z$, the identity at $L$, across time and space.
-If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise.
-Recombination between two loci is modeled as a poisson process with rate $r$.
-The identity $z$ for locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
+i.e., where their lineages fall across geography as we look further back towards the time of first contact.
+We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$ at the time of secondary contact,
+i.e., if its lineage traces back to to the $A$ side.
+A block of genome is of ancestry $A$ if every locus in it does the same;
+recombination events may cause us to follow more than one lineage back to determine the block's ancestry,
+and the block will be of ancestry $A$ if all lineages trace back to the $A$ side.
+This process, in which recombination events cause such lineages to \emph{branch},
+is depicted in Figure~\ref{Fig:schematic}.
+We say that a block of genome is on the $A$ \emph{background} if it is physically linked to an allele of ancestry $A$ at the selected locus;
+if the block includes the selected site then it is necessarily on the $A$ background.
+As the identity of the selected allele determines how selection acts on the haplotype,
+and linked variation can only move between the $A$ and the $B$ background in heterozygotes,
+a key factor in these models is the density and fecundity of heterozygotes.
 
@@
 \caption{A cartoon representation of a haplotype under the model.
-The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day ($\tau=0$).
+The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day.
 The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$.
@@
 Here, yellow lineages are linked to $L_A$ , and blue lineages are linked to $L_B$.
-Based on position at time $\tau=100$, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
+Based on position 100 generations after contact, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
 \end{figure}
@@
 
-\plr{The aim here is to give both the heuristics, for how a lineage moves around
-and what that tells us about the scale on which clines happen;
-as well as writing down the PDE.  These should probably be better separated?}
+In this section we aim to give both heuristics for how lineages move, 
+that are helpful in establishing order-of-magnitude estimates,
+and analytics, mostly in terms of partial differential equations needing numerical solution.
 
-Perhaps start out with the main points:
-the cline takes time $s$ to spread out,
-and after time $\tau$ will have clines out to genetic distance of order $1/\tau$,
-maybe more. \alisa{has this not already been stated by Nick somewhere?}
-
-\paragraph{Establishment of  \alisa{a cline at the selected locus}}
-Our simple model of underdominance considers a single locus, with alternate alleles fixed in the two parental populations,
-for which heterozygotes have fitness $1-s$, and homozygotes are of equal fitness.
+\paragraph{Establishment of the cline at the selected locus}
 The deterministic theory predicts that after secondary contact,
-these alleles will form a stable cline \citep{Barton1979}.
-To study gene flow past this cline, we \alisa{turn our attention to} how the cline itself is formed.
+the two alleles at the selected locus will form a stable cline \citep{Barton1979},
+affecting neighboring loci as well.
+First, we turn our attention to how the cline at the selected locus itself is formed.
 Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
-Suppose that secondary contact occurred at time $t=0$, \alisa{(Will switching between backwards and forwards in time not get confusing?)}
+Suppose that secondary contact occurred at time $t=0$, 
 and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
-(In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
-Let $\sigma^2$ be the mean squared distance between parent and offspring \alisa{(i.e. the variance of the distribution of dispersal distances)}.
-Assuming that: (i) alleles locally assort into diploids randomly, (ii) habitat and dispersal are homogeneous, and (iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
+% (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
+Let $\sigma^2$ be the mean squared distance between parent and offspring 
+(i.e., the variance of the distribution of dispersal distance).
+Assuming that: (i) alleles locally assort into diploids randomly, 
+(ii) habitat and dispersal are homogeneous, and 
+(iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
 \begin{align} \label{eqn:cline_pde}
@@
 where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as derived in \citep{Bazykin1969}; for discussion of diploidy see \citep{diploidcline} \alisa{not sure which reference this is}.
-Although there is not an exact analytic solution to this equation, a good approximation to the equilibrium solution is that
+as derived in \citet{Bazykin1969}; for discussion of diploidy see \citep{christiansen1995genotypic}.
+Although there is no known exact analytic solution to this equation, even in steady state, 
+a good approximation to the equilibrium solution is that
 $\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
-This relies on numerous approximations, but importantly, the stable cline has width of order $\sigma/\sqrt{s}$,
-and decays exponentially.
-Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} does not affect this solution, and so the cline establishes over a timescale of the order $1/s$.
+This relies on numerous approximations, but the general conclusions should be quite robust: 
+the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
+Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} results in a dimensionless equation,
+implying that the cline establishes over a timescale of the order $1/s$. 
 In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes. 
 \alisa{I don't think I understand the independence from $\sigma$?}
+\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}
 
@@
 Now suppose that the frequency profile of the selected allele $p(x,t)$ is known.
-We will understand how lineages of sampled individuals behave,
-both at and away from the selected site,
-by following lineages as they move between selected backgrounds,
-as done for instance in \citet{selectioncoal,Ralph2015}.
-
+We will understand first how lineages of sampled individuals behave at the selected locus,
+from which lineages at linked loci require only a minor modification.
+% PLR edited to here
 Consider the collection of $A$ alleles found at geographic location $x$.
@@
 \includegraphics{figs/alleleFrequencies_sim}
-\caption{Frequency of ancestry B at a given genomic and physical position at $\tau=100$.
+\caption{Frequency of ancestry B at a given genomic and physical position at $T=100$ generations after secondary contact.
 Every line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
@@
 and the relative slowness of diffusive movement \citep{Sedghifar2015}.
-We are interested, for samples across a cline $\tau$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
+We are interested, for samples across a cline $T$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
 
@@
 at some point recombined from an individual carrying the $B$ allele.
-This implies that loci closer than $1/\tau$ to the selected site will not have a substantially wider cline than the selected cline.
+This implies that loci closer than $1/T$ to the selected site will not have a substantially wider cline than the selected cline.
 Of course, this calculation assumes that every generation provides an opportunity for recombination on the other background, but the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote.
@@
 and lineages \alisa{on the ``correct" side?} are generally pushed away from the cline but have no bias far away,
-the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
+the amount of time a lineage spends in heterozygotes should grow as $\sqrt{T}$ for large $T$,
 and so the width of the genomic region showing clines about the selected locus could be 
-substantially larger than $1/\tau$. However, this distinction appears hard to observe for realistic parameter values.
+substantially larger than $1/T$. However, this distinction appears hard to observe for realistic parameter values.
 
@@
 \includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000}
-\caption{Randomly sampled chromosomes  across a hybrid zone of age $\tau=1000$.
+\caption{Randomly sampled chromosomes  across a hybrid zone of age $T=1000$.
 Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
@@
 Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
-In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
+In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
 This is a reflection of the average size of a haplotype family.
-We find that, as distance from the zone center increases, the average family size of an ancestor from time $\tau$ decreases (Figure~\ref{Fig:family_size}).
+We find that, as distance from the zone center increases, the average family size of an ancestor from time $T$ decreases (Figure~\ref{Fig:family_size}).
 One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.
@@
 \includegraphics{figs/number_of_ancestors_tau1000}
-\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space.
+\caption{The number of individuals of ancestry B per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
 Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$).
@@
 \section*{Discussion}
+
+\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+    which we have done using both theory and simulation.
+    Our most useful observations are that clines establish over time $1/s$,
+    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+    tend to move back towards the side where they are common
+    at speed $\sigma/\sqrt{s}$.
+    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+}
 
@@
 
+\plr{(selected lineages drift back, so haplotypes are longer)}
 Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing.
@@
 %The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
+\plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
 \yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
@@
 
+\plr{(selection increases drift and migrant families are small; need better words/description here.)}
 We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a relatively large number of independent ancestors per individual of that ancestry.
@@
 
+\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
 In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
@@
 
+\plr{(we'll be able to see this soon)}
 These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown.
@@
 \subsection*{Paths to inference}
+
+\plr{(combine with haplotype length stuff above)}
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
-\alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, with some simplifying approximations, it is possible that a future framework could be constructed for inference of migration rate, hybrid zone age and selection.??}
 
 \subsection*{Assumptions}
+
+\plr{(het. disadvantage)}
 We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations.
@@
 
+\plr{(one locus only)}
 The assumption about the density of selected loci, however, may have greater consequences.
@@
 
+\plr{(1D)}
 For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape.
 This allows us to derive analytic expressions that track lineages, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}.
-However, this assumes a large population density limit that excludes recombination and makes linages Markov \alisa{not sure about what point this is making?
+\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
+However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
 Is there a better way to state this for our intended audience?}.
+\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
 In contrast, our simulations model regularly spaced demes.
 In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
-While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.\\
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
 
@@
 
-\subsection{*Islands of divergence}  
+\subsection*{Islands of divergence}
+
+\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
 The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
@@
 \subsection*{Future directions}
+
+\plr{Perhaps delete this paragraph? A small note saying should check power with simulations up where we discuss the statistic would do.}
 Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
@@
 Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
-We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $\tau$
+We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $T$
 
+\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
 Throughout, we have focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
@@
 
-\subsection*{Conclusion}
-SOMETHING AWESOME
+\subsection*{Theory and simulation}
+In this work, we have taken two complementary approaches,
+using both simulation and theory, and comparing the two.
+As usual, simulations make fewer biological simplifications,
+while theory provides more generalizable conclusions.
+To do this, we have described the branching diffusion process that approximates
+the lineages along which haplotypes are inherited.
+Since the expected motion of a lineage depends on the local frequencies of the selected alleles,
+these diffusions are time-inhomogeneous.
 
-Despite being a fundamental to evolution, the process of speciation remain poorly understood.
-There is great excitement that population genomic analysis of hybrid populations will provide a means by which to learn more about the biology of speciation.
-In particular, the work we present can be applied towards learning more about gene flow between species and how this is impacted by selection.
+The diffusion model for lineages predicts that quantities of interest solve sets of coupled partial differential equations (PDE),
+which we have written down.
+As there are no known analytical solutions to these PDE,
+we have constructed numerical solutions (and provide the source code for doing this).
+A main role of these solutions in our work has been to verify that theory based on the diffusion model of lineages
+matches the simulations.
+These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci.
+However, due to the high dimensionality of the haplotype problem (spatial position $\times$ time $\times$ endpoints of the haplotype),
+numerical solutions for mean haplotype lengths along the genome can be as computationally intensive as simulations
+(although are substantially less noisy).
+More work could be done to develop more efficient methods of solution,
+but it may be better to perform more biologically realistic forwards-time simulations that includes coalescence and drift.
+Nonetheless, the PDE approach is easily modified to provide predictions for spatial and temporally inhomogeneous systems --
+for instance, across maps of real landscapes.
 
-%% yaniv says:
-\paragraph{Signals of block length}
+\subsection*{Simulation and analysis code}
 
-\plr{or, perhaps for introduction?}
-
-\paragraph{Solutions to PDE}
-Numerical solution for the PDE can be used on real, non-flat landscapes that change with time;
-we have code and notes on solving them on github.
+These are written primarily in R, and are available HERE.
 
@@
 \bibliography{library}
+\bibliography{references}
 

### 2016-02-24 drafts/hybrid_zone.tex
peter edits

@@
 
+\newcommand{\deriv}[1]{\frac{d}{d#1}}
+\newcommand{\dderiv}[1]{\frac{d^2}{d#1^2}}
 \newcommand{\given}{\,\vert\,}
@@
 
-\usepackage{xr}
-\externaldocument{supplement}
+% doesn't work with hyperref
+% \usepackage{xr}
+% \externaldocument{supplement}
 
@@
 
+\plr{cite selection-equals-subdivided-population stuff}
+
 %%%%%%% %%%%%%%%% %%%%%%%%
@@
 from which lineages at linked loci require only a minor modification.
-% PLR edited to here
+
 Consider the collection of $A$ alleles found at geographic location $x$.
-The expected number, among these alleles, that had a parent at location $y$ in the previous generation is proportional to the number of  $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$.
+The expected number, among these alleles, that had a parent at location $y$ in the previous generation 
+is proportional to the number of $A$ alleles that were at $y$ multiplied by the per-generation probability of dispersing from $y$ to $x$. 
 In other words, lineages move as a random walk determined by the dispersal kernel,
 but biased towards locations where the selected allele they carry is at higher frequency.
+(Lineages are also biased towards regions with higher fecundity,
+and thus away from the hybrid zone, but we assume that $s$ is small and ignore this.)
 
-\alisa{Maybe describe, in a few words, how you're constructing the model that is in the Appendix.}
 Making the same assumptions as for equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
 how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill \alisa{better word?} on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
+with drift $\sigma^2 \grad \log p(x,t)$ (i.e., it moves at speed $\sigma^2$ in the potential $-\log p$).
+% \alisa{note to self to read through appendix when written}
+To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
+lived at position $x+r$ is proportional to $\P\{R=r\} p(x+r)$, where $R$ is the random dispersal distance;
+and so the mean displacement from offspring to parent is $\E[R p(x+R)]/\E[p(x+R)]$.
+% $$ = \E[ R ( p(x) + R p'(x) ) ]/ p(x) + O(R^3) $$
+Using that $\E[R]=0$ and $\E[R^2]=\sigma^2$
+and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
+\plr{Does that help at all?}
+
 At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
-Therefore, roughly, lineages on 
-then roughly, lineages on ``their own'' side wander randomly,
+Therefore, roughly, lineages on ``their own'' side wander randomly,
 while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
-back towards the side where they are more common. \alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
+back towards the side where they are more common. 
+\alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
+\plr{This isn't a construction, it's a consequence of the model, so you're right, it is maybe a "result", but I think the explanation of why it happens belongs here.}
 Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
@@
 Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. 
-\alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}
+% \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}
+
+Note that this description gives more information than the steady-state solution,
+since in that analysis only the ratio $\sigma/\sqrt{s}$ appears.
+Here we see that lineages with $\sigma=10$ and $s=.16$ move quite differently
+to lineages with $\sigma=1$ and $s=.0016$,
+reflecting strong differences in selection against heterozygotes,
+even though the stable clines have the same form.
 
@@
 except that when in heterozygotes the lineage may move between backgrounds:
-if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
+if we follow back through time a lineage at a locus we find linked to an $A$ allele, 
 it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
@@
 
-Suppose we sample an allele today ($t=0$) at a locus at recombination distance $r$ to the selected site.
-If $X_t$ is the geographic location of its ancestor $t$ generations ago,
-and $Z_t$ is the identity of the selected allele that ancestor carried,
+Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
+We will follow its lineage back through time, using $\tau$ to denote ``time ago'' to avoid confusion.
+If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
+and $Z_\tau$ is the identity of the selected allele that ancestor carried,
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ (same as previously described movement of a selected allele),
@@
 
-In symbols,
+We can describe this process formally in It\^o notation:
+with $B_\tau$ a standard Brownian motion,
+$T_B$ the first time $Z_\tau = B$, and likewise for $T_A$,
 \begin{align}
     \begin{aligned} \label{eqn:lineage_motion}
-        dX_t &= \begin{cases}
-             s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
-             s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
+        dX_\tau &= \begin{cases}
+             \sigma^2 \grad \log(p(X_\tau,\tau)) d\tau + \sigma dB_\tau \qquad & \text{if } Z_\tau = A \\
+             \sigma^2 \grad \log(1-p(X_\tau,\tau)) d\tau + \sigma dB_\tau \qquad & \text{if } Z_\tau = B 
         \end{cases} \\
-        \P\{ Z_{t+\epsilon} = B \given Z_t = A \} &= \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
-        \P\{ Z_{t+\epsilon} = A \given Z_t = B \} &= \epsilon r p(X_t,t) + O(\epsilon^2)  .
+        % \P\{ Z_{\tau+\epsilon} = B \given Z_\tau = A \} &= \epsilon r (1-p(X_\tau,\tau)) + O(\epsilon^2) \\
+        % \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B \} &= \epsilon r p(X_\tau,\tau) + O(\epsilon^2)  .
+        \P\{ T_B &> u \given Z_\tau = A \} = \exp\left( - r \int_0^u (1-p(X_s,s)) ds \right) \\
+        \P\{ T_A &> u \given Z_\tau = B \} = \exp\left( - r \int_0^u p(X_s,s) ds \right) .
     \end{aligned}
@@
 \alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}
+\plr{Connect these up more.}
 
 \paragraph{PDE for linked clines}
-Clines at linked sites are given by
-the probability that 
+We can use this diffusion model for lineages to compute expected clines in ancestry,
+i.e., the expected proportion of individuals who inherit from species $A$,
+as a function of space, time, and position on the genome.
+Precisely, we need the probability that 
 an allele sampled $t$ generations after secondary contact at location $x$,
-at recombination distance $r$ to a selected allele of type $z$
+at recombination distance $r$ to a selected allele of type $z$,
 is inherited from an individual of ancestry $A$,
@@
 \end{align}
-The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation. \alisa{Is the above equation true? Shouldn't it depend on $r$?}
+(Recall that $Z$ depends implicitly on $r$.)
+The description above implies that $q_z$ solves the following Kolmogorov backward equation:
 \begin{align}
@@
     \dot q_A(x,t,r) 
-            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
+            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                 + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
@@
                 r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
-    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
+    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
             + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
             \\ &\qquad {} + 
-            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
-    \end{aligned} \\
+            r p(x,t) (q_A(x,t,r)-q_B(x,t,r))  ,
+    \end{aligned} 
 \end{align}
-with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$. \alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
-This can be expressed as differential operators, and written as:
+with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
+and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
+\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
+\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
+applied to a backwards-time process!}
 
 We will have more use for the differential operators on the right-hand sides of these equations,
-so write these also as 
-$$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$$
+so define these as
+$G^A = \sigma^2 \grad \log(p(x,t)) \cdot \grad + (\sigma^2/2) \Delta$
 and
-$$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$$.
-\alisa{A few words about what $G$ is?}
+$G^B = \sigma^2 \grad \log(1-p(x,t)) \cdot \grad + (\sigma^2/2) \Delta$,
+and then equation \eqref{eqn:q_pde} can be written more compactly as
+\begin{align*}
+    \dot q_A &= G^A q_A + r (1-p) (q_A-q_B)
+    \dot q_B &= G^B q_B + r (1-p) (q_B-q_A) .
+\end{align*}
+In technical terms, $G^A$ and $G^B$ are the generators of the diffusions of $A$ and $B$ lineages, respectively.
 
@@
 To determine how the cline at a linked locus is expected to relax,
-we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically,
-producing, for instance, a heatmap of expected local frequencies of $A$ ancestry
-across space and time, as in Figure \ref{fig:pde_clines} \alisa{which shows}
- the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
-\alisa{denoted} $p(x,t,r)$, and can be computed as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. \alisa{Some details on numerical solution (e.g. R package?)}
-
-%%%%%%
-\alisa{MOVED figure here to supplement}
+we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically.
+For instance, Figure \ref{fig:pde_clines} shows a heatmap of $p(x,t,r)$, 
+the expected frequency of ancestry $A$ at location $x$ and time $t$ 
+at a site at recombination distance $r$ from the selected site,
+which is computed as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 
+The equations are solved numerically in R, using the ReacTran package \citep{soetaert2012reactive}.
 
 \paragraph{Haplotype lengths}
-We can also find the probability of seeing an entire block of genome (haplotype)
+We can also find the probability of seeing an \emph{entire} block of genome (haplotype)
 of a single ancestry at a given location and time.
 To quantify haplotype abundances, we start by considering a segment of the genome between positions $a$ and $b$, relative to the selected site. 
-We are interested in the probability  $g_A(x,t;a,b)$ that the entire segment $(a,b)$ in an individual sampled at spatial location $x$ at time $t$ after the cline has formed is of ancestry $A$.
+We are interested in the probability $g_z(x,t;a,b)$ 
+of finding an entire segment $(a,b)$ of ancestry $A$ in an individual having selected allele of type $z$,
+sampled at spatial location $x$ at time $t$ after the cline has formed.
 
-\paragraph{General haplotype probabilities}
-As in \citet{Sedghifar2015}, we need only keep track of both lineages when a recombination occurs within a block (Figure~\ref{Fig:schematic}),
-so that lineages will act like a particular class of labeled, branching diffusions,
-where the total branching rate is conserved. We will not consider subsequent coalescence.
-The general description of the process is as follows:
-The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
+As in \citet{Sedghifar2015}, 
+a given block of genome is inherited along a single lineage ever since the most recent recombination event
+that fell within that block.
+Prior to this, there are two lineages to follow (see Figure~\ref{Fig:schematic}),
+and so lineages act like a particular class of labeled, branching diffusions,
+where the total branching rate is conserved. 
+We do not consider subsequent coalescence.
+The general description of the process, again looking backwards in time, is as follows:
+The lineage of a segment of genome moves as a linked locus described in equations \eqref{eqn:lineage_motion},
 with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
@@
 a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
-each of which proceeds as before, independently. $g_A(x,t;a,b)$ is then the  probability that all branches are found on the same side when the cline formed.
+each of which proceeds as before, independently. 
+Then, $g_z(x,T;a,b)$ is then the probability that all branches are found on the $A$ side of the hybrid zone
+at the time of secondary contact $T$ units of time ago.
 
-This is similar to equation \eqref{eqn:q_pde} with a term added for branching, so that 
-\plr{write this equation more compactly?}
+The resulting equation si similar to \eqref{eqn:q_pde} with a term added for branching,
+which is written (omitting the $(x,t)$ for conciseness):
 \begin{align}
     \begin{aligned} \label{eqn:g_pde}
-        \dot g_A(x,t;a,b) 
-            &= G_{r(a,b)}^A g_A(x,t;a,b) 
-            - (b-a) g_A(x,t;a,b) 
+        \dot g_A(a,b) 
+        &= G^A g_A(a,b) + r(a,b) (1-p) (g_A(a,b)-g_B(a,b))
             \\ {} & \qquad 
-            + p(x,t) \int_a^b {
-                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
-                - g_A(x,t;a,b) .
+            + (1-p) \int_a^0 \left(
+                g_B(a,\theta) g_A(\theta,b)
+                + g_A(a,\theta) g_B(\theta,b)
+            \right) d\theta,
+            \\ {} & \qquad 
+            + p \int_a^b {
+                g_A(a,\theta) g_A(\theta,b) 
             } d\theta
-            \\ {} & \qquad 
-            + (1-p(x,t)) \int_a^0 {
-                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
-            } d\theta
-            \\ {} & \qquad 
-            + (1-p(x,t)) \int_0^b {
-                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
-            } d\theta,
-    \end{aligned}
+            - (b-a) g_A(a,b)  
+    \end{aligned} ,
 \end{align}
-
-\alisa{where $\theta$ is the genomic position of a recombination event?}, and the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
+where $\theta$ is the genomic position of a recombination event, 
+the integral $\int_a^0$ is zero if $a>0$ and likewise $\int_0^b$ is zero if $b<0$;
 and $r(a,b)$ is the distance from the segment to the selected site:
@@
 are hierarchical in $[a,b]$:
-the equation for haplotype identity probabilities on a segment $[a,b]$ depends only on those probabilities for smaller segments.
+the equation for haplotype identity probabilities on a segment $[a,b]$ depends only on those probabilities for segments contained in $[a,b]$.
 This allows for efficient numerical solutions,
@@
 the model described above requires only minor modifications,
-and our numerical code can be used directly.
+and our numerical code can be used directly;
+for more detail see Appendix~\ref{apx:haplotype_calcs}.
 
@@
 
-We implemented forwards-time simulations of diploid individuals on a one-dimensional row of demes
+We implemented forwards-time simulations of diploid individuals on a one-dimensional grid of demes
 with non-overlapping generations and fixed population sizes (a Wright-Fisher model).
@@
 (The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
-Migrants past either end of the range remained at the terminal demes.
+Migrants past either end of the range remain at the terminal demes.
 Then, 
-fitnesses were computed (heterozygotes at the selected locus have fitness $1-s$; homozygotes have fitness 1),
+fitnesses are computed (heterozygotes at the selected locus have fitness $1-s$; homozygotes have fitness 1),
 and in each deme $N_d$ pairs of parents are chosen, with replacement,
@@
 
-Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$.
+Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1$ and $s=0.01$.
 From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$).
 These were compared to the output of neutral simulations with $s=0$
+\plr{missing end of sentence?}
 
 The simulations are executed in R, with scripts available on
-
- https://github.com/asedghifar/hybridzones.
+\url{https://github.com/asedghifar/hybridzones}.
 
@@
 \includegraphics{figs/alleleFrequencies_sim}
-\caption{Frequency of ancestry B at a given genomic and physical position at $T=100$ generations after secondary contact.
-Every line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
-Grey lines represent the same positions from a similar simulation with no selection.}\label{alleleFreq_tau100}
+\caption{
+Frequency of ancestry B,
+across geography and at several physical positions on the genome,
+at $T=100$ generations after secondary contact. 
+Each line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
+Grey lines represent the same positions from a simulation with
+identical parameters except $s=0$.
+}\label{alleleFreq_tau100}
 \end{figure}
@@
 \subsection*{Measures of introgression}
+
+\plr{This needs to be more precise, and should define and give a name to the ratio-of-adjacents statistic here, too.}
+
 With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci.
@@
 
-From previous results, we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, the width of clines at closely linked sites are impacted (Fig:~\ref{alleleFreq_tau100}).
+From previous results, we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, clines at closely linked sites are narrower.
+This is seen, for instance, in Figure~\ref{alleleFreq_tau100}.
 
@@
 This implies that loci closer than $1/T$ to the selected site will not have a substantially wider cline than the selected cline.
-Of course, this calculation assumes that every generation provides an opportunity for recombination on the other background, but the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote.
 
-Since heterozygotes for the selected allele are only found at high frequency in the cline,
-and lineages \alisa{on the ``correct" side?} are generally pushed away from the cline but have no bias far away,
-the amount of time a lineage spends in heterozygotes should grow as $\sqrt{T}$ for large $T$,
+In principle, the genomic window about the selected site
+in which clines remain narrow could be quite a bit wider,
+since the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote,
+and heterozygotes for the selected allele are only found at high frequency in the cline.
+The majority of lineages are generally pushed away from the cline but have no bias far away,
+so the amount of time a lineage spends in heterozygotes should grow as $\sqrt{T}$ for large $T$,
 and so the width of the genomic region showing clines about the selected locus could be 
@@
 
+%%%%% %%%%%%%%%%
 \subsection*{The distribution of ancestry tract lengths}
+
 The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone.
 We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome.
+The raw data -- ancestry assignments for a few individuals sampled from across space --
+is shown in Figure~\ref{Fig:resistanceToIntrogression1000g},
+showing distinct patterns across geography and the genome.
 
-As expected, regions surrounding a locus under selection are more resistant to introrgression (Fig~\ref{Fig:resistanceToIntrogression1000g},~\ref{Fig:resistanceToIntrogression100g}).
-When present, however, we expect an unfit migrant haplotype to be longer than the genome-wide average, and indeed we observe a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Fig.~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
-This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
-A signal is also present for, but less well resolved when considering $l_\cdot$ (i.e., when ancestry is not taken into account, Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
+As expected, regions surrounding a locus under selection are more resistant to introgression,
+as seen in figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}).
+When present, however, we expect haplotypes including the locally less common allele
+to be longer than the genome-wide average.
+Indeed,
+as shown in Figures \ref{Fig:blockLengths} and \ref{Supp:blockLengthHeatmap},
+the mean length of such haplotype blocks are up to three times longer
+than the average for that geographic location.
+\plr{Write down in the section on statistics what exactly you normalize by:
+is it the mean for that geographic location?  Or, what?}
+% This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
+The mean haplotype length found without conditioning on ancestry ($l_\cdot$)
+also shows an small increase near the selected locus,
+but it is much smaller (see Figures \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}).
 
-In the context of our model, we expect that\alisa{, in the absence of selection,} $l_A$ across the genome will tend to be shorter the further one into the $B$ side of the cline,
+We expect that, in the absence of selection, blocks of $A$ ancestry across the genome 
+will tend to be shorter the further one goes into the $B$ side of the cline,
 because they have had more opportunities to recombine with $B$ haplotypes. 
 However, we expect that stretches of $A$ ancestry containing a selected site will be longer
-than those that do not at the same spatial location,
-because lineages containing the selected site must have been inherited from the $A$ side of the cline recently. 
-Concretely, these lineages move at speed roughly $\sigma/\sqrt{s}$,
-so we expect (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago \alisa{(compared to $\sigma$?? for a neutral allele)},
+than those that do not contain the selected site at the same spatial location,
+because lineages containing the selected site have usually been inherited from the $A$ side of the cline recently. 
+As discussed above, we expect these lineages move at speed roughly $\sigma/\sqrt{s}$,
+so (selected) $A$ alleles at distance $x$ from the cline center
+to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago 
+(compared to $x^2/\sigma^2$ for a neutral allele),
 and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
-\plr{(check predicted decay of mean length $1/x$ in simulations).}
+\plr{(check predicted decay of mean length $1/x$ in simulations?)}
 For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
@@
 \subsection*{Lengths of adjacent blocks}
-We find that $l_A(k\pm)$, the length of blocks adjacent to blocks containing $k$, are shorter than average when $k$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side" of the side of the zone.
+
+\plr{Define this statistic and put in the Summary Stats section. Also, move this up with the haplotype length stuff?}
+We find that $l_A(k\pm)$, the length of blocks adjacent to blocks containing $k$, are shorter than average when $k$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
 (Figure~\ref{Supp:adjacentBlocks}).
@@
 \subsection*{The number of `unfit' ancestors}
+
+\plr{This needs a bit more precise discussion.}
+
 Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
@@
 Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing.
-Thus we can interpret this pattern as Brownian motion with drift in which an ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the "correct" side of the hybrid zone.
+Thus we can interpret this pattern as Brownian motion with drift in which an ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct'' side of the hybrid zone.
 In other words, selected alleles that have, by chance, encroached deep into the other side of the hybrid zone have likely done so quite recently, as they cannot survive long in such hostile territory.
@@
 The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-The metaphors of ``islands", ``continents" and ``archipelagoes" of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
 REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
@@
 \bibliographystyle{molecularEcology}
-\bibliography{library}
-\bibliography{references}
+\bibliography{library,references}
 
 \appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
-\section{Rescaling a dicrete model to obtain the lineage motion}
+\section{Rescaling a discrete model to obtain the lineage motion}
 \label{apx:lineage_derivation}
 
-For completeness, here we provide one possible derivation of equation \eqref{eqn:cline_pde},
-which gives the expected time evolution of $p(x,t)$
-under a particular concrete model.
-Other derivations are given in \citet{thesepapers};
-see especially \citet{thatone}, which does not assume locally random mating.
-A derivation in the same style is found in \citet{ralph2015patchy}.
+For concreteness, here we describe a discrete model that rescales to the continuous model we consider.
 
-For concreteness, we will discuss a population of outcrossing plants.
-Suppose that the population is composed of a finite number of discrete demes,
-each having $N$ diploid individuals.
-Suppose that each individual at $x$ chooses to reproduce at random, 
-exponentially distributed intervals,
-producing a random number $X$ of seeds
-using pollen coming from a random location $x+R_p$.
-Each seed then disperses to a random location $x+R_s$,
-where it replaces a randomly chosen extant individual.
-Suppose that
+In this discrete model, the total number of individuals at location $x$ is $N(x)$,
+and we count ``individuals'' as \emph{haploid},
+so each individual is of type either A or B at the selected locus,
+and the proportion of individuals of type A and location $x$ and time $t$ is $p(x,t)$
+(but, we often neglect the $t$).
+Suppose that type A individuals at location $x$ reproduce at rate $s_A(x)$, 
+and likewise type B at rate $s_B(x)$.
+Assuming locally random mating, we will then have that
+$s_A(x) = 1 - s (1-p(x))$ and $s_B(x) = 1 - s p(x)$.
 
-Let $\Phi_{AA}(x,t)$ be the proportion of individuals at location $x$ and time $t$
-with genotype $AA$, and likewise for $\Phi_{AB}$ and $\Phi_{BB}$.
-\begin{align*}
-    \frac{\partial}{\partial t}
-    \E[ f(\Phi_t^N) ]
-    &=
-    N \sum_x
-    \left\{
-        \E\left[ 
-            \Phi_t(x-R_s) \Phi_t(x-R_s-R_p)  % AA
-            + \left( \Phi_t(x-R_s) (1-\Phi_t(x-R_s-R_p)) + (1-\Phi_t(x-R_s)) \Phi_t(x-R_s-R_p) \right)
-            + (1-\Phi_t(x-R_s)) (1-\Phi_t(x-R_s-R_p))
-        \right]
-    \right\}
-\end{align*}
+At reproduction, individuals recombine with others in the same location,
+with recombination occuring between the locus we follow and the selected locus with probability $r$,
+and the offspring choose a new location $y$ with probability $m(x,y)$.
+The population dynamics are random, 
+but suppose that $N(x)$ is sufficiently large that these do not vary substantially with time.
+Suppose this is a Moran model.
+There are four things that can happen:
+\begin{enumerate}
+    \item[$x\xrightarrow{AA}y$] One type A individual at location $x$ reproduces, 
+        either does not recombine or recombines with another type A,
+        and sends the offspring to $y$.
+    \item[$x\xrightarrow{AB}y$] An individual at location $x$ reproduces, 
+        recombines with the other type,
+        and sends to $y$ an offspring
+        who inherits at the selected locus from the type $B$ parent 
+        and the at the neutral locus from the type A parent.
+    \item[$x\xrightarrow{BA}y$] An individual at location $x$ reproduces, 
+        recombines with the other type,
+        and sends to $y$ an offspring
+        who inherits at the selected locus from the type A parent 
+        and the at the neutral locus from the type $B$ parent.
+    \item[$x\xrightarrow{BB}y$] One type $B$ individual at location $x$ reproduces, 
+        either does not recombine or recombines with another type $B$,
+        and sends the offspring to $y$.
+\end{enumerate}
+These four things happen at rates:
+\begin{align}
+    & x\xrightarrow{AA}y & \qquad w_{AA}(x,y) &= p(x) s_A(x) \left(1 - r (1-p(x)) \right)  N(x) m(x,y) \\
+    & x\xrightarrow{AB}y & \qquad w_{AB}(x,y) &= r p(x) (1-p(x)) \frac{s_A(x)+s_B(x)}{2} N(x) m(x,y) \\
+    & x\xrightarrow{BA}y & \qquad w_{BA}(x,y) &= r p(x) (1-p(x)) \frac{s_A(x)+s_B(x)}{2} N(x) m(x,y) \\
+    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= (1-p(x)) s_B(x) \left(1 - r p(x) \right) N(x) m(x,y) 
+\end{align}
 
-Here put derivation of the PDE for $p(x,t)$ and also the SDE of a lineage motion.
+Note that at equilibrium, we require $N$ to solve
+\begin{align}
+  0 = \sum_y N(y) m(y,x) \left\{ p(y)s_A(y)(1-p(x)) - (1-p(y))s_B(y)p(x) \right\} .
+\end{align}
 
-\section{Numerical solutions for haplotype probabilities}
+\subsection*{Lineage movement}
+
+These rates tell us the rates at which a lineage will move, backwards in time.
+For instance, the rate at which a lineage at the selected locus
+currently in a type A individual at location $x$
+jumps to another type A individual at location $y$ is equal to the rate of influx of migrants from $y$
+divided by the number of A alleles at $x$,
+or
+\begin{align}
+  r_A(x,y) &= \frac{w_{AA}(y,x)}{p(x)N(x)} \\
+  &= N(y) p(y) s_A(y) m(y,x) \frac{ 1 }{ N(x) p(x) } .
+\end{align}
+
+Let $X_t$ denote the position of the lineage of a selected locus of time $A$ at time $t$ in the past,
+and let $f$ be test function with $f(\rho)=0$.
+Then,
+\begin{align}
+    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y r_{A}(y,x) ( f(y)-f(x) ) \\
+    &= \frac{1}{N(x)p(x)} \sum_y  N(y) p(y) s_A(y) m(y,x) ( f(y) - f(x)  ) . \label{eqn:discrete_generator}
+\end{align}
+
+\subsection*{Diffusion limit}
+% from notes/lineage-movement.tex
+
+Now suppose that $m(x,y)$ is symmetric, and depends on a parameter $\sigma$ so that as $\sigma \to 0$,
+the associated random walk converges to Brownian motion, so that for an arbitrary smooth function $f$,
+\begin{align}
+    \lim_{\sigma \to 0} \sum_y \frac{ m(x,y) ( f(y) - f(x) ) }{\sigma^2} = \frac{1}{2} \dderiv{x} f(x) .
+\end{align}
+Write $f'(x) = \deriv{x}f(x)$, and note that
+\begin{align}
+    \frac{1}{\sigma^2} \sum_y g(y) m(y,x) (f(y)-f(x)) 
+    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) + (g(x)-g(y)) f(x) \right) \\
+    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) \right) \\
+    & \qquad - f(x) \frac{1}{\sigma^2} \sum_y m(y,x) (g(y)-g(x)) \\
+    &\xrightarrow{\sigma \to 0} \frac{1}{2} \dderiv{x}\left( g(x)f(x) \right) - \frac{1}{2} f(x) \dderiv{x} g(x) \\
+    &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) \\
+    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) . \label{eqn:deriv_limit}
+\end{align}
+
+Under these assumptions, combining \eqref{eqn:discrete_generator} and \eqref{eqn:deriv_limit},
+\begin{align}
+    \deriv{t} \E[f(X_{t/\sigma^2}) \given X_0=x ] &\xrightarrow{\sigma \to 0} 
+    \frac{1}{2} s_A(x) \dderiv{x} f(x) 
+    + \frac{1}{N(x)p(x)} \deriv{x} \left\{ N(x)p(x) s_A(x) \right\}  \deriv{x} f(x) \\
+    &= \frac{1}{2} s_A(x) f''(x) + \left( s_A'(x) + \log(N(x)p(x))' s_A(x) \right) f'(x) ,
+\end{align}
+i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with mean displacement
+(``drift'' in diffusion terminology) $\frac{1}{ N(x)p(x)} \deriv{x} ( N(x) p(x) s_A(x) )$ and killed at rate $\rho k(x)$.
+
+In our case, since $s_A(x) = 1 - s (1-p(x))$, the drift is $p'(x)/p(x)$ to first order in $s$;
+the time scaling by $\sigma^2$ implies that the Brownian noise and the mean displacement
+should both be scaled by $\sigma^2$.
+
+%%%%%%% %%%%%%%%%%%%%
+\section{Numerical calculation of haplotype probabilities}
 \label{apx:haplotype_calcs}.
 
-Here put some details about the numerical solutions of haplotype probabilities.
+Numerical details.
+
+\section{Questions:} 
+\begin{enumerate}
+
+    \item Do we get the same system using, say, a Wright--Fisher model?  (I assume so?)
+
+    \item Can we prove convergence of the finite model to this one, under suitable assumptions?
+A main difficulty here would be showing that the lineage spends all its time in the part of the range where $p(x)$ does not fluctuate significantly, i.e.\ is not too small\ldots but if this is not true, it would be even more interesting.
+
+\end{enumerate}
+
+%% supplement
+\section{Supplementary Figures}
+\input{supplement.tex}
 

### 2016-02-24 drafts/hybrid_zone.tex
notes from conversation

@@
 \usepackage{lineno}
+
+% from http://tex.stackexchange.com/questions/43648/why-doesnt-lineno-number-a-paragraph-when-it-is-followed-by-an-align-equation/55297#55297
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
 
@@
 \begin{align*}
-    \dot q_A &= G^A q_A + r (1-p) (q_A-q_B)
+    \dot q_A &= G^A q_A + r (1-p) (q_A-q_B) \\
     \dot q_B &= G^B q_B + r (1-p) (q_B-q_A) .
@@
 at $T=100$ generations after secondary contact. 
+\plr{what is $s$?}
 Each line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
@@
 identical parameters except $s=0$.
+    \plr{Add same figure for some different time points and maybe values of $s$ to the supplement.} 
 }\label{alleleFreq_tau100}
@@
 This is seen, for instance, in Figure~\ref{alleleFreq_tau100}.
+
+\plr{When discussing haplotype lengths, say what happens at different $s$ and $T$.}
 
@@
 Regions surrounding the selected site are failing to introgress.
+    \plr{Add same figure for some different time points and maybe values of $s$ to the supplement.} 
 }\label{Fig:resistanceToIntrogression1000g}
@@
 \caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
-Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.}\label{Fig:blockLengths}
+Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.
+    \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
+}\label{Fig:blockLengths}
 \end{figure}
@@
 
-\section{Questions:} 
-\begin{enumerate}
-
-    \item Do we get the same system using, say, a Wright--Fisher model?  (I assume so?)
-
-    \item Can we prove convergence of the finite model to this one, under suitable assumptions?
-A main difficulty here would be showing that the lineage spends all its time in the part of the range where $p(x)$ does not fluctuate significantly, i.e.\ is not too small\ldots but if this is not true, it would be even more interesting.
-
-\end{enumerate}
-
 %% supplement

### 2016-02-25 drafts/hybrid_zone.tex
.

--- 2016-02-23 drafts/hybrid_zone.tex
@@
 \usepackage{setspace}
-\usepackage[hidelinks]{hyperref}
 
@@
 
+% from http://tex.stackexchange.com/questions/43648/why-doesnt-lineno-number-a-paragraph-when-it-is-followed-by-an-align-equation/55297#55297
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
 \usepackage{authblk}
+
+\usepackage[hidelinks]{hyperref}
+\AtBeginDocument{\let\textlabel\label}
 
@@
 
-\plr{
-Main points, make sure they appear in the abstract:
-clines establish over time $1/s$, extend along genome to at least $1/T$;
-locally disadvantageous alleles exist as recent, small families (moving at speed $\sigma/\sqrt{s}$);
-haplotype chunks longest around introgressing selected bits, flanked by shorter chunks.
-}
+Main points:
+\begin{itemize}
+    \item 
+\end{itemize}
 
@@
 --- labeled \emph{Species $A$} and \emph{Species $B$} ---  
-that came into contact $T$ generations in the past.  
-We will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.  
+that come into contact $\tau$ generations in the past.  
+To fix notation, we will say that species $A$ was initially on the ``left'' of the zone of contact, which corresponds to spatial positions $x<0$.  
 After contact, the populations live across continuous geographical space,
@@
 We model selection through only a single, underdominant locus:
-at this locus, there are two alleles, one corresponding to each of the ancestral populations,
-and individuals who are heterozygous at this locus produce on average $1-s$ times as many offspring than either homozygote.
-Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993},
-or complex epistatic systems \plr{CITE}.
+at this locus, there are two alleles, each fixed in one of the ancestral populations,
+and heterozygous individuals produce on average $1-s$ times as many offspring as either homozygote.
+Although most selection in hybrid zones is likely more complex than the single locus intrinsic under-dominance we consider, previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993} which also generate clines like those generated under complex epistatic systems.
 
@@
 we gain substantial insight by considering where the ancestors of modern-day individuals lived,
-i.e., where their lineages fall across geography as we look further back towards the time of first contact.
-We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$ at the time of secondary contact,
-i.e., if its lineage traces back to to the $A$ side.
-A block of genome is of ancestry $A$ if every locus in it does the same;
-recombination events may cause us to follow more than one lineage back to determine the block's ancestry,
-and the block will be of ancestry $A$ if all lineages trace back to the $A$ side.
-This process, in which recombination events cause such lineages to \emph{branch},
-is depicted in Figure~\ref{Fig:schematic}.
-We say that a block of genome is on the $A$ \emph{background} if it is physically linked to an allele of ancestry $A$ at the selected locus;
-if the block includes the selected site then it is necessarily on the $A$ background.
-As the identity of the selected allele determines how selection acts on the haplotype,
-and linked variation can only move between the $A$ and the $B$ background in heterozygotes,
-a key factor in these models is the density and fecundity of heterozygotes.
+i.e., where their linages fall across geography as we look further back towards the time of first contact.
+we consider a lineage sampled in the present day, and determine its ancestry by its position in space at the time of first contact, $\tau$ generations in the past (Figure~\ref{Fig:schematic}). 
+The ancestry at locus $L$ is determined by its state, so that a lineage with identity $L_A$ is of ancestry $A$ and a lineage with identity $L_B$ is of ancestry $B$ regardless of its sampling location (we assume that at time $\tau$ the two species are fixed for alternate alleles at this locus).
+For every other locus, we keep track of $z$, the identity at $L$, across time and space.
+If at time $\tau$ a locus is on the same background as $L_A$, it is considered to be of ancestry $A$, and of ancestry $B$ otherwise.
+Recombination between two loci is modeled as a poisson process with rate $r$.
+The identity $z$ for locus $K$ changes when a recombination event occurs between $L$ and $K$ in an individual that is heterozygous at locus $L$.
 
@@
 \caption{A cartoon representation of a haplotype under the model.
-The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day.
+The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day ($\tau=0$).
 The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$.
@@
 Here, yellow lineages are linked to $L_A$ , and blue lineages are linked to $L_B$.
-Based on position 100 generations after contact, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
+Based on position at time $\tau=100$, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
 \end{figure}
@@
 
-In this section we aim to give both heuristics for how lineages move, 
-that are helpful in establishing order-of-magnitude estimates,
-and analytics, mostly in terms of partial differential equations needing numerical solution.
+\plr{The aim here is to give both the heuristics, for how a lineage moves around
+and what that tells us about the scale on which clines happen;
+as well as writing down the PDE.  These should probably be better separated?}
 
-\paragraph{Establishment of the cline at the selected locus}
+Perhaps start out with the main points:
+the cline takes time $s$ to spread out,
+and after time $\tau$ will have clines out to genetic distance of order $1/\tau$,
+maybe more. \alisa{has this not already been stated by Nick somewhere?}
+
+\paragraph{Establishment of  \alisa{a cline at the selected locus}}
+Our simple model of underdominance considers a single locus, with alternate alleles fixed in the two parental populations,
+for which heterozygotes have fitness $1-s$, and homozygotes are of equal fitness.
 The deterministic theory predicts that after secondary contact,
-the two alleles at the selected locus will form a stable cline \citep{Barton1979},
-affecting neighboring loci as well.
-First, we turn our attention to how the cline at the selected locus itself is formed.
+these alleles will form a stable cline \citep{Barton1979}.
+To study gene flow past this cline, we \alisa{turn our attention to} how the cline itself is formed.
 Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
-Suppose that secondary contact occurred at time $t=0$, 
+Suppose that secondary contact occurred at time $t=0$, \alisa{(Will switching between backwards and forwards in time not get confusing?)}
 and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
-% (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
-Let $\sigma^2$ be the mean squared distance between parent and offspring 
-(i.e., the variance of the distribution of dispersal distance).
-Assuming that: (i) alleles locally assort into diploids randomly, 
-(ii) habitat and dispersal are homogeneous, and 
-(iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
+(In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
+Let $\sigma^2$ be the mean squared distance between parent and offspring \alisa{(i.e. the variance of the distribution of dispersal distances)}.
+Assuming that: (i) alleles locally assort into diploids randomly, (ii) habitat and dispersal are homogeneous, and (iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
 \begin{align} \label{eqn:cline_pde}
@@
 where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as derived in \citet{Bazykin1969}; for discussion of diploidy see \citep{christiansen1995genotypic}.
-Although there is no known exact analytic solution to this equation, even in steady state, 
-a good approximation to the equilibrium solution is that
+as derived in \citep{Bazykin1969}; for discussion of diploidy see \citep{diploidcline} \alisa{not sure which reference this is}.
+Although there is not an exact analytic solution to this equation, a good approximation to the equilibrium solution is that
 $\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}.
-This relies on numerous approximations, but the general conclusions should be quite robust: 
-the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
-Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} results in a dimensionless equation,
-implying that the cline establishes over a timescale of the order $1/s$. 
+This relies on numerous approximations, but importantly, the stable cline has width of order $\sigma/\sqrt{s}$,
+and decays exponentially.
+Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} does not affect this solution, and so the cline establishes over a timescale of the order $1/s$.
 In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes.
 \alisa{I don't think I understand the independence from $\sigma$?}
-\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}
 
@@
 Now suppose that the frequency profile of the selected allele $p(x,t)$ is known.
-We will understand first how lineages of sampled individuals behave at the selected locus,
-from which lineages at linked loci require only a minor modification.
-% PLR edited to here
+We will understand how lineages of sampled individuals behave,
+both at and away from the selected site,
+by following lineages as they move between selected backgrounds,
+as done for instance in \citet{selectioncoal,Ralph2015}.
+
 Consider the collection of $A$ alleles found at geographic location $x$.
@@
 \includegraphics{figs/alleleFrequencies_sim}
-\caption{Frequency of ancestry B at a given genomic and physical position at $T=100$ generations after secondary contact.
+\caption{Frequency of ancestry B at a given genomic and physical position at $\tau=100$.
 Every line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
@@
 and the relative slowness of diffusive movement \citep{Sedghifar2015}.
-We are interested, for samples across a cline $T$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
+We are interested, for samples across a cline $\tau$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
 
@@
 at some point recombined from an individual carrying the $B$ allele.
-This implies that loci closer than $1/T$ to the selected site will not have a substantially wider cline than the selected cline.
+This implies that loci closer than $1/\tau$ to the selected site will not have a substantially wider cline than the selected cline.
 Of course, this calculation assumes that every generation provides an opportunity for recombination on the other background, but the only way to move linked lineages between selected backgrounds is via recombination in a heterozygote.
@@
 and lineages \alisa{on the ``correct" side?} are generally pushed away from the cline but have no bias far away,
-the amount of time a lineage spends in heterozygotes should grow as $\sqrt{T}$ for large $T$,
+the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
 and so the width of the genomic region showing clines about the selected locus could be 
-substantially larger than $1/T$. However, this distinction appears hard to observe for realistic parameter values.
+substantially larger than $1/\tau$. However, this distinction appears hard to observe for realistic parameter values.
 
@@
 \includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000}
-\caption{Randomly sampled chromosomes  across a hybrid zone of age $T=1000$.
+\caption{Randomly sampled chromosomes  across a hybrid zone of age $\tau=1000$.
 Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
@@
 Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
-In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
+In particular, we wish to know the relative number of ancestors from $\tau$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
 This is a reflection of the average size of a haplotype family.
-We find that, as distance from the zone center increases, the average family size of an ancestor from time $T$ decreases (Figure~\ref{Fig:family_size}).
+We find that, as distance from the zone center increases, the average family size of an ancestor from time $\tau$ decreases (Figure~\ref{Fig:family_size}).
 One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.
@@
 \includegraphics{figs/number_of_ancestors_tau1000}
-\caption{The number of individuals of ancestry B per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
+\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space.
 Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$).
@@
 \section*{Discussion}
-
-\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-    which we have done using both theory and simulation.
-    Our most useful observations are that clines establish over time $1/s$,
-    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-    tend to move back towards the side where they are common
-    at speed $\sigma/\sqrt{s}$.
-    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-}
 
@@
 
-\plr{(selected lineages drift back, so haplotypes are longer)}
 Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing.
@@
 %The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
-\plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
 \yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
@@
 
-\plr{(selection increases drift and migrant families are small; need better words/description here.)}
 We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a relatively large number of independent ancestors per individual of that ancestry.
@@
 
-\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
 In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
@@
 
-\plr{(we'll be able to see this soon)}
 These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown.
@@
 \subsection*{Paths to inference}
-
-\plr{(combine with haplotype length stuff above)}
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
+\alisa{While numerical solutions of the PDEs presented here may be computationally restrictive, with some simplifying approximations, it is possible that a future framework could be constructed for inference of migration rate, hybrid zone age and selection.??}
 
 \subsection*{Assumptions}
-
-\plr{(het. disadvantage)}
 We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations.
@@
 
-\plr{(one locus only)}
 The assumption about the density of selected loci, however, may have greater consequences.
@@
 
-\plr{(1D)}
 For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape.
 This allows us to derive analytic expressions that track lineages, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}.
-\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
-However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
+However, this assumes a large population density limit that excludes recombination and makes linages Markov \alisa{not sure about what point this is making?
 Is there a better way to state this for our intended audience?}.
-\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
 In contrast, our simulations model regularly spaced demes.
 In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
-While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.\\
 
@@
 
-\subsection*{Islands of divergence}
-
-\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
+\subsection{*Islands of divergence}  
 The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
@@
 \subsection*{Future directions}
-
-\plr{Perhaps delete this paragraph? A small note saying should check power with simulations up where we discuss the statistic would do.}
 Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
@@
 Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
-We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $T$
+We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $\tau$
 
-\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
 Throughout, we have focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
@@
 
-\subsection*{Theory and simulation}
-In this work, we have taken two complementary approaches,
-using both simulation and theory, and comparing the two.
-As usual, simulations make fewer biological simplifications,
-while theory provides more generalizable conclusions.
-To do this, we have described the branching diffusion process that approximates
-the lineages along which haplotypes are inherited.
-Since the expected motion of a lineage depends on the local frequencies of the selected alleles,
-these diffusions are time-inhomogeneous.
+\subsection*{Conclusion}
+SOMETHING AWESOME
 
-The diffusion model for lineages predicts that quantities of interest solve sets of coupled partial differential equations (PDE),
-which we have written down.
-As there are no known analytical solutions to these PDE,
-we have constructed numerical solutions (and provide the source code for doing this).
-A main role of these solutions in our work has been to verify that theory based on the diffusion model of lineages
-matches the simulations.
-These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci.
-However, due to the high dimensionality of the haplotype problem (spatial position $\times$ time $\times$ endpoints of the haplotype),
-numerical solutions for mean haplotype lengths along the genome can be as computationally intensive as simulations
-(although are substantially less noisy).
-More work could be done to develop more efficient methods of solution,
-but it may be better to perform more biologically realistic forwards-time simulations that includes coalescence and drift.
-Nonetheless, the PDE approach is easily modified to provide predictions for spatial and temporally inhomogeneous systems --
-for instance, across maps of real landscapes.
+Despite being a fundamental to evolution, the process of speciation remain poorly understood.
+There is great excitement that population genomic analysis of hybrid populations will provide a means by which to learn more about the biology of speciation.
+In particular, the work we present can be applied towards learning more about gene flow between species and how this is impacted by selection.
 
-\subsection*{Simulation and analysis code}
+%% yaniv says:
+\paragraph{Signals of block length}
 
-These are written primarily in R, and are available HERE.
+\plr{or, perhaps for introduction?}
+
+\paragraph{Solutions to PDE}
+Numerical solution for the PDE can be used on real, non-flat landscapes that change with time;
+we have code and notes on solving them on github.
 
@@
 \bibliography{library}
-\bibliography{references}
 

### 2016-02-25 sims/compare-to-theory.Rmd

@@
 rgrid <- setup.grid.1D(x.up=-0.5, x.down=0.5, N=41)
-# hap.soln <- forwards_backwards_haplotypes(s=theory.params$s, times=tt, xgrid=xgrid, rgrid=rgrid,
-#                            sigma=theory.params$sigma, 
-#                            fwds.grid=fgrid, fwds.soln=fwds.soln )
+hap.soln <- forwards_backwards_haplotypes(s=theory.params$s, times=tt, xgrid=xgrid, rgrid=rgrid,
+                           sigma=theory.params$sigma, 
+                           fwds.grid=fgrid, fwds.soln=fwds.soln )
 ```

### 2016-02-25 sims/compare-to-theory.Rmd
hap viz in compare-to-theory

@@
 We'll do this against $x$ at a few values of $r$ because that's what we have.
+(Don't mess with this next code chunk, it'll invalidate the cache, which was expensive to make.)
 ```{r theory_haps, cache=TRUE}
@@
 ```
+Here are haplotype probabilities centered on the selected site:
+```{r look_at_haps}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.soln, left=-haplen/2, right=haplen/2 )
+    image(hap.probs, zlim=c(0,1), legend=TRUE, yaxt='n',
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```
+and here are the same offset by `r rgrid$dx[1]`
+```{r look_at_haps_offset}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.soln, left=rgrid$dx[1], right=rgrid$dx[1]+haplen )
+    image(hap.probs, zlim=c(0,1), legend=TRUE, yaxt='n',
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```
+
+Now, let's get **mean lengths** as best we can.

### 2016-02-25 sims/compare-to-theory.Rmd
mean hap lengths works?

--- 2016-02-25 sims/compare-to-theory.Rmd
@@
 ```
+
+Now let's compare this to simulation:
+```{r sim_hap_lens}
+sim_lens <- sapply( chunks, sapply, mean )
+sim_cond_lens <- sapply( chunks, sapply, function (x) { if (any(x>0)) { mean(x[x>0]) } else { NA } } )
+
+hap.len.mat <- sapply( chunks.rvals, function (crv) { 
+               haplens <- mean_hap_len( hap.soln, loc=if(crv[1]==1){crv[2]-0.5}else{max(rgrid$x.mid)} ) 
+               haplens[length(tt),1+seq_along(xgrid$x.mid)]
+           } )
+
+matplot(xx, sim_lens, type='l', lty=sapply(chunks.rvals,"[",1),
+        main="mean haplotype length",
+        xlab="spatial position", ylab="mean B haplotype length (incl. zeros)")
+matlines( rev(xgrid$x.mid), hap.len.mat, lty=3 )
+legend("topleft", lty=c(sapply(chunks.rvals,"[",1),3), col=c(rep_len(1:6,ncol(sim_cond_lens)),1),
+        legend=c( sapply( chunks.rvals, function (x) { sprintf("chr %d: %0.3f",x[1],x[2]) } ), "theory") )
+```
+
+```{r get_mean_hap_lens}
+hap.pdf <- hap_cdf_to_pdf( hap.soln )
+```
+
+Here are haplotype probabilities (**not** cumulative):
+```{r look_at_haps}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.pdf, left=-haplen/2, right=haplen/2 )
+    image(hap.probs, legend=TRUE, yaxt='n',
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```
+and here are the same offset by `r rgrid$dx[1]`
+```{r look_at_haps_offset}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.pdf, left=rgrid$dx[1], right=rgrid$dx[1]+haplen )
+    image(hap.probs, legend=TRUE, yaxt='n',
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```
+
+Here are also heatmaps of **mean lengths**:
+```{r hap_lens}
+hap.centers <- c(0.0, 0.02, 0.04, 0.08, 0.16, 0.32)
+for (kloc in seq_along(hap.centers)) {
+    hap.len <- mean_hap_len( hap.pdf, loc=hap.centers[kloc] )
+    image(hap.len, legend=TRUE, yaxt='n',
+          main=paste(c("A:","B:"),sprintf("mean haplotype length at %f",hap.centers[kloc])) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```

### 2016-02-25 notes/cline-fns.R

@@ -292,7 +292,14 @@ forwards_backwards_haplotypes <- function (s, times, xgrid, rgrid, sigma=1,
     #  - spatial location
     #  - linked background
     #  - left/right endpoints (in upper triangular order)
-
+    ###
+    # so can be indexed e.g. by (where hap.soln is the output)
+    # hap.coords <- expand.grid( x=xgrid$x.mid, AB=c("A","B"), rk=1:nrow(attr(hap.soln,"r")) )
+    # hap.coords$p <- hap.soln[nrow(hap.soln),-1]
+    # hap.coords$right <- attr(hap.soln,"r")[hap.coords$rk,"right"]
+    # hap.coords$left <- attr(hap.soln,"r")[hap.coords$rk,"left"]
+    # hap.coords$len <- hap.coords$right-hap.coords$left
+    ####
     # precompute some things
     rgrid$eps <- unique(diff(rgrid$x.mid))[1]
     rgrid$zeroind <- which(rgrid$x.mid>0)[1]
@@ -387,8 +394,9 @@ AD_from_soln <- function (soln,xgrid) {
 
 
 get_hap_probs <- function ( hap.soln, left, right ) {
-    # number of sites in spatial grid
-    nx <- attr(hap.soln,"soln.dims")[1]/2
+    # Return a deSolve-class object giving haplotype probabilites for the given (left,right) endpoints,
+    # across time, space, and linked to A and B.
+    nx <- attr(hap.soln,"soln.dims")[1]/2 # number of sites in spatial grid
     # find closest interval in rgrid (hack!)
     k.ab <- which.min( (attr(hap.soln,"r")[,"left"]-left)^2 + (attr(hap.soln,"r")[,"right"]-right)^2 )
     # columns of hap.soln are time, then ordered by space, linked allele, rows of r
@@ -398,5 +406,59 @@ get_hap_probs <- function ( hap.soln, left, right ) {
     attr(dummy.soln,"r") <- attr(hap.soln,"r")[k.ab,]
     attr(dummy.soln,"nspec") <- 2
     attr(dummy.soln,"dimens") <- attr(hap.soln,"dimens")
+    return(dummy.soln)
+}
+
+hap_cdf_to_pdf <- function ( hap.soln ) {
+    # we compute probability that (a,b) is within the maximal segment
[…]

### 2016-02-25 sims/compare-to-theory.Rmd
.

--- 2016-02-25 sims/compare-to-theory.Rmd
@@
 ```
+
+Now let's compare this to simulation:
+```{r sim_hap_lens}
+sim_lens <- sapply( chunks, sapply, mean )
+sim_cond_lens <- sapply( chunks, sapply, function (x) { if (any(x>0)) { mean(x[x>0]) } else { NA } } )
+
+hap.len.mat <- sapply( chunks.rvals, function (crv) { 
+               haplens <- mean_hap_len( hap.pdf, loc=if(crv[1]==1){crv[2]-0.5}else{max(rgrid$x.mid)} ) 
+               haplens[length(tt),1+seq_along(xgrid$x.mid)]
+           } )
+
+matplot(xx, sim_lens, type='l', lty=sapply(chunks.rvals,"[",1),
+        main="mean haplotype length",
+        xlab="spatial position", ylab="mean B haplotype length (incl. zeros)")
+matlines( rev(xgrid$x.mid), hap.len.mat, lty=3 )
+legend("topleft", lty=c(sapply(chunks.rvals,"[",1),3), col=c(rep_len(1:6,ncol(sim_cond_lens)),1),
+        legend=c( sapply( chunks.rvals, function (x) { sprintf("chr %d: %0.3f",x[1],x[2]) } ), "theory") )
+```
+
+```{r get_mean_hap_lens}
+hap.pdf <- hap_cdf_to_pdf( hap.soln )
+```
+
+Here are haplotype length distributions:
+```{r hap_len_distributions}
+hap.coords <- expand.grid( x=xgrid$x.mid, AB=c("A","B"), rk=1:nrow(attr(hap.pdf,"r")) )
+hap.coords$p <- hap.pdf[nrow(hap.pdf),-1]
+hap.coords$cump <- hap.soln[nrow(hap.pdf),-1]
+hap.coords$right <- attr(hap.pdf,"r")[hap.coords$rk,"right"]
+hap.coords$left <- attr(hap.pdf,"r")[hap.coords$rk,"left"]
+hap.coords$len <- hap.coords$right-hap.coords$left
+
+```
+
+Here are haplotype probabilities (**not** cumulative):
+```{r look_at_haps}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.pdf, left=-haplen/2, right=haplen/2 )
+    image(hap.probs, legend=TRUE, yaxt='n', zlim=range(hap.probs[,-1]),
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+    locator(1)
+}
+```
+and here are the same offset by `r rgrid$dx[1]`
+```{r look_at_haps_offset}
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+    hap.probs <- get_hap_probs( hap.pdf, left=rgrid$dx[1], right=rgrid$dx[1]+haplen )
+    image(hap.probs, legend=TRUE, yaxt='n', zlim=range(hap.probs[,-1]),
+       main=paste(c("A:","B:"),paste(round(attr(hap.probs,"r"),3),collapse=" to ")) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```
+
+Here are also heatmaps of **mean lengths**:
+```{r hap_lens}
+hap.centers <- c(0.0, 0.02, 0.04, 0.08, 0.16, 0.32)
+for (kloc in seq_along(hap.centers)) {
+    hap.len <- mean_hap_len( hap.pdf, loc=hap.centers[kloc] )
+    image(hap.len, legend=TRUE, yaxt='n', zlim=range(hap.len[,-1]),
+          main=paste(c("A:","B:"),sprintf("mean haplotype length at %f",hap.centers[kloc])) )
+    axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
+}
+```

### 2016-02-25 notes/cline-fns.R

@@ -415,6 +415,7 @@ hap_cdf_to_pdf <- function ( hap.soln ) {
     # prob(a,>b) = prob(<a,>b) - prob(<a-1,>b) )
     # prob(a,b) = prob(a,>b) - prob(a,>b+1)
     #             prob(<a,>b) - prob(<a-1,>b) ) - prob(<a,>b+1) + prob(<a-1,>b+1) )
+    # except at boundaries: so, turn NAs below into zeros.
     r.coords <- attr(hap.soln,"r")
     rvals <- sort(unique(as.vector(r.coords)))
     segs.k <- cbind( left=match(r.coords[,"left"],rvals), right=match(r.coords[,"right"],rvals) )
@@ -423,11 +424,10 @@ hap_cdf_to_pdf <- function ( hap.soln ) {
             left = match( .hash(segs.k[,1]-1,segs.k[,2]), .hash(segs.k[,1],segs.k[,2]) ),    # (a-1,b)
             right = match( .hash(segs.k[,1],segs.k[,2]+1), .hash(segs.k[,1],segs.k[,2]) ),   # (a,b+1)
             both = match( .hash(segs.k[,1]-1,segs.k[,2]+1), .hash(segs.k[,1],segs.k[,2]) ) ) # (a-1,b+1)
-    dothese.segs <- (rowSums(is.na(segs.neighbors))==0)
     # here's a nrow(r.coords) x nrow(r.coords) matrix to do the conversion
-    cdf.to.pdf <- sparseMatrix( i = c((1:nrow(r.coords))[dothese.segs],as.vector(segs.neighbors[dothese.segs,])), 
-                                j = rep((1:nrow(r.coords))[dothese.segs],4), 
-                                x = rep(c(1,-1,-1,1),each=sum(dothese.segs)),
+    cdf.to.pdf <- sparseMatrix( i = c((1:nrow(r.coords)),segs.neighbors[!is.na(segs.neighbors)]), 
+                                j = rep((1:nrow(r.coords)),4)[c(rep(TRUE,nrow(r.coords)),!is.na(segs.neighbors))], 
+                                x = rep(c(1,-1,-1,1),c(nrow(r.coords),colSums(!is.na(segs.neighbors)))),
                                 dims = c(nrow(r.coords),nrow(r.coords)) )
     nxab <- attr(hap.soln,"soln.dims")[1]  # number of space x AB combinations
     for (xk in 1:nxab) {

### 2016-02-26 sims/compare-to-theory.Rmd
haplotype length comparison works!

@@
 
+```{r get_mean_hap_lens}
+hap.pdf <- hap_cdf_to_pdf( hap.soln )
+```
+
 Now let's compare this to simulation:
@@
 matplot(xx, sim_lens, type='l', lty=sapply(chunks.rvals,"[",1),
-        main="mean haplotype length",
+        main="mean haplotype length", ylim=range(sim_lens,hap.len.mat),
         xlab="spatial position", ylab="mean B haplotype length (incl. zeros)")
@@
         legend=c( sapply( chunks.rvals, function (x) { sprintf("chr %d: %0.3f",x[1],x[2]) } ), "theory") )
-```
-
-```{r get_mean_hap_lens}
-hap.pdf <- hap_cdf_to_pdf( hap.soln )
 ```
@@
 
+x.inds <- floor(seq(1,length(xgrid$x.mid),length.out=11))
+xx <- xgrid$x.mid[x.inds]
+nxab <- attr(hap.pdf,"soln.dims")[1]  # number of space x AB combinations
+r.lens <- attr(hap.pdf,"r")[,"right"] - attr(hap.pdf,"r")[,"left"]
+lenvals <- seq(0,1,length.out=21)
+hap.distrns <- sapply( x.inds, function (xk) {
+        kk <- 1 + which( ( (0:(ncol(hap.pdf)-2)) %% nxab )+1 == xk )
+        y <- tapply( hap.pdf[nrow(hap.pdf),kk], match(r.lens,sort(unique(r.lens))), sum )
+        # not really a density: need to divide
+        diff(c(0,approx( sort(unique(r.lens)), cumsum(y), xout=lenvals )$y))
+    } )
+matplot( lenvals, hap.distrns, type='l', xlab="haplotype length", ylab="probability density" )
+legend("topright", lty=1:5, col=1:6, legend=sprintf("location = %0.2f", xx))
 ```
@@
 ```{r look_at_haps}
-for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32, 0.64) ) {
     hap.probs <- get_hap_probs( hap.pdf, left=-haplen/2, right=haplen/2 )
@@
     axis(2,at=(pretty(xgrid$x.mid)-min(xgrid$x.mid))/diff(range(xgrid$x.mid)),labels=pretty(xgrid$x.mid))
-    locator(1)
 }
@@
 ```{r look_at_haps_offset}
-for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32) ) {
+for (haplen in c(0.02, 0.04, 0.08, 0.16, 0.32, 0.64) ) {
     hap.probs <- get_hap_probs( hap.pdf, left=rgrid$dx[1], right=rgrid$dx[1]+haplen )

### 2016-02-26 notes/cline-fns.R

@@ -419,7 +419,7 @@ hap_cdf_to_pdf <- function ( hap.soln ) {
     r.coords <- attr(hap.soln,"r")
     rvals <- sort(unique(as.vector(r.coords)))
     segs.k <- cbind( left=match(r.coords[,"left"],rvals), right=match(r.coords[,"right"],rvals) )
-    .hash <- function (a,b) { a*length(rvals)+b }
+    .hash <- function (a,b) { a*2*length(rvals)+b }  # factor of two avoids wraparound
     segs.neighbors <- cbind(
             left = match( .hash(segs.k[,1]-1,segs.k[,2]), .hash(segs.k[,1],segs.k[,2]) ),    # (a-1,b)
             right = match( .hash(segs.k[,1],segs.k[,2]+1), .hash(segs.k[,1],segs.k[,2]) ),   # (a,b+1)

### 2016-02-26 drafts/hybrid_zone.tex
changes to discussion

@@
 \usepackage{setspace}
+%\usepackage[hidelinks]{hyperref}
 
@@
 
-% from http://tex.stackexchange.com/questions/43648/why-doesnt-lineno-number-a-paragraph-when-it-is-followed-by-an-align-equation/55297#55297
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-  \renewenvironment{#1}%
-     {\linenomath\csname old#1\endcsname}%
-     {\csname oldend#1\endcsname\endlinenomath}}% 
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-  \patchAmsMathEnvironmentForLineno{#1}%
-  \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
-
 \usepackage{authblk}
-
-\usepackage[hidelinks]{hyperref}
-\AtBeginDocument{\let\textlabel\label}
 
@@
 
-Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
-Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
-\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g.
-][]{Hellenthal2014}.
-\alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
-Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}
+%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+%   which we have done using both theory and simulation.
+%   Our most useful observations are that clines establish over time $1/s$,
+%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+%    tend to move back towards the side where they are common
+%   at speed $\sigma/\sqrt{s}$.
+%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+%}
+
+\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
+We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
+This occurs at speed $\sigma/\sqrt{s}$.
+Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
+This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
+
+\alisa{\paragraph{Targets of selection}}
+Most attempts to identify loci under selection in hybrid zones identify alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
+Current genomic approaches can provide an additional rich source of information when applied to the study of hybrid zones.
+In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture \citep[e.g.
+][]{Hellenthal2014,Schumer2014}.
+In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone.
+Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
 

### 2016-02-26 drafts/hybrid_zone.tex

--- 2016-02-24 drafts/hybrid_zone.tex
@@
 \usepackage{setspace}
-\usepackage[hidelinks]{hyperref}
+
+%\usepackage[hidelinks]{hyperref}
 
@@
 \usepackage{authblk}
+
+\usepackage[hidelinks]{hyperref}
+\AtBeginDocument{\let\textlabel\label}
 
@@
 
-\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-    which we have done using both theory and simulation.
-    Our most useful observations are that clines establish over time $1/s$,
-    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-    tend to move back towards the side where they are common
-    at speed $\sigma/\sqrt{s}$.
-    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-}
-
-Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
-Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
-\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g.
-][]{Hellenthal2014}.
-\alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
-Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}
+%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+%   which we have done using both theory and simulation.
+%   Our most useful observations are that clines establish over time $1/s$,
+%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+%    tend to move back towards the side where they are common
+%   at speed $\sigma/\sqrt{s}$.
+%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+%}
+
+\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
+We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
+This occurs at speed $\sigma/\sqrt{s}$.
+Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
+This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
+
+\alisa{\paragraph{Targets of selection}}
+Most attempts to identify loci under selection in hybrid zones identify alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
+Current genomic approaches can provide an additional rich source of information when applied to the study of hybrid zones.
+In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture \citep[e.g.
+][]{Hellenthal2014,Schumer2014}.
+In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone.
+Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
 

### 2016-02-26 drafts/hybrid_zone.tex
heatmaps

@@
 \usepackage{setspace}
-
-%\usepackage[hidelinks]{hyperref}
 
@@
 
-\paragraph{PDE for linked clines}
+\begin{figure}
+    \includegraphics{figs/linked-frequencies}
+    \caption{
+        \textbf{Probabilities of $A$ ancestry,}
+        across space (vertical axis, in units of $\sigma$) 
+        and time (horizontal axis, in generations).
+        In each plot, color corresponds to the expected frequency of $A$ ancestry
+        at a particular location in time and space.
+        The selection coefficient is $s=.02$.
+        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
+        on a time scale of $1/s=50$ generations.
+        \textbf{Bottom left:} at an unliked site,
+        with cline flattening continuing with $\sqrt{t}$.
+        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
+        on the ancestry at the selected site,
+        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
+        as described in the text (see definition of $q_z(x,t,r)$).
+        See figure \ref{sfig:linked_cline_heatmaps_longtime}
+        for the same figure over a longer period of time.
+    }
+    \label{fig:linked_cline_heatmaps}
+\end{figure}
+
+\paragraph{Linked clines}
 We can use this diffusion model for lineages to compute expected clines in ancestry,
@@
 
-%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-%   which we have done using both theory and simulation.
-%   Our most useful observations are that clines establish over time $1/s$,
-%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-%    tend to move back towards the side where they are common
-%   at speed $\sigma/\sqrt{s}$.
-%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-%}
-
-\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
-We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
-This occurs at speed $\sigma/\sqrt{s}$.
-Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
-This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
-
-\alisa{\paragraph{Targets of selection}}
-Most attempts to identify loci under selection in hybrid zones identify alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
-Current genomic approaches can provide an additional rich source of information when applied to the study of hybrid zones.
-In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture \citep[e.g.
-][]{Hellenthal2014,Schumer2014}.
-In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone.
-Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
+\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+    which we have done using both theory and simulation.
+    Our most useful observations are that clines establish over time $1/s$,
+    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+    tend to move back towards the side where they are common
+    at speed $\sigma/\sqrt{s}$.
+    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+}
+
+Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
+Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
+\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g.
+][]{Hellenthal2014}.
+\alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
+Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}
 

### 2016-02-26 drafts/supplement.tex
peter edits
.
heatmaps

@@
-\documentclass[12pt,titlepage]{article}
+% \documentclass[12pt,titlepage]{article}
+% 
+% %% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
+% %% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
+% \usepackage{graphicx}
+% %\usepackage{figcaps}
+% \usepackage[nomarkers,notablist,nofiglist,tablesfirst]{endfloat}
+% \usepackage{amsfonts}
+% \usepackage{sectsty}
+% %\usepackage{subfigure}
+% %\usepackage{subcaption}
+% \usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
+% \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+% %\allsectionsfont{\sffamily}
+% %% THE MANUSCRIPT TITLE
+% 
+% \usepackage{caption}
+% \usepackage[labelformat=simple]{subcaption}
+% 
+% \usepackage{bibentry}
+% 
+% \usepackage{xr}
+% \externaldocument{hybrid_zone}
+% 
+% \newcommand{\alisa}[1]{{\em \color{red} #1}}
+% \newcommand{\plr}[1]{{\em \color{blue} #1}}
+% \newcommand{\yb}[1]{{\em \color{magenta} #1}}
+% 
+% 
+% \newcommand{\given}{\,\vert\,}
+% \newcommand{\st}{\,\colon\,}
+% \renewcommand{\and}{\,\&\,}
+% 
+% 
+% \begin{document}
+% \setcounter{table}{0}
+% \renewcommand{\thetable}{S\arabic{table}}
+% \setcounter{figure}{0}
+% \renewcommand{\thefigure}{S\arabic{figure}}
 
-%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
-%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
-\usepackage{graphicx}
-%\usepackage{figcaps}
-\usepackage[nomarkers,notablist,nofiglist,tablesfirst]{endfloat}
-\usepackage{amsfonts}
-\usepackage{sectsty}
-%\usepackage{subfigure}
-%\usepackage{subcaption}
-\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-%\allsectionsfont{\sffamily}
-%% THE MANUSCRIPT TITLE
-
-\usepackage{caption}
-\usepackage[labelformat=simple]{subcaption}
-
-\usepackage{bibentry}
-
-\usepackage{xr}
-\externaldocument{hybrid_zone}
-
-\newcommand{\alisa}[1]{{\em \color{red} #1}}
-\newcommand{\plr}[1]{{\em \color{blue} #1}}
-\newcommand{\yb}[1]{{\em \color{magenta} #1}}
-
-\newcommand{\given}{\,\vert\,}
-\newcommand{\st}{\,\colon\,}
-\renewcommand{\and}{\,\&\,}
-
-\begin{document}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
+\begin{figure}
+    \includegraphics{figs/linked-frequencies-longtime}
+    \caption{
+        \textbf{Probabilities of $A$ ancestry,}
+        across space (vertical axis, in units of $\sigma$) 
+        and time (horizontal axis, in generations).
+        In each plot, color corresponds to the expected frequency of $A$ ancestry
+        at a particular location in time and space.
+        The selection coefficient is $s=.02$.
+        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
+        on a time scale of $1/s=50$ generations.
+        \textbf{Bottom left:} at an unliked site,
+        with cline flattening continuing with $\sqrt{t}$.
+        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
+        on the ancestry at the selected site,
+        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
+        as described in the text (see definition of $q_z(x,t,r)$).
+        See figure \ref{sfig:linked_cline_heatmaps}
+        for the same figure over a shorter period of time.
+    }
+    \label{sfig:linked_cline_heatmaps_longtime}
+\end{figure}
 
@@
 
-\end{document}
+% \end{document}

### 2016-02-26 drafts/figs/theory-figs.R

@@ -1,0 +1,86 @@
+library(ReacTran)
+library(Matrix)
+library(RColorBrewer)
+source("../../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../../notes/cline-fns.R",chdir=TRUE)
+
+
+###
+# styles
+
+prop.pal <- colorRampPalette(c(rev(brewer.pal(n=6,name="Blues")),brewer.pal(n=6,name="Reds")))
+# prop.pal <- colorRampPalette(rev(brewer.pal(n=6,name="Blues")))
+
+#####
+## from pde-clines.Rmd
+
+## Time for the cline to get set up
+sigma <- 1
+xgrid <- setup.grid.1D(x.up=-floor(30*sigma), x.down=floor(30*sigma), N = 500)
+
+r <- 0.1
+s <- 0.02
+zone.age <- 150  # this is tau
+tt <- seq(0,zone.age,length.out=500)
+
+# selected frequency
+fwds.soln <- forwards_pde(s=s, times=tt, grid=extend_grid(xgrid), sigma=sigma)
+
+# linked frequencies
+rr <- c(0, 0.01, 0.02, 0.04, 0.08, 0.16, 0.32, 0.5)
+linked.solns <- lapply( rr, function (r) {
+            forwards_backwards_pde(s=s, times=tt, grid=xgrid, 
+                    r=r, sigma=sigma,
+                    log.p=TRUE, fwds.soln=fwds.soln)
+        } )
+linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
+
+.imgplot <- function (x,which=1,
+        xlab="generation", ylab="space", ...) {
+    xmat <- x[,1+(which-1)*attr(x,"dimens")+1:attr(x,"dimens")]
+    graphics::image( xmat, zlim=c(0,1), 
+            xaxt='n', yaxt='n',
+            ylim=c(0.3,0.7),
+            ylab=ylab, xlab=xlab,
+            col=prop.pal(100), 
+           ... )
+    graphics::contour( xmat, add=TRUE )
+    add_grid_axis(attr(x,"grid"))
+    add_grid_axis(side=1,x=x[,1])
+}
+
+# from deSolve:::drawlegend
+.drawlegend <- function ( zlim=c(0,1), colfn=prop.pal, horizontal=FALSE ) {
+    iy <- 1
+    minz <- zlim[1]
+    maxz <- zlim[2]
[…]

### 2016-02-26 notes/cline-fns.R

@@ -57,9 +57,8 @@ clineplot <- function (soln, thegrid, x=thegrid$x.mid,
     return( invisible( t(soln[times,-1]) ) )
 }
 
-add_grid_axis <- function (thegrid,side=2,...) {
-    axis(side=side,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),
-            labels=pretty(thegrid$x.mid))
+add_grid_axis <- function (thegrid,side=2,x=thegrid$x.mid,...) {
+    axis( side=side, at=(pretty(x)-min(x))/diff(range(x)), labels=pretty(x))
 }
 
 cline_interp <- function (t,soln) {

### 2016-02-27 drafts/figs/Plot_frequency_clines.R
various

@@ -1,0 +1,94 @@
+
+source("../../sims/sim-fns.R",chdir=TRUE)
+library(ReacTran)
+library(Matrix)
+source("../../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../../notes/cline-fns.R",chdir=TRUE)
+source("../../sims/chunks_fns.R",chdir=TRUE)
+
+
+###
+# Plot clines at selected and linked sites
+
+get_simparams <- function (fname) {
+    strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
+    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",fname)) } )
+}
+
+
+positions = seq(0.2,0.8,0.01)
+
+sel.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
+load(sel.sim.file)
+#CHUNKS_SEL = get.chunks.at.positions(positions,CHR=1)
+sel.params <- get_simparams( sel.sim.file )
+
+# get frequencies for selected simulation
+demeID = rep(1:sel.params$ndemes, each=sel.params$ninds/sel.params$ndemes)
+B_freq = get.ancestry.freqs(
+        IND_DATA=sims.sums[[1]]$ind.ancest, 
+        demeID=demeID, 
+        VECTOR_OF_POS=positions, 
+        CHR=1, 
+        ndemes=sel.params$ndemes)
+
+
+neutral.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"
+load(neutral.sim.file)
+#CHUNKS_NEU = get.chunks.at.positions(positions,CHR=2)
+neu.params <- get_simparams( neutral.sim.file )
+
+# get frequencies for neutral simulation
+demeID = rep(1:neu.params$ndemes, each=neu.params$ninds/neu.params$ndemes)
+B_neu = get.ancestry.freqs(
+        IND_DATA=sims.sums[[1]]$ind.ancest, 
+        demeID=demeID, 
+        VECTOR_OF_POS=positions, 
+        CHR=2, 
+        ndemes=neu.params$ndemes)
+
+
[…]

### 2016-02-27 drafts/figs/theory-figs.R

@@ -14,14 +14,15 @@ prop.pal <- colorRampPalette(c(rev(brewer.pal(n=6,name="Blues")),brewer.pal(n=6,
 #####
 ## from pde-clines.Rmd
 
+########
 ## Time for the cline to get set up
 sigma <- 1
-xgrid <- setup.grid.1D(x.up=-floor(30*sigma), x.down=floor(30*sigma), N = 500)
+xgrid <- setup.grid.1D(x.up=-floor(30*sigma), x.down=floor(30*sigma), N = 200)
 
 r <- 0.1
 s <- 0.02
 zone.age <- 150  # this is tau
-tt <- seq(0,zone.age,length.out=500)
+tt <- seq(0,zone.age,length.out=200)
 
 # selected frequency
 fwds.soln <- forwards_pde(s=s, times=tt, grid=extend_grid(xgrid), sigma=sigma)
@@ -43,6 +44,7 @@ linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
             ylim=c(0.3,0.7),
             ylab=ylab, xlab=xlab,
             col=prop.pal(100), 
+            useRaster=TRUE,
            ... )
     graphics::contour( xmat, add=TRUE )
     add_grid_axis(attr(x,"grid"))
@@ -71,16 +73,51 @@ linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
 }
 
 
-pdf( file="linked-frequencies.pdf", width=6.5, height= 4.2, pointsize=10 )
+pdf( file="linked-frequencies.pdf", width=6.5, height=4.2, pointsize=10 )
+# png( file="linked-frequencies.png", width=6.5*144, height=4.2*144, pointsize=10, res=144 )
 par(mar=c(3.5,3,2,0)+.1, mgp=c(2.3,1,0))
 layout(matrix(c(1:9,9),nrow=2),widths=c(rep(1,4),0.4))
 .imgplot( fwds.soln, main=expression(p(x,t)) )
 .imgplot( linked.clines[[match(0.5,rr)]], main=expression(q(x,t,r==0.5)) )
 for (k in c(2,4,8)) {
-    .imgplot( linked.solns[[k]], which=1, 
+    .imgplot( linked.solns[[k]], which=1, xlab='',
             main=as.expression(substitute( q[A](x,t,r==thisr), list(thisr=rr[k]))) )
     .imgplot( linked.solns[[k]], which=2,
[…]

### 2016-02-27 drafts/figs/Plot_frequency_clines.R
bug in cline_from_soln

@@ -19,33 +19,41 @@ get_simparams <- function (fname) {
 positions = seq(0.2,0.8,0.01)
 
 sel.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
-load(sel.sim.file)
-#CHUNKS_SEL = get.chunks.at.positions(positions,CHR=1)
 sel.params <- get_simparams( sel.sim.file )
 
-# get frequencies for selected simulation
-demeID = rep(1:sel.params$ndemes, each=sel.params$ninds/sel.params$ndemes)
-B_freq = get.ancestry.freqs(
-        IND_DATA=sims.sums[[1]]$ind.ancest, 
-        demeID=demeID, 
-        VECTOR_OF_POS=positions, 
-        CHR=1, 
-        ndemes=sel.params$ndemes)
-
-
 neutral.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"
-load(neutral.sim.file)
-#CHUNKS_NEU = get.chunks.at.positions(positions,CHR=2)
 neu.params <- get_simparams( neutral.sim.file )
 
-# get frequencies for neutral simulation
-demeID = rep(1:neu.params$ndemes, each=neu.params$ninds/neu.params$ndemes)
-B_neu = get.ancestry.freqs(
-        IND_DATA=sims.sums[[1]]$ind.ancest, 
-        demeID=demeID, 
-        VECTOR_OF_POS=positions, 
-        CHR=2, 
-        ndemes=neu.params$ndemes)
+if (!file.exists("plot_frequency_clines.Robj")) {
+    load(sel.sim.file)
+    #CHUNKS_SEL = get.chunks.at.positions(positions,CHR=1)
+
+    # get frequencies for selected simulation
+    demeID = rep(1:sel.params$ndemes, each=sel.params$ninds/sel.params$ndemes)
+    B_freq = get.ancestry.freqs(
+            IND_DATA=sims.sums[[1]]$ind.ancest, 
+            demeID=demeID, 
+            VECTOR_OF_POS=positions, 
+            CHR=1, 
+            ndemes=sel.params$ndemes)
+
+
+    load(neutral.sim.file)
+    #CHUNKS_NEU = get.chunks.at.positions(positions,CHR=2)
+
[…]

### 2016-02-27 drafts/figs/theory-figs.R

@@ -109,12 +109,12 @@ pdf( file="linked-frequencies-longtime.pdf", width=6.5, height=4.2, pointsize=10
 # png( file="linked-frequencies.png", width=6.5*144, height=4.2*144, pointsize=10, res=144 )
 par(mar=c(3.5,3,2,0)+.1, mgp=c(2.3,1,0))
 layout(matrix(c(1:9,9),nrow=2),widths=c(rep(1,4),0.4))
-.imgplot( fwds.soln, main=expression(p(x,t)) )
+.imgplot( fwds.soln, main=expression(p(x,t)), xlab='' )
 .imgplot( linked.clines[[match(0.5,rr)]], main=expression(q(x,t,r==0.5)) )
 for (k in c(2,4,8)) {
-    .imgplot( linked.solns[[k]], which=1, xlab='',
+    .imgplot( linked.solns[[k]], which=1, xlab='', ylab='',
             main=as.expression(substitute( q[A](x,t,r==thisr), list(thisr=rr[k]))) )
-    .imgplot( linked.solns[[k]], which=2,
+    .imgplot( linked.solns[[k]], which=2, ylab='',
             main=as.expression(substitute( q[B](x,t,r==thisr), list(thisr=rr[k]))) )
 }
 .drawlegend()

### 2016-02-27 notes/cline-fns.R

@@ -243,8 +243,9 @@ cline_from_soln <- function (soln, grid, clinefn=attr(soln,"u") ) {
     #  - and the next N numbers gives the value linked to Bs
     A.inds <- 1+(1:grid$N)
     B.inds <- 1+grid$N+(1:grid$N)
-    sel.cline <- outer( grid$x.mid, soln[,1], clinefn )
-    pde.cline <- cbind( soln[,1], soln[,A.inds]*sel.cline[col(soln[,A.inds])] + soln[,B.inds]*(1-sel.cline)[col(soln[,B.inds])] )
+    sel.cline <- t( outer( grid$x.mid, soln[,1], clinefn ) )
+    #pde.cline <- cbind( soln[,1], soln[,A.inds]*sel.cline[col(soln[,A.inds])] + soln[,B.inds]*(1-sel.cline)[col(soln[,B.inds])] )
+    pde.cline <- cbind( soln[,1], soln[,A.inds]*sel.cline + soln[,B.inds]*(1-sel.cline) )
     attributes(pde.cline) <- c( attributes(pde.cline), attributes(soln)[setdiff(names(attributes(soln)),c("dim","dimnames"))] )
     attr(pde.cline,"nspec")<-1
     return(pde.cline)

### 2016-02-27 drafts/hybrid_zone.tex
various
added supp fig on comparison of clines t

@@
 \url{https://github.com/asedghifar/hybridzones}.
+\plr{if I'm to be the corresp author, put this to \url{https://github.com/petrelharp/clinal-lineages}?}
 
@@
     across geography and at several physical positions on the genome,
-at $T=100$ generations after secondary contact. 
-\plr{what is $s$?}
-Each line represents a locus some distance $r$ away from the true target of selection, and $r=0$ represents the locus that is under selection.
-Grey lines represent the same positions from a simulation with
+    at $T=100$ generations after secondary contact,
+    and with $s=0.1$.
+    Each line represents a locus some distance $r$ away from the true target of selection, 
+    and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
     identical parameters except $s=0$.
+    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison}.
     \plr{Add same figure for some different time points and maybe values of $s$ to the supplement.} 

### 2016-02-27 drafts/supplement.tex
added supp fig on comparison of clines t

@@
 \begin{figure}
+    \includegraphics{figs/alleleFrequencies_sim_comparison}
+    \caption{
+        \textbf{Comparison of simulated to theoretical clines:}
+        As in Figure \ref{alleleFreq_tau100},
+        frequency of ancestry B,
+        across geography and at several physical positions on the genome,
+        at $T=100$ generations after secondary contact,
+        and at $s=0.1$.
+        Dotted lines are expected frequencies of $B$ ancestry $1-q(x,t,r)$
+        (unconditional on the ancestry of the linked allele), computed numerically.
+        Each line represents a locus some distance $r$ away from the true target of selection, 
+        and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
+        identical parameters except $s=0$.
+    }\label{sfig:alleleFreq_tau100_comparison}
+\end{figure}
+
+\begin{figure}
     \begin{center}

### 2016-02-27 drafts/figs/Plot_frequency_clines.R

@@ -89,8 +89,10 @@ fgrid <- extend_grid(xgrid)
 fwds.soln <- forwards_pde(s=theory.params$s, times=tt, grid=fgrid, sigma=theory.params$sigma )
 fwds.neutral.soln <- forwards_pde(s=0, times=tt, grid=fgrid, sigma=theory.params$sigma )
 
-compare.positions <- positions[ 1+(0:(floor(length(positions)/5)-1))*5 ]
-compare.positions <- compare.positions[compare.positions>=0.5]
+# compare.k <- 1+(0:(floor(length(positions)/5)-1))*5 
+# compare.k <- compare.k[ positions[compare.k]>=0.5 ]
+compare.k <- 31+c(0,1,2,3,5,8,13,21) 
+compare.positions <- positions[ compare.k ]
 compare.rvals <- compare.positions - 0.5
 
 linked.solns <- lapply( compare.rvals, function (r) {
@@ -98,62 +100,39 @@ linked.solns <- lapply( compare.rvals, function (r) {
                            sigma=theory.params$sigma, fwds.soln=fwds.soln, log.p=TRUE)
     } )
 linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
+tfreqs <- sapply( linked.clines, function (lnc) { lnc[nrow(lnc),-1] } )
 
 linked.neutral.solns <- lapply( compare.rvals, function (r) {
         forwards_backwards_pde(s=0, times=tt, grid=xgrid, r=r, 
                            sigma=theory.params$sigma, fwds.soln=fwds.neutral.soln, log.p=TRUE)
     } )
 linked.neutral.clines <- lapply( linked.neutral.solns, cline_from_soln, grid=xgrid )
+neutral.tfreqs <- sapply( linked.neutral.clines, function (lnc) { lnc[nrow(lnc),-1] } )
 
-pdf(file="alleleFrequencies_sim_comparison.pdf", width=6, height=4, pointsize=10)
+pdf(file="alleleFrequencies_sim_comparison.pdf", width=6, height=8, pointsize=10)
 layout((1:2))
 par(mar=c(3.5,3.5,0.5,0.5))
 
-# Neutral comparison
-# frequencies from theory
-tfreqs <- sapply( linked.neutral.clines, function (lnc) { lnc[nrow(lnc),-1] } )
[…]

### 2016-02-27 drafts/hybrid_zone.tex
removed old supp fig

@@
 we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically.
-For instance, Figure \ref{fig:pde_clines} shows a heatmap of $p(x,t,r)$, 
+For instance, Figure \ref{fig:linked_cline_heatmaps} shows a heatmap of $p(x,t,r)$, 
 the expected frequency of ancestry $A$ at location $x$ and time $t$ 

### 2016-02-27 drafts/hybrid_zone.tex
edits, mostly to discussion

--- 2016-02-26 drafts/hybrid_zone.tex
@@
 \usepackage{setspace}
+
+%\usepackage[hidelinks]{hyperref}
 
@@
 centered on their point of contact \citep{Barton1985}. 
-\plr{(check that neutral clines discussed in barton '85)}
+\plr{(check that neutral clines discussed in barton '85)}\alisa{This is simply stated in Barton1985 without any justification or citation. Can we just cite Sedghifar2015?}
 If the populations are sufficiently diverged,
@@
 are broken up by historical recombination events,
-leading to substantial correlations in ancestry between distant markers
-(termed ``ancestry LD'').
-\plr{(but, ``ancestry linkage disequilibrium'' doesn't work;
-how about just ``correlations in ancestry''?}
+leading to substantial correlations in ancestry between distant markers.
 
@@
 Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
+For this, we build on the neutral framework presented in \citet{Sedghifar2015} to derive the expected distribution of ancestry block lengths across a contact zone.
 \yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
 
-\plr{explain relationship to \citet{Sedghifar2015}}
+%\plr{explain relationship to \citet{Sedghifar2015}}
 
@@
 In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes. 
-\alisa{I don't think I understand the independence from $\sigma$?}
-\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}
+%\alisa{I don't think I understand the independence from $\sigma$?}
+%\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}
 
@@
 back towards the side where they are more common. 
-\alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
-\plr{This isn't a construction, it's a consequence of the model, so you're right, it is maybe a "result", but I think the explanation of why it happens belongs here.}
 Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
@@
 
-\begin{figure}
-    \includegraphics{figs/linked-frequencies}
-    \caption{
-        \textbf{Probabilities of $A$ ancestry,}
-        across space (vertical axis, in units of $\sigma$) 
-        and time (horizontal axis, in generations).
-        In each plot, color corresponds to the expected frequency of $A$ ancestry
-        at a particular location in time and space.
-        The selection coefficient is $s=.02$.
-        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
-        on a time scale of $1/s=50$ generations.
-        \textbf{Bottom left:} at an unliked site,
-        with cline flattening continuing with $\sqrt{t}$.
-        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
-        on the ancestry at the selected site,
-        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
-        as described in the text (see definition of $q_z(x,t,r)$).
-        See figure \ref{sfig:linked_cline_heatmaps_longtime}
-        for the same figure over a longer period of time.
-    }
-    \label{fig:linked_cline_heatmaps}
-\end{figure}
-
-\paragraph{Linked clines}
+\paragraph{PDE for linked clines}
 We can use this diffusion model for lineages to compute expected clines in ancestry,
@@
 and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
-\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
-\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
-applied to a backwards-time process!}
+%\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
+%\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
+%applied to a backwards-time process!}
 
@@
 From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$).
-These were compared to the output of neutral simulations with $s=0$
-\plr{missing end of sentence?}
-
-The simulations are executed in R, with scripts available on
+These were compared to the output of neutral simulations with $s=0$.
+
+The simulations were executed in R, with scripts available on
 \url{https://github.com/asedghifar/hybridzones}.
@@
 
-Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
+%Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
 
 Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$, that flank the block of unbroken ancestry containing $m$.
+Since there appears to be a signal at these regions also, we further define a statistic
+
+$$C =  \frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$$
+
+which is the ratio of mean block length around $k_i$ and mean length of blocks directly flanking the block surrounding $k_i$.
 
@@
 
+\begin{figure}
+\includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
+\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
+Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.
+    \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
+}\label{Fig:blockLengths}
+\end{figure}
+
 \subsection*{The number of `unfit' ancestors}
@@
 
-\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-    which we have done using both theory and simulation.
-    Our most useful observations are that clines establish over time $1/s$,
-    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-    tend to move back towards the side where they are common
-    at speed $\sigma/\sqrt{s}$.
-    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-}
-
-Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
-Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.
-\alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g.
+%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+%   which we have done using both theory and simulation.
+%   Our most useful observations are that clines establish over time $1/s$,
+%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+%    tend to move back towards the side where they are common
+%   at speed $\sigma/\sqrt{s}$.
+%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+%}
+
+\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
+We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
+This occurs at speed $\sigma/\sqrt{s}$.
+Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
+This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
+
+\subsection*{Genomic signals associated with targets of selection}
+
+There is great interest in identifying the loci that are experiencing selection in hybrid zones.
+Popular approaches to this involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
+Current genomic resources, however, can provide an additional rich information when applied to the study of hybrid zones.
+In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture from present day samples \citep[e.g.
 ][]{Hellenthal2014}.
-\alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention.
-Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}
+In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone.
+Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
+This pattern reflects the intuition that regions surrounding selected loci introgress less readily, and loci under selection can therefore be thought as moving under Brownian motion with drift whereby a lineage that is presently on the `` wrong" side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct" side.
+In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong" side, are likely to have done so relatively recently.
+Haplotypes surrounding the selected locus therefore have had relatively few opportunities for recombination with different ancestries and this is reflected in longer blocks of contiguous ancestry.
+
+Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short.
+This is measured by the length of the block adjacent to the block containing the selected site.
+A simple and intuitive explanation for this patter in that loci physically linked to a selected site can only survive on the wrong side of the hybrid zone if they recombine off the background of the unfit variant.
+A disproportionate majority of the subsequent recombination events it experiences will be with the more fit ancestry, and so the ancestry block will decay rapidly. 
+This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.
+
+Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
+Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary.
+This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
+
+While these statistics presented here are currently detectable only in the few systems for which dense genetic markers and phase information is known, as the cost of sequencing continues to drop and long-phased reads become more common, the patterns described here could further aid in the identification of selected sites in hybrid zones.
+In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available low-density and unphased genomic data, could offer a less-powerful alternative to these findings.
 
@@
 
-\plr{(selected lineages drift back, so haplotypes are longer)}
-Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing.
-Thus we can interpret this pattern as Brownian motion with drift in which an ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct'' side of the hybrid zone.
-In other words, selected alleles that have, by chance, encroached deep into the other side of the hybrid zone have likely done so quite recently, as they cannot survive long in such hostile territory.
-\alisa{As a result, the haplotype that the selected locus resides on has few opportunities for recombination on the `wrong' side, which is reflected in long blocks of contiguous ancestry.}
-%The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
+\alisa{These seem like an orphan statement. Any idea where to stick these?}
+We further find that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will be descended from a relatively large number of independent ancestors, per individual, of the less fit ancestry.
+This result reflects that, on average, less fit haplotypes tend to be recent migrants with few descendants.
+
+\subsection*{Assumptions}
+
+We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations with respect to the architecture of selection, the geographic distribution of populations, and relatedness of individuals.
+
+Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}).
+We expect, therefore, our findings here to be somewhat generalizable.
+
+Our assumption of a single selected locus may have greater consequences.
+While our model is likely to be relevant to cases in which there are only a small number of targets of selection scattered throughout the genome, our predictions may differ significantly from situations in which the density of selected sites is higher.
+Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination.
+Both factors are expected to result in regions that are surrounded by even greater chunks of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
+ $s$ and relatively short map distances between linked incompatibilities)
+
+%Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
+This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
+
 \plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
@@
 
-\subsection*{Additional signatures around selected loci}
-
-\plr{(selection increases drift and migrant families are small; need better words/description here.)}
-We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a relatively large number of independent ancestors per individual of that ancestry.
-(Figure~\ref{Fig:family_size}).
-% is this right
-This result also reflects the fact that, on average, haplotypes on the ``wrong'' side of the hybrid zone tend to be recent migrants and have therefore had few descendants.
-
-\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
-In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
+%\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
+%In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
 We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
@@
 \yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}
-
-\plr{(we'll be able to see this soon)}
-These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown.
-However, as the const of sequencing continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.
-%\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
-\alisa{In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available data, surrounding selected loci could offer an approximation to these findings.}
-
-\subsection*{Paths to inference}
-
-\plr{(combine with haplotype length stuff above)}
-Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
-Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary.
-This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
-
-\subsection*{Assumptions}
-
-\plr{(het. disadvantage)}
-We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations.
-%Thus before our results can be profitably used for inference, it is worthwhile to  reflect on our assumptions and their consequences.
-Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
-This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
-\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
-
-\plr{(one locus only)}
-The assumption about the density of selected loci, however, may have greater consequences.
-Our model, which focuses on a single selected site is likely to be still relevant to cases in which there are only a small number of selected loci scattered throughout the genome.
-However, our predictions may differ significantly from scenarios in which the density of selected sites is high.
-Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \citep{Barton1983,Kruuk1999}.
-Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting  greater $s$ and relatively short map distance between linked incompatibilities -- \alisa{and will impact expectations of genome-wide patterns of clines and block lengths.
-In terms of detecting targets of selection, such sites will be a common limitation among many approaches to identifying selection.}
 

### 2016-02-27 drafts/supplement.tex

@@
 \begin{figure}
-    \begin{center}
-       \includegraphics[width=0.7\textwidth]{figs/example_cline}
-    \end{center}
-    \caption{
-        \textbf{Numerically computed clines,} at
-        \textbf{(A)} the selected site, which forms over time $1/s$ and then remains stable, and
-        \textbf{(B)} a linked site at distance $r=0.01$, which continues to smooth out as $\sqrt{t}$.
-        \textbf{(C)} a linked site at $r=0.1$
-        \textbf{(D)} an unlinked site.
-        Other parameters.
-       \plr{Do we want this figure? This one is just a placeholder; I'd put in two, that showed the formation and spread better.
-       This is from the PDE (we don't have time courses from sims).}
-        \label{fig:pde_clines}
-    }
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=\textwidth]{figs/plot_chromosomes_tau100}

### 2016-02-27 sims/compare-to-theory.Rmd

@@
     strings <- c( sigma="SIGMA", ninds="Ninds", ndemes="ndemes", s="s", tau="tau" )
-    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",simchunks.file)) } )
+    lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",fname)) } )
 }

### 2016-02-28 drafts/hybrid_zone.tex
more comparison figs

--- 2016-02-27 drafts/hybrid_zone.tex
@@
 }
+
+\plr{say what population density in simulation figures}
 
@@
 
-\section*{Single-locus behavior}
+\subsection*{Comparison of theory and simulation}
+
+The theoretical results described above matched well with our individual-based simulations,
+across a large range of parameter values.
+This can be seen by comparing
+clines in frequency of $A$ ancestry at linked loci,
+either unconditional on the identity of the linked selected allele (Figure \ref{sfig:alleleFreq_tau100_comparison}),
+or conditional on its identity 
+(Figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}).
+
+\subsection*{Single-locus behavior}
 
@@
 
+\plr{note that when drift starts to kick in for clines (around $\tau=320$) we start to get better power with haplotype lengths}
+
 \plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
@@
 matches the simulations.
-These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci.
+These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci:
+about 1 second to compute predicted clines, as opposed to hours for the full simulation.
 However, due to the high dimensionality of the haplotype problem (spatial position $\times$ time $\times$ endpoints of the haplotype),

### 2016-02-28 sims/sim-fns.R

@@ -1,6 +1,6 @@
 
 getRuns = function(index){
-    #This returns an array in which rows are the the beggingns and end of a TRUE/FALSE vector
+    #This returns an array in which rows are the the beginnings and end of a TRUE/FALSE vector
     # (note: more efficient to use rle() probably)
     starts  = which(c(0,index)[-1]==1 & c(0,index)[-length(index)]!=1)
     stops   = which(c(index,0)[-1]!=1 & c(index,0)[-(1+length(index))]==1)
@@ -235,9 +235,8 @@ trueSigma <- function (sigma,n=1e4) {
 
 
 ###########################
-#To get distribution of chunks of ancestry B to the (wlog) right of a locus, for instance:
-# testB_far = lapply(1:params$ndemes,function(X){getDemeChunks(DEME=X,ancA=FALSE,POS=0.01)})
-
+# Getting lengths of chunks of ancestry
+###
 
 getChunk = function( IND_DATA, CHR, POS ) {
 	# return the length and identity of the interval containing the focal site for a (haploid) individual	

### 2016-02-28 drafts/figs/theory-hap-lens.R
code for hap lengths

@@ -1,0 +1,30 @@
+# Get fine-scale haplotype length predictions.
+
+source("../../sims/sim-fns.R",chdir=TRUE)
+library(ReacTran)
+library(Matrix)
+source("../../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../../notes/cline-fns.R",chdir=TRUE)
+source("../../sims/chunks_fns.R",chdir=TRUE)
+
+theory.params <- list( 
+              sigma=1,
+              s=0.01,
+              tau=1000,
+              density=100,
+              width=50 )
+
+tt <- seq(0,theory.params$tau,length.out=21)
+xgrid <- setup.grid.1D(x.up=-theory.params$width/2, x.down=theory.params$width/2, N=50)
+fgrid <- extend_grid(xgrid)
+fwds.soln <- forwards_pde(s=theory.params$s, times=tt, grid=fgrid, sigma=theory.params$sigma )
+
+# do it on a fine scale: for tau=1000, looks to be over +/- 5 or 6 cM
+rgrid <- setup.grid.1D(x.up=-0.05, x.down=0.05, N=40)
+hap.soln <- forwards_backwards_haplotypes(s=theory.params$s, times=tt, xgrid=xgrid, rgrid=rgrid,
+                           sigma=theory.params$sigma, 
+                           fwds.grid=fgrid, fwds.soln=fwds.soln )
+
+dir.create("cache",showWarnings=FALSE)
+save(list=ls(), file=sprintf("cache/haplotypes_SIGMA%d_density%d_ndemes%d_s%0.3f_tau%d.Robj", 
+                             theory.params$sigma, theory.params$density, theory.params$width, theory.params$s, theory.params$tau ) )

### 2016-02-28 drafts/hybrid_zone.tex
added section on numerical soln of pde

@@
             \\ {} & \qquad 
-            + (1-p) \int_a^0 \left(
+            + (1-p) \left( \int_a^0 
                 g_B(a,\theta) g_A(\theta,b)
-                + g_A(a,\theta) g_B(\theta,b)
-            \right) d\theta,
+                + \int_0^b g_A(a,\theta) g_B(\theta,b)
+            d\theta \right)
             \\ {} & \qquad 
@@
 
-Numerical details.
+In this section, we give details for how we found numerical solutions
+to the partial differential equations (PDE) of the text,
+which are all of reaction-diffusion type.
+The R code, with worked examples, is available in our git repository.
+Spatial grids were usually chosen to have at least four grid sites per dispersal distance,
+but using substantially finer grids did not affect the results.
+
+The forwards-time evolution of the selected alleles,
+equation \eqref{eqn:cline_pde}, presents no difficulty;
+we use the ReacTran package \citep{soetaert2012reactive} to compute discrete approximations to the diffusion term,
+and the deSolve package \citep{soetaert2010solving} to solve the equation.
+
+The equations \eqref{eqn:q_pde} describing probabilities that a lineage descends from $A$ ancestry,
+conditional on the linked selected allele, required some more attention.
+First, $s>0.1$ the system of equations can be \emph{stiff} 
+(as is commonly observed for reaction-diffusion equations),
+and hence slow to solve,
+because of the extremely steep slope of the selected allele frequency $p(x,t)$.
+In practice we used $s<0.1$.
+Second, the ReacTran function tran.1D that converts the diffusion portion of the PDE into a system of ODE
+contains a term like 
+\begin{align*}
+    \frac{1}{A(x)} \frac{d}{dx} A(x) f(x) = \frac{d}{dx} f(x) + f(x) \frac{d}{dx} \log A(x) ,
+\end{align*}
+where $A(x)$ is the interface area between grid cells, and $f(x)$ is the flux.
+(See \citet{sotaert2012reactive} for more details.)
+Discrete approximations of the left-hand side, as implemented in ReacTran, 
+run into numerical difficulties if $A(x)$ is small,
+and in our case, $A(x)$ is equal to $p$, the local frequency of the selected $A$ allele.
+To avoid this, we made minor modifications to the tran.1D to provide a discrete approximation to the right-hand side.
+
+Haplotype probabilities are obtained from equations \eqref{eqn:g_pde},
+which is a coupled system of integro-differential equations in three variables plus time.
+One method for solution would be via a Wild sum over the number of recombination events, 
+as we did in \citet{Sedghifar2015}.
+Here, we solved the equations numerically,
+again discretizing the equations and using the ode.1D function of the deSolve package.
+The reaction-diffusion part is the same as for equations \eqref{eqn:q_pde}.
+The functions $g_A(x,t;a,b)$ and $g_B(x,t;a,b)$ are functions of space ($x$), time ($t$), 
+and the endpoints of the block in question ($a$ and $b$, with $a<b$).
+The second integral in \eqref{eqn:g_pde} is
+\begin{align} \label{eqn:unifprod}
+    \int_a^b { g_A(a,\theta) g_A(\theta,b) } d\theta .
+\end{align}
+Conceptually, this is an integral transformation
+$g(a,b) \mapsto \int_a^b g(a,\theta) g(\theta,b) d\theta$;
+the reason it appears here is that for the entire segment $(a,b)$ to be of ancestry $A$,
+if there was a recombination at $\theta$,
+the two segments $(a,\theta)$ and $(\theta,b)$ must both be of ancestry $A$
+(and, ignoring coalescence, these probabilities are independent).
+Suppose we have divided the segment of chromosome into a regular grid, say, $r_1 < \ldots < r_n$.
+The natural discretization approximates this transformation by a sum,
+and is equivalent to keeping track of only a finite number of loci.
+Writing $g_A(i,j)$ for $g_A(x,t;r_i,r_j)$,
+the discrete transformation we use corresponding to \eqref{eqn:unifprod}
+is $g_A(i,j) \mapsto {}$ the sum
+\begin{align*}
+    \sum_{k=i}^{j-1} (r_{k+1}-r_k) g_A(i,k) g_A(k+1,j) .
+\end{align*}
+This is the correct term for the process only tracking loci on the grid, because
+for all the alleles at $r_i, r_{i+1}, \ldots, r_j$ to be of ancestry $A$,
+when a recombination occurrs between $r_k$ and $r_{k+1}$,
+the sequences of alleles at $r_i, \ldots, r_{k}$ and at $r_{k+1}, \ldots, r_j$ must each be of ancestry $A$.
+The first integral term in \eqref{eqn:g_pde} is
+%       p * unifprod_ut(y=yA,ncx=nabvals)
+%       + (1-p) * biprod_ut(yA=yA,yB=yB,zeroind=zeroind,ncx=nabvals)
+\begin{align*} \label{eqn:biprod}
+    \int_a^0 g_B(a,\theta) g_A(\theta,b) d\theta
+    + \int_0^b g_A(a,\theta) g_B(\theta,b) d\theta 
+\end{align*}
+We now require that one of the grid points along the chromosome is exactly at the selected site;
+say this is $r_\ell = 0$.
+The discrete term corresponding to \eqref{eqn:biprod} is
+\begin{align*}
+    \sum_{k=i}^{\ell-1} (r_{k+1}-r_k) g_B(i,k) g_A(k+1,j) +
+    \sum_{k=\ell}^{j-1} (r_{k+1}-r_k) g_A(i,k) g_B(k+1,j) .
+\end{align*}
+Since these are not easily vectorizable in R, for efficiency
+we implemented the discrete transformations in C, using the Rcpp package \citep{Eddelbuettel2011}.
+In implementing these,
+we kept track of $g_A(i,j)$ in a vector in the order that the upper triangular elements of a matrix are encountered
+when traversing the matrix column-wise,
+allowing for efficient computation of the sum.
+
+\paragraph{A note on symmetry:}
+The equations we present have the symmetry that they are invariant after exchanging $A$ and $B$ and reversing space.
+For instance, $q_A(x,t,r) = 1 - q_B(-x,t,r)$.
+Using this fact would speed up the code by a factor of two,
+at the cost of generality:
+as written, it would be easy to modify the code to allow space to be inhomogeneous 
+(which would break this symmetry).
 

### 2016-02-28 drafts/hybrid_zone.tex
edits

--- 2016-02-27 drafts/hybrid_zone.tex
@@
 
-\alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}
+\alisa{This process of a haplotype sampled in the present day tracing it's path $X_\tau$ backwards in time, while splitting and changing backgrounds $Z_\tau$ through recombination, is shown in Figure~\ref{Fig:schematic}.}
 \plr{Connect these up more.}
@@
 The simulations were executed in R, with scripts available on
-\url{https://github.com/asedghifar/hybridzones}.
-\plr{if I'm to be the corresp author, put this to \url{https://github.com/petrelharp/clinal-lineages}?}
+\url{https://github.com/petrelharp/clinal-lineages}
 
@@
 
-With the above theory and simulations, we are able to make predictions about patterns surrounding selected loci.
+The above theory and simulations allow us to make predictions about patterns surrounding selected loci.
 In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection.
-The most straightforward measure is $l_I(m)$ the length of haplotype chunk that a sampled chunk containing genomic position $b$, conditioned on ancestry $I$ at the site.
-Based on our expectation of the distribution of block lengths, we compute $l_B(m)$ for $m$ across a chromosome.
-
+The most straightforward measure is $l_I(m)$,  the mean length of all haplotype chunks that contain genomic position $m$, given that $m$ is of ancestry $I$, where $I\in\{A,B\}$.
+We then focus on $I_B(m)$ for $m$ across a chromosome. 
 %Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
 
-Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$, that flank the block of unbroken ancestry containing $m$.
+Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$ that flank, to the left and right respectively, the block of unbroken ancestry containing $m$.
 Since there appears to be a signal at these regions also, we further define a statistic
 
-$$C =  \frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$$
-
-which is the ratio of mean block length around $k_i$ and mean length of blocks directly flanking the block surrounding $k_i$.
+$$C(m) =  \frac{2\sum_i{l_B(m_i)}}{\sum_i{l_A(m_i+)+l_A(m_i-)}}$$
+
+which is the ratio of mean block length around $m_i$ and mean length of blocks directly flanking the block surrounding $m_i$ (i.e. adjacent blocks).
+
+We are interested in identifying regions of the genome with abnormal distributions of block lengths, and where indicated normalize $l_I(m)$ and $C(m)$ by dividing respectively by mean $l_I$ and $C$  within a deme.
+While there is useful information contained in the geographic distribution of block length patterns, obtaining good spatial sampling can often be difficult, and this allows us to predict patterns within relatively few populations.
+
+Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
+In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
+This is a reflection of the average size of a haplotype family, and is potentially a source of additional information.
+Within each deme $X$, we calculated $F_B(m,X)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
 
@@
 than the average for that geographic location.
-\plr{Write down in the section on statistics what exactly you normalize by:
-is it the mean for that geographic location?  Or, what?}
+
 % This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
@@
 but it is much smaller (see Figures \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}).
+
+Additionally, we find that $l_A(m\pm)$, the length of blocks adjacent to blocks containing $m$, are shorter than average when $m$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
+(Figure~\ref{Supp:adjacentBlocks}).
+This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by measuring $C(m)$  (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
 
@@
 
-\subsection*{Lengths of adjacent blocks}
-
-\plr{Define this statistic and put in the Summary Stats section. Also, move this up with the haplotype length stuff?}
-We find that $l_A(k\pm)$, the length of blocks adjacent to blocks containing $k$, are shorter than average when $k$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
-(Figure~\ref{Supp:adjacentBlocks}).
-This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by taking the ratio $\frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
-
 %This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
@@
 \includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
-Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.
+\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
+Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m)$ along the chromosome in the deme.
     \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
@@
 \plr{This needs a bit more precise discussion.}
-
-Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
-In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
-This is a reflection of the average size of a haplotype family.
-We find that, as distance from the zone center increases, the average family size of an ancestor from time $T$ decreases (Figure~\ref{Fig:family_size}).
-One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection.
-This can potentially provide additional signal for characterizing lineages strongly influenced by selection.
+As expected, we find that $F_B(m)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate.
+We find that at the selected locus, $F_B(m,X)$ decreases with X, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
+As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
+
+One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families.
+In conjunction with block length, measuring $F$ can provide an additional signal for characterizing lineages strongly influenced by selection, by looking at levels of haplotype diversity in populations along a hybrid zone.
 
@@
 
+\paragraph{The nature of selection}
 Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}).
@@
 
+\paragraph{Geographic space and other sources of stochasticity}
+For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape, allowing us to derive analytic expressions that track lineages.
+This assumes a large population density limit that excludes coalescence and makes lineages Markov.  
+In contrast, our simulations model regularly spaced demes. 
+In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur. 
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
+
+Of particular concern that stems from the above assumption is that our model ignores potential, and likely, sources of stochasticity arising from geographic structure and coalescence/pedigree structure.
+The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \cite{Liang2014}.
+With respect to genealogy, the broad agreement between our analytic predictions and simulations (which include an explicit pedigree structure), suggests that this assumption does not have a severe qualitative impact on our major results when populations are moderately high, especially for all but the most recently formed hybrid zones.
+
+Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
+This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.
+
+%, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}. 
+%\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
+%However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
+Is there a better way to state this for our intended audience?}.
+%\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
+%In contrast, our simulations model regularly spaced demes.
+In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
+
+%Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
+%The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
+
 %Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
@@
 \alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
-
-\plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
-\yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
-This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.}
 
@@
 
-\plr{(1D)}
-For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape.
-This allows us to derive analytic expressions that track lineages, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}.
-\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
-However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
-Is there a better way to state this for our intended audience?}.
-\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
-In contrast, our simulations model regularly spaced demes.
-In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
-While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
-
-Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
-\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
-The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
-
 %One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
@@
 
-\subsection*{Islands of divergence}
-
-\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
-The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
-REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
-We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
-% in terms of the processes of migration, and selection influencing patterns of ancestry across the genome of hybridizing populations.
+\subsection*{Patterns of divergence}  
+A number of studies have described heterogeneous patterns of genetic divergence across the genome.
+These so-called ``islands" of divergence \citep{Turner2005,Nosil2009} have been largely descriptive \citep{Cruickshank2014,Noor2009}, and as we begin to rigorously model patterns of divergence in the genome to begin to more  explore these patterns.
+The work presented here may offer ways in which to think about the process of heterogeneous divergence more concretely in terns of how processes such as migration and selection influence patterns across the genome of hybridizing populations.
 Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.
 Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection rather than high-level summaries that are somewhat abstracted from the evolutionary process.
+
+%\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
+%The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
+%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+%REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
+%We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
+% in terms of the processes of migration, and selection influencing patterns of ancestry across the genome of hybridizing populations.
 
@@
 
-\subsection*{Simulation and analysis code}
-
-These are written primarily in R, and are available HERE.
+%\subsection*{Simulation and analysis code}
+%These are written primarily in R, and are available HERE.
 

### 2016-02-28 drafts/hybrid_zone.tex

@@
 
+\plr{note that when drift starts to kick in for clines (around $\tau=320$) we start to get better power with haplotype lengths}
+
+\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+    which we have done using both theory and simulation.
+    Our most useful observations are that clines establish over time $1/s$,
+    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+    tend to move back towards the side where they are common
+    at speed $\sigma/\sqrt{s}$.
+    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+}
+
 %However, how selection influences the length of tracts of ancestry around selected sites in hybrid zones has received much less attention.

### 2016-02-28 drafts/supplement.tex

@@
 
+\begin{figure}
+    \includegraphics{figs/cond_freqs_comparison_tau_80}
+    \caption{
+        \textbf{Conditional frequencies of ancestry $A$ at $\tau=80$},
+        comparing simulation and theory, with $\sigma=3$ deme spacings, $s=0.05$, and 50 individuals per deme.
+        The top left figure shows observed and expected genotype frequencies for the two homozygotes and the heterozygote at the selected locus;
+        expected genotype counts were obtained assuming random mating,
+        and by solving equation \eqref{eqn:cline_pde} numerically.
+        The remaining figures show observed and expected frequencies of $A$ ancestry,
+        separately conditioned on the identity of the linked allele at the selected site.
+        Observed frequencies become much noisier where the linked allele becomes rarer.
+    } \label{sfig:condAlleleFreq_tau80_comparison}
+\end{figure}
+
+\begin{figure}
+    \includegraphics{figs/cond_freqs_comparison_tau_320}
+    \caption{
+        \textbf{Conditional frequencies of ancestry $A$ at $\tau=320$},
+        as in figure \ref{sfig:condAlleleFreq_tau80_comparison}.
+        Deviations are larger than at $\tau=80$, 
+        due to genetic drift.
+    } \label{sfig:condAlleleFreq_tau320_comparison}
+\end{figure}
+
+\begin{figure}
+    \includegraphics{figs/cond_freqs_comparison_tau_1280}
+    \caption{
+        \textbf{Conditional frequencies of ancestry $A$ at $\tau=1280$},
+        as in figure \ref{sfig:condAlleleFreq_tau80_comparison}.
+        Deviations are larger still than at $\tau=320$, 
+        due to genetic drift.
+    } \label{sfig:condAlleleFreq_tau1280_comparison}
+\end{figure}
+
 % \end{document}

### 2016-02-28 drafts/hybrid_zone.tex
edits through to end of discussion

@@
 
-%, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}. 
-%\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
-%However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
-Is there a better way to state this for our intended audience?}.
-%\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
-%In contrast, our simulations model regularly spaced demes.
-In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
-While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
-
-%Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
-\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
-%The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
-
-%Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
-This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
-\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
-
-%\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
-%In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
-We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
-Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
-A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
-Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly.
-\yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}
-
-%One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
-%Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
-%We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
-%An additional assumption of our analytical model is that we ignore coalescence.  
-%The broad agreement between our analytical and theoretical results suggests that this assumption will also have negligible effect on our results.
-
-%Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
-
-%\alisa{epistasis}
-%\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
-
-\subsection*{Patterns of divergence}  
-A number of studies have described heterogeneous patterns of genetic divergence across the genome.
-These so-called ``islands" of divergence \citep{Turner2005,Nosil2009} have been largely descriptive \citep{Cruickshank2014,Noor2009}, and as we begin to rigorously model patterns of divergence in the genome to begin to more  explore these patterns.
-The work presented here may offer ways in which to think about the process of heterogeneous divergence more concretely in terns of how processes such as migration and selection influence patterns across the genome of hybridizing populations.
-Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.
-Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection rather than high-level summaries that are somewhat abstracted from the evolutionary process.
-
-%\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
-%The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
-%REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
-%We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
-% in terms of the processes of migration, and selection influencing patterns of ancestry across the genome of hybridizing populations.
-
-\subsection*{Future directions}
-
-\plr{Perhaps delete this paragraph? A small note saying should check power with simulations up where we discuss the statistic would do.}
-Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
-We suggest that such signatures could be combined in an inference framework that aims to identify incompatibilities from genome scans of hybrid zones.  
-However, doing so requires careful evaluation of potential signals of an incompatibility. 
-For example, should we consider jointly lengths of ancestry and allele frequencies to identify incompatibilities?
-Will dividing haplotype length by the length of adjacent haplotypes add power to tests because of the extra information contained in this measure, or add noise because of chance variation in these statistics.
-Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
-We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $T$
-
-\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
-Throughout, we have focussed on hybrid incompatibilities, however, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.   
-The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin.
-\citep{Song2011}.
-Future work could dress expected genomic signatures of adaptive introgression in hybridizing populations using the number and length of haplotypes of non-native ancestry. 
-Doing so requires an understanding of how positive selection shapes ancestry block lengths in these cases.
-
 \subsection*{Theory and simulation}
@@
 for instance, across maps of real landscapes.
+
+%, and is an example of the more general idea of duality for particle systems \alisa{I think this may be beyond the average ME reader?}. 
+%\plr{Tracking lineages is duality, has nothing to do with 1D, though.}
+%However, this assumes a large population density limit that excludes \plr{coalescence} and makes linages Markov \alisa{not sure about what point this is making?
+Is there a better way to state this for our intended audience?}.
+%\plr{It means we're ignoring drift/coalescence, an important source of stochasticity; but that's the next paragraph too}
+%In contrast, our simulations model regularly spaced demes.
+In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur.
+While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
+
+%Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
+%The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
+
+%Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus.
+This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met.
+\alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}
+
+%\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
+%In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding.
+We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}).
+Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short.
+A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site.
+Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly.
+\yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}
+
+%One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
+%Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
+%We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
+%An additional assumption of our analytical model is that we ignore coalescence.  
+%The broad agreement between our analytical and theoretical results suggests that this assumption will also have negligible effect on our results.
+
+%Also, parent-offspring dispersal is not the same thing as successful-offspring-parent dispersal.
+
+%\alisa{epistasis}
+%\alisa{selection here is STRONG, not sure how well this will work on systems in which selection on any given locus is small (most systems)}
+
+\subsection*{Patterns of divergence}  
+A number of studies have described heterogeneous patterns of genetic divergence across the genome.
+These so-called ``islands" of divergence \citep{Turner2005,Nosil2009} have been largely descriptive \citep{Cruickshank2014,Noor2009}, and as we begin to rigorously model patterns of divergence in the genome to begin to more  explore these patterns.
+The work presented here may offer ways in which to think about the process of heterogeneous divergence more concretely in terns of how processes such as migration and selection influence patterns across the genome of hybridizing populations.
+Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.
+Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection rather than high-level summaries that are somewhat abstracted from the evolutionary process.
+
+%\plr{seems like a nice point but should be more positive and less snarky.  "Thanks to studies like these, it is time to start explicit modeling, as we have done here..."}
+%The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
+%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+%REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
+%We hope that our work provides a path towards thinking about the process of heterogeneous divergence more concretely.
+% in terms of the processes of migration, and selection influencing patterns of ancestry across the genome of hybridizing populations.
+
+%\subsection*{Future directions}
+
+%\plr{Perhaps delete this paragraph? A small note saying should check power with simulations up where we discuss the statistic would do.}
+%Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
+%We suggest that such signatures could be combined in an inference framework that aims to identify incompatibilities from genome scans of hybrid zones.  
+%However, doing so requires careful evaluation of potential signals of an incompatibility. 
+%For example, should we consider jointly lengths of ancestry and allele frequencies to identify incompatibilities?
+%Will dividing haplotype length by the length of adjacent haplotypes add power to tests because of the extra information contained in this measure, or add noise because of chance variation in these statistics.
+%Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
+We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $T$
+
+\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
+\subsection*{The hybrid filter}
+\alisa{Could not find a reference to the hybrid filter}
+
+We have demonstrated here the potential utility for block length distributions in hybrid zone populations, particularly in populations that are found some distance from the center of the zone.
+While the focus here, has been on hybrid incompatibilities, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.
+Future work could take a similar approach to understand how positive selection shapes ancestry block lengths, and predict  signatures of adaptive introgression in hybridizing populations using similar statistics presented here.
+These could eventually be combined to gain a fuller understanding of the forces shaping patterns of introgression in hybrid zones.
+
+%The best characterized example of adaptive introgression comes from the vitamin K epoxide reductase subcomponent 1 (vkorc1), which seems to have spread from Algerian mice \emph{Mus spretus} to house mice \emph{Mus musculus} from because it confers resistance to the anticoagulant rodenticide, warfarin.
+\citep{Song2011}.
+%Future work could dress expected genomic signatures of adaptive introgression in hybridizing populations using the number and length of haplotypes of non-native ancestry. 
+%Doing so requires an understanding of how positive selection shapes ancestry block lengths in these cases.
 

### 2016-02-28 drafts/hybrid_zone.tex
added more plots to run_sims.R

--- 2016-02-28 drafts/hybrid_zone.tex
@@
 where $A(x)$ is the interface area between grid cells, and $f(x)$ is the flux.
-(See \citet{sotaert2012reactive} for more details.)
+(See \citet{soetaert2012reactive} for more details.)
 Discrete approximations of the left-hand side, as implemented in ReacTran, 

### 2016-02-28 sims/run_sims.R

@@ -202,3 +202,199 @@ LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:param
 pdf(file = paste(outfile,"_LD.pdf",sep=""))
 matplot(xx,LD_matrix,type="l",col=rainbow(20),xlim=c(-10,10),lty=1)
 dev.off()
+
+#########
+#########
+#########
+## stuff from potential_statistics: haplotype lengths, etcetera.
+
+chromosome = "chr1"
+plot_start = 0.48
+plot_stop=0.52
+increments = 0.001
+
+positions = seq(plot_start,plot_stop,increments)
+
+outstring = with( list2env(params), sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_dir/tau%s_dir/results_runid_%s_chr%s_start%g_stop%g_by%g",
+                                            SIGMA, ninds, ndemes, S, zone_age, run.id,  
+                                            chromosome, plot_start, plot_stop, increments ) )
+
+deme_ID = rep( 1:params$ndemes, each=params$deme_size )
+
+
+intervalSizes = lapply(positions,function(POS){
+                do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
+                    get.interval.size(IND_DATA=IND,CHR=chromosome,POS=POS,ancA=TRUE)	
+                }))
+            })
+
+intervalSizes_allAncs = lapply(positions,function(POS){
+            do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
+				get.interval.size(IND_DATA=IND,CHR=chromosome,POS=POS,ancA=TRUE,restrict.anc=FALSE)
+            }))
+        })
+
+#[[POS]][ind,chr]
+
+# from deSolve:::drawlegend
+.drawlegend <- function ( zlim=c(0,1), colfn=heat.colors, horizontal=TRUE ) {
+    iy <- 1
+    minz <- zlim[1]
+    maxz <- zlim[2]
+    binwidth <- (maxz - minz)/64
+    ix <- seq(minz + binwidth/2, maxz - binwidth/2, by = binwidth)
[…]

### 2016-02-28 sims/sim-fns.R

@@ -252,11 +252,14 @@ getChunk = function( IND_DATA, CHR, POS ) {
 
 
 # here usually IND_DATA=sims.sums[[1]]$ind.ancest[[1]]
-get.interval.size = function(IND_DATA,CHR=1,POS=0.5,ancA = TRUE){
+get.interval.size = function(IND_DATA,CHR=1,POS=0.5,ancA=TRUE,restrict.anc=TRUE){
 	#return the interval containing focal site for a individual	
 	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
-	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
-	replace(chunk,which(identity==ancA),0)
+    if (restrict.anc) {
+        identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+        replace(chunk,which(identity==ancA),0)
+    }
+    return(chunk)
 }
 
 

### 2016-02-28 drafts/hybrid_zone.tex
a few more edits

@@
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are effectively indistinguishable from more complex models of ecological selection \citep{Barton1989, Barton1993},
-or complex epistatic systems \plr{CITE}.
+previous work has shown that such models generate clines that are effectively indistinguishable from models of ecological selection or epistatic systems \cite{Kruuk1999,Barton2000a}.
 
@@
 
+\alisa{For very strong $s$, very few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may only be present at geographic locations close to the center of the zone.
+The overall length of the haplotypes decreases as the hybrid zone ages, but the signal remains strong relative to the rest of the genome and detectability will largely depend on the density of ancestry-informative markers.}
+
 Additionally, we find that $l_A(m\pm)$, the length of blocks adjacent to blocks containing $m$, are shorter than average when $m$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
@@
 Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short.
-This is measured by the length of the block adjacent to the block containing the selected site.
-A simple and intuitive explanation for this patter in that loci physically linked to a selected site can only survive on the wrong side of the hybrid zone if they recombine off the background of the unfit variant.
-A disproportionate majority of the subsequent recombination events it experiences will be with the more fit ancestry, and so the ancestry block will decay rapidly. 
+This is measured by the mean length of the blocks adjacent to the block containing the selected site.
+A simple and intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond.
+Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby.
 This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.
@@
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
-Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary.
+Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $C(k)$, could provide a simple and informative summary. 
 This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.

### 2016-02-29 drafts/hybrid_zone.tex
minor comments

@@
 Since there appears to be a signal at these regions also, we further define a statistic
-
-$$C(m) =  \frac{2\sum_i{l_B(m_i)}}{\sum_i{l_A(m_i+)+l_A(m_i-)}}$$
-
+\begin{align*}
+C(m) =  \frac{2\sum_i{l_B(m_i)}}{\sum_i{l_A(m_i+)+l_A(m_i-)}}
+\end{align*}
 which is the ratio of mean block length around $m_i$ and mean length of blocks directly flanking the block surrounding $m_i$ (i.e. adjacent blocks).
-
-We are interested in identifying regions of the genome with abnormal distributions of block lengths, and where indicated normalize $l_I(m)$ and $C(m)$ by dividing respectively by mean $l_I$ and $C$  within a deme.
+\plr{what is the sum over? define $m_i$.}
+
+We are interested in identifying regions of the genome with abnormal distributions of block lengths, and where indicated normalize $l_I(m)$ and $C(m)$ by dividing respectively by \plr{the empirical} mean \plr{of} $l_I$ and $C$  within \plr{the corresponding} deme.
+\plr{which deme? should say that this is computed separately by deme?}
 While there is useful information contained in the geographic distribution of block length patterns, obtaining good spatial sampling can often be difficult, and this allows us to predict patterns within relatively few populations.
+\plr{I'm not sure what you mean by "predict patterns".}
 
@@
 Within each deme $X$, we calculated $F_B(m,X)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
+\plr{how about $x$ instead of $X$? above $x$ refers to spatial location.}
 
@@
 Indeed,
-as shown in Figures \ref{Fig:blockLengths} and \ref{Supp:blockLengthHeatmap},
+as shown in Figures \ref{Fig:blockLengths}, \ref{Fig:blockLengthsZoom} and \ref{Supp:blockLengthHeatmap},
 the mean length of such haplotype blocks are up to three times longer
@@
     \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
-}\label{Fig:blockLengths}
+}\label{Fig:blockLengthsZoom}
 \end{figure}
@@
 \plr{This needs a bit more precise discussion.}
-As expected, we find that $F_B(m)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate.
+As expected, we find that $F_B(m,X)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
 We find that at the selected locus, $F_B(m,X)$ decreases with X, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
@@
 \alisa{Could not find a reference to the hybrid filter}
+\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}
 
@@
 The first integral term in \eqref{eqn:g_pde} is
-%       p * unifprod_ut(y=yA,ncx=nabvals)
-%       + (1-p) * biprod_ut(yA=yA,yB=yB,zeroind=zeroind,ncx=nabvals)
-\begin{align*} \label{eqn:biprod}
+\begin{align} \label{eqn:biprod}
     \int_a^0 g_B(a,\theta) g_A(\theta,b) d\theta
     + \int_0^b g_A(a,\theta) g_B(\theta,b) d\theta 
-\end{align*}
+\end{align}
 We now require that one of the grid points along the chromosome is exactly at the selected site;

### 2016-02-29 sims/simulation_SIGMA3_Ninds6000_ndemes120_s0.01_dir/compile-pdfs.sh

@@ -1,2 +1,10 @@
 
 dirs="tau50_dir tau100_dir tau150_dir tau250_dir tau750_dir tau1000_dir"
+
+figs="adjacentBlocksAlongChromAncBConditioning.pdf adjacentBlocksAlongChromNoConditioning.pdf blocksAlongChromAncBConditioning.pdf blocksAlongChromHeatmapAncBConditioning.pdf blocksAlongChromHeatmap.pdf blocksAlongChromNoConditioning.pdf ratioAdjacentBlocksAlongChromAncBConditioning.pdf ratioAdjacentBlocksAlongChromHeatmapAncBConditioning.pdf ratioAdjacentBlocksAlongChromHeatmapNoConditioning.pdf ratioAdjacentBlocksAlongChromNoConditioning.pdf density.pdf ecdf.pdf mean.pdf freqplot.pdf LD.pdf"
+
+for figname in $figs
+do
+    pdfs=$((for dirz in $dirs; do find $dirz -name "*$figname"; done)|tr '\n' ' ')
+    pdfjoin --outfile $figname $pdfs
+done

### 2016-02-29 drafts/hybrid_zone.tex
more edits

@@
 Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space.
-\yb{We find that ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks}, especially when removed from the center of the hybrid zone.
+\alisa{We describe the properties of clines across the genomes as they are formed, and find that locally disadvantageous alleles exist as recent, small families that move relatively rapidly across space}.
+\yb{We find that, within a population, ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks} within the genome, especially when removed from the center of the hybrid zone.
 The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
@@
 }
-
-\plr{say what population density in simulation figures}
 
@@
     Frequency of ancestry B,
-    across geography and at several physical positions on the genome,
-    at $T=100$ generations after secondary contact,
+    across geography and at several physical positions on the genome, simulated for a hybrid zone
+     $T=100$ generations after secondary contact,
     and with $s=0.1$.
+        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
     Each line represents a locus some distance $r$ away from the true target of selection, 
@@
     Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison}.
-    \plr{Add same figure for some different time points and maybe values of $s$ to the supplement.} 
 }\label{alleleFreq_tau100}
@@
 \subsection*{Measures of introgression}
-
-\plr{This needs to be more precise, and should define and give a name to the ratio-of-adjacents statistic here, too.}
 
@@
 \end{align*}
-which is the ratio of mean block length around $m_i$ and mean length of blocks directly flanking the block surrounding $m_i$ (i.e. adjacent blocks).
-\plr{what is the sum over? define $m_i$.}
-
-We are interested in identifying regions of the genome with abnormal distributions of block lengths, and where indicated normalize $l_I(m)$ and $C(m)$ by dividing respectively by \plr{the empirical} mean \plr{of} $l_I$ and $C$  within \plr{the corresponding} deme.
-\plr{which deme? should say that this is computed separately by deme?}
-While there is useful information contained in the geographic distribution of block length patterns, obtaining good spatial sampling can often be difficult, and this allows us to predict patterns within relatively few populations.
-\plr{I'm not sure what you mean by "predict patterns".}
+where $m_i$ is the block length in individual $i$, and $C$ is  the mean block length around $m$ in the population, divided by the mean lengths of  the two blocks directly flanking the block containing $m$.
+
+We are interested in identifying regions of the genome with abnormal distributions of block lengths, and within each deme compute $l_B(m)$ and $C(m)$ across the genome.
+Where indicated, we normalize $l_B(m)$ and $C(m)$ by dividing by the empirical mean of $I_B$ and $C$ across the genome for a given deme.
+While there is useful information contained in the geographic distribution of block length patterns, obtaining good spatial sampling can often be difficult, and this allows us to search for  patterns if block lengths within relatively few populations.
+%\plr{I'm not sure what you mean by "predict patterns".}
 
@@
 This is a reflection of the average size of a haplotype family, and is potentially a source of additional information.
-Within each deme $X$, we calculated $F_B(m,X)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
-\plr{how about $x$ instead of $X$? above $x$ refers to spatial location.}
+Within each deme $x$, we calculated $F_B(m,x)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
+%\plr{how about $x$ instead of $X$? above $x$ refers to spatial location.}
 
@@
 From previous results, we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, clines at closely linked sites are narrower.
-This is seen, for instance, in Figure~\ref{alleleFreq_tau100}.
+This is seen, for instance, in Figures~\ref{alleleFreq_tau100},~\ref{alleel_Freq_tau100_weaker_s} and ~\ref{alleleFreq_tau1000}.
 
@@
 Red blocks along the chromosome denote ancestry $B$, and orange blocks are ancestry $A$.
-Regions surrounding the selected site are failing to introgress.
-    \plr{Add same figure for some different time points and maybe values of $s$ to the supplement.} 
+    50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
 }\label{Fig:resistanceToIntrogression1000g}
@@
 Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m)$ along the chromosome in the deme.
-    \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
+Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
 }\label{Fig:blockLengths}
@@
 \includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
-\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$.
-Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.
-    \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
+\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
+Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m)$ along the chromosome in the deme.
+Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
+This figure represents the same simulation and statistic as Fig.~\ref{Fig:blockLengths}, on a finer genomic scale.
 }\label{Fig:blockLengthsZoom}
@@
 \caption{The number of individuals of ancestry B per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
+The hybrid zone is 50 demes, each containing 500 diploid individuals.
  Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$).
@@
 \plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
-\subsection*{The hybrid filter}
+\subsection*{Other patterns of introgression}
 \alisa{Could not find a reference to the hybrid filter}

### 2016-02-29 drafts/supplement.tex
minor comments
more edits

@@
         as described in the text (see definition of $q_z(x,t,r)$).
-        See figure \ref{sfig:linked_cline_heatmaps}
+        See figure \ref{fig:linked_cline_heatmaps}
         for the same figure over a shorter period of time.
@@
         Each line represents a locus some distance $r$ away from the true target of selection, 
+        and $r=0$ represents the locus that is under selection.    }\label{sfig:alleleFreq_tau100_comparison}
+\end{figure}
+
+\begin{figure}
+\includegraphics{figs/alleleFrequencies_sim_s01_tau100_closely_linked.pdf}
+\caption{
+    Frequency of ancestry B,
+    across geography and at several physical positions on the genome, simulated for a hybrid zone
+     $T=100$ generations after secondary contact,
+    and with $s=0.01$.
+        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+    Each line represents a locus some distance $r$ away from the true target of selection, 
     and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
     identical parameters except $s=0$.
-    }\label{sfig:alleleFreq_tau100_comparison}
+}\label{alleleFreq_tau100_weaker_s}
+\end{figure}
+
+\begin{figure}
+\includegraphics{figs/alleleFrequencies_sim_s1_tau1000.pdf}
+\caption{
+    Frequency of ancestry B,
+    across geography and at several physical positions on the genome, simulated for a hybrid zone
+     $T=1000$ generations after secondary contact,
+    and with $s=0.1$.
+        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+    Each line represents a locus some distance $r$ away from the true target of selection, 
+    and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
+    identical parameters except $s=0$.
+ }\label{alleleFreq_tau1000}
 \end{figure}

### 2016-02-29 drafts/hybrid_zone.tex
some comments

@@
 The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
+\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
+
 }
@@
 centered on their point of contact \citep{Barton1985}. 
-\plr{(check that neutral clines discussed in barton '85)}\alisa{This is simply stated in Barton1985 without any justification or citation. Can we just cite Sedghifar2015?}
+%\plr{(check that neutral clines discussed in barton '85)}
 If the populations are sufficiently diverged,
@@
 the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
-The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.
-\plr{(check Gravel; isn't that approximate?)}
+The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.
 
@@
 
+\alisa{how small is too small for s? when do we start seeing that peak (1/tau)? Bring paragraph here. Look at smaller spatial scale for the small s simulations}
+
 Additionally, we find that $l_A(m\pm)$, the length of blocks adjacent to blocks containing $m$, are shorter than average when $m$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
@@
 and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
-\plr{(check predicted decay of mean length $1/x$ in simulations?)}
+%\plr{(check predicted decay of mean length $1/x$ in simulations?)}
 For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
@@
 %%%%%%
-\begin{figure}
-    \begin{center}
-    \end{center}
-    \caption{
-        \textbf{Haplotype length distributions},
-        comparing numerical solutions to simulations.
-        At one time, plot 
-        probability of seeing a haplotype of length at least $r$ covering that position on the genome,
-        against (a) position on the genome, at a few spatial locations;
-        or (b) spatial location, at a few positions on the genome.
-        \plr{maybe?}
-        \label{fig:haplotype_lengths}
-    }
-\end{figure}
+%\begin{figure}
+ %   \begin{center}
+ %   \end{center}
+%    \caption{
+  %      \textbf{Haplotype length distributions},
+  %      comparing numerical solutions to simulations.
+   %     At one time, plot 
+   %     probability of seeing a haplotype of length at least $r$ covering that position on the genome,
+    %    against (a) position on the genome, at a few spatial locations;
+   %     or (b) spatial location, at a few positions on the genome.
+   %     \plr{maybe?}
+    %    \label{fig:haplotype_lengths}
+   % }
+%\end{figure}
 
@@
 \plr{This needs a bit more precise discussion.}
-As expected, we find that $F_B(m,X)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
-We find that at the selected locus, $F_B(m,X)$ decreases with X, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
+\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
+As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
+% $F_B(m,x)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
+We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
 As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
@@
 \plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
-\subsection*{Other patterns of introgression}
+\subsection*{Adaptive introgression}
 \alisa{Could not find a reference to the hybrid filter}

### 2016-02-29 drafts/hybrid_zone.tex
alisa's edits

@@
 Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space.
-\alisa{We describe the properties of clines across the genomes as they are formed, and find that locally disadvantageous alleles exist as recent, small families that move relatively rapidly across space}.
-\yb{We find that, within a population, ancestral blocks surrounding single-locus incompatibilities} are longer than average \yb{ancestry blocks} within the genome, especially when removed from the center of the hybrid zone.
+\alisa{We describe the properties of clines across the genomes as they are formed, as well as correlations in ancestry.
+Clines are expected to form over time $1/s$, where $s$ is the strength of selection against hybrids.
+Linked clines persist over a  physical scale of $1/T$, where $T$ is the age of the hybrid zone.
+Additionally, we find that locally disadvantageous alleles move at speed $\sigma\sqrt{s}$ (where $\sigma$ is the variance in dispersal), and tend to exist as relatively recent, small families.
+The length of continuous tracts of ancestry provide an additional source of information; within a population, blocks of ancestry surrounding single-locus incompatibilities} are longer than the genome-wide average block length, especially when removed from the center of the hybrid zone.
 The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
-\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
+%\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
 
@@
 
-\plr{
-Main points, make sure they appear in the abstract:
-clines establish over time $1/s$, extend along genome to at least $1/T$;
-locally disadvantageous alleles exist as recent, small families (moving at speed $\sigma/\sqrt{s}$);
-haplotype chunks longest around introgressing selected bits, flanked by shorter chunks.
-}
+%\plr{
+%Main points, make sure they appear in the abstract:
+%clines establish over time $1/s$, extend along genome to at least $1/T$;
+%locally disadvantageous alleles exist as recent, small families (moving at speed $\sigma/\sqrt{s}$);
+%haplotype chunks longest around introgressing selected bits, flanked by shorter chunks.
+%}
 
@@
 before reproductive isolation is complete, gene flow is still possible. 
-As two such populations interbreed,
+As two such populations interbreed, tracts of
 individuals' genomes can be labeled according to which ancestral population they are inherited from,
 and local interbreeding and migration of such populations
-creates a gradient of ancestry assignments across geography,
+create a gradient of ancestry assignments across geographic space,
 centered on their point of contact \citep{Barton1985}. 
@@
 leading to substantial correlations in ancestry between distant markers.
-
-Gene flow will eventually erase 
-both geographic clines in ancestry, and strong correlations in ancestry.
-Subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
+The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci.
+%the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
 The mixing of ancestries of a given genome is mediated by 
 the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
-The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.
-
+Patterns of correlated ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.
+
+While for much of the genome gene flow will eventually erase 
+both geographic clines in ancestry, and strong correlations in ancestry,
+subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
 % Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}.
@@
 at loci close to selected loci, and wider clines at less tightly linked loci.
-This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}.
+This provides some expectation for how selection shapes covariance in ancestry at a single locus over \emph{geographic space}.
 Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}.
@@
 Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
-Our goal is to understand how selection against hybrids interacts with dispersal and recombination to produce observed patterns ancestry around selected sites in recently established hybrid zones.
+Our goal is to understand how selection against hybrids interacts with dispersal and recombination to produce observed patterns of ancestry around selected sites in recently established hybrid zones.
 In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
 In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one.
-If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and  patterns of introgression can provide information about selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and providing information about the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}.
 \yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
@@
 We therefore expect \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.
+As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus is expected to be relatively long when far away from the zone center.
 This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
 
-Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
+Using a combination of theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
 For this, we build on the neutral framework presented in \citet{Sedghifar2015} to derive the expected distribution of ancestry block lengths across a contact zone.
@@
 
-We model selection through only a single, underdominant locus:
-at this locus, there are two alleles, one corresponding to each of the ancestral populations,
-and individuals who are heterozygous at this locus produce on average $1-s$ times as many offspring than either homozygote.
+We model selection through a single, underdominant locus:
+at this locus, there are two alleles, one corresponding to each of the ancestral populations.
+Individuals who are heterozygous at this locus produce on average $1-s$ times as many offspring than either homozygote.
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are effectively indistinguishable from models of ecological selection or epistatic systems \cite{Kruuk1999,Barton2000a}.
+previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \cite{Kruuk1999,Barton2000a}.
 
 As in \citet{Sedghifar2015}, 
-we gain substantial insight by considering where the ancestors of modern-day individuals lived,
+we gain insight by considering where the ancestors of modern-day individuals lived,
 i.e., where their lineages fall across geography as we look further back towards the time of first contact.
 We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$ at the time of secondary contact,
-i.e., if its lineage traces back to to the $A$ side.
+i.e., if its lineage traces back to to the $x<0$.
 A block of genome is of ancestry $A$ if every locus in it does the same;
@@
 This process, in which recombination events cause such lineages to \emph{branch},
-is depicted in Figure~\ref{Fig:schematic}.
+is illustrated in Figure~\ref{Fig:schematic}.
 We say that a block of genome is on the $A$ \emph{background} if it is physically linked to an allele of ancestry $A$ at the selected locus;
-if the block includes the selected site then it is necessarily on the $A$ background.
+if a block of ancestry $A$ includes the selected site then it is necessarily on the $A$ background.
 As the identity of the selected allele determines how selection acts on the haplotype,
@@
 \paragraph{Establishment of the cline at the selected locus}
-The deterministic theory predicts that after secondary contact,
+Deterministic theory predicts that after secondary contact,
 the two alleles at the selected locus will form a stable cline \citep{Barton1979},
@@
 Now suppose that the frequency profile of the selected allele $p(x,t)$ is known.
-We will understand first how lineages of sampled individuals behave at the selected locus,
+We seek to first understand how lineages of sampled individuals behave at the selected locus,
 from which lineages at linked loci require only a minor modification.
@@
 Using that $\E[R]=0$ and $\E[R^2]=\sigma^2$
-and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
-\plr{Does that help at all?}
+and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$.
 
@@
 Therefore, roughly, lineages on ``their own'' side wander randomly,
-while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
+while lineages on ``the wrong'' side are pushed at constant speed $\sigma\sqrt{s}$ 
 back towards the side where they are more common. 
@@
 Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
-We will follow its lineage back through time, using $\tau$ to denote ``time ago'' to avoid confusion.
+We will follow its lineage back through time, using $\tau$ to denote ``generations ago'' to avoid confusion.
 If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
@@
 \end{align}
-
-\alisa{This process of a haplotype sampled in the present day tracing it's path $X_\tau$ backwards in time, while splitting and changing backgrounds $Z_\tau$ through recombination, is shown in Figure~\ref{Fig:schematic}.}
-\plr{Connect these up more.}
 
@@
 Notice that equations \eqref{eqn:g_pde}
-are hierarchical in $[a,b]$:
-the equation for haplotype identity probabilities on a segment $[a,b]$ depends only on those probabilities for segments contained in $[a,b]$.
+are hierarchical in $(a,b)$:
+the equation for haplotype identity probabilities on a segment $(a,b)$ depends only on those probabilities for segments contained in $(a,b)$.
 This allows for efficient numerical solutions,
@@
 \caption{
-    Frequency of ancestry B,
-    across geography and at several physical positions on the genome, simulated for a hybrid zone
+    Frequency of ancestry $B$,
+    across geography at different physical positions on the genome, simulated for a hybrid zone
      $T=100$ generations after secondary contact,
     and with $s=0.1$.
-        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+        $50$ demes, each with population size $N_d=500$ diploid individuals, were simulated across the hybrid zone.
     Each line represents a locus some distance $r$ away from the true target of selection, 
-    and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
+with $r=0$ representing the locus under selection. Grey lines represent the same positions from a simulation with
     identical parameters except $s=0$.
@@
 The most straightforward measure is $l_I(m)$,  the mean length of all haplotype chunks that contain genomic position $m$, given that $m$ is of ancestry $I$, where $I\in\{A,B\}$.
-We then focus on $I_B(m)$ for $m$ across a chromosome. 
+We then focus on $l_B(m)$ for $m$ across a chromosome. 
 %Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
@@
 In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
-This is a reflection of the average size of a haplotype family, and is potentially a source of additional information.
+This is a reflection of the average size of a haplotype family, and is potentially a source of additional information about selection.
 Within each deme $x$, we calculated $F_B(m,x)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
@@
 \paragraph{The timescale over which linked clines relax:}
-Although loci not under selection can in principle spread across the cline unimpeded,
+Although loci not under direct selection can in principle spread across the cline unimpeded,
 in practice it can take quite some time for even unlinked neutral loci to homogenize,
@@
 
-\plr{Figure showing this? 
-    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
-    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
-    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
-}
-
 %%%%% %%%%%%%%%%
@@
 showing distinct patterns across geography and the genome.
+
+We expect that, in the absence of selection, blocks of $A$ ancestry across the genome 
+will tend to be shorter the further one goes into the $B$ side of the cline,
+because they have had more opportunities to recombine with $B$ haplotypes. 
+However, we expect that stretches of $A$ ancestry containing a selected site will be longer
+than those that do not contain the selected site at the same spatial location,
+because lineages containing the selected site have usually been inherited from the $A$ side of the cline recently. 
+As discussed above, we expect these lineages move at speed roughly $\sigma\sqrt{s}$,
+so (selected) $A$ alleles at distance $x$ from the cline center
+to have last had an ancestor on the $A$ side of the cline around $x /\sqrt{s}\sigma$ generations ago 
+(compared to $x^2/\sigma^2$ for a neutral allele),
+and so the enclosing $A$ haplotype is no longer than $\sigma\sqrt{s}/x $.
+%\plr{(check predicted decay of mean length $1/x$ in simulations?)}
+For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
 
@@
 % This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
-The mean haplotype length found without conditioning on ancestry ($l_\cdot$)
+The mean haplotype length found without conditioning on ancestry ($l_\cdot(m)$)
 also shows an small increase near the selected locus,
@@
 
-\alisa{For very strong $s$, very few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may only be present at geographic locations close to the center of the zone.
+\alisa{For weaker selection ($s=0.001$), then the relative speed of retreat is not fast enough compared to the genomic background, and the strength of signal from elevated $l_B(m)$ is diminished.
+The resolution of the peak is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
+Since clines in ancestry establish over the timescale $1/s$, it is difficult to intuit the time it takes for a signal of elevated $l_B(m)$ to form around the selected locus.
+
+We note that for very strong $s$, few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may be present only at geographic locations close to the center of the zone.}
+
+%\alisa{For very strong $s$, very few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may only be present at geographic locations close to the center of the zone.
 The overall length of the haplotypes decreases as the hybrid zone ages, but the signal remains strong relative to the rest of the genome and detectability will largely depend on the density of ancestry-informative markers.}
@@
 This pattern provides an avenue for further enhancement of the block-length signal around a selected locus, by measuring $C(m)$  (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap}).
-
-We expect that, in the absence of selection, blocks of $A$ ancestry across the genome 
-will tend to be shorter the further one goes into the $B$ side of the cline,
-because they have had more opportunities to recombine with $B$ haplotypes. 
-However, we expect that stretches of $A$ ancestry containing a selected site will be longer
-than those that do not contain the selected site at the same spatial location,
-because lineages containing the selected site have usually been inherited from the $A$ side of the cline recently. 
-As discussed above, we expect these lineages move at speed roughly $\sigma/\sqrt{s}$,
-so (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago 
-(compared to $x^2/\sigma^2$ for a neutral allele),
-and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
-%\plr{(check predicted decay of mean length $1/x$ in simulations?)}
-For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
 
@@
 \includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
-Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m)$ along the chromosome in the deme.
+\caption{Mean block length, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
+Here each line represents  $l_B(m)$, in a given deme, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
 Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
@@
 \includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
-\caption{Mean block length, $l_B(m_i)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
-Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(m)$ along the chromosome in the deme.
+\caption{Mean block length, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
+Here each line represents  $l_B(m)$, in a given deme, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
 Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
@@
 \plr{This needs a bit more precise discussion.}
-\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
-As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
+
+It is possible that there is additional information contained in patterns of within-ancestry haplotype diversity throughout the hybrid zone.
+Looking at the number of ancestral haplotypes present in the population would provide informationa bout whether the migrant contribution is through relatively few, successful migrants, or through many migrants that each contribute relatively little.
+
+We find that the average family size of a $B$ allele ($F_B(m,x)$) decrease with distance into the $A$ side of the hybrid zone, and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}).
+As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
+This pattern is expected and consistent with the prediction in our models that unfit lineages tend to be recent migrants, which will have smaller families.
+In conjunction with block length, measuring $F_B$ can provide an additional signal for characterizing lineages strongly influenced by selection.
+
+%\alisa{add motivation-- additional information in haplotype diversity. Present alternate hypotheses. This ties in with our world view within our model}
+%As expected, we find that the average family size of a $B$ allele decreases with distance into the $A$ side of the hybrid zone. 
 % $F_B(m,x)$ is highest on the size of the hybrid zone where individuals of ancestry $B$ predominate. 
-We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
+%We find that at the selected locus, $F_B(m,x)$ decreases with x, and is lowest far away from the zone center, where ancestry $B$ is at low frequency  (Figure~\ref{Fig:family_size}).
 As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
 
-One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families.
+%One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families.
 In conjunction with block length, measuring $F$ can provide an additional signal for characterizing lineages strongly influenced by selection, by looking at levels of haplotype diversity in populations along a hybrid zone.
@@
 \includegraphics{figs/number_of_ancestors_tau1000}
-\caption{The number of individuals of ancestry B per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
+\caption{The number of individuals of ancestry $B$ per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
 The hybrid zone is 50 demes, each containing 500 diploid individuals.
@@
 We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
-This occurs at speed $\sigma/\sqrt{s}$.
+This occurs at speed $\sigma\sqrt{s}$.
 Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
@@
 
-\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-    which we have done using both theory and simulation.
-    Our most useful observations are that clines establish over time $1/s$,
-    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-    tend to move back towards the side where they are common
-    at speed $\sigma/\sqrt{s}$.
-    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-}
+%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
+%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
+%    which we have done using both theory and simulation.
+%    Our most useful observations are that clines establish over time $1/s$,
+%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
+ %   tend to move back towards the side where they are common
+ %   at speed $\sigma/\sqrt{s}$.
+ %   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
+%}
 
@@
 We find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
-
-\alisa{These seem like an orphan statement. Any idea where to stick these?}
-We further find that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will be descended from a relatively large number of independent ancestors, per individual, of the less fit ancestry.
-This result reflects that, on average, less fit haplotypes tend to be recent migrants with few descendants.
 
@@
 
-\plr{Remove "Future directions", title this "The hybrid filter", and discuss the hybrid filter ideas we had?}
 \subsection*{Adaptive introgression}
-\alisa{Could not find a reference to the hybrid filter}
-\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}
+%\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}
 

### 2016-03-01 drafts/hybrid_zone.tex
typo hunt

@@
 
-While for much of the genome gene flow will eventually erase 
-both geographic clines in ancestry, and strong correlations in ancestry,
+While for much of the genome, gene flow will eventually erase 
+both geographic clines in ancestry, and strong correlations in ancestry across linked markers, 
 subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
@@
 We consider two isolated populations
---- labeled \emph{Species $A$} and \emph{Species $B$} ---  
+--- labeled species $A$ and species $B$ ---   % yb: for consistency with the rest of the MS
 that came into contact $T$ generations in the past.  
@@
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \cite{Kruuk1999,Barton2000a}.
+previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,Barton2000a}.
 
@@
 where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as derived in \citet{Bazykin1969}; for discussion of diploidy see \citep{christiansen1995genotypic}.
+as derived in \citet{Bazykin1969}; for discussion of diploidy see \citet{christiansen1995genotypic}.
 Although there is no known exact analytic solution to this equation, even in steady state, 
@@
 how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with drift $\sigma^2 \grad \log p(x,t)$ (i.e., it moves at speed $\sigma^2$ in the potential $-\log p$).
+with drift $\sigma^2 \grad \log p(x,t)$ (i.e., it moves at speed $\sigma^2$ in the potential $-\log p$, note that we use ``drift'' as directional movement as in the terminology of diffusion, rather than its population genetic meaning).
+% YB for clarity
 % \alisa{note to self to read through appendix when written}
@@
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ (same as previously described movement of a selected allele),
-and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$)or $r p$ (for $B$ to $A$).
+and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (for $B$ to $A$).
 
@@
 We can use this diffusion model for lineages to compute expected clines in ancestry,
-i.e., the expected proportion of individuals who inherit from species $A$,
+i.e., the expected proportion of individuals who inherit ancestry from species $A$,
 as a function of space, time, and position on the genome.
@@
 
-The resulting equation si similar to \eqref{eqn:q_pde} with a term added for branching,
+The resulting equation is similar to \eqref{eqn:q_pde} with a term added for branching,
 which is written (omitting the $(x,t)$ for conciseness):
@@
 
-Each deme has exactly $N_d$ diploid individuals at the start of each generation.
+Each deme has exactly $N_d$ diploid hermaphroditic individuals at the start of each generation.
 At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
@@
 and the relative slowness of diffusive movement \citep{Sedghifar2015}.
-We are interested, for samples across a cline $T$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
+We wish to know, for samples across a cline $T$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
 
@@
 The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone.
-We are interested specifically in the how tracts of ancestry surrounding the selected locus compares to the rest of the genome.
+We are interested specifically in how the tracts of ancestry surrounding the selected locus compares to the rest of the genome.
 The raw data -- ancestry assignments for a few individuals sampled from across space --
@@
 It is possible that there is additional information contained in patterns of within-ancestry haplotype diversity throughout the hybrid zone.
-Looking at the number of ancestral haplotypes present in the population would provide informationa bout whether the migrant contribution is through relatively few, successful migrants, or through many migrants that each contribute relatively little.
+Looking at the number of ancestral haplotypes present in the population would provide information about whether the migrant contribution is through relatively few, successful migrants, or through many migrants that each contribute relatively little.
 
@@
 Of particular concern that stems from the above assumption is that our model ignores potential, and likely, sources of stochasticity arising from geographic structure and coalescence/pedigree structure.
-The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \cite{Liang2014}.
+The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{Liang2014}.
 With respect to genealogy, the broad agreement between our analytic predictions and simulations (which include an explicit pedigree structure), suggests that this assumption does not have a severe qualitative impact on our major results when populations are moderately high, especially for all but the most recently formed hybrid zones.

### 2016-03-01 drafts/hybrid_zone.tex
one minor typo

@@
 Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
-This pattern reflects the intuition that regions surrounding selected loci introgress less readily, and loci under selection can therefore be thought as moving under Brownian motion with drift whereby a lineage that is presently on the `` wrong" side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct" side.
+This pattern reflects the intuition that regions surrounding selected loci introgress less readily, and loci under selection can therefore be thought as moving under Brownian motion with drift whereby a lineage that is presently on the ``wrong" side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct" side.
 In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong" side, are likely to have done so relatively recently.

### 2016-03-03 drafts/hybrid_zone.tex
added small note about deme size to Assumptions section
peter edited 2/3 through

@@
 
-\title{Blocks of extended ancestry around selected loci in hybrid zones}
+% other title ideas:
+% {Blocks of extended ancestry around selected loci in hybrid zones}
+\title{Beyond clines: lineages and haplotype blocks in hybrid zones}
 \author{}
@@
 \abstract{
-Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation.
-\yb{Some form of selection against hybrid genotypes is likely common in many hybrid zones.
-Here, we use a combination of analytical theory and explicit forward simulations} to gain a better understanding of how this impacts patterns of introgression \yb{across genomic and geographic} space.
-\alisa{We describe the properties of clines across the genomes as they are formed, as well as correlations in ancestry.
-Clines are expected to form over time $1/s$, where $s$ is the strength of selection against hybrids.
-Linked clines persist over a  physical scale of $1/T$, where $T$ is the age of the hybrid zone.
-Additionally, we find that locally disadvantageous alleles move at speed $\sigma\sqrt{s}$ (where $\sigma$ is the variance in dispersal), and tend to exist as relatively recent, small families.
-The length of continuous tracts of ancestry provide an additional source of information; within a population, blocks of ancestry surrounding single-locus incompatibilities} are longer than the genome-wide average block length, especially when removed from the center of the hybrid zone.
+Hybrid zones formed between recently diverged populations offer an opportunity 
+to better understand the mechanisms underlying reproductive isolation and the process of speciation. 
+Here, we use a combination of analytical theory and explicit forward simulations
+to gain a better understanding of how selection against hybrid gentypes impacts patterns of introgression 
+across genomic and geographic space. 
+By understanding how lineages move across the hybrid zone, in a model without coalescence, 
+we describe how clines form and how parental haplotypes are broken up during introgression.
+Clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
+and linked clines persist over a  physical scale of $1/T$, where $T$ is the age of the hybrid zone. 
+Locally disadvantageous alleles tend to exist as small families,
+whose lineages trace back to the side they originated on at speed $\sqrt{s}$ dispersal distances per generation.
+The lengths of continuous tracts of ancestry provide an additional source of information:
+blocks of ancestry surrounding single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location. 
 The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
 %\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
-
 }
@@
 \doublespacing
-
-%\plr{
-%Main points, make sure they appear in the abstract:
-%clines establish over time $1/s$, extend along genome to at least $1/T$;
-%locally disadvantageous alleles exist as recent, small families (moving at speed $\sigma/\sqrt{s}$);
-%haplotype chunks longest around introgressing selected bits, flanked by shorter chunks.
-%}
 
@@
 
-The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
-When formerly isolated populations come back into contact 
-before reproductive isolation is complete, gene flow is still possible. 
-As two such populations interbreed, tracts of
-individuals' genomes can be labeled according to which ancestral population they are inherited from,
+The process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
+Gene flow between formerly isolated populations is possible if reproductive isolation is not complete.
+As two such populations interbreed, each portion of each individual's genome 
+can be labeled according to which ancestral population is is inherited from,
 and local interbreeding and migration of such populations
-create a gradient of ancestry assignments across geographic space,
-centered on their point of contact \citep{Barton1985}. 
+creates a gradient of ancestry assignments across geographic space,
+centered on their point of contact \citep[reviewed in][]{Barton1985}. 
 %\plr{(check that neutral clines discussed in barton '85)}
@@
 are broken up by historical recombination events,
-leading to substantial correlations in ancestry between distant markers.
-The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci.
-%the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}.
-The mixing of ancestries of a given genome is mediated by 
-the process of recombination, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
+forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003},
+leading to substantial correlations in ancestry (and hence, genotype) between distant markers.
 Patterns of correlated ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.
@@
 While for much of the genome, gene flow will eventually erase 
-both geographic clines in ancestry, and strong correlations in ancestry across linked markers, 
+both geographic clines in ancestry and strong correlations in ancestry across linked markers, 
 subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
-% Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}.
-% While these patterns of co-ancestry have been described in multi-locus systems, the \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}
-\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones.
-} %GREAT SENTENCE!!
-Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.
-
-The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \citep{Barton1979a,Barton1986}.  
+Studies of hybrid zones have primarily been based on single-site statistics such as clines,
+but today, 
+the availability of dense genotype data in combination with linkage information 
+makes it now possible to use patterns of the block length distribution
+to better understand the processes operating in hybrid zones.
+
+\plr{Literature: Haldane, Nagylaki, Slatkin, \citep{May1975,Slatkin1973}}
+\plr{Better summary of what Barton did.}
+Population genetic modeling of hybrid zones has focused largely on the 
+equilibrium states of clines at selected loci \citep{Barton1979a,Barton1986}.  
 Since stronger selection reduces the amount of genetic mixing,
@@
 at loci close to selected loci, and wider clines at less tightly linked loci.
-This provides some expectation for how selection shapes covariance in ancestry at a single locus over \emph{geographic space}.
-Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}.
-
-Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
-Our goal is to understand how selection against hybrids interacts with dispersal and recombination to produce observed patterns of ancestry around selected sites in recently established hybrid zones.
-In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness.
+This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
+Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.
+
+The arrangement of segments of genome inheriting from particular ancestral populations
+provides information about the number of recombination events that have taken place between ancestral genotypes, 
+and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, 
+this could be especially important for understanding the impact of selection on the mixing of ancestral genomes and, consequently, the maintenance of species distinctness. 
 In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and providing information about the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}.
-\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
-\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}
+If hybrids have non-zero fitness, some amount of introgression can occur, providing information about the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
+The resulting heterogeneous patterns of ancestry across the genome have been used to identify putative targets of selection in hybridizing populations,
+by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}.
 
 Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
-We therefore expect \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
+We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
 As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus is expected to be relatively long when far away from the zone center.
-This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
-
-Using a combination of theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
-For this, we build on the neutral framework presented in \citet{Sedghifar2015} to derive the expected distribution of ancestry block lengths across a contact zone.
-\yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
-
-%\plr{explain relationship to \citet{Sedghifar2015}}
-
-\plr{cite selection-equals-subdivided-population stuff}
+This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
+as many ancestors of the locally common type as do neutral haplotypes.
+
+Using a combination of theory and simulated hybrid zones, we describe genome-wide patterns of coancestry 
+as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
+In \citet{Sedghifar2015} we described how lineages move across a neutral zone of secondary contact,
+obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
+In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
+by describing how lineages move conditional on the frequencies of the selected allele
+and allowing lineages to recombine between selected background 
+(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
 
@@
 We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$ at the time of secondary contact,
-i.e., if its lineage traces back to to the $x<0$.
+i.e., if its lineage traces back to a spatial position $x<0$.
 A block of genome is of ancestry $A$ if every locus in it does the same;
@@
 
+%%%%%%%%
 \begin{figure}
 \includegraphics{figs/BM_schematic}
-\caption{A cartoon representation of a haplotype under the model.
-The black line represents the path, backwards in time, of a site under selection ($L_B$) that is sampled in the present day.
-The center of the hybrid zone is defined at geographic position $x=0$, such that ancestry A segregates at lower frequency for $x>0$.
-Here, because of negative selection, $L_B$ migrates rapidly back to $x>0$ from a position $x<0$.
-Red dots represent recombination events along the entire haplotype.
-When a recombination separates two chunks of the sampled haplotype takes place, a branching event occurs and these splitting lineages are represented by colored lines (here the initial haplotype is broken up into 3 chunks over 100 generations).
-Subsequent recombination events can switch the selected background to which a chunk is linked.
-Here, yellow lineages are linked to $L_A$ , and blue lineages are linked to $L_B$.
-Based on position 100 generations after contact, from right to left the chunks are of ancestry $(A,B,B)$.}\label{Fig:schematic}
+\caption{
+    \textbf{A schematic figure of the lineages along which a segment of genome has been inherited,}
+    i.e., showing where the ancestors lived from whom the haplotype was inherited (horizontal axis),
+    and how long ago they lived (vertical axis).
+    The segment contains the $B$ alelle at the selected locus.
+    The black line shows the lineage at the selected site in the sampled individual in the present day (at $\tau=0$).
+    The center of the hybrid zone is at geographic position $x=0$; in the region $x>0$, ancestry $A$ is less common,
+    and because of negative selection, the selected lineage spends little time there.
+    Red dots represent ancestral recombination events that fall in the given segment of genome in an ancestor;
+    these can cause the lineage to branch (lineages at linked sites are depicted as colored lines),
+    or can switch lineages at linked sites between selected backgrounds 
+    (yellow lineages are linked to the $A$ selected locus; blue are linked to $B$).
+    Here the initial haplotype is broken up into 3 chunks over 100 generations,
+    which switch between selected backgrounds once.
+    Based on position of the lineages at $\tau=100$ generations after contact, 
+    from left to right the segments are of ancestry $(A,B,B)$.
+}\label{Fig:schematic}
 \end{figure}
-
-%As an approximation, we consider a cline of narrow width (look up cline width for underdominance, Barton/May?), such that the frequency of migrant alleles is exponential and negligible at distance away from the center of the hybrid zone.
-This means that recombination events that occur away from the center of the zone do not result in a change in $a$.
 
@@
 $\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
-This relies on numerous approximations, but the general conclusions should be quite robust: 
+This relies on several approximations, but the general conclusions should be quite robust: 
 the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
@@
 implying that the cline establishes over a timescale of the order $1/s$. 
-In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes. 
-%\alisa{I don't think I understand the independence from $\sigma$?}
-%\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}
+In other words, the cline is established through diffusion, despite being slowed down somewhat by selection against heterozygotes.
 
 \paragraph{Lineages at the selected locus}
-Now suppose that the frequency profile of the selected allele $p(x,t)$ is known.
+Now suppose that the frequency profile of the selected allele, $p(x,t)$, is known.
 We seek to first understand how lineages of sampled individuals behave at the selected locus,
@@
 how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with drift $\sigma^2 \grad \log p(x,t)$ (i.e., it moves at speed $\sigma^2$ in the potential $-\log p$, note that we use ``drift'' as directional movement as in the terminology of diffusion, rather than its population genetic meaning).
-% YB for clarity
+with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
+(i.e., it moves at speed $\sigma^2$ in the potential $-\log p$).
 % \alisa{note to self to read through appendix when written}
 To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
-lived at position $x+r$ is proportional to $\P\{R=r\} p(x+r)$, where $R$ is the random dispersal distance;
+lived at position $x-r$ is proportional to $\P\{R=r\} p(x+r)$, where $R$ is the random dispersal distance;
 and so the mean displacement from offspring to parent is $\E[R p(x+R)]/\E[p(x+R)]$.
@@
 Using that $\E[R]=0$ and $\E[R^2]=\sigma^2$
-and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$.
+and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
+which is $\sigma^2$ multiplied by the gradient of $\log p(x)$.
 
@@
 at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
-Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. 
 % \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}
 
-Note that this description gives more information than the steady-state solution,
-since in that analysis only the ratio $\sigma/\sqrt{s}$ appears.
+Note that this description gives more information than the steady-state cline $p(x)$,
+which depends only on $\sigma/\sqrt{s}$.
 Here we see that lineages with $\sigma=10$ and $s=.16$ move quite differently
@@
 it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
-but if it spends time in the clinal region where recombination may happen,
+but if it spends time in the hybrid zone, where recombination between selected backgrounds may happen,
 it may be found to have been inherited from a $B$-carrying individual,
@@
 
-Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
+Suppose we sample an allele today at a locus at recombination distance $r$ from the selected site.
 We will follow its lineage back through time, using $\tau$ to denote ``generations ago'' to avoid confusion.
@@
 and $Z_\tau$ is the identity of the selected allele that ancestor carried,
-then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ (same as previously described movement of a selected allele),
-and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (for $B$ to $A$).
+then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ 
+(just as previously described for a selected allele),
+and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$).
 
@@
         % \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B \} &= \epsilon r p(X_\tau,\tau) + O(\epsilon^2)  .
-        \P\{ T_B &> u \given Z_\tau = A \} = \exp\left( - r \int_0^u (1-p(X_s,s)) ds \right) \\
-        \P\{ T_A &> u \given Z_\tau = B \} = \exp\left( - r \int_0^u p(X_s,s) ds \right) .
+        \P\{ T_B &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
+        \P\{ T_A &> \tau+u \given Z_\tau = B \} = \exp\left( - r \int_\tau^{\tau+u} p(X_s,s) ds \right) .
     \end{aligned}
@@
 \paragraph{Linked clines}
-We can use this diffusion model for lineages to compute expected clines in ancestry,
+We can use this diffusion model for lineages to find expected clines in ancestry,
 i.e., the expected proportion of individuals who inherit ancestry from species $A$,
@@
 We denote this probability $q_z(x,t,r)$.
-In other words,
+In the notation above,
 \begin{align}
@@
 $G^B = \sigma^2 \grad \log(1-p(x,t)) \cdot \grad + (\sigma^2/2) \Delta$,
-and then equation \eqref{eqn:q_pde} can be written more compactly as
+so that equation \eqref{eqn:q_pde} can be written more compactly as
 \begin{align*}
-    \dot q_A &= G^A q_A + r (1-p) (q_A-q_B) \\
-    \dot q_B &= G^B q_B + r (1-p) (q_B-q_A) .
+    \dot q_A &= G^A q_A + r (1-p) (q_B-q_A) \\
+    \dot q_B &= G^B q_B + r p (q_A-q_B) .
 \end{align*}
-In technical terms, $G^A$ and $G^B$ are the generators of the diffusions of $A$ and $B$ lineages, respectively.
+In technical terms, $G^A$ and $G^B$ are the generators of the diffusions of lineages of selected alleles of ancestry $A$ and $B$, respectively.
 
@@
 at a site at recombination distance $r$ from the selected site,
-which is computed as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 
+which is computed as $q(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 
 The equations are solved numerically in R, using the ReacTran package \citep{soetaert2012reactive}.
@@
 of a single ancestry at a given location and time.
-To quantify haplotype abundances, we start by considering a segment of the genome between positions $a$ and $b$, relative to the selected site. 
+Consider a segment of the genome between positions $a$ and $b$, relative to the selected site. 
 We are interested in the probability $g_z(x,t;a,b)$ 
@@
 Additionally, at rate equal to the genetic length of the segment,
-a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
+a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location between $a$ and $b$,
 each of which proceeds as before, independently. 
@@
             \\ {} & \qquad 
-            + (1-p) \left( \int_a^0 
-                g_B(a,\theta) g_A(\theta,b)
-                + \int_0^b g_A(a,\theta) g_B(\theta,b)
-            d\theta \right)
+            + (1-p) \left( 
+                \int_a^0 g_B(a,\theta) g_A(\theta,b) d\theta
+                + \int_0^b g_A(a,\theta) g_B(\theta,b) d\theta 
+            \right)
             \\ {} & \qquad 
@@
 that the chromosome inherited that region of genome from;
-genotyping at individual loci was performed by looking up the ancestor's genotype.
-
-Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1$ and $s=0.01$.
-From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$).
-These were compared to the output of neutral simulations with $s=0$.
-
-The simulations were executed in R, with scripts available on
-\url{https://github.com/petrelharp/clinal-lineages}
+ancestry assignment at individual loci was performed by looking up the ancestor's genotype.
+
+The simulations were executed in R, with scripts available at
+\url{https://github.com/petrelharp/clinal-lineages}.
 
@@
     across geography at different physical positions on the genome, simulated for a hybrid zone
-     $T=100$ generations after secondary contact,
-    and with $s=0.1$.
-        $50$ demes, each with population size $N_d=500$ diploid individuals, were simulated across the hybrid zone.
+     $T=100$ generations after secondary contact, with $s=0.1$,
+     using 50 demes, each with 500 diploid individuals and $\sigma=1$.
     Each line represents a locus some distance $r$ away from the true target of selection, 
-with $r=0$ representing the locus under selection. Grey lines represent the same positions from a simulation with
+    with $r=0$ representing the locus under selection. 
+    Grey lines represent the same positions from a simulation with
     identical parameters except $s=0$.
-    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison}.
+    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
+    the same plot is shown with weaker selection in Figure \ref{alleleFreq_tau100_weaker_s} and at a longer time in Figure \ref{alleleFreq_tau1000}.
 }\label{alleleFreq_tau100}
@@
 The above theory and simulations allow us to make predictions about patterns surrounding selected loci. 
-In reality, however, such loci are not known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection.
-The most straightforward measure is $l_I(m)$,  the mean length of all haplotype chunks that contain genomic position $m$, given that $m$ is of ancestry $I$, where $I\in\{A,B\}$.
-We then focus on $l_B(m)$ for $m$ across a chromosome. 
-%Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.
-
-Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$ that flank, to the left and right respectively, the block of unbroken ancestry containing $m$.
-Since there appears to be a signal at these regions also, we further define a statistic
+In reality, however, such loci are not usually known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection. 
+The most straightforward measure is $l_B(m,x)$,  
+the mean length of all contiguous segments of ancestry $B$ sampled at position $x$ that contain genomic position $m$.
+Likewise, $l_A(m,x)$ is the mean length of segments of ancestry $A$.
+We compute similarly the unconditioned mean block length $l(m,x)$
+by averaging all segment lengths without regard to ancestral identity.
+
+We also look at the mean length of the two chunks, $m-$ and $m+$, that flank, to the left and right respectively, 
+the block of unbroken ancestry containing $m$. 
+As described below, these blocks tend to be \emph{shorter} than average,
+motivating us to define the statistic
 \begin{align*}
-C(m) =  \frac{2\sum_i{l_B(m_i)}}{\sum_i{l_A(m_i+)+l_A(m_i-)}}
+     C(m,x) =  \frac{2\sum_i{l_B(m_i,x)}}{\sum_i{l_A(m_i+,x)+l_A(m_i-,x)}}
 \end{align*}
-where $m_i$ is the block length in individual $i$, and $C$ is  the mean block length around $m$ in the population, divided by the mean lengths of  the two blocks directly flanking the block containing $m$.
-
-We are interested in identifying regions of the genome with abnormal distributions of block lengths, and within each deme compute $l_B(m)$ and $C(m)$ across the genome.
-Where indicated, we normalize $l_B(m)$ and $C(m)$ by dividing by the empirical mean of $I_B$ and $C$ across the genome for a given deme.
-While there is useful information contained in the geographic distribution of block length patterns, obtaining good spatial sampling can often be difficult, and this allows us to search for  patterns if block lengths within relatively few populations.
-%\plr{I'm not sure what you mean by "predict patterns".}
+where $m_i$ is the block length in individual $i$, and the sum is over individuals at location $x$.
+The statistic $C$ is the mean block length around $m$ in the population, 
+divided by the mean lengths of the two blocks directly flanking the block containing $m$.
+
+We are interested in identifying regions of the genome with abnormal distributions of block lengths, 
+and at each spatial location $x$, compute $l_B(m,x)$ and $C(m,x)$ at a grid of positions across the genome. 
+Where indicated, we normalize $l_B(m,x)$ and $C(m,x)$ by dividing by the empirical mean of $l_B$ and $C$ across the genome for location $x$.
+While there is useful information contained in the geographic distribution of block length patterns, 
+obtaining good spatial sampling can often be difficult, 
+and this allows us to search for patterns if block lengths are only known within relatively few populations.
 
@@
 Within each deme $x$, we calculated $F_B(m,x)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
-%\plr{how about $x$ instead of $X$? above $x$ refers to spatial location.}
 
@@
 
+%%%%%%%%%%%%%%
 \section*{Results}
 
-\subsection*{Comparison of theory and simulation}
-
-The theoretical results described above matched well with our individual-based simulations,
-across a large range of parameter values.
-This can be seen by comparing
-clines in frequency of $A$ ancestry at linked loci,
-either unconditional on the identity of the linked selected allele (Figure \ref{sfig:alleleFreq_tau100_comparison}),
-or conditional on its identity 
-(Figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}).
-
-\subsection*{Single-locus behavior}
-
-From previous results, we expect that far away from the zone center, variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, and correspondingly, clines at closely linked sites are narrower.
-This is seen, for instance, in Figures~\ref{alleleFreq_tau100},~\ref{alleel_Freq_tau100_weaker_s} and ~\ref{alleleFreq_tau1000}.
-
 \plr{When discussing haplotype lengths, say what happens at different $s$ and $T$.}
 
-\paragraph{The timescale over which linked clines relax:}
+\subsection*{Single-loci: clines}
+
+As expected, clines in simulations form and begin to flatten, 
+with ancestry frequencies at the selected site maintaining a stable shape,
+and tightly linked sites flattening more slowly than distant sites.
+This is seen, for instance, in Figures \ref{alleleFreq_tau100},
+\ref{alleleFreq_tau100_weaker_s} and \ref{alleleFreq_tau1000}. 
+Clines in ancestry frequency matched expected values found from deterministic theory
+up to stochasticity due to genetic drift, which is more pronounced at lower population densities.
+Figure \ref{sfig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
+match simulated clines in Figure \ref{alleleFreq_tau100} quite well.
+Figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}
+show this comparison with a lower population density than in Figure \ref{alleleFreq_tau100},
+and separating linked clines based on the identity of the linked selected allele.
+The agreement is good over hundreds of generations, but by $T=1280$ generations,
+stochasticity has substantially moved allele frequencies rangewide
+(which is unsuprising, as the total population size is only 5,000 diploids).
+
 Although loci not under direct selection can in principle spread across the cline unimpeded,
@@
 due to the decreased fitness of heterozygotes \citep{Barton1986}
-and the relative slowness of diffusive movement \citep{Sedghifar2015}.
-We wish to know, for samples across a cline $T$ units of time after secondary contact, how far away from the selected locus one has to look to see clines that resemble the genome-wide mean.
-
-A linked cline will begin to flatten out if there is an appreciable chance that a lineage at the linked locus initially coupled with an $A$ allele 
-at some point recombined from an individual carrying the $B$ allele.
-This implies that loci closer than $1/T$ to the selected site will not have a substantially wider cline than the selected cline.
+and the relative slowness of diffusive movement.
+The spread of genotypes can be seen across space and time in Figure \ref{fig:linked_cline_heatmaps}.
+The cline in ancestry at a locus recombination distance $r$ from the selected site
+will have flattened out to distance $x$ (say, on the $B$ side)
+if there is a god chance that the corresponding lineage that begins linked to a $B$ allele
+traces back to an $A$ alelle on the opposite side of the hybrid zone.
+Since lineages linked to $B$ alleles move nearly as unbiased Brownian motion on the $B$ side,
+this is only possible if Brownian motion has had enough time to dravel distance $x$,
+i.e., if $T>\sqrt{x}$.
+This square-root flattening is seen in Figure \ref{fig:linked_cline_heatmaps}.
+A linked lineage must also spend at least $1/r$ generations in heterozygotes
+to have a good chance of recombining,
+so it is certain that clines with $r<1/T$ will still resemble the selected cline,
+which can also be seen in Figure \ref{fig:linked_cline_heatmaps}.
 
@@
 %%%%% %%%%%%%%%%
-\subsection*{The distribution of ancestry tract lengths}
-
-The distribution of contiguous ancestry block lengths necessarily contains more information than allele frequency alone.
+\subsection*{Blocks of ancestry}
+
+The distribution of contiguous ancestry block lengths contains more information than allele frequency alone.
 We are interested specifically in how the tracts of ancestry surrounding the selected locus compares to the rest of the genome.
-The raw data -- ancestry assignments for a few individuals sampled from across space --
-is shown in Figure~\ref{Fig:resistanceToIntrogression1000g},
-showing distinct patterns across geography and the genome.
+Ideal raw data -- true ancestry assignments for a few simulated individuals sampled from across space --
+are shown in Figures \ref{Fig:resistanceToIntrogression100g} ($T=100$ generations) and \ref{Fig:resistanceToIntrogression1000g} ($T=1000$ generations).
+For the more recent hybrid zone ($T=100$), the selected cline has established,
+but linked clines are still flattening.
+After a longer period ($T=1000$),
+clines over much of the chromsome are flat
+(since the width of the entire population is less than $1/\sqrt{T}=31.6\sigma$),
+but a distinct enrichment of each ancestry is observed around the selected site.
 
@@
 because they have had more opportunities to recombine with $B$ haplotypes. 
+This is seen in Figure \ref{Fig:meanBlockLengthsUnconditioned}.
 However, we expect that stretches of $A$ ancestry containing a selected site will be longer
@@
 so (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x /\sqrt{s}\sigma$ generations ago 
-(compared to $x^2/\sigma^2$ for a neutral allele),
-and so the enclosing $A$ haplotype is no longer than $\sigma\sqrt{s}/x $.
+to have last had an ancestor on the $A$ side of the cline around $x/(\sqrt{s}\sigma)$ generations ago 
+(compared to $x^2/\sigma^2$ for a neutral allele).
+This implies the enclosing $A$ haplotype should be no longer than $\sigma\sqrt{s}/x$.
 %\plr{(check predicted decay of mean length $1/x$ in simulations?)}
-For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.
 
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{{figs/results_runid_753324_chr1_start0.48_stop0.52_by0.001_blocksAlongChromNoConditioning_nonnormalized}.pdf}
+\caption{
+    \textbf{Mean enclosing block length $l(m,x)$}, across the genome (horizontal axis)
+    and at different geographic positions (different colored lines).
+    Results are from a simulation with 120 demes of 200 diploids each, selection $s=.05$, dispersal $\sigma=3$, 
+    and after $T=500$ generations.
+}\label{Fig:meanBlockLengthsUnconditioned}
+\end{figure}
+
+\begin{figure}
 \includegraphics{figs/blocksAlongChromAncBConditioning}
@@
 The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{Liang2014}.
+\alisa{We have simulated with relatively high within-deme population density ($Nd=500$), and smaller population sizes could have a large impact on the strength of signal due to increased noise.}
+
 With respect to genealogy, the broad agreement between our analytic predictions and simulations (which include an explicit pedigree structure), suggests that this assumption does not have a severe qualitative impact on our major results when populations are moderately high, especially for all but the most recently formed hybrid zones.

### 2016-03-03 drafts/hybrid_zone.tex
.

@@
 
-\plr{Literature: Haldane, Nagylaki, Slatkin, \citep{May1975,Slatkin1973}}
-\plr{Better summary of what Barton did.}
+\alisa{Better summary of what Barton did.
 Population genetic modeling of hybrid zones has focused largely on the 
-equilibrium states of clines at selected loci \citep{Barton1979a,Barton1986}.  
+equilibrium states of clines at selected loci \citep{Nagylaki1975,Barton1979a,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
 Since stronger selection reduces the amount of genetic mixing,
@@
 at loci close to selected loci, and wider clines at less tightly linked loci.
+In particular \citet{Barton1979,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
 This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
-Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.
+Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.}
 
@@
 
-\plr{When discussing haplotype lengths, say what happens at different $s$ and $T$.}
-
 \subsection*{Single-loci: clines}
@@
 
+\paragraph{Identifying selected loci}
+The statistic $C(m,x)$ shows promise for identifying selected loci under some circumstances.
 As expected, regions surrounding a locus under selection are more resistant to introgression,
 as seen in figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}).
-When present, however, we expect haplotypes including the locally less common allele
+When present, we expect haplotypes that contain the locally less common allele
 to be longer than the genome-wide average.
@@
 as shown in Figures \ref{Fig:blockLengths}, \ref{Fig:blockLengthsZoom} and \ref{Supp:blockLengthHeatmap},
-the mean length of such haplotype blocks are up to three times longer
-than the average for that geographic location.
-
-% This pattern provides a potential avenue for identifying targets of selection by looking for a signal of increased $l_B$ or $l_A$ for positions surrounding a selected site, compared to the genome-wide background  at geographic sites far from the center of the zone (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}).
-The mean haplotype length found without conditioning on ancestry ($l_\cdot(m)$)
-also shows an small increase near the selected locus,
-but it is much smaller (see Figures \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}).
-
-\alisa{For weaker selection ($s=0.001$), then the relative speed of retreat is not fast enough compared to the genomic background, and the strength of signal from elevated $l_B(m)$ is diminished.
+the mean length of such haplotype blocks is up to three times longer
+than the average for that geographic location,
+peaking quite sharply around the location of the selected site.
+
+The mean haplotype length found without conditioning on ancestry 
+shows a smaller increase near the selected locus,
+shown in Figures \ref{Fig:meanBlockLengthsUnconditioned}, \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}.
+
+\alisa{
+For weaker selection ($s=0.001$), then the relative speed of retreat is not fast enough compared to the genomic background, and the strength of signal from elevated $l_B(m)$ is diminished.
 The resolution of the peak is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
@@
 
-We note that for very strong $s$, few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may be present only at geographic locations close to the center of the zone.}
+We note that for very strong $s$, few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may be present only at geographic locations close to the center of the zone.
+}
 
@@
 
+\yaniv{remove this? or, what figure?}
 Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short.
@@
  $s$ and relatively short map distances between linked incompatibilities)
+ \yaniv{provide citation.}
 
@@
 
+\yaniv{move this earlier}
 Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.

### 2016-03-03 sims/extract-mean-length-matrix.R
.

@@ -1,0 +1,45 @@
+# run this in a directory to compile information from all *intervalSizes.Robj and *intervalSizes_allAncs.Robj files within that directory.
+
+source("sim-fns.R")
+
+B.files <- list.files(".",".*intervalSizes.Robj",recursive=TRUE,full.names=TRUE)
+all.files <- list.files(".",".*intervalSizes_allAncs.Robj",recursive=TRUE,full.names=TRUE)
+
+B.outfile <- "mean-haplotype-lengths_AncB.csv"
+all.outfile <- "mean-haplotype-lengths_AllAnc.csv"
+
+get_simparams <- function (fname) {
+    strings <- c( SIGMA="SIGMA", ninds="Ninds", ndemes="ndemes", S="s", tau="tau", run.id="runid_", start="start", stop="stop", by="by", chromosome="chr" )
+    out <- lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",fname)) } )
+    out$zone_age <- out$tau
+    out$deme_size <- out$ninds/out$ndemes
+    return(out)
+}
+
+B.lengths <- lapply( B.files, function (filename) {
+            load(filename)
+
+            # this is a (locus x deme) matrix giving mean lengths
+            #   for haplotype chunks of B ancestry
+            # mean_per_deme_AncB = 
+            do.call(rbind, lapply(intervalSizes,function(P){tapply(1:length(deme_ID),deme_ID,function(Z){all_sites = P[Z,]; mean(all_sites[which(all_sites>0)])})}))
+        } )
+
+
+all.lengths <- lapply( all.files, function (filename) {
+            load(filename)
+            # this is a (locus x deme) matrix giving mean lengths
+            #  for haplotype chunks of ALL ANCESTRIES
+            # mean_per_deme = 
+            do.call(rbind, lapply(intervalSizes_allAncs,function(P){tapply(1:length(deme_ID),deme_ID,function(Z){mean(P[Z,])})}))
+        } )
+
+positions = with( list2env(params), seq(plot_start,plot_stop,increments) )
+
+outstring = with( list2env(params), sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_dir/tau%s_dir/results_runid_%s_%s_start%g_stop%g_by%g",
+                                            SIGMA, ninds, ndemes, S, zone_age, run.id,  
[…]

### 2016-03-04 drafts/hybrid_zone.tex
.

--- 2016-03-03 drafts/hybrid_zone.tex
@@
 
-\subsection*{The number of `unfit' ancestors}
-
-\plr{This needs a bit more precise discussion.}
-
-It is possible that there is additional information contained in patterns of within-ancestry haplotype diversity throughout the hybrid zone.
-Looking at the number of ancestral haplotypes present in the population would provide information about whether the migrant contribution is through relatively few, successful migrants, or through many migrants that each contribute relatively little.
-
-We find that the average family size of a $B$ allele ($F_B(m,x)$) decrease with distance into the $A$ side of the hybrid zone, and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}).
-As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
-This pattern is expected and consistent with the prediction in our models that unfit lineages tend to be recent migrants, which will have smaller families.
-In conjunction with block length, measuring $F_B$ can provide an additional signal for characterizing lineages strongly influenced by selection.
+%%%%%%
+\subsection*{The size of migrant families}
+
+Within-ancestry haplotype diversity, i.e., the number of ancestral haplotypes of each type,
+could provide additional information about whether introgression is through relatively few, successful migrants, 
+or through many migrants that each contribute relatively little. 
+In our simulations, the average local family size of a selected $B$ allele ($F_B(m,x)$) decreases with distance into the $A$ side of the hybrid zone, 
+and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}). 
+Unlinked loci have similar local family sizes similar to neutral simulations,
+and loci linked to the selected locus have intermediate sizes.
+This pattern is consistent with the prediction in our models that unfit lineages tend to be recent migrants, 
+which will have smaller families.
 
@@
 As $m$ moves further away from the target of selection, the average family size recovers to expected values under neutrality.
-
 %One interpretation of this observation is that migrants tend to leave few offspring before being removed from the population by selection, leaving behind small families.
@@
 
-\yaniv{remove this? or, what figure?}
+\yb{remove this? or, what figure?}
 Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short.
@@
  $s$ and relatively short map distances between linked incompatibilities)
- \yaniv{provide citation.}
+ \yb{provide citation.}
 
@@
 
-\yaniv{move this earlier}
+\yb{move this earlier}
 Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.

### 2016-03-04 drafts/hybrid_zone.tex
through assumptions

@@
 
-The mean haplotype length found without conditioning on ancestry 
-shows a smaller increase near the selected locus,
-shown in Figures \ref{Fig:meanBlockLengthsUnconditioned}, \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}.
-
-\alisa{
-For weaker selection ($s=0.001$), then the relative speed of retreat is not fast enough compared to the genomic background, and the strength of signal from elevated $l_B(m)$ is diminished.
-The resolution of the peak is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
-Since clines in ancestry establish over the timescale $1/s$, it is difficult to intuit the time it takes for a signal of elevated $l_B(m)$ to form around the selected locus.
-
-We note that for very strong $s$, few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may be present only at geographic locations close to the center of the zone.
-}
-
-%\alisa{For very strong $s$, very few haplotypes are found on the ``wrong" side, and so the signal of longer haplotypes around selected loci may only be present at geographic locations close to the center of the zone.
-The overall length of the haplotypes decreases as the hybrid zone ages, but the signal remains strong relative to the rest of the genome and detectability will largely depend on the density of ancestry-informative markers.}
-
-\alisa{how small is too small for s? when do we start seeing that peak (1/tau)? Bring paragraph here. Look at smaller spatial scale for the small s simulations}
-
-Additionally, we find that $l_A(m\pm)$, the length of blocks adjacent to blocks containing $m$, are shorter than average when $m$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
+Genomes are inherited from the ancestors of $T$ generations ago
+in blocks of size roughly $1/T$,
+so there will be a long segment of $A$ ancestry
+if many such adjacent blocks are of the same ancestry.
+Ancestries of neighboring blocks are correlated,
+due to the branching process described above,
+but if we assume they are independent, 
+then since the block about location $r$ in an individual at $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
+we'd expect to see on average $2q_A(x,T,r)/(1-q_A(x,T,r))$ blocks of ancestry $A$ about a given site unlinked to the selected site.
+This implies the mean length $l(x,m)$ would be $2q_A(x,T,r)/(T(1-q_A(x,T,r)))$,
+and the mean conditional length $l_A(x,m)$ would be $2/(T(1-q_A(x,T,r))$.
+Furthermore, we know that if $x=0$ or if $T$ is large and $r$ is not too small, that $q_A$ is close to $1/2$;
+and for $r<1/T$ that $q_A$ looks like the selected cline.
+Also, we know from the discussion above that the lineage of a \emph{selected} $A$ allele,
+if it is in the region where $A$ is rare,
+moves at speed roughly $\sigma \sqrt{s}$ back towards the $A$ side of the zone,
+returning to the region where $A$ is common in about $x/(\sigma \sqrt{s})$ generations.
+Therefore, a selected $A$ allele found on the $B$ side of the zone
+should carry with it a haplotype of average length $\sigma \sqrt{s}/x$
+that looks like haplotypes from the center of the zone.
+The analysis above suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$;
+this is indeed what is seen at distant sites, for instance, in figure \ref{sfig:blocksAlongChromAncBConditioning_nonnormalized}.
+Haplotypes at the selected site are expected to be longer, but still of a length proportional to $1/T$,
+suggesting that the normalization in the statistic $C(x,m)$ is appropriate,
+as shown in Figure \ref{Supp:adjacentBlocks},
+although a numericl prediction of the value of $C(x,m)$ is elusive.
+
+The mean haplotype length found without conditioning on ancestry, $l(x,m)$,
+shows a smaller increase near the selected locus, because most blocks will be of the locally common type, and so do not trace back to regions of different block lengths.
+This is shown in Figures \ref{Fig:meanBlockLengthsUnconditioned}, \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}.
+
+Power will be optimal at intermediate values of $s$.
+If selection is too strong, it may be difficult to observe the signal
+due to a lack of introgressed selective sites,
+while if selection is week, the selected lineage does not move very fast,
+and so the strength of signal from elevated $l_B(m)$ is diminished. 
+Indeed, if $s$ is less than $x/(\sigma T)$, then only one block of ancestry $A$ is expected to be seen about a selected $A$ allele,
+and $l_A(x,m)$ is expected to be $2/T$.
+For similar reasons, power and resolution are best at intermediate $T$.
+
+% Additionally, we find that $l_A(m\pm)$, the length of blocks adjacent to blocks containing $m$, are shorter than average when $m$ is in a block of unbroken ancestry surrounding a selected site on the ``wrong side'' of the side of the zone.
 (Figure~\ref{Supp:adjacentBlocks}).
@@
 
-%%%%%%
-%\begin{figure}
- %   \begin{center}
- %   \end{center}
-%    \caption{
-  %      \textbf{Haplotype length distributions},
-  %      comparing numerical solutions to simulations.
-   %     At one time, plot 
-   %     probability of seeing a haplotype of length at least $r$ covering that position on the genome,
-    %    against (a) position on the genome, at a few spatial locations;
-   %     or (b) spatial location, at a few positions on the genome.
-   %     \plr{maybe?}
-    %    \label{fig:haplotype_lengths}
-   % }
-%\end{figure}
-
-%This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})
-
-%We are interested specifically in the distribution of the length of continuous tracts of ancestry surrounding the selected locus.
-Far away from the zone center variants under negative selection will be rare \citep[for demonstration of this theoretical result, see e.g.][]{May1975,Slatkin1973,Barton??}, however, when present an unfit migrant haplotype is expected to be longer than a neutral migrant haplotype.
-Figure X shows the estimated probability density (pdf), of unbroken tract of ancestry B surrounding the selected locus, conditioning on a haplotype being of ancestry B at the selected locus.
-Compared to the expected density of tract lengths surrounding neutral loci, tracts of ancestry B around the selected locus are longer.
-
-%Of particular interest is whether or targets of selection could potentially be identified by a signal of larger ancestry tracts surrounding them compared to a genome-wide background.
-To demonstrate this, we simulated a chromosome of length 1M with a selected site at position 0.5M.
-For a set of positions $\mathbf{m}$ along the chromosome, we computed the length $l_B(m_i)$ of the surrounding unbroken block of ancestry.
-There is a detectable signal of increased block length surrounding selected loci when conditioning on the less-fit ancestry (Figure~\ref{Fig:blockLengths},~\ref{Supp:blockLengthHeatmap}) at geographic sites far from the center of the zone.
-A signal is also present for, but less well resolved when ancestry is not taken into account (Figures~\ref{Supp:blockLengthHeatmapNoAnc},~\ref{Supp:blockLengthNoAnc}).
-
 \begin{figure}
 \includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000}
-\caption{Randomly sampled chromosomes  across a hybrid zone of age $T=1000$.
-Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
+\caption{
+    \textbf{Ancestry assignment for randomly sampled chromosomes} across a hybrid zone of age $T=1000$. 
+    Here we compare chromosomes of length 1M from a neutral zone to a zone that has a single under-dominant locus with $s=0.1$
+    in the middle of the chromosome (indicated by black arrow). 
     Red blocks along the chromosome denote ancestry $B$, and orange blocks are ancestry $A$. 
-    50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+    The simulation was performed in a population with 50 demes, 
+    each with 500 diploids, and $\sigma=1$.
+    An analogous figure at $T=100$ is shown in Figure \ref{Fig:resistanceToIntrogression100g}.
 }\label{Fig:resistanceToIntrogression1000g}
@@
 \includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{Mean block length, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
-Here each line represents  $l_B(m)$, in a given deme, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
-Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
+\caption{
+    \textbf{Normalized mean enclosing block length}, $C(m,x)$, after $T=1000$ generations,
+    against position relative to the selected locus (horizontal axis)
+    located in the center of a 1M chromosome.
+    Each line shows the mean block length at that spatial and genomic position
+    divided by the mean over the chromosome at that location;
+    the simulation was run with $s=1$ and $\sigma=1$,
+    50 demes, each containing 500 diploid individuals.
 }\label{Fig:blockLengths}
@@
 
-%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-%   which we have done using both theory and simulation.
-%   Our most useful observations are that clines establish over time $1/s$,
-%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
-%    tend to move back towards the side where they are common
-%   at speed $\sigma/\sqrt{s}$.
-%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-%}
-
+\plr{COME BACK TO THIS}
 \alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
@@
 This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
+The resolution of the peak is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
 
@@
 
-There is great interest in identifying the loci that are experiencing selection in hybrid zones.
-Popular approaches to this involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.
-Current genomic resources, however, can provide an additional rich information when applied to the study of hybrid zones.
-In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture from present day samples \citep[e.g.
-][]{Hellenthal2014}.
-In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone.
-Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
-This pattern reflects the intuition that regions surrounding selected loci introgress less readily, and loci under selection can therefore be thought as moving under Brownian motion with drift whereby a lineage that is presently on the ``wrong" side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct" side.
+Popular approaches to identifying the loci that are experiencing selection in hybrid zones 
+involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
+The availability of genomic data has made it possible to use recombination as an additional source of information.
+In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} 
+has been used extensively to infer the demographic history of hybridization/admixture from present day samples \citep[e.g. ][]{Hellenthal2014}. 
+In this paper, we have described how selection against hybrid incompatibilities 
+results in long contiguous blocks of ancestry around these loci,
+because regions surrounding selected loci introgress less readily, 
+and so introgressed alleles have been inherited from ``their'' side of the zone more recently.
+We get quantitative understanding fromt he fact that lineages of
+loci under selection moving as Brownian motion that is pushed at speed $\sigma \sqrt{s}$ towards its ancestral side.
 In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong" side, are likely to have done so relatively recently.
@@
 \yb{remove this? or, what figure?}
-Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short.
-This is measured by the mean length of the blocks adjacent to the block containing the selected site.
-A simple and intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond.
+Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location.
+An intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. 
 Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby.
-This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.
 
@@
 Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $C(k)$, could provide a simple and informative summary. 
-This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.
+The ratio of mean adjacent block lengths $C(x,m)$ shows a very sharp peak at the selected site,
+suggesting that, despite the fact that haplotypes surrounding incompatibilities might be quite long, 
+genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.
 
@@
 
-\plr{note that when drift starts to kick in for clines (around $\tau=320$) we start to get better power with haplotype lengths}
-
-%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
-%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
-%    which we have done using both theory and simulation.
-%    Our most useful observations are that clines establish over time $1/s$,
-%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
- %   tend to move back towards the side where they are common
- %   at speed $\sigma/\sqrt{s}$.
- %   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
-%}
-
-%However, how selection influences the length of tracts of ancestry around selected sites in hybrid zones has received much less attention.
-We we present an expectation of ancestry block lengths in hybrid zone.
-We find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
+Ancestry assignments have the additional benefit that relatively unbiased estimates are possible even with markers of problematic ascertainment
+such as those on SNP arrays.
+The power and resolution of these approaches depends strongly on the strength of selection,
+the time since secondary contact,
+and the strength of genetic drift:
+in our simulations,
+we found good power and resolution at $s=.01$, $T=1000$ generations after secondary contact,
+and with hundreds of individuals per dispersal distance.
 
@@
 
-We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations with respect to the architecture of selection, the geographic distribution of populations, and relatedness of individuals.
+Here we review the assumptions we made and their likely impact.
 

### 2016-03-04 drafts/hybrid_zone.tex
peter edited almost everything

@@
 are broken up by historical recombination events,
-forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003},
+forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution},
 leading to substantial correlations in ancestry (and hence, genotype) between distant markers.
-Patterns of correlated ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,Sedghifar2015}.
+Patterns of correlated ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
 
@@
 as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
-In \citet{Sedghifar2015} we described how lineages move across a neutral zone of secondary contact,
+In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
 obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
@@
 
-As in \citet{Sedghifar2015}, 
+As in \citet{sedghifar2015spatial}, 
 we gain insight by considering where the ancestors of modern-day individuals lived,
@@
 
-As in \citet{Sedghifar2015}, 
+As in \citet{sedghifar2015spatial}, 
 a given block of genome is inherited along a single lineage ever since the most recent recombination event
@@
 To compute correlations in local ancestry
-(i.e., ``ancestry disequilibrium'', as in \citet{Pool2015,Schumer2016}),
+(i.e., ``ancestry disequilibrium'', as in \citet{pool2015mosaic,Schumer2016}),
 we need only follow lineages at two sites, instead of an entire region,
@@
 
-\plr{COME BACK TO THIS}
 \alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
@@
 loci under selection moving as Brownian motion that is pushed at speed $\sigma \sqrt{s}$ towards its ancestral side.
-In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong" side, are likely to have done so relatively recently.
+In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong'' side, are likely to have done so relatively recently.
 Haplotypes surrounding the selected locus therefore have had relatively few opportunities for recombination with different ancestries and this is reflected in longer blocks of contiguous ancestry.
@@
 Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}).
-We expect, therefore, our findings here to be somewhat generalizable.
 
@@
 Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination.
-Both factors are expected to result in regions that are surrounded by even greater chunks of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
- $s$ and relatively short map distances between linked incompatibilities)
+Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
 \yb{provide citation.}
 
-\paragraph{Geographic space and other sources of stochasticity}
-For ease of calculations, our analytic model assumes a continuous and smooth one-dimensional spatial structure of the geographic landscape, allowing us to derive analytic expressions that track lineages.
-This assumes a large population density limit that excludes coalescence and makes lineages Markov.  
-In contrast, our simulations model regularly spaced demes. 
+To speed calculations and simulations, throughout we have assumed a homogeneous, one-dimensional geographical range.
+Theoretical work also assumed a large population density, 
+effectively working with a deterministic model that ignores coalescence and associated stochasticity.
+In contrast, our simulations model regularly spaced demes of finite size.
 In reality, populations may be patchily connected, especially at the edges of species ranges where hybrid zones may occur. 
-While we still believe that the genomic signature of long ancestry chunks around selected sites will still hold in these situations, the strength of this expectation is unclear.
-
-Of particular concern that stems from the above assumption is that our model ignores potential, and likely, sources of stochasticity arising from geographic structure and coalescence/pedigree structure.
-The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{Liang2014}.
-\alisa{We have simulated with relatively high within-deme population density ($Nd=500$), and smaller population sizes could have a large impact on the strength of signal due to increased noise.}
-
-With respect to genealogy, the broad agreement between our analytic predictions and simulations (which include an explicit pedigree structure), suggests that this assumption does not have a severe qualitative impact on our major results when populations are moderately high, especially for all but the most recently formed hybrid zones.
-
-\yb{move this earlier}
-Lastly, while our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms.
-This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.
+The degree to which inhomogeneous geography would affect the predictions depends on how patchy the zone is;
+the differential equations provide a way to evaluate this in specific circumstances.
+
+It is particularly unfortunate 
+that our analytic calculations do not capture stochasticity arising from coalescence/pedigree structure. 
+The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{liang2014lengths}. 
+Indeed, correlated fluctuations visible in simulations (e.g., Figure \ref{Fig:blockLengths})
+are likely due to coupling due to demographic stochasticity;
+and simulations at lower density show larger fluctuations than those at higher density.
+The agreement between analytic predictions and simulations (which include an explicit pedigree structure)
+suggests that this assumption does not have a severe impact on our major results when population densities are moderately high.
 
@@
 A main role of these solutions in our work has been to verify that theory based on the diffusion model of lineages
-matches the simulations.
+matches reality.
 These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci:
@@
 %Lastly, the analytic model presented here lacks \yb{both coalescence and a true pedigree structure}.
-\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{Liang2014}.}
+\alisa{This can result in an underestimate of covariance in ancestry, as we have ignored sharing through genealogy \cite{liang2014lengths}.}
 %The broad agreement between our analytic predictions and simulations that include an explicit pedigree structure suggests that this assumption does not have a severe qualitative impact on our major results, especially for all but the most recently formed hybrid zones.
@@
 A number of studies have described heterogeneous patterns of genetic divergence across the genome. 
-These so-called ``islands" of divergence \citep{Turner2005,Nosil2009} have been largely descriptive \citep{Cruickshank2014,Noor2009}, and as we begin to rigorously model patterns of divergence in the genome to begin to more  explore these patterns.
-The work presented here may offer ways in which to think about the process of heterogeneous divergence more concretely in terns of how processes such as migration and selection influence patterns across the genome of hybridizing populations.
-Specifically, we provide a natural interpretation of large contiguous chunks of ancestry, as sites linked to those experiencing selection.
-Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection rather than high-level summaries that are somewhat abstracted from the evolutionary process.
+Work on these ``islands of divergence'' \citep{turner2005genomic,Nosil2009} and related patterns have been largely descriptive \citep{Cruickshank2014,Noor2009}.
+This paper contributes to a model-based understanding of 
+how migration and selection influence such patterns across the genome of hybridizing populations. 
+Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection 
+rather than high-level summaries that are somewhat abstracted from the evolutionary process.
 
@@
 %The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{Turner2005, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{turner2005genomic, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
 %REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
@@
 
-%\subsection*{Future directions}
-
-%\plr{Perhaps delete this paragraph? A small note saying should check power with simulations up where we discuss the statistic would do.}
-%Our work has highlighted distinct genomic signatures accompanying hybrid incompatibilities. 
-%We suggest that such signatures could be combined in an inference framework that aims to identify incompatibilities from genome scans of hybrid zones.  
-%However, doing so requires careful evaluation of potential signals of an incompatibility. 
-%For example, should we consider jointly lengths of ancestry and allele frequencies to identify incompatibilities?
-%Will dividing haplotype length by the length of adjacent haplotypes add power to tests because of the extra information contained in this measure, or add noise because of chance variation in these statistics.
-%Validation of the effectiveness of our proposed statistics, especially under demographic models that differ from the one dimensional hybrid zone explored above, will  provide a guide on how our results can be used to learn about selection from the distribution of ancestry tract lengths.
-We note, however, that power to detect regions under selection will depend on a number of parameters, such as $s$ and $T$
-
 \subsection*{Adaptive introgression}
@@
 
-We have demonstrated here the potential utility for block length distributions in hybrid zone populations, particularly in populations that are found some distance from the center of the zone.
-While the focus here, has been on hybrid incompatibilities, the possibility that some loci adaptively introgress across hybrid zones has excited many biologists \citep{Arnold2004}.
+While our focus has been on hybrid incompatibilities, 
+unconditionally adaptive loci are also known to introgress across hybrid zones \citep{Barton1986,martinsen2001hybrid,arnold2004transfer}. 
 Future work could take a similar approach to understand how positive selection shapes ancestry block lengths, and predict  signatures of adaptive introgression in hybridizing populations using similar statistics presented here.
 These could eventually be combined to gain a fuller understanding of the forces shaping patterns of introgression in hybrid zones.
+In particular, beneficial alleles tightly linked to incompatibilities
+cannot introgress until recombination separates them;
+our model describes at least roughly how quickly this should happen.
 
@@
 %Doing so requires an understanding of how positive selection shapes ancestry block lengths in these cases.
-
-%\subsection*{Simulation and analysis code}
-%These are written primarily in R, and are available HERE.
 
@@
 One method for solution would be via a Wild sum over the number of recombination events, 
-as we did in \citet{Sedghifar2015}.
+as we did in \citet{sedghifar2015spatial}.
 Here, we solved the equations numerically,

### 2016-03-05 drafts/hybrid_zone.tex

@@
 Hybrid zones formed between recently diverged populations offer an opportunity 
-to better understand the mechanisms underlying reproductive isolation and the process of speciation. 
+to study the mechanisms underlying reproductive isolation and the process of speciation. % too much `understanding' here
 Here, we use a combination of analytical theory and explicit forward simulations
-to gain a better understanding of how selection against hybrid gentypes impacts patterns of introgression 
+to show how selection against hybrid genotypes impacts patterns of introgression  % too much `understanding' here
 across genomic and geographic space. 
@@
 Locally disadvantageous alleles tend to exist as small families,
-whose lineages trace back to the side they originated on at speed $\sqrt{s}$ dispersal distances per generation.
+whose lineages trace back to the side from which they originated at speed $\sqrt{s}$ dispersal distances per generation.
 The lengths of continuous tracts of ancestry provide an additional source of information:
@@
 
-The process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
-Gene flow between formerly isolated populations is possible if reproductive isolation is not complete.
-As two such populations interbreed, each portion of each individual's genome 
-can be labeled according to which ancestral population is is inherited from,
-and local interbreeding and migration of such populations
-creates a gradient of ancestry assignments across geographic space,
+Much of the process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
+% Gene flow between formerly isolated populations is possible if reproductive isolation is not complete.
+However, when formerly isolated populations come into contact before reproductive isolation is complete, some gene flow is possible.
+%As two such populations interbreed , each portion of each individual's genome 
+%can be labeled according to which ancestral population is is inherited from, and local 
+Interbreeding and migration between such populations  
+% YB It seems like the fact that such ancestry can be labeled is more important later, but probably shouldn't be a focus here. 
+creates a gradient of ancestry across geographic space, %YB removed `assignments' as it seems like it is the ancestry, not its assignment, that is of interest
 centered on their point of contact \citep[reviewed in][]{Barton1985}. 
@@
 are broken up by historical recombination events,
-forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution},
-leading to substantial correlations in ancestry (and hence, genotype) between distant markers.
-Patterns of correlated ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
+forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution}. 
+Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
+patterns of correlated ancestry in admixed populations has been previously described the history of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
 

### 2016-03-05 drafts/supplement.tex
peter edited 2/3 through

@@
 % \begin{document}
-% \setcounter{table}{0}
-% \renewcommand{\thetable}{S\arabic{table}}
-% \setcounter{figure}{0}
-% \renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
         Each line represents a locus some distance $r$ away from the true target of selection, 
-        and $r=0$ represents the locus that is under selection.    }\label{sfig:alleleFreq_tau100_comparison}
+        and $r=0$ represents the locus that is under selection.    
+    }\label{sfig:alleleFreq_tau100_comparison}
 \end{figure}
@@
  }\label{Fig:resistanceToIntrogression100g}
+\end{figure}
+
+\begin{figure}
+    \includegraphics{{figs/simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_blocksAlongChromAncBConditioning_nonnormalized}.pdf}
+    \caption{
+        \textbf{Mean haplotype lengths of $A$ haplotypes}, $l(x,m)$,
+        across the genome (horizontal axis) and at differntial spatial locations (colored lines),
+        from a simulation with 50 demes having 500 individuals each, $s=0.01$, $\sigma=1$, and after $T=1000$ generations.
+    } \label{sfig:blocksAlongChromAncBConditioning_nonnormalized}
 \end{figure}

### 2016-03-05 drafts/hybrid_zone.tex
YB edited the entire intro

@@
 
-The arrangement of segments of genome inheriting from particular ancestral populations
+The arrangement of segments of genome inherited from particular ancestral populations
 provides information about the number of recombination events that have taken place between ancestral genotypes, 
-and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+and therefore contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. % too much `provides information'
 In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, 
-this could be especially important for understanding the impact of selection on the mixing of ancestral genomes and, consequently, the maintenance of species distinctness. 
+this could be especially important for understanding the impact of selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
 In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, some amount of introgression can occur, providing information about the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
-The resulting heterogeneous patterns of ancestry across the genome have been used to identify putative targets of selection in hybridizing populations,
-by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}.
-
-Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
+If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has be used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+In addition to facilitating demographic inference, heterogeneous patterns of ancestry across the genome can help identify putative targets of selection in hybridizing populations  \citep{Porter1997, Gompert2012}.
+
+If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
 We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.

### 2016-03-05 drafts/hybrid_zone.tex
YB edited through Lineages at the selected locus

@@
 blocks of ancestry surrounding single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location. 
-The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.
+Our findings suggest that the density of ancestry junctions provide a means for characterizing hybrid zones and identifying candidate targets of selection.
 %\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
@@
 
-As in \citet{sedghifar2015spatial}, 
-we gain insight by considering where the ancestors of modern-day individuals lived,
+As in \citet{sedghifar2015spatial}, consider where the ancestors of modern-day individuals lived,
 i.e., where their lineages fall across geography as we look further back towards the time of first contact.
@@
 A block of genome is of ancestry $A$ if every locus in it does the same; 
-recombination events may cause us to follow more than one lineage back to determine the block's ancestry,
-and the block will be of ancestry $A$ if all lineages trace back to the $A$ side.
+	this occurs if there is no recombination in this block, or if all lineages generated by recombination events 
+	in this block trace back to the $A$ side. 
 This process, in which recombination events cause such lineages to \emph{branch},
@@
 if a block of ancestry $A$ includes the selected site then it is necessarily on the $A$ background.
-As the identity of the selected allele determines how selection acts on the haplotype,
-and linked variation can only move between the $A$ and the $B$ background in heterozygotes,
+Because the identity of the selected allele determines how selection acts on the haplotype,
+and because linked variation can only move between the $A$ and the $B$ background in heterozygotes, 
 a key factor in these models is the density and fecundity of heterozygotes.

### 2016-03-05 drafts/hybrid_zone.tex
YB edited through haplotype length. major changes in P1 of Lineages at selected Loci

@@
 Now suppose that the frequency profile of the selected allele, $p(x,t)$, is known.
-We seek to first understand how lineages of sampled individuals behave at the selected locus,
-from which lineages at linked loci require only a minor modification.
+We seek to first understand how lineages of sampled individuals behave at the selected locus. 
+These results allow us to extend our analysis to lineages at loci linked to the selected locus.
 
@@
 % $$ = \E[ R ( p(x) + R p'(x) ) ]/ p(x) + O(R^3) $$
-Using that $\E[R]=0$ and $\E[R^2]=\sigma^2$
+Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$
 and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
@@
 back towards the side where they are more common. 
-Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
+Since an $A$ allele must, by definition, have been inherited from the $A$ side of the barrier 
 at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
@@
 \paragraph{Lineages at linked loci}
-The behavior of a lineage at a linked locus is similar to a selected locus,
-except that when in heterozygotes the lineage may move between backgrounds:
-if we follow back through time a lineage at a locus we find linked to an $A$ allele, 
-it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
-but if it spends time in the hybrid zone, where recombination between selected backgrounds may happen,
-it may be found to have been inherited from a $B$-carrying individual,
+%YB DID SUBSTANTIAL EDITING OF THIS PARAGRAPH
+The behavior of a lineage at a linked locus is similar to a selected locus. 
+However, there is one important difference -- in heterozygotes the lineage lineage linked to the selected locus may move between backgrounds. 
+Therefore, if we follow back through time a lineage at a locus linked to an $A$ allele, 
+it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left). 
+However, with sufficient time in the hybrid zone, recombination allows this linked locus 
+to have been inherited from a $B$-carrying individual,
 whose ancestors will tend to be more from the right.
 
-Suppose we sample an allele today at a locus at recombination distance $r$ from the selected site.
+Suppose we sample an allele today $r$ Morgans from the selected site. 
 We will follow its lineage back through time, using $\tau$ to denote ``generations ago'' to avoid confusion.
@@
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ 
-(just as previously described for a selected allele),
+(as described for a selected allele),
 and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$).

### 2016-03-05 drafts/hybrid_zone.tex
YB edited up to  Simulations

@@
 \paragraph{Haplotype lengths}
-We can also find the probability of seeing an \emph{entire} block of genome (haplotype)
+We now find the probability of maintaining an \emph{entire} block of genome (haplotype) % thought `seeing' was too much about our inference rather than the the biological process
 of a single ancestry at a given location and time.
@@
 of finding an entire segment $(a,b)$ of ancestry $A$ in an individual having selected allele of type $z$,
-sampled at spatial location $x$ at time $t$ after the cline has formed.
+sampled at spatial location $x$ at time $t$ after the initiation of gene flow. % I think people sometimes use 'cline formed' to mean when the equilibrium was established, changed for clarity
 
@@
 Additionally, at rate equal to the genetic length of the segment,
-a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location between $a$ and $b$,
+recombination occurs, at which point the segment splits in two at a uniformly chosen location between $a$ and $b$,
 each of which proceeds as before, independently. 
-Then, $g_z(x,T;a,b)$ is then the probability that all branches are found on the $A$ side of the hybrid zone
+$g_z(x,T;a,b)$ is then the probability that all branches are found on the $A$ side of the hybrid zone
 at the time of secondary contact $T$ units of time ago.
@@
 the equation for haplotype identity probabilities on a segment $(a,b)$ depends only on those probabilities for segments contained in $(a,b)$.
-This allows for efficient numerical solutions,
+This allows for numerical solutions, %`efficient` may be an overstatement
 described in more detail in Appendix \ref{apx:haplotype_calcs}.
@@
 (i.e., ``ancestry disequilibrium'', as in \citet{pool2015mosaic,Schumer2016}),
-we need only follow lineages at two sites, instead of an entire region,
-the model described above requires only minor modifications,
-and our numerical code can be used directly;
-for more detail see Appendix~\ref{apx:haplotype_calcs}.
+we need only follow lineages at two sites, instead of an entire region. 
+Doing so requires only minor modifications of the model, above. 
+Moreover, direct use of our numerical code provides a measure of the correlation in ancestry between markers. 
+See Appendix~\ref{apx:haplotype_calcs} for more detail.
 

### 2016-03-05 drafts/hybrid_zone.tex
YB edited up to Blocks of Ancestry

@@
 
-We implemented forwards-time simulations of diploid individuals on a one-dimensional grid of demes
+We implemented forwards-time simulations of a one-dimensional grid of demes
 with non-overlapping generations and fixed population sizes (a Wright-Fisher model).
 Individuals are diploid, with haploid number $n=2$ chromosomes, each of length 1 Morgan.
-One chromosome pair harbors in its center (at position 0.5M) a single locus that reduces fitness by $s$ in heterozygotes, while the other contains no sites under direct selection.
+One chromosome pair harbors, at position 0.5M, a single locus that reduces fitness by $s$ in heterozygotes, while the other contains no sites under direct selection.
 
 Each deme has exactly $N_d$ diploid hermaphroditic individuals at the start of each generation.
-At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
+Then, every individual can migrate to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
 where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
@@
 and combining the resulting gametes such that each pair of parents leaves one descendant.
-Meiosis results in alternating blocks of the gamete's chromsome
+Meiosis results in alternating blocks of the gamete's chromosome
 being inherited from the two parental chromosomes,
@@
 and the order of the parental chromosomes chosen randomly.
-The simulation software works by recording for each chromosome
-a list of ancestry breakpoints and the index of the ancestor in the first generation
-that the chromosome inherited that region of genome from;
-ancestry assignment at individual loci was performed by looking up the ancestor's genotype.
+The simulation software works by recording, for each chromosome, 
+a list of ancestry breakpoints, and the index of the ancestor from which 
+the chromosome inherited that genomic region.
+We the assigned ancestry at individual loci by looking up the ancestor's genotype.
 
@@
 
-The above theory and simulations allow us to make predictions about patterns surrounding selected loci. 
+The above theory and simulations generate predictions about patterns of ancestry surrounding selected loci. 
 In reality, however, such loci are not usually known, so it is useful to have per-site statistics that may allow for detection of candidate targets of selection. 
@@
 The agreement is good over hundreds of generations, but by $T=1280$ generations,
-stochasticity has substantially moved allele frequencies rangewide
-(which is unsuprising, as the total population size is only 5,000 diploids).
-
-Although loci not under direct selection can in principle spread across the cline unimpeded,
-in practice it can take quite some time for even unlinked neutral loci to homogenize,
+stochasticity has substantially moved allele frequencies range-wide
+(which is unsurprising, as the total population size is only 5,000 diploids).
+
+Loci not under direct selection can in principle spread across the cline unimpeded. 
+In practice, however, it can take quite some time for even unlinked neutral loci to homogenize,
 due to the decreased fitness of heterozygotes \citep{Barton1986}
 and the relative slowness of diffusive movement.
-The spread of genotypes can be seen across space and time in Figure \ref{fig:linked_cline_heatmaps}.
-The cline in ancestry at a locus recombination distance $r$ from the selected site
+We display the spread of ancestry across space and time in Figure \ref{fig:linked_cline_heatmaps}. 
+The cline in ancestry at a locus $r$ Morgans from the selected site
 will have flattened out to distance $x$ (say, on the $B$ side)
-if there is a god chance that the corresponding lineage that begins linked to a $B$ allele
-traces back to an $A$ alelle on the opposite side of the hybrid zone.
+if there is a good chance that the corresponding lineage that begins linked to a $B$ allele
+traces back to an $A$ allele on the opposite side of the hybrid zone.
 Since lineages linked to $B$ alleles move nearly as unbiased Brownian motion on the $B$ side,
-this is only possible if Brownian motion has had enough time to dravel distance $x$,
+this is only possible if Brownian motion has had enough time to travel distance $x$,
 i.e., if $T>\sqrt{x}$.

### 2016-03-05 drafts/hybrid_zone.tex
YB edited up to Discussion

@@
 The distribution of contiguous ancestry block lengths contains more information than allele frequency alone.
-We are interested specifically in how the tracts of ancestry surrounding the selected locus compares to the rest of the genome.
-Ideal raw data -- true ancestry assignments for a few simulated individuals sampled from across space --
+We are specifically interested in how the tracts of ancestry surrounding the selected locus compares to the rest of the genome.
+Ideal information -- true ancestry assignments for a few simulated individuals sampled from across space -- % YB changed  `raw-data' to information, because ancestry will never be raw-data -- ATGC will
 are shown in Figures \ref{Fig:resistanceToIntrogression100g} ($T=100$ generations) and \ref{Fig:resistanceToIntrogression1000g} ($T=1000$ generations).
@@
 After a longer period ($T=1000$),
-clines over much of the chromsome are flat
+clines over much of the chromosome are flat
 (since the width of the entire population is less than $1/\sqrt{T}=31.6\sigma$),
@@
 As expected, regions surrounding a locus under selection are more resistant to introgression,
-as seen in figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}).
+as seen in Figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}).
 When present, we expect haplotypes that contain the locally less common allele
@@
 
-Genomes are inherited from the ancestors of $T$ generations ago
+Genomic regions are inherited from ancestors $T$ generations ago
 in blocks of size roughly $1/T$,
@@
 Ancestries of neighboring blocks are correlated,
-due to the branching process described above,
-but if we assume they are independent, 
-then since the block about location $r$ in an individual at $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
-we'd expect to see on average $2q_A(x,T,r)/(1-q_A(x,T,r))$ blocks of ancestry $A$ about a given site unlinked to the selected site.
+due to the branching process described above. 
+But, if we assume they are independent, 
+then since the block about location $r$ in an individual at geographic location $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
+we'd expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ blocks of ancestry $A$ about a given site unlinked to the selected site.
 This implies the mean length $l(x,m)$ would be $2q_A(x,T,r)/(T(1-q_A(x,T,r)))$,
@@
 that looks like haplotypes from the center of the zone.
-The analysis above suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$;
-this is indeed what is seen at distant sites, for instance, in figure \ref{sfig:blocksAlongChromAncBConditioning_nonnormalized}.
+This analysis suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$;
+this is indeed what is seen at distant sites, for instance, in Figure \ref{sfig:blocksAlongChromAncBConditioning_nonnormalized}.
 Haplotypes at the selected site are expected to be longer, but still of a length proportional to $1/T$,
@@
 as shown in Figure \ref{Supp:adjacentBlocks},
-although a numericl prediction of the value of $C(x,m)$ is elusive.
+although a numerical prediction of the value of $C(x,m)$ is elusive.
 
@@
 due to a lack of introgressed selective sites,
-while if selection is week, the selected lineage does not move very fast,
+while if selection is weak, the selected lineage does not move very fast,
 and so the strength of signal from elevated $l_B(m)$ is diminished. 
@@
 and is lowest far away from the zone center, where ancestry $B$ is at low frequency (Figure~\ref{Fig:family_size}). 
-Unlinked loci have similar local family sizes similar to neutral simulations,
+Unlinked loci have local family sizes similar to neutral simulations,
 and loci linked to the selected locus have intermediate sizes.
-This pattern is consistent with the prediction in our models that unfit lineages tend to be recent migrants, 
+This pattern is consistent with our theoretical results showing that unfit lineages tend to be recent migrants, 
 which will have smaller families.

### 2016-03-05 drafts/hybrid_zone.tex
YB finished editing

@@
 This occurs at speed $\sigma\sqrt{s}$.
-Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline.
-This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}
-The resolution of the peak is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
+Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline, than those surrounding neutral loci.
+This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency. 
+Additionally, our work suggests that the ancestry block length distribution can help detect targets of purifying selection in hybrid zones.}  
+The resolution of this approach is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
 
@@
 
-Popular approaches to identifying the loci that are experiencing selection in hybrid zones 
+Popular approaches to identifying loci under selection in hybrid zones 
 involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
-The availability of genomic data has made it possible to use recombination as an additional source of information.
+The availability of genomic data has made it possible to use local ancestry as an additional source of information. 
 In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} 
-has been used extensively to infer the demographic history of hybridization/admixture from present day samples \citep[e.g. ][]{Hellenthal2014}. 
-In this paper, we have described how selection against hybrid incompatibilities 
-results in long contiguous blocks of ancestry around these loci,
-because regions surrounding selected loci introgress less readily, 
-and so introgressed alleles have been inherited from ``their'' side of the zone more recently.
-We get quantitative understanding fromt he fact that lineages of
-loci under selection moving as Brownian motion that is pushed at speed $\sigma \sqrt{s}$ towards its ancestral side.
+can inform the demographic history of hybridization/admixture from present day samples \citep[e.g. ][]{Hellenthal2014}. 
+We described how selection against hybrid incompatibilities  
+results in long contiguous blocks of ancestry around these loci. 
+This is because regions surrounding selected loci do not readily introgress, 
+and so introgressed alleles have been inherited from ``their'' side of the zone recently. %removed `more and less bc we are not comparing anything in this sentence'
+We get quantitative understanding from the fact that lineages of
+loci under selection move as Brownian motion pushed at speed $\sigma \sqrt{s}$ towards its ancestral side.
 In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong'' side, are likely to have done so relatively recently.
@@
 
-While these statistics presented here are currently detectable only in the few systems for which dense genetic markers and phase information is known, as the cost of sequencing continues to drop and long-phased reads become more common, the patterns described here could further aid in the identification of selected sites in hybrid zones.
-In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available low-density and unphased genomic data, could offer a less-powerful alternative to these findings.
+% many modifications here
+\yb{Our results focus on the length distribution of ancestry blocks --- a statistic that can only be obtained in the few systems for which dense genetic markers and phase information is known. 
+However, as the cost of sequencing continues to drop and long-phased reads become more common, the patterns described here could further aid in the identification of selected sites in hybrid zones in many non-model taxa.
+In the meantime, patterns of elevated pairwise LD, which is easily computed from readily available low-density and unphased genomic data, could offer an alternative path forward for empirical work. 
+IImportantly, the numerical solutions derived above readily and efficiently predict correlations in ancestry, and so our results can be applied immediately to genotype data for any taxa with markers placed on a genetic map.}
 
@@
 Our assumption of a single selected locus may have greater consequences.
-While our model is likely to be relevant to cases in which there are only a small number of targets of selection scattered throughout the genome, our predictions may differ significantly from situations in which the density of selected sites is higher.
+Our model is most relevant to scenarios with few targets of selection scattered throughout the genome, and therefore, our predictions may differ significantly from situations in which the density of selected sites is higher.
 Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination.
@@
 
-To speed calculations and simulations, throughout we have assumed a homogeneous, one-dimensional geographical range.
-Theoretical work also assumed a large population density, 
+To speed calculations and simulations, we assumed a homogeneous, one-dimensional geographical range.
+Our analytical results further assumed a large population density, 
 effectively working with a deterministic model that ignores coalescence and associated stochasticity.
@@
 
-It is particularly unfortunate 
-that our analytic calculations do not capture stochasticity arising from coalescence/pedigree structure. 
-The latter can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{liang2014lengths}. 
+%It is particularly unfortunate 
+%that our analytic calculations do not capture stochasticity arising from coalescence/pedigree structure. 
+Extending our analytical results to capture stochasticity arising from coalescence/pedigree structure represents an important future direction.  
+Ignoring pedigree structure can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{liang2014lengths}. 
 Indeed, correlated fluctuations visible in simulations (e.g., Figure \ref{Fig:blockLengths})
@@
 \subsection*{Theory and simulation}
-In this work, we have taken two complementary approaches,
+We have taken two complementary approaches,
 using both simulation and theory, and comparing the two.
@@
 More work could be done to develop more efficient methods of solution, 
-but it may be better to perform more biologically realistic forwards-time simulations that includes coalescence and drift.
-Nonetheless, the PDE approach is easily modified to provide predictions for spatial and temporally inhomogeneous systems --
-for instance, across maps of real landscapes.
+but it may be better to perform more biologically realistic forwards-time simulations that includes coalescence and drift to better characterize the block length distribution.
+If however the correlation in ancestry, rather than the block-length distribution, is of interest the PDE approach may be preferable because it is easily modified to provide predictions for spatial and temporally inhomogeneous systems -- for instance, across maps of real landscapes.
 
@@
 \section{Numerical calculation of haplotype probabilities}
-\label{apx:haplotype_calcs}.
+\label{apx:haplotype_calcs}
 

### 2016-03-05 drafts/hybrid_zone.tex
YB minor change to legend of Fig4
minor edits

@@
 Clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
-and linked clines persist over a  physical scale of $1/T$, where $T$ is the age of the hybrid zone. 
+and linked clines persist over a  physical scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. 
 Locally disadvantageous alleles tend to exist as small families,
@@
 
-If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions.
+If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression of surrounding genomic regions.
 We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
@@
 stochasticity has substantially moved allele frequencies range-wide
-(which is unsurprising, as the total population size is only 5,000 diploids).
+(which is unsurprising, as the total population size is only 5,000 diploids) \citep{Polechova2011}.
 
@@
 then since the block about location $r$ in an individual at geographic location $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
-we'd expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ blocks of ancestry $A$ about a given site unlinked to the selected site.
+we'd expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ \alisa{consecutive} blocks of ancestry $A$ about a given site unlinked to the selected site.
 This implies the mean length $l(x,m)$ would be $2q_A(x,T,r)/(T(1-q_A(x,T,r)))$,
@@
 that looks like haplotypes from the center of the zone.
-This analysis suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$;
+This analysis suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$ \alisa{should this be $2/T$?};
 this is indeed what is seen at distant sites, for instance, in Figure \ref{sfig:blocksAlongChromAncBConditioning_nonnormalized}.
@@
 \caption{
-    \textbf{Ancestry assignment for randomly sampled chromosomes} across a hybrid zone of age $T=1000$. 
+    \textbf{Ancestry blocks for randomly sampled chromosomes} across a hybrid zone of age $T=1000$.
+% YB changed from `assignment' to `blocks' because we are interested in the blocks not the assignment of them...
     Here we compare chromosomes of length 1M from a neutral zone to a zone that has a single under-dominant locus with $s=0.1$
@@
 In the meantime, patterns of elevated pairwise LD, which is easily computed from readily available low-density and unphased genomic data, could offer an alternative path forward for empirical work. 
-IImportantly, the numerical solutions derived above readily and efficiently predict correlations in ancestry, and so our results can be applied immediately to genotype data for any taxa with markers placed on a genetic map.}
+Importantly, the numerical solutions derived above readily and efficiently predict correlations in ancestry, and so our results can be applied immediately to genotype data for any taxa with markers placed on a genetic map.}
 

### 2016-03-05 drafts/hybrid_zone.tex
notes on refs

@@
 
+%%%
+%% these deal with environmental variation in fitness
+% Haldane1948 : does 1D stable cline including dominance for step fitness gradient
+% Fisher1950 : continuous gradient in selection
+% Hansen1966 : does numerical solns to discretized ODE including observing minimal patch size
+% Nagylaki1975 : writes down equations in 2D and conditions for stable solns
+%%%
+%% these deal with hybrid zones
+% Bazykin1969 : writes down the stable solution for the one-dimensional cline
+% barton1979dynamics of hybrid zones: looks at how hybrid zones move using a potential; also looks at the strength of stochastic effects. seems to NOT say how fast they relax?
+% barton1979gene flow past a cline : step in frequency of linked neutral allele 
+%    proportional to gradient in frequency and inverse to recomb rate
+%    adaptive introgression occurs quickly 
+% barton1986 and bengtsson : looks at many linked incompatibilities
+%    with barrier B, neutral allele delayed by (B/sigma)^2, 
+%    while adaptively introgressing allele delayed only if (B/sigma)^2 > 1/s.
+%    also does size of step of allele frequency
+% barton2008effect of a barrier to gene flow : on fluctuations in allele frequency, using diffusion approximation; 
+%    finds is more sensitive in 2d than 1d
+%%%
+%% this deals with both:
+% Slatkin1973 : does environmental cline and minimal width of a patch;
+%    shows discontinuous freqs at barriers to dispersal
+%    AND observes that het disadvantage leads to steeper clines,
+%    so maybe environment determines location of cline; het disadvantage the shape
+% May1975 : spatial resolution of adaptation; also does het disadvantage
 \alisa{Better summary of what Barton did.
@@
 as derived in \citet{Bazykin1969}; for discussion of diploidy see \citet{christiansen1995genotypic}.
-Although there is no known exact analytic solution to this equation, even in steady state, 
-a good approximation to the equilibrium solution is that
-$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
+Although there is no known exact analytic solution to this equation, 
+the steady-state solution is
+$\lim_{t \to \infty} p(x,t) = (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
 This relies on several approximations, but the general conclusions should be quite robust: 

### 2016-03-07 drafts/hybrid_zone.tex
lit review notes
peter added paragraph

--- 2016-03-05 drafts/hybrid_zone.tex
@@
 centered on their point of contact \citep[reviewed in][]{Barton1985}. 
-%\plr{(check that neutral clines discussed in barton '85)}
 If the populations are sufficiently diverged, 
@@
 
-\alisa{Better summary of what Barton did.
-Population genetic modeling of hybrid zones has focused largely on the 
-equilibrium states of clines at selected loci \citep{Nagylaki1975,Barton1979a,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
-Since stronger selection reduces the amount of genetic mixing,
-these models predict the narrowest clines 
-(i.e., the most rapid changes in allele frequencies)
-at loci close to selected loci, and wider clines at less tightly linked loci.
-In particular \citet{Barton1979,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
-This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
-Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.}
+There is substantial theoretical work on genetic clines.
+We study those maintained \emph{only} by selection against hybrids -- \emph{tension zones} --
+as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
+Clines may also be maintained by environmental gradients \citep{Haldane1948,Fisher1950};
+although these are in principle different,
+\citet{Slatkin1973} pointed out (among many other things)
+that while the environment may determine the location of a cline,
+the fitness of the heterozygote will likely determine the shape of the cline,
+and \citet{Barton1986} show that the shapes of several sorts of stable clines are similar.
+There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
+and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,barton1989adaptation,Barton1993,Barton2000}.
+In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
+in particular describing haplotype patterns.
+Although the selected cline establishes quickly, transient patterns can persist for quite some time,
+and haplotypes can potentially carry substantial information about the formation of recent zones.
+To do this, we take the reverse-time perspective,
+understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
+This is particularly suited to understanding transient patterns,
+although because of the well-known problems with spatial coalescence \citep{felsenstein1978pain,barton2002spatial},
+incorporating genetic drift is not straightforward.
+Common to nearly all studies is the use of the
+reaction-diffusion equations that govern the deterministic, high-density limit \citep[][as in]{Nagylaki1975},
+or the discrete analogues \citep{Hansen1966}.
+
+% \alisa{
+% Population genetic modeling of hybrid zones has focused largely on the 
+% equilibrium states of clines at selected loci \citep{Nagylaki1975,Barton1979a,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
+% Since stronger selection reduces the amount of genetic mixing,
+% these models predict the narrowest clines 
+% (i.e., the most rapid changes in allele frequencies)
+% at loci close to selected loci, and wider clines at less tightly linked loci.
+% In particular \citet{Barton1979,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
+% This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
+% Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.}
 

### 2016-03-07 drafts/hybrid_zone.tex
added refs

@@
 as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
-Clines may also be maintained by environmental gradients \citep{Haldane1948,Fisher1950};
+Clines may also be maintained by environmental gradients \citep{Haldane1948,fisher1950frequencies};
 although these are in principle different,
@@
 There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
-and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,barton1989adaptation,Barton1993,Barton2000}.
+and quantifying the barrier to gene flow caused by clines \citep{Barton1979,Barton1989,Barton1993,Barton2000}.
 In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
@@
 This is particularly suited to understanding transient patterns,
-although because of the well-known problems with spatial coalescence \citep{felsenstein1978pain,barton2002spatial},
+although because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002neutral},
 incorporating genetic drift is not straightforward.
@@
 reaction-diffusion equations that govern the deterministic, high-density limit \citep[][as in]{Nagylaki1975},
-or the discrete analogues \citep{Hansen1966}.
+or the discrete analogues \citep{hanson1966effects}.
 

### 2016-03-07 drafts/hybrid_zone.tex
minor tweaks

@@
 % YB It seems like the fact that such ancestry can be labeled is more important later, but probably shouldn't be a focus here. 
-creates a gradient of ancestry across geographic space, %YB removed `assignments' as it seems like it is the ancestry, not its assignment, that is of interest
+creates a gradient of alleles derived from the two populations across geographic space, %YB removed `assignments' as it seems like it is the ancestry, not its assignment, that is of interest
 centered on their point of contact \citep[reviewed in][]{Barton1985}. 
@@
 are broken up by historical recombination events,
-forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution}. 
-Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
-patterns of correlated ancestry in admixed populations has been previously described the history of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
+forming chromosomal {junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution}. 
+% Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
+These
+patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
 
@@
 both geographic clines in ancestry and strong correlations in ancestry across linked markers, 
-subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
+subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{barton1979gene}. 
+% previously had wrong Barton '79 here
 Studies of hybrid zones have primarily been based on single-site statistics such as clines,
@@
 There is substantial theoretical work on genetic clines.
-We study those maintained \emph{only} by selection against hybrids -- \emph{tension zones} --
+We study those maintained only by selection against hybrids -- \emph{tension zones} --
 as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
@@
 There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
-and quantifying the barrier to gene flow caused by clines \citep{Barton1979,Barton1989,Barton1993,Barton2000}.
+and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
 In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
@@
 Common to nearly all studies is the use of the
-reaction-diffusion equations that govern the deterministic, high-density limit \citep[][as in]{Nagylaki1975},
+reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
 or the discrete analogues \citep{hanson1966effects}.
@@
 % Population genetic modeling of hybrid zones has focused largely on the 
-% equilibrium states of clines at selected loci \citep{Nagylaki1975,Barton1979a,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
+% equilibrium states of clines at selected loci \citep{Nagylaki1975,barton1979dynamics,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
 % Since stronger selection reduces the amount of genetic mixing,
@@
 % at loci close to selected loci, and wider clines at less tightly linked loci.
-% In particular \citet{Barton1979,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
+% In particular \citet{barton1979gene,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
 % This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
@@
 Because the identity of the selected allele determines how selection acts on the haplotype,
-and because linked variation can only move between the $A$ and the $B$ background in heterozygotes, 
+and because linked alleles can only move between the $A$ and the $B$ background in heterozygotes, 
 a key factor in these models is the density and fecundity of heterozygotes.
@@
 Deterministic theory predicts that after secondary contact,
-the two alleles at the selected locus will form a stable cline \citep{Barton1979},
+the two alleles at the selected locus will form a stable cline \citep{barton1979gene},
 affecting neighboring loci as well.
@@
 Assuming that: (i) alleles locally assort into diploids randomly, 
-(ii) habitat and dispersal are homogeneous, and 
-(iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
+(ii) habitat and dispersal are homogeneous,
+(iii) $s$ and $\sigma$ are small, 
+and (iv) population density is large,
+then the commonly-used equation that $p$ solves is
 \begin{align} \label{eqn:cline_pde}
@@
 where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as derived in \citet{Bazykin1969}; for discussion of diploidy see \citet{christiansen1995genotypic}.
+as in \citet{Bazykin1969}; 
+this can be derived as in \citet{Nagylaki1975};
+for discussion of diploidy see \citet{christiansen1995genotypic}.
 Although there is no known exact analytic solution to this equation, 
@@
 the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
-Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} results in a dimensionless equation,
+Note that rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
+results in a dimensionless equation \citep{Slatkin1973},
 implying that the cline establishes over a timescale of the order $1/s$. 
@@
 Now suppose that the frequency profile of the selected allele, $p(x,t)$, is known.
-We seek to first understand how lineages of sampled individuals behave at the selected locus. 
-These results allow us to extend our analysis to lineages at loci linked to the selected locus.
+We seek to first understand how lineages of sampled individuals behave at the selected locus,
+and next extend the analysis to lineages at loci linked to the selected locus.
 
@@
 Making the same assumptions as for equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
-how the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
+that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
 with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient

### 2016-03-07 drafts/hybrid_zone.tex
edits

@@
 
-As in \citet{sedghifar2015spatial}, consider where the ancestors of modern-day individuals lived,
+As in \citet{sedghifar2015spatial}, we consider where the ancestors of modern-day individuals lived,
 i.e., where their lineages fall across geography as we look further back towards the time of first contact.
@@
     \textbf{A schematic figure of the lineages along which a segment of genome has been inherited,}
-    i.e., showing where the ancestors lived from whom the haplotype was inherited (horizontal axis),
-    and how long ago they lived (vertical axis).
+    i.e., showing where (horizontal axis) and when (vertical axis) the ancestors, from whom the haplotype was inherited, lived.
     The segment contains the $B$ alelle at the selected locus.
@@
 Deterministic theory predicts that after secondary contact,
-the two alleles at the selected locus will form a stable cline \citep{barton1979gene},
-affecting neighboring loci as well.
+the two alleles at the selected locus will form a stable cline,
+affecting neighboring loci as well \citep{barton1979gene, Barton1986}.
 First, we turn our attention to how the cline at the selected locus itself is formed.
@@
 The behavior of a lineage at a linked locus is similar to a selected locus. 
-However, there is one important difference -- in heterozygotes the lineage lineage linked to the selected locus may move between backgrounds. 
+However, there is one important difference -- in heterozygotes the lineage linked to the selected locus may move between backgrounds. 
 Therefore, if we follow back through time a lineage at a locus linked to an $A$ allele, 
@@
     Grey lines represent the same positions from a simulation with
-    identical parameters except $s=0$.
+    identical parameters except that $s=0$.
     Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
@@
 so (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x/(\sqrt{s}\sigma)$ generations ago 
+to have last had an ancestor on the $A$ side of the cline around $x/(\sigma\sqrt{s})$ generations ago 
 (compared to $x^2/\sigma^2$ for a neutral allele).
@@
 due to the branching process described above. 
-But, if we assume they are independent, 
+But, if we assume they are \emph{independent}, 
 then since the block about location $r$ in an individual at geographic location $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
-we'd expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ \alisa{consecutive} blocks of ancestry $A$ about a given site unlinked to the selected site.
+we would expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ \alisa{consecutive} blocks of ancestry $A$ about a given site unlinked \alisa{unlinked, or linked?} to the selected site.
 This implies the mean length $l(x,m)$ would be $2q_A(x,T,r)/(T(1-q_A(x,T,r)))$,
@@
 that looks like haplotypes from the center of the zone.
-This analysis suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$ \alisa{should this be $2/T$?};
+This analysis suggests that $A$ haplotypes in the center of the zone should be of average length $4/T$;
 this is indeed what is seen at distant sites, for instance, in Figure \ref{sfig:blocksAlongChromAncBConditioning_nonnormalized}.
@@
 although a numerical prediction of the value of $C(x,m)$ is elusive.
+\alisa{Above argument assumes independence between lineages. Does this make block lengths conservative? }
 
@@
 If selection is too strong, it may be difficult to observe the signal
-due to a lack of introgressed selective sites,
+due to a lack of introgressed selected sites,
 while if selection is weak, the selected lineage does not move very fast,
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{{figs/results_runid_753324_chr1_start0.48_stop0.52_by0.001_blocksAlongChromNoConditioning_nonnormalized}.pdf}
-\caption{
-    \textbf{Mean enclosing block length $l(m,x)$}, across the genome (horizontal axis)
-    and at different geographic positions (different colored lines).
-    Results are from a simulation with 120 demes of 200 diploids each, selection $s=.05$, dispersal $\sigma=3$, 
-    and after $T=500$ generations.
-}\label{Fig:meanBlockLengthsUnconditioned}
-\end{figure}
-
-\begin{figure}
-\includegraphics{figs/blocksAlongChromAncBConditioning}
-\caption{
-    \textbf{Normalized mean enclosing block length}, $C(m,x)$, after $T=1000$ generations,
-    against position relative to the selected locus (horizontal axis)
-    located in the center of a 1M chromosome.
-    Each line shows the mean block length at that spatial and genomic position
-    divided by the mean over the chromosome at that location;
-    the simulation was run with $s=1$ and $\sigma=1$,
-    50 demes, each containing 500 diploid individuals.
-}\label{Fig:blockLengths}
-\end{figure}
-
-\begin{figure}
 \includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
-\caption{Mean block length, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
+\caption{\textbf{Normalized mean enclosing block length}, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
 Here each line represents  $l_B(m)$, in a given deme, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
@@
 \includegraphics{figs/number_of_ancestors_tau1000}
-\caption{The number of individuals of ancestry $B$ per number of ancestors from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
+\caption{\textbf{Mean family size of haplotypes.} The number of individuals of ancestry $B$ per number of ancestors ($F_B(m,x)$) from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
 The hybrid zone is 50 demes, each containing 500 diploid individuals.
@@
 
-\yb{remove this? or, what figure?}
-Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location.
+%\yb{remove this? or, what figure?}
+Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location (Figure~\ref{Supp:adjacentBlocks}).
 An intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. 
@@
 Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination.
-Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
-\yb{provide citation.}
+Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection \citep{Slatkin1975b}.
 
@@
 Work on these ``islands of divergence'' \citep{turner2005genomic,Nosil2009} and related patterns have been largely descriptive \citep{Cruickshank2014,Noor2009}.
-This paper contributes to a model-based understanding of 
-how migration and selection influence such patterns across the genome of hybridizing populations. 
+Our study here contributes to a model-based understanding of 
+how migration and selection may influence such patterns across the genome of hybridizing populations. 
 Overall, focusing on lengths of ancestry blocks across the genome brings focus to the processes of migration and selection 

### 2016-03-07 drafts/supplement.tex
edits

@@
         As in Figure \ref{alleleFreq_tau100},
-        frequency of ancestry B,
+        frequency of ancestry $B$,
         across geography and at several physical positions on the genome,
@@
 \caption{
-    Frequency of ancestry B,
+    Frequency of ancestry $B$,
     across geography and at several physical positions on the genome, simulated for a hybrid zone
@@
     and with $s=0.01$.
-        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+        50 demes, each with a population size of 500 diploid individuals.
     Each line represents a locus some distance $r$ away from the true target of selection, 
     and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
-    identical parameters except $s=0$.
+    identical parameters except that $s=0$.
 }\label{alleleFreq_tau100_weaker_s}
@@
     and with $s=0.1$.
-        50 demes, each with a population size of 500 diploid individuals, were simulated across the hybrid zone.
+        50 demes, each with a population size of 500 diploid individuals.
     Each line represents a locus some distance $r$ away from the true target of selection, 
     and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
-    identical parameters except $s=0$.
+    identical parameters except that $s=0$.
  }\label{alleleFreq_tau1000}
@@
 \includegraphics[width=\textwidth]{figs/plot_chromosomes_tau100}
-\caption{Randomly sampled chromosomes  across a hybrid zone of age $\tau=100$.
+\caption{Randomly sampled chromosomes  across a hybrid zone of age $T=100$.
 Here we compare chromosomes of length 1M from a neutral zone to one that has a single under-dominant locus ($s=0.1$) in the middle of the chromosome (indicated by black arrow).
 Red blocks along the chromosome denote ancestry $B$, and orange blocks are ancestry $A$.
-Regions surrounding the selected site are failing to introgress.
  }\label{Fig:resistanceToIntrogression100g}
@@
     \caption{
-        \textbf{Mean haplotype lengths of $A$ haplotypes}, $l(x,m)$,
+        \textbf{Mean haplotype lengths of $A$ haplotypes} \alisa{Should this be $B$ haplotypes?
+That is what the figure title seems to indicate}, $l(x,m)$ \alisa{this would need to be changed to either $l_A$ or $l_B$},
         across the genome (horizontal axis) and at differntial spatial locations (colored lines),
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{{figs/results_runid_753324_chr1_start0.48_stop0.52_by0.001_blocksAlongChromNoConditioning_nonnormalized}.pdf}
+\caption{
+    \textbf{Mean enclosing block length $l(m,x)$}, across the genome (horizontal axis)
+    and at different geographic positions (different colored lines).
+    Results are from a simulation with 120 demes of 200 diploids each, selection $s=.05$, dispersal $\sigma=3$, 
+    and after $T=500$ generations.
+}\label{Fig:meanBlockLengthsUnconditioned}
+\end{figure}
+
+\begin{figure}
+\includegraphics{figs/blocksAlongChromAncBConditioning}
+\caption{
+    \textbf{Normalized mean enclosing block length}, $l_B(m,x)$, after $T=1000$ generations,
+    against position relative to the selected locus (horizontal axis)
+    located in the center of a 1M chromosome.
+    Each line shows the mean block length at that spatial and genomic position
+    divided by the mean over the chromosome at that location;
+    the simulation was run with $s=1$ and $\sigma=1$,
+    50 demes, each containing 500 diploid individuals.
+}\label{Fig:blockLengths}
+\end{figure}
+
+\begin{figure}
 \includegraphics{figs/blocksAlongChromHeatmapAncBConditioning}
-\caption{Heatmap of mean block length $l_B(m_i)$ along a simulated chromosome. }\label{Supp:blockLengthHeatmap}
+\caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome. }\label{Supp:blockLengthHeatmap}
 \end{figure}
@@
 \includegraphics{figs/blocksAlongChromNoConditioning}
-\caption{Heatmap of mean block length $l_\bullet(m_i)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, \tau=1000$). }\label{Supp:blockLengthNoAnc}
+\caption{Heatmap of mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
 \end{figure}
@@
 \includegraphics{figs/blocksAlongChromHeatmap}
-\caption{Heatmap of mean block length $l_\bullet(m_i)$ along a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
+\caption{Heatmap of mean block length $l(m)$ along a simulated chromosome with a single underdominant site ($s=0.01, T=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
 \end{figure}
@@
 \includegraphics{figs/adjacentBlocksAlongChromAncBConditioning}
-\caption{Mean block length of $l_A(m_i\cdot)$ across chromosome with single under dominant site ($s=0.01, \tau=1000$)}\label{Supp:adjacentBlocks}
+\caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus. ($s=0.01, T=1000$)}\label{Supp:adjacentBlocks}
 \end{figure}
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning}
-\caption{Ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and adjacent block lengths across a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$).
+\caption{Ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
 Each line represents a deme and is normalized by mean block length across the chromosome in the deme.}\label{Supp:ratioBlockAdjacent}
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning}
-\caption{Heatmap of $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site ($s=0.01, \tau=1000$). }\label{Supp:ratioBlockAdjacentHeatmap}
+\caption{Heatmap of $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+}\label{Supp:ratioBlockAdjacentHeatmap}
 \end{figure}
@@
 \begin{figure}
-    \includegraphics{figs/cond_freqs_comparison_tau_80}
+    \includegraphics[height=0.8\textheight]{figs/cond_freqs_comparison_tau_80}
     \caption{

### 2016-03-07 drafts/figs/Plot_frequency_clines.R

@@ -1,10 +1,10 @@
 
-source("../../sims/sim-fns.R",chdir=TRUE)
+source("../sims/sim-fns.R",chdir=TRUE)
 library(ReacTran)
 library(Matrix)
-source("../../notes/tran.1D.R",chdir=TRUE) # fix for log.A
-source("../../notes/cline-fns.R",chdir=TRUE)
-source("../../sims/chunks_fns.R",chdir=TRUE)
+source("../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../notes/cline-fns.R",chdir=TRUE)
+source("../sims/chunks_fns.R",chdir=TRUE)
 
 
 ###
@@ -17,6 +17,7 @@ get_simparams <- function (fname) {
 
 
 positions = seq(0.2,0.8,0.01)
+positions = seq(0.48,0.52,0.001)
 
 sel.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
 sel.params <- get_simparams( sel.sim.file )
@@ -47,7 +48,7 @@ if (!file.exists("plot_frequency_clines.Robj")) {
             IND_DATA=sims.sums[[1]]$ind.ancest, 
             demeID=demeID, 
             VECTOR_OF_POS=positions, 
-            CHR=2, 
+            CHR=1, 
             ndemes=neu.params$ndemes)
 
     save(B_freq, B_neu, file="plot_frequency_clines.Robj")
@@ -56,16 +57,16 @@ if (!file.exists("plot_frequency_clines.Robj")) {
 }
 
 
-pdf(file="alleleFrequencies_sim.pdf",height=4,width=6,pointsize=10)
+pdf(file="alleleFrequencies_sim_s0.01_tau100_closely_linked.pdf",height=4,width=6,pointsize=10)
 par(mar=c(3.5,3.5,0.5,0.5))
-    matplot(B_neu/1000,xlab="",ylab="",type="l",lty=1,col="grey",lwd=0.5,x=seq(-24.5,24.6,1))
+    matplot(1-B_neu/1000,xlab="",ylab="",type="l",lty=1,col="grey",lwd=0.5,x=seq(-24.5,24.6,1),ylim=c(0,1))
     for(i in 1:30){
[…]

### 2016-03-07 sims/run_sims.R

@@ -6,8 +6,8 @@ source("sim-fns.R")
 
 params = list(
     SIGMA = 3,
-    ndemes = 10,
-    deme_size = 10,
+    ndemes = 100,
+    deme_size = 50,
     S = 0.05,
     zone_age = 10
 )
@@ -86,11 +86,24 @@ dev.off()
 
 focal_sites = lapply(qtl,function(Z){Z$pos})
 
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
+	#return the interval containing focal site for a individual	
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancA),0)
+}
+
+
+get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
+	#get distribution of chunks within a deme
+	INDS = IND_DATA[which(sims.full[[1]]$deme==DEME)]
+	intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
+	return(intervals)
+}
+
 #output is a nind*2 matrix, where each column is a chromosome. 
 #EXAMPLE: testB = lapply(1:50,function(X){get.deme.chunks(DEME=X,ancA=FALSE)})
 #D = 23; hist(testB[[D]][which(testB[[D]]<1 & testB[[D]]>0)], col="black",breaks=seq(0,1,0.05))
-
-deme_ID = rep( 1:params$ndemes, each=params$deme_size )
 
 chunks.rvals = list( 
         selected=c(chr=1,pos=0.5),
@@ -102,10 +115,7 @@ chunks.rvals = list(
         unlinked=c(chr=2,pos=0.5),
         unlinked0.4=c(chr=2,pos=0.9) )
 chunks <- lapply( chunks.rvals, function (xx) {
-            lapply(1:params$ndemes, function (X) { 
-                       get.deme.chunks(IND_DATA=sims.sums[[1]]$ind.ancest,demeID=deme_ID,
[…]

### 2016-03-08 drafts/hybrid_zone.tex
minor fixes

--- 2016-03-07 drafts/hybrid_zone.tex
@@
 although these are in principle different,
-\citet{Slatkin1973} pointed out (among many other things)
+\citet{Slatkin1973} pointed out, in a far-reaching general analysis,
 that while the environment may determine the location of a cline,
 the fitness of the heterozygote will likely determine the shape of the cline,
-and \citet{Barton1986} show that the shapes of several sorts of stable clines are similar.
-There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
+and \citet{barton1986barrier} show that the shapes of several sorts of stable clines are similar.
+There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
 and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
@@
 This is particularly suited to understanding transient patterns,
-although because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002neutral},
+although because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
 incorporating genetic drift is not straightforward.
@@
 % Population genetic modeling of hybrid zones has focused largely on the 
-% equilibrium states of clines at selected loci \citep{Nagylaki1975,barton1979dynamics,Barton1986,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
+% equilibrium states of clines at selected loci \citep{Nagylaki1975,barton1979dynamics,barton1986barrier,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
 % Since stronger selection reduces the amount of genetic mixing,
@@
 % at loci close to selected loci, and wider clines at less tightly linked loci.
-% In particular \citet{barton1979gene,Barton1986} have described patterns of clinal variation at loci linked to targets of selection.  
+% In particular \citet{barton1979gene,barton1986barrier} have described patterns of clinal variation at loci linked to targets of selection.  
 % This provides some expectation for how selection shapes covariance in ancestry at a single locus over geographic space.
@@
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,Barton2000a}.
+previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}.
 
@@
 for discussion of diploidy see \citet{christiansen1995genotypic}.
+As we show below the solutions provide an good approximation across the range of realistic values of $s$.
 Although there is no known exact analytic solution to this equation, 
@@
 \paragraph{Lineages at linked loci}
-%YB DID SUBSTANTIAL EDITING OF THIS PARAGRAPH
 The behavior of a lineage at a linked locus is similar to a selected locus. 
-However, there is one important difference -- in heterozygotes the lineage lineage linked to the selected locus may move between backgrounds. 
+However, there is one important difference -- 
+in heterozygotes the lineage linked to the selected locus may move between backgrounds. 
 Therefore, if we follow back through time a lineage at a locus linked to an $A$ allele, 
@@
 (as described for a selected allele),
-and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$).
+and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$),
+by the assumption of locally random assortment of alleles
+(see Appendix \ref{apx:lineage_derivation}).
 
@@
 with $B_\tau$ a standard Brownian motion,
-$T_B$ the first time $Z_\tau = B$, and likewise for $T_A$,
+$T_B$ the most recent time that $Z_\tau = B$, and likewise for $T_A$,
 \begin{align}
@@
         with cline flattening continuing with $\sqrt{t}$.
-        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
+        \textbf{Remaining figures} show frequencies of $A$ ancestry \emph{conditional}
         on the ancestry at the selected site,
@@
 We can use this diffusion model for lineages to find expected clines in ancestry,
-i.e., the expected proportion of individuals who inherit ancestry from species $A$,
+i.e., the expected proportion of individuals who inherit from species $A$,
 as a function of space, time, and position on the genome.
@@
 \paragraph{Haplotype lengths}
-We now find the probability of maintaining an \emph{entire} block of genome (haplotype) % thought `seeing' was too much about our inference rather than the the biological process
-of a single ancestry at a given location and time.
+We now find the frequency at which \emph{entire} blocks of genome (haplotype) 
+% YB: thought `seeing' was too much about our inference rather than the the biological process
+% PLR: hm, it's not 'maintaining' either, since the block might have been broken up along the way.
+of a single ancestry are found, at a given location and time.
 Consider a segment of the genome between positions $a$ and $b$, relative to the selected site. 
@@
 each of which proceeds as before, independently. 
+In this description,  % don't like to have math at the start of a sentence
 $g_z(x,T;a,b)$ is then the probability that all branches are found on the $A$ side of the hybrid zone
@@
 we need only follow lineages at two sites, instead of an entire region. 
-Doing so requires only minor modifications of the model, above. 
-Moreover, direct use of our numerical code provides a measure of the correlation in ancestry between markers. 
-See Appendix~\ref{apx:haplotype_calcs} for more detail.
+Doing so only requires computing correlations in ancestry between markers,
+which can be done directly using our numerical code;
+see Appendix~\ref{apx:haplotype_calcs} for more detail.
 
@@
 the chromosome inherited that genomic region.
-We the assigned ancestry at individual loci by looking up the ancestor's genotype.
+We then assigned ancestry at individual loci by looking up which side of the zone the ancestor lived on.
+% removed "ancestor's genotype" because it's not really genotype
 
@@
 stochasticity has substantially moved allele frequencies range-wide
-(which is unsurprising, as the total population size is only 5,000 diploids) \citep{Polechova2011}.
+\citep[which is unsurprising, as the total population size is only 5,000 diploids][]{Polechova2011}.
 
@@
 In practice, however, it can take quite some time for even unlinked neutral loci to homogenize,
-due to the decreased fitness of heterozygotes \citep{Barton1986}
+due to the decreased fitness of heterozygotes \citep{barton1986barrier}
 and the relative slowness of diffusive movement.
@@
 to have a good chance of recombining,
-so it is certain that clines with $r<1/T$ will still resemble the selected cline,
+so clines with $r<1/T$ will still resemble the selected cline,
 which can also be seen in Figure \ref{fig:linked_cline_heatmaps}.
@@
 so (selected) $A$ alleles at distance $x$ from the cline center
-to have last had an ancestor on the $A$ side of the cline around $x/(\sqrt{s}\sigma)$ generations ago 
+have last had an ancestor on the $A$ side of the cline around $x/(\sqrt{s}\sigma)$ generations ago 
 (compared to $x^2/\sigma^2$ for a neutral allele).
@@
 Genomic regions are inherited from ancestors $T$ generations ago
-in blocks of size roughly $1/T$,
-so there will be a long segment of $A$ ancestry
-if many such adjacent blocks are of the same ancestry.
+in blocks of size roughly $1/T$.
+There will be a long segment of $A$ ancestry
+if many such adjacent blocks are all inherited from the $A$ side.
 Ancestries of neighboring blocks are correlated,
@@
 But, if we assume they are independent, 
-then since the block about location $r$ in an individual at geographic location $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
-we'd expect to see, on average, $2q_A(x,T,r)/(1-q_A(x,T,r))$ \alisa{consecutive} blocks of ancestry $A$ about a given site unlinked to the selected site.
-This implies the mean length $l(x,m)$ would be $2q_A(x,T,r)/(T(1-q_A(x,T,r)))$,
-and the mean conditional length $l_A(x,m)$ would be $2/(T(1-q_A(x,T,r))$.
-Furthermore, we know that if $x=0$ or if $T$ is large and $r$ is not too small, that $q_A$ is close to $1/2$;
+then since the block surrounding location $r$ in an individual at geographic location $x$ is of ancestry $A$ with probability $q_A(x,T,r)$,
+we'd expect to see, roughly, 
+$q_A(x,T,r)(1+2q_A(x,T,r))/(1-q_A(x,T,r))$ 
+% P(N=0) = q and if N>0 then is 1 + two independent q-geometrics with mean q/(1-q),
+% so is q ( 1 + 2q/(1-q) ) = q(1+2q)/(1-q) .
+consecutive blocks of ancestry $A$ about a given site unlinked to the selected site.
+(This assumes, crudely, that the number of $A$-blocks on either side of the enclosing block
+has a Geometric distribution with parameter $q_A(x,T,r)$.)
+This implies the mean length $l(x,r)$ would be $q_A(x,T,r)(1+2q_A(x,T,r))/((1-q_A(x,T,r))T)$,
+and the mean conditional length $l_A(x,r)$ would be $(1+2q_A(x,T,r))/(T(1-q_A(x,T,r))$.
+Furthermore, we know that if $x=0$ or if $T$ is large and $r$ is not too small, that $q_A$ is close to $1/2$
+(so $l_A(x,r) \approx 4/T$);
 and for $r<1/T$ that $q_A$ looks like the selected cline.
@@
 and loci linked to the selected locus have intermediate sizes.
-This pattern is consistent with our theoretical results showing that unfit lineages tend to be recent migrants, 
+This pattern is consistent with the prediction 
+% our theoretical results showing   % sounds pretentious to call the obvious observation 'our results'
+that unfit lineages tend to be recent migrants,
 which will have smaller families.
@@
 
-\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
+Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
 We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
@@
 This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency. 
-Additionally, our work suggests that the ancestry block length distribution can help detect targets of purifying selection in hybrid zones.}  
+Additionally, our work suggests that the ancestry block length distribution can help detect targets of purifying selection in hybrid zones.  
 The resolution of this approach is expected to scale with $1/T$, as this is the physical scale over which linked clines persist.
@@
 
-\yb{remove this? or, what figure?}
-Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location.
-An intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. 
-Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby.
-
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 
-% many modifications here
-\yb{Our results focus on the length distribution of ancestry blocks --- a statistic that can only be obtained in the few systems for which dense genetic markers and phase information is known. 
+Our results focus on the length distribution of ancestry blocks --- a statistic that can only be obtained in the few systems for which dense genetic markers and phase information is known. 
 However, as the cost of sequencing continues to drop and long-phased reads become more common, the patterns described here could further aid in the identification of selected sites in hybrid zones in many non-model taxa.
 In the meantime, patterns of elevated pairwise LD, which is easily computed from readily available low-density and unphased genomic data, could offer an alternative path forward for empirical work. 
-Importantly, the numerical solutions derived above readily and efficiently predict correlations in ancestry, and so our results can be applied immediately to genotype data for any taxa with markers placed on a genetic map.}
+Importantly, the numerical solutions derived above readily and efficiently predict correlations in ancestry, and so our results can be applied immediately to genotype data for any taxa with markers placed on a genetic map.
 
@@
 Our model is most relevant to scenarios with few targets of selection scattered throughout the genome, and therefore, our predictions may differ significantly from situations in which the density of selected sites is higher.
-Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination.
+Having numerous selected sites within one ancestry block increases the strength of selection \citep{barton1986barrier}, 
+and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination \citep{barton1986effects}. 
 Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
-\yb{provide citation.}
+% \yb{provide citation.} % grr, can't find it.
 
@@
 Extending our analytical results to capture stochasticity arising from coalescence/pedigree structure represents an important future direction.  
-Ignoring pedigree structure can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{liang2014lengths}. 
 Indeed, correlated fluctuations visible in simulations (e.g., Figure \ref{Fig:blockLengths})
@@
 and simulations at lower density show larger fluctuations than those at higher density.
-The agreement between analytic predictions and simulations (which include an explicit pedigree structure)
-suggests that this assumption does not have a severe impact on our major results when population densities are moderately high.
+Furthermore, ignoring pedigree structure can result in an underestimate of covariance in ancestry, as we have ignored additional sharing of ancestry through shared genealogy \citep{liang2014lengths};
+but there is nonetheless good agreement between analytic predictions and simulations (which include an explicit pedigree structure).
+% suggests that this assumption does not have a severe impact on our major results when population densities are moderately high.
 
@@
 \subsection*{Adaptive introgression}
-%\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{Barton1986}}
+%\plr{don't have to call it the hybrid filter, and it looks like maybe I made up that term, but after reading this paper: \citet{martinsen2001hybrid}; these sorts of things are discussed in \citet{barton1986barrier}}
 
 While our focus has been on hybrid incompatibilities, 
-unconditionally adaptive loci are also known to introgress across hybrid zones \citep{Barton1986,martinsen2001hybrid,arnold2004transfer}. 
+unconditionally adaptive loci are expected to easily introgress across hybrid zones \citep{barton1979gene,barton1986barrier,martinsen2001hybrid,arnold2004transfer}. 
 Future work could take a similar approach to understand how positive selection shapes ancestry block lengths, and predict  signatures of adaptive introgression in hybridizing populations using similar statistics presented here.

### 2016-03-08 drafts/hybrid_zone.tex

@@
 
+Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location (Figure~\ref{Supp:adjacentBlocks}).
+An intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. 
+Compared to other chromosomes in their new geographic location, these migrants will have on average longer $B$ haplotypes at the selected locus, and shorter $A$ haplotypes nearby.
+
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
@@
 and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination \citep{barton1986effects}. 
-Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection.
-% \yb{provide citation.} % grr, can't find it.
+Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection \citep{Slatkin1975b,Barton1983}.
 

### 2016-03-08 drafts/supplement.tex
last fix

@@
     \caption{
-        \textbf{Mean haplotype lengths of $A$ haplotypes} \alisa{Should this be $B$ haplotypes?
-That is what the figure title seems to indicate}, $l(x,m)$ \alisa{this would need to be changed to either $l_A$ or $l_B$},
+        \textbf{Mean haplotype lengths of $B$ haplotypes}, $l_B(x,m)$,
         across the genome (horizontal axis) and at differntial spatial locations (colored lines),

### 2016-03-08 drafts/hybrid_zone.tex
minor fixes

@@
 Hybrid zones formed between recently diverged populations offer an opportunity 
-to study the mechanisms underlying reproductive isolation and the process of speciation. % too much `understanding' here
+to study the mechanisms underlying reproductive isolation and the process of speciation.
 Here, we use a combination of analytical theory and explicit forward simulations
-to show how selection against hybrid genotypes impacts patterns of introgression  % too much `understanding' here
+to show how selection against hybrid genotypes impacts patterns of introgression
 across genomic and geographic space. 
@@
 The lengths of continuous tracts of ancestry provide an additional source of information:
-blocks of ancestry surrounding single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location. 
-Our findings suggest that the density of ancestry junctions provide a means for characterizing hybrid zones and identifying candidate targets of selection.
+blocks of ancestry surrounding single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location,
+an observation that might be used to characterize the age of hybrid zones
+and identify candidate targets of selection.
 %\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
@@
     \textbf{A schematic figure of the lineages along which a segment of genome has been inherited,}
-    i.e., showing where (horizontal axis) and when (vertical axis) the ancestors, from whom the haplotype was inherited, lived.
+    showing the location (horizontal axis) and time (vertical axis) at which the ancestors 
+    of a hypothetical sampled haplotype lived.
     The segment contains the $B$ alelle at the selected locus.
     The black line shows the lineage at the selected site in the sampled individual in the present day (at $\tau=0$).
-    The center of the hybrid zone is at geographic position $x=0$; in the region $x>0$, ancestry $A$ is less common,
+    The center of the hybrid zone is at geographic position $x=0$; in the region $x<0$, ancestry $B$ is less common,
     and because of negative selection, the selected lineage spends little time there.
@@
     which switch between selected backgrounds once.
-    Based on position of the lineages at $\tau=100$ generations after contact, 
+    Based on position of the lineages at $\tau=100$ generations ago,
     from left to right the segments are of ancestry $(A,B,B)$.
@@
      using 50 demes, each with 500 diploid individuals and $\sigma=1$.
-    Each line represents a locus some distance $r$ away from the true target of selection, 
-    with $r=0$ representing the locus under selection. 
+    Each line represents a locus some distance $r$ away from the true target of selection.
     Grey lines represent the same positions from a simulation with
@@
 As expected, regions surrounding a locus under selection are more resistant to introgression,
-as seen in Figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}).
+as seen in Figures~\ref{Fig:resistanceToIntrogression1000g} and \ref{Fig:resistanceToIntrogression100g}.
 When present, we expect haplotypes that contain the locally less common allele
@@
 Indeed,
-as shown in Figures \ref{Fig:blockLengths}, \ref{Fig:blockLengthsZoom} and \ref{Supp:blockLengthHeatmap},
+as shown in Figures \ref{Fig:blockLengthsZoom}, \ref{Fig:blockLengths} and \ref{Supp:blockLengthHeatmap},
 the mean length of such haplotype blocks is up to three times longer
@@
 although a numerical prediction of the value of $C(x,m)$ is elusive.
-\alisa{Above argument assumes independence between lineages. Does this make block lengths conservative? }
+% \alisa{Above argument assumes independence between lineages. Does this make block lengths conservative? }
+% PR: yes, maybe, but that correction seems like a small one, and it's not getting very quantitative predictions;
+% a more serious issue is that it's ignoring that the next block has a different $r$.
 

### 2016-03-08 drafts/supplement.tex

@@
 \begin{figure}
-    \includegraphics{figs/linked-frequencies-longtime}
+    \includegraphics{figs/linked-frequencies-longtime.pdf}
     \caption{
@@
 \begin{figure}
-    \includegraphics{figs/alleleFrequencies_sim_comparison}
+    \includegraphics{figs/alleleFrequencies_sim_comparison.pdf}
     \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau100}
+\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau100.pdf}
 \caption{Randomly sampled chromosomes  across a hybrid zone of age $T=100$.
@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromAncBConditioning}
+\includegraphics{figs/blocksAlongChromAncBConditioning.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromHeatmapAncBConditioning}
+\includegraphics{figs/blocksAlongChromHeatmapAncBConditioning.pdf}
 \caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome. }\label{Supp:blockLengthHeatmap}
@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromNoConditioning}
+\includegraphics{figs/blocksAlongChromNoConditioning.pdf}
 \caption{Heatmap of mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromHeatmap}
+\includegraphics{figs/blocksAlongChromHeatmap.pdf}
 \caption{Heatmap of mean block length $l(m)$ along a simulated chromosome with a single underdominant site ($s=0.01, T=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
@@
 \begin{figure}
-\includegraphics{figs/adjacentBlocksAlongChromAncBConditioning}
+\includegraphics{figs/adjacentBlocksAlongChromAncBConditioning.pdf}
 \caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus. ($s=0.01, T=1000$)}\label{Supp:adjacentBlocks}
@@
 \begin{figure}
-\includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning}
+\includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning.pdf}
 \caption{Ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
@@
 \begin{figure}
-\includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning}
+\includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning.pdf}
 \caption{Heatmap of $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
@@
 \begin{figure}
-    \includegraphics[height=0.8\textheight]{figs/cond_freqs_comparison_tau_80}
+    \includegraphics[height=0.8\textheight]{figs/cond_freqs_comparison_tau_80.pdf}
     \caption{
@@
 \begin{figure}
-    \includegraphics{figs/cond_freqs_comparison_tau_320}
+    \includegraphics{figs/cond_freqs_comparison_tau_320.pdf}
     \caption{
@@
 \begin{figure}
-    \includegraphics{figs/cond_freqs_comparison_tau_1280}
+    \includegraphics{figs/cond_freqs_comparison_tau_1280.pdf}
     \caption{

### 2016-04-14 drafts/hybrid_zone.tex
added .pdf to figs

@@
 %\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
+
+\textbf{Keywords:} hybridization, tension zone, introgression, haplotypes, diffusion methods.
 }
@@
 \begin{figure}
-\includegraphics{figs/BM_schematic}
+\includegraphics{figs/BM_schematic.pdf}
 \caption{
@@
 \begin{figure}
-    \includegraphics{figs/linked-frequencies}
+    \includegraphics{figs/linked-frequencies.pdf}
     \caption{
@@
 \begin{figure}
-\includegraphics{figs/alleleFrequencies_sim}
+\includegraphics{figs/alleleFrequencies_sim.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000}
+\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
+\includegraphics{figs/blocksAlongChromAncBConditioningHighRes.pdf}
 \caption{\textbf{Normalized mean enclosing block length}, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
@@
 \begin{figure}
-\includegraphics{figs/number_of_ancestors_tau1000}
+\includegraphics{figs/number_of_ancestors_tau1000.pdf}
 \caption{\textbf{Mean family size of haplotypes.} The number of individuals of ancestry $B$ per number of ancestors ($F_B(m,x)$) from secondary contact occurring $T=1000$ generations ago, represented across geographic space.
@@
 \bibliography{library,references}
+
+\section*{Data Accessibility}
+
+All scripts used in production of this paper are available at 
+\url{https://github.com/petrelharp/clinal-lineages}
+under the GPLv3 software license,
+and on Data Dryad \url{http://datadryad.org/XXXX}.
 

### 2016-04-21 drafts/hybrid_zone.tex
more math explanation

@@
 In this section we aim to give both heuristics for how lineages move, 
-that are helpful in establishing order-of-magnitude estimates,
+that are helpful in providing intuition and for establishing order-of-magnitude estimates,
 and analytics, mostly in terms of partial differential equations needing numerical solution.
@@
 First, we turn our attention to how the cline at the selected locus itself is formed.
-Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
+Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
 Suppose that secondary contact occurred at time $t=0$, 
-and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
+and the $A$ allele is initially fixed on the left, so
+$p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
 % (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
-Let $\sigma^2$ be the mean squared distance between parent and offspring 
-(i.e., the variance of the distribution of dispersal distance).
+We assume that dispersal is local and unbiased;
+and that the mean squared displacement between parent and offspring
+in the direction perpendicular to the cline
+is $\sigma^2$.
 Assuming that: (i) alleles locally assort into diploids randomly, 
@@
 and (iv) population density is large,
-then the commonly-used equation that $p$ solves is
+then as in \citet{Bazykin1969},
+the commonly-used equation that $p$ solves is
 \begin{align} \label{eqn:cline_pde}
-    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
-\end{align}
-where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
-as in \citet{Bazykin1969}; 
-this can be derived as in \citet{Nagylaki1975};
+    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) .
+\end{align}
+where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian.
+\plr{change $\dot p$ to $\frac{d}{dt}p$, and $\Delta$ to the second derivative?}
+This says that the local rate of change of the frequency of $A$ ($\dot p$)
+is equal to the net rate at which $A$ alleles arrive by local dispersal ($\sigma^2/2 \Delta p$)
+and a logistic term proportional to the difference in reproductive rate between $A$ and $a$ alleles
+due to local proportions of heterozygotes,
+which is $s(2p-1)$ (up to an error of size $s^2$).
+This can be derived as in \citet{Nagylaki1975};
 for discussion of diploidy see \citet{christiansen1995genotypic}.
@@
 implying that the cline establishes over a timescale of the order $1/s$. 
-In other words, the cline is established through diffusion, despite being slowed down somewhat by selection against heterozygotes.
+Since this is how long it takes diffusion at rate $\sigma^2$ to smooth across a region of width $\sigma/\sqrt{s}$,
+this means we can think of
+the cline as established by diffusion, 
+despite being slowed down somewhat by selection against heterozygotes.
 
@@
 Making the same assumptions as for equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
-that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
-with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
+that the lineage of an $A$ allele moves as Brownian motion driven by dispersal
+with mean displacement proportional to the derivative of $\log p$.
+This has the effect of pulling $A$ lineages towards regions where they are more common
+% with speed $\sigma$
+% with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
 (i.e., it moves at speed $\sigma^2$ in the potential $-\log p$).
@@
 To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
-lived at position $x-r$ is proportional to $\P\{R=r\} p(x+r)$, where $R$ is the random dispersal distance;
-and so the mean displacement from offspring to parent is $\E[R p(x+R)]/\E[p(x+R)]$.
-% $$ = \E[ R ( p(x) + R p'(x) ) ]/ p(x) + O(R^3) $$
+lived at position $x-r$ is proportional to $\P\{R=-r\} p(x-r)$, where $R$ is the random dispersal distance;
+and so the mean displacement from offspring to parent is $-\E[R p(x-R)]/\E[p(x-R)]$.
+% $$ = \E[ -R ] - \E[R]^2 p'(x)/p(x) + \E[R^2] p'(x)/p(x) + O(R^3) $$
 Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$
@@
 which is $\sigma^2$ multiplied by the gradient of $\log p(x)$.
-
-At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
+This description holds even when the frequency profile of the $A$ allele changes with time (replacing $p(x)$ by $p(x,t)$).
+
+At equilibrium, the $A$ allele is nearly fixed at a geographic position far to the left ($p \approx 1$),
+and is rare far to the right, with mean frequency at distance $x$ proportional to $\exp(-x\sqrt{s}/\sigma)$.
 Therefore, roughly, lineages on ``their own'' side wander randomly,
 while lineages on ``the wrong'' side are pushed at constant speed $\sigma\sqrt{s}$ 
-back towards the side where they are more common. 
+back towards the side where they are more common
+(since here $\log p$ is $\sqrt{s}/\sigma$ and the speed is $\sigma^2$). 
 Since an $A$ allele must, by definition, have been inherited from the $A$ side of the barrier 
@@
 However, there is one important difference -- 
-in heterozygotes the lineage linked to the selected locus may move between backgrounds. 
+in heterozygotes the lineage linked to the selected locus may
+recombine onto the other selected background.
 Therefore, if we follow back through time a lineage at a locus linked to an $A$ allele, 
@@
 Suppose we sample an allele today $r$ Morgans from the selected site. 
-We will follow its lineage back through time, using $\tau$ to denote ``generations ago'' to avoid confusion.
+We will follow its lineage back through time, using $\tau$ to denote ``generations ago''
+and reserving $t$ for time measured in the usual direction.
 If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
@@
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ 
-(as described for a selected allele),
+(as described above for a selected allele),
 and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$),
 by the assumption of locally random assortment of alleles
-(see Appendix \ref{apx:lineage_derivation}).
+(see Appendix \ref{apx:lineage_derivation} for a more precise description).
 
@@
 with $B_\tau$ a standard Brownian motion,
-$T_B$ the most recent time that $Z_\tau = B$, and likewise for $T_A$,
+$T_B(\tau)$ the most recent time before $\tau$ that $Z_\tau = B$, and likewise for $T_A$,
 \begin{align}
     \begin{aligned} \label{eqn:lineage_motion}
-        dX_\tau &= \begin{cases}
-             \sigma^2 \grad \log(p(X_\tau,\tau)) d\tau + \sigma dB_\tau \qquad & \text{if } Z_\tau = A \\
-             \sigma^2 \grad \log(1-p(X_\tau,\tau)) d\tau + \sigma dB_\tau \qquad & \text{if } Z_\tau = B 
+        dX_\tau &= dB_\tau + \begin{cases}
+             \sigma^2 \grad \log(p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = A \\
+             \sigma^2 \grad \log(1-p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = B 
         \end{cases} \\
@@
         % \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B \} &= \epsilon r p(X_\tau,\tau) + O(\epsilon^2)  .
-        \P\{ T_B &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
-        \P\{ T_A &> \tau+u \given Z_\tau = B \} = \exp\left( - r \int_\tau^{\tau+u} p(X_s,s) ds \right) .
+        \P\{ T_B(\tau) &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
+        \P\{ T_A(\tau) &> \tau+u \given Z_\tau = B \} = \exp\left( - r \int_\tau^{\tau+u} p(X_s,s) ds \right) .
     \end{aligned}
 \end{align}
+In the first expression, giving the distribution of how the lineage location changes,
+the first term ($\sigma dB_\tau$) is Brownian noise driven by dispersal,
+and the second is the mean displacement, which moves the lineage ``downhill'' on either $-\log p$ or $-\log (1-p)$
+($\grad$ is the gradient),
+depending on which selected allele the lineage is linked to.
+The second and third expressions say that a lineage switches from an $A$ to $B$ background
+at rate $r$ multiplied by the local proportion of $B$ alleles, which is $(1-p)$.
 
@@
 and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
+The three terms in these equations come from Brownian movement of a linages,
+the tendency of a lineage to inherit from regions where its type is more common,
+and recombination between selected backgrounds, respectively.
+Equivalently,
+the first term ($\Delta q$) reflects the smoothing action of local dispersal,
+the second is the net flux induced by reduced fitness in the hybrid zone,
+and the third is the contribution of recombination.
 %\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
@@
 \end{align*}
-In technical terms, $G^A$ and $G^B$ are the generators of the diffusions of lineages of selected alleles of ancestry $A$ and $B$, respectively.
+In technical terms, $G^A$ and $G^B$ are the generators of the diffusions of lineages of selected alleles of ancestry $A$ and $B$, respectively;
+informally, they encode the stochastic motion of lineages at the selected loci.
 
@@
 at a site at recombination distance $r$ from the selected site,
-which is computed as $q(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 
+which is computed as the frequency of ancestry $A$ on each selected background weighted by the freequencies of each background:
+$q(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 
 The equations are solved numerically in R, using the ReacTran package \citep{soetaert2012reactive}.
@@
 \paragraph{Haplotype lengths}
-We now find the frequency at which \emph{entire} blocks of genome (haplotype) 
-% YB: thought `seeing' was too much about our inference rather than the the biological process
-% PLR: hm, it's not 'maintaining' either, since the block might have been broken up along the way.
+We now find the frequency at which \emph{entire} blocks of genome (haplotypes) 
 of a single ancestry are found, at a given location and time.
@@
 of finding an entire segment $(a,b)$ of ancestry $A$ in an individual having selected allele of type $z$,
-sampled at spatial location $x$ at time $t$ after the initiation of gene flow. % I think people sometimes use 'cline formed' to mean when the equilibrium was established, changed for clarity
+sampled at spatial location $x$ at time $t$ after the initiation of gene flow.
 
@@
 Prior to this, there are two lineages to follow (see Figure~\ref{Fig:schematic}),
-and so lineages act like a particular class of labeled, branching diffusions,
+and so lineages behave as labeled, branching diffusions,
 where the total branching rate is conserved. 
@@
 each of which proceeds as before, independently. 
-In this description,  % don't like to have math at the start of a sentence
+In this description,
 $g_z(x,T;a,b)$ is then the probability that all branches are found on the $A$ side of the hybrid zone
 at the time of secondary contact $T$ units of time ago.
+We will write $r(a,b)$ for the distance from the segment to the selected site:
+always taking $a \le b$,
+$r(a,b)=0$ if $a \le 0 \le b$ and $r(a,b)=\min(|a|,|b|)$ otherwise.
 
@@
             + (1-p) \left( 
-                \int_a^0 g_B(a,\theta) g_A(\theta,b) d\theta
-                + \int_0^b g_A(a,\theta) g_B(\theta,b) d\theta 
+                \int_a^{\min(0,b)} g_B(a,\theta) g_A(\theta,b) d\theta
+                + \int_{\max(0,a)}^b g_A(a,\theta) g_B(\theta,b) d\theta 
             \right)
@@
             } d\theta
-            - (b-a) g_A(a,b)  
-    \end{aligned} ,
-\end{align}
-where $\theta$ is the genomic position of a recombination event, 
-the integral $\int_a^0$ is zero if $a>0$ and likewise $\int_0^b$ is zero if $b<0$;
-and $r(a,b)$ is the distance from the segment to the selected site:
-$r(a,b)=0$ if $ab<0$ and $r(a,b)=\min(|a|,|b|)$ otherwise.
+            - (b-a) g_A(a,b)  ,
+    \end{aligned} 
+\end{align}
+Here the first term ($G^A g_A$) represents spatial mixing,
+and the second term results from recombinations \emph{between} the block and the selected site 
+in heterozygotes, which switch the identity of the linked, selected locus
+(the factor $(1-p)$ is the probability that the block, initially linked to an $A$ allele, encounters a $B$ allele
+under the assumption of locally random mating).
+In the integrals, $\theta$ is the genomic position of the recombination event occuring \emph{inside} the block.
+The third term accounts for such recombinations in a heterozygote: 
+the portion of the block nearest the selected site remains linked to an $A$ allele, 
+and the remaining portion becomes linked to a $B$ allele
+(and we say that the integral $\int_a^{\min(0,b)}$ is zero if $a>0$ 
+and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$).
+The last integral results from recombination inside the block in homozygotes for $A$,
+and the final term balances the outflux due to all recombinations.
 The equation for $g_B(x,t;a,b)$ is identical after exchanging $A \leftrightarrow B$ and

### 2016-04-21 drafts/hybrid_zone.tex
some minor edits

--- 2016-04-14 drafts/hybrid_zone.tex
@@
     of a hypothetical sampled haplotype lived.
-    The segment contains the $B$ alelle at the selected locus.
-    The black line shows the lineage at the selected site in the sampled individual in the present day (at $\tau=0$).
+    At the time of sampling ($\tau=0$), the segment contains the $B$ allele at the selected locus.
+    Paths of lineages in space, backwards in time, are depicted by blue and yellow lines. Splitting 
+    events  represent recombination events along the sampled segment.
+    The thick blue line represents the path of the selected locus. The colors of the other paths indicate 
+    the identity at the linked selected locus (blue: linked to allele $B$. yellow: linked to allele $A$). 
     The center of the hybrid zone is at geographic position $x=0$; in the region $x<0$, ancestry $B$ is less common,
     and because of negative selection, the selected lineage spends little time there.
-    Red dots represent ancestral recombination events that fall in the given segment of genome in an ancestor;
-    these can cause the lineage to branch (lineages at linked sites are depicted as colored lines),
-    or can switch lineages at linked sites between selected backgrounds 
-    (yellow lineages are linked to the $A$ selected locus; blue are linked to $B$).
-    Here the initial haplotype is broken up into 3 chunks over 100 generations,
-    which switch between selected backgrounds once.
+  %  Red dots represent ancestral recombination events that fall in the given segment of genome in an ancestor;
+  %  these can cause the lineage to branch (lineages at linked sites are depicted as colored lines),
+  %  or can switch lineages at linked sites between selected backgrounds 
+  %  (yellow lineages are linked to the $A$ selected locus; blue are linked to $B$).
+  %  Here the initial haplotype is broken up into 3 chunks over 100 generations,
+  %  which switch between selected backgrounds once.
     Based on position of the lineages at $\tau=100$ generations ago,
@@
 (ii) habitat and dispersal are homogeneous,
-(iii) $s$ and $\sigma$ are small, 
+(iii) $s$ and $\sigma$ are small, \alisa{relative to $N$ and $\tau$ respectively?}
 and (iv) population density is large,
@@
 stochasticity has substantially moved allele frequencies range-wide
-\citep[which is unsurprising, as the total population size is only 5,000 diploids][]{Polechova2011}.
+\citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000.pdf}
+\includegraphics[width=\textwidth]{figs/plot_chromosomes_tau1000_zoomed.pdf}
 \caption{
@@
 
-Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location (Figure~\ref{Supp:adjacentBlocks}).
+Through our simulations we find also that blocks of ancestry that have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are shorter than average for the spatial location (Figure~\ref{Supp:adjacentBlocks}).
 An intuitive explanation for this pattern is that loci physically linked to a selected site have recently come from the center of the zone or beyond. 

### 2016-04-21 drafts/hybrid_zone.tex
draft responses and some changes

@@
 
+\section*{Responses}
+Editor Comments to Author:
+
+Both reviewers appreciate the paper, but also make several criticisms.
+I am considerably more positive, and believe that those criticisms that are justified can be dealt with quite easily (see below).
+If the revision can be returned (very) quickly, it might be possible to still make the special issue, but otherwise, the paper would appear separately.
+
+Reviewer 1 would like to see development of the finding that ancestral blocks are longer around a selected locus into a practical test.
+I think that this would be a substantial project for the future, and that it is valuable to first explain and understand the theoretical prediction, as is done here.
+Those analysing actual data can of course look for this pattern without needing an already implemented test.
+
+Rev.
+1 is also concerned that only under dominance is simulated.
+However, it is clear that any form of selection that gives the same marginal selection coefficient will give essentially the same effect: a large body of work on clines shows that different selective mechanisms are indistinguishable, and it is reasonable for the authors to rely on that.
+Theother main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).
+I don't know of a reference where they have been written in exactly this form, but the recombination terms and the diffusion terms in Eq.
+4 both go back a long way.
+
+Rev 2 is mainly concerned about the fit between theory and simulations.
+I am not worried about this, since the deterministic mathematics stands independently, and the numerical computations are straightforward.
+The stochastic simulations are useful mainly as illustrations of the process.
+Theory and simulations are compared in the supplementary material; it should be stated more clearly how closely these match.
+(A precise fit is not expected anyway because there is substantial drift).
+Rev.
+2 also worries that the simulations are over a rather narrow range; I suspect that the authors should have results that address sensitivity to the boundaries.
+Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that he authors suggest otherwise.
+
+Sincerely,
+
+Prof. Nick Barton
+Subject Editor, Molecular Ecology
+
+AE comments
+
+This is a nice analysis of how neutral genome introgresses past a cline maintained by selection against heterozygotes.
+There has been a good deal of work on related problems, but mostly assuming equilibrium, and mostly on the forwards evolution of allele frequencies, rather than tracing back the ancestry of blocks of genome.
+The theory presented here is therefore new, and directly relevant to interpreting genomes sampled from hybrid zones.
+
+Fig 1 caption - "alelle"
+\alisa{fixed}
+
+116 - "s and sigma are small" - but relative to what?  This is ot at all obvious.
+\alisa{I think this is relative to $N$ and $\tau$ respectively?}
+
+156 - Surely if time is rescaled as wel as distance and genome length, everything should look the same?  All the results, and to a good approximation the simulations, are in the 
+continuous regime where we can rescale.
+\alisa{I think our point is that these are different for the same scaling of $\tau$?}
+
+- A comment on two dimensions would be helpful.  Ancestral lineages simply diffuse transversely at a rate sigma, independent of their motion parallel to the cline.
+\alisa{Not sure where this would go? Maybe in the discussion? Do we just say that we ignore two dimensions and this is an ok approximation?}
+
+240 - This is a discrete deme model, so dispersal cannt follow a continuouus Gaussian.
+I guess what is happening is that sigma is chosen larger relative to deme spacing.
+However, there is some approximation here, and the approach will fail if sigma is small relatve to deme spacing.
+\alisa{I think that what we have is ok? Of course if $\sigma$ is tiny then maybe not, but obviously we have chosen the deme spacing to work with our choice of $\sigma$}
+
+299 - Missing semicolon?
+\alisa{fixed}
+
+420 - "are have"
+\alisa{fixed}
+
+495 - "reality" means "simulation" here??
+\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
+
+- Some appendices might be transferred to supplementary material - the derivations are fairly standard.
+\alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology}
+
+Reviewer Comments to Author:
+Reviewer: 1
+
+Recommendation: Reject, encourage resubmission
+
+Comments:
+
+This manuscript presents a discussion of patterns of linkage and allele frequency clines in a hybrid zone between two populations, under a scenario of mild hybrid incompatibility.
+It presents deterministic equations for the distribution of ancestral haplotypes along the genome and argues that the lengths of ancestry segments can be used to infer selection.
+
+The idea of using the lengths of ancestry blocks to detect selection in hybrid zones is interesting, and I had not encountered it before.
+This article will likely be of most interest to theorists, however, as the discussion of the practical confounders of the selection test is limited.
+A serious limitation of the test is that it only detects selection against heterozygotes.
+The claim that this would not affect results, on l.453-456:
+
+"Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models"
+
+is unconvincing.
+
+\alisa{We agree with the comments of the AE on the matter, and have added some additional text to the introduction that hopefully convinces that this is ok.}
+
+From a theoretical perspective, the article makes modest but valid contributions, and  Equation 4 and 5 are potentially interesting extensions to equation (1).
+The practical benefit of using these equations rather than forward simulations appears limited.
+
+\alisa{This is a point we have considered ourselves.
+We present these equations as a way of building our model up to address haplotype lengths.
+Both analytical expressions and simulations are presented so that they each support the conclusions of the other.}
+
+While reading the manuscript, I found it difficult to distinguish the most important observations from more tangential observations.
+The manuscript reads like a long list of observations, and would benefit from more focus on the main findings and more structure.
+\alisa{I have no good comment for this right now, but am trying to edit the manuscript to address this.}
+
+Specific points:
+1-
+Since the recombination-based terms in Eqs 4 and 5 do not depend on the spatially explicit part of the model or on the selection term, I would be surprised if they had not been derived somewhere.
+
+2-
+Eq5:
+If $a<b<\theta<0$, then I interpret $gtheta,b) = 0$. If correct, this should be specified.  Alternately, the bounds of the integral could be modified to account for this situation.
+\alisa{I don't think that this is true? I think that the situation described here is accounted for by the other terms?}
+
+3-
+As an example of a section that felt unstructured, consider the discussion starting at l.347.
+This long paragraph starts with a long list of observations, without giving a hint as to where we are going.
+Beginning the paragraph with the main idea for the paragraph would be useful.
+\alisa{Working on how to fix this in the ms now}
+
+4-
+Minor issues:
+
+"In the notation above, $qz(x, t, r) = Px{Zt = A}$. "
+I did not see the notation $P^x$ above, and the RHS does not depend on small z.
+\alisa{(This is at l180) Should the $=$ be a $:=$}
+
+l.120 As we show below the solutions provide an good approximation
+\alisa{fixed}
+
+l.144: It seems like the x+r and x-r should have the same sign.
+\alisa{I think may be correct?}
+l 199:
+
+$g_z(x, T, a,b)$ is defined as: " We are interested in the probability gz(x, t; a, b) of finding an entire segment (a, b) of ancestry A in an individual having selected allele of type z, ... "
+
+This definition does not make it clear that the probability is conditional on z; P(A|z) vs P(A,z).
+
+\alisa{I dont't actually agree with this... thoughts?}
+
+l.606: Reference issues
+\alisa{This is a reference that is expected to appear in the same issue, so we do not have page numbers. The hope is that this can be filled in during final proofs for the issue}
+
+Additional Questions:
+By selecting ?Yes?
+you are opting in to the Publons service and data about this review (including your name and the review itself) will be transferred to Publons.
+You may opt-out of the service at any time.
+: No
+
+Quality of Science: Experimentally and/or theoretically excellent reliable data, no fatal flaws
+
+Importance of Science: Research addresses a consequential question in ecology, evolution, behaviour, or conservation
+
+Quality of Presentation: Ideas and methods mostly clear, but grammar and/or spelling is poor, format does not follow guidelines, and /or there is redundancy between sections
+
+Does this manuscript require significant reduction in length? If 'Yes", please indicate where shortening is required in your specific comments: Yes
+
+Does the Data Accessibility section list all the datasets needed to recreate the results in the manuscript?
+If ?No?, please specify which additional data are needed in your comments to the authors.: Yes
+
+Reviewer: 2
+
+Recommendation: Reject, encourage resubmission
+
+Comments:
+REVIEW:
+
+In this paper, the authors extend previous work on the diffusion of neutral alleles in a secondary-contact zone to include selection against hybrids.
+Building on the classic diffusion model for the dynamics of a cline at a selected locus, they formalize the dynamics of linked neutral loci and investigate the distribution of ancestry block-lengths around the selected locus.
+
+The authors obtain numerical results for the flattening of the cline at a neutral locus linked to an under-dominant selected locus in the absence of genetic drift.
+They also outline scaling arguments for the rates of movement and length of the introgressing blocks around the selected locus.
+Simulations are set up to test the theoretical predictions and illustrate other interesting patterns such as the distribution of haplotype (ancestry block?) lengths.
+
+Unfortunately, the fit of the theoretical predictions to simulations is largely not shown, and those comparisons in the manuscript are limited to the frequency dynamics of the neutral alleles (Figure S3).
+The predictive theoretical results are based on scaling arguments - but it is not clear which of these results are uniquely derived in this paper.
+
+\alisa{Our hope was that the simulations and theory would provide qualitative support for one another.
+Some predictions made by the theory e.g.
+time and spatial scales do seem to correspond.
+We will try to do a better job of signposting our unique contributions??}
+
+The authors also aim to characterize the movement of the genomic block that contains a selected locus through a hybrid zone.
+Yet, the size of the simulated habitat barely covers the cline width, and the simulation results may be sensitive to the boundaries.
+
+\alisa{Should we be running more sims? Nick seems to think we have done this, but I don't think we have.}
+
+Overall, this MS presents some interesting insights, and the formalization of the block dynamics can be useful.
+However, I think that more work is needed to explain and test (by simulations) the novel predictions.
+Currently, the theory and the matching simulations are largely presented separately.
+
+Detailed comments:
+
+Figure 1: Plotting the haplotypes to the side of the plot would help.
+\alisa{We have remade Figure 1, with additional information that we hope will help with interpretation}
+
+Simulations;
+l.
+245.
+As the theory is set up in the absence of genetic drift, and, correspondingly, the local population size is large, it is unclear how a deme with 500 individuals could become empty in a Wright-Fisher model.
+The (narrow) width of the habitat also needs justification.
+\alisa{It is true that the demes are never empty. However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}
+
+Figure 3:
+What are the replicate lines for the same ?r??
+Why re the lines for r = 0 all slightly displaced to the right?
+Also, it would appear that the color scheme in associated Figure S2 (bottom) is mismatched with the legend, shifted to the red part of the spectrum.
+After fixing this, I think that the Figure S2 should be in the main text instead of Fig.
+3.
+
+\alisa{We believe that the displacement is likely due to drift. (We could do a repeated sim to show this?)}
+
+Results
+l.
+299 Total population size is irrelevant when predicting the effects of genetic drift (besides, some simulations only have 50 demes across, not 100).
+In a one-dimensional habitat, one uses N sigma sqrt(s) to judge whether the effect of genetic drift is likely to be negligible.
+\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}
+
+Scaling arguments, p.
+14, 15: Please give a citation with the known scaling, and take more care with the novel ones.
+l.339: I think the sentence about the maximum length of the enclosing haplotype, sigma sqrt(s) / x is mis-phrased: presumably this should be a statement on the distribution (expected size?) of the introgressing block length at distance x, as a lucky block can travel many sigmas in one generation?
+
+\alisa{I don't think this is misphrased?}
+
+Discussion:
+It is important to make the reader aware that even neutral clines do not just keep flattening ?
+genetic drift leads to a finite cline width even for neutral clines.
+Despite the occasional mention of genetic drift, the MS does not clearly state that the parameter space is chosen such that genetic drift is weak and does not indicate when this is a reasonable assumption (see comment on l.
+299).
+\alisa{We should add a line or two in the discussion about drift.
+However, IBD is perhaps not relevant over the timescales we are considering and is not necessarily ancestry informative, and since we are not interested in the equilibrium state, it is not clear that we need to worry about this limit.
+}
+
+p.27 rates, e.g.
+Eq.
+(7,8 ): I find the explanation of the rates a bit unclear.
+I understand that the recombinant AB can arise either by A reproducing at rate $s_A$ and recombining with B, or B reproducing at rate $s_B$ and recombining with A ?
+such that the neutral locus of the type A is kept.
+Or alternatively, the BA recombinant arises, where the neutral locus at B is kept.
+Which one of the two happens has a probability of 1/2.
+\alisa{We'll work on making this more understandable.}
+
+The "" Data Accessibility section" is missing.
+The folder at ....
+appears to be a working folder with lots of other files and the r-scripts hidden under a ?sim?
+directory.
+The README file appears insufficient to allow orientation in the scripts, and worryingly, it has an ?oh, wrong recombination distance?
+note to it.
+\alisa{Ok. This probably needs to be fixed.}
+
+The Acknowledgements section is missing.
+\alisa{I had noticed this. This will be added.}
+
+Additional Questions:
+Do you want to get credit for reviewing this manuscript on Publons?
+By selecting ?Yes?
+you are opting in to the Publons service and data about this review (including your name and the review itself) will be transferred to Publons.
+You may opt-out of the service at any time.
+: Yes
+
+Quality of Science: Mostly competent, but suffering from flaws of a technical or analytical nature
+
+Importance of Science: Research addresses a consequential question in ecology, evolution, behaviour, or conservation
+
+Quality of Presentation: Writing is clear, methods and data analyses are transparent, ideas make sense, proper grammar and spelling is employed, redundancy is avoided
+
+Does this manuscript require significant reduction in length? If 'Yes", please indicate where shortening is required in your specific comments: No
+
+Does the Data Accessibility section list all the datasets needed to recreate the results in the manuscript?
+If ?No?, please specify which additional data are needed in your comments to the authors.: No
+
 %%%%% %%%%%% %%%%%%%
@@
 
-Much of the process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
+The process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
 % Gene flow between formerly isolated populations is possible if reproductive isolation is not complete.
@@
 % previously had wrong Barton '79 here
-Studies of hybrid zones have primarily been based on single-site statistics such as clines,
-but today, 
-the availability of dense genotype data in combination with linkage information 
+Studies of hybrid zones have primarily been based on single-site statistics (e.g. clines),
+but the recent availability of dense genotype data in combination with linkage information 
 makes it now possible to use patterns of the block length distribution
@@
 There is substantial theoretical work on genetic clines.
-We study those maintained only by selection against hybrids -- \emph{tension zones} --
+We focus on those maintained only by selection against hybrids -- \emph{tension zones} --
 as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
@@
 the fitness of the heterozygote will likely determine the shape of the cline,
-and \citet{barton1986barrier} show that the shapes of several sorts of stable clines are similar.
+and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
 There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
-and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
+and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000} \alisa{and much of this focuses on equilibrium states}.
 In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
@@
 Although the selected cline establishes quickly, transient patterns can persist for quite some time,
-and haplotypes can potentially carry substantial information about the formation of recent zones.
+and \alisa{ancestry} haplotypes can potentially carry substantial information about the formation of recent zones.
 To do this, we take the reverse-time perspective,
 understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
-This is particularly suited to understanding transient patterns,
-although because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
+This is particularly suited to understanding transient patterns 
+although, because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
 incorporating genetic drift is not straightforward.
@@
 In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has be used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Hellenthal2014}.
+If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
 In addition to facilitating demographic inference, heterogeneous patterns of ancestry across the genome can help identify putative targets of selection in hybridizing populations  \citep{Porter1997, Gompert2012}.
@@
 This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes.
+as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological (e.g. local adaptation) }
 
@@
 \begin{figure}
-\includegraphics{figs/BM_schematic.pdf}
+\includegraphics[clip,width=\textwidth,trim=0cm 0cm 0cm 0.05cm]{figs/BM_schematic_new2.pdf}
 \caption{
@@
 that are helpful in providing intuition and for establishing order-of-magnitude estimates,
-and analytics, mostly in terms of partial differential equations needing numerical solution.
+and analytics, mostly in terms of partial differential equations needing numerical solution. \alisa{For simplicity, we ignore coalescence.}
 
@@
 First, we turn our attention to how the cline at the selected locus itself is formed. 
+\alisa{This has been previously described, but is presented here as it provides a framework for further development of our model}
 Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 for discussion of diploidy see \citet{christiansen1995genotypic}.
-As we show below the solutions provide an good approximation across the range of realistic values of $s$.
+As we show below the solutions provide a good approximation across the range of realistic values of $s$.
 Although there is no known exact analytic solution to this equation, 
@@
 \paragraph{Lineages at the selected locus}
-Now suppose that the frequency profile of the selected allele, $p(x,t)$, is known.
+Now suppose that the frequency profile of the selected allele, $p(x,t)$, is known as a function of time and space.
 We seek to first understand how lineages of sampled individuals behave at the selected locus,
@@
 (Lineages are also biased towards regions with higher fecundity,
-and thus away from the hybrid zone, but we assume that $s$ is small and ignore this.)
+and thus away from the hybrid zone, but we assume that $s$ is small and ignore this.) \alisa{could this reference to small $s$ be confusing?}
 
@@
 % with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
-(i.e., it moves at speed $\sigma^2$ in the potential $-\log p$).
+(i.e., it moves at speed $\sigma^2$ in the potential $-\log p$). \alisa{What does potential here mean?}
 % \alisa{note to self to read through appendix when written}
@@
 % $$ = \E[ -R ] - \E[R]^2 p'(x)/p(x) + \E[R^2] p'(x)/p(x) + O(R^3) $$
-Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$
+Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$ \alisa{perhaps a brief explanation of how this works}
 and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
@@
 \end{align}
+\alisa{Explain in words}
 In the first expression, giving the distribution of how the lineage location changes,
@@
 \end{align}
+\alisa{Use our words}
 with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
-and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
+and where $\grad$ is the gradient, and $\Delta$ is the Laplacian \alisa{describing the rate of change}. 
 The three terms in these equations come from Brownian movement of a linages,
@@
 \paragraph{Numerical computation}
-To determine how the cline at a linked locus is expected to relax,
+To determine how the cline at a linked locus is expected to relax, \alisa{and to confirm the behavior of our model to previously established results,}
 we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically.
@@
 \paragraph{Haplotype lengths}
-We now find the frequency at which \emph{entire} blocks of genome (haplotypes) 
+We now develop our model further to find the frequency at which \emph{entire} blocks of genome (haplotypes) 
 of a single ancestry are found, at a given location and time.
@@
 \end{align}
+
+\alisa{Need to be related to Figure 1}
 Here the first term ($G^A g_A$) represents spatial mixing,
@@
 and the remaining portion becomes linked to a $B$ allele
-(and we say that the integral $\int_a^{\min(0,b)}$ is zero if $a>0$ 
-and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$).
+(\alisa{to consider cases where the selected locus lies outside the interval [a,b], i.e.,} we say that the integral $\int_a^{\min(0,b)}$ is zero if  $a>0$ 
+and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$) .
 The last integral results from recombination inside the block in homozygotes for $A$,
@@
 but linked clines are still flattening.
-After a longer period ($T=1000$),
+After a longer period ($T=1000$), \alisa{I see, this is what reviewer was saying about flatness, but shouldn't our sims produce IBD?}
 clines over much of the chromosome are flat
@@
 This is seen in Figure \ref{Fig:meanBlockLengthsUnconditioned}.
-However, we expect that stretches of $A$ ancestry containing a selected site will be longer
+However, we expect that stretches of $A$ ancestry \emph{containing a selected site} will be longer
 than those that do not contain the selected site at the same spatial location,
@@
 (compared to $x^2/\sigma^2$ for a neutral allele).
-This implies the enclosing $A$ haplotype should be no longer than $\sigma\sqrt{s}/x$.
+This implies the enclosing $A$ haplotype surrounding the locus should be no longer than $\sigma\sqrt{s}/x$.
 %\plr{(check predicted decay of mean length $1/x$ in simulations?)}
@@
 
-Genomic regions are inherited from ancestors $T$ generations ago
+\alisa{To provide some intuition about the physical scale of this signal, we note that}
+genomic regions are inherited from ancestors $T$ generations ago
 in blocks of size roughly $1/T$.
@@
 The mean haplotype length found without conditioning on ancestry, $l(x,m)$,
-shows a smaller increase near the selected locus, because most blocks will be of the locally common type, and so do not trace back to regions of different block lengths.
-This is shown in Figures \ref{Fig:meanBlockLengthsUnconditioned}, \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}.
+shows a smaller increase near the selected locus, because most blocks will be of the locally common type, and so do not trace back to regions of different block lengths (Figures \ref{Fig:meanBlockLengthsUnconditioned}, \ref{Supp:blockLengthHeatmapNoAnc} and \ref{Supp:blockLengthNoAnc}).
 
@@
 Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
-Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $C(k)$, could provide a simple and informative summary. 
+Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage --  and similarly $C(k)$, could provide a simple and informative summary. 
 The ratio of mean adjacent block lengths $C(x,m)$ shows a very sharp peak at the selected site,
@@
 cannot introgress until recombination separates them;
-our model describes at least roughly how quickly this should happen.
+our model \alisa{provides some rough expectation on how quickly this should happen}.
 
@@
 %Doing so requires an understanding of how positive selection shapes ancestry block lengths in these cases.
+
+\section*{Acknowledgements}
+We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Nick Barton and two anonymous reviewers for helpful feedback and suggestions. \alisa{More info here}
 

### 2016-04-22 drafts/review-responses-april-2016.tex
moved review responses to review-respones-april-2106.tex

% \documentclass{article}
% \usepackage{fullpage}
% \input{review-response-commands}
% 
% \begin{document}

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it Molecular Ecology}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Alisa's address\\
    Yaniv's address\\
    \emph{and} Molecular and Computational Biology\\
    University of Southern California \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We did all the things, thanks.

\begin{quote}
Reviewer 1 would like to see development of the finding that ancestral blocks are longer around a selected locus into a practical test.  I think that this would be a substantial project for the future, and that it is valuable to first explain and understand the theoretical prediction, as is done here.  Those analysing actual data can of course look for this pattern without needing an already implemented test.
\end{quote}

\begin{quote}
Rev. 1 is also concerned that only under dominance is simulated.  However, it is clear that any form of selection that gives the same marginal selection coefficient will give essentially the same effect: a large body of work on clines shows that different selective mechanisms are indistinguishable, and it is reasonable for the authors to rely on that.  Theother main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).  I don't know of a reference where they have been written in exactly this form, but the recombination terms and the diffusion terms in Eq. 4 both go back a long way.
\end{quote}

\begin{quote}
Rev 2 is mainly concerned about the fit between theory and simulations.  I am not worried about this, since the deterministic mathematics stands independently, and the numerical computations are straightforward.  The stochastic simulations are useful mainly as illustrations of the process.  Theory and simulations are compared in the supplementary material; it should be stated more clearly how closely these match. (A precise fit is not expected anyway because there is substantial drift).  Rev. 2 also worries that the simulations are over a rather narrow range; I suspect that the authors should have results that address sensitivity to the boundaries.  Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that he authors suggest otherwise.
\end{quote}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Alisa Sedghifar,
Yaniv Brandvain, and
Peter Ralph
}\\
\end{minipage}

\vskip 4em

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
This is a nice analysis of how neutral genome introgresses past a cline maintained by selection against heterozygotes.  There has been a good deal of work on related problems, but mostly assuming equilibrium, and mostly on the forwards evolution of allele frequencies, rather than tracing back the ancestry of blocks of genome.  The theory presented here is therefore new, and directly relevant to interpreting genomes sampled from hybrid zones.
\end{quote}

\begin{point}{Fig 1 caption}
``alelle''
\end{point}

\reply
Fixed.

\begin{point}{l.116}
``s and sigma are small'' - but relative to what?  This is not at all obvious.
\end{point}

\reply
\alisa{I think this is relative to $N$ and $\tau$ respectively?}

\begin{point}{l.156}
Surely if time is rescaled as wel as distance and genome length, everything should look the same?  All the results, and to a good approximation the simulations, are in the continuous regime where we can rescale.
\end{point}

\reply
\alisa{I think our point is that these are different for the same scaling of $\tau$?}

\begin{point}{}
 A comment on two dimensions would be helpful.  Ancestral lineages simply diffuse transversely at a rate sigma, independent of their motion parallel to the cline.
\end{point}

\reply
\alisa{Not sure where this would go? Maybe in the discussion? Do we just say that we ignore two dimensions and this is an ok approximation?}

\begin{point}{l.240}
This is a discrete deme model, so dispersal cannt follow a continuouus Gaussian. I guess what is happening is that sigma is chosen larger relative to deme spacing.  However, there is some approximation here, and the approach will fail if sigma is small relatve to deme spacing.
\end{point}

\reply
\alisa{I think that what we have is ok? Of course if $\sigma$ is tiny then maybe not, but obviously we have chosen the deme spacing to work with our choice of $\sigma$}

\begin{point}{l.299}
Missing semicolon?
\end{point}

\reply
Fixed.

\begin{point}{l.420}
``are have''
\end{point}

\reply
Fixed.

\begin{point}{l.495}
``reality'' means ``simulation'' here??
\end{point}

\reply
\alisa{I think we do mean ``reality" here, but alternate opinions welcome}

\begin{point}{}
 Some appendices might be transferred to supplementary material -- the derivations are fairly standard.
\end{point}

\reply
\alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology}

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
    This manuscript presents a discussion of patterns of linkage and allele frequency clines in a hybrid zone between two populations, under a scenario of mild hybrid incompatibility. It presents deterministic equations for the distribution of ancestral haplotypes along the genome and argues that the lengths of ancestry segments can be used to infer selection.

    The idea of using the lengths of ancestry blocks to detect selection in hybrid zones is interesting, and I had not encountered it before. This article will likely be of most interest to theorists, however, as the discussion of the practical confounders of the selection test is limited. A serious limitation of the test is that it only detects selection against heterozygotes. The claim that this would not affect results, on l.453-456:
    \textit{``Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models''}
    is unconvincing.
\end{quote}

\alisa{We agree with the comments of the AE on the matter, and have added some additional text to the introduction that hopefully convinces that this is ok.}

\begin{quote}
    From a theoretical perspective, the article makes modest but valid contributions, and  Equation 4 and 5 are potentially interesting extensions to equation (1). The practical benefit of using these equations rather than forward simulations appears limited.
\end{quote}

\alisa{This is a point we have considered ourselves. We present these equations as a way of building our model up to address haplotype lengths. Both analytical expressions and simulations are presented so that they each support the conclusions of the other.}

\begin{quote}
    While reading the manuscript, I found it difficult to distinguish the most important observations from more tangential observations. The manuscript reads like a long list of observations, and would benefit from more focus on the main findings and more structure.
\end{quote}

\begin{point}{}
Since the recombination-based terms in Eqs 4 and 5 do not depend on the spatially explicit part of the model or on the selection term, I would be surprised if they had not been derived somewhere.
\end{point}

\reply

\begin{point}{Eq.5}
If $a<b<theta<0$, then I interpret $g(theta,b) = 0$. If correct, this should be specified.  Alternately, the bounds of the integral could be modified to account for this situation.
\end{point}

\reply

\begin{point}{}
As an example of a section that felt unstructured, consider the discussion starting at l.347.  This long paragraph starts with a long list of observations, without giving a hint as to where we are going. Beginning the paragraph with the main idea for the paragraph would be useful.
\end{point}

\reply
\alisa{Working on how to fix this in the ms now}

\begin{point}{``In the notation above, $q_z(x, t, r) = P^x{Z_t = A}$.''}
I did not see the notation $P^x$ above, and the RHS does not depend on small z.
\end{point}

\reply
\alisa{(This is at l180) Should the $=$ be a $:=$}

\begin{point}{l.120} 
As we show below the solutions provide an good approximation
\end{point}

\reply

\begin{point}{l.144}
It seems like the x+r and x-r should have the same sign.
\end{point}

\begin{point}{l.199}
$g_z(x, T, a,b)$ is defined as: \textit{``We are interested in the probability $g_z(x, t; a, b)$ of finding an entire segment $(a, b)$ of ancestry $A$ in an individual having selected allele of type $z$, \ldots ''}

This definition does not make it clear that the probability is conditional on $z$; $\P(A|z)$ vs $\P(A,z)$.
\end{point}

\begin{point}{l.606}
Reference issues
\end{point}

\reply
\alisa{This is a reference that is expected to appear in the same issue, so we do not have page numbers. The hope is that this can be filled in during final proofs for the issue}

% Quality of Science: Experimentally and/or theoretically excellent reliable data, no fatal flaws
% 
% Importance of Science: Research addresses a consequential question in ecology, evolution, behaviour, or conservation
% 
% Quality of Presentation: Ideas and methods mostly clear, but grammar and/or spelling is poor, format does not follow guidelines, and /or there is redundancy between sections
% 
% Does this manuscript require significant reduction in length? If 'Yes', please indicate where shortening is required in your specific comments: Yes
% 
% Does the Data Accessibility section list all the datasets needed to recreate the results in the manuscript? If `No', please specify which additional data are needed in your comments to the authors.: Yes

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
    In this paper, the authors extend previous work on the diffusion of neutral alleles in a secondary-contact zone to include selection against hybrids. Building on the classic diffusion model for the dynamics of a cline at a selected locus, they formalize the dynamics of linked neutral loci and investigate the distribution of ancestry block-lengths around the selected locus.

    The authors obtain numerical results for the flattening of the cline at a neutral locus linked to an under-dominant selected locus in the absence of genetic drift. They also outline scaling arguments for the rates of movement and length of the introgressing blocks around the selected locus. Simulations are set up to test the theoretical predictions and illustrate other interesting patterns such as the distribution of haplotype (ancestry block?) lengths.

    Unfortunately, the fit of the theoretical predictions to simulations is largely not shown, and those comparisons in the manuscript are limited to the frequency dynamics of the neutral alleles (Figure S3). The predictive theoretical results are based on scaling arguments - but it is not clear which of these results are uniquely derived in this paper.
\end{quote}

\alisa{Our hope was that the simulations and theory would provide qualitative support for one another. Some predictions made by the theory e.g. time and spatial scales do seem to correspond. We will try to do a better job of signposting our unique contributions??}

\begin{quote}
    The authors also aim to characterize the movement of the genomic block that contains a selected locus through a hybrid zone. Yet, the size of the simulated habitat barely covers the cline width, and the simulation results may be sensitive to the boundaries.
\end{quote}

\alisa{Should we be running more sims? Nick seems to think we have done this, but I don't think we have.}

\begin{quote}
Overall, this MS presents some interesting insights, and the formalization of the block dynamics can be useful. However, I think that more work is needed to explain and test (by simulations) the novel predictions. Currently, the theory and the matching simulations are largely presented separately.
\end{quote}

\begin{point}{Figure 1}
 Plotting the haplotypes to the side of the plot would help.
\end{point}

\reply
\alisa{We have remade Figure 1, with additional information that we hope will help with interpretation}

\begin{point}{l. 245}
 As the theory is set up in the absence of genetic drift, and, correspondingly, the local population size is large, it is unclear how a deme with 500 individuals could become empty in a Wright-Fisher model.  The (narrow) width of the habitat also needs justification.
\end{point}

\reply
\alisa{It is true that the demes are never empty. However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}

\begin{point}{Figure 3:}
What are the replicate lines for the same $r$? Why re the lines for $r = 0$ all slightly displaced to the right? Also, it would appear that the color scheme in associated Figure S2 (bottom) is mismatched with the legend, shifted to the red part of the spectrum. After fixing this, I think that the Figure S2 should be in the main text instead of Fig. 3.
\end{point}

\reply
\alisa{We believe that the displacement is likely due to drift. (We could do a repeated sim to show this?)}

\begin{point}{l. 299}
Total population size is irrelevant when predicting the effects of genetic drift (besides, some simulations only have 50 demes across, not 100). In a one-dimensional habitat, one uses $N \sigma \sqrt{s}$ to judge whether the effect of genetic drift is likely to be negligible.
\end{point}

\reply
\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}

\begin{point}{Scaling arguments, p. 14, 15:}
 Please give a citation with the known scaling, and take more care with the novel ones.  
\end{point}

\reply
\alisa{I don't think this is misphrased?}

\begin{point}{l.339}
 I think the sentence about the maximum length of the enclosing haplotype, $\sigma \sqrt{s} / x$ is mis-phrased: presumably this should be a statement on the distribution (expected size?) of the introgressing block length at distance x, as a lucky block can travel many sigmas in one generation?
\end{point}

\reply

\begin{point}{Discussion:}
It is important to make the reader aware that even neutral clines do not just keep flattening -- genetic drift leads to a finite cline width even for neutral clines. Despite the occasional mention of genetic drift, the MS does not clearly state that the parameter space is chosen such that genetic drift is weak and does not indicate when this is a reasonable assumption (see comment on l. 299).
\end{point}

\reply
\alisa{We should add a line or two in the discussion about drift. However, IBD is perhaps not relevant over the timescales we are considering and is not necessarily ancestry informative, and since we are not interested in the equilibrium state, it is not clear that we need to worry about this limit. }

\begin{point}{p.27 rates, e.g. Eq. (7,8):}
 I find the explanation of the rates a bit unclear. I understand that the recombinant AB can arise either by A reproducing at rate $s_A$ and recombining with B, or B reproducing at rate $s_B$ and recombining with A -- such that the neutral locus of the type A is kept. Or alternatively, the BA recombinant arises, where the neutral locus at B is kept. Which one of the two happens has a probability of 1/2.
\end{point}

\reply
\alisa{We'll work on making this more understandable.}

\begin{point}
The ``Data Accessibility section'' is missing. The folder at \href{http://github.com_petrelharp_clinal-lineages}  appears to be a working folder with lots of other files and the r-scripts hidden under a `sim' directory. The README file appears insufficient to allow orientation in the scripts, and worryingly, it has an `oh, wrong recombination distance' note to it.
\end{point}

\reply
\alisa{Ok. This probably needs to be fixed.}

\begin{point}
The Acknowledgements section is missing.
\end{point}

\reply

% Quality of Science: Mostly competent, but suffering from flaws of a technical or analytical nature
% 
% Importance of Science: Research addresses a consequential question in ecology, evolution, behaviour, or conservation
% 
% Quality of Presentation: Writing is clear, methods and data analyses are transparent, ideas make sense, proper grammar and spelling is employed, redundancy is avoided
% 
% Does this manuscript require significant reduction in length? If 'Yes", please indicate where shortening is required in your specific comments: No
% 
% Does the Data Accessibility section list all the datasets needed to recreate the results in the manuscript? If `No', please specify which additional data are needed in your comments to the authors.: No

% \end{document}

### 2016-04-22 drafts/hybrid_zone.tex
moved review responses to review-respones-april-2106.tex

--- 2016-04-21 drafts/hybrid_zone.tex
@@
 \renewcommand{\and}{\,\&\,}
-
-% other title ideas:
-% {Blocks of extended ancestry around selected loci in hybrid zones}
-\title{Beyond clines: lineages and haplotype blocks in hybrid zones}
-\author{}
-\date{} % delete this line to display the current date
 
@@
 
+% Responses to reviews
+\input{review-response-commands}
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
+\newcommand{\responsefile}{review-responses-april-2016.tex}  % name of the review reponses file
+
+% other title ideas:
+% {Blocks of extended ancestry around selected loci in hybrid zones}
+\title{Beyond clines: lineages and haplotype blocks in hybrid zones}
+\author{}
+\date{} % delete this line to display the current date
+
 \usepackage{authblk}
@@
 
+%% response to reviews
+% to remove these, change 'ifreviewresponses' up top
+\includereviews
+
 \end{document}

### 2016-04-22 drafts/review-responses-april-2016.tex
went through AE comments

@@
 % \begin{document}
+
+%%%%%%
+%%
+%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
+%%
+%%%%%
 
@@
 
-\begin{point}{l.116}
-``s and sigma are small'' - but relative to what?  This is not at all obvious.
+\begin{point}{\revref}
+``$s$ and $\sigma$ are small'' - but relative to what?  This is not at all obvious.
 \end{point}
@@
 \reply
-\alisa{I think this is relative to $N$ and $\tau$ respectively?}
+We just meant that in the rescaling argument for a discrete model this is required formally,
+without giving it enough thought.
+We've added more thoughtful discussion.
 
-\begin{point}{l.156}
-Surely if time is rescaled as wel as distance and genome length, everything should look the same?
+\begin{point}{\revref}
+Surely if time is rescaled as well as distance and genome length, everything should look the same?  
 All the results, and to a good approximation the simulations, are in the continuous regime where we can rescale.
@@
 \reply
-\alisa{I think our point is that these are different for the same scaling of $\tau$?}
+Good point; we in fact outlined this rescaling at \llname{ll:rescaling}.
+It still seems fair to say there is ``more information'' here, because we can pick up the difference in time scaling
+using haplotypes and linkage;
+we've edited this paragraph to make the point better.
 
 \begin{point}{}
- A comment on two dimensions would be helpful.  Ancestral lineages simply diffuse transversely at a rate sigma, independent of their motion parallel to the cline.
+A comment on two dimensions would be helpful.  
+Ancestral lineages simply diffuse transversely at a rate $\sigma$, independent of their motion parallel to the cline.
 \end{point}
@@
 \reply
-\alisa{Not sure where this would go? Maybe in the discussion? Do we just say that we ignore two dimensions and this is an ok approximation?}
+We've added a short note on this (with more references) \revref.
 
-\begin{point}{l.240}
-This is a discrete deme model, so dispersal cannt follow a continuouus Gaussian.
-I guess what is happening is that sigma is chosen larger relative to deme spacing.
-However, there is some approximation here, and the approach will fail if sigma is small relatve to deme spacing.
+\begin{point}{\revref}
+This is a discrete deme model, so dispersal can't follow a continuous Gaussian. 
+I guess what is happening is that $\sigma$ is chosen larger relative to deme spacing.  
+However, there is some approximation here, and the approach will fail if $\sigma$ is small relatve to deme spacing.
 \end{point}
@@
 \reply
-\alisa{I think that what we have is ok? Of course if $\sigma$ is tiny then maybe not, but obviously we have chosen the deme spacing to work with our choice of $\sigma$}
+There might be some confusion about our use of the floor function, which we now explicitly explain.
+It's a good point more generally that if $\sigma$ is smaller than the deme spacing this scheme won't get dispersal with $\sigma$ close to the $\sigma$ that we plug into the Gaussian;
+it's for this reason that we compute numerically what $\sigma$ actually is when comparing simulation to theory.
 
-\begin{point}{l.299}
+\begin{point}{}
 Missing semicolon?
@@
 
-\begin{point}{l.420}
+\begin{point}{}
 ``are have''
@@
 
-\begin{point}{l.495}
+\begin{point}{\revref}
 ``reality'' means ``simulation'' here??
@@
 \reply
+Well, we hope to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted the wording.
 \alisa{I think we do mean ``reality" here, but alternate opinions welcome}
+\plr{I changed to ``realistic, individual-based models''.}
 
@@
 \alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology}
+\plr{I'm fine with moving some; not super clear on the difference anyhow...}
 

### 2016-04-22 drafts/hybrid_zone.tex
went through AE comments

@@
 or the discrete analogues \citep{hanson1966effects}.
+
+\plr{Need to cite Durrett and Cox's stuff here, especially \citet{durrett2007width}.}
 
@@
 (ii) habitat and dispersal are homogeneous,
-(iii) $s$ and $\sigma$ are small, \alisa{relative to $N$ and $\tau$ respectively?}
-and (iv) population density is large,
+and (iii) population density is large,
+% and (iv) $s$ and $\sigma$ are small, \alisa{relative to $N$ and $\tau$ respectively?}
 then as in \citet{Bazykin1969},
-the commonly-used equation that $p$ solves is
+the commonly-used equation that $p$ approximately solves is
 \begin{align} \label{eqn:cline_pde}
-    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) .
+    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
 \end{align}
-where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian.
-\plr{change $\dot p$ to $\frac{d}{dt}p$, and $\Delta$ to the second derivative?}
-This says that the local rate of change of the frequency of $A$ ($\dot p$)
+where $\Delta p$ is the Laplacian.
+\plr{change $\Delta$ to the second derivative?}
+The equation is approximate because it 
+omits terms of order $s^2$
+(so deviations from the prediction occur over $1/s^2$ generations),
+and relies on the Central Limit Theorem to approximate generation-to-generation dispersal by the Gaussian
+(so if dispersal is non-Gaussian, will fit best over longer time scales, say, tens of generations).  \revpoint{AE}{2}
+This says that the local rate of change of the frequency of $A$ ($\frac{d}{dt} p$)
 is equal to the net rate at which $A$ alleles arrive by local dispersal ($\sigma^2/2 \Delta p$)
@@
 the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
-Note that rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
-results in a dimensionless equation \citep{Slatkin1973},
+Rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
+results in a dimensionless equation \citep{Slatkin1973}, \linelabel{ll:rescaling}
 implying that the cline establishes over a timescale of the order $1/s$. 
@@
 despite being slowed down somewhat by selection against heterozygotes.
+
+For simplicity we mostly work with one-dimensional clines,
+but the mathematics applies as well to two-dimensional systems with some modifications:
+for instance, if the hybrid zone is a straight line, 
+the description above applies to motion of a lineage transverse to the hybrid zone.  \revpoint{AE}{4}
+It is for this reason that we use $\Delta$ for the second derivative and $\grad$ for the first derivative,
+as these carry over without change to two dimensions;
+in fact, if the landscape is heterogeneous or dispersal is non-Gaussian,
+the theory still holds replacing $\Delta$ by the appropriate operator.
+Note, however, that in two dimensions drift can have a much stronger effect \citep{cox1995hybrid,barton2010phylogeography,durrett2007width}.
 
@@
 
-Note that this description gives more information than the steady-state cline $p(x)$,
+This description gives more information than the steady-state cline $p(x)$,
 which depends only on $\sigma/\sqrt{s}$.
-Here we see that lineages with $\sigma=10$ and $s=.16$ move quite differently
-to lineages with $\sigma=1$ and $s=.0016$,
+Here we see that lineages with $\sigma=10$ and $s=.16$ move much faster
+than lineages with $\sigma=1$ and $s=.0016$,
 reflecting strong differences in selection against heterozygotes,
 even though the stable clines have the same form.
+Even though a rescaling of space and time as described above can make the models equivalent,
+this difference in lineage speed can be seen through the action of recombination, which we explore next.
+\revpoint{AE}{3}
 
@@
     \begin{aligned}  \label{eqn:q_pde}
-    \dot q_A(x,t,r) 
+    \frac{d}{dt} q_A(x,t,r) 
             &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
@@
                 r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
-    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
+    \frac{d}{dt} q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
             + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
@@
 \begin{align*}
-    \dot q_A &= G^A q_A + r (1-p) (q_B-q_A) \\
-    \dot q_B &= G^B q_B + r p (q_A-q_B) .
+    \frac{d}{dt} q_A &= G^A q_A + r (1-p) (q_B-q_A) \\
+    \frac{d}{dt} q_B &= G^B q_B + r p (q_A-q_B) .
 \end{align*}
@@
     \begin{aligned} \label{eqn:g_pde}
-        \dot g_A(a,b) 
+        \frac{d}{dt} g_A(a,b) 
         &= G^A g_A(a,b) + r(a,b) (1-p) (g_A(a,b)-g_B(a,b))
@@
 where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
-(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
+(Here $\lfloor z \rfloor$ is the largest integer less than or equal to $z$,
+so this equivalently takes a centered Gaussian displacement and assigns it to the nearest deme. \revpoint{AE}{5}
+The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
 Migrants past either end of the range remain at the terminal demes.
@@
 A main role of these solutions in our work has been to verify that theory based on the diffusion model of lineages
-matches reality.
+matches realistic, individual-based models.  \revpoint{AE}{8}
 These solutions easily and quickly provide predictions of joint frequencies at small numbers of loci:
@@
 % to remove these, change 'ifreviewresponses' up top
+\singlespacing
 \includereviews

### 2016-04-22 drafts/review-responses-april-2016.tex
responses ready to be edited

@@
 
+\plr{these aren't points because they are all points below in Reviewer 1 \& 2's sections; we'll just say here that we did 'em, see below}
+
 \begin{quote}
@@
 However, it is clear that any form of selection that gives the same marginal selection coefficient will give essentially the same effect: a large body of work on clines shows that different selective mechanisms are indistinguishable, and it is reasonable for the authors to rely on that.
-Theother main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).
+The other main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).
 I don't know of a reference where they have been written in exactly this form, but the recombination terms and the diffusion terms in Eq.
@@
 \begin{quote}
-Rev 2 is mainly concerned about the fit between theory and simulations.
+Reviewer 2 is mainly concerned about the fit between theory and simulations.
 I am not worried about this, since the deterministic mathematics stands independently, and the numerical computations are straightforward.
@@
 \alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology}
-\plr{I'm fine with moving some; not super clear on the difference anyhow...}
+\plr{good start...}
 
@@
 This article will likely be of most interest to theorists, however, as the discussion of the practical confounders of the selection test is limited.
-A serious limitation of the test is that it only detects selection against heterozygotes.
-The claim that this would not affect results, on l.453-456:
+\end{quote}
+
+\begin{point}{}
+    A serious limitation of the test is that it only detects selection against heterozygotes. The claim that this would not affect results, on l.453-456:
     \textit{``Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models''}
     is unconvincing.
-\end{quote}
+\end{point}
 
+\reply
 \alisa{We agree with the comments of the AE on the matter, and have added some additional text to the introduction that hopefully convinces that this is ok.}
 
-\begin{quote}
+\begin{point}{}
     From a theoretical perspective, the article makes modest but valid contributions, and  Equation 4 and 5 are potentially interesting extensions to equation (1).
 The practical benefit of using these equations rather than forward simulations appears limited.
-\end{quote}
+\end{point}
 
+\reply
 \alisa{This is a point we have considered ourselves.
@@
 
-\begin{quote}
+\begin{point}{}
     While reading the manuscript, I found it difficult to distinguish the most important observations from more tangential observations.
 The manuscript reads like a long list of observations, and would benefit from more focus on the main findings and more structure.
-\end{quote}
+\end{point}
+
+\reply
 
@@
 \reply
+\plr{yaniv doing that}
 
 \begin{point}{Eq.5}
-If $a<b<theta<0$, then I interpret $g(theta,b) = 0$. If correct, this should be specified.  Alternately, the bounds of the integral could be modified to account for this situation.
+If $a<b<\theta<0$, then I interpret $g(\theta,b) = 0$. If correct, this should be specified.  Alternately, the bounds of the integral could be modified to account for this situation.
 \end{point}
@@
 \reply
+\plr{peter did this}
 
@@
 \alisa{(This is at l180) Should the $=$ be a $:=$}
+\plr{peter will fix this}
 
-\begin{point}{l.120} 
-As we show below the solutions provide an good approximation
+\begin{point}{} 
+``As we show below the solutions provide an good approximation''
 \end{point}
@@
 \reply
+Fixed.
 
@@
 \end{point}
+
+\reply
+\plr{peter did this}
 
@@
 \end{point}
+
+\reply
+\plr{peter will do this}
 
@@
 Simulations are set up to test the theoretical predictions and illustrate other interesting patterns such as the distribution of haplotype (ancestry block?) lengths.
+\end{quote}
 
+\begin{point}{}
     Unfortunately, the fit of the theoretical predictions to simulations is largely not shown, and those comparisons in the manuscript are limited to the frequency dynamics of the neutral alleles (Figure S3).
 The predictive theoretical results are based on scaling arguments - but it is not clear which of these results are uniquely derived in this paper.
-\end{quote}
+\end{point}
 
+\reply
 \alisa{Our hope was that the simulations and theory would provide qualitative support for one another.
@@
 
-\begin{quote}
+\plr{peter to add sentence saying this is standard, ref slatkin, barton}
+
+\plr{alisa to see if we can add any more from what we've done}
+
+\begin{point}{}
     The authors also aim to characterize the movement of the genomic block that contains a selected locus through a hybrid zone.
 Yet, the size of the simulated habitat barely covers the cline width, and the simulation results may be sensitive to the boundaries.
-\end{quote}
 
-\alisa{Should we be running more sims? Nick seems to think we have done this, but I don't think we have.}
-
-\begin{quote}
 Overall, this MS presents some interesting insights, and the formalization of the block dynamics can be useful.
@@
 Currently, the theory and the matching simulations are largely presented separately.
-\end{quote}
+\end{point}
+
+\reply
+\alisa{Should we be running more sims? Nick seems to think we have done this, but I don't think we have.}
+\plr{run sims with longer range}
+\plr{also run sims in part of param range where nothing's supposed to happen?}
+\plr{alisa to add $s=.001$ and the like}
 
@@
 \alisa{We have remade Figure 1, with additional information that we hope will help with interpretation}
+\plr{alisa to refer to these events in discussion of terms in figure 5}
 
@@
 \alisa{It is true that the demes are never empty. However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}
+\plr{peter to add note about width of sims in the section describing simulations}
 
@@
 
+\plr{alisa to clarify figure legend}
+\plr{peter or alisa fix S2}
+
 \begin{point}{l. 299}
@@
 \alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}
+\plr{I agree: they are probably thinking of selected loci, not linked ones. We're talking about global drift not local drift. Yaniv to investigate.}
 
@@
 \reply
-\alisa{I don't think this is misphrased?}
+\plr{see above}
 
@@
 \reply
+\alisa{I don't think this is misphrased?}
+\plr{peter will rephrase: it is stochastic}
 
@@
 \reply
+\plr{see above; add something to discussion.}
+
 \alisa{We should add a line or two in the discussion about drift.
@@
 \alisa{We'll work on making this more understandable.}
+\plr{peter to do this}
 
-\begin{point}
+\begin{point}{}
 The ``Data Accessibility section'' is missing.
@@
 \alisa{Ok. This probably needs to be fixed.}
+\plr{alisa to write a good README: what scripts to run to get a simulation; where each figure gets produced}
 
-\begin{point}
+\begin{point}{}
 The Acknowledgements section is missing.

### 2016-04-23 drafts/review-responses-april-2016.tex
peter went through responses once

@@
 2 also worries that the simulations are over a rather narrow range; I suspect that the authors should have results that address sensitivity to the boundaries.
-Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that he authors suggest otherwise.
+Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that the authors suggest otherwise.
 \end{quote}
@@
 
-\begin{point}{Eq.5}
+\begin{point}{Eq.\ref{eqn:g_pde}}
 If $a<b<\theta<0$, then I interpret $g(\theta,b) = 0$. If correct, this should be specified.  Alternately, the bounds of the integral could be modified to account for this situation.
@@
 \reply
-\plr{peter did this}
+Thanks for catching the mistake; we modified the bounds of the integral.
 
@@
 
-\begin{point}{``In the notation above, $q_z(x, t, r) = P^x{Z_t = A}$.''}
+\begin{point}{\revref}
+``In the notation above, $q_z(x, t, r) = P^x{Z_t = A}$.''
 I did not see the notation $P^x$ above, and the RHS does not depend on small z.
@@
 \reply
-\alisa{(This is at l180) Should the $=$ be a $:=$}
-\plr{peter will fix this}
+Good point; we've used a more common notation to specify that $X_0=x$ (and $Z_0=z$ now).
 
@@
 
-\begin{point}{l.144}
+\begin{point}{\revref}
 It seems like the x+r and x-r should have the same sign.
@@
 \reply
-\plr{peter did this}
+That's correct; we've fixed the sign issues.
 
-\begin{point}{l.199}
+\begin{point}{\revref}
 $g_z(x, T, a,b)$ is defined as: \textit{``We are interested in the probability $g_z(x, t; a, b)$ of finding an entire segment $(a, b)$ of ancestry $A$ in an individual having selected allele of type $z$, \ldots ''}
-
 This definition does not make it clear that the probability is conditional on $z$; $\P(A|z)$ vs $\P(A,z)$.
@@
 \reply
-\plr{peter will do this}
+This was indeed confusing; it's rewritten to be hopefully less confusing.
 
@@
 
-\plr{peter to add sentence saying this is standard, ref slatkin, barton}
+The rescaling arguments we provide  are of a standard sort in the literature;
+we've inserted comments to this effect, at \llname{ll:rescaling} \plr{and elsewhere?}.
 
@@
 
-\begin{point}{l. 245}
+\begin{point}{\revref}
  As the theory is set up in the absence of genetic drift, and, correspondingly, the local population size is large, it is unclear how a deme with 500 individuals could become empty in a Wright-Fisher model.
@@
 \reply
-\alisa{It is true that the demes are never empty. However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}
-\plr{peter to add note about width of sims in the section describing simulations}
+Although most of the simulations shown in the main text have high population densities,
+we have in fact run simulations with considerably lower densities, 
+as shown for instance in figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}.
+\alisa{It is true that the demes are never empty.}
+\plr{(It happened in sims that I ran.)}
+\alisa{However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}
+\plr{add reference to more low pop density figures if added}
 
@@
 \alisa{We believe that the displacement is likely due to drift. (We could do a repeated sim to show this?)}
+\plr{I'm ok with switching S2 and 3; that seems like a good idea.
+About the horizontal shift: I think that when I noticed this we figured out that it was because the center of the cline wasn't being plotted quite at zero for some reason?
+If I recall correctly, we thought maybe it was drift but it wasn't.}
+\plr{Alisa to clarify figure legend.}
+\plr{peter or alisa fix S2: this is in \url{Plot_frequency_clines.R}}
 
-\plr{alisa to clarify figure legend}
-\plr{peter or alisa fix S2}
-
-\begin{point}{l. 299}
+\begin{point}{\revref}
 Total population size is irrelevant when predicting the effects of genetic drift (besides, some simulations only have 50 demes across, not 100).
@@
 
-\begin{point}{l.339}
- I think the sentence about the maximum length of the enclosing haplotype, $\sigma \sqrt{s} / x$ is mis-phrased: presumably this should be a statement on the distribution (expected size?) of the introgressing block length at distance x, as a lucky block can travel many sigmas in one generation?
+\begin{point}{\revref}
+ I think the sentence about the maximum length of the enclosing haplotype, $\sigma \sqrt{s} / x$ is mis-phrased: presumably this should be a statement on the distribution (expected size?) of the introgressing block length at distance $x$, as a lucky block can travel many sigmas in one generation?
 \end{point}
@@
 \reply
-\alisa{I don't think this is misphrased?}
-\plr{peter will rephrase: it is stochastic}
+Good point, it was meant to be a statement about an order of magnitude bound to a stochastic quantity rather than a precise bound; it's been rephrased.
 
@@
     It is important to make the reader aware that even neutral clines do not just keep flattening -- genetic drift leads to a finite cline width even for neutral clines.
-Despite the occasional mention of genetic drift, the MS does not clearly state that the parameter space is chosen such that genetic drift is weak and does not indicate when this is a reasonable assumption (see comment on l.
-299).
+Despite the occasional mention of genetic drift, the MS does not clearly state that the parameter space is chosen such that genetic drift is weak and does not indicate when this is a reasonable assumption (see comment above [about \revreffull{2}{6}]).
 \end{point}
@@
 
-\begin{point}{p.27 rates, e.g. Eq. (7,8):}
+\begin{point}{rates at \revref, e.g. Eq. (7,8):}
  I find the explanation of the rates a bit unclear.
@@
 \reply
-\alisa{We'll work on making this more understandable.}
-\plr{peter to do this}
+Right; we've added more discussion of this.
 

### 2016-04-23 drafts/hybrid_zone.tex
peter went through responses once

@@
 and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
+\plr{(What forms of selection?)}
 There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
 and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000} \alisa{and much of this focuses on equilibrium states}.
+
 In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
@@
 and \alisa{ancestry} haplotypes can potentially carry substantial information about the formation of recent zones.
+\plr{(I think we just want to say ``haplotypes'', because ``ancestry haplotypes'' isn't a thing, and although we do only study the sort of information about haplotype patterns we can get by painting by ancestry, the statement holds in general.)}
 To do this, we take the reverse-time perspective,
@@
 This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological (e.g. local adaptation) }
+as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological (e.g. local adaptation).}
+\plr{(ecological what?)}
 
@@
 that are helpful in providing intuition and for establishing order-of-magnitude estimates,
-and analytics, mostly in terms of partial differential equations needing numerical solution. \alisa{For simplicity, we ignore coalescence.}
+and analytics, mostly in terms of partial differential equations needing numerical solution. 
+\alisa{For simplicity, we ignore coalescence.}
+\plr{(say why and what that means: e.g., deterministic dynamics)}
 
@@
 First, we turn our attention to how the cline at the selected locus itself is formed. 
-\alisa{This has been previously described, but is presented here as it provides a framework for further development of our model}
+\alisa{This has been previously described, but is presented here as it provides a framework for further development of our model.}
+\plr{(I presume it's been described, but I don't think we found a reference that actually did this, right?
+So maybe say it's ``well-known in the field'' or something?
+Also, no-one seems to have looked at lineages, and as our final point says, we actually get some more information out of this perspective.)}
 Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 Rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
-results in a dimensionless equation \citep{Slatkin1973}, \linelabel{ll:rescaling}
+results in a dimensionless equation,
 implying that the cline establishes over a timescale of the order $1/s$. 
+(This sort of rescaling is common in the literature, e.g., \citep{fisher1950frequencies,Slatkin1973}.)  \linelabel{ll:rescaling}
 Since this is how long it takes diffusion at rate $\sigma^2$ to smooth across a region of width $\sigma/\sqrt{s}$,
@@
 but biased towards locations where the selected allele they carry is at higher frequency.
-(Lineages are also biased towards regions with higher fecundity,
-and thus away from the hybrid zone, but we assume that $s$ is small and ignore this.) \alisa{could this reference to small $s$ be confusing?}
+% (Lineages are also biased towards regions with higher fecundity,
+% and thus away from the hybrid zone, but we assume that $s$ is small and ignore this.) 
+% \alisa{could this reference to small $s$ be confusing?}
+% \plr{it's confusing because it's differences in fecundity that bias the lineage since they create the net flux of A alleles to the right, even though there's a different term in how a lineage moves for the difference in previous-generation fecundity, that we're ignoring.
+I'm removing the statement.}
 
@@
 % with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
-(i.e., it moves at speed $\sigma^2$ in the potential $-\log p$). \alisa{What does potential here mean?}
-% \alisa{note to self to read through appendix when written}
+(the analogous physical system is a particle that moves at speed $\sigma^2$ in the potential $-\log p$). 
+\alisa{What does potential here mean?}
+\plr{(It means like a gravity well.  Clear now that it says ``particle''?)}
 To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
 lived at position $x-r$ is proportional to $\P\{R=-r\} p(x-r)$, where $R$ is the random dispersal distance;
-and so the mean displacement from offspring to parent is $-\E[R p(x-R)]/\E[p(x-R)]$.
+and so the mean displacement from offspring to parent is $-\E[R p(x-R)]/\E[p(x-R)]$.  \revpoint{1}{9}
 % $$ = \E[ -R ] - \E[R]^2 p'(x)/p(x) + \E[R^2] p'(x)/p(x) + O(R^3) $$
-Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$ \alisa{perhaps a brief explanation of how this works}
+Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$ 
+\alisa{(perhaps a brief explanation of how this works)}
+\plr{(That's what the rest of the sentence is? Work it out, and if you still think it needs more detail, add it?)}
 and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
@@
 \end{align}
-\alisa{Explain in words}
+\alisa{Explain in words.}
+\plr{(I hope that comment came before I added these words? Or do you want different ones?)}
 In the first expression, giving the distribution of how the lineage location changes,
@@
 \begin{align}
-    q_z(x,t,r) = \P^x \{Z_t = A\} .
+    q_z(x,t,r) = \P^x \{Z_t = A \given X_0 = x, \; Z_0 = z\} .
 \end{align}
-(Recall that $Z$ depends implicitly on $r$.)
+(Recall that $Z$ depends implicitly on $r$.) \revpoint{1}{7}
 The description above implies that $q_z$ solves the following Kolmogorov backward equation:
@@
 \end{align}
-\alisa{Use our words}
+\alisa{Use our words} \plr{(what does that mean)}
 with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
-and where $\grad$ is the gradient, and $\Delta$ is the Laplacian \alisa{describing the rate of change}. 
+and where $\grad$ is the gradient, and $\Delta$ is the Laplacian 
+\alisa{describing the rate of change}. 
+\plr{(the Laplacian is describing the rate of change? I don't follow.)}
 The three terms in these equations come from Brownian movement of a linages,
@@
 \paragraph{Numerical computation}
-To determine how the cline at a linked locus is expected to relax, \alisa{and to confirm the behavior of our model to previously established results,}
+To determine how the cline at a linked locus is expected to relax, 
+\alisa{and to confirm the behavior of our model to previously established results,}
+\plr{(got a reference for that?)}
 we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically.
@@
 of a single ancestry are found, at a given location and time.
-Consider a segment of the genome between positions $a$ and $b$, relative to the selected site. 
+Suppose we sample an individual at spatial location $x$ and time $t$ after the initiation of gene flow,
+and genotype them on the genomic segment between positions $a$ and $b$, relative to the selected site. 
 We are interested in the probability $g_z(x,t;a,b)$ 
-of finding an entire segment $(a,b)$ of ancestry $A$ in an individual having selected allele of type $z$,
-sampled at spatial location $x$ at time $t$ after the initiation of gene flow.
+of finding an entire segment $(a,b)$ of ancestry $A$ 
+given that the individual has selected allele of type $z$.
+\revpoint{1}{10}
+For instance, $g_B(0,10;0,.2)$ is the probability that an individual carrying a selected $B$ allele sampled at the center of the zone
+at $t=10$ has a block of $A$ ancestry for 0.2 Morgans to the right of the selected site.
 
@@
 \end{align}
-
 \alisa{Need to be related to Figure 1}
@@
 the portion of the block nearest the selected site remains linked to an $A$ allele, 
-and the remaining portion becomes linked to a $B$ allele
-(\alisa{to consider cases where the selected locus lies outside the interval [a,b], i.e.,} we say that the integral $\int_a^{\min(0,b)}$ is zero if  $a>0$ 
-and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$) .
+and the remaining portion becomes linked to a $B$ allele;
+\alisa{to account for cases where the selected locus lies outside the interval [a,b],} 
+we say that the integral $\int_a^{\min(0,b)}$ is zero if  $a>0$ 
+and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$.
 The last integral results from recombination inside the block in homozygotes for $A$,
@@
 in this case, parents are chosen from other demes with probability proportional to fitness
-multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
+multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme. \revpoint{2}{4}
 The next generation is formed by carrying out meiosis in each parent
@@
 
-\subsection*{Single-loci: clines}
+\subsection*{Single locus clines}
 
@@
 \citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
-
+\revpoint{2}{6}
+
+\plr{(This next bit is one of these scaling arguments that Reviewer 2 wants clarified.
+Again, I don't know who's described how the cline relaxes, do you?
+I think at least the "speed of a lineage" bit is new, but don't want to make a big deal out of it.
+Not sure how to signpost this.}
 Loci not under direct selection can in principle spread across the cline unimpeded. 
@@
 (compared to $x^2/\sigma^2$ for a neutral allele).
-This implies the enclosing $A$ haplotype surrounding the locus should be no longer than $\sigma\sqrt{s}/x$.
+This implies the scale on which $A$ haplotypes are found surrounding the locus should be no longer than about $\sigma\sqrt{s}/x$.
+\revpoint{2}{8}
 %\plr{(check predicted decay of mean length $1/x$ in simulations?)}
@@
 \end{align}
+For instance,  \revpoint{2}{10}
+there are $p(x) N(x)$ type $A$ individuals at $x$, that reproduce at rate $s_A(x)$;
+the chance that each reproduction is with a type $B$ is $(1-p(x))$, 
+and the chance there is a recombination that gives the offspring genotype $AB$ is $r/2$;
+the offspring has probability $m(x,y)$ to disperse to $y$;
+the total rate at which individuals at $x$ with the $A$ allele produce offspring at $y$ with the $B$ allele is the product of these terms,
+$p(x) N(x) s_A(x) (1-p(x)) r / 2$.
+The second and third rates (for $x\xrightarrow{AB}y$ and $x\xrightarrow{BA}y$) are the same;
+they are this value plus the corresponding term
+for individuals carrying the $B$ allele producing offspring that carry the $A$ allele.
 
 Note that at equilibrium, we require $N$ to solve
-\begin{align}
+\begin{align*}
   0 = \sum_y N(y) m(y,x) \left\{ p(y)s_A(y)(1-p(x)) - (1-p(y))s_B(y)p(x) \right\} .
-\end{align}
+\end{align*}
 
@@
 or
-\begin{align}
+\begin{align*}
   r_A(x,y) &= \frac{w_{AA}(y,x)}{p(x)N(x)} \\
   &= N(y) p(y) s_A(y) m(y,x) \frac{ 1 }{ N(x) p(x) } .
-\end{align}
+\end{align*}
 
@@
 the associated random walk converges to Brownian motion, so that for an arbitrary smooth function $f$,
-\begin{align}
+\begin{align*}
     \lim_{\sigma \to 0} \sum_y \frac{ m(x,y) ( f(y) - f(x) ) }{\sigma^2} = \frac{1}{2} \dderiv{x} f(x) .
-\end{align}
+\end{align*}
 Write $f'(x) = \deriv{x}f(x)$, and note that
-\begin{align}
+\begin{align*}
     \frac{1}{\sigma^2} \sum_y g(y) m(y,x) (f(y)-f(x)) 
@@
     &\xrightarrow{\sigma \to 0} \frac{1}{2} \dderiv{x}\left( g(x)f(x) \right) - \frac{1}{2} f(x) \dderiv{x} g(x) \\
-    &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) \\
-    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) . \label{eqn:deriv_limit}
+    &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) ,
+\end{align*}
+which tells us the differential operator that best approximates the discrete sum:
+\begin{align}
+    \frac{1}{\sigma^2} \sum_y g(y) m(y,x) (f(y)-f(x)) 
+    &\xrightarrow{\sigma \to 0} \frac{1}{2} g(x) f''(x) + g'(x) f'(x) . \label{eqn:deriv_limit}
 \end{align}
@@
 Under these assumptions, combining \eqref{eqn:discrete_generator} and \eqref{eqn:deriv_limit},
-\begin{align}
+\begin{align*}
     \deriv{t} \E[f(X_{t/\sigma^2}) \given X_0=x ] &\xrightarrow{\sigma \to 0} 
@@
     &= \frac{1}{2} s_A(x) f''(x) + \left( s_A'(x) + \log(N(x)p(x))' s_A(x) \right) f'(x) ,
-\end{align}
+\end{align*}
 i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with mean displacement

### 2016-04-23 drafts/hybrid_zone.tex
.pdf problems

--- 2016-04-22 drafts/hybrid_zone.tex
@@
 the two alleles at the selected locus will form a stable cline,
-affecting neighboring loci as well \citep{barton1979gene, Barton1986}.
+affecting neighboring loci as well \citep{barton1979gene, barton1986barrier}.
 First, we turn our attention to how the cline at the selected locus itself is formed. 

### 2016-04-24 drafts/review-responses-april-2016.tex

@@
 Well, we hope to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted the wording.
-\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
-\plr{I changed to ``realistic, individual-based models''.}
+%\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
+%\plr{I changed to ``realistic, individual-based models''.}
 
@@
 \reply
-\alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology}
+\alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology.
+One of our concerns is that the math would be largely overlooked if moved into the supplement.
+Having consulted a number of non-theoretician biologists, we decided that it would be helpful to have even the fairly standard derivations easily accessible.}
 \plr{good start...}
@@
 \alisa{We have remade Figure 1, with additional information that we hope will help with interpretation}
-\plr{alisa to refer to these events in discussion of terms in figure 5}
+\plr{alisa to refer to these events in discussion of terms in figure 5} \alisa{Did we mean Eq. 5?}
 
@@
 as shown for instance in figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}.
-\alisa{It is true that the demes are never empty.}
+\alisa{It is true that the demes are never empty for very large populations. We however, did encounter empty demes in some simulations will smaller populations sizes.}
 \plr{(It happened in sims that I ran.)}

### 2016-04-24 drafts/hybrid_zone.tex

--- 2016-04-23 drafts/hybrid_zone.tex
@@
 Although the selected cline establishes quickly, transient patterns can persist for quite some time,
-and \alisa{ancestry} haplotypes can potentially carry substantial information about the formation of recent zones.
-\plr{(I think we just want to say ``haplotypes'', because ``ancestry haplotypes'' isn't a thing, and although we do only study the sort of information about haplotype patterns we can get by painting by ancestry, the statement holds in general.)}
+and  haplotypes can potentially carry substantial information about the formation of recent zones.
 To do this, we take the reverse-time perspective,
@@
 This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological (e.g. local adaptation).}
+as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological selection (e.g. local adaptation).}
 \plr{(ecological what?)}
@@
 \begin{figure}
-\includegraphics[clip,width=\textwidth,trim=0cm 0cm 0cm 0.05cm]{figs/BM_schematic_new2.pdf}
+\includegraphics[clip,width=\textwidth,trim=0cm 8cm 0cm 0.05cm]{figs/BM_schematic_new2.pdf}
 \caption{
@@
 and analytics, mostly in terms of partial differential equations needing numerical solution. 
-\alisa{For simplicity, we ignore coalescence.}
+\alisa{For simplicity, we ignore coalescence, which equates to ignoring the effects of drift and pedigree.}
 \plr{(say why and what that means: e.g., deterministic dynamics)}
@@
 So maybe say it's ``well-known in the field'' or something?
-Also, no-one seems to have looked at lineages, and as our final point says, we actually get some more information out of this perspective.)}
+Also, no-one seems to have looked at lineages, and as our final point says, we actually get some more information out of this perspective.)} \alisa{So, when I sati's been previously described, I'm thinking about the traditional cline theory for a single locus (e.g.
+Slatkin, May), which describes the time-forward diffusion.}
 Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 \end{align}
-where $\Delta p$ is the Laplacian.
-\plr{change $\Delta$ to the second derivative?}
+where $\Delta p$ is the Laplacian\alisa{, or second derivative.}
 The equation is approximate because it 
@@
 Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$ 
-\alisa{(perhaps a brief explanation of how this works)}
+\alisa{(perhaps a brief explanation of how this works; in particular I wasn't sure why were were taking the ratio of two expectations...)}
 \plr{(That's what the rest of the sentence is? Work it out, and if you still think it needs more detail, add it?)}
@@
 \end{align}
-\alisa{Explain in words.}
-\plr{(I hope that comment came before I added these words? Or do you want different ones?)}
 In the first expression, giving the distribution of how the lineage location changes,
-the first term ($\sigma dB_\tau$) is Brownian noise driven by dispersal,
-and the second is the mean displacement, which moves the lineage ``downhill'' on either $-\log p$ or $-\log (1-p)$
+the first term ($\sigma dB_\tau$) is Brownian noise driven by dispersal, \alisa{This term is missing the $\sigma$ above...}
+and the second is the mean displacement, which moves the lineage ``downhill'' \alisa{towards their ancestral range} on either $-\log p$ or $-\log (1-p)$
 ($\grad$ is the gradient),
 depending on which selected allele the lineage is linked to.
-The second and third expressions say that a lineage switches from an $A$ to $B$ background
-at rate $r$ multiplied by the local proportion of $B$ alleles, which is $(1-p)$.
+The second and third expressions \alisa{give the expected time for a lineage to switch from} an $A$ to $B$ (or $B$ to $A$) background as a result of recombination.
+This occurs at rate $r$ multiplied by the local proportion of $B$ alleles, which is $(1-p)$.
 
@@
 \end{align}
-\alisa{Use our words} \plr{(what does that mean)}
 with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
-and where $\grad$ is the gradient, and $\Delta$ is the Laplacian 
-\alisa{describing the rate of change}. 
-\plr{(the Laplacian is describing the rate of change? I don't follow.)}
+and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
 The three terms in these equations come from Brownian movement of a linages,
 the tendency of a lineage to inherit from regions where its type is more common,
-and recombination between selected backgrounds, respectively.
+and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
 Equivalently,
@@
 and the second term results from recombinations \emph{between} the block and the selected site 
-in heterozygotes, which switch the identity of the linked, selected locus
+in heterozygotes, which switch the identity of the linked, selected locus \alisa{without splitting the block}
 (the factor $(1-p)$ is the probability that the block, initially linked to an $A$ allele, encounters a $B$ allele
@@
 the portion of the block nearest the selected site remains linked to an $A$ allele, 
-and the remaining portion becomes linked to a $B$ allele;
+and the remaining portion becomes linked to a $B$ allele (as is the case in the most recent recombination event shown in Fig.~\ref{Fig:schematic}).
 \alisa{to account for cases where the selected locus lies outside the interval [a,b],} 
@@
 and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$.
-The last integral results from recombination inside the block in homozygotes for $A$,
+The last integral results from recombination inside the block in homozygotes for $A$ (as is the case for the second recombination event in Fig.~\ref{Fig:schematic}),
 and the final term balances the outflux due to all recombinations.

### 2016-04-24 drafts/hybrid_zone.tex
made some revisions
yb edited intro

@@
 patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.  
-
-While for much of the genome, gene flow will eventually erase 
-both geographic clines in ancestry and strong correlations in ancestry across linked markers, 
-subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{barton1979gene}. 
+\yb{We show that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is, 
+because selection rapidly removes incompatible alleles from heterospecific genomes, hybrid incompatibilities will be surrounded by disproportionately long ancestry blocks.}
+
+% YB REMOVED
+%While for much of the genome, gene flow will eventually erase 
+%both geographic clines in ancestry and strong correlations in ancestry across linked markers, 
+%subsets of the genome can maintain stable clines in ancestry if selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{barton1979gene}. 
       % previously had wrong Barton '79 here
-Studies of hybrid zones have primarily been based on single-site statistics (e.g. clines),
-but the recent availability of dense genotype data in combination with linkage information 
-makes it now possible to use patterns of the block length distribution
-to better understand the processes operating in hybrid zones.
-
-There is substantial theoretical work on genetic clines.
-We focus on those maintained only by selection against hybrids -- \emph{tension zones} --
+%Studies of hybrid zones have primarily been based on single-site statistics (e.g. clines),
+%but the recent availability of dense genotype data in combination with linkage information 
+%makes it now possible to use patterns of the block length distribution
+%to better understand the processes operating in hybrid zones.
+
+\ybr{Our work builds on the large body of theory on genetic clines. Specifically,  
+we focus on genetic clines maintained} by selection against hybrids -- \emph{tension zones} --
 as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}. 
-Clines may also be maintained by environmental gradients \citep{Haldane1948,fisher1950frequencies};
-although these are in principle different,
-\citet{Slatkin1973} pointed out, in a far-reaching general analysis,
-that while the environment may determine the location of a cline,
-the fitness of the heterozygote will likely determine the shape of the cline,
-and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
-\plr{(What forms of selection?)}
+\yb{Clines may also be maintained by differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies}, generating \emph{ecotones}. 
+Although these are, in principle different, theoretical analyses consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, with the primary difference being that the center of a tension zone wobbles over time \citep{Barton1979a} because stochastic changes in allele frequency can shit the cline center, while centers of ecotones do not change unless the environmental gradient changes.
+}
+%\citet{Slatkin1973} pointed out, in a far-reaching general analysis,
+%that while the environment may determine the location of a cline,
+%the fitness of the heterozygote will likely determine the shape of the cline,
+%and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
+%\plr{(What forms of selection?)}
 There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
@@
 in particular describing haplotype patterns.
-Although the selected cline establishes quickly, transient patterns can persist for quite some time,
+Although the selected allele frequency cline establishes quickly, transient patterns can persist for quite some time,
 and  haplotypes can potentially carry substantial information about the formation of recent zones.
@@
 understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
-This is particularly suited to understanding transient patterns 
-although, because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
-incorporating genetic drift is not straightforward.
-Common to nearly all studies is the use of the
+This is particularly suited to understanding transient patterns.  
+\yb{However, because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
+incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.   
+This approach} is common to nearly all studies is the use of the
 reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
@@
 \plr{Need to cite Durrett and Cox's stuff here, especially \citet{durrett2007width}.}
+\yb{While we cannot analytically derive expectations for finite populations, our simulation results (below) suggest that genetic drift does not substantially influence our conclusions, perhaps because of the short time-scales that we examine.}
 
@@
 The arrangement of segments of genome inherited from particular ancestral populations
-provides information about the number of recombination events that have taken place between ancestral genotypes, 
-and therefore contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. % too much `provides information'
-In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, 
-this could be especially important for understanding the impact of selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
-In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
-In addition to facilitating demographic inference, heterogeneous patterns of ancestry across the genome can help identify putative targets of selection in hybridizing populations  \citep{Porter1997, Gompert2012}.
+is shaped by the number of recombination events that have taken place between ancestral genotypes. 
+Therefore the length distribution of such segments contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
+% too much `provides information'
+The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Gravel2012,sedghifar2015spatial}.
+In addition to allowing researchers to interpret non-equilibrium patterns of diversity in hybrid zones, 
+the length distribution of ancestry tracts could reveal the impact of ongoing selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
+%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
+%If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent %of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
+\yb{We show that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.
+][]{Porter1997, Gompert2012}.
 
@@
 This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes. \alisa{This reasoning is applicable to both instances of under dominance and ecological selection (e.g. local adaptation).}
-\plr{(ecological what?)}
+as many ancestors of the locally common type as do neutral haplotypes.
+\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotones generated by extrinsic ecological selection (e.g.
+local adaptation).}
+%\plr{(ecological what?)}
 
@@
 \begin{figure}
-\includegraphics{figs/alleleFrequencies_sim.pdf}
+\includegraphics{figs/alleleFrequencies_sim_comparison_revision_original.pdf}
 \caption{

### 2016-04-24 drafts/hybrid_zone.tex
some changes

--- 2016-04-24 drafts/hybrid_zone.tex
@@
 \begin{figure}
-\includegraphics{figs/alleleFrequencies_sim.pdf}
+\includegraphics{figs/alleleFrequencies_sim_comparison_revision_original.pdf}
     \caption{
-    Frequency of ancestry $B$,
-    across geography at different physical positions on the genome, simulated for a hybrid zone
-     $T=100$ generations after secondary contact, with $s=0.1$,
-     using 50 demes, each with 500 diploid individuals and $\sigma=1$.
-    Each line represents a locus some distance $r$ away from the true target of selection.
-    Grey lines represent the same positions from a simulation with
-    identical parameters except that $s=0$.
-    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
-    the same plot is shown with weaker selection in Figure \ref{alleleFreq_tau100_weaker_s} and at a longer time in Figure \ref{alleleFreq_tau1000}.
-}\label{alleleFreq_tau100}
+        \textbf{Comparison of simulated to theoretical clines:}
+        As in Figure \ref{alleleFreq_tau100},
+        frequency of ancestry $B$,
+        across geography and at several physical positions on the genome,
+        at $T=100$ generations after secondary contact,
+        and at $s=0.1$.
+        Dotted lines are expected frequencies of $B$ ancestry $1-q(x,t,r)$
+        (unconditional on the ancestry of the linked allele), computed numerically.
+        Each line represents a locus some distance $r$ away from the true target of selection, 
+        and $r=0$ represents the locus that is under selection.    
+    }\label{sfig:alleleFreq_tau100_comparison}
 \end{figure}
@@
 The hybrid zone is 50 demes, each containing 500 diploid individuals.
- Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$).
+ Each red or orange line represents a site some distance away (ranging from 0 - 0.5cM) from the selected site (here $s=0.01$ and $\sigma=1$).
 Yellow lines are corresponding positions on an unlinked chromosome with no selected loci, and grey lines are corresponding positions in a simulation with no selection.

### 2016-04-24 drafts/supplement.tex
made some revisions
some changes

@@
 \begin{figure}
-    \includegraphics{figs/alleleFrequencies_sim_comparison.pdf}
+    \includegraphics{figs/alleleFrequencies_sim.pdf}
     \caption{
-        \textbf{Comparison of simulated to theoretical clines:}
-        As in Figure \ref{alleleFreq_tau100},
-        frequency of ancestry $B$,
-        across geography and at several physical positions on the genome,
-        at $T=100$ generations after secondary contact,
-        and at $s=0.1$.
-        Dotted lines are expected frequencies of $B$ ancestry $1-q(x,t,r)$
-        (unconditional on the ancestry of the linked allele), computed numerically.
-        Each line represents a locus some distance $r$ away from the true target of selection, 
-        and $r=0$ represents the locus that is under selection.    
-    }\label{sfig:alleleFreq_tau100_comparison}
+    Frequency of ancestry $B$,
+    across geography at different physical positions on the genome, simulated for a hybrid zone
+     $T=100$ generations after secondary contact, with $s=0.1$,
+     using 50 demes, each with 500 diploid individuals and $\sigma=1$.
+    Each line represents a locus some distance $r$ away from the true target of selection.
+    Grey lines represent the same positions from a simulation with
+    identical parameters except that $s=0$.
+    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
+    the same plot is shown with weaker selection in Figure \ref{alleleFreq_tau100_weaker_s} and at a longer time in Figure \ref{alleleFreq_tau1000}.
+}\label{alleleFreq_tau100}
 \end{figure}
@@
 
+\begin{figure}
+    \includegraphics{figs/s001_adjacentBlocksAlongChromAncBConditioning}
+    \includegraphics{figs/s001_adjacentBlocksAlongChromNoConditioningHighRes}
+        \caption{
+     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (upper) and without conditioning (lower) alisa{is this perhaps an instance where we would need to have more demes to see a signal?
+    }
+    } \label{sfig:blockLengthPlot_t1000_s0.001}
+\end{figure}
+
 % \end{document}

### 2016-04-24 drafts/figs/Plot_frequency_clines.R

@@ -27,7 +27,7 @@ sel.params <- get_simparams( sel.sim.file )
 neutral.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"
 neu.params <- get_simparams( neutral.sim.file )
 
-if (!file.exists("plot_frequency_clines_revised.Robj")) {
+if (!file.exists("plot_frequency_clines_revision.Robj")) {
     load(sel.sim.file)
     #CHUNKS_SEL = get.chunks.at.positions(positions,CHR=1)
 
@@ -53,9 +53,9 @@ if (!file.exists("plot_frequency_clines_revised.Robj")) {
             CHR=1, 
             ndemes=neu.params$ndemes)
 
-    save(B_freq, B_neu, file="plot_frequency_clines_revised.Robj")
+    save(B_freq, B_neu, file="plot_frequency_clines_revision.Robj")
 } else {
-    load("plot_frequency_clines_revised.Robj")
+    load("plot_frequency_clines_revision.Robj")
 }
 
 

### 2016-04-24 drafts/figs/number_of_inds.R

@@ -1,3 +1,6 @@
+start_coordinate=0.49
+end_coordinate=0.51
+
 load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_run2016-01-27_rawoutput.Robj")	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.ids","deme","pars","QTL")]]
 		#where:
 			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
@@ -35,8 +38,8 @@ ancestry.at.position = function(IND=1,DATA=sims.s0.1,POS = 0.5,CHROM="chr1",SP="
 
 ninds=25000
 
-selected_chromsome_ancestor_bydeme = lapply(seq(0.495,0.505,0.00025),function(P){tapply(1:ninds,sims.s0.1[[1]][["deme"]],function(X){lapply(X,ancestry.at.position,DATA=sims.s0.1,POS=P,CHROM="chr1",SP="B")})})
-unlinked_chromsome_ancestor_bydeme = lapply(seq(0.495,0.505,0.00025),function(P){tapply(1:ninds,sims.s0.1[[1]][["deme"]],function(X){lapply(X,ancestry.at.position,DATA=sims.s0.1,POS=P,CHROM="chr2",SP="B")})})
+selected_chromsome_ancestor_bydeme = lapply(seq(start_coordinate,end_coordinate,0.00025),function(P){tapply(1:ninds,sims.s0.1[[1]][["deme"]],function(X){lapply(X,ancestry.at.position,DATA=sims.s0.1,POS=P,CHROM="chr1",SP="B")})})
+unlinked_chromsome_ancestor_bydeme = lapply(seq(start_coordinate,end_coordinate,0.00025),function(P){tapply(1:ninds,sims.s0.1[[1]][["deme"]],function(X){lapply(X,ancestry.at.position,DATA=sims.s0.1,POS=P,CHROM="chr2",SP="B")})})
 
 selected_locus_AncB_ancestors = lapply(selected_chromsome_ancestor_bydeme,function(POS){
 	lapply(POS,function(X){table_of_inds=do.call(rbind,lapply(X,function(Z){
@@ -48,7 +51,7 @@ unlinked_locus_AncB_ancestors = lapply(unlinked_chromsome_ancestor_bydeme,functi
 
 
[…]

### 2016-04-24 drafts/hybrid_zone.tex

--- 2016-04-24 drafts/hybrid_zone.tex
@@
     \caption{
-    Frequency of ancestry $B$,
-    across geography at different physical positions on the genome, simulated for a hybrid zone
-     $T=100$ generations after secondary contact, with $s=0.1$,
-     using 50 demes, each with 500 diploid individuals and $\sigma=1$.
-    Each line represents a locus some distance $r$ away from the true target of selection.
-    Grey lines represent the same positions from a simulation with
-    identical parameters except that $s=0$.
-    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
-    the same plot is shown with weaker selection in Figure \ref{alleleFreq_tau100_weaker_s} and at a longer time in Figure \ref{alleleFreq_tau1000}.
-}\label{alleleFreq_tau100}
+        \textbf{Comparison of simulated to theoretical clines:}
+        As in Figure \ref{alleleFreq_tau100},
+        frequency of ancestry $B$,
+        across geography and at several physical positions on the genome,
+        at $T=100$ generations after secondary contact,
+        and at $s=0.1$.
+        Dotted lines are expected frequencies of $B$ ancestry $1-q(x,t,r)$
+        (unconditional on the ancestry of the linked allele), computed numerically.
+        Each line represents a locus some distance $r$ away from the true target of selection, 
+        and $r=0$ represents the locus that is under selection.    
+    }\label{sfig:alleleFreq_tau100_comparison}
 \end{figure}

### 2016-04-24 drafts/hybrid_zone.tex
added graham comments

@@
 Clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
-and linked clines persist over a  physical scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. 
+and linked clines persist over a  genomic scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. 
 Locally disadvantageous alleles tend to exist as small families,
@@
 
-\ybr{Our work builds on the large body of theory on genetic clines. Specifically,  
+\yb{Our work builds on the large body of theory on genetic clines. Specifically,  
 we focus on genetic clines maintained} by selection against hybrids -- \emph{tension zones} --
@@
 ][]{Porter1997, Gompert2012}.
-
-If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression of surrounding genomic regions.
+}
+
+If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected allele will be removed rapidly from the population, preventing introgression of surrounding genomic regions.
 We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
@@
 Making the same assumptions as for equation \eqref{eqn:cline_pde}, the derivation in Appendix \ref{apx:lineage_derivation} shows 
-that the lineage of an $A$ allele moves as Brownian motion driven by dispersal
+that the lineage of an $A$ allele at position $x$ moves as Brownian motion driven by dispersal
 with mean displacement proportional to the derivative of $\log p$.
@@
 reflecting strong differences in selection against heterozygotes,
-even though the stable clines have the same form.
+even though the stable clines have the same form, and therefore have additional information about the hybrid zone.
 Even though a rescaling of space and time as described above can make the models equivalent,
@@
 the block of unbroken ancestry containing $m$. 
-As described below, these blocks tend to be \emph{shorter} than average,
+As described below, these blocks tend to be \emph{shorter} than average when $m$ is the selected site,
 motivating us to define the statistic
-\begin{align*}
+\begin{align}
      C(m,x) =  \frac{2\sum_i{l_B(m_i,x)}}{\sum_i{l_A(m_i+,x)+l_A(m_i-,x)}}
-\end{align*}
+\end{align}
 where $m_i$ is the block length in individual $i$, and the sum is over individuals at location $x$.
@@
 \caption{\textbf{Normalized mean enclosing block length}, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
-Here each line represents  $l_B(m)$, in a given deme, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
+Here each line represents  $l_B(m)$, in a given deme some distance from the zone center, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
 Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals.
@@
 \begin{figure}
-\includegraphics{figs/number_of_ancestors_tau1000.pdf}
+\includegraphics{figs/family_size_tau1000_revision.pdf}
 \caption{\textbf{Mean family size of haplotypes.} The number of individuals of ancestry $B$ per number of ancestors ($F_B(m,x)$) from secondary contact occurring $T=1000$ generations ago, represented across geographic space.

### 2016-04-24 drafts/hybrid_zone.tex
yb biologized maths

@@
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}.
-
-As in \citet{sedghifar2015spatial}, we consider where the ancestors of modern-day individuals lived,
-i.e., where their lineages fall across geography as we look further back towards the time of first contact.
-We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$ at the time of secondary contact,
-i.e., if its lineage traces back to a spatial position $x<0$.
+previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}, \yb{because, selection on an allele depends on its marginal effect on fitness.
+Exceptionally strong epistatic selection which acts more quickly than free recombination may therefore, yield results that differ slightly from ours. 
+Likewise, while we do not model ecological selection, previous theory shows that clines formed by under dominant selection and extransic selection are essential indistinguishable, especially over short time scales when there is little time for clines to wobble  \citep{Barton1979a}}.
+
+As in \citet{sedghifar2015spatial}, we consider where the ancestors of modern-day 
+fall across geography as we look further back towards the time of first contact.  
+We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$,
+i.e., if its lineage traces back to a spatial position $x<0$  at the time of secondary contact.
 A block of genome is of ancestry $A$ if every locus in it does the same; 
@@
 and analytics, mostly in terms of partial differential equations needing numerical solution. 
-\alisa{For simplicity, we ignore coalescence, which equates to ignoring the effects of drift and pedigree.}
-\plr{(say why and what that means: e.g., deterministic dynamics)}
+\yb{Because of challenges in modeling spatial coalescence, we ignore the effects of drift and the pedigree. 
+Therefore, while our model incorporates stochastic migration, the genetics displays deterministic dynamics.}
+%\plr{(say why and what that means: e.g., deterministic dynamics)}
 
@@
 \begin{align} \label{eqn:cline_pde}
-    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
+    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p +  2 p (1-p)  \times s (\frac{p}{2} - 1), %YB Rearranged this equation to allow for cleaner description
+%    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
 \end{align}
-where $\Delta p$ is the Laplacian\alisa{, or second derivative.}
+\yb{where $ \frac{d}{dt} p$ is the local rate of change of the frequency of $A$ ($\frac{d}{dt} p$), and  $\Delta p$ is the Laplacian\alisa{, i.e.
+the second derivative of allele frequency over space ($x$), at time $t$.}
+The first term,  $\sigma^2/2 \Delta p$, describes the net rate at which $A$ alleles arrive by local dispersal ($\sigma^2/2 \Delta p$). 
+The second term,  $2 p (1-p)  s (p/2-1)$, describes the impact of selection on local allele frequency as the product in local genotypic variance, $2 p (1-p)$, and selection on allele $A$, $s (p/2-1)$.
+The term $s (p/2-1)$ clearly describes  underdominance --- $A$ is favored when common ($p>.5$) and disfavored when rare ($p<.5$).}
+%YB: Pardon the extreme changes here, I tried to make things more obvious to a biologist
+%and a logistic term proportional to the difference in reproductive rate between $A$ and $a$ alleles
+%due to local proportions of heterozygotes,
+%which is $s(2p-1)$ (up to an error of size $s^2$).
+%This can be derived as in \citet{Nagylaki1975};
+%for discussion of diploidy see \citet{christiansen1995genotypic}.
+
 The equation is approximate because it 
@@
 (so if dispersal is non-Gaussian, will fit best over longer time scales, say, tens of generations).  \revpoint{AE}{2} 
-This says that the local rate of change of the frequency of $A$ ($\frac{d}{dt} p$)
-is equal to the net rate at which $A$ alleles arrive by local dispersal ($\sigma^2/2 \Delta p$)
-and a logistic term proportional to the difference in reproductive rate between $A$ and $a$ alleles
-due to local proportions of heterozygotes,
-which is $s(2p-1)$ (up to an error of size $s^2$).
-This can be derived as in \citet{Nagylaki1975};
-for discussion of diploidy see \citet{christiansen1995genotypic}.
-As we show below the solutions provide a good approximation across the range of realistic values of $s$.
+As we show below, the solutions provide a good approximation across the range of realistic values of $s$.
 Although there is no known exact analytic solution to this equation, 
@@
 with mean displacement proportional to the derivative of $\log p$.
-This has the effect of pulling $A$ lineages towards regions where they are more common
+This has the effect of pulling $A$ lineages towards regions where they are more common i.e. the left side of the range.   
 % with speed $\sigma$
@@
 (the analogous physical system is a particle that moves at speed $\sigma^2$ in the potential $-\log p$). 
-\alisa{What does potential here mean?}
-\plr{(It means like a gravity well.  Clear now that it says ``particle''?)}
+%\alisa{What does potential here mean?}
+%\plr{(It means like a gravity well.  Clear now that it says ``particle''?)}
 To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
@@
 Using the fact that $\E[R]=0$ and $\E[R^2]=\sigma^2$ 
-\alisa{(perhaps a brief explanation of how this works; in particular I wasn't sure why were were taking the ratio of two expectations...)}
-\plr{(That's what the rest of the sentence is? Work it out, and if you still think it needs more detail, add it?)}
+%\alisa{(perhaps a brief explanation of how this works; in particular I wasn't sure why were were taking the ratio of two expectations...)}
+%\plr{(That's what the rest of the sentence is? Work it out, and if you still think it needs more detail, add it?)}
 and expanding $p()$ to first order about $x$ shows that this mean displacement is approximately $\sigma^2 p'(x)/p(x)$,
@@
 
-Suppose we sample an allele today $r$ Morgans from the selected site. 
-We will follow its lineage back through time, using $\tau$ to denote ``generations ago''
-and reserving $t$ for time measured in the usual direction.
+\yb{Suppose we sample an allele today $r$ Morgans from the selected site, and  
+follow its lineage back through time, using $\tau$ to denote ``generations ago'' (r
+reserving $t$ for time measured in the usual direction).
 If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
-and $Z_\tau$ is the identity of the selected allele that ancestor carried,
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ 
-(as described above for a selected allele),
-and that $Z$ jumps between $A$ and $B$ at rate either $r (1-p)$  (from $A$ to $B$) or $r p$ (from $B$ to $A$),
-by the assumption of locally random assortment of alleles
-(see Appendix \ref{apx:lineage_derivation} for a more precise description).
+(as described for a selected allele in equation \eqref{eqn:cline_pde}).
+Following this lineage back in time, the identity of the selected allele that ancestor carried at time $\tau$, $Z_\tau$, 
+ jumps between $A$ and $B$ with recombination events between the neutral and selected locus (with frequency $r$), that result in a change in the  selected background.    
+Thus, by the assumption of locally random assortment of alleles, $Z$ shifts from $A$ to $B$  at rate $r (1-p)$, and $B$ to $A$ at rate $r p$ (see Appendix \ref{apx:lineage_derivation} for a more precise description).}
 
@@
 and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
-The three terms in these equations come from Brownian movement of a linages,
-the tendency of a lineage to inherit from regions where its type is more common,
+The three terms in these equations come from Brownian movement of a linages (i.e. the smoothing action of local dispersal),
+the tendency of a lineage to inherit from regions where its type is more common (i.e. the net flux induced by reduced fitness in the hybrid zone),
 and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
-Equivalently,
-the first term ($\Delta q$) reflects the smoothing action of local dispersal,
-the second is the net flux induced by reduced fitness in the hybrid zone,
-and the third is the contribution of recombination. 
 %\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
@@
 To determine how the cline at a linked locus is expected to relax, 
-\alisa{and to confirm the behavior of our model to previously established results,}
-\plr{(got a reference for that?)}
+%\alisa{and to confirm the behavior of our model to previously established results,}
+%\plr{(got a reference for that?)} 
 we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically.
@@
 at a site at recombination distance $r$ from the selected site,
-which is computed as the frequency of ancestry $A$ on each selected background weighted by the freequencies of each background:
+which is computed as the frequency of ancestry $A$ on each selected background weighted by the frequencies of each background:
 $q(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. 

### 2016-04-25 drafts/review-responses-april-2016.tex
made some revisions
yb did his best to deal with reviews

@@
 \begin{minipage}[b]{2.5in}
-    Alisa's address\\
+    Lewis-Sigler institute for integrative genomics\\
+    Princeton University\\
     Yaniv's address\\
@@
 \reply
-Well, we hope to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted the wording.
+Well, we hope to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted the wording to say ``\emph{realistic, individual-based models.}''.
 %\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
@@
 \alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology. 
-One of our concerns is that the math would be largely overlooked if moved into the supplement.
-Having consulted a number of non-theoretician biologists, we decided that it would be helpful to have even the fairly standard derivations easily accessible.}
-\plr{good start...}
+Thus, while some results are fairly standard, they might not be familiar to our readership.  
+One of our concerns is that the math would be largely overlooked if moved into the supplement (which is not automatically printed with the main document) and this could create a barrier to understanding among the readers.}
 
@@
 We present these equations as a way of building our model up to address haplotype lengths.
-Both analytical expressions and simulations are presented so that they each support the conclusions of the other.}
+Both analytical expressions and simulations are presented so that they each support the conclusions of the other.} \yb{However, we also believe that analytical results can allow for a deeper understanding than simulation output.
+To aid our readers in developing this understanding, we have gone through considerable pains to map terms in our equations to the biological processes of migration, selection and recombination. 
+We believe that our efforts will help build a better mechanistic understanding in our readers in addition to making them aware of our results.}
 
@@
 \reply
+\yb{We agree that a long set of theoretical results can be tedious.
+We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts'' and result ``'It shortens them''.}
 
@@
 \reply
-\plr{yaniv doing that}
+\yb{We agree that this representation has likely been derived many times.
+We turned to Hartl and Clark and found a similar derivation.
+We therefore signpost that this is a standard result by pointing readers to that text.}
 
@@
 \reply
-\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}
-\plr{I agree: they are probably thinking of selected loci, not linked ones. We're talking about global drift not local drift. Yaniv to investigate.}
+\yb{Thanks for spotting this.
+Because we do nothing quantitative with respect to drift here, we simply state the our observed wobble of the center of the tension zone is consistent with previous theory.
+However, more broadly throughout the manuscript we are not sure about  relying on $N \sigma \sqrt{s}$ as we are concerned with loci linked to the selected site, and we are not aware of results extending this rule to an unlinked locus.}
+%\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}
+%\plr{I agree: they are probably thinking of selected loci, not linked ones. We're talking about global drift not local drift. Yaniv to investigate.}
 
@@
 
-\reply
+\reply This has been added. Thanks.
 

### 2016-04-25 drafts/hybrid_zone.tex
yb did his best to deal with reviews

@@
 and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
+\yb{We note that because recombination occurs within a deme, the recombination terms in equation \eqref{eqn:q_pde} are quite standard  \citep{HartlClark}}
 %\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
@@
 The agreement is good over hundreds of generations, but by $T=1280$ generations,
-stochasticity has substantially moved allele frequencies range-wide
-\citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
+stochasticity has substantially moved allele frequencies range-wide, \yb{which is in accordance with previous theory showing that tension zones wobble at a rate proportional $\sqrt{T}$ \citep{Barton1979a}.
+}
+%\citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
 \revpoint{2}{6}

### 2016-04-25 drafts/hybrid_zone.tex
reworked the intro -- please read

@@
 creates a gradient of alleles derived from the two populations across geographic space, %YB removed `assignments' as it seems like it is the ancestry, not its assignment, that is of interest
-centered on their point of contact \citep[reviewed in][]{Barton1985}. 
+centered on their point of contact \citep[reviewed by][]{Barton1985}. 
 If the populations are sufficiently diverged, 
@@
 % Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
-These
-patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.  
-\yb{We show that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is, 
+These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
+
+The mixing of genomes across space, combined with selection, results in the formation of genetic clines.
+These patterns have been the traditional focus of both empirical and theoretical studies into hybrid zones.
+While here we focus on hybrid zones maintained by selection against hybrids -- \emph{tension zones} --
+as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}, clines may also be differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies}, generating \emph{ecotones}.
+Although these are, in principle different, theoretical analyses consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, with the primary difference being that the center of a tension zone fluctuates over time \citep{Barton1979a} because stochastic changes in allele frequency can shit the cline center, while centers of ecotones do not change unless the environmental gradient changes.
+
+Our work here builds on the large body of theory describing tension zones.
+While much of the previous work describes equilibrium states of genetic clines, we focus on \emph{transient} patterns after the initial formation of a tension zone, in particular describing \emph{haplotype} patterns.
+Although the selected allele frequency cline establishes quickly, transient patterns can persist for quite some time, and  haplotypes can potentially carry substantial information about the formation of recent zones.
+To do this, we take the reverse-time perspective,
+understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
+This is particularly suited to understanding transient patterns.
+
+The arrangement of segments of genome inherited from particular ancestral populations
+is shaped by the number of recombination events that have taken place between ancestral genotypes. 
+Therefore the length distribution of such segments contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact both in single populations and in hybrid zones \citep{Price2009, Gravel2012,sedghifar2015spatial}.
+In addition to allowing researchers to interpret non-equilibrium patterns of diversity in hybrid zones,
+the length distribution of ancestry tracts could reveal the impact of ongoing selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
+%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
+%If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent %of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
+\yb{We show here that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.
+][]{Porter1997, Gompert2012}.
+Specifically, we find that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is,
 because selection rapidly removes incompatible alleles from heterospecific genomes, hybrid incompatibilities will be surrounded by disproportionately long ancestry blocks.}
+
+Using a combination of theory and simulated hybrid zones, we describe genome-wide patterns of coancestry 
+as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
+In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
+obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
+In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
+by describing how lineages move conditional on the frequencies of the selected allele
+and allowing lineages to recombine between selected background 
+(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
+
+If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected allele will be removed rapidly from the population, preventing introgression of surrounding genomic regions.
+We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
+As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus is expected to be relatively long when far away from the zone center.
+This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
+as many ancestors of the locally common type as do neutral haplotypes.
+\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotone models of extrinsic ecological selection (e.g.
+local adaptation).}
+%\plr{(ecological what?)}
+
+\yb{Because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
+incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.   
+This approach} is common to nearly all studies is the use of the
+reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
+or the discrete analogues \citep{hanson1966effects}. There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
+and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
+\plr{Need to cite Durrett and Cox's stuff here, especially \citet{durrett2007width}.}
+\yb{While we cannot analytically derive expectations for finite populations, our simulation results (below) suggest that genetic drift does not substantially influence our conclusions, perhaps because of the short time-scales that we examine.} \alisa{Our results present a possible approach for identifying genomic regions experiencing selection in hybrid zones.}
+
+%\citet{Slatkin1973} pointed out, in a far-reaching general analysis,
+%that while the environment may determine the location of a cline,
+%the fitness of the heterozygote will likely determine the shape of the cline,
+%and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
+%\plr{(What forms of selection?)}
 
@@
 %to better understand the processes operating in hybrid zones.
-
-\yb{Our work builds on the large body of theory on genetic clines. Specifically,  
-we focus on genetic clines maintained} by selection against hybrids -- \emph{tension zones} --
-as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}. 
-\yb{Clines may also be maintained by differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies}, generating \emph{ecotones}. 
-Although these are, in principle different, theoretical analyses consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, with the primary difference being that the center of a tension zone wobbles over time \citep{Barton1979a} because stochastic changes in allele frequency can shit the cline center, while centers of ecotones do not change unless the environmental gradient changes.
-}
-%\citet{Slatkin1973} pointed out, in a far-reaching general analysis,
-%that while the environment may determine the location of a cline,
-%the fitness of the heterozygote will likely determine the shape of the cline,
-%and \citet{barton1986barrier} show that \alisa{clines under several different forms of selection are similar in shape.}
-%\plr{(What forms of selection?)}
-There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
-and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000} \alisa{and much of this focuses on equilibrium states}.
-
-In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
-in particular describing haplotype patterns.
-Although the selected allele frequency cline establishes quickly, transient patterns can persist for quite some time,
-and  haplotypes can potentially carry substantial information about the formation of recent zones.
-To do this, we take the reverse-time perspective,
-understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
-This is particularly suited to understanding transient patterns.  
-\yb{However, because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
-incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.   
-This approach} is common to nearly all studies is the use of the
-reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
-or the discrete analogues \citep{hanson1966effects}.
-
-\plr{Need to cite Durrett and Cox's stuff here, especially \citet{durrett2007width}.}
-\yb{While we cannot analytically derive expectations for finite populations, our simulation results (below) suggest that genetic drift does not substantially influence our conclusions, perhaps because of the short time-scales that we examine.}
 
@@
 % Our aim here is to extend this to include expectations for how selection shapes covariance in ancestry across the genome.}
-
-The arrangement of segments of genome inherited from particular ancestral populations
-is shaped by the number of recombination events that have taken place between ancestral genotypes. 
-Therefore the length distribution of such segments contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome.
-% too much `provides information'
-The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact \citep{Price2009, Gravel2012,sedghifar2015spatial}.
-In addition to allowing researchers to interpret non-equilibrium patterns of diversity in hybrid zones, 
-the length distribution of ancestry tracts could reveal the impact of ongoing selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
-%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
-%If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent %of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
-\yb{We show that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.
-][]{Porter1997, Gompert2012}.
-}
-
-If hybrid ancestry at a given locus is disfavored, migrant haplotypes containing the selected allele will be removed rapidly from the population, preventing introgression of surrounding genomic regions.
-We therefore expect a deficit of short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
-As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus is expected to be relatively long when far away from the zone center.
-This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes.
-\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotones generated by extrinsic ecological selection (e.g.
-local adaptation).}
-%\plr{(ecological what?)}
-
-Using a combination of theory and simulated hybrid zones, we describe genome-wide patterns of coancestry 
-as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
-In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
-obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
-In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
-by describing how lineages move conditional on the frequencies of the selected allele
-and allowing lineages to recombine between selected background 
-(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
 

### 2016-04-25 drafts/review-responses-april-2016.tex
added blockLengthPlots

@@
 \yb{We agree that a long set of theoretical results can be tedious.
-We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts'' and result ``'It shortens them''.}
+We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts?'' }
 
@@
 \reply
-\alisa{Working on how to fix this in the ms now}
+We added an introductory sentence here to provide some structure. We have also gone through to clarify other sections that felt unclear.
 
@@
 \plr{also run sims in part of param range where nothing's supposed to happen?}
-\plr{alisa to add $s=.001$ and the like}
 

### 2016-04-25 drafts/hybrid_zone.tex
added blockLengthPlots

@@
 \end{align}
-\alisa{Need to be related to Figure 1}
 Here the first term ($G^A g_A$) represents spatial mixing,

### 2016-04-25 drafts/review-responses-april-2016.tex
some changes

@@
 \reply
-\alisa{We have remade Figure 1, with additional information that we hope will help with interpretation}
-\plr{alisa to refer to these events in discussion of terms in figure 5} \alisa{Did we mean Eq. 5?}
+\alisa{We have remade Figure 1, with additional information that we hope will help with interpretation. We have also related this figure to the terms in Equation 5.}
 

### 2016-04-25 drafts/supplement.tex
some changes

@@
 \includegraphics{figs/blocksAlongChromNoConditioning.pdf}
-\caption{Heatmap of mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
+\caption{Mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
 \end{figure}
@@
         \caption{
-     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (upper) and without conditioning (lower) alisa{is this perhaps an instance where we would need to have more demes to see a signal?
-    }
-    } \label{sfig:blockLengthPlot_t1000_s0.001}
+     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (top) and without conditioning (bottom) } \label{sfig:blockLengthPlot_t1000_s0.001}
 \end{figure}

### 2016-04-25 drafts/hybrid_zone.tex
some more minor changes

@@
 First, we turn our attention to how the cline at the selected locus itself is formed. 
-\alisa{This has been previously described, but is presented here as it provides a framework for further development of our model.}
+\alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,May1975,Nagylaki1975}}
 \plr{(I presume it's been described, but I don't think we found a reference that actually did this, right?
@@
 
-Power will be optimal at intermediate values of $s$.
-If selection is too strong, it may be difficult to observe the signal
+Power will be optimal at intermediate values of $s$ \alisa{relative to $\sigma$}.
+If selection is too strong, it may be difficult to observe this signal
 due to a lack of introgressed selected sites,
-while if selection is weak, the selected lineage does not move very fast,
-and so the strength of signal from elevated $l_B(m)$ is diminished. 
+while if selection is weak, the selected lineage does not move very fast \alisa{back in time},
+and so the strength of signal from elevated $l_B(m)$ is diminished \alisa{(Fig.~\ref{sfig:blockLengthPlot_t1000_s0.001})}. 
 Indeed, if $s$ is less than $x/(\sigma T)$, then only one block of ancestry $A$ is expected to be seen about a selected $A$ allele,

### 2016-04-25 drafts/review-responses-april-2016.tex
fixed response 2.5

@@
 \reply
-\alisa{We believe that the displacement is likely due to drift. (We could do a repeated sim to show this?)}
+\alisa{We have switched figures S2 and 3 and fixed the colors.
+The colored lines are not all represented in the legend which seems to have caused some confusion and we have tried to address this in the figure legends.
+We don't interpret the mismatch as a displacement to the right, as translating the lines horizontally will not improve the overall fit, and attribute this to drift in the simulations.}
 \plr{I'm ok with switching S2 and 3; that seems like a good idea.
@@
 If I recall correctly, we thought maybe it was drift but it wasn't.}
-\plr{Alisa to clarify figure legend.}
-\plr{peter or alisa fix S2: this is in \url{Plot_frequency_clines.R}}
 
@@
 \reply
-\alisa{Ok. This probably needs to be fixed.}
-\plr{alisa to write a good README: what scripts to run to get a simulation; where each figure gets produced}
+\alisa{We have created a new repository with minimally required scripts and a descriptive README file}
 

### 2016-04-25 sims/make-plots.R
plotting

@@ -1,0 +1,22 @@
+# run haplotype-plots.R
+# on all _simsums.Robj files passed in on the command line
+
+the.files <- commandArgs(TRUE)
+
+get_simparams <- function (fname) {
+    strings <- c( SIGMA="SIGMA", ninds="Ninds", ndemes="ndemes", S="s", tau="tau", run.id="runid_" )
+    out <- lapply( strings, function (x) { as.numeric(gsub(sprintf(".*[/_]%s([0-9.]+)_.*",x),"\\1",fname)) } )
+    out$zone_age <- out$tau
+    out$deme_size <- out$ninds/out$ndemes
+    return(out)
+}
+
+for (fname in the.files) {
+    if (!file.exists(fname)) {
+        warn(sprintf("File %s does not exist.\n",fname))
+        break
+    }
+    params <- get_simparams(fname)
+    load(fname)
+    source("haplotype-plots.R")
+}

### 2016-04-25 drafts/supplement.tex

@@
      using 50 demes, each with 500 diploid individuals and $\sigma=1$.
-    Each line represents a locus some distance $r$ away from the true target of selection.
+    Each line represents a locus some distance $r$ away from the true target of selection with colors corresponding to different values of $r$, transitioning from red (tight linkage to selected site) to blue (distantly linked).
     Grey lines represent the same positions from a simulation with

### 2016-04-25 drafts/hybrid_zone.tex
fixed response 2.5
notes and ref to may, endler, mcmutrie

@@
 % Population genetic modeling of hybrid zones has focused largely on the 
-% equilibrium states of clines at selected loci \citep{Nagylaki1975,barton1979dynamics,barton1986barrier,Slatkin1973,Slatkin1975,May1975,Haldane1948}.  
+% equilibrium states of clines at selected loci \citep{Nagylaki1975,barton1979dynamics,barton1986barrier,Slatkin1973,Slatkin1975,may1975gene,Haldane1948}.  
 % Since stronger selection reduces the amount of genetic mixing,
@@
 First, we turn our attention to how the cline at the selected locus itself is formed. 
-\alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,May1975,Nagylaki1975}}
+\alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,may1975gene,Nagylaki1975}}
 \plr{(I presume it's been described, but I don't think we found a reference that actually did this, right?
@@
         Each line represents a locus some distance $r$ away from the true target of selection, 
-        and $r=0$ represents the locus that is under selection.    
+        and $r=0$ represents the locus that is under selection. Colors correspond to different values of $r$, transitioning from red (tight linkage to selected site) to blue (distantly linked).
     }\label{sfig:alleleFreq_tau100_comparison}

### 2016-04-25 drafts/review-responses-april-2016.tex
minor edits

@@
 \reply
-Well, we hope to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted the wording to say ``\emph{realistic, individual-based models.}''.
+Well, we hoped to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted this to say ``realistic, individual-based models.''.
 %\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
@@
 \reply
-\alisa{Our decision to keep the derivations in appendices was with consideration for the readership of Molecular Ecology. 
-Thus, while some results are fairly standard, they might not be familiar to our readership.  
-One of our concerns is that the math would be largely overlooked if moved into the supplement (which is not automatically printed with the main document) and this could create a barrier to understanding among the readers.}
+We're not strongly against moving the appendices to supplementary material,
+but we'd prefer to keep them as appendices (and thus distributed with the document)
+because although the results are fairly standard, 
+they might not be familiar to our readership.  
+We think it is also useful to have a more detailed description of a concrete model
+that falls in the domain of attraction that we discuss here.
 
@@
 
-\begin{point}{}
-    A serious limitation of the test is that it only detects selection against heterozygotes. The claim that this would not affect results, on l.453-456:
+\begin{point}{\revref}
+    A serious limitation of the test is that it only detects selection against heterozygotes. They claim that this would not affect results, [previously at] l.453-456:
     \textit{``Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models''}
@@
 \reply
-\alisa{We agree with the comments of the AE on the matter, and have added some additional text to the introduction that hopefully convinces that this is ok.}
+\alisa{We agree with the comments of the AE on the matter, and have added additional text to the introduction \plr{(where?)} and to the referenced section \revref.}
+\plr{cite \citet{may1975gene} for this}
 
@@
 \reply
-\alisa{This is a point we have considered ourselves.
-We present these equations as a way of building our model up to address haplotype lengths.
-Both analytical expressions and simulations are presented so that they each support the conclusions of the other.} \yb{However, we also believe that analytical results can allow for a deeper understanding than simulation output.
+We also pondered this point, and feel that the equations contribute substantially
+to building intuitive understanding for how to address haplotype lengths,
+and aim to present analytical expressions and simulations so that they each support the conclusions of the other.
+We believe that analytical results can allow for a deeper understanding than simulation output,
+as for instance the intuitive discussion at \llname{ll:discussion_summary},
+which mostly derives from the analytical model.
 To aid our readers in developing this understanding, we have gone through considerable pains to map terms in our equations to the biological processes of migration, selection and recombination. 
-We believe that our efforts will help build a better mechanistic understanding in our readers in addition to making them aware of our results.}
+We believe that our efforts will help build a better mechanistic understanding in our readers in addition to making them aware of our results.
 
@@
 \reply
-\yb{We agree that a long set of theoretical results can be tedious.
-We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts?'' }
+\yb{
+    We agree that a long set of theoretical results can be tedious.
+We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts?''
+}
 
@@
 \reply
-\yb{We agree that this representation has likely been derived many times.
-We turned to Hartl and Clark and found a similar derivation.
-We therefore signpost that this is a standard result by pointing readers to that text.}
+We agree that this representation has likely been derived many times. 
+We turned to Hartl and Clark and found a similar derivation, which is now cited in this point in the text.
 
@@
 \begin{point}{}
-As an example of a section that felt unstructured, consider the discussion starting at l.347.
+As an example of a section that felt unstructured, consider the discussion starting at \revref.
 This long paragraph starts with a long list of observations, without giving a hint as to where we are going.
@@
 \reply
-We added an introductory sentence here to provide some structure. We have also gone through to clarify other sections that felt unclear.
+\plr{(comment is more general than that particular location; say something about this)}
+We added an introductory sentence at \revref to provide some structure. 
+We have also gone through to clarify other sections that felt unclear.
 
@@
 we've inserted comments to this effect, at \llname{ll:rescaling} \plr{and elsewhere?}.
+\plr{cite \citet{may1975gene} for the time-scaling}
 
@@
 Although most of the simulations shown in the main text have high population densities,
-we have in fact run simulations with considerably lower densities, 
+we have run simulations with considerably lower densities, 
 as shown for instance in figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}.
-\alisa{It is true that the demes are never empty for very large populations. We however, did encounter empty demes in some simulations will smaller populations sizes.}
-\plr{(It happened in sims that I ran.)}
-\alisa{However, this is the way the simulation is coded in order to account for the possibility if smaller population sizes are ever simulated.}
+Demes are never empty for very large populations, but
+we did encounter empty demes in some simulations with smaller population sizes.
 \plr{add reference to more low pop density figures if added}
@@
 We don't interpret the mismatch as a displacement to the right, as translating the lines horizontally will not improve the overall fit, and attribute this to drift in the simulations.}
-\plr{I'm ok with switching S2 and 3; that seems like a good idea.
-About the horizontal shift: I think that when I noticed this we figured out that it was because the center of the cline wasn't being plotted quite at zero for some reason?
-If I recall correctly, we thought maybe it was drift but it wasn't.}
 
@@
 However, more broadly throughout the manuscript we are not sure about  relying on $N \sigma \sqrt{s}$ as we are concerned with loci linked to the selected site, and we are not aware of results extending this rule to an unlinked locus.}
+\plr{(Do you mean "linked locus"?)}
 %\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}

### 2016-04-25 drafts/hybrid_zone.tex
minor edits

@@
 and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
-\yb{We note that because recombination occurs within a deme, the recombination terms in equation \eqref{eqn:q_pde} are quite standard  \citep{HartlClark}}
+\yb{We note that because recombination occurs within a deme, the recombination terms in equation \eqref{eqn:q_pde} are quite standard \citep{HartlClark}.} \revpoint{1}{4}
 %\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
@@
 genomic regions are inherited from ancestors $T$ generations ago
-in blocks of size roughly $1/T$.
+in blocks of size roughly $1/T$. \revpoint{1}{6}
 There will be a long segment of $A$ ancestry
@@
 
-Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
+Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e.
+non-equilibrium) hybrid zone.
+\linelabel{ll:discussion_summary}
 We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
@@
 \paragraph{The nature of selection}
-Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}).
+Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, 
+we believe that this is unlikely to drastically influence our findings, 
+as previous studies have demonstrated that models of underdominant selection 
+share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}). 
+\revpoint{1}{1}
+\plr{(more detail here or reference intro. would like to say `indistinguishable'; who can we cite?)}
 

### 2016-04-26 drafts/review-responses-april-2016.tex
peter added refs to intro and misc

@@
 \reply
-\alisa{We agree with the comments of the AE on the matter, and have added additional text to the introduction \plr{(where?)} and to the referenced section \revref.}
-\plr{cite \citet{may1975gene} for this}
+We had failed to cite the substantial literature on the matter (referred to by the AE);
+we have now added pointers to this in the introduction \llname{ll:all_clines_are_the_same} and to the referenced section \revref.
+The point is made quite well by \citet{Slatkin1973}: the details of the shape are determined mostly by the mean fitness of heterozygotes,
+a parameter that collapses many models (approximately) onto each other,
+at least after drift and environmental heterogeneity is considered.
 

### 2016-04-26 drafts/hybrid_zone.tex
peter added refs to intro and misc

@@
 Here, we use a combination of analytical theory and explicit forward simulations
-to show how selection against hybrid genotypes impacts patterns of introgression
+to describe how selection against hybrid genotypes impacts patterns of introgression
 across genomic and geographic space. 
@@
 These patterns have been the traditional focus of both empirical and theoretical studies into hybrid zones. 
-While here we focus on hybrid zones maintained by selection against hybrids -- \emph{tension zones} --
-as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}, clines may also be differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies}, generating \emph{ecotones}.
-Although these are, in principle different, theoretical analyses consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, with the primary difference being that the center of a tension zone fluctuates over time \citep{Barton1979a} because stochastic changes in allele frequency can shit the cline center, while centers of ecotones do not change unless the environmental gradient changes.
+While here we focus on hybrid zones maintained by selection against hybrids 
+-- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
+clines may also be due to differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies,hanson1966effects},
+% generating \emph{ecotones}.  % I think ecotone means something somewhat different: see Barton & Hewitt 1985, p. 132, for instance
+Although these are in principle different,  \linelabel{ll:all_clines_are_the_same}
+theoretical and simulation studies show that
+genomic clines seen in the two situations are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999},
+with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
+% consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, 
+% \plr{(I can't find discussion of similarity of different types of clines in \citet{barton1986barrier}: put the citation back in if I missed it.)}
+%% don't think we need this next bit
+% The primary difference being that the center of a tension zone fluctuates over time \citep{barton1979dynamics} 
+% because stochastic changes in allele frequency can shit the cline center, 
+% while centers of ecotones do not change unless the environmental gradient changes.
 
 Our work here builds on the large body of theory describing tension zones. 
-While much of the previous work describes equilibrium states of genetic clines, we focus on \emph{transient} patterns after the initial formation of a tension zone, in particular describing \emph{haplotype} patterns.
-Although the selected allele frequency cline establishes quickly, transient patterns can persist for quite some time, and  haplotypes can potentially carry substantial information about the formation of recent zones.
+While much of this previous work describes equilibrium states of genetic clines, 
+we focus on \emph{transients}, in particular describing haplotype patterns. 
+Although the cline in selected allele frequency establishes quickly, transient patterns can persist for quite some time, 
+and haplotypes can potentially carry substantial information about the formation of recent zones.
 To do this, we take the reverse-time perspective,
-understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
-This is particularly suited to understanding transient patterns.
-
-The arrangement of segments of genome inherited from particular ancestral populations
+understanding the temporal dynamics of the system by describing properties of the lineages tracing back from the zone.
+% This is particularly suited to understanding transient patterns.
+
+The arrangement of segments of genome inherited from particular ancestral populations,
+which may be identifiable as distinct haplotypes,
 is shaped by the number of recombination events that have taken place between ancestral genotypes. 
@@
 %If hybrids have non-zero fitness, some amount of introgression can occur, and the distribution of ancestry tract length in such populations has been used to infer the extent %of hybridization and timing of secondary contact \citep{Price2009, Gravel2012}.
-\yb{We show here that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.
-][]{Porter1997, Gompert2012}.
+
+\yb{We show here that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.,][]{Porter1997, Gompert2012}.
 Specifically, we find that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is,
@@
 \yb{Because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
-incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.   
-This approach} is common to nearly all studies is the use of the
-reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
-or the discrete analogues \citep{hanson1966effects}. There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,barton2008effect,Polechova2011},
+incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.}
+This approach is common to nearly all theoretical work,
+that study the reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
+or the discrete analogues \citep{hanson1966effects}. 
+There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,durrett2007width,barton2008effect,Polechova2011},
 and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
-\plr{Need to cite Durrett and Cox's stuff here, especially \citet{durrett2007width}.}
-\yb{While we cannot analytically derive expectations for finite populations, our simulation results (below) suggest that genetic drift does not substantially influence our conclusions, perhaps because of the short time-scales that we examine.} \alisa{Our results present a possible approach for identifying genomic regions experiencing selection in hybrid zones.}
+\plr{(why are references to 'barrier to gene flow' in the paragraph on drift? perhaps move to discussion?)}
+While we do not attempt to account for finite population size in our analytic results,
+our simulation results (below) suggest that although genetic drift strongly affects local noisiness of the system,
+it does not substantially influence our conclusions at moderate population densities
+(hundreds within one dispersal distance),
+perhaps because of the short time-scales that we examine.
+\plr{(maybe remove this next sentence? but want to say something about the very low densities we don't examine.)}
+At much lower population densities, local extinction and recolonization are expected to have a stronger effect.
+
+\plr{(nice wrap-up sentence shouldn't be at the end of paragraph on drift:)}
+\alisa{Our results present a possible approach for identifying genomic regions experiencing selection in hybrid zones.}
 
@@
 \section*{Acknowledgements}
-We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Nick Barton and two anonymous reviewers for helpful feedback and suggestions. \alisa{More info here}
+We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Nick Barton and two anonymous reviewers for helpful feedback and suggestions.
+PR was supported in part by NSF grant DBI-1262645 and a Sloan Research Fellowship.
+\alisa{More info here}
 

### 2016-04-26 drafts/review-responses-april-2016.tex
peter edited through

@@
 \reply
-There might be some confusion about our use of the floor function, which we now explicitly explain.
+There might have been some confusion about our use of the floor function?
+We now describe the dispersal distribution in a more straightforward way \revref.
 It's a good point more generally that if $\sigma$ is smaller than the deme spacing this scheme won't get dispersal with $\sigma$ close to the $\sigma$ that we plug into the Gaussian;
@@
 We believe that our efforts will help build a better mechanistic understanding in our readers in addition to making them aware of our results.
+
+Furthermore, as noted at \llname{ll:ld_comps}, the equations can be used to generate expectations for LD much more efficiently than forward-time simulation.
 
@@
 The rescaling arguments we provide  are of a standard sort in the literature;
-we've inserted comments to this effect, at \llname{ll:rescaling} \plr{and elsewhere?}.
-\plr{cite \citet{may1975gene} for the time-scaling}
-
-\plr{alisa to see if we can add any more from what we've done}
+we've inserted comments to this effect, at \llname{ll:rescaling} and \llname{ll:may_citation}.
 
@@
 \reply
-\yb{Thanks for spotting this.
-Because we do nothing quantitative with respect to drift here, we simply state the our observed wobble of the center of the tension zone is consistent with previous theory.
+\plr{We meant that the allele frequencies in the \emph{entire population} had shifted, which is determined by the total (effective) population size.  
+However, we decied to }
+% \yb{Thanks for spotting this. Because we do nothing quantitative with respect to drift here, we}
+\yb{simply state the our observed wobble of the center of the tension zone is consistent with previous theory.
 However, more broadly throughout the manuscript we are not sure about relying on $N \sigma \sqrt{s}$ as we are concerned with loci linked to the selected site, and we are not aware of results extending this rule to an unlinked locus.}
@@
 \reply
-\plr{see above}
+As noted above, we have provided some citations to similar arguments in the literature
+(although such arguments are often used without citation; see, for instance, \citet{durrett2007width}).
 
@@
 \reply
-\plr{see above; add something to discussion.}
+We now have much more discussion about drift, in particular at \revref.
 
+\plr{I don't understand this:}
 \alisa{We should add a line or two in the discussion about drift.

### 2016-04-26 drafts/hybrid_zone.tex
peter edited through

@@
 as many ancestors of the locally common type as do neutral haplotypes.
-\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotone models of extrinsic ecological selection (e.g.
-local adaptation).}
+\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotone models of extrinsic ecological selection (e.g., local adaptation).}
 %\plr{(ecological what?)}
@@
 \yb{Because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
-incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.}
+incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.} \revpoint{2}{9}
 This approach is common to nearly all theoretical work,
@@
       % previously had wrong Barton '79 here
-%Studies of hybrid zones have primarily been based on single-site statistics (e.g. clines),
+%Studies of hybrid zones have primarily been based on single-site statistics (e.g., clines),
 %but the recent availability of dense genotype data in combination with linkage information 
@@
 Although most selection in hybrid zones is likely more complex than this simple situation, 
-previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}, \yb{because, selection on an allele depends on its marginal effect on fitness.
+previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}, 
+\yb{because selection on an allele depends on its marginal effect on fitness.  
 Exceptionally strong epistatic selection which acts more quickly than free recombination may therefore, yield results that differ slightly from ours. 
 Likewise, while we do not model ecological selection, previous theory shows that clines formed by under dominant selection and extransic selection are essential indistinguishable, especially over short time scales when there is little time for clines to wobble \citep{Barton1979a}}.
-
-As in \citet{sedghifar2015spatial}, we consider where the ancestors of modern-day 
+\plr{(Merge paragraph from into into here?)}
+
+As in \citet{sedghifar2015spatial}, we consider where the ancestors of modern-day individuals
 fall across geography as we look further back towards the time of first contact.  
-We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an individual of species $A$,
+We say that a locus in a sampled individual is of \emph{ancestry} $A$ if it has been inherited from an ancestor of species $A$,
 i.e., if its lineage traces back to a spatial position $x<0$  at the time of secondary contact.
@@
     At the time of sampling ($\tau=0$), the segment contains the $B$ allele at the selected locus.
-    Paths of lineages in space, backwards in time, are depicted by blue and yellow lines. Splitting 
-    events  represent recombination events along the sampled segment.
-    The thick blue line represents the path of the selected locus. The colors of the other paths indicate 
-    the identity at the linked selected locus (blue: linked to allele $B$. yellow: linked to allele $A$). 
+    Paths traced by lineages across space are depicted by blue and yellow lines. 
+    Branching events occur when there was a recombination event within the sampled segment.
+    The thick blue line represents the path of the selected locus. 
+    \plr{(All lines look the same width: no thick and thin.)}
+    The colors of the other paths indicate the identity at the linked selected locus 
+    (blue: linked to allele $B$; yellow: linked to allele $A$). 
     The center of the hybrid zone is at geographic position $x=0$; in the region $x<0$, ancestry $B$ is less common,
-    and because of negative selection, the selected lineage spends little time there.
+    and because of selection, the selected lineage spends little time there.
   %  Red dots represent ancestral recombination events that fall in the given segment of genome in an ancestor;
@@
     from left to right the segments are of ancestry $(A,B,B)$.
+    % \plr{(Add blue/yellow bars at the bottom?)}
 }\label{Fig:schematic}
@@
 and analytics, mostly in terms of partial differential equations needing numerical solution. 
-\yb{Because of challenges in modeling spatial coalescence, we ignore the effects of drift and the pedigree. 
-Therefore, while our model incorporates stochastic migration, the genetics displays deterministic dynamics.}
-%\plr{(say why and what that means: e.g., deterministic dynamics)}
+\plr{As noted above, we ignore the effects of drift and the pedigree,
+so that the underlying allele frequency dynamics are deterministic.
+This should provide a good model for stochastic motion of a single lineage (our main focus),
+but does not model correlations between lineages.}
 
@@
 First, we turn our attention to how the cline at the selected locus itself is formed.
-\alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,may1975gene,Nagylaki1975}}
-\plr{(I presume it's been described, but I don't think we found a reference that actually did this, right?
-So maybe say it's ``well-known in the field'' or something?
-Also, no-one seems to have looked at lineages, and as our final point says, we actually get some more information out of this perspective.)} \alisa{So, when I sati's been previously described, I'm thinking about the traditional cline theory for a single locus (e.g.
-Slatkin, May), which describes the time-forward diffusion.}
+%% we don't need to cite these here, since we cite Bazykin for the equation
+% \alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,may1975gene,Nagylaki1975}}
 Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 \begin{align} \label{eqn:cline_pde}
-    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p +  2 p (1-p)  \times s (\frac{p}{2} - 1), %YB Rearranged this equation to allow for cleaner description
-%    \frac{d}{dt} p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
+    \frac{d}{dt} p = \frac{\sigma^2}{2} \frac{d^2}{dx^2} p +  2 p (1-p)  \times s (\frac{p}{2} - 1), %YB Rearranged this equation to allow for cleaner description
+    %    \frac{d}{dt} p = \frac{\sigma^2}{2} \frac{d}{dx^2} p + s p (1-p) (2p-1) ,
 \end{align}
-\yb{where $ \frac{d}{dt} p$ is the local rate of change of the frequency of $A$ ($\frac{d}{dt} p$), and  $\Delta p$ is the Laplacian\alisa{, i.e.
-the second derivative of allele frequency over space ($x$), at time $t$.}
-The first term,  $\sigma^2/2 \Delta p$, describes the net rate at which $A$ alleles arrive by local dispersal ($\sigma^2/2 \Delta p$). 
-The second term,  $2 p (1-p)  s (p/2-1)$, describes the impact of selection on local allele frequency as the product in local genotypic variance, $2 p (1-p)$, and selection on allele $A$, $s (p/2-1)$.
-The term $s (p/2-1)$ clearly describes  underdominance --- $A$ is favored when common ($p>.5$) and disfavored when rare ($p<.5$).}
+where $ \frac{d}{dt} p$ is the local rate of change of the frequency of $A$.
+The first term, $\frac{\sigma^2}{2} \frac{d^2}{dx^2} p$, describes the net rate at which $A$ alleles arrive by local dispersal.
+The second term % ,  $2 p (1-p)  s (p/2-1)$,  % only need this term in the sentence once?
+describes the impact of selection on local allele frequency as the product of local genotypic variance, 
+$2 p (1-p)$, and selection on allele $A$, $s (p/2-1)$. 
+\plr{ I don't think 'genotypic variance' is the right interpretation?  Let me try again: 
+The second term is the product of the local frequency of heterozygotes, $2p(1-p)$,
+and the strength of selection on allele $A$, $s (p/2-1)$.  }
+The term $s (p/2-1)$ describes underdominance -- $A$ is favored when common ($p>1/2$) and disfavored when rare ($p<1/2$).
 %YB: Pardon the extreme changes here, I tried to make things more obvious to a biologist
@@
 %which is $s(2p-1)$ (up to an error of size $s^2$).
-%This can be derived as in \citet{Nagylaki1975};
-%for discussion of diploidy see \citet{christiansen1995genotypic}.
+The equation can be derived as in \citet{Nagylaki1975};
+the assumption of locally random mating was shown to have only a small effect \citet{christiansen1995genotypic}.
 
@@
 the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially.
-Rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
+
+As noted by others \citep[e.g.][]{Slatkin1973,may1975gene}, \linelabel{ll:rescaling}
+rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
 results in a dimensionless equation,
 implying that the cline establishes over a timescale of the order $1/s$. 
-(This sort of rescaling is common in the literature, e.g., \citep{fisher1950frequencies,Slatkin1973}.)  \linelabel{ll:rescaling}
 Since this is how long it takes diffusion at rate $\sigma^2$ to smooth across a region of width $\sigma/\sqrt{s}$,
@@
 the description above applies to motion of a lineage transverse to the hybrid zone.  \revpoint{AE}{4}
-It is for this reason that we use $\Delta$ for the second derivative and $\grad$ for the first derivative,
-as these carry over without change to two dimensions;
-in fact, if the landscape is heterogeneous or dispersal is non-Gaussian,
-the theory still holds replacing $\Delta$ by the appropriate operator.
-Note, however, that in two dimensions drift can have a much stronger effect \citep{cox1995hybrid,barton2010phylogeography,durrett2007width}.
+In this case, $d^2/dx^2$ should be replaced by the Laplacian operator,
+and $d/dx$ by the gradient.
+In fact, if the landscape is heterogeneous or dispersal is non-Gaussian,
+the theory still holds replacing the Laplacian by the appropriate operator.
+Note, however, that with low population densities, 
+drift can have a strong effect that differs between one and two dimensional systems \citep{cox1995hybrid,barton2010phylogeography,durrett2007width}.
 
@@
 with mean displacement proportional to the derivative of $\log p$.
-This has the effect of pulling $A$ lineages towards regions where they are more common i.e. the left side of the range.   
+This has the effect of pulling $A$ lineages towards regions where they are more common, i.e., the left side of the range.   
 % with speed $\sigma$
@@
 To see why this is true, note that the probability that the parent of an $A$ allele found at $x$
-lived at position $x-r$ is proportional to $\P\{R=-r\} p(x-r)$, where $R$ is the random dispersal distance;
-and so the mean displacement from offspring to parent is $-\E[R p(x-R)]/\E[p(x-R)]$.  \revpoint{1}{9}
+lived at position $x-r$ is proportional to $\P\{R=r\} p(x-r)$, where $R$ is the random dispersal distance;
+and so the mean displacement from offspring to parent is $\E[R p(x-R)]/\E[p(x-R)]$.  \revpoint{1}{9}
 % $$ = \E[ -R ] - \E[R]^2 p'(x)/p(x) + \E[R^2] p'(x)/p(x) + O(R^3) $$
@@
 back towards the side where they are more common
-(since here $\log p$ is $\sqrt{s}/\sigma$ and the speed is $\sigma^2$). 
+(since here $\frac{d}{dx}\log p(x) \approx \sqrt{s}/\sigma$ and the speed is $\sigma^2$). 
 Since an $A$ allele must, by definition, have been inherited from the $A$ side of the barrier 
@@
 
-This description gives more information than the steady-state cline $p(x)$,
+This description gives more information than the steady-state cline,
 which depends only on $\sigma/\sqrt{s}$.
@@
 \yb{Suppose we sample an allele today $r$ Morgans from the selected site, and  
-follow its lineage back through time, using $\tau$ to denote ``generations ago'' (r
-reserving $t$ for time measured in the usual direction).
+follow its lineage back through time, using $\tau$ to denote ``generations ago''
+(reserving $t$ for time measured in the usual direction).
 If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
@@
  jumps between $A$ and $B$ with recombination events between the neutral and selected locus (with frequency $r$), that result in a change in the  selected background.    
-Thus, by the assumption of locally random assortment of alleles, $Z$ shifts from $A$ to $B$  at rate $r (1-p)$, and $B$ to $A$ at rate $r p$ (see Appendix \ref{apx:lineage_derivation} for a more precise description).}
+Thus, by the assumption of locally random assortment of alleles, $Z$ shifts from $A$ to $B$  at rate $r (1-p)$, and $B$ to $A$ at rate $r p$ 
+(see Appendix \ref{apx:lineage_derivation} for a more precise description).}
 
@@
     \begin{aligned} \label{eqn:lineage_motion}
-        dX_\tau &= dB_\tau + \begin{cases}
-             \sigma^2 \grad \log(p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = A \\
-             \sigma^2 \grad \log(1-p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = B 
+        dX_\tau &= \sigma dB_\tau + \begin{cases}
+            \sigma^2 \frac{d}{dx} \log(p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = A \\
+             \sigma^2 \frac{d}{dx} \log(1-p(X_\tau,\tau)) d\tau \qquad & \text{if } Z_\tau = B 
         \end{cases} \\
@@
 In the first expression, giving the distribution of how the lineage location changes,
-the first term ($\sigma dB_\tau$) is Brownian noise driven by dispersal, \alisa{This term is missing the $\sigma$ above...}
-and the second is the mean displacement, which moves the lineage ``downhill'' \alisa{towards their ancestral range} on either $-\log p$ or $-\log (1-p)$
-($\grad$ is the gradient),
+the first term ($\sigma dB_\tau$) is Brownian noise driven by dispersal,
+and the second is the mean displacement, which moves the lineage ``downhill'' towards its selected allele's ancestral range on either $-\log p$ or $-\log (1-p)$,
 depending on which selected allele the lineage is linked to.
-The second and third expressions \alisa{give the expected time for a lineage to switch from} an $A$ to $B$ (or $B$ to $A$) background as a result of recombination.
-This occurs at rate $r$ multiplied by the local proportion of $B$ alleles, which is $(1-p)$.
+(In two dimensions, $d/dx$ is replaced by the gradient.)
+\plr{
+The second expression gives the probability that a lineage on the $A$ background at time $\tau$
+does not switch to the $B$ background across $u$ generations,
+i.e., the probability that there was no recombination between the lineage and the selected site in those $u$ generations.
+Since the rate of recombination between the two is $r$,
+and the probability the lineage is in a heterozygote at time $s$ 
+is equal to the local proportion of $B$ alleles, $1-p(X_s,s)$,
+the mean number of recombinations in heterozygotes across that time period is $r \int_\tau^{\tau+u} (1-p(X_s,s)) ds$;
+assuming recombinations are Poisson, the probability of no such recombinations is the expression given.
+The third expression is the same, except for switching $B$ to $A$.
+}
+% gives the rate at which a lineage switches from an $A$ to $B$ background as a result of recombination,
+% which occurs at rate $r$ multiplied by the local proportion of $B$ alleles, which is $(1-p)$.
+% The third expression is the corresponding rate for switching from $B$ to $A$.
 
@@
     \frac{d}{dt} q_A(x,t,r) 
-            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
-                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
+            &= 
+                \frac{\sigma^2}{2} \frac{d^2}{dx^2} q_A(x,t,r) 
+                + \sigma^2 \frac{d}{dx} \log(p(x,t)) \cdot \frac{d}{dx} q_A(x,t,r) 
             \\ &\qquad {} + 
                 r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
-    \frac{d}{dt} q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
-            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
+        \frac{d}{dt} q_B(x,t,r) 
+            &= 
+                \frac{\sigma^2}{2} \frac{d^2}{dx^2} q_B(x,t,r)
+                + \sigma^2 \frac{d}{dx} \log(1-p(x,t)) \cdot \frac{d}{dx} q_B(x,t,r) 
             \\ &\qquad {} + 
@@
 \end{align}
-with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
-and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
-The three terms in these equations come from Brownian movement of a linages (i.e. the smoothing action of local dispersal),
-the tendency of a lineage to inherit from regions where its type is more common (i.e. the net flux induced by reduced fitness in the hybrid zone),
+with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$.
+% and where $\grad$ is the gradient.
+The three terms in these equations come from Brownian movement of a linages 
+(i.e., the smoothing action of local dispersal),
+the tendency of a lineage to inherit from regions where its type is more common 
+(i.e., the net flux induced by reduced fitness in the hybrid zone),
 and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
@@
 so define these as
-$G^A = \sigma^2 \grad \log(p(x,t)) \cdot \grad + (\sigma^2/2) \Delta$
+$G^A = \sigma^2 \frac{d}{dx} \log(p(x,t)) \cdot \frac{d}{dx} + (\sigma^2/2) \frac{d}{dx^2}$
 and
-$G^B = \sigma^2 \grad \log(1-p(x,t)) \cdot \grad + (\sigma^2/2) \Delta$,
+$G^B = \sigma^2 \frac{d}{dx} \log(1-p(x,t)) \cdot \frac{d}{dx} + (\sigma^2/2) \frac{d}{dx^2}$,
 so that equation \eqref{eqn:q_pde} can be written more compactly as
@@
 and the remaining portion becomes linked to a $B$ allele (as is the case in the most recent recombination event shown in Fig.~\ref{Fig:schematic}).
-\alisa{to account for cases where the selected locus lies outside the interval [a,b],} 
+To account for cases where the selected locus lies outside the interval [a,b],
 we say that the integral $\int_a^{\min(0,b)}$ is zero if  $a>0$ 
@@
 which can be done directly using our numerical code;
-see Appendix~\ref{apx:haplotype_calcs} for more detail.
+see Appendix~\ref{apx:haplotype_calcs} for more detail. \linelabel{ll:ld_comps}
 
@@
 Each deme has exactly $N_d$ diploid hermaphroditic individuals at the start of each generation.
-Then, every individual can migrate to a new deme offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
-where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
-(Here $\lfloor z \rfloor$ is the largest integer less than or equal to $z$,
-so this equivalently takes a centered Gaussian displacement and assigns it to the nearest deme. \revpoint{AE}{5}
-The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
+Then, every individual disperses to a (possibly) new deme
+by choosing a Gaussian displacement with mean zero and variance $\sigma^2$,
+then dispersing to the nearest deme. \revpoint{AE}{5}
+% can migrate to a new deme offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
+% where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
+% (Here $\lfloor z \rfloor$ is the largest integer less than or equal to $z$,
+% so this equivalently takes a centered Gaussian displacement and assigns it to the nearest deme. 
+(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
 Migrants past either end of the range remain at the terminal demes.
@@
         and $r=0$ represents the locus that is under selection. Colors correspond to different values of $r$, transitioning from red (tight linkage to selected site) to blue (distantly linked).
-    }\label{sfig:alleleFreq_tau100_comparison}
+    }\label{fig:alleleFreq_tau100_comparison}
 \end{figure}
@@
 
-We are interested in identifying regions of the genome with abnormal distributions of block lengths, 
-and at each spatial location $x$, compute $l_B(m,x)$ and $C(m,x)$ at a grid of positions across the genome. 
-Where indicated, we normalize $l_B(m,x)$ and $C(m,x)$ by dividing by the empirical mean of $l_B$ and $C$ across the genome for location $x$.
+\plr{Below we use $l_B$ to mean the normalized version also.  I am defining consistent notation, $\bar l_B$.}
+
+To identify regions of the genome with abnormal distributions of block lengths, 
+we compute $l_B(m,x)$ and $C(m,x)$ at a grid of positions across geography and across the genome. 
+From these, we compute normalized versions $\bar l_B(m,x)$ and $\bar C(m,x)$
+by dividing by the empirical mean across the genome for each geographic location $x$:
+for instance, $\bar l_B(m,x) = n_x l_B(m,x)/\sum_i l_B(m_i,x)$,
+where the sum is over the $n_x$ individuals at location $x$.
 While there is useful information contained in the geographic distribution of block length patterns, 
@@
 
-Although our analytical model does not account for coalescence, we are able to partially trace the genealogy of haplotypes in our simulations.
+We are able to partially trace the genealogy of haplotypes in our simulations. 
 In particular, we wish to know the relative number of ancestors from $T$ generations ago that are represented in present day populations at a given locus, given the frequency of a particular ancestry.
-This is a reflection of the average size of a haplotype family, and is potentially a source of additional information about selection.
-Within each deme $x$, we calculated $F_B(m,x)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, divided by the number of individuals of ancestry $B$ at site $m$.
+This is a reflection of the average size of a haplotype's family, 
+and is potentially a source of additional information about selection. 
+Within each deme $x$, we calculated $F_B(m,x)$, which is the number of independent ancestors from time $T$ contributing to the pool of alleles of ancestry $B$ at site $m$, 
+divided by the number of individuals of ancestry $B$ at site $m$.
 
@@
 and tightly linked sites flattening more slowly than distant sites.
-This is seen, for instance, in Figures \ref{alleleFreq_tau100},
-\ref{alleleFreq_tau100_weaker_s} and \ref{alleleFreq_tau1000}. 
+This is seen in Figure \ref{fig:alleleFreq_tau100_comparison},
+and supplementary figures \ref{alleleFreq_tau100}, \ref{alleleFreq_tau100_weaker_s} and \ref{alleleFreq_tau1000}. 
 Clines in ancestry frequency matched expected values found from deterministic theory
 up to stochasticity due to genetic drift, which is more pronounced at lower population densities.
-Figure \ref{sfig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
+Figure \ref{fig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
 match simulated clines in Figure \ref{alleleFreq_tau100} quite well.
@@
 and separating linked clines based on the identity of the linked selected allele.
-The agreement is good over hundreds of generations, but by $T=1280$ generations,
-stochasticity has substantially moved allele frequencies range-wide, \yb{which is in accordance with previous theory showing that tension zones wobble at a rate proportional $\sqrt{T}$ \citep{Barton1979a}.
-}
-%\citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
-\revpoint{2}{6}
-
-\plr{(This next bit is one of these scaling arguments that Reviewer 2 wants clarified.
-Again, I don't know who's described how the cline relaxes, do you?
-I think at least the "speed of a lineage" bit is new, but don't want to make a big deal out of it.
-Not sure how to signpost this.}
+The agreement is good over hundreds of generations, \revpoint{2}{6}
+although as the clines flatten they are seen to wobble
+(which happens at rate proportional to $\sqrt{T}$; \citet{Barton1979a}).
+%% \plr{I tried to rephrase this:}
+% but by $T=1280$ generations,
+% stochasticity has substantially moved allele frequencies range-wide,
+% \yb{which is in accordance with previous theory showing that tension zones wobble at a rate proportional to $\sqrt{T}$ \citep{Barton1979a}.}
+% %\citep[which is unsurprising, as the total population size is only 5,000 diploids:][]{Polechova2011}.
+% \plr{(With effective population size only thousands, once you get to 1,280 generations, it doesn't matter that you're simulating a spatial population;
+% drift will kick in and move frequencies \emph{range-wide}.
+% We see that in figure \ref{sfig:condAlleleFreq_tau1280_comparison}.
+% That's what I was trying to say here.
+% Total population size is not irrelevant.)}
+
 Loci not under direct selection can in principle spread across the cline unimpeded. 
@@
 i.e., if $T>\sqrt{x}$.
-This square-root flattening is seen in Figure \ref{fig:linked_cline_heatmaps}.
+This square-root flattening is seen in Figure \ref{fig:linked_cline_heatmaps}
+(and is discussed for environmentally determined clines by \citet{may1975gene}; but see \citet{durrett2007width}). \linelabel{ll:may_citation}
 A linked lineage must also spend at least $1/r$ generations in heterozygotes
@@
 Ideal information -- true ancestry assignments for a few simulated individuals sampled from across space -- % YB changed  `raw-data' to information, because ancestry will never be raw-data -- ATGC will
-are shown in Figures \ref{Fig:resistanceToIntrogression100g} ($T=100$ generations) and \ref{Fig:resistanceToIntrogression1000g} ($T=1000$ generations).
+are shown in Figure \ref{Fig:resistanceToIntrogression100g} ($T=100$ generations) and Figure \ref{Fig:resistanceToIntrogression1000g} ($T=1000$ generations).
 For the more recent hybrid zone ($T=100$), the selected cline has established,
 but linked clines are still flattening.
-After a longer period ($T=1000$), \alisa{I see, this is what reviewer was saying about flatness, but shouldn't our sims produce IBD?}
+After a longer period ($T=1000$),
 clines over much of the chromosome are flat
@@
 
-Power will be optimal at intermediate values of $s$ \alisa{relative to $\sigma$}.
+Power will be optimal at intermediate values of $s$.
 If selection is too strong, it may be difficult to observe this signal
 due to a lack of introgressed selected sites,
-while if selection is weak, the selected lineage does not move very fast \alisa{back in time},
+while if selection is weak, the selected lineage does not move very fast,
 and so the strength of signal from elevated $l_B(m)$ is diminished \alisa{(Fig.~\ref{sfig:blockLengthPlot_t1000_s0.001})}. 
@@
 \includegraphics{figs/blocksAlongChromAncBConditioningHighRes.pdf}
-\caption{\textbf{Normalized mean enclosing block length}, $l_B(m)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$.
-Here each line represents  $l_B(m)$, in a given deme some distance from the zone center, and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme.
+\caption{\textbf{Normalized mean enclosing block length}, $\bar l_B(m,x)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$. 
+    Here each line shows $\bar l_B(m,x)$, in a given deme some distance from the zone center. % , and is normalized by dividing by mean $l_B(m)$ along the chromosome within the deme. 
     Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals. 
-This figure represents the same simulation and statistic as Fig.~\ref{Fig:blockLengths}, on a finer genomic scale.
+    The same simulation and statistic are shown in Fig.~\ref{Fig:blockLengths}, on coarser genomic scale.
 }\label{Fig:blockLengthsZoom}
@@
 
-Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e.
-non-equilibrium) hybrid zone.
+Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e., non-equilibrium) hybrid zone.
 \linelabel{ll:discussion_summary}
-We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
+We show why clines establish over time $1/s$, 
+and why lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they are unfit. 
 This occurs at speed $\sigma\sqrt{s}$.
@@
 In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} 
-can inform the demographic history of hybridization/admixture from present day samples \citep[e.g. ][]{Hellenthal2014}. 
+can inform the demographic history of hybridization/admixture from present day samples \citep[e.g.][]{Hellenthal2014}. 
 We described how selection against hybrid incompatibilities  
@@
 
-Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
-Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage --  and similarly $C(k)$, could provide a simple and informative summary. 
-The ratio of mean adjacent block lengths $C(x,m)$ shows a very sharp peak at the selected site,
+Our results suggest that the statistic $l(x,m)$ 
+-- the mean genetic length of ancestry surrounding a site of a given lineage --
+could help identify selected loci in hybrid zones. 
+The ratio of mean adjacent block lengths, $C(x,m)$, alsot shows a very sharp peak at the selected site,
 suggesting that, despite the fact that haplotypes surrounding incompatibilities might be quite long, 
@@
 Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, 
-we believe that this is unlikely to drastically influence our findings, 
-as previous studies have demonstrated that models of underdominant selection 
-share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}). 
+we believe that this is unlikely to drastically influence our findings.
+As discussed in the Introduction,
+previous studies have demonstrated that models of underdominant selection 
+share similar properties with more realistic models, e.g.,
+selection on an ecological gradient or multilocus, epistatic selection
+\citep{Slatkin1973,Barton1993,Kruuk1999}.
 \revpoint{1}{1}
-\plr{(more detail here or reference intro. would like to say `indistinguishable'; who can we cite?)}
 
@@
 
-%One major assumption of our model is that selection acts against heterozygotes --i.e. that loci are underdominant. 
+%One major assumption of our model is that selection acts against heterozygotes -- i.e., that loci are underdominant. 
 %Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 

### 2016-04-26 drafts/supplement.tex
peter edited through

@@
     identical parameters except that $s=0$. 
-    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{sfig:alleleFreq_tau100_comparison};
+    Corresponding theoretical quantities are shown juxtaposed in Figure \ref{fig:alleleFreq_tau100_comparison};
     the same plot is shown with weaker selection in Figure \ref{alleleFreq_tau100_weaker_s} and at a longer time in Figure \ref{alleleFreq_tau1000}.
@@
         \textbf{Mean haplotype lengths of $B$ haplotypes}, $l_B(x,m)$,
-        across the genome (horizontal axis) and at differntial spatial locations (colored lines),
+        across the genome (horizontal axis) and at different spatial locations (colored lines),
         from a simulation with 50 demes having 500 individuals each, $s=0.01$, $\sigma=1$, and after $T=1000$ generations.
@@
 \caption{
-    \textbf{Normalized mean enclosing block length}, $l_B(m,x)$, after $T=1000$ generations,
+    \textbf{Normalized mean enclosing block length}, $\bar l_B(m,x)$, after $T=1000$ generations,
     against position relative to the selected locus (horizontal axis)
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning.pdf}
-\caption{Ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{The statistic $\bar C(m,x)$: the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
 Each line represents a deme and is normalized by mean block length across the chromosome in the deme.}\label{Supp:ratioBlockAdjacent}
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning.pdf}
-\caption{Heatmap of $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{Heatmap of $C(m,x)=\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
 }\label{Supp:ratioBlockAdjacentHeatmap}

### 2016-04-26 drafts/hybrid_zone.tex
reworked the intro

@@
 Clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
-and linked clines persist over a  genomic scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. 
+and linked clines persist over a  genomic scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. \alisa{This is in May1975? Should it be removed?}
 Locally disadvantageous alleles tend to exist as small families,
@@
 % Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
-These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,sedghifar2015spatial}.
-
-The mixing of genomes across space, combined with selection, results in the formation of genetic clines. 
-These patterns have been the traditional focus of both empirical and theoretical studies into hybrid zones. 
-While here we focus on hybrid zones maintained by selection against hybrids 
--- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
-clines may also be due to differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies,hanson1966effects},
-% generating \emph{ecotones}.  % I think ecotone means something somewhat different: see Barton & Hewitt 1985, p. 132, for instance
-Although these are in principle different,  \linelabel{ll:all_clines_are_the_same}
-theoretical and simulation studies show that
-genomic clines seen in the two situations are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999},
-with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
+These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2014,sedghifar2015spatial}.
+
+%The mixing of genomes across space, combined with selection, results in the formation of genetic clines. 
+%These patterns have been the traditional focus of both empirical and theoretical studies into hybrid zones. 
+%While here we focus on hybrid zones maintained by selection against hybrids 
+-%- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
+%clines may also be due to differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies,hanson1966effects},
+%% generating \emph{ecotones}.  % I think ecotone means something somewhat different: see Barton & Hewitt 1985, p. 132, for instance
+%Although these are in principle different,  \linelabel{ll:all_clines_are_the_same}
+%theoretical and simulation studies show that
+%genomic clines seen in the two situations are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999},
+%with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
 % consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, 
@@
 
-Our work here builds on the large body of theory describing tension zones. 
-While much of this previous work describes equilibrium states of genetic clines, 
-we focus on \emph{transients}, in particular describing haplotype patterns. 
-Although the cline in selected allele frequency establishes quickly, transient patterns can persist for quite some time, 
-and haplotypes can potentially carry substantial information about the formation of recent zones.
+Our work here builds on the large body of theory describing tension zones \alisa{and aims to describe \emph{transient} dynamics with a focus on \emph{haplotype} patterns.}
+%While much of this previous work describes equilibrium states of genetic clines, 
+%we focus on \emph{transients}, in particular describing haplotype patterns. 
+%Although the cline in selected allele frequency establishes quickly, transient patterns can persist for quite some time, 
+%and haplotypes can potentially carry substantial information about the formation of recent zones.
 To do this, we take the reverse-time perspective,
@@
 
-The arrangement of segments of genome inherited from particular ancestral populations,
-which may be identifiable as distinct haplotypes,
+The arrangement of segments of genome inherited from particular ancestral populations%, which may be identifiable as distinct haplotypes,
 is shaped by the number of recombination events that have taken place between ancestral genotypes. 
-Therefore the length distribution of such segments contains clues concerning the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
-The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact both in single populations and in hybrid zones \citep{Price2009, Gravel2012,sedghifar2015spatial}.
+The length distribution of such segments therefore contains information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+%The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact both in single populations and in hybrid zones \citep{Price2009, Gravel2012,sedghifar2015spatial}.
 In addition to allowing researchers to interpret non-equilibrium patterns of diversity in hybrid zones,
-the length distribution of ancestry tracts could reveal the impact of ongoing selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
+%the length distribution of ancestry tracts could reveal the impact of ongoing selection on the mixing of ancestral genomes. %and, consequently, the maintenance of species distinctness. 
 %In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
@@
 as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
-In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
-obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
-In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
-by describing how lineages move conditional on the frequencies of the selected allele
-and allowing lineages to recombine between selected background 
-(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
+\alisa{We extend the framework presented by \cite{Sedghifar2015} to describe how lineages move across a zone of secondary contact.
+We consider the influence of selection against heterozygotes at a single locus on the movement of lineages across space, conditioning on the frequencies of the selected allele and allowing lineages to recombine between selected background } (analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
+%In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
+%obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
+%In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
+%by describing how lineages move conditional on the frequencies of the selected allele
+%and allowing lineages to recombine between selected background 
+%(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
 
@@
 or the discrete analogues \citep{hanson1966effects}. 
-There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,durrett2007width,barton2008effect,Polechova2011},
+There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Nagylaki1978,durrett2007width,barton2008effect,Polechova2011},
 and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
@@
 
-\plr{(nice wrap-up sentence shouldn't be at the end of paragraph on drift:)}
-\alisa{Our results present a possible approach for identifying genomic regions experiencing selection in hybrid zones.}
+\plr{(nice wrap-up sentence shouldn't be at the end of paragraph on drift:)} \alisa{Is it weird to end intro on above paragraph? Could this wrap-up sentence hold up on its own? }
+\alisa{Our results look beyond the traditional signature of selection observed in the width of genetic clines, and present a possible new approach for identifying genomic regions experiencing selection in hybrid zones.}
 
@@
     Branching events occur when there was a recombination event within the sampled segment.
-    The thick blue line represents the path of the selected locus. 
-    \plr{(All lines look the same width: no thick and thin.)}
+    The \alisa{rightmost} blue line represents the path of the selected locus. 
+   % \plr{(All lines look the same width: no thick and thin.)}
     The colors of the other paths indicate the identity at the linked selected locus 

### 2016-04-26 drafts/hybrid_zone.tex
pulled together the arguments about cline generality

@@
 previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}, 
-\yb{because selection on an allele depends on its marginal effect on fitness.  
+\yb{because selection on an allele depends on its marginal effect on fitness.  }
 Exceptionally strong epistatic selection which acts more quickly than free recombination may therefore, yield results that differ slightly from ours. 
-Likewise, while we do not model ecological selection, previous theory shows that clines formed by under dominant selection and extransic selection are essential indistinguishable, especially over short time scales when there is little time for clines to wobble \citep{Barton1979a}}.
-\plr{(Merge paragraph from into into here?)}
+Likewise, while we do not model ecological selection, theoretical and simulation studies show that clines formed by underdominant selection and extrinsic selection are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999}, especially over short time scales \citep{Barton1979a}, \linelabel{ll:all_clines_are_the_same}
+with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
 

### 2016-04-26 drafts/hybrid_zone.tex
minor fixen

@@
 % Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
-These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2014,sedghifar2015spatial}.
+These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}.
 
@@
 %While here we focus on hybrid zones maintained by selection against hybrids 
--%- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
+%-- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
 %clines may also be due to differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies,hanson1966effects},
@@
 as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
-\alisa{We extend the framework presented by \cite{Sedghifar2015} to describe how lineages move across a zone of secondary contact.
+\alisa{We extend the framework presented by \cite{sedghifar2015spatial} to describe how lineages move across a zone of secondary contact.
 We consider the influence of selection against heterozygotes at a single locus on the movement of lineages across space, conditioning on the frequencies of the selected allele and allowing lineages to recombine between selected background } (analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
@@
 
-We can describe this process formally in It\^o notation:
+We can describe this Markov process formally in It\^o notation:
 with $B_\tau$ a standard Brownian motion,
@@
         \end{cases} \\
-        % \P\{ Z_{\tau+\epsilon} = B \given Z_\tau = A \} &= \epsilon r (1-p(X_\tau,\tau)) + O(\epsilon^2) \\
-        % \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B \} &= \epsilon r p(X_\tau,\tau) + O(\epsilon^2)  .
-        \P\{ T_B(\tau) &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
-        \P\{ T_A(\tau) &> \tau+u \given Z_\tau = B \} = \exp\left( - r \int_\tau^{\tau+u} p(X_s,s) ds \right) .
+        \P\{ Z_{\tau+\epsilon} = B \given Z_\tau = A, \; X_\tau=x \} &= \epsilon r (1-p(x,\tau)) + O(\epsilon^2) \\
+        \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B, \; X_\tau=x \} &= \epsilon r p(x,\tau) + O(\epsilon^2)  .
+        % \P\{ T_B(\tau) &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
+        % \P\{ T_A(\tau) &> \tau+u \given Z_\tau = B \} = \exp\left( - r \int_\tau^{\tau+u} p(X_s,s) ds \right) .
+        % \E[ T_B(\tau) &> \tau+u \given Z_\tau = A ] = r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \\
+        % \E[ T_A(\tau) &> \tau+u \given Z_\tau = B ] = r \int_\tau^{\tau+u} p(X_s,s) ds .
     \end{aligned}
@@
 \plr{
-The second expression gives the probability that a lineage on the $A$ background at time $\tau$
-does not switch to the $B$ background across $u$ generations,
-i.e., the probability that there was no recombination between the lineage and the selected site in those $u$ generations.
-Since the rate of recombination between the two is $r$,
-and the probability the lineage is in a heterozygote at time $s$ 
-is equal to the local proportion of $B$ alleles, $1-p(X_s,s)$,
-the mean number of recombinations in heterozygotes across that time period is $r \int_\tau^{\tau+u} (1-p(X_s,s)) ds$;
-assuming recombinations are Poisson, the probability of no such recombinations is the expression given.
-The third expression is the same, except for switching $B$ to $A$.
+The second expression says that the probability a lineage on the $A$ background at $x$
+recombines onto the $B$ background
+is equal to the recombination rate, $r$, multiplied by the local proportion of $B$ alleles, $(1-p(x,\tau))$,
+per generation.
 }
+% \plr{
+% The second expression gives the probability that a lineage on the $A$ background at time $\tau$
+% does not switch to the $B$ background across $u$ generations,
+% i.e., the probability that there was no recombination between the lineage and the selected site in those $u$ generations.
+% Since the rate of recombination between the two is $r$,
+% and the probability the lineage is in a heterozygote at time $s$ 
+% is equal to the local proportion of $B$ alleles, $1-p(X_s,s)$,
+% the mean number of recombinations in heterozygotes across that time period is $r \int_\tau^{\tau+u} (1-p(X_s,s)) ds$;
+% assuming recombinations are Poisson, the probability of no such recombinations is the expression given.
+% The third expression is the same, except for switching $B$ to $A$.
+% }
 % gives the rate at which a lineage switches from an $A$ to $B$ background as a result of recombination,

### 2016-04-26 new_repo/scripts/figures/blockLengthPlotsS0.001.R
added README

@@ -1,0 +1,221 @@
+
+####s=0.001:
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.001_tau1000_chr1_start0.49_stop0.51_by0.0001_flanking.blocks.by.ind.Robj")
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.001_tau1000_chr1_start0.49_stop0.51_by0.0001_flanking.blocks.by.ind.AncB.Robj")
+
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.001_tau1000_chr1_start0.49_stop0.51_by0.0001_intervalSizes_allAncs.Robj")
+load('simulation_SIGMA1_Ninds25000_ndemes50_s0.001_tau1000_chr1_start0.49_stop0.51_by0.0001_intervalSizes.Robj')
+
+
+
+Ndemes=50
+deme_ID = do.call(c,lapply(1:Ndemes,function(X){rep(X,Ninds/Ndemes)}))
+
+#positions=seq(0.495,0.505,0.0001)
+positions=seq(0.49,0.51,0.0001)
+
+flanking.blocks.deme.matrix = do.call(rbind,lapply(1:50,function(DEME){sapply(flanking.blocks.by.ind,function(POS){mean(POS[[DEME]])})}))
+pdf(height=4, width=6.25,file="s0.001_adjacentBlocksAlongChromNoConditioningHighRes.pdf")
+par(mar=c(3.5,3.5,0.5,0.5))
+matplot(t(flanking.blocks.deme.matrix/apply(flanking.blocks.deme.matrix,1,mean))[,50:1],xaxt="n",yaxt="n",xlab="",ylab="",lty=1,col=rainbow(50),type="l",x=positions,ylim=c(0.5,1.75))
+axis(1,at=seq(0.49,0.51,0.001),labels=seq(-0.01,0.01,0.001)*100,cex.axis=0.9)
+axis(2,at=seq(0.5,3,0.5),cex.axis=0.9,las=2)
+mtext("Distance from selected locus (cM)",side=1,line=2.5)
+mtext("Normalized mean block length",side=2,line=2.5)
[…]

### 2016-04-26 new_repo/scripts/figures/number_of_ancestors.R

@@ -1,0 +1,28 @@
+load("~/Downloads/number_of_ancestors_tau1000.Robj")
+
+number_of_founders_per_ind_neutral = do.call(cbind,lapply(number_of_ancestors[[1]],function(X){sapply(X,function(Z){length(unique(Z))/length(Z)})
+}))
+number_of_founders_per_ind_unlinked = do.call(cbind,lapply(number_of_ancestors[[2]],function(X){sapply(X,function(Z){length(unique(Z))/length(Z)})
+}))
+number_of_founders_per_ind_linked = do.call(cbind,lapply(number_of_ancestors[[3]],function(X){sapply(X,function(Z){length(unique(Z))/length(Z)})
+}))
+
+r2y=colorRampPalette(c("red","yellow"))
+
+pdf(file="fmaily_size_tau1000.pdf_revision.pdf",width=6,height=5)
+par(mar=c(3,3.5,0.5,0.5))
+matplot(1/number_of_founders_per_ind_neutral[,seq(20,61,2)],type="l",lty=1,col="grey",xaxt="n",yaxt="n",ylab="",xlab="",ylim=c(0,30))
+matpoints(1/number_of_founders_per_ind_unlinked[,seq(20,61,2)],type="l",lty=1,col="yellow",xaxt="n",yaxt="n",ylab="",xlab="")
+points(1/number_of_founders_per_ind_unlinked[,41],type="l",col="yellow",lwd=5)
+
+matpoints(1/number_of_founders_per_ind_linked[,seq(20,61,2)],type="l",lty=1,col=c(rev(r2y(15)[3:12]),"red",r2y(15)[3:12]),lwd=1.5,xaxt="n",yaxt="n",ylab="",xlab="")
+points(1/number_of_founders_per_ind_linked[,41],type="l",col="red",lwd=5)
+
+
+legend("bottomright",legend = c("neutral","unlinked","linked"),col=c("grey","yellow","red"),lty=1,horiz=TRUE)
+axis(1,at=c(seq(0.5,20.5,5),25.5,seq(30.5,50.5,5)),labels=c(seq(-25,-5,5),0,seq(5,25,5)),cex.axis=0.8)
+mtext("Distance from HZ center",side=1,line=2)
+axis(2,at=seq(0,30,5),cex.axis=0.8,las=2)
+mtext("mean number of descendents per haplotype",side=2,line=2.5)
+dev.off()
+

### 2016-04-26 new_repo/scripts/figures/number_of_inds.R

@@ -1,0 +1,79 @@
+start_coordinate=0.49
+end_coordinate=0.51
+
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_run2016-01-27_rawoutput.Robj")	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.ids","deme","pars","QTL")]]
+		#where:
+			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+			#"sp.ids" is a character vector of length ninds, giving the ancestry given an ancestor ID
+			#"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
+			#"pars" is an integer that gives the number of generations of the simulation
+			#"QTL" returns the QTL vector used in the simulation
+
+
+#Let's get the number of ancestors represented in chunks in each deme:
+
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
+	#return the interval containing focal site for a individual	
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancA),0)
+}
+
+
+get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
+	#get distribution of chunks within a deme
+	INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
+	intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
+	return(intervals)
+}
+
+
+
+
+#returns table (ancestor at locus, and whether it is AncB (T/F))
+ancestry.at.position = function(IND=1,DATA=sims.s0.1,POS = 0.5,CHROM="chr1",SP="B"){
[…]

### 2016-04-26 new_repo/scripts/figures/potential_statistics.R

@@ -1,0 +1,199 @@
+#load("~/Documents/Hybrid_Zones/clinal-lineages/sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau1000_simsums.Robj")
+#load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_run2016-01-27_simsums.Robj")
+load("simulation_SIGMA1_Ninds25000_ndemes50_s0.001_tau1000_run2016-02-01_simsums.Robj")
+#load("simulation_SIGMA1_Ninds25000_ndemes50_s0_tau1000_simsums.Robj")
+#load("simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau200_run2016-01-28_simsums.Robj")
+
+chromosome = "chr1"
+plot_start = 0.495
+plot_stop=0.505
+increments = 0.0001
+Ngen = 1000
+sel=0.001
+SIGMA=1
+Ninds=25000
+Ndemes = 50
+
+positions = seq(plot_start,plot_stop,increments)
+
+outstring = sprintf("simulation_SIGMA%d_Ninds%d_ndemes50_s%g_tau%d_%s_start%g_stop%g_by%g",SIGMA,Ninds,sel,Ngen,chromosome,plot_start,plot_stop,increments)
+
+deme_ID = do.call(c,lapply(1:Ndemes,function(X){rep(X,Ninds/Ndemes)}))
+
+#1. For each site, add up scores of all overlapping haplotypes
+#2. Ways to score:
+#	1. length
+#	2. score = $\log$ of ECDF of length (i.e. log of prob that haplotype is longer than this one)
+
+#lengths:
+#lengths given a particular ancestry
+get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=chromosome,POS=0.5,ancB=TRUE){
+	chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})
+	identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
+	replace(chunk,which(identity==ancB),0)
+	
+}
+
+intervalSizes = lapply(positions,function(POS){
+				do.call(rbind,lapply(sims.sums[[1]]$ind.ancest,function(IND){
+				get.interval.size(IND_DATA=IND,CHR=chromosome,POS=POS,ancB=TRUE)	
+				}))
+				})
+				
[…]

### 2016-04-26 new_repo/scripts/notes/cline-fns.R

@@ -1,0 +1,464 @@
+source("unifprod.R")  # for the C part
+
+pcline <- function (x,r,p=recombfn(x,2*r,...),recombfn,...) {
+    (1/2)*( 1 - sign(x)*p )
+}
+
+tanh_cline <- function (x,s,sigma=1,log=FALSE,...) {
+    if (log) {
+        - (4*x*sqrt(s)/sigma) - log1p(exp(-4*x*sqrt(s)/sigma))
+    } else {
+        (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
+    }
+}
+
+tanh_drift <- function (x,s,sigma=1) {
+    # log-derivative of tanh_cline
+    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
+}
+
+tanh_het <- function (x,s,sigma=1) {
+    # prob of being in a het = (1-s)(1-p)/(1-s(1-p))
+    pp <- tanh_cline(x,s,sigma=sigma)
+    (1-s)*(1-pp)/(1-s*(1-pp))
+}
+
+zeta.reflected <- function (x,r) {
+    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
+}
+
+sim_diffusion <- function (drift,tau,nwalks,ntimes,xinit=0,...) {
+    # want total time to be tau, so:
+    dt <- tau/ntimes
+    # put the noise in rws
+    rws <- matrix( c( rep_len(xinit,nwalks), # initial conditions
+                 rnorm(nwalks*(ntimes-1)) ), nrow=ntimes, ncol=nwalks, byrow=TRUE )
+    # and then integrate over it
+    for (k in 2:ntimes) {
+        rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],...)
+    }
+    attr(rws,"dt") <- dt
+    return(rws)
+}
+
+weighted_time <- function (rws, weight.fn, use.t=TRUE, ...) {
+    colSums( weight.fn(rws[use.t,],...) ) * attr(rws,"dt")
+}
+
+clineplot <- function (soln, thegrid, x=thegrid$x.mid,
+               times=seq(1,nrow(soln),length.out=ntimes), ntimes=20,
[…]

### 2016-04-26 new_repo/scripts/notes/tran.1D.R

@@ -1,0 +1,301 @@
+##
+#
+# Trivially modified from ReacTran::tran.1D
+# to allow passing in A in log scale.
+#
+##
+
+tran.1D <- function (C, C.up = C[1], C.down = C[length(C)], flux.up = NULL, 
+    flux.down = NULL, a.bl.up = NULL, a.bl.down = NULL, D = 0, 
+    v = 0, AFDW = 1, VF = 1, A = 1, dx, 
+    log.A = FALSE,   # TRUE if A is given in log units
+    full.check = FALSE, full.output = FALSE) 
+{
+    if (is.null(dx)) 
+        stop("error: dx should be inputted ")
+    N <- length(C)
+    if (N == 0) 
+        stop("C should be a vector with numeric values")
+    if (is.null(C.up)) 
+        C.up <- C[1]
+    if (is.null(C.down)) 
+        C.down <- C[N]
+    if (!is.list(AFDW)) 
+        AFDW <- list(int = rep_len(AFDW,N+1))
+    if (!is.list(D)) 
+        D <- list(int = rep_len(D,N+1))
+    if (!is.list(v)) 
+        v <- list(int = rep_len(v,N+1))
+    if (!is.list(VF)) 
+        VF <- list(int = rep(VF, length.out = (N + 1)), mid = 0.5 * 
+            (rep(VF, length.out = (N + 1))[1:N] + rep(VF, length.out = (N + 
+                1))[2:(N + 1)]))
+    if (!is.list(A)) 
+        A <- list(int = rep(A, length.out = (N + 1)), 
+                  mid = 0.5 * (rep(A, length.out = (N + 1))[1:N] + rep(A, length.out = (N + 1))[2:(N + 1)]))
+    if (is.list(dx)) 
+        grid <- dx
+    if (!is.list(dx)) 
+        grid <- list(dx = rep(dx, length.out = N), dx.aux = 0.5 * 
+            (c(0, rep(dx, length.out = N)) + c(rep(dx, length.out = N), 
+                0)))
+    if (full.check) {
+        gn <- names(AFDW)
+        if (!"int" %in% gn) 
+            stop("error: AFDW should be a list that contains 'int', the AFDW values at the interface of the grid cell ")
+        if (is.null(AFDW$int)) 
+            stop("error: AFDW is NULL, should contain (numeric) values")
+        if (!is.null(AFDW$int)) {
+            if (!((length(AFDW$int) == 1) || (length(AFDW$int) == 
[…]

### 2016-04-26 new_repo/scripts/sims/chunks_fns.R

@@ -1,0 +1,117 @@
+##
+# Functions for extracting information from a _simsums_chunks object.
+##
+
+genotype_counts <- function (ch) { 
+    gcounts <- t(sapply( ch, function (x) { tabulate( 1+rowSums( x>0 ), nbins=3 ) } )) 
+    colnames(gcounts) <- c("AA","AB","BB")
+    return(gcounts)
+}
+
+conditional_genotype_counts <- function (x,y) { 
+    # returns joint genotype counts across space from two chunks objects
+    #  a (ndemes) x 4 x 4 array (4 phased genotypes)
+    deme <- rep(seq_along(x),sapply(x,NROW))
+    x.alleles <- ifelse(do.call(rbind,x)==0,"A","B")
+    y.alleles <- ifelse(do.call(rbind,y)==0,"A","B")
+    x.genotypes <- factor( paste(x.alleles[,1],x.alleles[,2],sep=""), levels=c("AA","AB","BA","BB") )
+    y.genotypes <- factor( paste(y.alleles[,1],y.alleles[,2],sep=""), levels=c("AA","AB","BA","BB") )
+    return( as.array( table( deme, x=x.genotypes, y=y.genotypes ) ) )
+}
+
+dip.to.hap <- rbind( A=c(AA=2,AB=1,BB=0), B=c(AA=0,AB=1,BB=2) )
+
+bidip.to.bihap <- rbind(
+            "A-A" = c( 2, 1, 1, 0, 1, 1, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0 ),
+            "B-A" = c( 0, 1, 1, 2, 0, 0, 1, 1, 0, 1, 0, 1, 0, 0, 0, 0 ),
+            "A-B" = c( 0, 0, 0, 0, 1, 0, 1, 0, 1, 1, 0, 0, 2, 1, 1, 0 ),
+            "B-B" = c( 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 1, 1, 0, 1, 1, 2 ) )
+colnames( bidip.to.bihap ) <- as.vector( outer( c("AA","AB","BA","BB"), c("AA","AB","BA","BB"), paste, sep="-" ) )
+
+conditional_haplotype_counts <- function (gcounts) {
+    # uses output of conditional_genotype_counts to get conditional haplotype counts
+    # a (ndemes) x 2 x 2 array
[…]

### 2016-04-26 new_repo/scripts/sims/run_sims.R

@@ -1,0 +1,207 @@
+#!/usr/bin/Rscript
+
+source("sim-fns.R")
+
+# Parameters for simulation:
+
+params = list(
+    SIGMA = 3,        # sigma for dispersal, in deme spacings
+    ndemes = 100,     # number of demes
+    deme_size = 50,   # number of diploid individuals per deme
+    S = 0.05,         # selection coefficient, s
+    zone_age = 10     # number of generations to run simulation for
+)
+params$ninds = params$ndemes * params$deme_size
+
+run.id = sprintf("%0.0f",1e6*runif(1))  # unique run id
+outfile = with( list2env(params), sprintf("simulation_SIGMA%s_Ninds%s_ndemes%s_s%s_dir/tau%s_dir/results_runid_%s",SIGMA,ninds,ndemes,S,zone_age,run.id) )
+dir.create(dirname(outfile),recursive=TRUE)
+
+#PARAMS for parsing:
+loci = list(seq(0.5,1,0.05),seq(0.5,1,0.05))
+xx = (1:params$ndemes)-0.5-params$ndemes/2
+
+
+#DEFINE selection against genotypes:
+qtl=list(
+        chr1=data.frame(
+                traits=c("underdominant"), 
+                s=params$S,
+                pos=c(0.5)),
+        chr2=data.frame(
+                traits=c("underdominant"), 
+                s = 0,pos=c(0.5))
+        )
+
+# save out parameters
+save( params, outfile, loci, xx, qtl, file=paste0(outfile,"_params.Robj") )
+
+#SIMULATE the populations
+sims.full = lapply(params$zone_age, sim.zone, n.ind=params$ninds, n.deme=params$ndemes, sigma=params$SIGMA)
+	#output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
+		#where:
+			#"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
+			#"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
[…]

### 2016-04-26 new_repo/scripts/sims/sim-fns.R

@@ -1,0 +1,308 @@
+
+getRuns = function(index){
+    #This returns an array in which rows are the the beginnings and end of a TRUE/FALSE vector
+    # (note: more efficient to use rle() probably)
+    starts  = which(c(0,index)[-1]==1 & c(0,index)[-length(index)]!=1)
+    stops   = which(c(index,0)[-1]!=1 & c(index,0)[-(1+length(index))]==1)
+    if(index[length(index)] == 1 & index[(length(index)-1)] !=1){starts = c(starts,length(index))}
+    return (cbind(starts,stops))
+}
+
+
+namesList = function(these.names){
+    # this is a way to get a list where each elelment has a name equalto its value, so that when I lapply, I get back a [correctly] named list
+    all.things = as.list(these.names)
+    names(all.things)=these.names
+    return(all.things)
+}
+
+#get breaks between species' ancestry
+spBreaks = function(sim){
+    ind.ancest = lapply( sim$inds,function(IND){ 
+            lapply(IND,function(CHR){
+               lapply(CHR,function(PAR){
+                    if(nrow(PAR)==1){return(data.frame(starts=0,stops=1,sp2 = PAR$ancest > length(sim[[1]])))}
+                    sp1 = getRuns(as.numeric( PAR$ancest <= length(sim[[1]])))
+                    sp2 = getRuns(as.numeric( PAR$ancest > length(sim[[1]])))
+                    tmp = rbind(data.frame(sp1,sp2=rep(FALSE,nrow(sp1))),data.frame(sp2,sp2=rep(TRUE,nrow(sp2))))
+                    tmp = tmp[order(tmp$starts),]
+                    tmp$starts = PAR$pos[tmp$starts]
+                    tmp$stops = c(PAR$pos[-1],1)[tmp$stops]
+                    return(tmp)
+                })
+            })
+        })
+    ancest.prop = t(sapply(ind.ancest,function(IND){
+        tmp = t(sapply(IND,function(CHR){ c(X1=with(CHR$X1,sum(sp2*(stops-starts))), X2 = with(CHR$X2,sum(sp2*(stops-starts))))  }))
+        return(colSums(tmp)/nrow(tmp))
+    }))		
+    return(list(ind.ancest = ind.ancest, ancest.prop = ancest.prop))
+}
+
+
[…]

### 2016-04-26 drafts/hybrid_zone.tex
modified abstract

@@
 across genomic and geographic space. 
-By understanding how lineages move across the hybrid zone, in a model without coalescence, 
-we describe how clines form and how parental haplotypes are broken up during introgression.
-Clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
-and linked clines persist over a  genomic scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. \alisa{This is in May1975? Should it be removed?}
+By describing how lineages move across the hybrid zone, in a model without coalescence, 
+we add to modern understanding of how clines form and how parental haplotypes are broken up during introgression.
+Working with lineages makes it easy to see that
+clines form in about $1/s$ generations, where $s$ is the strength of selection against hybrids,
+and linked clines persist over a genomic scale of $1/T$, where $T$ is the age, in generations, of the hybrid zone. 
 Locally disadvantageous alleles tend to exist as small families,

### 2016-04-26 drafts/review-responses-april-2016.tex
yb had a quick go at intro to response

@@
 
-We did all the things, thanks.
-
-\plr{these aren't points because they are all points below in Reviewer 1 \& 2's sections; we'll just say here that we did 'em, see below}
+\yb{We are pleased to submit a revision of our Manuscript, `Beyond clines: lineages and haplotype blocks in hybrid zones' (MEC-16-0266) for the special issue 'Genomics of Hybridization'.
+We believe that this version is much improved by changes we have made to address reviewer and AE comments.
+Below we provide a point-by-point response to reviewer and AE comments, describing how we addressed concerns raised by our readers.
+The major themes of our changes are (\emph{1}) Attempting to make our mathematical derivations more accessible for Molecular Ecology’s wide audience  (\emph{2}) Additional justification for using a genetic model of under dominance to model hybrid incompatibilities, in which we make heavy reference to the historical theory justifying this approach, and (\emph{3}) Additional simulations to show that our results are not driven by boundaries imposed in a defined range, and (\emph{4})  A streamlined introduction that makes clear our major goals and provides a stronger structure for our manuscript.
+}
 

### 2016-04-26 drafts/hybrid_zone.tex
peter edited intro

@@
 The process of speciation commonly involves populations diverging with relatively little gene flow \citep{Coyne2004}.  
-% Gene flow between formerly isolated populations is possible if reproductive isolation is not complete.
 However, when formerly isolated populations come into contact before reproductive isolation is complete, some gene flow is possible.
-%As two such populations interbreed , each portion of each individual's genome 
-%can be labeled according to which ancestral population is is inherited from, and local 
 Interbreeding and migration between such populations  
-% YB It seems like the fact that such ancestry can be labeled is more important later, but probably shouldn't be a focus here. 
-creates a gradient of alleles derived from the two populations across geographic space, %YB removed `assignments' as it seems like it is the ancestry, not its assignment, that is of interest
+creates a gradient of alleles derived from the two populations across geographic space,
 centered on their point of contact \citep[reviewed by][]{Barton1985}. 
@@
 forming chromosomal {junctions} in ancestry \citep{Fisher1954, Chapman2002, baird2003distribution}. 
-% Limited opportunities for recombination in recent hybrid zones generates substantial correlations in ancestry (and hence, genotype) between distant markers, and
-These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture \citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}.
-
-%The mixing of genomes across space, combined with selection, results in the formation of genetic clines. 
-%These patterns have been the traditional focus of both empirical and theoretical studies into hybrid zones. 
-%While here we focus on hybrid zones maintained by selection against hybrids 
-%-- \emph{tension zones}, as reviewed in \citet{Barton1985} --  % removed \citet{Bazykin1969}, it's cited later
-%clines may also be due to differential selection across environmental gradients \citep{Haldane1948,fisher1950frequencies,hanson1966effects},
-%% generating \emph{ecotones}.  % I think ecotone means something somewhat different: see Barton & Hewitt 1985, p. 132, for instance
-%Although these are in principle different,  \linelabel{ll:all_clines_are_the_same}
-%theoretical and simulation studies show that
-%genomic clines seen in the two situations are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999},
-%with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
-% consistently show ecotones and tension zones have very similar properties \citep{Slatkin1973, barton1986barrier,Kruuk1999}, 
-% \plr{(I can't find discussion of similarity of different types of clines in \citet{barton1986barrier}: put the citation back in if I missed it.)}
-%% don't think we need this next bit
-% The primary difference being that the center of a tension zone fluctuates over time \citep{barton1979dynamics} 
-% because stochastic changes in allele frequency can shit the cline center, 
-% while centers of ecotones do not change unless the environmental gradient changes.
-
-Our work here builds on the large body of theory describing tension zones \alisa{and aims to describe \emph{transient} dynamics with a focus on \emph{haplotype} patterns.}
-%While much of this previous work describes equilibrium states of genetic clines, 
-%we focus on \emph{transients}, in particular describing haplotype patterns. 
-%Although the cline in selected allele frequency establishes quickly, transient patterns can persist for quite some time, 
-%and haplotypes can potentially carry substantial information about the formation of recent zones.
-To do this, we take the reverse-time perspective,
-understanding the temporal dynamics of the system by describing properties of the lineages tracing back from the zone.
-% This is particularly suited to understanding transient patterns.
-
-The arrangement of segments of genome inherited from particular ancestral populations%, which may be identifiable as distinct haplotypes,
-is shaped by the number of recombination events that have taken place between ancestral genotypes. 
-The length distribution of such segments therefore contains information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture 
+\citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}.
+
 %The distribution of ancestry tract length in admixed or hybridizing populations has been used to infer the extent of hybridization and timing of secondary contact both in single populations and in hybrid zones \citep{Price2009, Gravel2012,sedghifar2015spatial}.
@@
 
-\yb{We show here that heterogeneous patterns of the ancestry length distribution across the genome may help identify putative targets of selection in hybridizing populations, providing another source of information in addition to gradients in allele frequencies  used in current approaches \citep[e.g.,][]{Porter1997, Gompert2012}.
+Our work here builds on the large body of theory describing tension zones 
+and aims to describe \emph{transient} dynamics with a focus on \emph{haplotype} patterns.
+% which may carry substantial information about the formation of recent zones.
+To do this, we take the reverse-time perspective,
+understanding the temporal dynamics of the system by describing how lineages tracing back from the hybrid zone
+moveme across space, conditioning on the frequencies of the selected allele and following them as they recombine between selected backgrounds
+\citep[similar to][]{hudson1988coalescent}.
+Using a combination of theory and simulated hybrid zones, 
+and extending the framework of \citet{sedghifar2015spatial},
+we describe genome-wide patterns of coancestry 
+as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.
+
+% In other words, the arrangement of ancestral haplotype segments
+% is shaped by the number of recombination events that have taken place between ancestral genotypes,
+% and so the length distribution of such segments contains information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. 
+In particular, we examine how heterogeneous patterns of the ancestry length distribution across the genome 
+may help identify putative targets of selection in hybridizing populations, 
+providing another source of information in addition to gradients in allele frequencies used in current approaches \citep[e.g.,][]{Porter1997, Gompert2012}. 
 Specifically, we find that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is, 
-because selection rapidly removes incompatible alleles from heterospecific genomes, hybrid incompatibilities will be surrounded by disproportionately long ancestry blocks.}
-
-Using a combination of theory and simulated hybrid zones, we describe genome-wide patterns of coancestry 
-as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. 
-\alisa{We extend the framework presented by \cite{sedghifar2015spatial} to describe how lineages move across a zone of secondary contact.
-We consider the influence of selection against heterozygotes at a single locus on the movement of lineages across space, conditioning on the frequencies of the selected allele and allowing lineages to recombine between selected background } (analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
-%In \citet{sedghifar2015spatial} we described how lineages move across a neutral zone of secondary contact,
-%obtaining predictions for distributions of lengths of blocks of ancestry, which we fit to several datasets.
-%In this work, we extend the same framework to the case of selection against heterozygotes at a single locus,
-%by describing how lineages move conditional on the frequencies of the selected allele
-%and allowing lineages to recombine between selected background 
-%(analogous to the structured coalescent applied to selection, as in \citet{hudson1988coalescent}).
+because selection rapidly removes incompatible alleles from heterospecific genomes, hybrid incompatibilities will be surrounded by disproportionately long ancestry blocks.
 
@@
 As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus is expected to be relatively long when far away from the zone center.
-This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, and therefore will not have
-as many ancestors of the locally common type as do neutral haplotypes.
-\alisa{This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotone models of extrinsic ecological selection (e.g., local adaptation).}
-%\plr{(ecological what?)}
-
-\yb{Because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous},
-incorporating genetic drift is not straightforward, and we therefore neglect genetic drift in our analytical results.} \revpoint{2}{9}
-This approach is common to nearly all theoretical work,
-that study the reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
+This is because an unfit haplotype is more likely to have been recently inherited from the other side of the hybrid zone, 
+and therefore will not have as many ancestors of the locally common type as do neutral haplotypes. 
+This reasoning is applicable to instances of both tension zones generated by intrinsic genetic incompatibilities, and ecotone models of extrinsic ecological selection (e.g., local adaptation).
+
+We neglect genetic drift in our analytical results \revpoint{2}{9}
+because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous}.
+This should provide a good model for stochastic motion of a single lineage (our main focus),
+but does not model correlations between lineages or non-Markovian dynamics of the lineages.
+This approach is common to much theoretical work,
+that studies the reaction-diffusion equations that govern the deterministic, high-density limit \citep[as in][]{Nagylaki1975},
 or the discrete analogues \citep{hanson1966effects}. 
-There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Nagylaki1978,durrett2007width,barton2008effect,Polechova2011},
-and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,Barton1989,Barton1993,Barton2000}.
-\plr{(why are references to 'barrier to gene flow' in the paragraph on drift? perhaps move to discussion?)}
-While we do not attempt to account for finite population size in our analytic results,
-our simulation results (below) suggest that although genetic drift strongly affects local noisiness of the system,
-it does not substantially influence our conclusions at moderate population densities
-(hundreds within one dispersal distance),
-perhaps because of the short time-scales that we examine.
-\plr{(maybe remove this next sentence? but want to say something about the very low densities we don't examine.)}
-At much lower population densities, local extinction and recolonization are expected to have a stronger effect.
-
-\plr{(nice wrap-up sentence shouldn't be at the end of paragraph on drift:)} \alisa{Is it weird to end intro on above paragraph? Could this wrap-up sentence hold up on its own? }
-\alisa{Our results look beyond the traditional signature of selection observed in the width of genetic clines, and present a possible new approach for identifying genomic regions experiencing selection in hybrid zones.}
+There is also substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Nagylaki1978,durrett2007width,barton2008effect,Polechova2011}.
+Our simulation results (below) suggest that although genetic drift strongly affects local noisiness of the system,
+it does not substantially influence our conclusions at the moderate population densities we examine,
+perhaps because we study relatively short time-scales.
+At much lower population densities, however, local extinction and recolonization may have a stronger effect. \linelabel{ll:much_lower_densities}
+
+There is also substantial work
+quantifying the barrier to gene flow caused by clines \citep{barton1979gene,barton1986effects,Barton2000}.
+We aim to complement this work,
+and to investigate a possible new approach for identifying genomic regions experiencing selection in hybrid zones.
 
@@
 and analytics, mostly in terms of partial differential equations needing numerical solution.
-\plr{As noted above, we ignore the effects of drift and the pedigree,
-so that the underlying allele frequency dynamics are deterministic.
-This should provide a good model for stochastic motion of a single lineage (our main focus),
-but does not model correlations between lineages.}
 
@@
 follow its lineage back through time, using $\tau$ to denote ``generations ago''
-(reserving $t$ for time measured in the usual direction).
+(reserving $t$ for time measured in the forward direction).
 If $X_\tau$ is the geographic location of its ancestor $\tau$ generations ago,
@@
 We then assigned ancestry at individual loci by looking up which side of the zone the ancestor lived on.
-% removed "ancestor's genotype" because it's not really genotype
+
+For efficiency, in most simulations the number of demes was only 3--5 times wider than the stable cline width. \linelabel{ll:narrow_sims}
+We verified that the size of the simulated region was not affecting dynamics in the center of the zone
+by running additional simulations on wider regions
+(as we did for numerical solutions of PDE).
 
@@
 and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination \citep{barton1986effects}. 
-Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection \citep{Slatkin1975b,Barton1983}.
+Both factors are expected to result in regions that are surrounded by even longer segments of unbroken ancestry, 
+and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection \citep{Slatkin1975b,Barton1983}.
 
@@
 \url{https://github.com/petrelharp/clinal-lineages}
-under the GPLv3 software license,
-and on Data Dryad \url{http://datadryad.org/XXXX}.
+under the GPLv3 software license
 

### 2016-04-26 drafts/supplement.tex
peter edited intro

@@
 \begin{figure}
-    \includegraphics{figs/alleleFrequencies_sim.pdf}
+    \includegraphics{{figs/alleleFrequencies_sim_s0.01_tau100_closely_linked}.pdf}
     \caption{
@@
 \includegraphics{figs/blocksAlongChromHeatmapAncBConditioning.pdf}
-\caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome. }\label{Supp:blockLengthHeatmap}
+\caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome under $T=1000$, $s=1$ and $\sigma=1$,
+    50 demes, each containing 500 diploid individuals. }\label{Supp:blockLengthHeatmap}
 \end{figure}

### 2016-04-26 drafts/figs/Plot_frequency_clines.R

@@ -1,10 +1,10 @@
 
-source("../sims/sim-fns.R",chdir=TRUE)
+source("../../sims/sim-fns.R",chdir=TRUE)
 library(ReacTran)
 library(Matrix)
-source("../notes/tran.1D.R",chdir=TRUE) # fix for log.A
-source("../notes/cline-fns.R",chdir=TRUE)
-source("../sims/chunks_fns.R",chdir=TRUE)
+source("../../notes/tran.1D.R",chdir=TRUE) # fix for log.A
+source("../../notes/cline-fns.R",chdir=TRUE)
+source("../../sims/chunks_fns.R",chdir=TRUE)
 
 
 ###
@@ -16,16 +16,19 @@ get_simparams <- function (fname) {
 }
 
 
-positions = seq(0.2,0.8,0.01)
-positions = seq(0.48,0.52,0.001)
+# positions = seq(0.2,0.8,0.01)
+# positions = seq(0.48,0.52,0.001)
 positions = c(0,0.01,0.02,0.05,0.1,0.2,0.3)
 #cols = c(2,3,6,11,21,31)
 
-sel.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
+sel.sim.file <- "../../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
 sel.params <- get_simparams( sel.sim.file )
 
-neutral.sim.file <- "../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"
+neutral.sim.file <- "../../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0_tau100_run2016-01-15_simsums.Robj"
 neu.params <- get_simparams( neutral.sim.file )
+
+long.sim.file <- "../../../sims/simulation_SIGMA3_Ninds250000_ndemes500_s0.01_dir/tau100_dir/results_runid_552675_simsums.Robj"
+long.params <- get_simparams(long.sim.file)
 
 if (!file.exists("plot_frequency_clines_revision.Robj")) {
     load(sel.sim.file)
@@ -53,7 +56,26 @@ if (!file.exists("plot_frequency_clines_revision.Robj")) {
             CHR=1, 
             ndemes=neu.params$ndemes)
 
[…]

### 2016-04-26 drafts/hybrid_zone.tex
minor
fixed fig 3 caption, more edits

--- 2016-04-26 drafts/hybrid_zone.tex
@@
 These patterns of correlated ancestry in admixed populations have been previously used to infer histories of hybridization and admixture 
-\citep[e.g.][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}.
+\citep[e.g.,][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}.
 
@@
 previous work has shown that such models generate clines that are very similar to models of ecological selection or epistatic systems \citep{Kruuk1999,barton2000epistasis}, 
-\yb{because selection on an allele depends on its marginal effect on fitness.  }
-Exceptionally strong epistatic selection which acts more quickly than free recombination may therefore, yield results that differ slightly from ours. 
-Likewise, while we do not model ecological selection, theoretical and simulation studies show that clines formed by underdominant selection and extrinsic selection are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999}, especially over short time scales \citep{Barton1979a}, \linelabel{ll:all_clines_are_the_same}
-with their shape determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}.
+because selection on an allele depends on its marginal effect on fitness.
+(More discussion of this point in the Discussion.) \linelabel{ll:all_clines_are_the_same}
 
@@
     Branching events occur when there was a recombination event within the sampled segment.
-    The \alisa{rightmost} blue line represents the path of the selected locus. 
-   % \plr{(All lines look the same width: no thick and thin.)}
+    The rightmost blue line represents the path of the selected locus. 
     The colors of the other paths indicate the identity at the linked selected locus 
@@
 %% we don't need to cite these here, since we cite Bazykin for the equation
-% \alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.][]{Slatkin1973,may1975gene,Nagylaki1975}}
+% \alisa{Equations describing the forward-time establishment of the cline have been previously described \cite[e.g.,][]{Slatkin1973,may1975gene,Nagylaki1975}}
 Write $p(x,t)$ for the frequency of the $A$ allele at spatial location $x$ and time $t$.  
@@
 \begin{align} \label{eqn:cline_pde}
-    \frac{d}{dt} p = \frac{\sigma^2}{2} \frac{d^2}{dx^2} p +  2 p (1-p)  \times s (\frac{p}{2} - 1), %YB Rearranged this equation to allow for cleaner description
+    \frac{d}{dt} p = \frac{\sigma^2}{2} \frac{d^2}{dx^2} p +  2 p (1-p)  \times s (\frac{p}{2} - 1).
     %    \frac{d}{dt} p = \frac{\sigma^2}{2} \frac{d}{dx^2} p + s p (1-p) (2p-1) ,
 \end{align}
-where $ \frac{d}{dt} p$ is the local rate of change of the frequency of $A$.
+Here the frequency $p=p(x,t)$ varies across time and space,
+and so $ \frac{d}{dt} p$ is the local rate of change of the frequency of $A$.
 The first term, $\frac{\sigma^2}{2} \frac{d^2}{dx^2} p$, describes the net rate at which $A$ alleles arrive by local dispersal.
 The second term % ,  $2 p (1-p)  s (p/2-1)$,  % only need this term in the sentence once?
-describes the impact of selection on local allele frequency as the product of local genotypic variance, 
-$2 p (1-p)$, and selection on allele $A$, $s (p/2-1)$. 
-\plr{ I don't think 'genotypic variance' is the right interpretation?  Let me try again: 
-The second term is the product of the local frequency of heterozygotes, $2p(1-p)$,
-and the strength of selection on allele $A$, $s (p/2-1)$.  }
+describes the impact of selection on local allele frequency as the product of local genotypic variance, $2p(1-p)$
+(equivalently, the local frequency of heterozygotes), 
+and selection on allele $A$, $s (p/2-1)$. 
 The term $s (p/2-1)$ describes underdominance -- $A$ is favored when common ($p>1/2$) and disfavored when rare ($p<1/2$).
@@
 
-As noted by others \citep[e.g.][]{Slatkin1973,may1975gene}, \linelabel{ll:rescaling}
+As noted by others \citep[e.g.,][]{Slatkin1973,may1975gene}, \linelabel{ll:rescaling}
 rescaling space and time by $\sigma/\sqrt{s}$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} 
@@
 the description above applies to motion of a lineage transverse to the hybrid zone.  \revpoint{AE}{4}
-In this case, $d^2/dx^2$ should be replaced by the Laplacian operator,
-and $d/dx$ by the gradient.
+In this case, $d^2/dx^2$ should be replaced by the Laplacian operator.
+% and $d/dx$ by the gradient.
 In fact, if the landscape is heterogeneous or dispersal is non-Gaussian,
@@
 that the lineage of an $A$ allele at position $x$ moves as Brownian motion driven by dispersal
-with mean displacement proportional to the derivative of $\log p$.
+with mean displacement proportional to the spatial derivative of $\log p$.
 This has the effect of pulling $A$ lineages towards regions where they are more common, i.e., the left side of the range.   
@@
 % with bias $\sigma^2 \grad \log p(x,t)$, where $\grad$ is the gradient
-(the analogous physical system is a particle that moves at speed $\sigma^2$ in the potential $-\log p$). 
+(The analogous physical system is a particle that moves at speed $\sigma^2$ in the potential $-\log p$). 
 %\alisa{What does potential here mean?}
@@
 reflecting strong differences in selection against heterozygotes,
-even though the stable clines have the same form, and therefore have additional information about the hybrid zone.
+even though the stable clines have the same form.
 Even though a rescaling of space and time as described above can make the models equivalent,
@@
 
-\yb{Suppose we sample an allele today $r$ Morgans from the selected site, and  
+Suppose we sample an allele today $r$ Morgans from the selected site, and  
 follow its lineage back through time, using $\tau$ to denote ``generations ago''
@@
 then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$ 
-(as described for a selected allele in equation \eqref{eqn:cline_pde}).
+(as described for a selected allele in the previous section).
 Following this lineage back in time, the identity of the selected allele that ancestor carried at time $\tau$, $Z_\tau$, 
-jumps between $A$ and $B$ with recombination events between the neutral and selected locus (with frequency $r$), that result in a change in the  selected background.    
+jumps between $A$ and $B$ with each recombination event that occurs between the focal locus and the selected locus (with frequency $r$), 
+that results in a change in the  selected background.    
 Thus, by the assumption of locally random assortment of alleles, $Z$ shifts from $A$ to $B$  at rate $r (1-p)$, and $B$ to $A$ at rate $r p$ 
-(see Appendix \ref{apx:lineage_derivation} for a more precise description).}
+(see Appendix \ref{apx:lineage_derivation} for a more precise description).
 
@@
 with $B_\tau$ a standard Brownian motion,
-$T_B(\tau)$ the most recent time before $\tau$ that $Z_\tau = B$, and likewise for $T_A$,
+$T_B(\tau)$ the most recent time before $\tau$ that $Z_\tau = B$, and likewise for $T_A(\tau)$,
 \begin{align}
@@
         \end{cases} \\
-        \P\{ Z_{\tau+\epsilon} = B \given Z_\tau = A, \; X_\tau=x \} &= \epsilon r (1-p(x,\tau)) + O(\epsilon^2) \\
-        \P\{ Z_{\tau+\epsilon} = A \given Z_\tau = B, \; X_\tau=x \} &= \epsilon r p(x,\tau) + O(\epsilon^2)  .
+        \P\{ Z_{\tau+\epsilon} = B &\given Z_\tau = A, \; X_\tau=x \} = \epsilon r (1-p(x,\tau)) + O(\epsilon^2) \\
+        \P\{ Z_{\tau+\epsilon} = A &\given Z_\tau = B, \; X_\tau=x \} = \epsilon r p(x,\tau) + O(\epsilon^2)  .
         % \P\{ T_B(\tau) &> \tau+u \given Z_\tau = A \} = \exp\left( - r \int_\tau^{\tau+u} (1-p(X_s,s)) ds \right) \\
@@
 (In two dimensions, $d/dx$ is replaced by the gradient.)
-\plr{
 The second expression says that the probability a lineage on the $A$ background at $x$
@@
 per generation.
-}
 % \plr{
@@
 (i.e., the net flux induced by reduced fitness in the hybrid zone),
-and recombination between selected backgrounds, respectively. \alisa{I think the first and second terms may have been switched??}
-\yb{We note that because recombination occurs within a deme, the recombination terms in equation \eqref{eqn:q_pde} are quite standard \citep{HartlClark}.} \revpoint{1}{4}
+and recombination between selected backgrounds, respectively.
+We note that because recombination occurs within a deme, the recombination terms in equation \eqref{eqn:q_pde} are quite standard \citep{HartlClark}. \revpoint{1}{4}
 %\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
@@
 so define these as
-$G^A = \sigma^2 \frac{d}{dx} \log(p(x,t)) \cdot \frac{d}{dx} + (\sigma^2/2) \frac{d}{dx^2}$
+$
+G^A = 
+(\sigma^2/2) \frac{d}{dx^2}
++ 
+\sigma^2 \frac{d}{dx} \log(p(x,t)) \cdot \frac{d}{dx} 
+$
 and
-$G^B = \sigma^2 \frac{d}{dx} \log(1-p(x,t)) \cdot \frac{d}{dx} + (\sigma^2/2) \frac{d}{dx^2}$,
+$
+G^B = 
+(\sigma^2/2) \frac{d}{dx^2}
++ 
+\sigma^2 \frac{d}{dx} \log(1-p(x,t)) \cdot \frac{d}{dx} 
+$
+,
 so that equation \eqref{eqn:q_pde} can be written more compactly as
@@
 Additionally, at rate equal to the genetic length of the segment,
-recombination occurs, at which point the segment splits in two at a uniformly chosen location between $a$ and $b$,
+recombination occurs within the segment, at which point it splits in two at a uniformly chosen location between $a$ and $b$,
 each of which proceeds as before, independently. 
@@
 and the second term results from recombinations \emph{between} the block and the selected site 
-in heterozygotes, which switch the identity of the linked, selected locus \alisa{without splitting the block}
+in heterozygotes, which switch the identity of the linked, selected locus without splitting the block
 (the factor $(1-p)$ is the probability that the block, initially linked to an $A$ allele, encounters a $B$ allele
 under the assumption of locally random mating).
-In the integrals, $\theta$ is the genomic position of the recombination event occuring \emph{inside} the block.
+In the integrals, $\theta$ is the genomic position of the recombination event.
 The third term accounts for such recombinations in a heterozygote: 
 the portion of the block nearest the selected site remains linked to an $A$ allele, 
-and the remaining portion becomes linked to a $B$ allele (as is the case in the most recent recombination event shown in Fig.~\ref{Fig:schematic}).
+and the remaining portion becomes linked to a $B$ allele 
+(e.g., the lower recombination event shown in Fig.~\ref{Fig:schematic}).
 To account for cases where the selected locus lies outside the interval [a,b],
@@
 and likewise $\int_{\max(0,a)}^b$ is zero if $b<0$.
-The last integral results from recombination inside the block in homozygotes for $A$ (as is the case for the second recombination event in Fig.~\ref{Fig:schematic}),
+The last integral results from recombination inside the block in homozygotes for $A$ 
+(e.g., the upper recombination event in Fig.~\ref{Fig:schematic}),
 and the final term balances the outflux due to all recombinations.
@@
 the equation for haplotype identity probabilities on a segment $(a,b)$ depends only on those probabilities for segments contained in $(a,b)$.
-This allows for numerical solutions, %`efficient` may be an overstatement
+This is useful for numerical solutions, %`efficient` may be an overstatement
 described in more detail in Appendix \ref{apx:haplotype_calcs}.
@@
 % so this equivalently takes a centered Gaussian displacement and assigns it to the nearest deme. 
-(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
+(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this discrete distribution.)
 Migrants past either end of the range remain at the terminal demes.
@@
 For efficiency, in most simulations the number of demes was only 3--5 times wider than the stable cline width. \linelabel{ll:narrow_sims}
-We verified that the size of the simulated region was not affecting dynamics in the center of the zone
+We verified that the size of the simulated region did not affect dynamics in the center of the zone
 by running additional simulations on wider regions
-(as we did for numerical solutions of PDE).
+(as we did for numerical solutions of PDE);
+an example is shown in Figure \ref{sfig:extra_long_cline}.
 
@@
         \textbf{Comparison of simulated to theoretical clines:}
-        As in Figure \ref{alleleFreq_tau100},
-        frequency of ancestry $B$,
-        across geography and at several physical positions on the genome,
-        at $T=100$ generations after secondary contact,
-        and at $s=0.1$.
+        Frequency of ancestry $B$,
+        across geography at different physical positions on the genome, simulated for a hybrid zone
+        $T=100$ generations after secondary contact, with $s=0.1$,
+        using 50 demes, each with 500 diploid individuals and $\sigma=1$.
+        In each figure,
+        each line represents a locus some distance $r$ away from the true target of selection, 
+        and $r=0$ represents the locus that is under selection. 
         Dotted lines are expected frequencies of $B$ ancestry $1-q(x,t,r)$
         (unconditional on the ancestry of the linked allele), computed numerically.
-        Each line represents a locus some distance $r$ away from the true target of selection, 
-        and $r=0$ represents the locus that is under selection. Colors correspond to different values of $r$, transitioning from red (tight linkage to selected site) to blue (distantly linked).
+        \textbf{Top:}
+        loci across a chromosome unlinked to the selected locus.
+        \textbf{Bottom:}
+        loci across the chromosome carrying the selected locus,
+        with colors corresponding to different values of $r$, from red (tight linkage to selected site) to blue (distantly linked).
+        The two sets of clines are shown superimposed in Figure \ref{alleleFreq_tau100}.
     }\label{fig:alleleFreq_tau100_comparison}
@@
 divided by the mean lengths of the two blocks directly flanking the block containing $m$.
-
-\plr{Below we use $l_B$ to mean the normalized version also.  I am defining consistent notation, $\bar l_B$.}
 
@@
 up to stochasticity due to genetic drift, which is more pronounced at lower population densities.
-Figure \ref{fig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
-match simulated clines in Figure \ref{alleleFreq_tau100} quite well.
+For instance, figure \ref{fig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
+match simulated clines quite well.
 Figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}
@@
 In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} 
-can inform the demographic history of hybridization/admixture from present day samples \citep[e.g.][]{Hellenthal2014}. 
+can inform the demographic history of hybridization/admixture from present day samples \citep[e.g.,][]{Hellenthal2014}. 
 We described how selection against hybrid incompatibilities  
@@
 Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, 
-we believe that this is unlikely to drastically influence our findings.
-As discussed in the Introduction,
-previous studies have demonstrated that models of underdominant selection 
-share similar properties with more realistic models, e.g.,
-selection on an ecological gradient or multilocus, epistatic selection
-\citep{Slatkin1973,Barton1993,Kruuk1999}.
+theoretical and simulation studies show that clines formed by underdominant selection and extrinsic selection 
+are effectively indistinguishable without additional information \citep{may1975gene,Barton1993,Kruuk1999}, 
+especially over short time scales \citep{Barton1979a},
+and their shape is determined mostly by the mean fitness of the heterozygote \citep{Slatkin1973}. 
 \revpoint{1}{1}
@@
 %Under dominance is quite rare in nature, because underdominant loci cannot invade when rare, and thus our major assumption is likely rarely met. 
-%We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar proprrties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
+%We believe, however, that this assumption is unlikely to drastically influence our results --- previous research has found that under dominant loci  have similar properties in hybrid zones as more realistic genetic models, such as ecological selection \yb{(CITE: (BARTON and HEWITT 1989; BARTON and GALE 1993)}.
 %An additional assumption of our analytical model is that we ignore coalescence.  

### 2016-04-26 drafts/supplement.tex
changed a word
fixed fig 3 caption, more edits

@@
     and with $s=0.01$.
-        50 demes, each with a population size of 500 diploid individuals.
+    The simulated zone had 50 demes, each with a population size of 500 diploid individuals.
     Each line represents a locus some distance $r$ away from the true target of selection, 
@@
     and with $s=0.1$.
-        50 demes, each with a population size of 500 diploid individuals.
+    The simulated zone had 50 demes, each with a population size of 500 diploid individuals.
     Each line represents a locus some distance $r$ away from the true target of selection, 
@@
  }\label{alleleFreq_tau1000}
+\end{figure}
+
+\begin{figure}
+    \includegraphics{{figs/alleleFrequencies_sim_SIGMA3_Ninds250000_ndemes500_s0.01_tau100}.pdf}
+\caption{
+    Frequency of ancestry $B$,
+    across geography and at several physical positions on the genome, simulated for a hybrid zone
+    $T=100$ generations after secondary contact,
+    and with $s=0.01$.
+    The simulated zone had 500 demes, each with a population size of 500 diploid individuals.
+    Each line represents a locus some distance $r$ away from the true target of selection, 
+    and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
+    identical parameters except that $s=0$.
+}\label{sfig:extra_long_cline}
 \end{figure}

### 2016-04-26 drafts/supplement.tex
fixing up the supplement

--- 2016-04-26 drafts/supplement.tex
@@
 \begin{figure}
-    \includegraphics{{figs/alleleFrequencies_sim_s0.01_tau100_closely_linked}.pdf}
+    \includegraphics{figs/alleleFrequencies_sim.pdf}
     \caption{
@@
 \begin{figure}
-\includegraphics{figs/ratioAdjacentBlocksAlongChromAncBConditioning.pdf}
+\includegraphics{figs/s0point01_ratioAdjacentBlocksAlongChromAncBConditioningHighRes.pdf}
     \caption{The statistic $\bar C(m,x)$: the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
-Each line represents a deme and is normalized by mean block length across the chromosome in the deme.}\label{Supp:ratioBlockAdjacent}
+Each line represents a deme and is normalized by mean block length across the chromosome in the deme.
+This  simulation corresponds to that represented in Figure~\ref{Fig:blockLengthsZoom}; note the difference in y-axis scale.}\label{Supp:ratioBlockAdjacent}
 \end{figure}

### 2016-04-26 drafts/review-responses-april-2016.tex

@@
 Although most of the simulations shown in the main text have high population densities,
-we have run simulations with considerably lower densities, 
+we have run simulations with lower densities, 
 as shown for instance in figures \ref{sfig:condAlleleFreq_tau80_comparison}, \ref{sfig:condAlleleFreq_tau320_comparison}, and \ref{sfig:condAlleleFreq_tau1280_comparison}.
@@
 we did encounter empty demes in some simulations with smaller population sizes.
-\plr{add reference to more low pop density figures if added}
+We don't examine the parameter regime with neighborhood size on the order of ten individuals,
+where we expect different dynamics to be at play,
+and have a note to this effect \llname{ll:much_lower_densities}.
+
+As for the width of the range, we had indeed also run additional simulations (and numerical solutions) on wider ranges to check for boundary effects;
+we now have a note to this effect at \llname{ll:narrow_sims}.
 

### 2016-04-26 drafts/hybrid_zone.tex
fixing up the supplement
minor fixen

@@
         as described in the text (see definition of $q_z(x,t,r)$).
-        See figure \ref{sfig:linked_cline_heatmaps_longtime}
+        See Figure \ref{sfig:linked_cline_heatmaps_longtime}
         for the same figure over a longer period of time.
@@
 This is seen in Figure \ref{fig:alleleFreq_tau100_comparison},
-and supplementary figures \ref{alleleFreq_tau100}, \ref{alleleFreq_tau100_weaker_s} and \ref{alleleFreq_tau1000}. 
+and supplementary Figures \ref{alleleFreq_tau100}, \ref{alleleFreq_tau100_weaker_s} and \ref{alleleFreq_tau1000}. 
 Clines in ancestry frequency matched expected values found from deterministic theory
 up to stochasticity due to genetic drift, which is more pronounced at lower population densities.
-For instance, figure \ref{fig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
+For instance, Figure \ref{fig:alleleFreq_tau100_comparison} shows that expected clines computed from equation \eqref{eqn:q_pde}
 match simulated clines quite well.
@@
 % drift will kick in and move frequencies \emph{range-wide}.
-% We see that in figure \ref{sfig:condAlleleFreq_tau1280_comparison}.
+% We see that in Figure \ref{sfig:condAlleleFreq_tau1280_comparison}.
 % That's what I was trying to say here.
@@
 \paragraph{Identifying selected loci}
-The statistic $C(m,x)$ shows promise for identifying selected loci under some circumstances.
+The statistics $\bar l(m,x)$ and $C(m,x)$ show promise for identifying selected loci under some circumstances.
 As expected, regions surrounding a locus under selection are more resistant to introgression,
@@
 
-\alisa{To provide some intuition about the physical scale of this signal, we note that}
-genomic regions are inherited from ancestors $T$ generations ago
-in blocks of size roughly $1/T$. \revpoint{1}{6}
+Using the fact that genomic regions are inherited from ancestors $T$ generations ago
+in blocks of size roughly $1/T$, \revpoint{1}{6}
+we can provide some intuition about the physical scale of the expected signal.
 There will be a long segment of $A$ ancestry

### 2016-04-26 drafts/supplement.tex
edits to the supplement

@@
     The simulated zone had 50 demes, each with a population size of 500 diploid individuals.
-    Each line represents a locus some distance $r$ away from the true target of selection, 
+    Each line represents a locus some distance $r$ away from the true target of selection 
     and $r=0$ represents the locus that is under selection. Grey lines represent the same positions from a simulation with
@@
 \begin{figure}
-    \includegraphics{{figs/alleleFrequencies_sim_SIGMA3_Ninds250000_ndemes500_s0.01_tau100}.pdf}
+   \includegraphics{{figs/polydiv_by_coverage_SD2-12}.pdf}
 \caption{
@@
 \begin{figure}
-    \includegraphics{{figs/simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_blocksAlongChromAncBConditioning_nonnormalized}.pdf}
+    \includegraphics{{figs/simulation_SIGMA1_Ninds25000_ndemes50_s0.01_tau1000_blocksAlongChromAncBConditioning_nonnormalized}}
     \caption{
@@
     divided by the mean over the chromosome at that location;
-    the simulation was run with $s=1$ and $\sigma=1$,
+    the simulation was run with $s=0.01$ and $\sigma=1$,
     50 demes, each containing 500 diploid individuals.
@@
 \includegraphics{figs/blocksAlongChromHeatmapAncBConditioning.pdf}
-\caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome under $T=1000$, $s=1$ and $\sigma=1$,
+\caption{Heatmap of mean block length $l_B(m)$ along a simulated chromosome under $T=1000$, $s=0.01$ and $\sigma=1$,
     50 demes, each containing 500 diploid individuals. }\label{Supp:blockLengthHeatmap}
@@
 \includegraphics{figs/blocksAlongChromNoConditioning.pdf}
-\caption{Mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
+\caption{Mean block length $l(m)$, not conditioning on ancestry, surrounding a given position along the genome with a single underdominant site with parameters as for Fig.~\ref{Fig:blockLengths} ($s=0.01, T=1000, \sigma=1$).
+}\label{Supp:blockLengthNoAnc}
 \end{figure}
@@
 \includegraphics{figs/blocksAlongChromHeatmap.pdf}
-\caption{Heatmap of mean block length $l(m)$ along a simulated chromosome with a single underdominant site ($s=0.01, T=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
+\caption{Heatmap of mean block length $l(m)$ (i.e.
+no conditioning on ancestry) along a simulated chromosome with a single underdominant site with parameters as for Fig.~\ref{Fig:blockLengths}($s=0.01, T=1000, \sigma=1$) }\label{Supp:blockLengthHeatmapNoAnc}
 \end{figure}
@@
 \includegraphics{figs/adjacentBlocksAlongChromAncBConditioning.pdf}
-\caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus. ($s=0.01, T=1000$)}\label{Supp:adjacentBlocks}
+\caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus and with parameters as for Fig.~\ref{Fig:blockLengths} ($s=0.01, T=1000, \sigma=1$)}\label{Supp:adjacentBlocks}
 \end{figure}
@@
 \includegraphics{figs/s0point01_ratioAdjacentBlocksAlongChromAncBConditioningHighRes.pdf}
-    \caption{The statistic $\bar C(m,x)$: the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{The statistic $\bar C(m,x)$: the normalized ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
 Each line represents a deme and is normalized by mean block length across the chromosome in the deme.
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning.pdf}
-    \caption{Heatmap of $C(m,x)=\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{Heatmap of $C(m,x)=\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000, \sigma=1$).
 }\label{Supp:ratioBlockAdjacentHeatmap}
@@
         \caption{
-     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (top) and without conditioning (bottom) } \label{sfig:blockLengthPlot_t1000_s0.001}
+     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (top) and without conditioning (bottom) and with parameters as for Fig.~\ref{Fig:blockLengths} } \label{sfig:blockLengthPlot_t1000_s0.001}
 \end{figure}

### 2016-04-26 drafts/hybrid_zone.tex
reset changes that were breaking supplement

@@
 while if selection is weak, the selected lineage does not move very fast,
-and so the strength of signal from elevated $l_B(m)$ is diminished \alisa{(Fig.~\ref{sfig:blockLengthPlot_t1000_s0.001})}. 
-Indeed, if $s$ is less than $x/(\sigma T)$, then only one block of ancestry $A$ is expected to be seen about a selected $A$ allele,
-and $l_A(x,m)$ is expected to be $2/T$.
+and the hybrid zone is wider, allowing more recombination,
+and so the strength of signal from elevated $l_B(m)$ is diminished.
+For instance, no signal is seen in Figure~\ref{sfig:blockLengthPlot_t1000_s0.001}.
+% Indeed, if $s$ is less than $x/(\sigma T)$, then only one block of ancestry $A$ is expected to be seen about a selected $A$ allele,
+% and $l_A(x,m)$ is expected to be $2/T$.
 For similar reasons, power and resolution are best at intermediate $T$.

### 2016-04-26 drafts/supplement.tex
added missing figures
fixed the supplement

--- 2016-04-26 drafts/supplement.tex
@@
 \includegraphics{figs/blocksAlongChromNoConditioning.pdf}
-\caption{Mean block length $l(m)$ surrounding a given position along the genome with a single underdominant site ($s=0.01, T=1000$). }\label{Supp:blockLengthNoAnc}
+\caption{Mean block length $l(m)$, without conditioning on ancestry, surrounding a given position along the genome with a single underdominant site with parameters as for Fig~\ref{Fig:blockLengths}.
+($s=0.01, T=1000, \sigma=1$).
+}\label{Supp:blockLengthNoAnc}
 \end{figure}
@@
 \includegraphics{figs/blocksAlongChromHeatmap.pdf}
-\caption{Heatmap of mean block length $l(m)$ along a simulated chromosome with a single underdominant site ($s=0.01, T=1000$) }\label{Supp:blockLengthHeatmapNoAnc}
+\caption{Heatmap of mean block length $l(m)$, without conditioning on ancestry, along a simulated chromosome with a single underdominant site, with parameters as for Fig~\ref{Fig:blockLengths} ($s=0.01, T=1000, \sigma=1$) }\label{Supp:blockLengthHeatmapNoAnc}
 \end{figure}
@@
 \includegraphics{figs/adjacentBlocksAlongChromAncBConditioning.pdf}
-\caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus. ($s=0.01, T=1000$)}\label{Supp:adjacentBlocks}
+\caption{Mean block length of $l_A(m\pm)$ across chromosome with single under dominant site, conditioning on ancestry $B$ at the selected locus, with parameters as for Fig~\ref{Fig:blockLengths} ($s=0.01, T=1000, \sigma=1$)}\label{Supp:adjacentBlocks}
 \end{figure}
@@
 \includegraphics{figs/s0point01_ratioAdjacentBlocksAlongChromAncBConditioningHighRes.pdf}
-    \caption{The statistic $\bar C(m,x)$: the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{The normalized statistic $\bar C(m,x)$: the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ of mean block length and mean adjacent block lengths across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
 Each line represents a deme and is normalized by mean block length across the chromosome in the deme.
@@
 \includegraphics{figs/ratioAdjacentBlocksAlongChromHeatmapAncBConditioning.pdf}
-    \caption{Heatmap of $C(m,x)=\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site ($s=0.01, T=1000$).
+    \caption{Heatmap of $C(m,x)=\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i-)+I_A(m_i+)}}$ across a simulated chromosome with a single underdominant site and conditioning on ancestry $B$ at the selected site and parameters as for Fig.~\ref{Fig:blockLengths} ($s=0.01, T=1000, \sigma=1$).
 }\label{Supp:ratioBlockAdjacentHeatmap}
@@
         \caption{
-     Block lengths along chromosome under $s=0.001$, conditioning on ancestry B (top) and without conditioning (bottom) } \label{sfig:blockLengthPlot_t1000_s0.001}
+     Block lengths along chromosome under $s=0.001$ and other parameters as for Fig.~\ref{Fig:blockLengths}, conditioning on ancestry B (top) and without conditioning (bottom) } \label{sfig:blockLengthPlot_t1000_s0.001}
 \end{figure}

### 2016-04-26 drafts/hybrid_zone.tex
save stuff for new fig

@@
 \begin{figure}
-\includegraphics{figs/blocksAlongChromAncBConditioningHighRes.pdf}
+\includegraphics{figs/s0point01_BlocksAlongChromAncBConditioning.pdf}
 \caption{\textbf{Normalized mean enclosing block length}, $\bar l_B(m,x)$, along a simulated chromosome of length $1M$ with selected locus of $s=0.01$ at position $0.5M$ in a hybrid zone of $T=1000$. 
@@
     Chromosome were sampled across the hybrid zone, which consists of 50 demes, each containing 500 diploid individuals. 
-    The same simulation and statistic are shown in Fig.~\ref{Fig:blockLengths}, on coarser genomic scale.
+    The same simulation and statistic are shown in Fig.~\ref{Fig:blockLengths}, on a coarser genomic scale.
 }\label{Fig:blockLengthsZoom}
@@
 
-Our results suggest that the statistic $l(x,m)$ 
--- the mean genetic length of ancestry surrounding a site of a given lineage --
+Our results suggest that the statistic $l_B(x,m)$ 
+-- the mean genetic length of ancestry surrounding a site of a given ancestry --
 could help identify selected loci in hybrid zones. 
-The ratio of mean adjacent block lengths, $C(x,m)$, alsot shows a very sharp peak at the selected site,
+The ratio of mean adjacent block lengths, $C(x,m)$, also shows a very sharp peak at the selected site,
 suggesting that, despite the fact that haplotypes surrounding incompatibilities might be quite long, 

### 2016-04-26 drafts/figs/Plot_frequency_clines.R

@@ -16,9 +16,9 @@ get_simparams <- function (fname) {
 }
 
 
-# positions = seq(0.2,0.8,0.01)
+positions = seq(0.2,0.8,0.01)
 # positions = seq(0.48,0.52,0.001)
-positions = c(0,0.01,0.02,0.05,0.1,0.2,0.3)
+# positions = c(0,0.01,0.02,0.05,0.1,0.2,0.3)
 #cols = c(2,3,6,11,21,31)
 
 sel.sim.file <- "../../../sims/simulation_SIGMA1_Ninds25000_ndemes50_s0.1_tau100_run2016-01-11_simsums.Robj"
@@ -56,6 +56,12 @@ if (!file.exists("plot_frequency_clines_revision.Robj")) {
             CHR=1, 
             ndemes=neu.params$ndemes)
 
+    save(B_freq, B_neu, file="plot_frequency_clines_revision.Robj")
+} else {
+    load("plot_frequency_clines_revision.Robj")
+}
+
+if (!file.exists("plot_frequency_clines_long_revision.Robj")) {
 
     load(long.sim.file)
     # get frequencies for simulation on wider range
@@ -75,9 +81,9 @@ if (!file.exists("plot_frequency_clines_revision.Robj")) {
 
 
 
-    save(B_freq, B_neu, B_long, B_long_neu, file="plot_frequency_clines_revision.Robj")
+    save(B_long, B_long_neu, file="plot_frequency_clines_long_revision.Robj")
 } else {
-    load("plot_frequency_clines_revision.Robj")
+    load("plot_frequency_clines_long_revision.Robj")
 }
 
 
@@ -170,11 +176,11 @@ xx = (1:long.params$ndemes)-0.5-long.params$ndemes/2
 
 pdf(file="alleleFrequencies_sim_SIGMA3_Ninds250000_ndemes500_s0.01_tau100.pdf",height=4,width=6,pointsize=10)
 par(mar=c(3.5,3.5,0.5,0.5))
-    matplot(xx, 1-B_long_neu/1000, xlab="", ylab="", type="l", lty=1, col="grey", lwd=0.5, ylim=c(0,1))
+    matplot(xx, B_long_neu/1000, xlab="", ylab="", type="l", lty=1, col="grey", lwd=0.5, ylim=c(0,1))
     for(i in 1:30){
[…]

### 2016-04-26 drafts/figs/Plot_frequency_clines.R
fixed fig

@@ -176,11 +176,11 @@ xx = (1:long.params$ndemes)-0.5-long.params$ndemes/2
 
 pdf(file="alleleFrequencies_sim_SIGMA3_Ninds250000_ndemes500_s0.01_tau100.pdf",height=4,width=6,pointsize=10)
 par(mar=c(3.5,3.5,0.5,0.5))
-    matplot(xx, B_long_neu/1000, xlab="", ylab="", type="l", lty=1, col="grey", lwd=0.5, ylim=c(0,1))
+    matplot(xx, B_long_neu[,1:30]/1000, xlab="", ylab="", type="l", lty=1, col="grey", lwd=0.5, ylim=c(0,1))
     for(i in 1:30){
-        matpoints(xx, B_long[,c(i,62-i)]/1000, type="l", col=(rainbow(43))[32-i], lty=1, lwd=0.75 )
+        matlines(xx, B_long[,i]/1000, col=(rainbow(43))[32-i], lty=1, lwd=0.75 )
     }
-    points(xx, B_long[,31]/1000, col="red", lwd=2, type="l" )
+    lines(xx, B_long[,31]/1000, col="red", lwd=2 )
     mtext("Geographic position",side=1,line=2)
     mtext(expression(p[B]),side=2,line=2.5)
     legend("bottomright",legend = c("r=0","r=0.005","r=0.01","r=0.02","r=0.03","no seln."),col=c("red",rainbow(43)[c(6,11,21,31)],"darkgrey"),lwd=c(2,rep(0.75,4),0.5),cex=0.8)

### 2016-04-26 drafts/hybrid_zone.tex
fixed bib format

@@
 may help identify putative targets of selection in hybridizing populations, 
-providing another source of information in addition to gradients in allele frequencies used in current approaches \citep[e.g.,][]{Porter1997, Gompert2012}. 
+providing another source of information in addition to gradients in allele frequencies used in current approaches \citep[e.g.,][]{porter1997pontia, Gompert2012}. 
 Specifically, we find that selection against hybrid incompatibility loci changes the extent of correlated ancestry --- that is, 
@@
 Popular approaches to identifying loci under selection in hybrid zones 
-involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
+involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{porter1997pontia,Gompert2012}.  
 The availability of genomic data has made it possible to use local ancestry as an additional source of information. 
@@
 A number of studies have described heterogeneous patterns of genetic divergence across the genome. 
-Work on these ``islands of divergence'' \citep{turner2005genomic,Nosil2009} and related patterns have been largely descriptive \citep{Cruickshank2014,Noor2009}.
+Work on these ``islands of divergence'' \citep{turner2005genomic,Nosil2009} and related patterns have been largely descriptive \citep{Cruickshank2014,noor2009islands}.
 Our study here contributes to a model-based understanding of 
@@
 %The past years have witnessed evocative metaphors of seascapes to describe heterogeneous patterns of genetic divergence across the genome.  
-%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{turner2005genomic, Nosil2009} have inspired much research (and controversy  \citep{Noor2009,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
+%The metaphors of ``islands'', ``continents'' and ``archipelagoes'' of divergence \citep{turner2005genomic, Nosil2009} have inspired much research (and controversy  \citep{noor2009islands,Cruickshank2014}) in speciation genomics, but have been largely descriptive.
 %REFS  http://www.ncbi.nlm.nih.gov/pubmed/16076241 http://www.ncbi.nlm.nih.gov/pubmed/19143936 http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2809014/ http://onlinelibrary.wiley.com/doi/10.1111/mec.12796/abstract
@@
 cannot introgress until recombination separates them;
-our model \alisa{provides some rough expectation on how quickly this should happen}.
+our model provides some rough expectation on how quickly this should happen.
 
@@
 PR was supported in part by NSF grant DBI-1262645 and a Sloan Research Fellowship.
-\alisa{More info here}
+% \alisa{More info here}
 

### 2016-04-26 drafts/review-responses-april-2016.tex
small changes

@@
 Below we provide a point-by-point response to reviewer and AE comments, describing how we addressed concerns raised by our readers.
-The major themes of our changes are (\emph{1}) Attempting to make our mathematical derivations more accessible for Molecular Ecology’s wide audience  (\emph{2}) Additional justification for using a genetic model of under dominance to model hybrid incompatibilities, in which we make heavy reference to the historical theory justifying this approach, and (\emph{3}) Additional simulations to show that our results are not driven by boundaries imposed in a defined range, and (\emph{4})  A streamlined introduction that makes clear our major goals and provides a stronger structure for our manuscript.
+The major themes of our changes are (\emph{1}) Attempting to make our mathematical derivations more accessible for Molecular Ecology'۪s wide audience  (\emph{2}) Additional justification for using a genetic model of under dominance to model hybrid incompatibilities, in which we make heavy reference to the historical theory justifying this approach, and (\emph{3}) Additional simulations to show that our results are not driven by boundaries imposed in a defined range, and (\emph{4})  A streamlined introduction that makes clear our major goals and provides a stronger structure for our manuscript.
 }

### 2016-04-26 drafts/supplement.tex
small changes
fixed S19 caption

@@
 \renewcommand{\thefigure}{S\arabic{figure}}
-
+\clearpage
 \begin{figure}
@@
 \end{figure}
-
+\clearpage
 \begin{figure}
@@
         \caption{
-     Block lengths along chromosome under $s=0.001$ and other parameters as for Fig.~\ref{Fig:blockLengths}, conditioning on ancestry B (top) and without conditioning (bottom) } \label{sfig:blockLengthPlot_t1000_s0.001}
+     Normalized ancestry $B$ block lengths (top) and all ancestry block lengths (bottom) along chromosome under $s=0.001$ and other parameters as for Fig.~\ref{Fig:blockLengths}.
+} \label{sfig:blockLengthPlot_t1000_s0.001}
 \end{figure}

### 2016-04-26 drafts/supplement.tex
fixed S19 caption

@@
         \caption{
-     Normalized ancestry $B$ block lengths (top) and all ancestry block lengths (bottom) along chromosome under $s=0.001$ and other parameters as for Fig.~\ref{Fig:blockLengths}.
+     Normalized ancestry $B$ block lengths (top) and all ancestry block lengths (bottom) along chromosome under $s=0.001$ and other parameters as for Fig.~\ref{Fig:blockLengths} ($\tau=1000,\sigma=1$).
 } \label{sfig:blockLengthPlot_t1000_s0.001}

### 2016-04-27 drafts/review-responses-april-2016.tex
responses done?

--- 2016-04-26 drafts/review-responses-april-2016.tex
@@
 \begin{minipage}[b]{2.5in}
-    Lewis-Sigler institute for integrative genomics\\
-    Princeton University\\
-    Yaniv's address\\
-    \emph{and} Molecular and Computational Biology\\
-    University of Southern California \\
+    Alisa Sedghifar \\
+    Yaniv Brandvain \\
+    \emph{and} Peter Ralph \\
   \today
@@
 
-\yb{We are pleased to submit a revision of our Manuscript, `Beyond clines: lineages and haplotype blocks in hybrid zones' (MEC-16-0266) for the special issue 'Genomics of Hybridization'.
+We are pleased to submit a revision of our Manuscript, `Beyond clines: lineages and haplotype blocks in hybrid zones' (MEC-16-0266) for the special issue 'Genomics of Hybridization'. 
 We believe that this version is much improved by changes we have made to address reviewer and AE comments. 
 Below we provide a point-by-point response to reviewer and AE comments, describing how we addressed concerns raised by our readers.
-The major themes of our changes are (\emph{1}) Attempting to make our mathematical derivations more accessible for Molecular Ecology’s wide audience  (\emph{2}) Additional justification for using a genetic model of under dominance to model hybrid incompatibilities, in which we make heavy reference to the historical theory justifying this approach, and (\emph{3}) Additional simulations to show that our results are not driven by boundaries imposed in a defined range, and (\emph{4})  A streamlined introduction that makes clear our major goals and provides a stronger structure for our manuscript.
-}
 
-\begin{quote}
-Reviewer 1 would like to see development of the finding that ancestral blocks are longer around a selected locus into a practical test.
-I think that this would be a substantial project for the future, and that it is valuable to first explain and understand the theoretical prediction, as is done here.
-Those analysing actual data can of course look for this pattern without needing an already implemented test.
-\end{quote}
+The major themes of our changes are 
+(\textbf{1}) Attempting to make our mathematical derivations more accessible for Molecular Ecology's wide audience  
+(\textbf{2}) Additional justification for using a genetic model of underdominance to model hybrid incompatibilities, in which we make heavy reference to the historical theory justifying this approach,
+(\textbf{3}) Additional simulations to show that our results are not driven by boundaries imposed in a defined range, and 
+(\textbf{4})  A streamlined introduction that makes clear our major goals and provides a stronger structure for our manuscript.
 
-\begin{quote}
-Rev.
-1 is also concerned that only under dominance is simulated.
-However, it is clear that any form of selection that gives the same marginal selection coefficient will give essentially the same effect: a large body of work on clines shows that different selective mechanisms are indistinguishable, and it is reasonable for the authors to rely on that.
-The other main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).
-I don't know of a reference where they have been written in exactly this form, but the recombination terms and the diffusion terms in Eq.
-4 both go back a long way.
-\end{quote}
-
-\begin{quote}
-Reviewer 2 is mainly concerned about the fit between theory and simulations.
-I am not worried about this, since the deterministic mathematics stands independently, and the numerical computations are straightforward.
-The stochastic simulations are useful mainly as illustrations of the process.
-Theory and simulations are compared in the supplementary material; it should be stated more clearly how closely these match.
-(A precise fit is not expected anyway because there is substantial drift).
-Rev.
-2 also worries that the simulations are over a rather narrow range; I suspect that the authors should have results that address sensitivity to the boundaries.
-Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that the authors suggest otherwise.
-\end{quote}
+Thanks to both reviewers and the AE for the useful (and prompt) feedback.
 
@@
 
+\pagebreak
+
 %%%%%%%%%%%%%%
 \reviewersection{AE}
+
+\begin{quote}
+Reviewer 1 would like to see development of the finding that ancestral blocks are longer around a selected locus into a practical test.
+I think that this would be a substantial project for the future, and that it is valuable to first explain and understand the theoretical prediction, as is done here.
+Those analysing actual data can of course look for this pattern without needing an already implemented test.
+\end{quote}
+
+We would also like very much to develop a practical test,
+but agree that falls outside of the scope of this paper.
+
+\begin{quote}
+Reviewer 1 is also concerned that only under dominance is simulated.
+However, it is clear that any form of selection that gives the same marginal selection coefficient will give essentially the same effect: a large body of work on clines shows that different selective mechanisms are indistinguishable, and it is reasonable for the authors to rely on that.
+The other main concern was about the novelty of Eqs 4, 5; these are indeed standard (which I think the authors intended by referring to them as the Kolmogorov backward equations).
+I don't know of a reference where they have been written in exactly this form, but the recombination terms and the diffusion terms in Eq.
+4 both go back a long way.
+\end{quote}
+
+As discussed in more detail below,
+we now provide citations to the literature showing that different underlying mechanims give similar clines,
+and have put in some citations to the literature making similar arguments (which we did indeed think of as standard).
+
+\begin{quote}
+Reviewer 2 is mainly concerned about the fit between theory and simulations.
+I am not worried about this, since the deterministic mathematics stands independently, and the numerical computations are straightforward.
+The stochastic simulations are useful mainly as illustrations of the process.
+Theory and simulations are compared in the supplementary material; it should be stated more clearly how closely these match.
+(A precise fit is not expected anyway because there is substantial drift).
+Rev.
+2 also worries that the simulations are over a rather narrow range; I suspect that the authors should have results that address sensitivity to the boundaries.
+Regarding specific point 1, I don't know of a previous derivation of exactly these equations, but similar equations have been used widely, and I don't think that the authors suggest otherwise.
+\end{quote}
+
+We have added more discussion in this direction, as described below.
+We had indeed run many more simulations than are shown in the text,
+including checking that the boundaries were not affecting things substantially (for both simulations and PDE solutions);
+now we have a statement to that effect.
 
@@
 \end{quote}
+
+Thanks for the encouraging words!
 
@@
 \reply
-There might have been some confusion about our use of the floor function?
+There might have been some confusion about our undocumented use of the floor function?
 We now describe the dispersal distribution in a more straightforward way \revref.
@@
 \reply
-Well, we hoped to mean ``reality'', but perhaps ``verify that theory matches reality'' is too strong; we've adjusted this to say ``realistic, individual-based models.''.
+Well, we hope to mean ``reality'', but you're right, ``verify that theory matches reality'' is too strong; we've changed this to say ``realistic, individual-based models.''.
 %\alisa{I think we do mean ``reality" here, but alternate opinions welcome}
@@
 \end{quote}
+
+We also hope to at least stimulate more exploration with real datasets
+(and collection of datasets this can be done on).
 
@@
 To aid our readers in developing this understanding, we have gone through considerable pains to map terms in our equations to the biological processes of migration, selection and recombination. 
-We believe that our efforts will help build a better mechanistic understanding in our readers in addition to making them aware of our results.
+We hope that our efforts will help readers to build a better mechanistic understanding of the underlying process.
 
@@
 \reply
-\yb{
     We agree that a long set of theoretical results can be tedious.
 We have therefore reshaped the introduction to highlight our major question ``How does selection against incompatibilities in a hybrid zone influence the length distribution of ancestry tracts?''
-}
 
@@
 \reply
-\plr{(comment is more general than that particular location; say something about this)}
-We added an introductory sentence at \revref to provide some structure. 
-We have also gone through to clarify other sections that felt unclear.
+We added an introductory sentence at \revref{} 
+to provide some structure. 
+We have also attempted to clarify other sections that felt unclear.
 
 \begin{point}{\revref}
-``In the notation above, $q_z(x, t, r) = P^x{Z_t = A}$.''
+``In the notation above, $q_z(x, t, r) = P^x\{Z_t = A\}$.''
 I did not see the notation $P^x$ above, and the RHS does not depend on small z.
@@
 \reply
-\alisa{This is a reference that is expected to appear in the same issue, so we do not have page numbers. The hope is that this can be filled in during final proofs for the issue}
+This is a reference that is expected to appear in the same issue, so we do not have page numbers. The hope is that this can be filled in during final proofs for the issue.
 
@@
 
-\begin{quote}
-    In this paper, the authors extend previous work on the diffusion of neutral alleles in a secondary-contact zone to include selection against hybrids.
+% \begin{quote}
+%     In this paper, the authors extend previous work on the diffusion of neutral alleles in a secondary-contact zone to include selection against hybrids.
 Building on the classic diffusion model for the dynamics of a cline at a selected locus, they formalize the dynamics of linked neutral loci and investigate the distribution of ancestry block-lengths around the selected locus.
-
-    The authors obtain numerical results for the flattening of the cline at a neutral locus linked to an under-dominant selected locus in the absence of genetic drift.
+% 
+%     The authors obtain numerical results for the flattening of the cline at a neutral locus linked to an under-dominant selected locus in the absence of genetic drift.
 They also outline scaling arguments for the rates of movement and length of the introgressing blocks around the selected locus.
 Simulations are set up to test the theoretical predictions and illustrate other interesting patterns such as the distribution of haplotype (ancestry block?) lengths.
-\end{quote}
+% \end{quote}
 
@@
 \reply
-\alisa{Our hope was that the simulations and theory would provide qualitative support for one another.
-Some predictions made by the theory e.g.
-time and spatial scales do seem to correspond.
-We will try to do a better job of signposting our unique contributions??}
+We have aimed to compare theoretical predictions to simulation everywhere that we have quantitative predictions.
+Unfortunately, expressions for mean lengths of haplotype blocks are only order-of-magnitude (but roughly agree, as we discuss),
+and obtaining numerical predictions from the PDE for these is rather daunting.
+Given the very good fit for linked clines, we are not concerned, however.
 
 The rescaling arguments we provide are of a standard sort in the literature;
-we've inserted comments to this effect, at \llname{ll:rescaling} and \llname{ll:may_citation}.
+we've inserted comments to this effect in several places, including at \llname{ll:rescaling} and \llname{ll:may_citation}.
 
@@
 \reply
-\alisa{Should we be running more sims? Nick seems to think we have done this, but I don't think we have.}
-\plr{run sims with longer range}
-\plr{also run sims in part of param range where nothing's supposed to happen?}
+As mentioned above, we had, indeed, run more simulations on longer habitats, to check the range was sufficient.
+As expected, the boundaries were not impacting dynamics in the center of the zone, as there isn't enough time for this to happen.
+Over longer periods, with small population densities, the total size of the range will affect rangewide drift,
+as perhaps seen in Figure \ref{sfig:condAlleleFreq_tau1280_comparison}; 
+but this is a relatively minor point.
 
@@
 What are the replicate lines for the same $r$?
-Why re the lines for $r = 0$ all slightly displaced to the right?
+Why are the lines for $r = 0$ all slightly displaced to the right?
 Also, it would appear that the color scheme in associated Figure S2 (bottom) is mismatched with the legend, shifted to the red part of the spectrum.
@@
 \reply
-\alisa{We have switched figures S2 and 3 and fixed the colors.
+We have switched figures S2 and 3 and fixed the colors.
 The colored lines are not all represented in the legend which seems to have caused some confusion and we have tried to address this in the figure legends.
-We don't interpret the mismatch as a displacement to the right, as translating the lines horizontally will not improve the overall fit, and attribute this to drift in the simulations.}
+We don't interpret the mismatch as a displacement to the right, as translating the lines horizontally will not improve the overall fit, and attribute this to drift in the simulations.
 
@@
 \reply
-\plr{We meant that the allele frequencies in the \emph{entire population} had shifted, which is determined by the total (effective) population size.  
-However, we decied to }
-% \yb{Thanks for spotting this. Because we do nothing quantitative with respect to drift here, we}
-\yb{simply state the our observed wobble of the center of the tension zone is consistent with previous theory.
-However, more broadly throughout the manuscript we are not sure about relying on $N \sigma \sqrt{s}$ as we are concerned with loci linked to the selected site, and we are not aware of results extending this rule to an unlinked locus.}
-\plr{(Do you mean "linked locus"?)}
-%\alisa{We are concerned here with linked loci, so the $N\sigma\sqrt{s}$ rule of thumb does not necessarily apply here? Also, within demes, population size is much smaller (500?)}
-%\plr{I agree: they are probably thinking of selected loci, not linked ones. We're talking about global drift not local drift. Yaniv to investigate.}
+We meant that the allele frequencies in the \emph{entire population} had shifted, which is determined by the total (effective) population size.  
+However, we decied to 
+simply state the our observed wobble of the center of the tension zone is consistent with previous theory.
+However, more broadly throughout the manuscript we are not sure about relying on $N \sigma \sqrt{s}$ as we are concerned with loci linked to the selected site, and we are not aware of results extending this rule to a linked locus.
 
@@
 As noted above, we have provided some citations to similar arguments in the literature
-(although such arguments are often used without citation; see, for instance, \citet{durrett2007width}).
+(although such arguments are often used without citation).
 
@@
 \reply
-We now have much more discussion about drift, in particular at \revref.
-
-\plr{I don't understand this:}
-\alisa{We should add a line or two in the discussion about drift.
-However, IBD is perhaps not relevant over the timescales we are considering and is not necessarily ancestry informative, and since we are not interested in the equilibrium state, it is not clear that we need to worry about this limit.
-}
+We agree this is an important point, and now have much more discussion about drift, in particular at \revref{}
+and a note about drift maintaining certain clines in old zones at \llname{ll:old_drift}.
+We would like to explore this more, but it seems beyond the scope of the paper.
 
@@
 \reply
-Right; we've added more discussion of this.
+That's correct; we've rewritten the explanation.
 
@@
 \reply
-\alisa{We have created a new repository with minimally required scripts and a descriptive README file}
+We have created the Data Accessibility section, and have written improved READMEs.
+We readily admit that many of the scripts are not terribly user-friendly,
+but there is an easy interface that will run a simulation,
+and sufficient documentation about the resulting files that users can recreate our results.
+In particular, we would like to have scripts that can quickly reproduce all figures from scratch,
+but this would require far too much computation time for the user, or storage space for us.
+The worrisome note should have been reassuring: `oh, wrong recombination distance' was the commit message for the last commit the README was changed;
+this indicates that we found and fixed a problem having to do with recombination distance.
 

### 2016-04-27 drafts/hybrid_zone.tex
responses done?

@@
 
+\paragraph{Geography}
 To speed calculations and simulations, we assumed a homogeneous, one-dimensional geographical range.
@@
 
+\paragraph{Coalescence}
 %It is particularly unfortunate 
@@
 % suggests that this assumption does not have a severe impact on our major results when population densities are moderately high.
+It could be particularly important for applications to understand to what degree drift within an old, stable hybrid zone
+can mimic the haplotype patterns seen in a recently formed zone, \linelabel{ll:old_drift}
+in the same way that drift can produce clines at neutral loci in old hybrid zones.
 
@@
 We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Nick Barton and two anonymous reviewers for helpful feedback and suggestions.
-PR was supported in part by NSF grant DBI-1262645 and a Sloan Research Fellowship.
-% \alisa{More info here}
+This work was supported by National Institute of Food and Agriculture (NIFA) project \#1007008 to YB
+and by National Science Foundation grant DBI-1262645 and a Sloan Research Fellowship to PR.
 
@@
 \url{https://github.com/petrelharp/clinal-lineages}
-under the GPLv3 software license
+under the GPLv3 software license.
 

### 2016-04-27 drafts/figs/Plot_frequency_clines.R

@@ -180,7 +180,7 @@ par(mar=c(3.5,3.5,0.5,0.5))
     for(i in 1:30){
         matlines(xx, B_long[,i]/1000, col=(rainbow(43))[32-i], lty=1, lwd=0.75 )
     }
-    lines(xx, B_long[,31]/1000, col="red", lwd=2 )
+    lines(xx, B_long[,30]/1000, col="red", lwd=2 )
     mtext("Geographic position",side=1,line=2)
     mtext(expression(p[B]),side=2,line=2.5)
     legend("bottomright",legend = c("r=0","r=0.005","r=0.01","r=0.02","r=0.03","no seln."),col=c("red",rainbow(43)[c(6,11,21,31)],"darkgrey"),lwd=c(2,rep(0.75,4),0.5),cex=0.8)

### 2016-04-27 drafts/review-responses-april-2016.tex
minor changes; looks good

@@
 \reply
-\alisa{We have remade Figure 1, with additional information that we hope will help with interpretation. We have also related this figure to the terms in Equation 5.}
+Figure 1 has been remade with additional information and details that we hope will aid in the understanding of our framework.
+We have related the recombination events in the figure to the terms in Equation 5.
+\llname{ll:eq5tofig1}
 

### 2016-04-27 drafts/hybrid_zone.tex
minor changes; looks good

@@
 and the remaining portion becomes linked to a $B$ allele 
-(e.g., the lower recombination event shown in Fig.~\ref{Fig:schematic}).
+\linelabel{ll:eq5tofig1} (e.g., the lower recombination event shown in Fig.~\ref{Fig:schematic}).
 To account for cases where the selected locus lies outside the interval [a,b],
@@
 \section*{Acknowledgements}
-We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Nick Barton and two anonymous reviewers for helpful feedback and suggestions.
+We thank Graham Coop, Nancy Chen, Emily Josephs, Kristin Lee, Simon Aeschbacher, Jeremy Berg, Vince Buffalo, Nick Barton and two anonymous reviewers for helpful feedback and suggestions.
 This work was supported by National Institute of Food and Agriculture (NIFA) project \#1007008 to YB

### 2016-04-27 drafts/hybrid_zone.tex
YB made minor changes to abstract and intro. Make sure to recompile [only pushed .tex]

@@
 The lengths of continuous tracts of ancestry provide an additional source of information:
-blocks of ancestry surrounding single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location,
-an observation that might be used to characterize the age of hybrid zones
-and identify candidate targets of selection.
+blocks of ancestry surrounding incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location, %single-locus incompatibilities can be substantially longer than the genome-wide average block length at the same spatial location, % YB Comment.
+I think single-locus is a red-herring as we hope this will work for singe loci involve in multi-locus incompatibilities
+an observation that might be used to % characterize the age of hybrid zones and %YB Comment - age seems like our last paper?
+identify candidate targets of selection.
 %\alisa{ascertainment bias. Tract lengths provide additional source of information, as they provide unbiased information even when used ascertained markers.}
@@
 \citep[e.g.,][]{Gravel2012,Hellenthal2014,Harris2013,sedghifar2015spatial}. 
+Here we show how natural selection against hybrid incompatibilities shapes patterns of coancestry around such loci.
+%YB Comment.
+I really want people to know what we are doing by the end of the 1st paragraph
 
@@
 
-Our work here builds on the large body of theory describing tension zones 
+Our work here builds on the large body of theory describing allele frequency clines in tension zones % YB added ' allele frequency clines'
 and aims to describe \emph{transient} dynamics with a focus on \emph{haplotype} patterns.
@@
 
-We neglect genetic drift in our analytical results \revpoint{2}{9}
+To address this idea we develop analytical theory and a set of stochastic simulations. % YB Comment. i thought a paragraph starting with neglect was a downer
+Our analytical theory provides a quantitative description of the problem, but in so doing, we neglect genetic drift  \revpoint{2}{9}
 because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous}.
@@
 There is also substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Nagylaki1978,durrett2007width,barton2008effect,Polechova2011}.
-Our simulation results (below) suggest that although genetic drift strongly affects local noisiness of the system,
+Our simulation results  suggest that although genetic drift strongly affects local noisiness of the system,
 it does not substantially influence our conclusions at the moderate population densities we examine,

### 2016-04-27 drafts/hybrid_zone.tex
submitting

@@
 Here we show how natural selection against hybrid incompatibilities shapes patterns of coancestry around such loci.
-%YB Comment.
-I really want people to know what we are doing by the end of the 1st paragraph
 
@@
 
-Our work here builds on the large body of theory describing allele frequency clines in tension zones % YB added ' allele frequency clines'
-and aims to describe \emph{transient} dynamics with a focus on \emph{haplotype} patterns.
+Our work here builds on the large body of theory describing allele frequency clines in tension zones,
+and aims to describe {transient} dynamics with a focus on {haplotype} patterns.
 % which may carry substantial information about the formation of recent zones.
@@
 
-To address this idea we develop analytical theory and a set of stochastic simulations. % YB Comment. i thought a paragraph starting with neglect was a downer
+To address this idea we develop analytical theory and a set of stochastic simulations.
 Our analytical theory provides a quantitative description of the problem, but in so doing, we neglect genetic drift \revpoint{2}{9}
-because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002continuous}.
+(because of the well-known problems with spatial coalescence; \citet{felsenstein1975pain,barton2002continuous}).
 This should provide a good model for stochastic motion of a single lineage (our main focus),
